### 2019-04-03 review_article.Rmd
Basic layout as discussed 29/3/19

---
title: "Review Manuscript"
author: "Stephanie"
date: "02/04/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

Utilising maps to present statistics has been widely used, for centuries.
Connecting the data to the geographical representation of areas that are already familiar.
It is not enough to be recognisable, if the value of the statistic cannot be seen. Cancer outcomes directly relate to the people living within a geographical area. In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer outcomes. This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes.

## Spatial Visualisation of Disease Cases

John Snow's use of a map for cholera, cases represented as blocks. Counts within region.

Plotted as points on a map, specific representation of each case.

## Spatial Visualisation of Disease Models

Driven by progressions in computing, statistical methods and collection of data.

Models summarise relationships between counts of diseases and other factors.

Choropleth Maps, relationship between regions and spatial location. Able to represent all the occurences within areas more cohesively, as total counts, or colours.

Urbanisation pushed people closer together, in workplaces and in homes.

Recognition of biases when utilising the geography to map people. Small areas contain more people, work towards representing regions of importance more clearly.

Cartograms stepped toward this goal by highlighting areas of interest. Made some areas bigger on the map space, kept relationship with neighbours.

Importance of the statistic in this area being communicated was more important than keeping geography recognisable.

This helps to draw attention to areas, relies of reader being able to see all areas on a map, and know the difference between initial geography and new layout given by a cartogram.

We were able to create cartograms by access to data, access to tools, greater demand for insights.

In theory these ideas have been used, but the people making them do not necessarily have not access or ability to create them.

Therefore the uptake of these maps has been due to education, not the validity of their use.

## Alternative Maps

If the goal of the map is the statistic.

If the distribution is the focus, the display should reflect that.

Replies on the idea that every area is important, no matter the size.

Recent methods have every area represented with the same map space.

This gives equal emphasis to every area, allows distributions and relationships between neighbours to become more clear.

### 2019-04-04 review_article.Rmd
additional author notes included in paper

@@
 ```
+<!--
+Author notes:
+6000 words maximum.
+focused  application  of  literature  to address  a  relevant  clinical  issue:
+Mapping cancer outcomes
+Cartogram applications should point to examples of cancer outcomes, with the main cartogram section discussing sources considered in grey literature review.
+Examples of political areas may be considered if cancer data is aggregated at these levels.
+
+Future steps should consider applications not currently utilised for cancer outcome displays, where their use may improve decision making
+-->
 
@@
 
-This helps to draw attention to areas, relies of reader being able to see all areas on a map, and know the difference between initial geography and new layout given by a cartogram.
+This helps to draw attention to areas, relies on reader being able to see all areas on a map, and know the difference between initial geography and new layout given by a cartogram.
 

### 2019-04-04 review_article.Rmd
additional author notes included in paper

@@
 Mapping cancer outcomes
-Cartogram applications should point to examples of cancer outcomes, with the main cartogram section discussing sources considered in grey literature review.
-Examples of political areas may be considered if cancer data is aggregated at these levels.
+All cartogram applications should point to examples of cancer outcomes, with the main cartogram section discussing sources considered in grey literature review
 
-Future steps should consider applications not currently utilised for cancer outcome displays, where their use may improve decision making
+Future steps should consider applications not utilised for cancer outcome displays, where their use may improve decision making
 -->

### 2019-04-04 review_article.Rmd
Bibliography information

--- 2019-04-04 review_article.Rmd
@@
 output: html_document
+
+toc: false
+fig_caption: yes
+number_sections: false
+fontsize: 11pt
+bibliography: reviewbib.bib
+biblio-style: vancouver
+link-citations: true
 ---

### 2019-05-06 review_article.Rmd
Structure of review article

--- 2019-04-04 review_article.Rmd
@@
 ---
-title: "Review Manuscript"
-author: "Stephanie"
-date: "02/04/2019"
+title: "Cartogram Mapping and Application to Cancer Data."
 output: html_document
+
+toc: true
+fig_caption: yes
+number_sections: false
+fontsize: 11pt
+bibliography: reviewbib.bib
+biblio-style: vancouver
+link-citations: true
 ---
@@
 
-## Spatial Visualisation of Disease Cases
+## Spatial Visualisation of Disease
 
-John Snow's use of a map for cholera, cases represented as blocks. Counts within region.
+### Choropleth Map (current go to)
 
-Plotted as points on a map, specific representation of each case.
+### counts versus predictions
 
-## Spatial Visualisation of Disease Models
+Observed, or predicted.
+Challenges for either.
 
@@
 
+## Cartograms
+
+General introduction:
 Cartograms stepped toward this goal by highlighting areas of interest. Made some areas bigger on the map space, kept relationship with neighbours.
+USA regular cartogram (UK too)
 
+### Contiguous
 Importance of the statistic in this area being communicated was more important than keeping geography recognisable.
@@
 
-## Alternative Maps
+### Non-Contiguous
+
+Dorling (Sized according to statistic)
+
+## Temporal displays
+
+## Alternatives
+
+Plotting centroids on top of geographies. (Size is kept constant)
+Hexmaps
 
@@
 This gives equal emphasis to every area, allows distributions and relationships between neighbours to become more clear.
+
+## Comparison of 
+(only worth including if it is possible for us to implement)
+Tabular form comparing and contrasting
+- Relationship to geography
+
+- Show using cancer examples
+
+## Animation
+
+lends to temporal pattern exploration
+
+## Software
+Commonly available tools to implement.

### 2019-05-08 review_article.Rmd
Review article sources contains notes on references and how they are relevant is each section.

@@
 
+@MACM
+geographically mapping cancer mortality uses maps of europe:
+Static choropleth maps, coloured from green to red, supplemented by histograms.
+Introduces Moran scatterplot "In the Moran’s scatterplot, each geographic unit is represented by a point, whose X coordinate is the value of the mortality in that area and the Y coordinate is equal to the average value of the theme in the bordering areas'
+And Moran's index: spatial correlation with neighbours
+
+'visual intuition coming from the inspection of maps must be complemented and validated by rigorous statistical analyses.'
+Does not acknowledge the map as a visual statistic, but a tool.
+
+cites: Cancer Mortality Atlas of EU and of the European Economic Area last published 8 years ago
+
 ### counts versus predictions
 
-Observed, or predicted.
-Challenges for either.
+@SEMA
+'Disease mapping is carried out to summarise spatial and spatio-temporal variation in risk.'
+Decriptive statistic of the relationship between geography and a disease.
+population, exposure, and health data may have pointdata or aggregated summaries (countdata)
+potentially be subject to a number of inaccuracies
+diagnostic error or misclassification
+
+Not only the counts, but population, migration, deaths and births are modelled across census years.
+
+rapid visual summary of complex geographical information, descriptive purposes, to generate hypotheses, for surveillance, and to aid policy formation and resource allocation.
+
+When using counts:
+ small populations imply large variability in estimated rates
+
+Decisions taken at this stage may be critical as apparent ‘clusters’ may depend crucially on the boundaries chosen in time or space: ‘The more narrowly the underlying population is defined, the less will be the number of expected cases, the greater will be the estimate of the excess rate, and often the more pronounced will be the statistical significance’
+
+Using prediction:
+assuming that neighbouring μi are correlated leads to a ‘clustering’ prior
 

### 2019-05-08 review_article_sources.Rmd
Review article sources contains notes on references and how they are relevant is each section.

--- 2019-05-08 review_article.Rmd
@@
 
-Utilising maps to present statistics has been widely used, for centuries.
-Connecting the data to the geographical representation of areas that are already familiar.
-It is not enough to be recognisable, if the value of the statistic cannot be seen.
-Cancer outcomes directly relate to the people living within a geographical area.
-In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer outcomes.
-This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes.
+## Cartogram Visualisations of Disease
 
-## Spatial Visualisation of Disease
+Cartograms utilise a map base, and distort using a statistic.
+
+There was a long history that arrived cartogram maps.
+
+@DMAHP
+Thematic cartography began around 1800, spot maps, did not consider population rates
+available data on the environment or society, weather crime rates. 
+Disease mapping, evaluate geographical patterns, identify risk factors
+For risks to be considered, need population for rates
 
 ### Choropleth Map (current go to)
+
+True maps, used for inspection to direct hypotheses.
+
+@CIBMUK
+-data exists within boundaries,
+-areas within are shaded to match level of value,
+-compare colour of areas to compare boundaries
+-visual perception tends to weight the shades from the largest areas more heavily since they occupy more of our visual space
+-display information in units of area, not population
+-can mislead and bias inference
 
@@
 ### counts versus predictions
+Need population, census data
 
 @SEMA
-'Disease mapping is carried out to summarise spatial and spatio-temporal variation in risk.'
-Decriptive statistic of the relationship between geography and a disease.
-population, exposure, and health data may have pointdata or aggregated summaries (countdata)
-potentially be subject to a number of inaccuracies
-diagnostic error or misclassification
+ - 'Disease mapping is carried out to summarise spatial and spatio-temporal variation in risk.'
+- Decriptive statistic of the relationship between geography and a disease.
+- rapid visual summary of complex geographical information, descriptive purposes, to generate hypotheses, for surveillance, and to aid policy formation and resource allocation.
+- population, exposure, and health data may have pointdata or aggregated summaries (countdata)
+- potentially be subject to a number of inaccuracies
+- diagnostic error or misclassification
 
-Not only the counts, but population, migration, deaths and births are modelled across census years.
-
-rapid visual summary of complex geographical information, descriptive purposes, to generate hypotheses, for surveillance, and to aid policy formation and resource allocation.
+- Not only the counts, but population, migration, deaths and births are modelled across census years.
 
@@
 
-Driven by progressions in computing, statistical methods and collection of data.
+@DMAHP 
+improve rates use age and sex
+Standardised rates allow choropleth implementation
 
-Models summarise relationships between counts of diseases and other factors.
+choice of variable:
+relative risk (or equivalent), possibly in conjunction with statistical significance.
+Other functions included case frequencies or statistical significance alone.
+The choice of function makes an enormous difference to the appearance of the map, but the rationale for the choice adopted was rarely discussed.
+The Finnish cancer atlas (Pukkala et al. 1987) was one of the earliest to use data smoothing, with a geometric centroid approach
+More recently, a Canadian atlas of cancer incidence has appeared (Le et al.
+1996; see Plate 1).
+Cancer is registered provincially, so considerable attention was paid to the quality and comparability of the data from contributing registries.
+moving toward incidence
 
-Choropleth Maps, relationship between regions and spatial location. Able to represent all the occurences within areas more cohesively, as total counts, or colours.
+Very skeptical of smoothing
 
-Urbanisation pushed people closer together, in workplaces and in homes.
-
-Recognition of biases when utilising the geography to map people. Small areas contain more people, work towards representing regions of importance more clearly.
+@HCUSMA The 1980s began computed generated cancer atlases
 
@@
 
+@DMAHP
+most disease atlases concerned cancer in developed countries
+iso-demographic map: size by pop, colour by stat
+@CIBMUK
+'geographical differences does not automatically mean
+that geography by itself makes the difference.'
+Cartograms overcome this problem by standardizing the visual information presented to population size,
+
+Issues: 'Many regions had such small expected frequencies that random variation would dominate the data.
+Interestingly, the smaller regions tended to occur in atlases that covered larger total populations.'
+Smaller regions are a key issue, often occur when population has to be broken into many areas.
+
 General introduction:
 Cartograms stepped toward this goal by highlighting areas of interest. Made some areas bigger on the map space, kept relationship with neighbours.
-USA regular cartogram (UK too)
 
+UK: 
+USA:
+AUS: @DACQ AUS Cancer atlas example, reasons for choosing choropleth
+
+@TVSSS Human cartography chapter
+'in order to show the spatial structure of society the conventional use of maps of physical geography has to be rejected.'
+strong opinions on lives of people not being accurately seen in geographical maps, not specifically disease related
+
+only with the advent of sophisticated computer graphics,
+visualization, has it become possible to do this on an easily replicable basis.
+MAny examples and references to different map displays
 ### Contiguous
+
+@CIBMUK 
+-a trade-off between informational accuracy and intuitive representation
+
 Importance of the statistic in this area being communicated was more important than keeping geography recognisable.
@@
 ## Comparison of
+
+@CD 
+discusses variable representation in maps
+'Maps are designed following the cartographic visualization process'
+Aknowledges map makers consider: methods and techniques, map type
+
+Open source and commerical software can help make the maps, but decision still needs to be made, "nonprofessionals" can now make their own
+
 (only worth including if it is possible for us to implement)
@@
 Commonly available tools to implement.
+
+@DMAHP
+"7.
+Increasing access to mapping technology The development of powerful computers has been a major force in the potential for disease mapping.
+It is likely that this trend will continue, so that larger databases and more rigorous analytical methods will become accessible to a wide circle of consumers.
+The availability of the European mortality atlas on the Internet is probably only the first such venture.
+In the future we will likely see capability for on-line manipulation of data to create maps to the user’s specifications, and an interface with other databases containing information on relevant covariables."
+
+# References

### 2019-05-08 review_article.Rmd
plan from today

@@
 output: html_document
-
-toc: true
 fig_caption: yes
@@
 
-## Spatial Visualisation of Disease
+## Disease mapping
 
-### Choropleth Map (current go to)
+### Choropleth Map (Steff)
+
+Cartographer lit, cancer atlas lit
 
@@
 
+### Public face (Jessie)
+
+what has been provided to the public, norms, common practices, all of Jessie's lit review
+
+## Cartograms (Steff)
+
+General introduction:
+Cartograms stepped toward this goal by highlighting areas of interest. Made some areas bigger on the map space, kept relationship with neighbours.
+USA regular cartogram (UK too)
+
+### Contiguous
+Importance of the statistic in this area being communicated was more important than keeping geography recognisable.
+
+This helps to draw attention to areas, relies on reader being able to see all areas on a map, and know the difference between initial geography and new layout given by a cartogram.
+
+We were able to create cartograms by access to data, access to tools, greater demand for insights.
+
+In theory these ideas have been used, but the people making them do not necessarily have not access or ability to create them.
+
+Therefore the uptake of these maps has been due to education, not the validity of their use.
+
+### Non-Contiguous
+
+Dorling (Sized according to statistic)
+
+<!--
+## Temporal displays
+-->
+
+### Centroid displays
+
+Plotting centroids on top of geographies. (Size is kept constant)
+Hexmaps
+
+If the goal of the map is the statistic.
+
+If the distribution is the focus, the display should reflect that.
+
+Replies on the idea that every area is important, no matter the size.
+
+Recent methods have every area represented with the same map space.
+
+This gives equal emphasis to every area, allows distributions and relationships between neighbours to become more clear.
+
+## A critique choropleth and cartograms
+
+(only worth including if it is possible for us to implement)
+Tabular form comparing and contrasting
+- Relationship to geography
+
+- Show using cancer examples
+
+## Animation
+
+lends to temporal pattern exploration
+
+## Acknowledgements
+
+What software did we use, for eg
+
+## References
+
+<!--
+## Software
+Commonly available tools to implement.
+-->
+
+<!--
 ### counts versus predictions
@@
 Recognition of biases when utilising the geography to map people. Small areas contain more people, work towards representing regions of importance more clearly.
-
-## Cartograms
-
-General introduction:
-Cartograms stepped toward this goal by highlighting areas of interest. Made some areas bigger on the map space, kept relationship with neighbours.
-USA regular cartogram (UK too)
-
-### Contiguous
-Importance of the statistic in this area being communicated was more important than keeping geography recognisable.
-
-This helps to draw attention to areas, relies on reader being able to see all areas on a map, and know the difference between initial geography and new layout given by a cartogram.
-
-We were able to create cartograms by access to data, access to tools, greater demand for insights.
-
-In theory these ideas have been used, but the people making them do not necessarily have not access or ability to create them.
-
-Therefore the uptake of these maps has been due to education, not the validity of their use.
-
-### Non-Contiguous
-
-Dorling (Sized according to statistic)
-
-## Temporal displays
-
-## Alternatives
-
-Plotting centroids on top of geographies. (Size is kept constant)
-Hexmaps
-
-If the goal of the map is the statistic.
-
-If the distribution is the focus, the display should reflect that.
-
-Replies on the idea that every area is important, no matter the size.
-
-Recent methods have every area represented with the same map space.
-
-This gives equal emphasis to every area, allows distributions and relationships between neighbours to become more clear.
-
-## Comparison of 
-(only worth including if it is possible for us to implement)
-Tabular form comparing and contrasting
-- Relationship to geography
-
-- Show using cancer examples
-
-## Animation
-
-lends to temporal pattern exploration
-
-## Software
-Commonly available tools to implement.
+-->

### 2019-06-13 review_article.Rmd
draft of choropleth development and use for cancer mapping

@@
 
-Cartographer lit, cancer atlas lit
+Choropleth mapping methods explore a relationship across a geographical space, utilising administrative or other boundaries.
+In thematic maps colours or patterns are used represent levels, or rates, of the information communicated.
+@CIBMUK believes these 'representations are highly intuitive' as the familiar shapes help those looking at the maps to infer spatial relationships.
+They are true maps, used for visual inspection and can help to direct hypotheses about the statial patterns across the physical space. 
+These true maps rely on visual perception, a fallible metric as it can 'weight the shades from the largest areas more heavily.' The inferences derived from visual inspection, and the hypotheses developed can suffer from 'bias'.
+The use of thematic disease mapping was motivated by a 'desire to evaluate geographical patterns' @DMAHP.
+This approach has been seen as a 'trigger for questions' which drives research of spatial distributions.
+@MACM also recognises the intuition that can be derived from choropleths.
+However, his critique is that they must be 'validated by rigorous statistical analyses.'
 
-@MACM
-geographically mapping cancer mortality uses maps of europe:
-Static choropleth maps, coloured from green to red, supplemented by histograms.
-Introduces Moran scatterplot "In the Moran’s scatterplot, each geographic unit is represented by a point, whose X coordinate is the value of the mortality in that area and the Y coordinate is equal to the average value of the theme in the bordering areas'
-And Moran's index: spatial correlation with neighbours
+As a cancer mapping tool choropleths are often used.
+Specifically in epidemiology, the study of the distribution of health-related states and events.
+Almost 100 years of cancer maps have explored the US and the UK with increased effectiveness in presentation of unbiased rates.
+@DMAHP cites Cruickshank's (1947) discussion of 'formal statistical assessment of the spatial pattern' as another major advancement.
 
-'visual intuition coming from the inspection of maps must be complemented and validated by rigorous statistical analyses.'
-Does not acknowledge the map as a visual statistic, but a tool.
+The data collection methods, and administrative control within regions lends itself to the presentation of cancer mortality rates across regions.
 

### 2019-06-14 review_article.Rmd
minor edits

@@
 Choropleth mapping methods explore a relationship across a geographical space, utilising administrative or other boundaries.
-In thematic maps colours or patterns are used represent levels, or rates, of the information communicated.
+In thematic maps colours, symbols or patterns are used represent levels, or rates, of the information communicated.
 @CIBMUK believes these 'representations are highly intuitive' as the familiar shapes help those looking at the maps to infer spatial relationships.
@@
 The data collection methods, and administrative control within regions lends itself to the presentation of cancer mortality rates across regions.
+Data is now presented often age adjusted, and as relative rates of risk. Some also present changes in the rates
 
@@
 USA regular cartogram (UK too)
+@DMAHP reduce the visual impact of large areas with small populations
+
+distortion of spatial units -> proportional to the denominator variable. iso-demographic map, for approx equal population density.
 

### 2019-06-19 review_article.Rmd
Choropleth mapping with references

@@
 
-Choropleth mapping methods explore a relationship across a geographical space, utilising administrative or other boundaries.
-In thematic maps colours, symbols or patterns are used represent levels, or rates, of the information communicated.
-@CIBMUK believes these 'representations are highly intuitive' as the familiar shapes help those looking at the maps to infer spatial relationships.
-They are true maps, used for visual inspection and can help to direct hypotheses about the statial patterns across the physical space. 
+Choropleth mapping methods explore a relationship across a geographical space.
+Thematic maps involve drawing administrative or other boundaries and filling the regions with colours, symbols or patterns to present levels, or rates, of the information communicated.
+@CIBMUK believes these 'representations are highly intuitive' as the familiar shapes help those looking at the maps to visually infer spatial relationships.
+The visualisations are true maps, used for visual inspection and can help to direct hypotheses about the spatial patterns across the physical space. 
 These true maps rely on visual perception, a fallible metric as it can 'weight the shades from the largest areas more heavily.' The inferences derived from visual inspection, and the hypotheses developed can suffer from 'bias'.
 The use of thematic disease mapping was motivated by a 'desire to evaluate geographical patterns' @DMAHP.
-This approach has been seen as a 'trigger for questions' which drives research of spatial distributions.
-@MACM also recognises the intuition that can be derived from choropleths.
-However, his critique is that they must be 'validated by rigorous statistical analyses.'
+This technique has been seen as a 'trigger for questions' which drives research of spatial distributions.
+@MACM also recognise the intuition that can be derived from choropleths with the critique that they must be 'validated by rigorous statistical analyses.'
 
-As a cancer mapping tool choropleths are often used.
-Specifically in epidemiology, the study of the distribution of health-related states and events.
+In epidemiology, choropleths are often used as a mapping tool in the study of the spatial distribution.
 Almost 100 years of cancer maps have explored the US and the UK with increased effectiveness in presentation of unbiased rates.
-@DMAHP cites Cruickshank's (1947) discussion of 'formal statistical assessment of the spatial pattern' as another major advancement.
+The data collection methods, and administrative control within regions lends itself choropleth visualisation, for the presentation of cancer mortality rates across regions.
+Mortality and incidence rates are now presented age adjusted, and as relative rates of risk across the population.
 
-The data collection methods, and administrative control within regions lends itself to the presentation of cancer mortality rates across regions.
-Data is now presented often age adjusted, and as relative rates of risk. Some also present changes in the rates
+@DMAHP cites Cruickshank's (1947) discussion of using visuals as 'formal statistical assessment of the spatial pattern' as a major advancement.
+@HEDP considered maps a research tool, dots maps use spots, choropleths and thematic maps shade areas and isopleths use isolines.
+These  techniques were commonly used before the advent of computer-assisted cartography.
+
+<!--
+-Atlases
+-methods they use
+-->
 
@@
 
+@HEDP argues that 'cancer occurs in people, not in geographical areas' and the map base should reflect this to avoid allocating 'undue prominence' to rural areas.
+geographical and demographic base maps could be used after digital computer methods became available.
+
 ### Contiguous
@@
 
+Used in displays of the UK by Howe in 1986 cited by @HEDP
 <!--
@@
 
+Computer generated map examples:  @HEDP
+(Hopps et al. 1968; Armstrong 1972). There
+has followed a flood of disease atlases, mainly concentrating on the modem problems of cancer and degenerative diseases from countries as scattered as the United States (Burbank 1971; Mason et al.
+1975, 1976; Pickle et al.
+1987), the Soviet Union (Levin 1980), Japan (Shigematsu 1977), the Federal Republic of Germany
+
 ### Centroid displays
@@
 Plotting centroids on top of geographies. (Size is kept constant)
-Hexmaps
+Hex maps
 
@@
 'Disease mapping is carried out to summarise spatial and spatio-temporal variation in risk.'
-Decriptive statistic of the relationship between geography and a disease.
-population, exposure, and health data may have pointdata or aggregated summaries (countdata)
+Descriptive statistic of the relationship between geography and a disease.
+population, exposure, and health data may have point data or aggregated summaries (count data)
 potentially be subject to a number of inaccuracies
@@
 
-Choropleth Maps, relationship between regions and spatial location. Able to represent all the occurences within areas more cohesively, as total counts, or colours.
+Choropleth Maps, relationship between regions and spatial location. Able to represent all the occurrences within areas more cohesively, as total counts, or colours.
 

### 2019-06-19 review_article.Rmd
include development of SMRs?

@@
 The data collection methods, and administrative control within regions lends itself choropleth visualisation, for the presentation of cancer mortality rates across regions.
-Mortality and incidence rates are now presented age adjusted, and as relative rates of risk across the population.
+Mortality rates are now presented as relative rates of risk across the population, an age adjusted to correct for the proportion of old people.
+Howe [-@HEDP] describes Stocks development of the standardised mortality ratios through the 1930s.
+The choropleth maps presented levels of cancer via hatchings on a black and white scale.
 
 @DMAHP cites Cruickshank's (1947) discussion of using visuals as 'formal statistical assessment of the spatial pattern' as a major advancement.
+
 @HEDP considered maps a research tool, dots maps use spots, choropleths and thematic maps shade areas and isopleths use isolines.
@@
 <!--
--Atlases
+Clear examples of:
+White paper example for published Atlases
 -methods they use
@@
 @HEDP argues that 'cancer occurs in people, not in geographical areas' and the map base should reflect this to avoid allocating 'undue prominence' to rural areas.
+'intimate part of the total human and social experience.'
 geographical and demographic base maps could be used after digital computer methods became available.
+Howe's National Atlas of Disease Mortality in the United Kingdom (Howe 1963) upgraded to a 
+a demographic base "map" (Howe 1970).
+
+introduction of electronic computer-assisted techniques created a flood of disease atlases
 

### 2019-06-19 review_article.Rmd
cartograms with advent of computation al methods

@@
 
-Choropleth mapping methods explore a relationship across a geographical space.
+Choropleth mapping methods effectively explore a relationship across a geographical space.
 Thematic maps involve drawing administrative or other boundaries and filling the regions with colours, symbols or patterns to present levels, or rates, of the information communicated.
@@
 <!--
+Current atlases utilising choropleth methods include Australia:
+
 Clear examples of:
@@
 
-General introduction:
-Cartograms stepped toward this goal by highlighting areas of interest. Made some areas bigger on the map space, kept relationship with neighbours.
-USA regular cartogram (UK too)
-@DMAHP reduce the visual impact of large areas with small populations
+Cartograms are a thematic map, presenting statistical and geographical information like choropleths. The difference is the augmentation of the geographical areas.
+This was originally implemented to allow a map base that represented the populations within regions. 
+@HEDP argues that 'cancer occurs in people, not in geographical areas' and the map base should reflect this to avoid allocating 'undue prominence' to rural areas.
+It makes areas with greater populations bigger on the map space, but kept relationship with neighbours.
+@DMAHP reduce the visual impact of large areas with small populations.
+The distortion of spatial areas is proportional to the denominator variable.
+This is implemented in an iso-demographic map, for approx equal population density.
 
-distortion of spatial units -> proportional to the denominator variable. iso-demographic map, for approx equal population density.
+Geographical and demographic base maps could be implemented after digital computer methods became available.
 
-@HEDP argues that 'cancer occurs in people, not in geographical areas' and the map base should reflect this to avoid allocating 'undue prominence' to rural areas.
-'intimate part of the total human and social experience.'
-geographical and demographic base maps could be used after digital computer methods became available.
-Howe's National Atlas of Disease Mortality in the United Kingdom (Howe 1963) upgraded to a 
-a demographic base "map" (Howe 1970).
-
-introduction of electronic computer-assisted techniques created a flood of disease atlases
+@HEDP shows the introduction of electronic computer-assisted techniques created a flood of disease atlases like Howe's National Atlas of Disease Mortality in the United Kingdom (Howe 1963) which upgraded to a demographic base "map" (Howe 1970).
 
 ### Contiguous
+
+Australia (McGlashan 1977),
+
 Importance of the statistic in this area being communicated was more important than keeping geography recognisable.

### 2019-06-20 review_article.Rmd
Cartogram algorithms and methods

@@
 -methods they use
+
+cites: Cancer Mortality Atlas of EU and of the European Economic Area last published 8 years ago
+@MTMSIH Bayesian disease mapping methods for areal data - explains key concepts applied to create an atlas
+Explores Australia as a case study, utilising QLD SA2 areas
+
+@DACQ
+
 -->
 
-cites: Cancer Mortality Atlas of EU and of the European Economic Area last published 8 years ago
+@SAMGIS suggests it is the "investigators' objectives" that drive the "representation of diseases on maps". There will be situations like infectious diseases that require choropleth methods
 
@@
 Geographical and demographic base maps could be implemented after digital computer methods became available.
+Alternatively, the area on the map space can represent the value.
+There have been many algorithms presented, @SAIC investigate these implementations and the "statistical accuracy, geographical accuracy, and topological accuracy".
 
@@
 
+@MTMSIH to uncover local-level inequalities frequently masked by health estimates from large areas
+
 ### Contiguous
+
+Keep neighbours by keeping borders intact.
+
+From a computer graphics perspective it is a problem of 'map deformation' to account for the value assigned to each area.
+@ACA discuss their implementation of three methods for creating value - by - area cartograms.
+Defining cells, and manipulating the display according to statistical value of the cell.
+@ECGC present the problem of maintaining shape presevation in non-contiguous cartograms.
+
+eg. state by population
+Highlight the value of animating contigous to see changes over time, US can be recognisable but animation aides interpretation
 
@@
 
-This helps to draw attention to areas, relies on reader being able to see all areas on a map, and know the difference between initial geography and new layout given by a cartogram.
+This helps to draw attention to large value areas, relies on reader being able to see all areas on a map, and know the difference between initial geography and new layout given by a cartogram.
 
@@
 -->
+
+@TFYCC Value-Area Cartograms.
+In these cartograms a region,country, or continent is subdivided into small regions, each ofwhich is represented by a rectangle.
+This rectangle is proportionate in area to the value which it represents in certain statistical distributions.
+The regions are grouped in approximately the same positions as they are on the map.
 

### 2019-06-20 review_article.Rmd
Find more sources of atlases using cartogram (contiguous and non contigous)

@@
 <!--
-Current atlases utilising choropleth methods include Australia:
+Current atlases utilising choropleth methods 
+- include Australia:
 
@@
 Explores Australia as a case study, utilising QLD SA2 areas
-
-@DACQ
 
@@
 
-Keep neighbours by keeping borders intact.
-
+<!-- Compress this to 2 sentence introduction about methods. Focus on which atlases have implemented this technique.--->
+Cartograms keep spatial relationships of neighbours intact by keeping borders when adjusting sizes.
 From a computer graphics perspective it is a problem of 'map deformation' to account for the value assigned to each area.
 @ACA discuss their implementation of three methods for creating value - by - area cartograms.
-Defining cells, and manipulating the display according to statistical value of the cell.
 @ECGC present the problem of maintaining shape presevation in non-contiguous cartograms.
 
+The intention is to allow the map space to highlight the distribution of the variable.
+However drawing attention to large value areas and allowing reader to see all significant areas on a map, relies on them also knowing the difference between initial geography and new layout given by a cartogram.
+Importance of the statistic in this area being communicated was more important than keeping geography recognisable.
+
 eg. state by population
-Highlight the value of animating contigous to see changes over time, US can be recognisable but animation aides interpretation
 
 Australia (McGlashan 1977),
-
-Importance of the statistic in this area being communicated was more important than keeping geography recognisable.
-
-This helps to draw attention to large value areas, relies on reader being able to see all areas on a map, and know the difference between initial geography and new layout given by a cartogram.
 
@@
 The regions are grouped in approximately the same positions as they are on the map.
+
+@CVBGSD presents bivariate maps, using size and colour for different variables.
 
@@
 lends to temporal pattern exploration
+
+@ECGC ?? Highlight the value of animating contigous to see changes over time, US can be recognisable but animation aides interpretation
 

### 2019-06-24 review_article.Rmd
Editing text

@@
 @CIBMUK believes these 'representations are highly intuitive' as the familiar shapes help those looking at the maps to visually infer spatial relationships.
+Using the topography as a base map is common for thematic maps of phenomena [@CD].
 The visualisations are true maps, used for visual inspection and can help to direct hypotheses about the spatial patterns across the physical space. 
-These true maps rely on visual perception, a fallible metric as it can 'weight the shades from the largest areas more heavily.' The inferences derived from visual inspection, and the hypotheses developed can suffer from 'bias'.
+These true maps rely on visual perception, a fallible metric as it can cause viewers to 'weight the shades from the largest areas more heavily.' The inferences derived from visual inspection, and the hypotheses developed can suffer from 'bias'.
 The use of thematic disease mapping was motivated by a 'desire to evaluate geographical patterns' @DMAHP.
-This technique has been seen as a 'trigger for questions' which drives research of spatial distributions.
+This technique has been seen as a 'trigger for questions' which drives research of spatial distributions of diseases.
 @MACM also recognise the intuition that can be derived from choropleths with the critique that they must be 'validated by rigorous statistical analyses.'
@@
 Almost 100 years of cancer maps have explored the US and the UK with increased effectiveness in presentation of unbiased rates. 
-The data collection methods, and administrative control within regions lends itself choropleth visualisation, for the presentation of cancer mortality rates across regions.
-Mortality rates are now presented as relative rates of risk across the population, an age adjusted to correct for the proportion of old people.
+@HEDP considered maps a research tool, dots maps use spots, choropleths and thematic maps shade areas and isopleths use isolines.
+These  techniques were commonly used before the advent of computer-assisted cartography.
+The data collection methods of cancer mortality rates across regions, and administrative control within regions lends itself choropleth visualisation.
+Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the proportion of old people.
 Howe [-@HEDP] describes Stocks development of the standardised mortality ratios through the 1930s.
@@
 @DMAHP cites Cruickshank's (1947) discussion of using visuals as 'formal statistical assessment of the spatial pattern' as a major advancement.
+@CD considers cartographic data analysis in practice. Presenting a clear guide for preparing cartograms for use with both qualitative and quantitative information.
+The authors recognise that the creation of cartograms was largely the work of professional cartographers until the innovation of geographical information systems that welcomed 
+map users as map creators. These systems are utilised depending on 'the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)'.
 
-@HEDP considered maps a research tool, dots maps use spots, choropleths and thematic maps shade areas and isopleths use isolines.
-These  techniques were commonly used before the advent of computer-assisted cartography.
+Examples of atlases:
+Kraak 1998, Kraak and Ormeling 2011
+Bertin 1967
 
@@
 
-Cartograms are a thematic map, presenting statistical and geographical information like choropleths. The difference is the augmentation of the geographical areas.
-This was originally implemented to allow a map base that represented the populations within regions. 
+Cartograms are a thematic map, presenting statistical and geographical information like choropleths.
+@AC suggests the difference is the intentional and desirable augmentation of the size, shape or distance of geographical areas.
+This was originally implemented to allow a map base that represented the proportion of the population within each region.
+@AC considers this design a 'more socially just form of mapping' by giving all members of the population 'equitable representation'.
+This is known as an iso-demographic map, and the design inherently lends itself to epidemiolgy.
 @HEDP argues that 'cancer occurs in people, not in geographical areas' and the map base should reflect this to avoid allocating 'undue prominence' to rural areas.
-It makes areas with greater populations bigger on the map space, but kept relationship with neighbours.
 @DMAHP reduce the visual impact of large areas with small populations.
 The distortion of spatial areas is proportional to the denominator variable.
-This is implemented in an iso-demographic map, for approx equal population density.
 
-Geographical and demographic base maps could be implemented after digital computer methods became available.
 Alternatively, the area on the map space can represent the value.
@@
 
-<!-- Compress this to 2 sentence introduction about methods. Focus on which atlases have implemented this technique.--->
-Cartograms keep spatial relationships of neighbours intact by keeping borders when adjusting sizes.
+<!-- Intentionally preserve shape-->
+
+@AC describes the process of creating a population cartogram for Canada, using the 1966 Census.
+An intentional goal of this project was to maintain contiguity, while attempting to keep the actual shape of places.
+The end result was a 'very accurate isodemo-graphic map of Canada'.
+This intentional design goal coincided with the rising interest in urban geography.
+
+Cartograms keep spatial relationships of neighbours intact by preserving borders when adjusting sizes.
 From a computer graphics perspective it is a problem of 'map deformation' to account for the value assigned to each area.
-@ACA discuss their implementation of three methods for creating value - by - area cartograms.
-@ECGC present the problem of maintaining shape presevation in non-contiguous cartograms.
+@ACA discuss their implementation of three methods for creating value-by-area cartograms.
 
@@
 However drawing attention to large value areas and allowing reader to see all significant areas on a map, relies on them also knowing the difference between initial geography and new layout given by a cartogram.
-Importance of the statistic in this area being communicated was more important than keeping geography recognisable.
 
-eg. state by population
+Requires knowledge of the methods
 
 Australia (McGlashan 1977),
-
-We were able to create cartograms by access to data, access to tools, greater demand for insights.
 
@@
 -->
+
+ @ECGC present the problem of maintaining shape presevation in non-contiguous cartograms.
 

### 2019-06-25 review_article.Rmd
Daniel dorling and spatial epidemiology

@@
 map users as map creators. These systems are utilised depending on 'the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)'.
+@SE reports the increase in disease mapping development due to the availability of geographic information system softwares.
+
+Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].
 
@@
 Cartograms are a thematic map, presenting statistical and geographical information like choropleths.
-@AC suggests the difference is the intentional and desirable augmentation of the size, shape or distance of geographical areas.
-This was originally implemented to allow a map base that represented the proportion of the population within each region.
-@AC considers this design a 'more socially just form of mapping' by giving all members of the population 'equitable representation'.
+@ACTUC suggests the difference is the intentional and desirable augmentation of the size, shape or distance of geographical areas.
+World projections reflect the frequent distortions seen from altering perspectives.
+Choropleth maps will always be distorted if they cover enough of the globe, as will photographs of the globe from space.
+Cartogram creation requires choosing a favourable distortion of the geography for presenting the set of spatial information.
+This concept was originally implemented to allow a map base that represented the proportion of the population within each region.
+@ACTUC considers this design a 'more socially just form of mapping' by giving all members of the population 'equitable representation'.
 This is known as an iso-demographic map, and the design inherently lends itself to epidemiolgy.
 @HEDP argues that 'cancer occurs in people, not in geographical areas' and the map base should reflect this to avoid allocating 'undue prominence' to rural areas.
-@DMAHP reduce the visual impact of large areas with small populations.
-The distortion of spatial areas is proportional to the denominator variable.
+@DMAHP reduce the visual impact of large areas with small populations, the distortion of spatial areas is proportional to the denominator variable.
 
-Alternatively, the area on the map space can represent the value.
-There have been many algorithms presented, @SAIC investigate these implementations and the "statistical accuracy, geographical accuracy, and topological accuracy".
+Alternatively, the area of the map space can represent the value.
+There have been many algorithms presented, @SAIC provided a framework to investigate implementations and the "statistical accuracy, geographical accuracy, and topological accuracy".
 
@@
 
-@MTMSIH to uncover local-level inequalities frequently masked by health estimates from large areas
+The presentation of small areas requires more thought during the implementation of maps, @MTMSIH encourage their use to uncover local-level inequalities frequently masked by health estimates from large areas.
 
@@
 
-@AC describes the process of creating a population cartogram for Canada, using the 1966 Census.
+@ACTUC presents the three methods for creating contiguous cartograms.
+John Hunter and Johnathan Young (1968), and Durham used phsyical accretion models, arranging wooden tiles by hand.
+Tobler (1973) used a computer programs.
+Skoda and Robertson (1972) developed a mechanical model utilising steel ball bearings.
+
+<!--describes the process of creating a population cartogram for Canada, using the 1966 Census.
 An intentional goal of this project was to maintain contiguity, while attempting to keep the actual shape of places.
@@
 This intentional design goal coincided with the rising interest in urban geography.
+-->
 
-Cartograms keep spatial relationships of neighbours intact by preserving borders when adjusting sizes.
+Cartograms keep spatial relationships of neighbours intact by preserving borders when adjusting sizes. Tobler's Conformal mapping means to preserve angles
+locally so that the shapes of very small areas on a traditional map and a
+cartogram would be similar
 From a computer graphics perspective it is a problem of 'map deformation' to account for the value assigned to each area.
@@
 
-The intention is to allow the map space to highlight the distribution of the variable.
-However drawing attention to large value areas and allowing reader to see all significant areas on a map, relies on them also knowing the difference between initial geography and new layout given by a cartogram.
+Their intention is to allow the map space to highlight the distribution of the variable.
+However a reader may have to know the difference between initial geography and new layout given by a cartogram, to be able to recognise the significant changes.
 
-Requires knowledge of the methods
+Tobler's method and the many implementations that 'elaborated' on it are derived from 'numerical approximations to a pair of equations'[@ACTUC].
+They all operate through incremental adjustments, and can produce wildly different outcomes from small changes in the inputs.
 
@@
 
+@ACTUC reiterates: 'There is no "best" cartogram or method of creating cartograms just as there is no "best" map' (Monmonier and Schnell, 1988).
+There are many alternatives to consider, the intended audience of the map, and its purpose are key points in cartogram use and creation.
+
 ### Non-Contiguous
 
-Dorling (Sized according to statistic)
+Dorling -@ACTUC puts forward a simple question:
+>If, for instance, it is desirable that areas on a map have boundaries which are as simple as possible, why not draw the areas as simple shapes in the first place?
+
+He answers this with his implementation of maps created with 'the simplest of all shapes'.
+While contiguous cartograms may be a 'more sophisticated' method, they produce 'very complex shapes'.
+Circular cartograms use the same shape for every region represented, and size them according to the population, or statistic represented.
+To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighbouring areas over many iterations.
+This implementation can work for up to 'one hundred thousand' areas.
+
+@ACTUC suggests 'population distribution is often extremely uneven in former British colonies'.
+
+'In Australia the urban federal constituencies occupy only a tenth of the land, but contain nine tenths of the people.
+It would be almost unthinkable to show election results for that country on a conventional equal land area map.'
+This 1966 cartogram uses mostly straight lines, and the result looks very little like the geographical shape of Australia.
+
+'Given the increasingly uneven population distribution of the United States and the growing social divides between the populations of neighbourhoods living at different densities, the need for cartograms like this is greater now than ever.'
 

### 2019-06-26 review_article.Rmd
CPISACA additions

@@
 This technique has been seen as a 'trigger for questions' which drives research of spatial distributions of diseases.
-@MACM also recognise the intuition that can be derived from choropleths with the critique that they must be 'validated by rigorous statistical analyses.'
+@CPISACA recommends Choropleths, or isopleths for 'measures that are continuous over space, such as elevation or temperature.' @MACM also recognise the intuition that can be derived from choropleths with the critique that they must be 'validated by rigorous statistical analyses.'
 
@@
 @SE reports the increase in disease mapping development due to the availability of geographic information system softwares.
+
+<!--
+@CPISACA
+The potential audience for the results of a spatial analysis of health data is no longer limited to scientists but now also includes the public, policymakers, the media, and a host of others.
+
+Using the recognisable geography, that is familiar, helps communicate the information.
+-->
 
@@
 
-Tobler's method and the many implementations that 'elaborated' on it are derived from 'numerical approximations to a pair of equations'[@ACTUC].
-They all operate through incremental adjustments, and can produce wildly different outcomes from small changes in the inputs.
-
 Australia (McGlashan 1977),
-
-In theory these ideas have been used, but the people making them do not necessarily have not access or ability to create them.
-
-Therefore the uptake of these maps has been due to education, not the validity of their use.
 
@@
 
+ @ECGC also present the problem of maintaining shape presevation in non-contiguous cartograms.
+
 @ACTUC suggests 'population distribution is often extremely uneven in former British colonies'.
@@
 
- @ECGC present the problem of maintaining shape presevation in non-contiguous cartograms.
+Tobler's method and the many implementations that 'elaborated' on it are derived from 'numerical approximations to a pair of equations'[@ACTUC].
+They all operate through incremental adjustments, and can produce wildly different outcomes from small changes in the inputs.
 
@@
 
-@CVBGSD presents bivariate maps, using size and colour for different variables.
-
 Computer generated map examples:  @HEDP
@@
 1987), the Soviet Union (Levin 1980), Japan (Shigematsu 1977), the Federal Republic of Germany
+
+<!--
+@CVBGSD presents bivariate maps, using size and colour for different variables. 
+-->
+
+@MDAC define the term 'mosaic cartograms'. 
+Compare amount of tiles to contrast population of regions.
+'Cartograms show a data value per input region by scaling each region such that its area is proportional to its data value.
+Mosaic cartograms show data in multiples of tiles, hence the input data must consist of, or be cast into, small integer units.'
 
@@
 ## A critique choropleth and cartograms
+
+>designing a map tailored to precise goals [is] easier than forcing a single map to accommodate diverse objectives - @CPISACA
+
+With that in mind, the intented user and message to communicate should drive map selection.
+
+'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
+
+<!-- @CPISACA contrasted several online atlases in a 2006 paper, consider for structure? -->
 

### 2019-06-26 review_article.Rmd
mapped out fir section

@@
 ---
-title: "Cartogram Mapping and Application to Cancer Data."
+title: "Cartogram Mapping and its Application to Cancer Data Visualisation"
+author: "Stephanie Kobakian, Jessie Roberts and Dianne Cook"
 output: html_document
@@
 
-### Choropleth Map (Steff)
+### Choropleth map (Steff)
 
-Choropleth mapping methods effectively explore a relationship across a geographical space. 
-Thematic maps involve drawing administrative or other boundaries and filling the regions with colours, symbols or patterns to present levels, or rates, of the information communicated.
-@CIBMUK believes these 'representations are highly intuitive' as the familiar shapes help those looking at the maps to visually infer spatial relationships.
-Using the topography as a base map is common for thematic maps of phenomena [@CD].
-The visualisations are true maps, used for visual inspection and can help to direct hypotheses about the spatial patterns across the physical space. 
-These true maps rely on visual perception, a fallible metric as it can cause viewers to 'weight the shades from the largest areas more heavily.' The inferences derived from visual inspection, and the hypotheses developed can suffer from 'bias'.
-The use of thematic disease mapping was motivated by a 'desire to evaluate geographical patterns' @DMAHP.
-This technique has been seen as a 'trigger for questions' which drives research of spatial distributions of diseases.
-@CPISACA recommends Choropleths, or isopleths for 'measures that are continuous over space, such as elevation or temperature.' @MACM also recognise the intuition that can be derived from choropleths with the critique that they must be 'validated by rigorous statistical analyses.'
+<!-- what is a choropleth map -->
+A choropleth map is used to show the spatial variability of measurements, with usage dating back to the 1800s.
+They are constructed by drawing administrative boundaries, and filling the polygon with colour to represent values of a measured variable.
+Early versions used symbols or patterns instead of colour.
+@CD provides a good overview (XXX check this is true), @CPISACA discusses the use of choropleth maps for visualising cancer data, and @DMAHP gives an overview of choropleth maps for displaying disease data.
+Figure 1 shows a choropleth map on XXX.
+What can be seen.
+A choropleth map is a true map, constructed for visual inspection of spatial patterns across a familiar geographic form, that might show trends in disease occurrence, or even localised outbreaks.
 
+< Figure 1 here: choropleth map >
+
+<!-- psychology -->
+The familiar shapes help viewers to visually infer spatial relationships, and hence intuitive to read [@CIBMUK].
+However, the different sizes of administrative areas can lead viewers to place more attention on the shades of the largest areas.
+The inferences derived, and the hypotheses developed can suffer from a large area "bias".
+<!--This technique has been seen as a 'trigger for questions' which drives research of spatial distributions of diseases.
+@MACM also recognise the intuition that can be derived from choropleths with the critique that they must be 'validated by rigorous statistical analyses.' -->
+
+<!-- Data shown -->
 In epidemiology, choropleths are often used as a mapping tool in the study of the spatial distribution. 
@@
 
+<!-- Usage and availability -->
 @DMAHP cites Cruickshank's (1947) discussion of using visuals as 'formal statistical assessment of the spatial pattern' as a major advancement.
@@
 -->
+
+<!-- Published atlases -->
 
@@
 
+<!-- Going from choropleth to cartogram -->
 Cartograms are a thematic map, presenting statistical and geographical information like choropleths.
@@
 
+<!-- Common variables used to create cartograms, eg population, mortality -->
+
+<!-- Overview of varieties -->
+
 Alternatively, the area of the map space can represent the value.
@@
 
+< Figure 2 here - basic cartogram - easy to see >
+
 ### Contiguous
+
+<!-- What is a contiguous cartogram -->
 
@@
 
+< Figure 3 here>
+
 ### Non-Contiguous
@@
 'Given the increasingly uneven population distribution of the United States and the growing social divides between the populations of neighbourhoods living at different densities, the need for cartograms like this is greater now than ever.'
+
+< Figure 3 here>
 
@@
 ### Centroid displays
+
+Dot plot: one dot for each region, coloured, and placed at centroid.
 

### 2019-07-01 usa_plots.R
Code to create USA figures

@@ -1,0 +1,69 @@
+#USA Cancer Data
+
+#  United States Cancer Statistics: Data Visualizations
+#The official federal statistics on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
+
+# https://gis.cdc.gov/cancer/USCS/DataViz.html
+
+
+# United States
+# Rate of New Cancers
+# Female and Male
+# Melanoma of the skin
+# 2012-2016
+
+# Export
+
+library(sugarbag)
+library(tidyverse)
+
+# read in data
+cancer <- read_csv("data/USCS.csv") %>% rename(NAME = Area)
+
+library(spData)
+library(tmap)
+library(maptools)
+
+# Join polygons to data
+data(us_states)
+cancer <- left_join(us_states, cancer, by = c("NAME"))
+us <- sf::as_Spatial(cancer)
+us <- spTransform(us, CRS("+init=epsg:3857"))
+
+# Histogram
+ggplot(cancer, aes(x = AgeAdjustedRate)) + geom_histogram(binwidth = 1) 
+ggplot(cancer, aes(x = AgeAdjustedRate)) + geom_density()
+
+
+# Create a choropleth
+ggchoro <- ggplot(cancer) + geom_sf(aes(fill = AgeAdjustedRate)) + scale_fill_distiller(type = "div", palette = "RdYlBu")
+ggsave(filename = "figures/ggchoro.png", device = "png", width = 12, height = 6)
+
+# Cartograms
+#install.packages("cartogram")
+library(cartogram)
+
+# Contiguous Cartograms
+cont <- cartogram_cont(us,
+  weight = "AgeAdjustedRate") %>% sf::st_as_sf()
+ggcont <- ggplot(cont) + 
+  geom_sf(aes(fill = AgeAdjustedRate)) + 
+  scale_fill_distiller(type = "div", palette = "RdYlBu")
+ggsave(filename = "figures/ggcont.png", device = "png", width = 12, height = 6)
+
+# Non - Contiguous Cartograms
+ncont <- cartogram_ncont(us,
+  weight = "AgeAdjustedRate") %>% sf::st_as_sf()
+ggncont <- ggplot(ncont) + 
+  geom_sf(aes(fill = AgeAdjustedRate)) + 
+  scale_fill_distiller(type = "div", palette = "RdYlBu")
[…]

### 2019-07-01 plots.R
Aus melanoma, cancer atlas data and colours

@@ -1,0 +1,69 @@
+# Create a choropleth of Australia
+library(sugarbag)
+library(tidyverse)
+
+# Find the shape file
+aus <- file.choose()
+
+# Read in shape file, and simplify
+aus_sf <- read_shape(shp_path = aus, simplify = TRUE, keep = 0.05)
+
+# Filter islands
+aus_sf <- aus_sf %>% 
+  filter(!(SA2_NAME11 %in% 
+      c("Christmas Island", "Cocos (Keeling) Islands", "Lord Howe Island")))
+
+# Join with cancer data
+SIR <- read_csv("data/SIR Downloadable Data.csv") 
+
+SIR <- SIR %>%
+  filter(Year == "2010-2014") 
+
+SIR_fem <- SIR %>% 
+  filter(Sex_name == "Females") %>%  
+  select(Year, SA2_NAME11 = SA2_name,
+    p50, Cancer_name) %>% 
+  spread(Cancer_name, p50)
+
+SIR_male <- SIR %>%
+  filter(Year == "2010-2014") %>% 
+  filter(Sex_name == "Males") %>%  
+  select(Year, SA2_NAME11 = SA2_name,
+    p50, Cancer_name) %>% 
+  spread(Cancer_name, p50)
+
+SIR_persons <- SIR %>%
+  filter(Year == "2010-2014") %>% 
+  filter(Sex_name == "Persons") %>%  
+  select(Year, SA2_NAME11 = SA2_name,
+    p50, Cancer_name) %>% 
+  spread(Cancer_name, p50)
+
+aus_sf <- aus_sf %>% 
+  left_join(SIR_persons) 
+
+# Create a colour scheme
+nac_colours <- c("A" = "#33809d",
+  "B" = "#aec6c7",
+  "C" = "#fff4bc",
+  "D" = "#ff9a64",
+  "E" = "#ff3500")
+
+colour_cat <- function(num){
+  case_when(
+    num < 0.75  ~ "A",
+    num >= 0.75 & num < 1  ~ "B",
+    num >= 1 & num < 1.25  ~ "C",
+    num >= 1.25 & num < 1.5  ~ "D",
+    num > 1.5  ~ "E")
+} 
+
+
+mel_data <- aus_sf %>% 
+  mutate(p50_col = factor(colour_cat(Melanoma), 
+    levels = c("A", "B", "C", "D", "E")))
+
+
+# Create a choropleth
[…]

### 2019-07-01 review_article.Rmd
Knit to html gives errors

@@
 author: "Stephanie Kobakian, Jessie Roberts and Dianne Cook"
-output: html_document
+bibliography: reviewbib.bib
+output:
+  html_document: default
+  pdf_document: default
+fontsize: 11pt
+link-citations: yes
+number_sections: no
 fig_caption: yes
-number_sections: false
-fontsize: 11pt
-bibliography: reviewbib.bib
 biblio-style: vancouver
-link-citations: true
 ---
@@
 
-Utilising maps to present statistics has been widely used, for centuries.
+Utilising maps to present statistics has been widely used for centuries.
 Connecting the data to the geographical representation of areas that are already familiar.
@@
 Early versions used symbols or patterns instead of colour.
-@CD provides a good overview (XXX check this is true), @CPISACA discusses the use of choropleth maps for visualising cancer data, and @DMAHP gives an overview of choropleth maps for displaying disease data.
+@CPISACA discusses the use of choropleth maps for visualising cancer data, and @DMAHP gives an overview of the development of the use of choropleth maps for displaying disease data.
 Figure 1 shows a choropleth map on XXX.

### 2019-07-01 review_article.Rmd
html needed pandoc update

@@
   html_document: default
-  pdf_document: default
 fontsize: 11pt

### 2019-07-01 review_article.Rmd
included choropleth

@@
 @CPISACA discusses the use of choropleth maps for visualising cancer data, and @DMAHP gives an overview of the development of the use of choropleth maps for displaying disease data.
-Figure 1 shows a choropleth map on XXX.
-What can be seen.
-A choropleth map is a true map, constructed for visual inspection of spatial patterns across a familiar geographic form, that might show trends in disease occurrence, or even localised outbreaks.
+Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of Melanoma of the skin from years 2012 to 2016.
+A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form, that might show trends in disease occurrence, or even localised outbreaks.
 
-< Figure 1 here: choropleth map >
+!["A choropleth map of the United States of America. Each state is coloured according the the average of the 20"](figures/ggchoro.png)
 

### 2019-07-01 review_article.Rmd
Cartogram figures

@@
 
-!["A choropleth map of the United States of America. Each state is coloured according the the average of the 20"](figures/ggchoro.png)
+!["A choropleth map of the United States of America."](figures/ggchoro.png)
 
@@
 
-< Figure 2 here - basic cartogram - easy to see >
-
 ### Contiguous
@@
 <!-- What is a contiguous cartogram -->
+
+!["A contiguous cartogram of the Unites States of America"](figures/ggcont.png)
 
@@
 
-< Figure 3 here>
-
 ### Non-Contiguous
 
+!["A Non - contiguous cartogram of the Unites States of America"](figures/ggncont.png)
+
+!["A dorling cartogram of the Unites States of America"](figures/ggdorl.png)
+
 Dorling -@ACTUC puts forward a simple question:
+
 >If, for instance, it is desirable that areas on a map have boundaries which are as simple as possible, why not draw the areas as simple shapes in the first place?

### 2019-07-02 review_article.Rmd
USA melanoma plots, coloured by rate, sized by population

@@
 link-citations: yes
-number_sections: no
+number_sections: yes
 fig_caption: yes
@@
 <!--
+
+Please note that the submission deadline is September 30, 2019.
+
 Author notes:
@@
 
-Future steps should consider applications not currently utilised for cancer outcome displays, where their use may improve decision making
+- Future steps should consider applications not currently utilised for cancer outcome displays, where their use may improve decision making
+- needs to have a clear relevance to spatial patterns in cancer epidemiology
+- presents alternative ways of visualising those spatial patterns, or the uncertainty in those patterns
+- a more applied-focussed paper (eg lots of cool images), rather than theoretical (eg lots of formula).
 -->
@@
 
-## Disease mapping
+## Introduction
+<!--
+Current best practices: Choropleths
+Current exploratory practices: Cartograms: cont, non cont, (tilegrams, dorlings)
+Suggested uses of 
+-->
+
+## Disease mapping methods
 

### 2019-07-02 usa_plots.R

@@ -16,54 +16,76 @@
 
 library(sugarbag)
 library(tidyverse)
+library(spData)
+library(ggthemes)
+library(maptools)
+library(sf)
 
 # read in data
 cancer <- read_csv("data/USCS.csv") %>% rename(NAME = Area)
 
-library(spData)
-library(tmap)
-library(maptools)
 
 # Join polygons to data
 data(us_states)
 cancer <- left_join(us_states, cancer, by = c("NAME"))
-us <- sf::as_Spatial(cancer)
-us <- spTransform(us, CRS("+init=epsg:3857"))
+cancer <- st_transform(cancer, 3857)
+
+b <- st_bbox(cancer)
 
 # Histogram
-ggplot(cancer, aes(x = AgeAdjustedRate)) + geom_histogram(binwidth = 1) 
-ggplot(cancer, aes(x = AgeAdjustedRate)) + geom_density()
+ggplot(cancer, aes(x = total_pop_15)) + geom_histogram() 
+ggplot(cancer, aes(x = total_pop_15 )) + geom_density()
 
+title <- ""
+  #"Average rate of incidence for melanoma of the skin for females and males in the United States 2012-2016, \n where the size of each state has been adjusted for population"
 
 # Create a choropleth
-ggchoro <- ggplot(cancer) + geom_sf(aes(fill = AgeAdjustedRate)) + scale_fill_distiller(type = "div", palette = "RdYlBu")
+ggchoro <- ggplot(cancer) + 
+  geom_sf(aes(fill = AgeAdjustedRate)) +
+  scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
+  ggtitle(title) +
+  coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
+  theme_void() +
+  theme(legend.position ="bottom")
 ggsave(filename = "figures/ggchoro.png", device = "png", width = 12, height = 6)
+ggchoro
 
 # Cartograms
 #install.packages("cartogram")
 library(cartogram)
 
 # Contiguous Cartograms
-cont <- cartogram_cont(us,
-  weight = "AgeAdjustedRate") %>% sf::st_as_sf()
+cont <- cartogram_cont(cancer,
+  weight = "total_pop_15") %>% st_as_sf()
[…]

### 2019-07-02 review_article.Rmd
lung and bronchus cancer

@@
 fontsize: 11pt
-link-citations: yes
-number_sections: yes
-fig_caption: yes
+link-citations: true
+number_sections: true
+fig_caption: true
 biblio-style: vancouver
@@
 @CPISACA discusses the use of choropleth maps for visualising cancer data, and @DMAHP gives an overview of the development of the use of choropleth maps for displaying disease data.
-Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of Melanoma of the skin from years 2012 to 2016.
+Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
 A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form, that might show trends in disease occurrence, or even localised outbreaks.

### 2019-07-02 usa_plots.R

@@ -9,7 +9,7 @@
 # United States
 # Rate of New Cancers
 # Female and Male
-# Melanoma of the skin
+# Lung and Bronchus
 # 2012-2016
 
 # Export
@@ -22,7 +22,7 @@ library(maptools)
 library(sf)
 
 # read in data
-cancer <- read_csv("data/USCS.csv") %>% rename(NAME = Area)
+cancer <- read_csv("data/USCSlung.csv") %>% rename(NAME = Area)
 
 
 # Join polygons to data
@@ -37,7 +37,7 @@ ggplot(cancer, aes(x = total_pop_15)) + geom_histogram()
 ggplot(cancer, aes(x = total_pop_15 )) + geom_density()
 
 title <- ""
-  #"Average rate of incidence for melanoma of the skin for females and males in the United States 2012-2016, \n where the size of each state has been adjusted for population"
+  #"Average rate of incidence for lung and bronchus for females and males in the United States 2012-2016, \n where the size of each state has been adjusted for population"
 
 # Create a choropleth
 ggchoro <- ggplot(cancer) + 

### 2019-07-04 review_article.Rmd
CTTMB notes and added section headings

@@
 
-### Choropleth map (Steff)
+### Current best practice map displays for cancer data
+<!--Choropleth map (Steff)-->
 
@@
 <!-- psychology -->
-The familiar shapes help viewers to visually infer spatial relationships, and hence intuitive to read [@CIBMUK].
+The familiar state shapes help viewers to visually infer spatial relationships, and hence intuitive to read [@CIBMUK].
 However, the different sizes of administrative areas can lead viewers to place more attention on the shades of the largest areas.
@@
 
-<!-- Data shown -->
-In epidemiology, choropleths are often used as a mapping tool in the study of the spatial distribution. 
-Almost 100 years of cancer maps have explored the US and the UK with increased effectiveness in presentation of unbiased rates. 
-@HEDP considered maps a research tool, dots maps use spots, choropleths and thematic maps shade areas and isopleths use isolines.
-These  techniques were commonly used before the advent of computer-assisted cartography.
+<!-- Displaying cancer data on choropleths -->
+In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality. 
+Almost 100 years of cancer maps have explored the United States and the United Kingdom with increased effectiveness in presentation of unbiased rates.
+The increasing development and use of disease maps can be attributed to the availability of geographic information system softwares [@SE].
 The data collection methods of cancer mortality rates across regions, and administrative control within regions lends itself choropleth visualisation.
@@
 Howe [-@HEDP] describes Stocks development of the standardised mortality ratios through the 1930s.
-The choropleth maps presented levels of cancer via hatchings on a black and white scale.
+The choropleth maps used presented levels of cancer via hatchings on a black and white scale.
+
+@CPISACA
+The potential audience for the results of a spatial analysis of health data is no longer limited to scientists but now also includes the public, policymakers, the media, and a host of others.
 
@@
 @CD considers cartographic data analysis in practice. Presenting a clear guide for preparing cartograms for use with both qualitative and quantitative information.
-The authors recognise that the creation of cartograms was largely the work of professional cartographers until the innovation of geographical information systems that welcomed 
-map users as map creators. These systems are utilised depending on 'the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)'.
-@SE reports the increase in disease mapping development due to the availability of geographic information system softwares.
-
-<!--
-@CPISACA
-The potential audience for the results of a spatial analysis of health data is no longer limited to scientists but now also includes the public, policymakers, the media, and a host of others.
+ The authors recognise that the creation of cartograms was largely the work of professional cartographers until the innovation of geographical information systems that welcomed map users as map creators.
+These systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’.
 
@@
 <!-- Going from choropleth to cartogram -->
-Cartograms are a thematic map, presenting statistical and geographical information like choropleths.
+<!-- Why transform?-->
+Cartograms aand choropleths both present statistical and geographical information.
 @ACTUC suggests the difference is the intentional and desirable augmentation of the size, shape or distance of geographical areas.
-World projections reflect the frequent distortions seen from altering perspectives.
+Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
+
+<!-- There are many alternative views of the world-->
+<!-- Tobler map projections-->
+Some consider choropleths true topological maps, however, if the land mass presented covers enough of the globe, there must be a transformation to visualise it in 2D.
+The world projections reflect the frequent distortions seen from altering perspectives.
 Choropleth maps will always be distorted if they cover enough of the globe, as will photographs of the globe from space.
-Cartogram creation requires choosing a favourable distortion of the geography for presenting the set of spatial information.
-This concept was originally implemented to allow a map base that represented the proportion of the population within each region.
+Choropleth creation requires choosing a favourable distortion of the geography for presenting the set of spatial information.
+
+!["Choropleth maps of the United States of America using four coordinate reference systems."](figures/ggchoroCRS.png)
+
+<!-- Cartograms extend transformations -->
+Cartograms may be seen as an extension of this concept, implementing a distorion of the geographical shape to allow a map base that represents the population within each region. 
 @ACTUC considers this design a 'more socially just form of mapping' by giving all members of the population 'equitable representation'. 
+@DMAHP reduce the visual impact of large areas with small populations, the distortion of spatial areas is proportional to the denominator variable.
 This is known as an iso-demographic map, and the design inherently lends itself to epidemiolgy.
-@HEDP argues that 'cancer occurs in people, not in geographical areas' and the map base should reflect this to avoid allocating 'undue prominence' to rural areas.
-@DMAHP reduce the visual impact of large areas with small populations, the distortion of spatial areas is proportional to the denominator variable.
 
 <!-- Common variables used to create cartograms, eg population, mortality -->
+
+@HEDP argues that 'cancer occurs in people, not in geographical areas' and the map base should reflect this to avoid allocating 'undue prominence' to rural areas.
+
+<!-- 
+@CTTMB
+choropleth maps of population characteristics exhibit serious graphic distortions when the population concerned varies greatly in density.
+The topological cartogram is proposed as a viable alternative in such situations
+
+there is likely 'to manifest a growing demand for urban thematic maps'
+
+paper presents 'the topological cartogram as a base map for the communication of the spatial variation of socio-economic data within urban areas. '
+
+'it is possible to adjust their boundaries so that the areas which they enclose become proportional to these quantities rather than to their areal extent.'
+
+'Firstly, that no such transformation has a unique solution, leaving the subjective choice in the hands of the individual map designer. '
+Design goals:
+- to maintain as much as possible of the intricate detail
+- to maximize the simplicity of the transformation
+- maintenance of directional accuracy
+- attempt to influence and direct the map user's attention,
+
+1) that the size of the electoral subdivisions should be made proportional to their respective electoral populations, 
+2) that the contiguity relationships between the DCU's should be maintained, 
+3) that true directions should be retained from the centre of the transformation, 
+4) that the main shape characteristics of the DCU's should be preserved, and 
+5) that the map should possess the overall shape of the built-up area. 
+-->
 

### 2019-07-04 usa_plots.R

@@ -89,3 +89,36 @@ ggdorl <- ggplot(dorl) +
   theme_void() + theme(legend.position ="bottom")
 ggsave(filename = "figures/ggdorl.png", device = "png", width = 12, height = 6)
 ggdorl
+
+
+
+# Choropleth projections
+# 3857, 2163, 4326, 2955
+ggchoro1 <- ggplot(st_transform(cancer, 3857)) + 
+  geom_sf(aes(fill = AgeAdjustedRate)) +
+  scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
+  ggtitle("The United States using ESPG: 3857") +
+  theme_void()+ guides(fill = FALSE)
+
+ggchoro2 <- ggplot(st_transform(cancer, 2163)) + 
+  geom_sf(aes(fill = AgeAdjustedRate)) +
+  scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
+  ggtitle("The United States using ESPG: 2163") +
+  theme_void()+ guides(fill = FALSE)
+
+ggchoro3 <- ggplot(st_transform(cancer, 4326)) + 
+  geom_sf(aes(fill = AgeAdjustedRate)) +
+  scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
+  ggtitle("The United States using ESPG: 4326") +
+  theme_void()+ guides(fill = FALSE)
+
+ggchoro4 <- ggplot(st_transform(cancer, 2955)) + 
+  geom_sf(aes(fill = AgeAdjustedRate)) +
+  scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
+  ggtitle("The United States using ESPG: 2955") +
+  theme_void() + guides(fill = FALSE)
+
+ggchoroCRS <- gridExtra::grid.arrange(ggchoro1, ggchoro2, ggchoro3, ggchoro4)
+
+ggsave(filename = "figures/ggchoroCRS.png", plot = ggchoroCRS,
+  device = "png", width = 12, height = 6)

### 2019-07-04 review_article.Rmd
consider urbanisation

@@
 <!-- psychology -->
-The familiar state shapes help viewers to visually infer spatial relationships, and hence intuitive to read [@CIBMUK].
-However, the different sizes of administrative areas can lead viewers to place more attention on the shades of the largest areas.
-The inferences derived, and the hypotheses developed can suffer from a large area "bias".
-<!--This technique has been seen as a 'trigger for questions' which drives research of spatial distributions of diseases.
-@MACM also recognise the intuition that can be derived from choropleths with the critique that they must be 'validated by rigorous statistical analyses.' -->
+Utilising the familiar state shapes can make a map intuitive to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data. 
+The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
+Just as geographers are no longer the only creators of maps, @CPISACA suggests the audience of spatial health data analysis has extended beyond researchers to the public, policymakers and the media.
+However, while they can be recognisable shapes, the different sizes of administrative areas can lead viewers to give more attention to the shades of the largest areas.
+Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].
+
+<!--
+Urbanisation has pushed people closer together,
+-in workplaces and in homes. 
+-this issue is not with the size of the administrative areas,
+- the increasing urbanisation is contributing to the changing density of population
+-->
 
@@
 
-@CPISACA
-The potential audience for the results of a spatial analysis of health data is no longer limited to scientists but now also includes the public, policymakers, the media, and a host of others.
-
 <!-- Usage and availability -->
@@
 These systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’.
-
-Using the recognisable geography, that is familiar, helps communicate the information.
--->
+@MACM also recognises the intuition that can be derived from choropleths must be 'validated by rigorous statistical analyses.'
 
 <!-- Published atlases -->
-
-Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].
 
@@
 
-Clear examples of:
-White paper example for published Atlases
--methods they use
-
 cites: Cancer Mortality Atlas of EU and of the European Economic Area last published 8 years ago
@@
 -->
-
-@SAMGIS suggests it is the "investigators' objectives" that drive the "representation of diseases on maps". There will be situations like infectious diseases that require choropleth methods
 
@@
 
+@SAMGIS suggests it is the "investigators' objectives" that drive the "representation of diseases on maps". 
+-infectious diseases may require choropleth methods
+
 'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
@@
 
-Choropleth Maps, relationship between regions and spatial location. Able to represent all the occurrences within areas more cohesively, as total counts, or colours.
-
-Urbanisation pushed people closer together, in workplaces and in homes.
-
 Recognition of biases when utilising the geography to map people. Small areas contain more people, work towards representing regions of importance more clearly.

### 2019-07-07 review_article.Rmd
add public facing sections - draft

--- 2019-07-04 review_article.Rmd
@@
 output:
+  pdf_document: default
   html_document: default
 fontsize: 11pt
-link-citations: true
-number_sections: true
-fig_caption: true
+link-citations: yes
+number_sections: yes
+fig_caption: yes
 biblio-style: vancouver
@@
 
-what has been provided to the public, norms, common practices, all of Jessie's lit review
+A grey literature review conducted by (ref) identified 33 Cancer Atlases publically available cancer atlases available on the internet and published between 01/01/2010 to 11/11/2015.
+
+All the maps identified in this review were chloropleth maps, in which the representation of the map geometry stayed true to the underlying geography.
+
+<!-- Could be nice to insert a grid of the different maps. Or at least one or two full maps -->
+
+#### Publishers {-}
+
+The majority of atlases were published by non-commercial organisations, including not-for-profits (NFPs), government, research organisations, advocacy groups or a partnership between an NFP & government.
+Only one map was published by a commercial entity (Maps of Cancer Mortality Rates in Spain), in this case a media organisation El Pais^[https://elpais.com/].
+
+#### Reported measures {-}
+
+The majority of maps identified in this study reported age adjusted rates of either incidence, mortality or survival.
+The list below summarises these measures, and provides a definition of each measure.
+
+#### Geographical coverage {-}
+
+Identified cancer atlases covered geographies from all around the world: 4 were global, 3 from Australia (AUS), 11 from the United States (US), 2 from Canada (CAN), 7 from the United Kingdom (UK), 2 from Spain, 1 from Switzerland, 1 from Germany, 1 from Norway, and 1 covering the European Union.
+Not all maps had a national focus and 10 covered a region or state rather than an entire nation.
+The states or counties/regions covered were South Australia (AUS), Queensland (AUS), Ontario (CAN), Valencia (Spain), Pennsylvania county Massachusetts (US), New Hampshire (US), Cape Cod (US), Missouri (US), Florida (US), New York State (US) and Arizona (US).
+
+\newpage
+Table: (\#tab:measures) Measures used to report cancer statistics
+
+|Measure |Details| 
+|:--------------------|:------------------------------------------------|  
+|1. IR (Incidence Ratio)|$(IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}$, |
+||Cancer incidence rate in region $i$ over the average cancer incidence rate for the total region|
+|2. SIR (Standardised Incidence Ratio)| IR standardised by age structure in each region $i$|
+|3. RER |$RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}$|
+|(Relative Excess Risk)|Represents the estimate of cancer related mortality within five years of diagnosis|
+||Also referred to as 'excess hazard ratio'|
+|4. Age Adjusted Relative Risk| RR standardised by age structure in each region $i$|   
+|5. Rate per 100,000|Cancer incidence per 100,000 population|  
+|6. Age Adjusted Rate per 100,000| #5 standardised by age structure or region|   
+|7. New cancer cases per 100,000|Specific methods could not be found|   
+|8. Count|Crude cancer counts|  
+|9. Below or above Expected|Alternative expression of the SIR|
+
+#### Statistical uncertainty
+
+Cancer atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map.
+Maps that only reported this information within the supplementary material were not considered to have directly attempted to report uncertainty.
+
+The review did not reveal any novel uncertainty visualisation approaches or visualisations.
+Maps used standard and well known measures including credible intervals and standard deviation, statistical significance, boxplots and distributions.
+These maps ranged from static pdfs or infographics to interactive online resources.
+The interactivity of the more modern maps enabled uncertainty information to be incorporated without cluttering the screen, such as in a tool tip feature.
+
+```{r ci-viz2, fig.cap = "Example of CI visualisation for uncertainty representation in cancer mapping (1/3).
+Source: Alberta Health IHDA Geographic.
+(2012) ", echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+
+knitr::include_graphics("figures/Alberta_CI_viz.png")
+```
+
+```{r ci-viz3, fig.cap = "Example of CI visualisation for uncertainty representation in cancer mapping (2/3).
+Source: Pensylvania Cancer Atlas ", echo=FALSE, message=FALSE, warning=FALSE, out.width='55%', fig.align='center'}
+
+knitr::include_graphics("figures/Pennsylvania_CI_viz.png")
+
+```
+
+Close to half of the atlases identified (42%, n=14) included some measure of uncertainty.
+The most common measure used to represent uncertainty were credible or confidence intervals (CIs).
+CIs were either visualised by including their bounds in a scatterplot or graph of estimates vs region (see Figures \@ref(fig:ci-viz2)^[Alberta Health IHDA Geographic.
+(2012).
+Age-Standardised Incidence Rate of COPD, 2011.
+Retrieved from: http://www.health.alberta.ca/health-info/IHDA-geographic/COPD/incidence-agestandard/atlas.html?epik=0GJSpE_IW34lx], and \@ref(fig:ci-viz3)^[Centres for Disease Control and Prevention (CDC).
+(n.d).
+United States Cancer Statistics: An Interactive Cancer Atlas (InCA).
+Retrieved from:https://nccd.cdc.gov/DCPC_INCA/] positioned next to the map, or reported numerically through the CI upper and lower bounds listed in a data table (see Figure 3.4^[Pennsylvania Cancer Atlas.
+(n.d).
+Retrieved from:  https://www.geovista.psu.edu/grants/CDC/?epik=0lJSpE_IW34lx]).
+Of those that visualised the CIs, 30% (n=10) embedded the visualisation within a tool tip function which visualised the CI when the mouse hovered over the relevant area (see Figure 3.5^[International Agency for Research on Cancer.
+(2017).
+Atlas of Cancer Mortality in the European Union and European Economic Area 1993-1997, Annex 4 - Cancer mortality maps by site.
+Retrieved from: http://www.iarc.fr/en/publications/pdfs-online/epi/sp159/]).
+
+```{r ci-tooltip, fig.cap= "Example of CI visualisation for uncertainty representation in cancer mapping (3/3).
+Source: Centres for Disease Control and Prevention (CDC).
+United States Cancer Statistics: An Interactive Cancer Atlas (InCA) ", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '60%', fig.align = 'center'}
+
+knitr::include_graphics("figures/CDC_CI_viz_tooltip.png")
+```
+
+```{r ci-table, fig.cap= "Example of an interactive data table with CI upper and lower bounds used in cancer mapping.
+Source: Pennsylvania Cancer Atlas", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '50%', fig.align = 'center'}
+
+knitr::include_graphics("figures/uncert_table.png")
+```
+
+Methods for representing sources of uncertainty information can be visualised or communicated in different ways, examples identified through this grey literature review are listed below.
+
+```{r stndvn-example, fig.cap= "Example of standard deviation visualised in cancer mapping", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '90%', fig.align = 'center'}
+
+knitr::include_graphics("figures/standard_deviation_large.png")
+```
+
+```{r boxplot-example, fig.cap= "Example of boxplot used in cancer mapping.
+Source: Source: Pensylvania Cancer Atlas ", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '40%', fig.align = 'center'}
+
+knitr::include_graphics("figures/Boxplot2.png")
+```
+
+Table: (\#tab:method-exp) Implicit and explicit measures of uncertainty.
+
+|Measure | Example | 
+|:-------------------------|:-------------------------------------------|
+|CI Interval |Figures 3.1, 3.2, 3.3|
+|Statistical Significance|Textured overlay on top of coloured regions used to indicate statistical significance|
+|Distribution|Figure 3.4|
+|Boxplots|Figures 3.4, 3.5|
+|Sample Size|Textured overlay or lack of colour on a region, was used to show regions with small sample size|
+|Standard deviation |Figure 3.6 - the second map in the bottom right corner shows standard deviation|
 

### 2019-07-07 review_article.Rmd
transforming

--- 2019-07-04 review_article.Rmd
@@
 <!-- Going from choropleth to cartogram -->
-<!-- Why transform?-->
-Cartograms aand choropleths both present statistical and geographical information.
-@ACTUC suggests the difference is the intentional and desirable augmentation of the size, shape or distance of geographical areas.
-Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
 
@@
 <!-- Tobler map projections-->
-Some consider choropleths true topological maps, however, if the land mass presented covers enough of the globe, there must be a transformation to visualise it in 2D.
+Choropleths may be considered true topological maps, however, if the land mass presented covers enough of the globe, there must be a transformation to visualise it in 2D.
 The world projections reflect the frequent distortions seen from altering perspectives.
@@
 Choropleth creation requires choosing a favourable distortion of the geography for presenting the set of spatial information.
+
+<!-- Why transform?-->
+Cartograms provide an alternative visualisation method for statistical and geographical information.
+The key difference is the intentional, and desirable, augmentation of the size, shape or distance of geographical areas [@ACTUC].
+Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
 

### 2019-07-07 usa_plots.R
created dot plot figure

@@ -67,16 +67,27 @@ ggsave(filename = "figures/ggcont.png", device = "png", width = 12, height = 6)
 ggcont
 
 # Non - Contiguous Cartograms
-ncont <- cartogram_ncont(cancer,
+# Needs a scaling factor
+
+cancer <- cancer %>% mutate(sva = sqrt(as.numeric(total_pop_15/AREA)))
+cancer %>% 
+  ggplot(.) +
+  geom_density(aes(x = sva)) + geom_vline(aes(xintercept = 5))
+
+# The state of Vermonet is used as the anchor unit
+
+ncont <- cartogram_ncont(cancer, k = 1/5,
   weight = "total_pop_15") %>% st_as_sf()
 ggncont <- ggplot(ncont) + 
   geom_sf(aes(fill = AgeAdjustedRate)) + 
+  geom_sf(data=cancer, fill = NA, colour = "grey") +
   ggtitle(title) + 
   scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
   coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
   theme_void() + theme(legend.position ="bottom")
 ggsave(filename = "figures/ggncont.png", device = "png", width = 12, height = 6)
 ggncont
+
 
 # Non - Contiguous Dorling Cartograms
 dorl <- cartogram_dorling(cancer,
@@ -91,6 +102,21 @@ ggsave(filename = "figures/ggdorl.png", device = "png", width = 12, height = 6)
 ggdorl
 
 
+
+# Centroid Cartograms
+cancer <- cancer %>% 
+  sf::st_centroid(geometry) %>% 
+  sf::st_coordinates() %>% as_tibble() %>% bind_cols(cancer, .)
+
+ggdot <- ggplot(cancer) + 
+  geom_point(aes(X,Y, colour = AgeAdjustedRate)) + 
+  geom_sf(data=cancer, fill = NA, colour = "grey") +
+  ggtitle(title) +
+  scale_colour_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
+  coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
+  theme_void() + theme(legend.position ="bottom")
[…]

### 2019-07-10 review_article.Rmd
Updates to sources and cartogram edits

@@
 
+<!--
+@GOINO major geographic storage device; map reading and map interpretation have been the basic methods of retrieval and analysis.
+Geography’s interests are in spatial distributions and associations, in areal differentiation
+and in spatial processes. All of these interests demand information that is geographically
+ordered. Geographically ordered data are those data which have their locational
+coordinates preserved.
+
+Certain limitations to the standard graphic techniques have also become apparent, however.
+They are slow, and do not permit instant retrieval of information.
+They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of
+spatial distributions visually.
+
+These opportunities will be lost if locational coding is not inserted into the data referencing system as such banks are built, however.
+
+There are six potential benefits to geographic research from computer-connected data banks in which the data have proper locational referents.
+These come under the general headings of mapping, partitioning and aggregation, merging of data systems, transformations of geographic spaces, descriprive statistics, and intermediate analysis.
+
+of some element, for example population density, or purchasing power, or agricultural fertility; reality, of course, is not like this.
+the background variable assumed to be uniform (such as population density) can be adjusted so that it behaves as if uniformity were the case
+-->
+
 ### Current best practice map displays for cancer data
+
 <!--Choropleth map (Steff)-->
@@
 Almost 100 years of cancer maps have explored the United States and the United Kingdom with increased effectiveness in presentation of unbiased rates.
-The increasing development and use of disease maps can be attributed to the availability of geographic information system softwares [@SE].
+The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
 The data collection methods of cancer mortality rates across regions, and administrative control within regions lends itself choropleth visualisation.
@@
 The choropleth maps used presented levels of cancer via hatchings on a black and white scale.
+
+<!--
+Computing mapping methods improved with access to necessary information: locations and associated data points.
+@GOINO discusses progression in computer mapping methods with the availabilty of census data with associated locations.
+-->
 
@@
 
-### Public face (Jessie)
+### Public face
 
@@
 
-## Cartograms (Steff)
+## Cartograms
 
@@
 <!-- Tobler map projections-->
-Choropleths may be considered true topological maps, however, if the land mass presented covers enough of the globe, there must be a transformation to visualise it in 2D.
+Choropleths may be considered true topological maps, however, if the land mass presented covers enough of the globe, there must be a transformation or distortion to display in 2D.
+The amount of distortion is related to the distance displayed @GAMP.
 The world projections reflect the frequent distortions seen from altering perspectives.
 Choropleth maps will always be distorted if they cover enough of the globe, as will photographs of the globe from space.
-Choropleth creation requires choosing a favourable distortion of the geography for presenting the set of spatial information.
+Choropleth creation requires choosing a map projection that shows a favourable distortion of the geography for presenting the set of spatial information.
+Diagrams that do not specify a projection can be considered to have some unknown projection. <!-- ?? cite tobler.-->
+
+!["Choropleth maps of the United States of America using four coordinate reference systems."](figures/ggchoroCRS.png)
 
@@
 Cartograms provide an alternative visualisation method for statistical and geographical information.
-The key difference is the intentional, and desirable, augmentation of the size, shape or distance of geographical areas [@ACTUC].
-Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
-
-!["Choropleth maps of the United States of America using four coordinate reference systems."](figures/ggchoroCRS.png)
-
-<!-- Cartograms extend transformations -->
-Cartograms may be seen as an extension of this concept, implementing a distorion of the geographical shape to allow a map base that represents the population within each region. 
-@ACTUC considers this design a 'more socially just form of mapping' by giving all members of the population 'equitable representation'. 
-@DMAHP reduce the visual impact of large areas with small populations, the distortion of spatial areas is proportional to the denominator variable.
-This is known as an iso-demographic map, and the design inherently lends itself to epidemiolgy.
+The key difference is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC].
+Cartograms may be seen as an extension of map tranformations and projections. The favourable distortion is proportional to a value other other than actual earth size area [@NAC].
 
@@
 
-@HEDP argues that 'cancer occurs in people, not in geographical areas' and the map base should reflect this to avoid allocating 'undue prominence' to rural areas.
+Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
+When implementing a distortion of the geographical shape according to population, an area cartogram [@NAC], or iso-demographic map is the result.
+When visualising population statistics @ACTUC considers this equitable representation design 'more socially just, giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
+@HEDP agrees that 'cancer occurs in people, not in geographical areas' and the map bases of population reflect this and avoid allocating 'undue prominence' to rural areas.
 
 <!--
+> The aim here is to see the whole, in as much detail as possible, at a glance.
+
+@AOCAFE
+An Online Census Atlas for everyone
+The familiar choropleth maps in the Online Census Atlas were produced using generalized boundaries.
+Great Britain is recognizable and it is relatively easy to identify the higher geography levels such as Counties and Local Authorities.
+A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, making interpretation of spatial patterns very difficult.
+
+The Atlas does not allow visualization of census
+data at ward level, but instead uses these boundaries
+aggregated to the six higher level geographies
+defined above for display of 88 census variables.
+
+To cartograms by:
+The creation of the map began
+with a computer-generated cartogram of parliamentary
+constituencies where each constituency was
+given equal weight.
+Using the circular cartogram
+algorithm (Dorling 1995), this resulted in a largely
+hexagonal tessellation of 641 circles, each circle
+representing a parliamentary constituency as first
+used in the 1997 general election.
+-->
+
 @CTTMB
@@
 
-there is likely 'to manifest a growing demand for urban thematic maps'
+@CTTMB believes there is likely 'to manifest a growing demand for urban thematic maps'.
 
@@
 
-'it is possible to adjust their boundaries so that the areas which they enclose become proportional to these quantities rather than to their areal extent.'
+<!-- Overview of varieties -->
+
+Alternatively, the area of the map space can represent the value.
+There have been many algorithms presented, @SAIC provided a framework to investigate implementations and the "statistical accuracy, geographical accuracy, and topological accuracy".
+
+@HEDP discusses the impact of electronic computer-assisted techniques. Disease atlases were implemented using both geographic (Howe 1963) and demographic bases (Howe 1970).
+
+The presentation of small areas requires more thought during the implementation of maps, @MTMSIH encourage their use to uncover local-level inequalities frequently masked by health estimates from large areas.
+
+<!--
 
@@
 
-<!-- Overview of varieties -->
-
-Alternatively, the area of the map space can represent the value.
-There have been many algorithms presented, @SAIC provided a framework to investigate implementations and the "statistical accuracy, geographical accuracy, and topological accuracy".
-
-@HEDP shows the introduction of electronic computer-assisted techniques created a flood of disease atlases like Howe's National Atlas of Disease Mortality in the United Kingdom (Howe 1963) which upgraded to a demographic base "map" (Howe 1970).
-
-The presentation of small areas requires more thought during the implementation of maps, @MTMSIH encourage their use to uncover local-level inequalities frequently masked by health estimates from large areas.
-
 ### Contiguous
@@
 
-@ACTUC presents the three methods for creating contiguous cartograms.
-John Hunter and Johnathan Young (1968), and Durham used phsyical accretion models, arranging wooden tiles by hand.
-Tobler (1973) used a computer programs.
-Skoda and Robertson (1972) developed a mechanical model utilising steel ball bearings.
-
 <!--describes the process of creating a population cartogram for Canada, using the 1966 Census.
@@
 -->
+@NAC Their intention is to allow the map space to highlight the distribution of the variable.
+However a reader may have to know the difference between initial geography and new layout given by a cartogram, to be able to recognise the significant changes.
 
-Cartograms keep spatial relationships of neighbours intact by preserving borders when adjusting sizes. Tobler's Conformal mapping means to preserve angles
+@CBATCC conflicting tasks: adjusting region sizes and retainingregion shapes. in context of agricultural applications
+
+Distortion of shape on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the noncontiguous cartogram but is replaced by the meaningful empty-space property.
+@NAC
+
+Cartograms keep spatial relationships of neighbours intact by preserving borders when adjusting sizes [-@NAC]. Tobler's Conformal mapping means to preserve angles
 locally so that the shapes of very small areas on a traditional map and a
 cartogram would be similar
+
 From a computer graphics perspective it is a problem of 'map deformation' to account for the value assigned to each area.
@@
 
-Their intention is to allow the map space to highlight the distribution of the variable.
-However a reader may have to know the difference between initial geography and new layout given by a cartogram, to be able to recognise the significant changes.
+@CBATCC Through the spatial transformation of map regions relative to the data, the cartogram
+prominently emphasizes data distribution instead of territorial size.
+-better visualized on a map more representative of population.
+-A continuous area cartogram is one in which the topology, or connectivity, of the map regions
+is retained while areas are resized.
+-explains Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method, Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., and Gusein-
+Zade and Tikunov’s Line Integral Method.
+- compare to their Constraint-Based Method (Kocmoud and House)
+
+presents 7 desirable characteristics
+1. Independent of region traversal order
+2. Independent of coordinate axes
+3. Conformal mapping
+4. Global displacements per iteration
+5. Intersection prevention
+6. Ability to fix (pin down) points
+7. User controls on area vs. shape
 
@@
 
+@ACTUC presents the three methods for creating contiguous cartograms.
+John Hunter and Johnathan Young (1968), and Durham used phsyical accretion models, arranging wooden tiles by hand.
+Tobler (1973) used a computer programs.
+Skoda and Robertson (1972) developed a mechanical model utilising steel ball bearings.
+
+<!--
+check this
+@ECGC also present the problem of maintaining shape presevation in non-contiguous cartograms. 
+-->
+
 ### Non-Contiguous
+
+<!--
+[-@NAC] observation units stay in more or less their correct location and maintain correct shape, but at least some units are separated by gaps from their neighbors.
+Each area in a non-contiguous cartogram is scaled down or up according to value for the specific area.
+A scaling factor specific to
+
+the empty area between units is meaningful. 
+the degree of difference from the original map that is the real message of a cartogram, @NAC
+
+A cartogram using the unit with the mean or median density as the anchor would also show relative discrepancies, but it is more difficult to produce by computer because many units increase in size and overlap.
+@NAC
+
+The success of the
+visual representation (contiguous or noncontiguous)
+depends, in fact, on the readers’
+recognition of the units shown and
+such recognizable units are generally fairly
+large for the area represented. @NAC
+
+constant shapes are easily recognized regardless of scale;
+-->
 
@@
 
- @ECGC also present the problem of maintaining shape presevation in non-contiguous cartograms.
-
 @ACTUC suggests 'population distribution is often extremely uneven in former British colonies'.
@@
 'Given the increasingly uneven population distribution of the United States and the growing social divides between the populations of neighbourhoods living at different densities, the need for cartograms like this is greater now than ever.'
-
-< Figure 3 here>
 
@@
 
+@NAC gives an example of
+
 Plotting centroids on top of geographies. (Size is kept constant)
-Hex maps
-
-If the goal of the map is the statistic.
-
-If the distribution is the focus, the display should reflect that.
 
 Replies on the idea that every area is important, no matter the size.
-
-Recent methods have every area represented with the same map space.
-
 This gives equal emphasis to every area, allows distributions and relationships between neighbours to become more clear.
@@
 With that in mind, the intented user and message to communicate should drive map selection.
+first is the nature or inherent properties of the device itself
+second is the ease and accuracy with which map users are able to extract information
+from it
 

### 2019-07-11 review_article.Rmd
Cartogram section edits

@@
   pdf_document: default
+  word_document: default
   html_document: default
@@
 <!-- Why transform?-->
+<!-- Common variables used to create cartograms, eg population, mortality -->
+Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
 Cartograms provide an alternative visualisation method for statistical and geographical information.
-The key difference is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC].
+The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC].
 Cartograms may be seen as an extension of map tranformations and projections. The favourable distortion is proportional to a value other other than actual earth size area [@NAC].
 
-<!-- Common variables used to create cartograms, eg population, mortality -->
-
-Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
+ @CTTMB argues that using choropleth maps for population characteristics requires graphic distortions when the population concerned varies greatly in density. 
 When implementing a distortion of the geographical shape according to population, an area cartogram [@NAC], or iso-demographic map is the result.
 When visualising population statistics @ACTUC considers this equitable representation design 'more socially just, giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
+@CTTMB suggest that spatial socio-economic data is best presented on a cartogram for urban areas.
 @HEDP agrees that 'cancer occurs in people, not in geographical areas' and the map bases of population reflect this and avoid allocating 'undue prominence' to rural areas.
+@MTMSIH encourage the use of cartograms to highlight small areas and uncover local-level inequalities.
+Using health estimates from large areas can prevent drawing attention to the inequalities.
 
@@
 > The aim here is to see the whole, in as much detail as possible, at a glance.
+
+@HEDP discusses the impact of electronic computer-assisted techniques. Disease atlases were implemented using both geographic (Howe 1963) and demographic bases (Howe 1970).
 
@@
 
-@CTTMB
-choropleth maps of population characteristics exhibit serious graphic distortions when the population concerned varies greatly in density.
-The topological cartogram is proposed as a viable alternative in such situations
-
-@CTTMB believes there is likely 'to manifest a growing demand for urban thematic maps'.
-
-paper presents 'the topological cartogram as a base map for the communication of the spatial variation of socio-economic data within urban areas. '
-
 <!-- Overview of varieties -->
 
-Alternatively, the area of the map space can represent the value.
+The area on the map space can also be used represent a value.
 There have been many algorithms presented, @SAIC provided a framework to investigate implementations and the "statistical accuracy, geographical accuracy, and topological accuracy".
+The spatial transformation of map regions relative to the data emphasises the data distribution instead of land size [@CBATCC].
 
-@HEDP discusses the impact of electronic computer-assisted techniques. Disease atlases were implemented using both geographic (Howe 1963) and demographic bases (Howe 1970).
+Cancer statistics and atlases
 
-The presentation of small areas requires more thought during the implementation of maps, @MTMSIH encourage their use to uncover local-level inequalities frequently masked by health estimates from large areas.
+<!-- Cartogram makers -->
 
-<!--
+The creation of cartograms was largely in the hands of cartographers. 
+@ACTUC discusses early approaches including John Hunter and Johnathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball bearing approach and Tobler's (1973) computer programs.
+
+There are many alternatives to consider, the intended audience of the map, and its purpose are key points in cartogram use and creation.
+@ACTUC reiterates: 'There is no "best" cartogram or method of creating cartograms just as there is no "best" map' (Monmonier and Schnell, 1988).
+
+<!-- Creation goals-->
 
 'Firstly, that no such transformation has a unique solution, leaving the subjective choice in the hands of the individual map designer. '
+
 Design goals:
@@
 5) that the map should possess the overall shape of the built-up area.
--->
 
@@
 
-<!--describes the process of creating a population cartogram for Canada, using the 1966 Census.
+A contiguous cartogram allows the map space to highlight the distribution of the variable.
+These displays keep spatial relationships of neighbours intact by preserving borders when adjusting sizes [@NAC].
+From a computer graphics perspective, @ACA believe it is a problem of 'map deformation' to account for the value assigned to each area, they provide three methods for creating value-by-area cartograms.
+A continuous area cartogram maintains connectivity of the map regions while areas are resized according to a statistic [@CBATCC].
+Examples include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [-@ACTUC], Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and Tikunov’s Line Integral Method, Constraint-Based Method (Kocmoud and House) [-@CBATCC].
+- compare to their
+
+describes the process of creating a population cartogram for Canada, using the 1966 Census.
 An intentional goal of this project was to maintain contiguity, while attempting to keep the actual shape of places.
@@
 This intentional design goal coincided with the rising interest in urban geography.
--->
-@NAC Their intention is to allow the map space to highlight the distribution of the variable.
-However a reader may have to know the difference between initial geography and new layout given by a cartogram, to be able to recognise the significant changes.
 
-@CBATCC conflicting tasks: adjusting region sizes and retainingregion shapes. in context of agricultural applications
-
-Distortion of shape on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the noncontiguous cartogram but is replaced by the meaningful empty-space property.
-@NAC
-
-Cartograms keep spatial relationships of neighbours intact by preserving borders when adjusting sizes [-@NAC]. Tobler's Conformal mapping means to preserve angles
-locally so that the shapes of very small areas on a traditional map and a
-cartogram would be similar
-
-From a computer graphics perspective it is a problem of 'map deformation' to account for the value assigned to each area.
-@ACA discuss their implementation of three methods for creating value-by-area cartograms.
-
-@CBATCC Through the spatial transformation of map regions relative to the data, the cartogram
-prominently emphasizes data distribution instead of territorial size.
--better visualized on a map more representative of population.
--A continuous area cartogram is one in which the topology, or connectivity, of the map regions
-is retained while areas are resized.
--explains Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method, Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., and Gusein-
-Zade and Tikunov’s Line Integral Method.
-- compare to their Constraint-Based Method (Kocmoud and House)
-
+<!--
 presents 7 desirable characteristics
@@
 
-@ACTUC reiterates: 'There is no "best" cartogram or method of creating cartograms just as there is no "best" map' (Monmonier and Schnell, 1988).
-There are many alternatives to consider, the intended audience of the map, and its purpose are key points in cartogram use and creation.
+-->
 
-@ACTUC presents the three methods for creating contiguous cartograms.
-John Hunter and Johnathan Young (1968), and Durham used phsyical accretion models, arranging wooden tiles by hand.
-Tobler (1973) used a computer programs.
-Skoda and Robertson (1972) developed a mechanical model utilising steel ball bearings.
-
-<!--
-check this
-@ECGC also present the problem of maintaining shape presevation in non-contiguous cartograms. 
--->
+To be able to recognise the significant changes, a reader will usually have to know the initial geography to find the differences in the new cartogram layout [@NAC].
+Tobler's Conformal mapping means to preserve angles
+locally so that the shapes of very small areas on a traditional map and a
+cartogram would be similar.
+@CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes.
+Distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the noncontiguous cartogram but is replaced by the meaningful empty-space property [@NAC, @ECGC].
 

### 2019-07-16 usa_plots.R
Choropleth titles

@@ -123,25 +123,25 @@ ggdot
 ggchoro1 <- ggplot(st_transform(cancer, 3857)) + 
   geom_sf(aes(fill = AgeAdjustedRate)) +
   scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
-  ggtitle("The United States using ESPG: 3857") +
+  ggtitle("The United States using EPSG: 3857") +
   theme_void()+ guides(fill = FALSE)
 
 ggchoro2 <- ggplot(st_transform(cancer, 2163)) + 
   geom_sf(aes(fill = AgeAdjustedRate)) +
   scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
-  ggtitle("The United States using ESPG: 2163") +
+  ggtitle("The United States using EPSG: 2163") +
   theme_void()+ guides(fill = FALSE)
 
 ggchoro3 <- ggplot(st_transform(cancer, 4326)) + 
   geom_sf(aes(fill = AgeAdjustedRate)) +
   scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
-  ggtitle("The United States using ESPG: 4326") +
+  ggtitle("The United States using EPSG: 4326") +
   theme_void()+ guides(fill = FALSE)
 
 ggchoro4 <- ggplot(st_transform(cancer, 2955)) + 
   geom_sf(aes(fill = AgeAdjustedRate)) +
   scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
-  ggtitle("The United States using ESPG: 2955") +
+  ggtitle("The United States using EPSG: 2955") +
   theme_void() + guides(fill = FALSE)
 
 ggchoroCRS <- gridExtra::grid.arrange(ggchoro1, ggchoro2, ggchoro3, ggchoro4)

### 2019-07-16 review_article.Rmd
GOINO maps as a storage device notes

@@
 
-<!--
-@GOINO major geographic storage device; map reading and map interpretation have been the basic methods of retrieval and analysis.
-Geography’s interests are in spatial distributions and associations, in areal differentiation
-and in spatial processes. All of these interests demand information that is geographically
-ordered. Geographically ordered data are those data which have their locational
-coordinates preserved.
-
-Certain limitations to the standard graphic techniques have also become apparent, however.
-They are slow, and do not permit instant retrieval of information.
-They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of
-spatial distributions visually.
-
-These opportunities will be lost if locational coding is not inserted into the data referencing system as such banks are built, however.
-
-There are six potential benefits to geographic research from computer-connected data banks in which the data have proper locational referents.
-These come under the general headings of mapping, partitioning and aggregation, merging of data systems, transformations of geographic spaces, descriprive statistics, and intermediate analysis.
-
-of some element, for example population density, or purchasing power, or agricultural fertility; reality, of course, is not like this.
-the background variable assumed to be uniform (such as population density) can be adjusted so that it behaves as if uniformity were the case
--->
-
 ### Current best practice map displays for cancer data
@@
 
+<!--
+@GOINO major geographic storage device; map reading and map interpretation have been the basic methods of retrieval and analysis.
+Geography’s interests are in spatial distributions and associations, in areal differentiation
+and in spatial processes. All of these interests demand information that is geographically
+ordered. Geographically ordered data are those data which have their locational
+coordinates preserved.
+-->
+
 !["A choropleth map of the United States of America."](figures/ggchoro.png)
@@
 Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].
+
+<!--
+Certain limitations to the standard graphic techniques have also become apparent, however.
+They are slow, and do not permit instant retrieval of information.
+They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually.
+
+These opportunities will be lost if locational coding is not inserted into the data referencing system as such banks are built, however.
+
+of some element, for example population density, or purchasing power, or agricultural fertility; reality, of course, is not like this.
+the background variable assumed to be uniform (such as population density) can be adjusted so that it behaves as if uniformity were the case
+-->
 

### 2019-07-16 review_article.Rmd
section numbering

@@
 output: 
-  pdf_document: default
-  word_document: default
+  pdf_document:
+    toc: true
+    toc_depth: 2
+    number_sections: true
   html_document: default
 fontsize: 11pt
-link-citations: yes
-number_sections: yes
+link-citations: true
 fig_caption: yes
@@
 
-## Abstract
+\section*{Abstract}
 
@@
 
-## Introduction
+# Introduction
 <!--
@@
 
-## Disease mapping methods
+# Disease mapping methods
 
-### Current best practice map displays for cancer data
+## Current best practice map displays for cancer data
 
@@
 A choropleth map is used to show the spatial variability of measurements, with usage dating back to the 1800s.
-They are constructed by drawing administrative boundaries, and filling the polygon with colour to represent values of a measured variable.
+They are constructed by drawing the given geographic or political boundaries, and filling the polygon with colour to represent values of a measured variable [@EI].
 Early versions used symbols or patterns instead of colour.
 @CPISACA discusses the use of choropleth maps for visualising cancer data, and @DMAHP gives an overview of the development of the use of choropleth maps for displaying disease data.
+A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form, that might show trends in disease occurrence, or even localised outbreaks.
 Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
-A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form, that might show trends in disease occurrence, or even localised outbreaks.
 
@@
 Just as geographers are no longer the only creators of maps, @CPISACA suggests the audience of spatial health data analysis has extended beyond researchers to the public, policymakers and the media.
-However, while they can be recognisable shapes, the different sizes of administrative areas can lead viewers to give more attention to the shades of the largest areas.
+However, while the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, different population densities and subject to changes over time.
+The different sizes of administrative areas can attract attention to the shades of the unpopulated, large physical areas [@EI].
 Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].
@@
 
-### Public face
+## Public face
 
@@
 
-#### Publishers {-}
+### Publishers {-}
 
@@
 
-#### Reported measures {-}
+### Reported measures {-}
 
@@
 
-#### Geographical coverage {-}
+### Geographical coverage {-}
 
@@
 
-#### Statistical uncertainty
+### Statistical uncertainty
 
@@
 A contiguous cartogram allows the map space to highlight the distribution of the variable.
-These displays keep spatial relationships of neighbours intact by preserving borders when adjusting sizes [@NAC].
+These cartograms maintain connectivity of the map regions while areas are resized according to a statistic, this often must occur at the expense of the shape [@CBATCC,@NAC].
 From a computer graphics perspective, @ACA believe it is a problem of 'map deformation' to account for the value assigned to each area, they provide three methods for creating value-by-area cartograms.
-A continuous area cartogram maintains connectivity of the map regions while areas are resized according to a statistic [@CBATCC].
 Examples include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [-@ACTUC], Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and Tikunov’s Line Integral Method, Constraint-Based Method (Kocmoud and House) [-@CBATCC].
@@
 While contiguous cartograms may be a 'more sophisticated' method, they produce 'very complex shapes'.
-Circular cartograms use the same shape for every region represented, and size them according to the population, or statistic represented.
+Circular cartograms use the same shape for every region represented, and size them according to the statistic represented or the population for a base map.
 To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighbouring areas over many iterations.
@@
 <!--
-## Temporal displays
+# Temporal displays
 -->
@@
 
-### Centroid displays
+## Centroid displays
 
@@
 
-## A critique choropleth and cartograms
+# A critique of mapping methods
 
@@
 
-## Animation
+# Animation
 
@@
 
-## Acknowledgements
+# Acknowledgements
 
@@
 
-## References
+# References
 
 <!--
-## Software
+# Software
 Commonly available tools to implement.
@@
 <!--
-### counts versus predictions
+## counts versus predictions
 

### 2019-07-18 review_article.Rmd
slight edit of Abstract

@@
 Utilising maps to present statistics has been widely used for centuries.
-Connecting the data to the geographical representation of areas that are already familiar.
-It is not enough to be recognisable, if the value of the statistic cannot be seen.
-Cancer outcomes directly relate to the people living within a geographical area.
-In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer outcomes.
-This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes.
+These maps connec data and/or statistics to the geographical representation of areas that are generally familiar to the audience.
+However, it is not enough to be recognisable, if the meaning of the statistic cannot be seen.
+This is particularly relevant to cancer statistics, as cancer outcomes directly relate to the people living within a geographical area.
+In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer statistics.
+This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes, and health outcomes more braodly.
 
@@
 
-A grey literature review conducted by (ref) identified 33 Cancer Atlases publically available cancer atlases available on the internet and published between 01/01/2010 to 11/11/2015.
+A grey literature review conducted by (ref) identified 33 publically available cancer atlases published on the internet between 01/01/2010 to 11/11/2015.
 

### 2019-07-18 review_article.Rmd
slight edits to section before Pubilc Face section

@@
 <!-- what is a choropleth map -->
-A choropleth map is used to show the spatial variability of measurements, with usage dating back to the 1800s.
-They are constructed by drawing the given geographic or political boundaries, and filling the polygon with colour to represent values of a measured variable [@EI].
-Early versions used symbols or patterns instead of colour.
+A choropleth map is used to show the spatial variability of measurements or statistics, with usage dating back to the 1800s.
+They are constructed by drawing the given geographic or political boundaries, and filling the resulting polygon with colour to represent values of a measured variable [@EI].
+Early versions of these maps used symbols or patterns instead of colour.
 @CPISACA discusses the use of choropleth maps for visualising cancer data, and @DMAHP gives an overview of the development of the use of choropleth maps for displaying disease data.
-A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form, that might show trends in disease occurrence, or even localised outbreaks.
+A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form, that might show trends in disease occurrences, or even localised outbreaks.
 Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
@@
 <!-- psychology -->
-Utilising the familiar state shapes can make a map intuitive to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data.
+Utilising the familiar state boundaries can make a map intuitive to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data, i.e.
+how cancer rate differs across the states.
 The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
-Just as geographers are no longer the only creators of maps, @CPISACA suggests the audience of spatial health data analysis has extended beyond researchers to the public, policymakers and the media.
-However, while the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, different population densities and subject to changes over time.
-The different sizes of administrative areas can attract attention to the shades of the unpopulated, large physical areas [@EI].
+Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
+However, while the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
+The different population and geographical sizes of administrative areas can attract attention to the shades of the unpopulated but large areas [@EI].
 Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].
@@
 In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality. 
-Almost 100 years of cancer maps have explored the United States and the United Kingdom with increased effectiveness in presentation of unbiased rates.
+Almost 100 years of cancer mapping in the United States and the United Kingdom has increased the effectiveness in the presentation of unbiased rates.
 The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
-The data collection methods of cancer mortality rates across regions, and administrative control within regions lends itself choropleth visualisation.
-Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the proportion of old people.
+The data collection methods of cancer mortality rates across regions, and administrative control within regions lends itself to choropleth visualisation.
+Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the the higher prevalence of cancers in older populations.
 Howe [-@HEDP] describes Stocks development of the standardised mortality ratios through the 1930s.
-The choropleth maps used presented levels of cancer via hatchings on a black and white scale.
+The choropleth maps used at this time presented levels of cancer via hatchings on a black and white scale.
 
@@
 @CD considers cartographic data analysis in practice. Presenting a clear guide for preparing cartograms for use with both qualitative and quantitative information.
-The authors recognise that the creation of cartograms was largely the work of professional cartographers until the innovation of geographical information systems that welcomed map users as map creators.
+The authors recognise that the creation of cartograms was largely the work of professional cartographers until the innovation of geographical information systems enabled map users to become map creators.
 These systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’.

### 2019-07-18 review_article.Rmd
start editing public face section

@@
 A grey literature review conducted by (ref) identified 33 publically available cancer atlases published on the internet between 01/01/2010 to 11/11/2015.
-
-All the maps identified in this review were chloropleth maps, in which the representation of the map geometry stayed true to the underlying geography.
-
-<!-- Could be nice to insert a grid of the different maps. Or at least one or two full maps -->
+Reflecting the high use of chloropleth maps founds in the literature, outlined in the section above, all the public facing maps identified in this review were also chloropleth maps.
 

### 2019-07-19 review_article.Rmd
Included table of contents

--- 2019-07-18 review_article.Rmd
@@
 fig_caption: yes
+yolo: true
 biblio-style: vancouver
@@
 Utilising maps to present statistics has been widely used for centuries.
-These maps connec data and/or statistics to the geographical representation of areas that are generally familiar to the audience.
-However, it is not enough to be recognisable, if the meaning of the statistic cannot be seen.
-This is particularly relevant to cancer statistics, as cancer outcomes directly relate to the people living within a geographical area.
-In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer statistics.
-This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes, and health outcomes more braodly.
+Connecting the data to the geographical representation of areas that are already familiar.
+It is not enough to be recognisable, if the value of the statistic cannot be seen.
+Cancer outcomes directly relate to the people living within a geographical area.
+In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer outcomes.
+This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes.
 
@@
 
+<!--
+Disease maps can rapidly highlight geographical
+patterns that would otherwise be hidden in a
+table, and are commonly used for hypothesis
+generating and as the first step in exploratory
+spatial data analysis. @SE
+Geography is concerned with the identification and
+explanation of spatial structure, pattern, and process,
+and with the analysis and explanation of the links
+between humans and the environment (2). @SAMGIS
+
+By providing a visual representation of
+cancer outcomes, geographic patterns of disease are able
+to be identified and effectively addressed
+-->
+
 ## Current best practice map displays for cancer data
@@
 <!-- what is a choropleth map -->
-A choropleth map is used to show the spatial variability of measurements, with usage dating back to the 1800s.
-They are constructed by drawing the given geographic or political boundaries, and filling the polygon with colour to represent values of a measured variable [@EI].
+A choropleth map is used to show the spatial relationship, or variability, of measurements.
+Created as a storage device which preserves locations for geographically ordered data, with usage dating back to the 1800s [@GOINO].
+They are constructed by drawing the given geographic or political boundaries, and filling the shapes with colours to represent values of a measured variable [@EI].
 Early versions used symbols or patterns instead of colour.
 @CPISACA discusses the use of choropleth maps for visualising cancer data, and @DMAHP gives an overview of the development of the use of choropleth maps for displaying disease data.
-A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form, that might show trends in disease occurrence, or even localised outbreaks.
+A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form.
 Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
@@
 <!--
-@GOINO major geographic storage device; map reading and map interpretation have been the basic methods of retrieval and analysis.
-Geography’s interests are in spatial distributions and associations, in areal differentiation
-and in spatial processes. All of these interests demand information that is geographically
-ordered. Geographically ordered data are those data which have their locational
-coordinates preserved.
+Thematic maps, on the other hand, do select some geographical concept or arrangement and attempt to focus attention entirely on the spatial characteristics of the entire distribution and upon its relation to other geographical factors.
 -->
@@
 <!-- psychology -->
-Utilising the familiar state shapes can make a map intuitive to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data.
-The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
+Utilising the familiar state shapes can make maps intuitive to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data.
+The value of the familiarity of the geography is worth consideration when presenting results of spatial analysis. 
 Just as geographers are no longer the only creators of maps, @CPISACA suggests the audience of spatial health data analysis has extended beyond researchers to the public, policymakers and the media.
 However, while the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, different population densities and subject to changes over time.
-The different sizes of administrative areas can attract attention to the shades of the unpopulated, large physical areas [@EI].
-Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].
-
-<!--
-Certain limitations to the standard graphic techniques have also become apparent, however.
-They are slow, and do not permit instant retrieval of information.
-They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually.
-
-These opportunities will be lost if locational coding is not inserted into the data referencing system as such banks are built, however.
-
-of some element, for example population density, or purchasing power, or agricultural fertility; reality, of course, is not like this.
-the background variable assumed to be uniform (such as population density) can be adjusted so that it behaves as if uniformity were the case
--->
+The different sizes of administrative areas can attract undue attention to the shades of the unpopulated, large physical areas [@EI].
+@GOINO believes there are "severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually".
+As the display may draw attention from the 'potentially more important results in the more populous communities' [@SE], choropleths can inhibit visual inference when presenting human related statistics.
 
@@
 - the increasing urbanisation is contributing to the changing density of population
+
+A particular spatial structure includes
+the individuals affected and how they are connected in
+communities, as well as the dynamics of these communities
+and their organization into larger units. @SAMGIS
 -->
@@
 <!-- Displaying cancer data on choropleths -->
+
 In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality. 
-Almost 100 years of cancer maps have explored the United States and the United Kingdom with increased effectiveness in presentation of unbiased rates.
-The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
-The data collection methods of cancer mortality rates across regions, and administrative control within regions lends itself choropleth visualisation.
-Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the proportion of old people.
-Howe [-@HEDP] describes Stocks development of the standardised mortality ratios through the 1930s.
+Almost 100 years of cancer maps have explored the United States and the United Kingdom with increased effectiveness in the presentation of unbiased rates.
+The increasing development and use of disease maps can be attributed to the availability of location data and associated data points [@GOINO], and geographic information system software [@SE].
+Choropleth visualisation suits the data collection methods of census data and cancer mortality rates across regions, and administrative control within regions.
+
+Mortality rates are now often presented as relative rates of risk across the population, and age adjusted.
+@HEDP describes Stocks' development of the standardised mortality ratios through the 1930s.
 The choropleth maps used presented levels of cancer via hatchings on a black and white scale.
@@
 <!--
-Computing mapping methods improved with access to necessary information: locations and associated data points.
-@GOINO discusses progression in computer mapping methods with the availabilty of census data with associated locations.
--->
+@DMAHP cites Cruickshank's (1947) discussion of using visuals as 'formal statistical assessment of the spatial pattern' as a major advancement.
+@MACM also recognises the intuition that can be derived from choropleths must be 'validated by rigorous statistical analyses.'
 
-<!-- Usage and availability -->
-@DMAHP cites Cruickshank's (1947) discussion of using visuals as 'formal statistical assessment of the spatial pattern' as a major advancement.
 @CD considers cartographic data analysis in practice. Presenting a clear guide for preparing cartograms for use with both qualitative and quantitative information.
@@
 These systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’.
-@MACM also recognises the intuition that can be derived from choropleths must be 'validated by rigorous statistical analyses.'
+-->
+
+<!--
+of some element, for example population density, or purchasing power, or agricultural fertility; reality, of course, is not like this.
+the background variable assumed to be uniform (such as population density) can be adjusted so that it behaves as if uniformity were the case
+@GOINO
+-->
 
@@
 
-A grey literature review conducted by (ref) identified 33 publically available cancer atlases published on the internet between 01/01/2010 to 11/11/2015.
+A grey literature review conducted by (ref) identified 33 Cancer Atlases publically available cancer atlases available on the internet and published between 01/01/2010 to 11/11/2015.
 
@@
 
-<!-- Going from choropleth to cartogram -->
+<!-- Going from choropleth to cartogram 
+A cartogram is a map purposely distorted so that its spatial properties represent quantities
+not directly associated with position on the globe. As thematic maps, cartograms emphasize
+the distribution of a variable by changing the area (or lengths) of objects on the map.
+Our approach maintains continuity, and preserves
+many local features of cartographic lines that provide visual clues to the identity of the distorted
+objects.
+@ACCAC
+-->
 
@@
 
- @CTTMB argues that using choropleth maps for population characteristics requires graphic distortions when the population concerned varies greatly in density. 
+<!--
+@SE Researchers mapping spatial patterns of disease
+for small areas are confronted with a number of
+challenges, including cartographic design, the
+modifiable area unit problem, and also the “small
+number” problem, where some places have no,
+or few, disease events and/or a low population at
+risk of contracting the disease.
+
+@TAAM variations in population density may obscure geographic patterns
+This factor may be eliminated by making the map area proportional to population rather than to actual land area.
+Such maps, or what have been termed "population-by-area cartograms," are well known in the
+cartographic literature
+-->
+Using choropleth maps for population characteristics requires graphic distortions when the population concerned varies greatly in density [@CTTMB].
 When implementing a distortion of the geographical shape according to population, an area cartogram [@NAC], or iso-demographic map is the result.
@@
 A contiguous cartogram allows the map space to highlight the distribution of the variable.
-These cartograms maintain connectivity of the map regions while areas are resized according to a statistic, this often must occur at the expense of the shape [@CBATCC,@NAC].
+These cartograms maintain connectivity of the map regions while areas are resized according to a statistic, this often must occur at the expense of the shape [@CBATCC,@NAC, @TAAM].
 From a computer graphics perspective, @ACA believe it is a problem of 'map deformation' to account for the value assigned to each area, they provide three methods for creating value-by-area cartograms.
 Examples include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [-@ACTUC], Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and Tikunov’s Line Integral Method, Constraint-Based Method (Kocmoud and House) [-@CBATCC].
-- compare to their
+
+<!--
+Medical application: TAAM 1965
+Actual shapes of counties were sacrificed to the extent necessary to retain as far as possible, common boundaries, relative locations, and the approximate contour of the State.
+
+There is no unique solution to the topologic
+problem of preparing a population-by-area
+cartogram for a given population and its jurisdictional
+boundaries. Although the map areas
+of the various regions are fixed by their popu¬
+lations, their shapes and relative locations are
+not completely determined.
+
+Other characteristics of the population
+may also serve as the base for adjustment; for
+example, sex, race, parity, socioeconomic level,
+or iminune status.
+
+@RSCW
+helps to correct misconceptions.
+Pupils are likely to over-estimate regions as Canada, Siberia, Africa, and our own West just because they loom up large on maps.
+The simple picture which the cartogram gives is easily comprehended and remembered and
+will be stimulating in making comparisons.
+
+@NISCC
+two advantages of the cartogram: (1) new relationships are often revealed visually, and ( 2 ) it is often more honest to use “other” space on the base map than land area.
+Finally, and most important for the present discussion, they suggest that the ability to preserve shape of the statistical units drops with an increase in the number of these units, and in situations where the difference is great between the smallest and largest values mapped.
+Until we know how the map reader compares it to an ordinary thematic map, we are experimenting with a graphic device that may be reducing (or at least hindering) the flow of information from the map to the map reader.
+-->
 
@@
 
+<!--
+-simple shapes of the circle, square, and triangle are relatively easily recognized.
+-sacrifice communication efficiency;
+-There is evidence, too, that angular shapes are more easily seeu than ones with smooth outlines.
+Shapes that contain the qualities of simplicity and reg,ularity are more readily perceived than others that do not possess these qualities.
+-Finally, psychologists tell us that shapes having familiarity are more quickly
+recognized than nonsense shapes or those that have no meaning.
+-The cartographer can oiily efficiently communicate with this kind of map if he has preserved particular elements of the original outline and incorporated them in the new shape.
+-
+@NISCC 
+-->
+
 @ACTUC suggests 'population distribution is often extremely uneven in former British colonies'.
@@
 This gives equal emphasis to every area, allows distributions and relationships between neighbours to become more clear.
+
+<!-- 
+TVSSS parliamentary cartogram where each constituency is represented by an oval of equal size. Many ovals overlap because the cartogram is still in the process of being made. Each
+-->
 

### 2019-07-19 review_article.Rmd

@@
 - presents alternative ways of visualising those spatial patterns, or the uncertainty in those patterns
-- a more applied-focussed paper (eg lots of cool images), rather than theoretical (eg lots of formula).
+- a more applied-focussed paper (e.g. lots of cool images), rather than theoretical (e.g. lots of formula).
 -->
@@
 Utilising maps to present statistics has been widely used for centuries.
-Connecting the data to the geographical representation of areas that are already familiar.
-It is not enough to be recognisable, if the value of the statistic cannot be seen.
-Cancer outcomes directly relate to the people living within a geographical area.
-In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer outcomes.
-This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes.
+These maps connect data and/or statistics to the geographical representation of areas that are generally familiar to the audience.
+However, it is not enough to be recognisable, if the meaning of the statistic cannot be seen.
+This is particularly relevant to cancer statistics, as cancer outcomes directly relate to the people living within a geographical area.
+In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer statistics.
+This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes, and health outcomes more broadly.
 
@@
 <!-- psychology -->
-Utilising the familiar state shapes can make maps intuitive to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data.
-The value of the familiarity of the geography is worth consideration when presenting results of spatial analysis. 
-Just as geographers are no longer the only creators of maps, @CPISACA suggests the audience of spatial health data analysis has extended beyond researchers to the public, policymakers and the media.
-However, while the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, different population densities and subject to changes over time.
-The different sizes of administrative areas can attract undue attention to the shades of the unpopulated, large physical areas [@EI].
-@GOINO believes there are "severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually".
-As the display may draw attention from the 'potentially more important results in the more populous communities' [@SE], choropleths can inhibit visual inference when presenting human related statistics.
+
+Utilising the familiar state boundaries can make a map intuitive to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data, i.e.
+how cancer rate differs across the states.
+The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
+Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
+However, while the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
+The different population and geographical sizes of administrative areas can attract attention to the shades of the unpopulated but large areas [@EI].
+Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].
+
+<!--
+Certain limitations to the standard graphic techniques have also become apparent, however.
+They are slow, and do not permit instant retrieval of information.
+They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually.
+
+These opportunities will be lost if locational coding is not inserted into the data referencing system as such banks are built, however.
+
+of some element, for example population density, or purchasing power, or agricultural fertility; reality, of course, is not like this.
+the background variable assumed to be uniform (such as population density) can be adjusted so that it behaves as if uniformity were the case
+-->
 
@@
 In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality.
-Almost 100 years of cancer maps have explored the United States and the United Kingdom with increased effectiveness in the presentation of unbiased rates.
-The increasing development and use of disease maps can be attributed to the availability of location data and associated data points [@GOINO], and geographic information system software [@SE].
-Choropleth visualisation suits the data collection methods of census data and cancer mortality rates across regions, and administrative control within regions.
+
+Almost 100 years of cancer mapping in the United States and the United Kingdom has increased the effectiveness in the presentation of unbiased rates.
+The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
+The data collection methods of cancer mortality rates across regions, and administrative control within regions lends itself to choropleth visualisation.
+Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the the higher prevalence of cancers in older populations.
+Howe [-@HEDP] describes Stocks development of the standardised mortality ratios through the 1930s.
+The choropleth maps used at this time presented levels of cancer via hatchings on a black and white scale.
 
@@
 
-A grey literature review conducted by (ref) identified 33 Cancer Atlases publically available cancer atlases available on the internet and published between 01/01/2010 to 11/11/2015.
-
-All the maps identified in this review were chloropleth maps, in which the representation of the map geometry stayed true to the underlying geography.
-
-<!-- Could be nice to insert a grid of the different maps. Or at least one or two full maps -->
+A grey literature review conducted by (ref) identified 33 publicly available cancer atlases published on the internet between 01/01/2010 to 11/11/2015.
+Reflecting the high use of chloropleth maps founds in the literature, outlined in the section above, all the public facing maps identified in this review were also chloropleth maps.
 
@@
 The review did not reveal any novel uncertainty visualisation approaches or visualisations.
-Maps used standard and well known measures including credible intervals and standard deviation, statistical significance, boxplots and distributions.
+Maps used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
 These maps ranged from static pdfs or infographics to interactive online resources.
@@
 <!-- Why transform?-->
-<!-- Common variables used to create cartograms, eg population, mortality -->
+<!-- Common variables used to create cartograms, e.g. population, mortality -->
 Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
@@
 The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC].
-Cartograms may be seen as an extension of map tranformations and projections. The favourable distortion is proportional to a value other other than actual earth size area [@NAC].
+Cartograms may be seen as an extension of map transformations and projections. The favourable distortion is proportional to a value other other than actual earth size area [@NAC].
 
@@
 The creation of cartograms was largely in the hands of cartographers. 
-@ACTUC discusses early approaches including John Hunter and Johnathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball bearing approach and Tobler's (1973) computer programs.
+@ACTUC discusses early approaches including John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball bearing approach and Tobler's (1973) computer programs.
 
@@
 
-There is no unique solution to the topologic
+There is no unique solution to the topology
 problem of preparing a population-by-area
@@
 boundaries. Although the map areas
-of the various regions are fixed by their popu¬
-lations, their shapes and relative locations are
+of the various regions are fixed by their populations, their shapes and relative locations are
 not completely determined.
@@
 may also serve as the base for adjustment; for
-example, sex, race, parity, socioeconomic level,
-or iminune status.
+example, sex, race, parity, socio-economic level,
+or immune status.
 
@@
 <!--
-[-@NAC] observation units stay in more or less their correct location and maintain correct shape, but at least some units are separated by gaps from their neighbors.
+[-@NAC] observation units stay in more or less their correct location and maintain correct shape, but at least some units are separated by gaps from their neighbours.
 Each area in a non-contiguous cartogram is scaled down or up according to value for the specific area.
@@
 -sacrifice communication efficiency;
--There is evidence, too, that angular shapes are more easily seeu than ones with smooth outlines.
-Shapes that contain the qualities of simplicity and reg,ularity are more readily perceived than others that do not possess these qualities.
+-There is evidence, too, that angular shapes are more easily seen than ones with smooth outlines.
+Shapes that contain the qualities of simplicity and regularity are more readily perceived than others that do not possess these qualities.
 -Finally, psychologists tell us that shapes having familiarity are more quickly
 recognized than nonsense shapes or those that have no meaning.
--The cartographer can oiily efficiently communicate with this kind of map if he has preserved particular elements of the original outline and incorporated them in the new shape.
+-The cartographer can efficiently communicate with this kind of map if he has preserved particular elements of the original outline and incorporated them in the new shape.
 -
@@
 @TFYCC Value-Area Cartograms.
-In these cartograms a region,country, or continent is subdivided into small regions, each ofwhich is represented by a rectangle.
+In these cartograms a region,country, or continent is subdivided into small regions, each of which is represented by a rectangle.
 This rectangle is proportionate in area to the value which it represents in certain statistical distributions.
@@
 
-With that in mind, the intented user and message to communicate should drive map selection.
+With that in mind, the intended user and message to communicate should drive map selection.
 first is the nature or inherent properties of the device itself
@@
 
-@ECGC ?? Highlight the value of animating contigous to see changes over time, US can be recognisable but animation aides interpretation
+@ECGC ?? Highlight the value of animating contiguous to see changes over time, US can be recognisable but animation aides interpretation
 
@@
 
-What software did we use, for eg
+What software did we use, for e.g.
 

### 2019-07-19 review_article.Rmd
edit intro to public face section.

@@
 output: 
+  html_document: default
   pdf_document:
@@
     number_sections: true
-  html_document: default
 fontsize: 11pt
@@
 
-A grey literature review conducted by (ref) identified 33 publicly available cancer atlases published on the internet between 01/01/2010 to 11/11/2015.
-Reflecting the high use of chloropleth maps founds in the literature, outlined in the section above, all the public facing maps identified in this review were also chloropleth maps.
-
-### Publishers {-}
-
-The majority of atlases were published by non-commercial organisations, including not-for-profits (NFPs), government, research organisations, advocacy groups or a partnership between an NFP & government.
-Only one map was published by a commercial entity (Maps of Cancer Mortality Rates in Spain), in this case a media organisation El Pais^[https://elpais.com/].
-
-### Reported measures {-}
-
-The majority of maps identified in this study reported age adjusted rates of either incidence, mortality or survival.
-The list below summarises these measures, and provides a definition of each measure.
+Cancer maps are effective communication tools for a general or non-expert audience and are commonly used in the public domain to communicate the output of sophisticated statistical analyses to these audiences.
+The heavy use of chloroplet maps within the research literature is reflected in the types of maps that are found in the public domain.
+A grey literature review conducted by (ref) identified 33 cancer atlases published on the internet between 01/01/2010 to 11/11/2015 all of which were chloropleth maps.
+These maps were mostly published by non-commercial organisations, including not-for-profits (NFPs), government, research organisations, advocacy groups or a partnership between an NFP & government.
+Only one map was published by a commercial entity (Maps of Cancer Mortality Rates in Spain), published by the media organisation El Pais^[https://elpais.com/].
 
@@
 
-\newpage
+### Reported measures {-}
+
+The majority of maps identified in this study reported age adjusted rates of either incidence, mortality or survival.
+The list below summarises these measures, and provides a definition of each measure.
+
 Table: (\#tab:measures) Measures used to report cancer statistics

### 2019-07-19 review_article.tex
edit intro to public face section.

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Cartogram Mapping and its Application to Cancer Data Visualisation},
            pdfauthor={Stephanie Kobakian, Jessie Roberts and Dianne Cook},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Cartogram Mapping and its Application to Cancer Data Visualisation}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Stephanie Kobakian, Jessie Roberts and Dianne Cook}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\section*{Abstract}

Utilising maps to present statistics has been widely used for centuries.
These maps connect data and/or statistics to the geographical
representation of areas that are generally familiar to the audience.
However, it is not enough to be recognisable, if the meaning of the
statistic cannot be seen. This is particularly relevant to cancer
statistics, as cancer outcomes directly relate to the people living
within a geographical area. In situations where the people are of
interest, not the land they live on, it is reasonable to explore views
that enhance the communication of the cancer statistics. This has
spurred innovations over the previous centuries to enhance the shapes
and maps presented to effectively communicate cancer outcomes, and
health outcomes more broadly.

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{disease-mapping-methods}{%
\section{Disease mapping methods}\label{disease-mapping-methods}}

\hypertarget{current-best-practice-map-displays-for-cancer-data}{%
\subsection{Current best practice map displays for cancer
data}\label{current-best-practice-map-displays-for-cancer-data}}

A choropleth map is used to show the spatial relationship, or
variability, of measurements. Created as a storage device which
preserves locations for geographically ordered data, with usage dating
back to the 1800s (Berry, Morrill, and Tobler,
\protect\hyperlink{ref-GOINO}{n.d.}). They are constructed by drawing
the given geographic or political boundaries, and filling the shapes
with colours to represent values of a measured variable (Tufte
\protect\hyperlink{ref-EI}{1990}). Early versions used symbols or
patterns instead of colour. Bell et al.
(\protect\hyperlink{ref-CPISACA}{2006}) discusses the use of choropleth
maps for visualising cancer data, and Walter
(\protect\hyperlink{ref-DMAHP}{2001}) gives an overview of the
development of the use of choropleth maps for displaying disease data. A
choropleth map is a true map of the topology, constructed for visual
inspection of spatial patterns across a familiar geographic form. Figure
1 shows a choropleth map where each state of the United States of
America is coloured by the average annual rate of new cases of lung and
bronchus cancer from years 2012 to 2016.

\begin{figure}
\centering
\includegraphics{figures/ggchoro.png}
\caption{``A choropleth map of the United States of America.''}
\end{figure}

Utilising the familiar state boundaries can make a map intuitive to read
(Brewster and Subramanian \protect\hyperlink{ref-CIBMUK}{2010}), and
allow viewers to visually infer the spatial relationships in the data,
i.e.~how cancer rate differs across the states. The familiarity of the
geography is a worthy consideration when presenting results of spatial
analysis. Just as geographers are no longer the only creators of maps,
Bell et al. (\protect\hyperlink{ref-CPISACA}{2006}) suggests the
audiences of spatial health data analysis have extended beyond
researchers to the public, policymakers and the media. However, while
the areas are recognisable shapes, they are often politically driven
boundaries with individual areas being of non-uniform size, containing
different population densities and subject to change over time. The
different population and geographical sizes of administrative areas can
attract attention to the shades of the unpopulated but large areas
(Tufte \protect\hyperlink{ref-EI}{1990}). Choropleths can inhibit visual
inference when presenting human related statistics as the display may
draw attention from the `potentially more important results in the more
populous communities' (Exeter \protect\hyperlink{ref-SE}{2017}).

In epidemiology, choropleths are often used as a tool to study the
spatial distribution of cancer incidence and mortality.

Almost 100 years of cancer mapping in the United States and the United
Kingdom has increased the effectiveness in the presentation of unbiased
rates. The increasing development and use of disease maps can be
attributed to the availability of geographic information system software
(Exeter \protect\hyperlink{ref-SE}{2017}). The data collection methods
of cancer mortality rates across regions, and administrative control
within regions lends itself to choropleth visualisation. Mortality rates
are now often presented as relative rates of risk across the population,
and age adjusted to correct for the the higher prevalence of cancers in
older populations. Howe (\protect\hyperlink{ref-HEDP}{1989}) describes
Stocks development of the standardised mortality ratios through the
1930s. The choropleth maps used at this time presented levels of cancer
via hatchings on a black and white scale.

Mortality rates are now often presented as relative rates of risk across
the population, and age adjusted. Howe
(\protect\hyperlink{ref-HEDP}{1989}) describes Stocks' development of
the standardised mortality ratios through the 1930s. The choropleth maps
used presented levels of cancer via hatchings on a black and white
scale.

Examples of atlases: Kraak 1998, Kraak and Ormeling 2011 Bertin 1967

\hypertarget{public-face}{%
\subsection{Public face}\label{public-face}}

A grey literature review conducted by (ref) identified 33 publicly
available cancer atlases published on the internet between 01/01/2010 to
11/11/2015. Reflecting the high use of chloropleth maps founds in the
literature, outlined in the section above, all the public facing maps
identified in this review were also chloropleth maps.

\hypertarget{publishers}{%
\subsubsection*{Publishers}\label{publishers}}
\addcontentsline{toc}{subsubsection}{Publishers}

The majority of atlases were published by non-commercial organisations,
including not-for-profits (NFPs), government, research organisations,
advocacy groups or a partnership between an NFP \& government. Only one
map was published by a commercial entity (Maps of Cancer Mortality Rates
in Spain), in this case a media organisation El Pais\footnote{\url{https://elpais.com/}}.

\hypertarget{reported-measures}{%
\subsubsection*{Reported measures}\label{reported-measures}}
\addcontentsline{toc}{subsubsection}{Reported measures}

The majority of maps identified in this study reported age adjusted
rates of either incidence, mortality or survival. The list below
summarises these measures, and provides a definition of each measure.

\hypertarget{geographical-coverage}{%
\subsubsection*{Geographical coverage}\label{geographical-coverage}}
\addcontentsline{toc}{subsubsection}{Geographical coverage}

Identified cancer atlases covered geographies from all around the world:
4 were global, 3 from Australia (AUS), 11 from the United States (US), 2
from Canada (CAN), 7 from the United Kingdom (UK), 2 from Spain, 1 from
Switzerland, 1 from Germany, 1 from Norway, and 1 covering the European
Union. Not all maps had a national focus and 10 covered a region or
state rather than an entire nation. The states or counties/regions
covered were South Australia (AUS), Queensland (AUS), Ontario (CAN),
Valencia (Spain), Pennsylvania county Massachusetts (US), New Hampshire
(US), Cape Cod (US), Missouri (US), Florida (US), New York State (US)
and Arizona (US).

\newpage

\begin{longtable}[]{@{}ll@{}}
\caption{(\#tab:measures) Measures used to report cancer
statistics}\tabularnewline
\toprule
\begin{minipage}[b]{0.28\columnwidth}\raggedright
Measure\strut
\end{minipage} & \begin{minipage}[b]{0.66\columnwidth}\raggedright
Details\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.28\columnwidth}\raggedright
Measure\strut
\end{minipage} & \begin{minipage}[b]{0.66\columnwidth}\raggedright
Details\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.28\columnwidth}\raggedright
1. IR (Incidence Ratio)\strut
\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright
\((IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}\),\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.28\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright
Cancer incidence rate in region \(i\) over the average cancer incidence
rate for the total region\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.28\columnwidth}\raggedright
2. SIR (Standardised Incidence Ratio)\strut
\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright
IR standardised by age structure in each region \(i\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.28\columnwidth}\raggedright
3. RER\strut
\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright
\(RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.28\columnwidth}\raggedright
(Relative Excess Risk)\strut
\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright
Represents the estimate of cancer related mortality within five years of
diagnosis\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.28\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright
Also referred to as `excess hazard ratio'\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.28\columnwidth}\raggedright
4. Age Adjusted Relative Risk\strut
\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright
RR standardised by age structure in each region \(i\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.28\columnwidth}\raggedright
5. Rate per 100,000\strut
\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright
Cancer incidence per 100,000 population\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.28\columnwidth}\raggedright
6. Age Adjusted Rate per 100,000\strut
\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright
\#5 standardised by age structure or region\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.28\columnwidth}\raggedright
7. New cancer cases per 100,000\strut
\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright
Specific methods could not be found\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.28\columnwidth}\raggedright
8. Count\strut
\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright
Crude cancer counts\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.28\columnwidth}\raggedright
9. Below or above Expected\strut
\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright
Alternative expression of the SIR\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{statistical-uncertainty}{%
\subsubsection{Statistical uncertainty}\label{statistical-uncertainty}}

Cancer atlases were considered to report uncertainty to the non-expert
user if they included a measure of statistical uncertainty either within
or alongside the map. Maps that only reported this information within
the supplementary material were not considered to have directly
attempted to report uncertainty.

The review did not reveal any novel uncertainty visualisation approaches
or visualisations. Maps used standard and well known measures including
credible intervals and standard deviation, statistical significance, box
plots and distributions. These maps ranged from static pdfs or
infographics to interactive online resources. The interactivity of the
more modern maps enabled uncertainty information to be incorporated
without cluttering the screen, such as in a tool tip feature.

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{figures/Alberta_CI_viz}

}

\caption{Example of CI visualisation for uncertainty representation in cancer mapping (1/3). Source: Alberta Health IHDA Geographic. (2012) }\label{fig:ci-viz2}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.55\linewidth]{figures/Pennsylvania_CI_viz}

}

\caption{Example of CI visualisation for uncertainty representation in cancer mapping (2/3). Source: Pensylvania Cancer Atlas }\label{fig:ci-viz3}
\end{figure}

Close to half of the atlases identified (42\%, n=14) included some
measure of uncertainty. The most common measure used to represent
uncertainty were credible or confidence intervals (CIs). CIs were either
visualised by including their bounds in a scatterplot or graph of
estimates vs region (see Figures @ref(fig:ci-viz2)\footnote{Alberta
  Health IHDA Geographic. (2012). Age-Standardised Incidence Rate of
  COPD, 2011. Retrieved from:
  \url{http://www.health.alberta.ca/health-info/IHDA-geographic/COPD/incidence-agestandard/atlas.html?epik=0GJSpE_IW34lx}},
and @ref(fig:ci-viz3)\footnote{Centres for Disease Control and
  Prevention (CDC). (n.d). United States Cancer Statistics: An
  Interactive Cancer Atlas (InCA). Retrieved
  from:\url{https://nccd.cdc.gov/DCPC_INCA/}} positioned next to the
map, or reported numerically through the CI upper and lower bounds
listed in a data table (see Figure 3.4\footnote{Pennsylvania Cancer
  Atlas. (n.d). Retrieved from:
  \url{https://www.geovista.psu.edu/grants/CDC/?epik=0lJSpE_IW34lx}}).
Of those that visualised the CIs, 30\% (n=10) embedded the visualisation
within a tool tip function which visualised the CI when the mouse
hovered over the relevant area (see Figure 3.5\footnote{International
  Agency for Research on Cancer. (2017). Atlas of Cancer Mortality in
  the European Union and European Economic Area 1993-1997, Annex 4 -
  Cancer mortality maps by site. Retrieved from:
  \url{http://www.iarc.fr/en/publications/pdfs-online/epi/sp159/}}).

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{figures/CDC_CI_viz_tooltip}

}

\caption{Example of CI visualisation for uncertainty representation in cancer mapping (3/3). Source: Centres for Disease Control and Prevention (CDC). United States Cancer Statistics: An Interactive Cancer Atlas (InCA) }\label{fig:ci-tooltip}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{figures/uncert_table}

}

\caption{Example of an interactive data table with CI upper and lower bounds used in cancer mapping. Source: Pennsylvania Cancer Atlas}\label{fig:ci-table}
\end{figure}

Methods for representing sources of uncertainty information can be
visualised or communicated in different ways, examples identified
through this grey literature review are listed below.

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{figures/standard_deviation_large}

}

\caption{Example of standard deviation visualised in cancer mapping}\label{fig:stndvn-example}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.4\linewidth]{figures/Boxplot2}

}

\caption{Example of boxplot used in cancer mapping. Source: Source: Pensylvania Cancer Atlas }\label{fig:boxplot-example}
\end{figure}

\begin{longtable}[]{@{}ll@{}}
\caption{(\#tab:method-exp) Implicit and explicit measures of
uncertainty.}\tabularnewline
\toprule
\begin{minipage}[b]{0.35\columnwidth}\raggedright
Measure\strut
\end{minipage} & \begin{minipage}[b]{0.59\columnwidth}\raggedright
Example\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.35\columnwidth}\raggedright
Measure\strut
\end{minipage} & \begin{minipage}[b]{0.59\columnwidth}\raggedright
Example\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.35\columnwidth}\raggedright
CI Interval\strut
\end{minipage} & \begin{minipage}[t]{0.59\columnwidth}\raggedright
Figures 3.1, 3.2, 3.3\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Statistical Significance\strut
\end{minipage} & \begin{minipage}[t]{0.59\columnwidth}\raggedright
Textured overlay on top of coloured regions used to indicate statistical
significance\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Distribution\strut
\end{minipage} & \begin{minipage}[t]{0.59\columnwidth}\raggedright
Figure 3.4\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Boxplots\strut
\end{minipage} & \begin{minipage}[t]{0.59\columnwidth}\raggedright
Figures 3.4, 3.5\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Sample Size\strut
\end{minipage} & \begin{minipage}[t]{0.59\columnwidth}\raggedright
Textured overlay or lack of colour on a region, was used to show regions
with small sample size\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Standard deviation\strut
\end{minipage} & \begin{minipage}[t]{0.59\columnwidth}\raggedright
Figure 3.6 - the second map in the bottom right corner shows standard
deviation\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{cartograms}{%
\subsection{Cartograms}\label{cartograms}}

Choropleths may be considered true topological maps, however, if the
land mass presented covers enough of the globe, there must be a
transformation or distortion to display in 2D. The amount of distortion
is related to the distance displayed Tobler
(\protect\hyperlink{ref-GAMP}{1963}). The world projections reflect the
frequent distortions seen from altering perspectives. Choropleth maps
will always be distorted if they cover enough of the globe, as will
photographs of the globe from space. Choropleth creation requires
choosing a map projection that shows a favourable distortion of the
geography for presenting the set of spatial information. Diagrams that
do not specify a projection can be considered to have some unknown
projection.

\begin{figure}
\centering
\includegraphics{figures/ggchoroCRS.png}
\caption{``Choropleth maps of the United States of America using four
coordinate reference systems.''}
\end{figure}

Waldo Tobler (\protect\hyperlink{ref-TVSSS}{2012}) explores many
graphical techniques, and suggests there are particular methods for
particular purposes. Cartograms provide an alternative visualisation
method for statistical and geographical information. The key difference
between a choropleth and a cartogram is the desirable augmentation of
the size, shape or distance of geographical areas (Dorling
\protect\hyperlink{ref-ACTUC}{2011}). Cartograms may be seen as an
extension of map transformations and projections. The favourable
distortion is proportional to a value other other than actual earth size
area (Olson \protect\hyperlink{ref-NAC}{1976}).

Using choropleth maps for population characteristics requires graphic
distortions when the population concerned varies greatly in density
(Griffin \protect\hyperlink{ref-CTTMB}{1980}). When implementing a
distortion of the geographical shape according to population, an area
cartogram (Olson \protect\hyperlink{ref-NAC}{1976}), or iso-demographic
map is the result. When visualising population statistics Dorling
(\protect\hyperlink{ref-ACTUC}{2011}) considers this equitable
representation design `more socially just, giving due attention to all
members of the population and reducing the visual impact of large areas
with small populations (Walter \protect\hyperlink{ref-DMAHP}{2001}).
Griffin (\protect\hyperlink{ref-CTTMB}{1980}) suggest that spatial
socio-economic data is best presented on a cartogram for urban areas.
Howe (\protect\hyperlink{ref-HEDP}{1989}) agrees that 'cancer occurs in
people, not in geographical areas' and the map bases of population
reflect this and avoid allocating `undue prominence' to rural areas.
Jahan et al. (\protect\hyperlink{ref-MTMSIH}{2018}) encourage the use of
cartograms to highlight small areas and uncover local-level
inequalities. Using health estimates from large areas can prevent
drawing attention to the inequalities.

The area on the map space can also be used represent a value. There have
been many algorithms presented, Nusrat and Kobourov
(\protect\hyperlink{ref-SAIC}{2016}) provided a framework to investigate
implementations and the ``statistical accuracy, geographical accuracy,
and topological accuracy''. The spatial transformation of map regions
relative to the data emphasises the data distribution instead of land
size (Kocmoud and House \protect\hyperlink{ref-CBATCC}{1998}).

Cancer statistics and atlases

The creation of cartograms was largely in the hands of cartographers.
Dorling (\protect\hyperlink{ref-ACTUC}{2011}) discusses early approaches
including John Hunter and Jonathan Young (1968) and Durham's wooden tile
method, Skoda and Robertson's (1972) steel ball bearing approach and
Tobler's (1973) computer programs.

There are many alternatives to consider, the intended audience of the
map, and its purpose are key points in cartogram use and creation.
Dorling (\protect\hyperlink{ref-ACTUC}{2011}) reiterates: `There is no
``best'' cartogram or method of creating cartograms just as there is no
``best'' map' (Monmonier and Schnell, 1988).

`Firstly, that no such transformation has a unique solution, leaving the
subjective choice in the hands of the individual map designer.'

Design goals: - to maintain as much as possible of the intricate detail
- to maximize the simplicity of the transformation - maintenance of
directional accuracy - attempt to influence and direct the map user's
attention,

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  that the size of the electoral subdivisions should be made
  proportional to their respective electoral populations,
\item
  that the contiguity relationships between the DCU's should be
  maintained,
\item
  that true directions should be retained from the centre of the
  transformation,
\item
  that the main shape characteristics of the DCU's should be preserved,
  and
\item
  that the map should possess the overall shape of the built-up area.
\end{enumerate}

\hypertarget{contiguous}{%
\subsubsection{Contiguous}\label{contiguous}}

\begin{figure}
\centering
\includegraphics{figures/ggcont.png}
\caption{``A contiguous cartogram of the Unites States of America''}
\end{figure}

A contiguous cartogram allows the map space to highlight the
distribution of the variable. These cartograms maintain connectivity of
the map regions while areas are resized according to a statistic, this
often must occur at the expense of the shape (Kocmoud and House
\protect\hyperlink{ref-CBATCC}{1998},@NAC, @TAAM). From a computer
graphics perspective, Min Ouyang and Revesz
(\protect\hyperlink{ref-ACA}{2000}) believe it is a problem of `map
deformation' to account for the value assigned to each area, they
provide three methods for creating value-by-area cartograms. Examples
include Tobler's Pseudo-Cartogram Method, Dorling's Cellular Automaton
Method (\protect\hyperlink{ref-ACTUC}{2011}), Radial Expansion Method of
Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and
Tikunov's Line Integral Method, Constraint-Based Method (Kocmoud and
House) (\protect\hyperlink{ref-CBATCC}{1998}).

describes the process of creating a population cartogram for Canada,
using the 1966 Census. An intentional goal of this project was to
maintain contiguity, while attempting to keep the actual shape of
places. The end result was a `very accurate isodemo-graphic map of
Canada'. This intentional design goal coincided with the rising interest
in urban geography.

To be able to recognise the significant changes, a reader will usually
have to know the initial geography to find the differences in the new
cartogram layout (Olson \protect\hyperlink{ref-NAC}{1976}). Tobler's
Conformal mapping means to preserve angles locally so that the shapes of
very small areas on a traditional map and a cartogram would be similar.
Kocmoud and House (\protect\hyperlink{ref-CBATCC}{1998}) presents this
issue as conflicting tasks or aims, to adjust region sizes and retain
region shapes. Distortion of region shapes on the contiguous cartogram
presents an additional hurdle to visual recognition and this hurdle is
not only eliminated on the noncontiguous cartogram but is replaced by
the meaningful empty-space property (Olson
\protect\hyperlink{ref-NAC}{1976}, @ECGC).

\hypertarget{non-contiguous}{%
\subsubsection{Non-Contiguous}\label{non-contiguous}}

\begin{figure}
\centering
\includegraphics{figures/ggncont.png}
\caption{``A Non - contiguous cartogram of the Unites States of
America''}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/ggdorl.png}
\caption{``A dorling cartogram of the Unites States of America''}
\end{figure}

Dorling (\protect\hyperlink{ref-ACTUC}{2011}) puts forward a simple
question:

\begin{quote}
If, for instance, it is desirable that areas on a map have boundaries
which are as simple as possible, why not draw the areas as simple shapes
in the first place?
\end{quote}

He answers this with his implementation of maps created with `the
simplest of all shapes'. While contiguous cartograms may be a `more
sophisticated' method, they produce `very complex shapes'. Circular
cartograms use the same shape for every region represented, and size
them according to the statistic represented or the population for a base
map. To produce a compelling map, a gravity model is applied to avoid
overlaps, and keep spatial relationships with neighbouring areas over
many iterations. This implementation can work for up to `one hundred
thousand' areas.

Dorling (\protect\hyperlink{ref-ACTUC}{2011}) suggests `population
distribution is often extremely uneven in former British colonies'.

`In Australia the urban federal constituencies occupy only a tenth of
the land, but contain nine tenths of the people. It would be almost
unthinkable to show election results for that country on a conventional
equal land area map.' This 1966 cartogram uses mostly straight lines,
and the result looks very little like the geographical shape of
Australia.

`Given the increasingly uneven population distribution of the United
States and the growing social divides between the populations of
neighbourhoods living at different densities, the need for cartograms
like this is greater now than ever.'

Used in displays of the UK by Howe in 1986 cited by Howe
(\protect\hyperlink{ref-HEDP}{1989})

Tobler's method and the many implementations that `elaborated' on it are
derived from `numerical approximations to a pair of equations'(Dorling
\protect\hyperlink{ref-ACTUC}{2011}). They all operate through
incremental adjustments, and can produce wildly different outcomes from
small changes in the inputs.

Tobler (\protect\hyperlink{ref-TFYCC}{2004}) Value-Area Cartograms. In
these cartograms a region,country, or continent is subdivided into small
regions, each of which is represented by a rectangle. This rectangle is
proportionate in area to the value which it represents in certain
statistical distributions. The regions are grouped in approximately the
same positions as they are on the map.

Computer generated map examples: Howe
(\protect\hyperlink{ref-HEDP}{1989}) (Hopps et al.~1968; Armstrong
1972). There has followed a flood of disease atlases, mainly
concentrating on the modem problems of cancer and degenerative diseases
from countries as scattered as the United States (Burbank 1971; Mason et
al.~1975, 1976; Pickle et al.~1987), the Soviet Union (Levin 1980),
Japan (Shigematsu 1977), the Federal Republic of Germany

Cano et al. (\protect\hyperlink{ref-MDAC}{2015}) define the term `mosaic
cartograms'. Compare amount of tiles to contrast population of regions.
`Cartograms show a data value per input region by scaling each region
such that its area is proportional to its data value. Mosaic cartograms
show data in multiples of tiles, hence the input data must consist of,
or be cast into, small integer units.'

\hypertarget{centroid-displays}{%
\subsection{Centroid displays}\label{centroid-displays}}

Dot plot: one dot for each region, coloured, and placed at centroid.

Olson (\protect\hyperlink{ref-NAC}{1976}) gives an example of

Plotting centroids on top of geographies. (Size is kept constant)

Replies on the idea that every area is important, no matter the size.
This gives equal emphasis to every area, allows distributions and
relationships between neighbours to become more clear.

\hypertarget{a-critique-of-mapping-methods}{%
\section{A critique of mapping
methods}\label{a-critique-of-mapping-methods}}

\begin{quote}
designing a map tailored to precise goals {[}is{]} easier than forcing a
single map to accommodate diverse objectives - Bell et al.
(\protect\hyperlink{ref-CPISACA}{2006})
\end{quote}

With that in mind, the intended user and message to communicate should
drive map selection. first is the nature or inherent properties of the
device itself second is the ease and accuracy with which map users are
able to extract information from it

Moore and Carpenter (\protect\hyperlink{ref-SAMGIS}{1999}) suggests it
is the ``investigators' objectives'' that drive the ``representation of
diseases on maps''. -infectious diseases may require choropleth methods

`Where control of the message is important, static maps will continue to
be the most effective, although good tables, graphs, and explanatory
text are still needed in order to ensure that different people will see
the same thing in the maps' Bell et al.
(\protect\hyperlink{ref-CPISACA}{2006})

(only worth including if it is possible for us to implement) Tabular
form comparing and contrasting - Relationship to geography

\begin{itemize}
\tightlist
\item
  Show using cancer examples
\end{itemize}

\hypertarget{animation}{%
\section{Animation}\label{animation}}

lends to temporal pattern exploration

Keim et al. (\protect\hyperlink{ref-ECGC}{2002}) ?? Highlight the value
of animating contiguous to see changes over time, US can be recognisable
but animation aides interpretation

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

What software did we use, for e.g.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-CPISACA}{}%
Bell, B Sue, Richard E Hoskins, Linda Williams Pickle, and Daniel
Wartenberg. 2006. ``Current Practices in Spatial Analysis of Cancer
Data: Mapping Health Statistics to Inform Policymakers and the Public.''
\emph{International Journal of Health Geographics} 5: 49.
\url{https://doi.org/10.1186/1476-072X-5-49}.

\leavevmode\hypertarget{ref-GOINO}{}%
Berry, Brian J. L., Richard L. Morrill, and Waldo R. Tobler. n.d.
``GEOGRAPHIC Ordering of Information: NEW Opportunities'' 16 (4).
Association of American Geographers: 39--44.
\url{https://doi.org/10.1111/j.0033-0124.1964.039_q.x}.

\leavevmode\hypertarget{ref-CIBMUK}{}%
Brewster, Mark B., and S. V. Subramanian. 2010. ``Cartographic Insights
into the Burden of Mortality in the United Kingdom: A Review of `The
Grim Reaper's Road Map'.'' Journal Article. \emph{International Journal
of Epidemiology} 39 (4): 1120--2.
\url{https://doi.org/10.1093/ije/dyp395}.

\leavevmode\hypertarget{ref-MDAC}{}%
Cano, R. G., K. Buchin, T. Castermans, A. Pieterse, W. Sonke, and B.
Speckmann. 2015. ``Mosaic Drawings and Cartograms.'' \emph{Computer
Graphics Forum} 34 (3): 361--70.

\leavevmode\hypertarget{ref-ACTUC}{}%
Dorling, Daniel. 2011. ``Area Cartograms: Their Use and Creation.'' In
\emph{Concepts and Techniques in Modern Geography (CATMOG)}, 59:252--60.
\url{https://doi.org/10.1002/9780470979587.ch33}.

\leavevmode\hypertarget{ref-TVSSS}{}%
---------. 2012. \emph{The Visualisation of Spatial Social Structure}.
John Wiley \& Sons Ltd.

\leavevmode\hypertarget{ref-SE}{}%
Exeter, Daniel J. 2017. ``Spatial Epidemiology.'' Journal Article, 1--4.
\url{https://doi.org/10.1002/9781118786352.wbieg0283}.

\leavevmode\hypertarget{ref-CTTMB}{}%
Griffin, T.L.C. 1980. ``Cartographic Transformation of the Thematic Map
Base.'' \emph{Cartography} 11 (3). Taylor \& Francis: 163--74.
\url{https://doi.org/10.1080/00690805.1980.10438102}.

\leavevmode\hypertarget{ref-HEDP}{}%
Howe, G. M. 1989. ``Historical Evolution of Disease Mapping in General
and Specifically of Cancer Mapping.'' In \emph{Cancer Mapping}, edited
by Peter Boyle, Calum S. Muir, and Ekkehard Grundmann, 1--21. Berlin,
Heidelberg: Springer Berlin Heidelberg.

\leavevmode\hypertarget{ref-MTMSIH}{}%
Jahan, Farzana, Earl Duncan, Susanna Cramb, Peter Baade, and Kerrie
Mengersen. 2018. ``Making More of Spatial Maps: A Bayesian Meta-Analysis
Approach.'' In.

\leavevmode\hypertarget{ref-ECGC}{}%
Keim, D.A, S.C North, C Panse, and J Schneidewind. 2002. ``Efficient
Cartogram Generation: A Comparison.'' In \emph{IEEE Symposium on
Information Visualization, 2002. INFOVIS 2002}, 2002-:33--36. IEEE.

\leavevmode\hypertarget{ref-CBATCC}{}%
Kocmoud, Christopher J., and Donald H. House. 1998. ``A Constraint-Based
Approach to Constructing Continuous Cartograms.'' In.

\leavevmode\hypertarget{ref-TAAM}{}%
Levison, M. E., and W. Haddon Jr. 1965. ``THE Area Adjusted Map. AN
Epidemiologic Device.'' Journal Article. \emph{Public Health Reports}
80: 55--59.

\leavevmode\hypertarget{ref-ACA}{}%
Min Ouyang, and P. Revesz. 2000. ``Algorithms for Cartogram Animation.''
In \emph{Proceedings 2000 International Database Engineering and
Applications Symposium (Cat. No.PR00789)}, 231--35.
\url{https://doi.org/10.1109/IDEAS.2000.880581}.

\leavevmode\hypertarget{ref-SAMGIS}{}%
Moore, Dale A., and Tim E. Carpenter. 1999. ``Spatial Analytical Methods
and Geographic Information Systems: Use in Health Research and
Epidemiology.'' \emph{Epidemiologic Reviews} 21 (2): 143--61.
\url{https://doi.org/10.1093/oxfordjournals.epirev.a017993}.

\leavevmode\hypertarget{ref-SAIC}{}%
Nusrat, Sabrina, and Stephen G. Kobourov. 2016. ``The State of the Art
in Cartograms.'' \emph{CoRR} abs/1605.08485.
\url{http://arxiv.org/abs/1605.08485}.

\leavevmode\hypertarget{ref-NAC}{}%
Olson, Judy M. 1976. ``NONCONTIGUOUS Area Cartograms.'' \emph{The
Professional Geographer} 28 (4). Routledge: 371--80.
\url{https://doi.org/10.1111/j.0033-0124.1976.00371.x}.

\leavevmode\hypertarget{ref-GAMP}{}%
Tobler, Waldo. 1963. ``Geographic Area and Map Projections,'' January.
American Geographical Society, 59--78.
\url{https://www.jstor.org/stable/212809}.

\leavevmode\hypertarget{ref-TFYCC}{}%
---------. 2004. ``Thirty Five Years of Computer Cartograms.''
\emph{Annals of the Association of American Geographers} 94 (1). Taylor
\& Francis Group: 58--73.

\leavevmode\hypertarget{ref-EI}{}%
Tufte, Edward R. 1990. \emph{Envisioning Information}. Graphics Press.

\leavevmode\hypertarget{ref-DMAHP}{}%
Walter, S. D. 2001. \emph{Disease Mapping: A Historical Perspective}.
Book. Oxford University Press: Oxford.

\end{document}

### 2019-07-22 review_article.Rmd
Measures of cancer

--- 2019-07-19 review_article.Rmd
@@
 
-<!-- Displaying cancer data on choropleths -->
+<!-- Displaying of cancer data on choropleths -->
 
 In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality. 
+Choropleth maps can be useful devices for communicating information to public on a familiar map base.
+A cancer atlas is a choropleth map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
+@MACM provide the definition and report that cancer atlases began with Haviland's maps in 1875.
+The data collection methods of cancer mortality rates across regions, and administrative control within regions lends itself to choropleth visualisation.
+The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
+Early work in US cancer atlases can be attributed to Burbank ???, and UK cancer atlases to Howe 1963 ???.
+The choropleth maps used at this time presented levels  via hatchings or dots on a black and white scale.
+These atlases were key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets @MACM.
+
+<!-- Measures of cancer -->
 
 Almost 100 years of cancer mapping in the United States and the United Kingdom has increased the effectiveness in the presentation of unbiased rates.
-The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
-The data collection methods of cancer mortality rates across regions, and administrative control within regions lends itself to choropleth visualisation.
 Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the the higher prevalence of cancers in older populations.
 Howe [-@HEDP] describes Stocks development of the standardised mortality ratios through the 1930s.
-The choropleth maps used at this time presented levels of cancer via hatchings on a black and white scale.
 
-Mortality rates are now often presented as relative rates of risk across the population, and age adjusted.
-@HEDP describes Stocks' development of the standardised mortality ratios through the 1930s.
-The choropleth maps used presented levels of cancer via hatchings on a black and white scale.
+<!--
+Literature regarding measures presented in public facing atlases.
+
+-->
 

### 2019-07-22 review_article.Rmd

@@
 
-Utilising maps to present statistics has been widely used for centuries.
-These maps connect data and/or statistics to the geographical representation of areas that are generally familiar to the audience.
-However, it is not enough to be recognisable, if the meaning of the statistic cannot be seen.
-This is particularly relevant to cancer statistics, as cancer outcomes directly relate to the people living within a geographical area.
-In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer statistics.
-This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes, and health outcomes more broadly.
-
 # Introduction
@@
 
+Maps have been adopted to present geospatial statistics for centuries.
+These maps connect data and/or statistics to the geographical representation of areas that are generally familiar to the audience.
+However, it is not enough for areas on maps to be recognisable, if the distribution of the statistic cannot be understood.
+This is particularly relevant to cancer statistics, as cancer outcomes directly relate to the people living within a geographical area.
+In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer statistics.
+This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes, and health outcomes more broadly.
+
 # Disease mapping methods
@@
 <!--
-Disease maps can rapidly highlight geographical
-patterns that would otherwise be hidden in a
-table, and are commonly used for hypothesis
-generating and as the first step in exploratory
-spatial data analysis. @SE
-Geography is concerned with the identification and
-explanation of spatial structure, pattern, and process,
-and with the analysis and explanation of the links
+Disease maps present geographic patterns that may have been overlooked in a table that obscures the geospatially related statistics [@SAMGIS].
+Often used as a first step to drive hypotheses
+
+Transforming statistics to visualisations begins with the coordinates of observations.
+The observation points can be aggregated to regions within boundaries.
+The polygon regions may be presented directly, or transformed first according to associated values.
+
+The visualisation method used will depend on the intended use of the map, namely the message a map will communicate.
+
+commonly used for hypothesis generating and as the first step in exploratory
+spatial data analysis. @SE Geography is concerned with the identification and
+explanation of spatial structure, pattern, and process, and with the analysis and explanation of the links
 between humans and the environment (2). @SAMGIS
 
-By providing a visual representation of
-cancer outcomes, geographic patterns of disease are able
+By providing a visual representation of cancer outcomes, geographic patterns of disease are able
 to be identified and effectively addressed
@@
 <!-- what is a choropleth map -->
-A choropleth map is used to show the spatial relationship, or variability, of measurements.
+A choropleth map is used to show the spatial characteritics of a 
+relationship, or variability, of measurements.
 Created as a storage device which preserves locations for geographically ordered data, with usage dating back to the 1800s [@GOINO].
@@
 <!--
+criticisms:
 Certain limitations to the standard graphic techniques have also become apparent, however.
@@
 They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually.
-
-These opportunities will be lost if locational coding is not inserted into the data referencing system as such banks are built, however.
-
-of some element, for example population density, or purchasing power, or agricultural fertility; reality, of course, is not like this.
-the background variable assumed to be uniform (such as population density) can be adjusted so that it behaves as if uniformity were the case
 -->
@@
 Literature regarding measures presented in public facing atlases.
+-->
 
--->
+### Reported measures in cancer atlases
+
+The list below summarises these measures, and provides a definition of each measure.
+
+Table: (\#tab:measures) Measures used to report cancer statistics
+
+|Measure |Details| 
+|:--------------------|:------------------------------------------------|  
+|1. IR (Incidence Ratio)|$(IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}$, |
+||Cancer incidence rate in region $i$ over the average cancer incidence rate for the total region|
+|2. SIR (Standardised Incidence Ratio)| IR standardised by age structure in each region $i$|
+|3. RER |$RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}$|
+|(Relative Excess Risk)|Represents the estimate of cancer related mortality within five years of diagnosis|
+||Also referred to as 'excess hazard ratio'|
+|4. Age Adjusted Relative Risk| RR standardised by age structure in each region $i$|   
+|5. Rate per 100,000|Cancer incidence per 100,000 population|  
+|6. Age Adjusted Rate per 100,000| #5 standardised by age structure or region|   
+|7. New cancer cases per 100,000|Specific methods could not be found|   
+|8. Count|Crude cancer counts|  
+|9. Below or above Expected|Alternative expression of the SIR|
 
@@
 -->
+
+### Supporting material in cancer atlases
+
+<!-- Moved Jessies stat uncertainty up-->
+Cancer atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map.
+Maps that only reported this information within the supplementary material were not considered to have directly attempted to report uncertainty.
+
+The review did not reveal any novel uncertainty visualisation approaches or visualisations.
+Maps used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
+These maps ranged from static pdfs or infographics to interactive online resources.
+The interactivity of the more modern maps enabled uncertainty information to be incorporated without cluttering the screen, such as in a tool tip feature.
 
@@
 
-A grey literature review conducted by (ref) identified 33 publicly available cancer atlases published on the internet between 01/01/2010 to 11/11/2015.
-Reflecting the high use of chloropleth maps founds in the literature, outlined in the section above, all the public facing maps identified in this review were also chloropleth maps.
+Cancer maps are effective communication tools for a general or non-expert audience and are commonly used in the public domain to communicate the output of sophisticated statistical analyses to these audiences.
+The heavy use of chloropleth maps within the research literature is reflected in the types of maps that are found in the public domain.
+A grey literature review conducted by (ref) identified 33 cancer atlases published on the internet between 01/01/2010 to 11/11/2015 all of which were chloropleth maps.
+These maps were mostly published by non-commercial organisations, including not-for-profits (NFPs), government, research organisations, advocacy groups or a partnership between an NFP & government.
+Only one map was published by a commercial entity (Maps of Cancer Mortality Rates in Spain), published by the media organisation El Pais^[https://elpais.com/].
 
-### Publishers {-}
+### Geographical coverage
 
-The majority of atlases were published by non-commercial organisations, including not-for-profits (NFPs), government, research organisations, advocacy groups or a partnership between an NFP & government.
-Only one map was published by a commercial entity (Maps of Cancer Mortality Rates in Spain), in this case a media organisation El Pais^[https://elpais.com/].
-
-### Reported measures {-}
-
-The majority of maps identified in this study reported age adjusted rates of either incidence, mortality or survival.
-The list below summarises these measures, and provides a definition of each measure.
-
-### Geographical coverage {-}
-
-Identified cancer atlases covered geographies from all around the world: 4 were global, 3 from Australia (AUS), 11 from the United States (US), 2 from Canada (CAN), 7 from the United Kingdom (UK), 2 from Spain, 1 from Switzerland, 1 from Germany, 1 from Norway, and 1 covering the European Union.
-Not all maps had a national focus and 10 covered a region or state rather than an entire nation.
+Identified cancer atlases covered geographies from all around the world, four were global atlases.
+Most focussed on single nations, the United States was considered by eleven atlases.
+The United Kingdom was considered by seven, followed by three of Australia, two of Canada, and one of each from Switzerland, Germany, Norway.
+One atlas covered the European Union.
+Not all maps had a national focus and ten covered a region or state rather than an entire nation.
 The states or counties/regions covered were South Australia (AUS), Queensland (AUS), Ontario (CAN), Valencia (Spain), Pennsylvania county Massachusetts (US), New Hampshire (US), Cape Cod (US), Missouri (US), Florida (US), New York State (US) and Arizona (US).
 
-\newpage
-Table: (\#tab:measures) Measures used to report cancer statistics
-
-|Measure |Details| 
-|:--------------------|:------------------------------------------------|  
-|1. IR (Incidence Ratio)|$(IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}$, |
-||Cancer incidence rate in region $i$ over the average cancer incidence rate for the total region|
-|2. SIR (Standardised Incidence Ratio)| IR standardised by age structure in each region $i$|
-|3. RER |$RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}$|
-|(Relative Excess Risk)|Represents the estimate of cancer related mortality within five years of diagnosis|
-||Also referred to as 'excess hazard ratio'|
-|4. Age Adjusted Relative Risk| RR standardised by age structure in each region $i$|   
-|5. Rate per 100,000|Cancer incidence per 100,000 population|  
-|6. Age Adjusted Rate per 100,000| #5 standardised by age structure or region|   
-|7. New cancer cases per 100,000|Specific methods could not be found|   
-|8. Count|Crude cancer counts|  
-|9. Below or above Expected|Alternative expression of the SIR|
-
-### Statistical uncertainty
-
-Cancer atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map.
-Maps that only reported this information within the supplementary material were not considered to have directly attempted to report uncertainty.
-
-The review did not reveal any novel uncertainty visualisation approaches or visualisations.
-Maps used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
-These maps ranged from static pdfs or infographics to interactive online resources.
-The interactivity of the more modern maps enabled uncertainty information to be incorporated without cluttering the screen, such as in a tool tip feature.
+### Statistical uncertainty in public facing atlases
 

### 2019-07-22 review_article.Rmd
Reported measures

@@
 
-<!-- Measures of cancer -->
+<!-- Measures reported in cancer atlases -->
 
-Almost 100 years of cancer mapping in the United States and the United Kingdom has increased the effectiveness in the presentation of unbiased rates.
+Almost 100 years of cancer mapping in the United States and the United Kingdom has seen increased effectiveness in the presentation of unbiased rates.
 Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the the higher prevalence of cancers in older populations.
@@
 
-### Reported measures in cancer atlases
-
-The list below summarises these measures, and provides a definition of each measure.
+The list below summarises the measures presented in published cancer atlases, and provides a definition of each measure.
 
@@
 
-<!--
 @DMAHP cites Cruickshank's (1947) discussion of using visuals as 'formal statistical assessment of the spatial pattern' as a major advancement.
-@MACM also recognises the intuition that can be derived from choropleths must be 'validated by rigorous statistical analyses.'
 
-@CD considers cartographic data analysis in practice. Presenting a clear guide for preparing cartograms for use with both qualitative and quantitative information.
-The authors recognise that the creation of cartograms was largely the work of professional cartographers until the innovation of geographical information systems that welcomed map users as map creators.
-These systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’.
--->
+When presenting cancer atlases, @MACM believes the intuition that can be derived must be 'validated by rigorous statistical analyses.' Additional statistics are often
 
@@
 -->
-
-### Supporting material in cancer atlases
-
-<!-- Moved Jessies stat uncertainty up-->
-Cancer atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map.
-Maps that only reported this information within the supplementary material were not considered to have directly attempted to report uncertainty.
-
-The review did not reveal any novel uncertainty visualisation approaches or visualisations.
-Maps used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
-These maps ranged from static pdfs or infographics to interactive online resources.
-The interactivity of the more modern maps enabled uncertainty information to be incorporated without cluttering the screen, such as in a tool tip feature.
 
@@
 <!-- Going from choropleth to cartogram
+
+A cartogram is an alternative map visual. 
+An intention distortion is applied to a geographic map base,
+
 A cartogram is a map purposely distorted so that its spatial properties represent quantities
@@
 @ACCAC
+
+@CD considers cartographic data analysis in practice. Presenting a clear guide for preparing cartograms for use with both qualitative and quantitative information.
+The authors recognise that the creation of cartograms was largely the work of professional cartographers until the innovation of geographical information systems that welcomed map users as map creators.
+These systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’.
+
+'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
+
 -->
@@
 
-With that in mind, the intended user and message to communicate should drive map selection.
-first is the nature or inherent properties of the device itself
-second is the ease and accuracy with which map users are able to extract information
-from it
+To choose an approporiate map display, the map creator must consider the intended user, and message the map will communicate.
+It is the objectives of the investigator that will drive the choice of representation [@CPISACA].
 
-@SAMGIS suggests it is the "investigators' objectives" that drive the "representation of diseases on maps". 
--infectious diseases may require choropleth methods
-
-'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
+There are two keys to drive the choice of display: the properties of the visualisation, and the ease or accuracy of information extraction for map users @SAMGIS.
 
@@
 
-# Animation
+# Animation and Interactivity
+
+Recent developments of technology allowed interactive
+
+'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
 
@@
 
-What software did we use, for e.g.
-
 # References
@@
 -->
+@tidyerse for data analysis, @spData and @sf to implement plotting of spatial data.
+@sugarbag to create hexagon tesselation. @ggthemes to enhance plot displays.
+@cartogram for contiguous and non-contiguous cartogram displays.
 
@@
 
-Driven by progressions in computing, statistical methods and collection of data.
-
-Models summarise relationships between counts of diseases and other factors.
-
-Recognition of biases when utilising the geography to map people. Small areas contain more people, work towards representing regions of importance more clearly.
 -->

### 2019-07-30 review_article.Rmd
Sentence edits

@@
   pdf_document:
-    toc: true
+    number_sections: yes
+    toc: yes
     toc_depth: 2
-    number_sections: true
   html_document: default
 fontsize: 11pt
-link-citations: true
+link-citations: yes
 fig_caption: yes
-yolo: true
 biblio-style: vancouver
+yolo: yes
 ---
@@
 
-Maps have been adopted to present geospatial statistics for centuries.
-These maps connect data and/or statistics to the geographical representation of areas that are generally familiar to the audience.
-However, it is not enough for areas on maps to be recognisable, if the distribution of the statistic cannot be understood.
+Maps have been adopted as a base to present geospatial statistics for centuries.
+These maps connect data and statistics to the geographical representation of areas that are generally familiar to the audience.
+However, it is not enough for areas on maps to be recognisable, if information regarding the distribution of the statistic cannot be gleaned.
 This is particularly relevant to cancer statistics, as cancer outcomes directly relate to the people living within a geographical area.
@@
 <!-- what is a choropleth map -->
-A choropleth map is used to show the spatial characteritics of a 
-relationship, or variability, of measurements.
-Created as a storage device which preserves locations for geographically ordered data, with usage dating back to the 1800s [@GOINO].
-They are constructed by drawing the given geographic or political boundaries, and filling the shapes with colours to represent values of a measured variable [@EI].
-Early versions used symbols or patterns instead of colour.
-@CPISACA discusses the use of choropleth maps for visualising cancer data, and @DMAHP gives an overview of the development of the use of choropleth maps for displaying disease data.
+A choropleth map is often used to display the spatial characteritics of a relationship, or variability, of measurements. 
+As an alternative storage device to a table, it preserves locations for geographically ordered data, with usage dating back to the 1800s [@GOINO].
+A choropleth is constructed by drawing the geographic or political boundaries, and filling the shapes with colours to represent values of a measured variable [@EI].
+Early versions of choropleth maps used symbols or patterns instead of colour.
+@CPISACA discusses their use in visualising cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
+Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
 A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form. 
@@
 
-<!--
-Thematic maps, on the other hand, do select some geographical concept or arrangement and attempt to focus attention entirely on the spatial characteristics of the entire distribution and upon its relation to other geographical factors.
--->
-
-!["A choropleth map of the United States of America."](figures/ggchoro.png)
+```{r choro, fig.cap = "A choropleth map of the United States of America.", echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+knitr::include_graphics("figures/ggchoro.png")
+```
 
 <!-- psychology -->
-
 Utilising the familiar state boundaries can make a map intuitive to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data, i.e.
@@
 The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
-Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
-However, while the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
+While the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
 The different population and geographical sizes of administrative areas can attract attention to the shades of the unpopulated but large areas [@EI]. 
 Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].
-
-<!--
-criticisms:
-Certain limitations to the standard graphic techniques have also become apparent, however.
-They are slow, and do not permit instant retrieval of information.
-They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually.
--->
-
-<!--
-Urbanisation has pushed people closer together,
--in workplaces and in homes. 
--this issue is not with the size of the administrative areas,
-- the increasing urbanisation is contributing to the changing density of population
-
-A particular spatial structure includes
-the individuals affected and how they are connected in
-communities, as well as the dynamics of these communities
-and their organization into larger units. @SAMGIS
--->
 
@@
 A cancer atlas is a choropleth map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
-@MACM provide the definition and report that cancer atlases began with Haviland's maps in 1875.
-The data collection methods of cancer mortality rates across regions, and administrative control within regions lends itself to choropleth visualisation.
+@MACM provide the definition and reports that cancer atlases began with Haviland's maps in 1875.
+The data collection methods of cancer mortality rates across regions, and the administrative control within regions lends itself to choropleth visualisation. 
 The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
-Early work in US cancer atlases can be attributed to Burbank ???, and UK cancer atlases to Howe 1963 ???.
-The choropleth maps used at this time presented levels  via hatchings or dots on a black and white scale.
+arly work in US cancer atlases can be attributed to Burbank ???, and UK cancer atlases to Howe 1963 ???.
+The choropleth maps presented levels via hatchings or dots on a black and white scale.
 These atlases were key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets @MACM.
@@
 
-Almost 100 years of cancer mapping in the United States and the United Kingdom has seen increased effectiveness in the presentation of unbiased rates.
+Almost 100 years of cancer mapping in the United States and the United Kingdom has seen increased effectiveness in the presentation cancer statistics.
 Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the the higher prevalence of cancers in older populations.
-Howe [-@HEDP] describes Stocks development of the standardised mortality ratios through the 1930s.
+Howe [-@HEDP] describes Stock's development of the standardised mortality ratios through the 1930s.
+Table 1 presents summarises the measures presented in published cancer atlases, and provides a definition of each measure.
 
-<!--
-Literature regarding measures presented in public facing atlases.
--->
-
-The list below summarises the measures presented in published cancer atlases, and provides a definition of each measure.
-
-Table: (\#tab:measures) Measures used to report cancer statistics
+<!-- ??? Table tags-->
+Table 1: Measures used to report cancer statistics
 
@@
 of some element, for example population density, or purchasing power, or agricultural fertility; reality, of course, is not like this. 
-the background variable assumed to be uniform (such as population density) can be adjusted so that it behaves as if uniformity were the case
-@GOINO
 -->
@@
 <!-- Going from choropleth to cartogram 
+the background variable assumed to be uniform (such as population density) can be adjusted so that it behaves as if uniformity were the case
+@GOINO
 
@@
 
+-->
+<!--
+criticisms:
+Certain limitations to the standard graphic techniques have also become apparent, however.
+They are slow, and do not permit instant retrieval of information.
+They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually.
+-->
+
+<!--
+Urbanisation has pushed people closer together,
+-in workplaces and in homes. 
+-this issue is not with the size of the administrative areas,
+- the increasing urbanisation is contributing to the changing density of population
+
+A particular spatial structure includes
+the individuals affected and how they are connected in
+communities, as well as the dynamics of these communities
+and their organization into larger units. @SAMGIS
 -->

### 2019-07-30 review_article.Rmd
Rearranged supporting material ideas

@@
 
-@DMAHP cites Cruickshank's (1947) discussion of using visuals as 'formal statistical assessment of the spatial pattern' as a major advancement.
-
-When presenting cancer atlases, @MACM believes the intuition that can be derived must be 'validated by rigorous statistical analyses.' Additional statistics are often
-
 <!--
-of some element, for example population density, or purchasing power, or agricultural fertility; reality, of course, is not like this. 
--->
-
-<!-- Published atlases -->
 
@@
 
-<!--
 Current atlases utilising choropleth methods 
@@
 
-<!-- Moved Jessies stat uncertainty up-->
-Cancer atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map.
+A map communicates quickly and draws attention to prominent geographies, but an atlas is often supplemented with supporting statistics.
+Cruickshank's (1947) as cited by @DMAHP, discusses using visuals as 'formal statistical assessment of the spatial pattern'.
+However, when presenting cancer atlases, @MACM believes the intuition derived must be 'validated by rigorous statistical analyses.'
+
+<!-- The supplementary statistics ??? -->
+These additional statistics often include a measure of the statistical uncertainty of the values of the statistics presented in a choropleth.
+In the review of atlases in the public domain, atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map.
 Maps that only reported this information within the supplementary material were not considered to have directly attempted to report uncertainty.
+The maps considered used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
+These maps ranged from static pdfs or infographics, to interactive online resources.
 
-The review did not reveal any novel uncertainty visualisation approaches or visualisations.
-Maps used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
-These maps ranged from static pdfs or infographics to interactive online resources.
-The interactivity of the more modern maps enabled uncertainty information to be incorporated without cluttering the screen, such as in a tool tip feature.
+<!-- Presentation methods of supplementary statistics ??? -->
+The interactivity of the more modern maps enabled supplementary information to be incorporated without cluttering the screen, such as in a tool tip feature.
 
@@
 
-Cancer maps are effective communication tools for a general or non-expert audience and are commonly used in the public domain to communicate the output of sophisticated statistical analyses to these audiences.
+Cancer maps are effective communication tools for a general or non-expert audience. They are commonly used in the public domain to communicate results of sophisticated statistical analyses.
 The heavy use of chloropleth maps within the research literature is reflected in the types of maps that are found in the public domain.
-A grey literature review conducted by (ref) identified 33 cancer atlases published on the internet between 01/01/2010 to 11/11/2015 all of which were chloropleth maps.
-These maps were mostly published by non-commercial organisations, including not-for-profits (NFPs), government, research organisations, advocacy groups or a partnership between an NFP & government.
-Only one map was published by a commercial entity (Maps of Cancer Mortality Rates in Spain), published by the media organisation El Pais^[https://elpais.com/].
+A grey literature review conducted by (ref) identified 33 cancer atlases published on the internet between January 2010 and November 2015.
+These chloropleth maps were mostly published by non-commercial organisations, including not-for-profits (NFPs), government, research organisations, advocacy groups or a partnership between an NFP & government.
+Only one map was published by a commercial entity.
 
-### Geographical coverage
-
-Identified cancer atlases covered geographies from all around the world, four were global atlases.
-Most focussed on single nations, the United States was considered by eleven atlases.
-The United Kingdom was considered by seven, followed by three of Australia, two of Canada, and one of each from Switzerland, Germany, Norway.
+<!-- geographic coverage-->
+The cancer atlases covered geographies from all around the world, four were global atlases.
+Most focussed on single nations, the United States was considered by eleven atlases, the United Kingdom by seven, followed by three of Australia, two of Canada, and one of each from Switzerland, Germany, Norway.
 One atlas covered the European Union.
@@
 
-### Statistical uncertainty in public facing atlases
+### Examples of supporting statistics in public facing atlases
+
+Close to half of the atlases identified (42%, n=14) included some measure of uncertainty. The most common measure used to represent uncertainty were credible or confidence intervals (CIs).
+
+<!--
+Create a table:
+cofactors: socio demographic status
+
+Supplementary stats
+stat        | Literature examples | Public atlas examples
+uncertainty |      ...            |    ...
+co factor(1)|      ...            |    ...
+co factor(2)|      ...            |    ...
+-->
 
@@
 Source: Alberta Health IHDA Geographic.
-(2012) ", echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+(2012) ", echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width='50%'}
 
@@
 ```{r ci-viz3, fig.cap = "Example of CI visualisation for uncertainty representation in cancer mapping (2/3).
-Source: Pensylvania Cancer Atlas ", echo=FALSE, message=FALSE, warning=FALSE, out.width='55%', fig.align='center'}
+Source: Pensylvania Cancer Atlas ", echo=FALSE, message=FALSE, warning=FALSE, out.width='50%', fig.align='center'}
 
@@
 
-Close to half of the atlases identified (42%, n=14) included some measure of uncertainty.
-The most common measure used to represent uncertainty were credible or confidence intervals (CIs).
-CIs were either visualised by including their bounds in a scatterplot or graph of estimates vs region (see Figures \@ref(fig:ci-viz2)^[Alberta Health IHDA Geographic.
-(2012).
-Age-Standardised Incidence Rate of COPD, 2011.
-Retrieved from: http://www.health.alberta.ca/health-info/IHDA-geographic/COPD/incidence-agestandard/atlas.html?epik=0GJSpE_IW34lx], and \@ref(fig:ci-viz3)^[Centres for Disease Control and Prevention (CDC).
-(n.d).
-United States Cancer Statistics: An Interactive Cancer Atlas (InCA).
-Retrieved from:https://nccd.cdc.gov/DCPC_INCA/] positioned next to the map, or reported numerically through the CI upper and lower bounds listed in a data table (see Figure 3.4^[Pennsylvania Cancer Atlas.
-(n.d).
-Retrieved from:  https://www.geovista.psu.edu/grants/CDC/?epik=0lJSpE_IW34lx]).
-Of those that visualised the CIs, 30% (n=10) embedded the visualisation within a tool tip function which visualised the CI when the mouse hovered over the relevant area (see Figure 3.5^[International Agency for Research on Cancer.
-(2017).
-Atlas of Cancer Mortality in the European Union and European Economic Area 1993-1997, Annex 4 - Cancer mortality maps by site.
-Retrieved from: http://www.iarc.fr/en/publications/pdfs-online/epi/sp159/]).
-
 ```{r ci-tooltip, fig.cap= "Example of CI visualisation for uncertainty representation in cancer mapping (3/3).
 Source: Centres for Disease Control and Prevention (CDC).
-United States Cancer Statistics: An Interactive Cancer Atlas (InCA) ", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '60%', fig.align = 'center'}
+United States Cancer Statistics: An Interactive Cancer Atlas (InCA) ", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '50%', fig.align = 'center'}
 

### 2019-08-01 review_article.Rmd
cofactors
editing introduction

@@
 
-Maps have been adopted as a base to present geospatial statistics for centuries.
-These maps connect data and statistics to the geographical representation of areas that are generally familiar to the audience.
+Maps have been adopted to present geospatial statistics for centuries.
+Maps connect data and statistics to the geographical representation of areas that are generally familiar to the audience.
 However, it is not enough for areas on maps to be recognisable, if information regarding the distribution of the statistic cannot be gleaned.
-This is particularly relevant to cancer statistics, as cancer outcomes directly relate to the people living within a geographical area.
 In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer statistics.
+Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS]. 
 This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes, and health outcomes more broadly.
+Cancer statistics directly represent and relate to the people living within individual geographical areas.
+
+The visualisation methods used to present cancer statistics will depend on the intended message and users of the map. 
+Transforming statistics to visualisations considers individual observations or observation points aggregated into communities and geographical units.
+Visualisations may be used for hypothesis generating and is often the first step in exploratory spatial data analysis. 
+Disease maps help to present geographic patterns that may have been overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
+By providing a visual representation of cancer outcomes, geographic patterns of disease are able to be identified and effectively addressed.
 
@@
 
-<!--
-Disease maps present geographic patterns that may have been overlooked in a table that obscures the geospatially related statistics [@SAMGIS].
-Often used as a first step to drive hypotheses
-
-Transforming statistics to visualisations begins with the coordinates of observations.
-The observation points can be aggregated to regions within boundaries.
-The polygon regions may be presented directly, or transformed first according to associated values.
-
-The visualisation method used will depend on the intended use of the map, namely the message a map will communicate.
-
-commonly used for hypothesis generating and as the first step in exploratory
-spatial data analysis. @SE Geography is concerned with the identification and
-explanation of spatial structure, pattern, and process, and with the analysis and explanation of the links
-between humans and the environment (2). @SAMGIS
-
-By providing a visual representation of cancer outcomes, geographic patterns of disease are able
-to be identified and effectively addressed
--->
-
 ## Current best practice map displays for cancer data
 
-<!--Choropleth map (Steff)-->
-
 <!-- what is a choropleth map -->
-A choropleth map is often used to display the spatial characteritics of a relationship, or variability, of measurements. 
+A choropleth map is used to display the spatial characteritics of a relationship, or variability, of measurements. 
 As an alternative storage device to a table, it preserves locations for geographically ordered data, with usage dating back to the 1800s [@GOINO].
@@
 
+<!--
+They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually.
+-->
+
 <!-- Displaying of cancer data on choropleths -->
@@
 The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
-arly work in US cancer atlases can be attributed to Burbank ???, and UK cancer atlases to Howe 1963 ???.
+Early work in US cancer atlases can be attributed to Burbank ???, and UK cancer atlases to Howe 1963 ???.
 The choropleth maps presented levels via hatchings or dots on a black and white scale.
@@
 
+An Online Census Atlas for everyone [@AOCAFE] uses familiar choropleth, produced using generalized boundaries for six higher level geographies.
+Great Britain is recognizable and it is relatively easy to identify the higher geography levels such as Counties and Local Authorities.
+
 <!--
+Medical application: TAAM 1965
+Actual shapes of counties were sacrificed to the extent necessary to retain as far as possible, common boundaries, relative locations, and the approximate contour of the State.
+
+There is no unique solution to the topology
+problem of preparing a population-by-area
+cartogram for a given population and its jurisdictional
+boundaries. Although the map areas
+of the various regions are fixed by their populations, their shapes and relative locations are not completely determined.
+
+Other characteristics of the population
+may also serve as the base for adjustment; for
+example, sex, race, parity, socio-economic level,
+or immune status.
 
@@
 However, when presenting cancer atlases, @MACM believes the intuition derived must be 'validated by rigorous statistical analyses.'
+
+<!--
+'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
+
+-->
 
@@
 Create a table:
-cofactors: socio demographic status
+cofactors: Greater capital city regions, states, socio demographic status, Remoteness
 
@@
 
-<!-- Going from choropleth to cartogram 
-the background variable assumed to be uniform (such as population density) can be adjusted so that it behaves as if uniformity were the case
-@GOINO
+> The aim here is to see the whole, in as much detail as possible, at a glance.
 
-A cartogram is an alternative map visual. 
-An intention distortion is applied to a geographic map base,
-
-A cartogram is a map purposely distorted so that its spatial properties represent quantities
-not directly associated with position on the globe. As thematic maps, cartograms emphasize
-the distribution of a variable by changing the area (or lengths) of objects on the map.
-Our approach maintains continuity, and preserves
-many local features of cartographic lines that provide visual clues to the identity of the distorted
-objects.
-@ACCAC
-
-@CD considers cartographic data analysis in practice. Presenting a clear guide for preparing cartograms for use with both qualitative and quantitative information.
-The authors recognise that the creation of cartograms was largely the work of professional cartographers until the innovation of geographical information systems that welcomed map users as map creators.
-These systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’.
-
-'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
-
--->
-<!--
-criticisms:
-Certain limitations to the standard graphic techniques have also become apparent, however.
-They are slow, and do not permit instant retrieval of information.
-They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually.
--->
-
-<!--
-Urbanisation has pushed people closer together,
--in workplaces and in homes. 
--this issue is not with the size of the administrative areas,
-- the increasing urbanisation is contributing to the changing density of population
-
-A particular spatial structure includes
-the individuals affected and how they are connected in
-communities, as well as the dynamics of these communities
-and their organization into larger units. @SAMGIS
--->
+<!-- What is a cartogram? -->
+Cartograms provide an alternative visualisation method for statistical and geographical information.
+A cartogram alters the map base with the intention of improving the presentation of a statistic.
+The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC].
+For a single variable of interest, each map area is changed to emphasise the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].
 
@@
 <!-- Tobler map projections-->
-Choropleths may be considered true topological maps, however, if the land mass presented covers enough of the globe, there must be a transformation or distortion to display in 2D.
-The amount of distortion is related to the distance displayed @GAMP.
-The world projections reflect the frequent distortions seen from altering perspectives.
-Choropleth maps will always be distorted if they cover enough of the globe, as will photographs of the globe from space.
+Choropleths may be considered true topological maps, however if the land mass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D.
+The amount of distortion is related to the distance covered by the landmass displayed @GAMP.
+World map projections reflect the frequent perspectives used to view the earth. Choropleth maps will always be distorted if they cover enough of the globe, just as photographs of the globe from space.
 Choropleth creation requires choosing a map projection that shows a favourable distortion of the geography for presenting the set of spatial information.
-Diagrams that do not specify a projection can be considered to have some unknown projection. <!-- ?? cite tobler.-->
+Diagrams that do not specify a projection can be considered to have some unknown projection. <!-- ??? cite tobler.-->
 
@@
 
-<!-- Why transform?-->
-<!-- Common variables used to create cartograms, e.g. population, mortality -->
-Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
-Cartograms provide an alternative visualisation method for statistical and geographical information.
-The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC].
-Cartograms may be seen as an extension of map transformations and projections. The favourable distortion is proportional to a value other other than actual earth size area [@NAC].
+<!-- Denominators -->
+If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed beyond choosing a projection. 
+The purposeful distortion of the map space is beneficial when a uniform density of the map base is desired, @ACTUC suggests 'population distribution is often extremely uneven in former British colonies', this makes the distortion necessary.
+When a map base is transformed according to population density, population becomes a uniformly distributed background for the statistic presented [@GOINO]. 
+Using choropleth maps for population characteristics requires graphic distortions when the population concerned varies greatly in density [@CTTMB].
+When implementing a distortion of the geographical shape according to population, an area cartogram [@NAC], population-by-area cartograms [@TAAM], or iso-demographic map is the result.
 
-<!--
-@SE Researchers mapping spatial patterns of disease
-for small areas are confronted with a number of
-challenges, including cartographic design, the
-modifiable area unit problem, and also the “small
-number” problem, where some places have no,
-or few, disease events and/or a low population at
-risk of contracting the disease.
-
-@TAAM variations in population density may obscure geographic patterns
-This factor may be eliminated by making the map area proportional to population rather than to actual land area.
-Such maps, or what have been termed "population-by-area cartograms," are well known in the
-cartographic literature
--->
-Using choropleth maps for population characteristics requires graphic distortions when the population concerned varies greatly in density [@CTTMB].
-When implementing a distortion of the geographical shape according to population, an area cartogram [@NAC], or iso-demographic map is the result.
+A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, making interpretation of spatial patterns very difficult [@AOCAFE].
+The distortion of a cartogram prevents the density obscuring the spatial patterns [@TAAM]. 
+The spatial transformation of map regions relative to the data emphasises the data distribution instead of land size [@CBATCC].
 When visualising population statistics @ACTUC considers this equitable representation design 'more socially just, giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
@@
 @MTMSIH encourage the use of cartograms to highlight small areas and uncover local-level inequalities.
-Using health estimates from large areas can prevent drawing attention to the inequalities.
 
 <!--
-> The aim here is to see the whole, in as much detail as possible, at a glance.
-
 @HEDP discusses the impact of electronic computer-assisted techniques. Disease atlases were implemented using both geographic (Howe 1963) and demographic bases (Howe 1970).
 
-@AOCAFE
-An Online Census Atlas for everyone
-The familiar choropleth maps in the Online Census Atlas were produced using generalized boundaries.
-Great Britain is recognizable and it is relatively easy to identify the higher geography levels such as Counties and Local Authorities.
-A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, making interpretation of spatial patterns very difficult.
+Design goals:
+- to maintain as much as possible of the intricate detail
+- to maximize the simplicity of the transformation
+- maintenance of directional accuracy
+- attempt to influence and direct the map user's attention,
 
-The Atlas does not allow visualization of census
-data at ward level, but instead uses these boundaries
-aggregated to the six higher level geographies
-defined above for display of 88 census variables.
-
-To cartograms by:
-The creation of the map began
-with a computer-generated cartogram of parliamentary
-constituencies where each constituency was
-given equal weight.
-Using the circular cartogram
-algorithm (Dorling 1995), this resulted in a largely
-hexagonal tessellation of 641 circles, each circle
-representing a parliamentary constituency as first
-used in the 1997 general election.
+2) that the contiguity relationships between the DCU's should be maintained, 
+3) that true directions should be retained from the centre of the transformation,
+4) that the main shape characteristics of the DCU's should be preserved, and 
+5) that the map should possess the overall shape of the built-up area. 
 -->
@@
 
-The area on the map space can also be used represent a value.
 There have been many algorithms presented, @SAIC provided a framework to investigate implementations and the "statistical accuracy, geographical accuracy, and topological accuracy".
-The spatial transformation of map regions relative to the data emphasises the data distribution instead of land size [@CBATCC].
-
-Cancer statistics and atlases
 
@@
 
-<!-- Creation goals-->
-
-'Firstly, that no such transformation has a unique solution, leaving the subjective choice in the hands of the individual map designer. '
-
-Design goals:
-- to maintain as much as possible of the intricate detail
-- to maximize the simplicity of the transformation
-- maintenance of directional accuracy
-- attempt to influence and direct the map user's attention,
-
-1) that the size of the electoral subdivisions should be made proportional to their respective electoral populations, 
-2) that the contiguity relationships between the DCU's should be maintained, 
-3) that true directions should be retained from the centre of the transformation, 
-4) that the main shape characteristics of the DCU's should be preserved, and 
-5) that the map should possess the overall shape of the built-up area.
-
 ### Contiguous
@@
 
-<!-- Intentionally preserve shape-->
+<!-- Intentionally preserve neighbours -->
 
-A contiguous cartogram allows the map space to highlight the distribution of the variable.
-These cartograms maintain connectivity of the map regions while areas are resized according to a statistic, this often must occur at the expense of the shape [@CBATCC,@NAC, @TAAM].
+A contiguous cartogram maintains connectivity of the map regions while areas are resized according to a statistic. This transformation often occurs at the expense of the shape [@CBATCC, @NAC, @TAAM].
 From a computer graphics perspective, @ACA believe it is a problem of 'map deformation' to account for the value assigned to each area, they provide three methods for creating value-by-area cartograms.
@@
 <!--
-Medical application: TAAM 1965
-Actual shapes of counties were sacrificed to the extent necessary to retain as far as possible, common boundaries, relative locations, and the approximate contour of the State.
-
-There is no unique solution to the topology
-problem of preparing a population-by-area
-cartogram for a given population and its jurisdictional
-boundaries. Although the map areas
-of the various regions are fixed by their populations, their shapes and relative locations are
-not completely determined.
-
-Other characteristics of the population
-may also serve as the base for adjustment; for
-example, sex, race, parity, socio-economic level,
-or immune status.
-
-@RSCW
-helps to correct misconceptions.
-Pupils are likely to over-estimate regions as Canada, Siberia, Africa, and our own West just because they loom up large on maps.
-The simple picture which the cartogram gives is easily comprehended and remembered and
-will be stimulating in making comparisons.
 
@@
 Until we know how the map reader compares it to an ordinary thematic map, we are experimenting with a graphic device that may be reducing (or at least hindering) the flow of information from the map to the map reader.
+
+The transformations required to create cartograms was historically the role of professional cartographers.  
+Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’[@CD].
+
 -->
@@
 constant shapes are easily recognized regardless of scale;
+
+Presenting an alternative view can help to correct interpretation of global statistics when some areas loom large on maps [@RSCW].
 -->
@@
 -->
-
-@ACTUC suggests 'population distribution is often extremely uneven in former British colonies'.
 
@@
 
+Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
 To choose an approporiate map display, the map creator must consider the intended user, and message the map will communicate.
@@
 -->
+
+<!--
+@SE Researchers mapping spatial patterns of disease
+for small areas are confronted with a number of
+challenges, including cartographic design, the
+modifiable area unit problem, and also the “small
+number” problem, where some places have no,
+or few, disease events and/or a low population at
+risk of contracting the disease.
+
+-->

### 2019-08-01 review_article.Rmd
Examples of cartograms in literature

@@
 The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
+Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
 While the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
@@
 <!-- Displaying of cancer data on choropleths -->
-
+Choropleth maps can be useful devices for communicating information to public on a familiar map base.
 In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality. 
-Choropleth maps can be useful devices for communicating information to public on a familiar map base.
 A cancer atlas is a choropleth map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
@@
 <!-- Measures reported in cancer atlases -->
-
 Almost 100 years of cancer mapping in the United States and the United Kingdom has seen increased effectiveness in the presentation cancer statistics.
@@
 
-An Online Census Atlas for everyone [@AOCAFE] uses familiar choropleth, produced using generalized boundaries for six higher level geographies.
-Great Britain is recognizable and it is relatively easy to identify the higher geography levels such as Counties and Local Authorities.
+<!-- Cancer atlas literature -->
 
 <!--
-Medical application: TAAM 1965
-Actual shapes of counties were sacrificed to the extent necessary to retain as far as possible, common boundaries, relative locations, and the approximate contour of the State.
-
-There is no unique solution to the topology
-problem of preparing a population-by-area
-cartogram for a given population and its jurisdictional
-boundaries. Although the map areas
-of the various regions are fixed by their populations, their shapes and relative locations are not completely determined.
-
-Other characteristics of the population
-may also serve as the base for adjustment; for
-example, sex, race, parity, socio-economic level,
-or immune status.
 
@@
 For a single variable of interest, each map area is changed to emphasise the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].
+
+<!--
+Medical application: TAAM 1965
+Actual shapes of counties were sacrificed to the extent necessary to retain as far as possible, common boundaries, relative locations, and the approximate contour of the State.
+
+There is no unique solution to the topology
+problem of preparing a population-by-area
+cartogram for a given population and its jurisdictional
+boundaries. Although the map areas
+of the various regions are fixed by their populations, their shapes and relative locations are not completely determined.
+
+Other characteristics of the population
+may also serve as the base for adjustment; for
+example, sex, race, parity, socio-economic level,
+or immune status.
+-->
 

### 2019-08-02 review_article.Rmd
added a reference
index of webpages

--- 2019-07-30 review_article.Rmd
@@
 These atlases were key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets @MACM.
+
+<!--
+A particular spatial structure includes
+the individuals affected and how they are connected in
+communities, as well as the dynamics of these communities
+and their organization into larger units. @SAMGIS
+-->
 
@@
 
+<!--
+'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
+
+-->
+
 <!-- The supplementary statistics ??? -->
@@
 
-<!-- Going from choropleth to cartogram 
-the background variable assumed to be uniform (such as population density) can be adjusted so that it behaves as if uniformity were the case
-@GOINO
-
-A cartogram is an alternative map visual. 
-An intention distortion is applied to a geographic map base,
-
-A cartogram is a map purposely distorted so that its spatial properties represent quantities
-not directly associated with position on the globe. As thematic maps, cartograms emphasize
-the distribution of a variable by changing the area (or lengths) of objects on the map.
-Our approach maintains continuity, and preserves
-many local features of cartographic lines that provide visual clues to the identity of the distorted
-objects.
-@ACCAC
-
-@CD considers cartographic data analysis in practice. Presenting a clear guide for preparing cartograms for use with both qualitative and quantitative information.
-The authors recognise that the creation of cartograms was largely the work of professional cartographers until the innovation of geographical information systems that welcomed map users as map creators.
-These systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’.
-
-'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
-
--->
-<!--
-criticisms:
-Certain limitations to the standard graphic techniques have also become apparent, however.
-They are slow, and do not permit instant retrieval of information.
-They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually.
--->
-
-<!--
-Urbanisation has pushed people closer together,
--in workplaces and in homes. 
--this issue is not with the size of the administrative areas,
-- the increasing urbanisation is contributing to the changing density of population
-
-A particular spatial structure includes
-the individuals affected and how they are connected in
-communities, as well as the dynamics of these communities
-and their organization into larger units. @SAMGIS
--->
+<!-- What is a cartogram? -->
+A cartogram alters the map base with the intention of improving the presentation of the statistic of interest.
+For a single variable of interest, each map area is changed to emphasise the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].
 
@@
 <!-- Tobler map projections-->
-Choropleths may be considered true topological maps, however, if the land mass presented covers enough of the globe, there must be a transformation or distortion to display in 2D.
-The amount of distortion is related to the distance displayed @GAMP.
-The world projections reflect the frequent distortions seen from altering perspectives.
-Choropleth maps will always be distorted if they cover enough of the globe, as will photographs of the globe from space.
+Choropleths may be considered true topological maps, however, if the land mass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D.
+The amount of distortion is related to the distance covered by the landmass displayed @GAMP.
+World map projections reflect the frequent perspectives used to view the earth.
+Choropleth maps will always be distorted if they cover enough of the globe, just as photographs of the globe from space.
 Choropleth creation requires choosing a map projection that shows a favourable distortion of the geography for presenting the set of spatial information.
-Diagrams that do not specify a projection can be considered to have some unknown projection. <!-- ?? cite tobler.-->
+Diagrams that do not specify a projection can be considered to have some unknown projection. <!-- ??? cite tobler.-->
+
+<!-- Denominators -->
+If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed beyond choosing a projection. 
+The purposeful distortion of the map space is beneficial when a uniform density of the map base is desired.
+When a map base is transformed according to population density, population becomes a uniformly distributed background for the statistic presented [@GOINO].
 
@@
 <!-- Common variables used to create cartograms, e.g. population, mortality -->
-Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
 Cartograms provide an alternative visualisation method for statistical and geographical information.
@@
 <!--
+
+The transformations required to create cartograms was historically the role of professional cartographers.  
+Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’[@CD].
+
 @SE Researchers mapping spatial patterns of disease

### 2019-08-02 review_article.Rmd

--- 2019-08-01 review_article.Rmd
@@
 Transforming statistics to visualisations considers individual observations or observation points aggregated into communities and geographical units.
-Visualisations may be used for hypothesis generating and is often the first step in exploratory spatial data analysis. 
+Visualising diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH]. 
 Disease maps help to present geographic patterns that may have been overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
@@
 <!-- Displaying of cancer data on choropleths -->
+
+<!-- Displaying of cancer data on choropleths -->
 Choropleth maps can be useful devices for communicating information to public on a familiar map base.
@@
 These atlases were key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets @MACM.
+
+<!--
+A particular spatial structure includes
+the individuals affected and how they are connected in
+communities, as well as the dynamics of these communities
+and their organization into larger units. @SAMGIS
+-->
+
+<!-- Measures reported in cancer atlases -->
 
@@
 
-> The aim here is to see the whole, in as much detail as possible, at a glance.
-
 <!-- What is a cartogram? -->
-Cartograms provide an alternative visualisation method for statistical and geographical information.
-A cartogram alters the map base with the intention of improving the presentation of a statistic.
-The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC].
+A cartogram alters the map base with the intention of improving the presentation of the statistic of interest.
 For a single variable of interest, each map area is changed to emphasise the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].
-
-<!--
-Medical application: TAAM 1965
-Actual shapes of counties were sacrificed to the extent necessary to retain as far as possible, common boundaries, relative locations, and the approximate contour of the State.
-
-There is no unique solution to the topology
-problem of preparing a population-by-area
-cartogram for a given population and its jurisdictional
-boundaries. Although the map areas
-of the various regions are fixed by their populations, their shapes and relative locations are not completely determined.
-
-Other characteristics of the population
-may also serve as the base for adjustment; for
-example, sex, race, parity, socio-economic level,
-or immune status.
--->
 
@@
 <!-- Tobler map projections-->
-Choropleths may be considered true topological maps, however if the land mass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D.
+Choropleths may be considered true topological maps, however, if the land mass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D.
 The amount of distortion is related to the distance covered by the landmass displayed @GAMP.
-World map projections reflect the frequent perspectives used to view the earth. Choropleth maps will always be distorted if they cover enough of the globe, just as photographs of the globe from space.
+World map projections reflect the frequent perspectives used to view the earth.
+Choropleth maps will always be distorted if they cover enough of the globe, just as photographs of the globe from space.
 Choropleth creation requires choosing a map projection that shows a favourable distortion of the geography for presenting the set of spatial information.
 Diagrams that do not specify a projection can be considered to have some unknown projection. <!-- ??? cite tobler.-->
+
+<!-- Denominators -->
+If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed beyond choosing a projection. 
+The purposeful distortion of the map space is beneficial when a uniform density of the map base is desired.
+When a map base is transformed according to population density, population becomes a uniformly distributed background for the statistic presented [@GOINO].
 
@@
 
-A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, making interpretation of spatial patterns very difficult [@AOCAFE].
-The distortion of a cartogram prevents the density obscuring the spatial patterns [@TAAM]. 
-The spatial transformation of map regions relative to the data emphasises the data distribution instead of land size [@CBATCC].
-When visualising population statistics @ACTUC considers this equitable representation design 'more socially just, giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
-@CTTMB suggest that spatial socio-economic data is best presented on a cartogram for urban areas.
-@HEDP agrees that 'cancer occurs in people, not in geographical areas' and the map bases of population reflect this and avoid allocating 'undue prominence' to rural areas. 
-@MTMSIH encourage the use of cartograms to highlight small areas and uncover local-level inequalities.
+<!-- Why transform?-->
+<!-- Common variables used to create cartograms, e.g. population, mortality -->
+Cartograms provide an alternative visualisation method for statistical and geographical information.
+The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC].
+Cartograms may be seen as an extension of map transformations and projections. The favourable distortion is proportional to a value other other than actual earth size area [@NAC].
+
+<!--
+
+The transformations required to create cartograms was historically the role of professional cartographers.  
+Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’[@CD].
+
+@SE Researchers mapping spatial patterns of disease
+for small areas are confronted with a number of
+challenges, including cartographic design, the
+modifiable area unit problem, and also the “small
+number” problem, where some places have no,
+or few, disease events and/or a low population at
+risk of contracting the disease.
 

### 2019-08-05 review_article.Rmd
NISCC additions

--- 2019-08-01 review_article.Rmd
@@
 The spatial transformation of map regions relative to the data emphasises the data distribution instead of land size [@CBATCC].
-When visualising population statistics @ACTUC considers this equitable representation design 'more socially just, giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
+When visualising population statistics @ACTUC considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
 @CTTMB suggest that spatial socio-economic data is best presented on a cartogram for urban areas.
@@
 <!--
-@HEDP discusses the impact of electronic computer-assisted techniques. Disease atlases were implemented using both geographic (Howe 1963) and demographic bases (Howe 1970).
+ Disease atlases were implemented using both geographic (Howe 1963) and demographic bases (Howe 1970).
 
@@
 
+<!--
+@NISCC
+two advantages of the cartogram: (1) new relationships are often revealed visually, and ( Finally, and most important for the present discussion, they suggest that the ability to preserve shape of the statistical units drops with an increase in the number of these units, and in situations where the difference is great between the smallest and largest values mapped.
+Until we know how the map reader compares it to an ordinary thematic map, we are experimenting with a graphic device that may be reducing (or at least hindering) the flow of information from the map to the map reader.
+-->
+
 <!-- Overview of varieties -->
+<!-- Cartogram makers -->
+
+The creation of cartograms was largely in the hands of professional cartographers [@CD]. 
+@ACTUC discusses early approaches including John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball bearing approach and Tobler's (1973) computer programs.
+@HEDP discusses the impact of electronic computer-assisted techniques.
+Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’[@CD].
 
 There have been many algorithms presented, @SAIC provided a framework to investigate implementations and the "statistical accuracy, geographical accuracy, and topological accuracy".
-
-<!-- Cartogram makers -->
-
-The creation of cartograms was largely in the hands of cartographers. 
-@ACTUC discusses early approaches including John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball bearing approach and Tobler's (1973) computer programs.
 
@@
 
-A contiguous cartogram maintains connectivity of the map regions while areas are resized according to a statistic. This transformation often occurs at the expense of the shape [@CBATCC, @NAC, @TAAM].
+A contiguous cartogram maintains connectivity of the map regions while areas are resized according to a statistic.
+This transformation often occurs at the expense of the shape of areas [@CBATCC, @NAC, @TAAM].
 From a computer graphics perspective, @ACA believe it is a problem of 'map deformation' to account for the value assigned to each area, they provide three methods for creating value-by-area cartograms.
@@
 
-<!--
-
-@NISCC
-two advantages of the cartogram: (1) new relationships are often revealed visually, and ( 2 ) it is often more honest to use “other” space on the base map than land area.
-Finally, and most important for the present discussion, they suggest that the ability to preserve shape of the statistical units drops with an increase in the number of these units, and in situations where the difference is great between the smallest and largest values mapped.
-Until we know how the map reader compares it to an ordinary thematic map, we are experimenting with a graphic device that may be reducing (or at least hindering) the flow of information from the map to the map reader.
-
-The transformations required to create cartograms was historically the role of professional cartographers.  
-Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’[@CD].
-
--->
-
-describes the process of creating a population cartogram for Canada, using the 1966 Census.
-An intentional goal of this project was to maintain contiguity, while attempting to keep the actual shape of places.
+An intentional goal when creating the 1966 Census population cartogram for Canada was to maintain contiguity, while attempting to keep the actual shape of places.
 The end result was a 'very accurate isodemo-graphic map of Canada'.

### 2019-08-05 review_article.Rmd

@@
 output:
+  html_document:
+    df_print: paged
+    toc: yes
+    toc_depth: '2'
   pdf_document:
@@
     toc_depth: 2
-  html_document: default
 fontsize: 11pt
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
@@
 Source: Alberta Health IHDA Geographic.
-(2012) ", echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width='50%'}
+(2012) ", echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width='50%', eval = FALSE}
 
@@
 ```{r ci-viz3, fig.cap = "Example of CI visualisation for uncertainty representation in cancer mapping (2/3).
-Source: Pensylvania Cancer Atlas ", echo=FALSE, message=FALSE, warning=FALSE, out.width='50%', fig.align='center'}
+Source: Pensylvania Cancer Atlas ", echo=FALSE, message=FALSE, warning=FALSE, out.width='50%', fig.align='center', eval = FALSE}
 
@@
 Source: Centres for Disease Control and Prevention (CDC).
-United States Cancer Statistics: An Interactive Cancer Atlas (InCA) ", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '50%', fig.align = 'center'}
+United States Cancer Statistics: An Interactive Cancer Atlas (InCA) ", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '50%', fig.align = 'center', eval = FALSE}
 
@@
 ```{r ci-table, fig.cap= "Example of an interactive data table with CI upper and lower bounds used in cancer mapping.
-Source: Pennsylvania Cancer Atlas", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '50%', fig.align = 'center'}
+Source: Pennsylvania Cancer Atlas", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '50%', fig.align = 'center', eval = FALSE}
 
@@
 ```{r boxplot-example, fig.cap= "Example of boxplot used in cancer mapping.
-Source: Source: Pensylvania Cancer Atlas ", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '40%', fig.align = 'center'}
+Source: Source: Pensylvania Cancer Atlas ", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '40%', fig.align = 'center', eval = FALSE}
 
@@
 
-A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, making interpretation of spatial patterns very difficult [@AOCAFE].
+<!-- Why transform?-->
+<!-- Common variables used to create cartograms, e.g. population, mortality -->
+Cartograms provide an alternative visualisation method for statistical and geographical information.
+The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC].
+Cartograms may be seen as an extension of map transformations and projections. The favourable distortion is proportional to a value other other than actual earth size area [@NAC].
+A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, making interpretation of spatial patterns very difficult.
 The distortion of a cartogram prevents the density obscuring the spatial patterns [@TAAM]. 
@@
 
-## Centroid displays
-
-Dot plot: one dot for each region, coloured, and placed at centroid.
-
-@NAC gives an example of
-
-Plotting centroids on top of geographies. (Size is kept constant)
-
-Replies on the idea that every area is important, no matter the size.
-This gives equal emphasis to every area, allows distributions and relationships between neighbours to become more clear.
-
-<!-- 
-TVSSS parliamentary cartogram where each constituency is represented by an oval of equal size. Many ovals overlap because the cartogram is still in the process of being made. Each
--->
-
 # A critique of mapping methods
@@
 
-There are two keys to drive the choice of display: the properties of the visualisation, and the ease or accuracy of information extraction for map users @SAMGIS.
+There are two keys presented by @SAMGIS to drive the choice of display: 
+ - the properties of the visualisation, and 
+ - the ease or accuracy of information extraction for map users
 
-<!-- @CPISACA contrasted several online atlases in a 2006 paper, consider for structure? -->
+<!-- @CPISACA contrasted several online atlases in a 2006 paper, consider for structure?
 
@@
 - Relationship to geography
-
 - Show using cancer examples
+-->
 
@@
 
-Recent developments of technology allowed interactive
+Recent developments in technology allowed interactive web atlases.
 
-'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
+> 'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
 
-lends to temporal pattern exploration
+The intention of interactivity and animation is to allow users to answer their own questions.
 
+<!--
+Using e-Delphi to Evaluate the Pennsylvania Cancer Atlas - explains PCA and reasons for interactivity.
+-->
+
+<!--
 @ECGC ?? Highlight the value of animating contiguous to see changes over time, US can be recognisable but animation aides interpretation 
+-->
 

### 2019-08-12 review_article.Rmd
add tile grid of chloropleth maps
first pass of chloropleth maps. May require a few of the images to be edited further so that sizes are uniform

--- 2019-08-02 review_article.Rmd
@@
 -->
+
+```{r chloropleth-grid, echo=FALSE, message=FALSE}
+
+library(grid)
+library(png)
+library(gridExtra)
+
+plot1 <- png::readPNG('cancer_map_images/chloropleths/1.png')
+plot2 <- png::readPNG('cancer_map_images/chloropleths/14.chloropleth1.png')
+plot3 <- png::readPNG('cancer_map_images/chloropleths/15.chloropleth.png')
+plot4 <- png::readPNG('cancer_map_images/chloropleths/3.chloropleth.png')
+plot5 <- png::readPNG('cancer_map_images/chloropleths/5.chloropleth.png')
+plot6 <- png::readPNG('cancer_map_images/chloropleths/6.chloropleth.png')
+plot6 <- png::readPNG('cancer_map_images/chloropleths/7.chloropleth.png')
+plot7 <- png::readPNG('cancer_map_images/chloropleths/8.chloropleth.png')
+plot8 <- png::readPNG('cancer_map_images/chloropleths/20.chloropleth1.png')
+plot9 <- png::readPNG('cancer_map_images/chloropleths/21.chloropleth2.png')
+plot10 <- png::readPNG('cancer_map_images/chloropleths/23.chloropleth.png')
+plot11 <- png::readPNG('cancer_map_images/chloropleths/24.chloropleth.png')
+plot12 <- png::readPNG('cancer_map_images/chloropleths/25.chloropleth.png')
+plot13 <- png::readPNG('cancer_map_images/chloropleths/26.chloropleth.png')
+plot14 <- png::readPNG('cancer_map_images/chloropleths/27.chloropleth.png')
+plot15 <- png::readPNG('cancer_map_images/chloropleths/29.chloropleth.png')
+plot16 <- png::readPNG('cancer_map_images/chloropleths/30.chloropleth.png')
+
+gridExtra::grid.arrange(grid::rasterGrob(plot1),grid::rasterGrob(plot2), grid::rasterGrob(plot3), 
+                        grid::rasterGrob(plot4), grid::rasterGrob(plot5), grid::rasterGrob(plot6),
+                        grid::rasterGrob(plot7), grid::rasterGrob(plot8), grid::rasterGrob(plot9),
+                        grid::rasterGrob(plot10),grid::rasterGrob(plot11), grid::rasterGrob(plot12), 
+                        grid::rasterGrob(plot13), grid::rasterGrob(plot14), grid::rasterGrob(plot15), 
+                        grid::rasterGrob(plot16), nrow=4)
+
+```
 

### 2019-08-13 review_article.Rmd
Fix quotes

--- 2019-08-05 review_article.Rmd
@@
 output:
+  pdf_document:
+    number_sections: yes
+    toc: yes
+    toc_depth: 2
   html_document:
@@
     toc_depth: '2'
-  pdf_document:
-    number_sections: yes
-    toc: yes
-    toc_depth: 2
 fontsize: 11pt
@@
 
-## Public face
+## Publicly available atlases using choropleth methods
 
@@
 
-## Cartograms
+## Cartograms as an alternative display
 
@@
 
->If, for instance, it is desirable that areas on a map have boundaries which are as simple as possible, why not draw the areas as simple shapes in the first place?
+>"If, for instance, it is desirable that areas on a map have boundaries which are as simple as possible, why not draw the areas as simple shapes in the first place?"
 
@@
 
->designing a map tailored to precise goals [is] easier than forcing a single map to accommodate diverse objectives - @CPISACA
+>"designing a map tailored to precise goals [is] easier than forcing a single map to accommodate diverse objectives" @CPISACA
 
@@
 
-> 'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
+>"Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps" @CPISACA
 

### 2019-08-13 review_article.tex
Fix quotes

@@
 
-Utilising maps to present statistics has been widely used for centuries.
-These maps connect data and/or statistics to the geographical
-representation of areas that are generally familiar to the audience.
-However, it is not enough to be recognisable, if the meaning of the
-statistic cannot be seen. This is particularly relevant to cancer
-statistics, as cancer outcomes directly relate to the people living
-within a geographical area. In situations where the people are of
-interest, not the land they live on, it is reasonable to explore views
-that enhance the communication of the cancer statistics. This has
-spurred innovations over the previous centuries to enhance the shapes
-and maps presented to effectively communicate cancer outcomes, and
-health outcomes more broadly.
-
 \hypertarget{introduction}{%
 \section{Introduction}\label{introduction}}
+
+Maps have been adopted to present geospatial statistics for centuries.
+Maps connect data and statistics to the geographical representation of
+areas that are generally familiar to the audience. However, it is not
+enough for areas on maps to be recognisable, if information regarding
+the distribution of the statistic cannot be gleaned. In situations where
+the people are of interest, not the land they live on, it is reasonable
+to explore views that enhance the communication of the cancer
+statistics. Identifying and explaining spatial structures, patterns, and
+processes involves considering the individuals in communities and
+organising communities into representable units (Moore and Carpenter
+\protect\hyperlink{ref-SAMGIS}{1999}). This has spurred innovations over
+the previous centuries to enhance the shapes and maps presented to
+effectively communicate cancer outcomes, and health outcomes more
+broadly. Cancer statistics directly represent and relate to the people
+living within individual geographical areas.
+
+The visualisation methods used to present cancer statistics will depend
+on the intended message and users of the map. Transforming statistics to
+visualisations considers individual observations or observation points
+aggregated into communities and geographical units. Visualising diseases
+on maps is often the first step in exploratory spatial data analysis and
+effectively helps in the formulation of hypotheses (Jahan et al.
+\protect\hyperlink{ref-MTMSIH}{2018}). Disease maps help to present
+geographic patterns that may have been overlooked in a table, obscuring
+the geospatially related statistics (Moore and Carpenter
+\protect\hyperlink{ref-SAMGIS}{1999}). By providing a visual
+representation of cancer outcomes, geographic patterns of disease are
+able to be identified and effectively addressed.
 
@@
 
-A choropleth map is used to show the spatial relationship, or
-variability, of measurements. Created as a storage device which
-preserves locations for geographically ordered data, with usage dating
-back to the 1800s (Berry, Morrill, and Tobler,
-\protect\hyperlink{ref-GOINO}{n.d.}). They are constructed by drawing
-the given geographic or political boundaries, and filling the shapes
+A choropleth map is used to display the spatial characteritics of a
+relationship, or variability, of measurements. As an alternative storage
+device to a table, it preserves locations for geographically ordered
+data, with usage dating back to the 1800s (Berry, Morrill, and Tobler,
+\protect\hyperlink{ref-GOINO}{n.d.}). A choropleth is constructed by
+drawing the geographic or political boundaries, and filling the shapes
 with colours to represent values of a measured variable (Tufte
-\protect\hyperlink{ref-EI}{1990}). Early versions used symbols or
-patterns instead of colour. Bell et al.
-(\protect\hyperlink{ref-CPISACA}{2006}) discusses the use of choropleth
-maps for visualising cancer data, and Walter
+\protect\hyperlink{ref-EI}{1990}). Early versions of choropleth maps
+used symbols or patterns instead of colour. Bell et al.
+(\protect\hyperlink{ref-CPISACA}{2006}) discusses their use in
+visualising cancer data, and Walter
 (\protect\hyperlink{ref-DMAHP}{2001}) gives an overview of the
-development of the use of choropleth maps for displaying disease data. A
-choropleth map is a true map of the topology, constructed for visual
-inspection of spatial patterns across a familiar geographic form. Figure
-1 shows a choropleth map where each state of the United States of
-America is coloured by the average annual rate of new cases of lung and
-bronchus cancer from years 2012 to 2016.
+development of these maps for displaying disease data. Just as
+geographers are no longer the only creators of maps, Bell et al.
+(\protect\hyperlink{ref-CPISACA}{2006}) suggests the audiences of
+spatial health data analysis have extended beyond researchers to the
+public, policymakers and the media. A choropleth map is a true map of
+the topology, constructed for visual inspection of spatial patterns
+across a familiar geographic form. Figure 1 shows a choropleth map where
+each state of the United States of America is coloured by the average
+annual rate of new cases of lung and bronchus cancer from years 2012 to
+2016.
 
 \begin{figure}
-\centering
-\includegraphics{figures/ggchoro.png}
-\caption{``A choropleth map of the United States of America.''}
+
+{\centering \includegraphics[width=0.8\linewidth]{figures/ggchoro}
+
+}
+
+\caption{A choropleth map of the United States of America.}\label{fig:choro}
 \end{figure}
@@
 audiences of spatial health data analysis have extended beyond
-researchers to the public, policymakers and the media. However, while
-the areas are recognisable shapes, they are often politically driven
-boundaries with individual areas being of non-uniform size, containing
-different population densities and subject to change over time. The
-different population and geographical sizes of administrative areas can
-attract attention to the shades of the unpopulated but large areas
-(Tufte \protect\hyperlink{ref-EI}{1990}). Choropleths can inhibit visual
+researchers to the public, policymakers and the media. While the areas
+are recognisable shapes, they are often politically driven boundaries
+with individual areas being of non-uniform size, containing different
+population densities and subject to change over time. The different
+population and geographical sizes of administrative areas can attract
+attention to the shades of the unpopulated but large areas (Tufte
+\protect\hyperlink{ref-EI}{1990}). Choropleths can inhibit visual
 inference when presenting human related statistics as the display may
@@
 
-In epidemiology, choropleths are often used as a tool to study the
-spatial distribution of cancer incidence and mortality.
+Choropleth maps can be useful devices for communicating information to
+public on a familiar map base. In epidemiology, choropleths are often
+used as a tool to study the spatial distribution of cancer incidence and
+mortality. A cancer atlas is a choropleth map, or collection of maps,
+representing cancer incidence and mortality for a country, or group of
+countries. d'Onofrio et al. (\protect\hyperlink{ref-MACM}{2016}) provide
+the definition and reports that cancer atlases began with Haviland's
+maps in 1875. The data collection methods of cancer mortality rates
+across regions, and the administrative control within regions lends
+itself to choropleth visualisation. The increasing development and use
+of disease maps can be attributed to the availability of geographic
+information system software (Exeter \protect\hyperlink{ref-SE}{2017}).
+Early work in US cancer atlases can be attributed to Burbank ???, and UK
+cancer atlases to Howe 1963 ???. The choropleth maps presented levels
+via hatchings or dots on a black and white scale. These atlases were key
+to developing hypotheses regarding areas with unusually high rates,
+geographic correlations, work related exposures, and high risk diets
+d'Onofrio et al. (\protect\hyperlink{ref-MACM}{2016}).
 
 Almost 100 years of cancer mapping in the United States and the United
-Kingdom has increased the effectiveness in the presentation of unbiased
-rates. The increasing development and use of disease maps can be
-attributed to the availability of geographic information system software
-(Exeter \protect\hyperlink{ref-SE}{2017}). The data collection methods
-of cancer mortality rates across regions, and administrative control
-within regions lends itself to choropleth visualisation. Mortality rates
-are now often presented as relative rates of risk across the population,
-and age adjusted to correct for the the higher prevalence of cancers in
-older populations. Howe (\protect\hyperlink{ref-HEDP}{1989}) describes
-Stocks development of the standardised mortality ratios through the
-1930s. The choropleth maps used at this time presented levels of cancer
-via hatchings on a black and white scale.
+Kingdom has seen increased effectiveness in the presentation cancer
+statistics. Mortality rates are now often presented as relative rates of
+risk across the population, and age adjusted to correct for the the
+higher prevalence of cancers in older populations. Howe
+(\protect\hyperlink{ref-HEDP}{1989}) describes Stock's development of
+the standardised mortality ratios through the 1930s. Table 1 presents
+summarises the measures presented in published cancer atlases, and
+provides a definition of each measure.
 
-Mortality rates are now often presented as relative rates of risk across
-the population, and age adjusted. Howe
-(\protect\hyperlink{ref-HEDP}{1989}) describes Stocks' development of
-the standardised mortality ratios through the 1930s. The choropleth maps
-used presented levels of cancer via hatchings on a black and white
-scale.
-
-Examples of atlases: Kraak 1998, Kraak and Ormeling 2011 Bertin 1967
-
-\hypertarget{public-face}{%
-\subsection{Public face}\label{public-face}}
-
-A grey literature review conducted by (ref) identified 33 publicly
-available cancer atlases published on the internet between 01/01/2010 to
-11/11/2015. Reflecting the high use of chloropleth maps founds in the
-literature, outlined in the section above, all the public facing maps
-identified in this review were also chloropleth maps.
-
-\hypertarget{publishers}{%
-\subsubsection*{Publishers}\label{publishers}}
-\addcontentsline{toc}{subsubsection}{Publishers}
-
-The majority of atlases were published by non-commercial organisations,
-including not-for-profits (NFPs), government, research organisations,
-advocacy groups or a partnership between an NFP \& government. Only one
-map was published by a commercial entity (Maps of Cancer Mortality Rates
-in Spain), in this case a media organisation El Pais\footnote{\url{https://elpais.com/}}.
-
-\hypertarget{reported-measures}{%
-\subsubsection*{Reported measures}\label{reported-measures}}
-\addcontentsline{toc}{subsubsection}{Reported measures}
-
-The majority of maps identified in this study reported age adjusted
-rates of either incidence, mortality or survival. The list below
-summarises these measures, and provides a definition of each measure.
-
-\hypertarget{geographical-coverage}{%
-\subsubsection*{Geographical coverage}\label{geographical-coverage}}
-\addcontentsline{toc}{subsubsection}{Geographical coverage}
-
-Identified cancer atlases covered geographies from all around the world:
-4 were global, 3 from Australia (AUS), 11 from the United States (US), 2
-from Canada (CAN), 7 from the United Kingdom (UK), 2 from Spain, 1 from
-Switzerland, 1 from Germany, 1 from Norway, and 1 covering the European
-Union. Not all maps had a national focus and 10 covered a region or
-state rather than an entire nation. The states or counties/regions
-covered were South Australia (AUS), Queensland (AUS), Ontario (CAN),
-Valencia (Spain), Pennsylvania county Massachusetts (US), New Hampshire
-(US), Cape Cod (US), Missouri (US), Florida (US), New York State (US)
-and Arizona (US).
-
-\newpage
+Table 1: Measures used to report cancer statistics
 
 \begin{longtable}[]{@{}ll@{}}
-\caption{(\#tab:measures) Measures used to report cancer
-statistics}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.28\columnwidth}\raggedright
-Measure\strut
-\end{minipage} & \begin{minipage}[b]{0.66\columnwidth}\raggedright
-Details\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
 \toprule
@@
 
-\hypertarget{statistical-uncertainty}{%
-\subsubsection{Statistical uncertainty}\label{statistical-uncertainty}}
+\hypertarget{supporting-material-in-cancer-atlases}{%
+\subsubsection{Supporting material in cancer
+atlases}\label{supporting-material-in-cancer-atlases}}
 
-Cancer atlases were considered to report uncertainty to the non-expert
-user if they included a measure of statistical uncertainty either within
-or alongside the map. Maps that only reported this information within
-the supplementary material were not considered to have directly
-attempted to report uncertainty.
+A map communicates quickly and draws attention to prominent geographies,
+but an atlas is often supplemented with supporting statistics.
+Cruickshank's (1947) as cited by Walter
+(\protect\hyperlink{ref-DMAHP}{2001}), discusses using visuals as
+`formal statistical assessment of the spatial pattern'. However, when
+presenting cancer atlases, d'Onofrio et al.
+(\protect\hyperlink{ref-MACM}{2016}) believes the intuition derived must
+be `validated by rigorous statistical analyses.'
 
-The review did not reveal any novel uncertainty visualisation approaches
-or visualisations. Maps used standard and well known measures including
-credible intervals and standard deviation, statistical significance, box
-plots and distributions. These maps ranged from static pdfs or
-infographics to interactive online resources. The interactivity of the
-more modern maps enabled uncertainty information to be incorporated
-without cluttering the screen, such as in a tool tip feature.
+These additional statistics often include a measure of the statistical
+uncertainty of the values of the statistics presented in a choropleth.
+In the review of atlases in the public domain, atlases were considered
+to report uncertainty to the non-expert user if they included a measure
+of statistical uncertainty either within or alongside the map. Maps that
+only reported this information within the supplementary material were
+not considered to have directly attempted to report uncertainty. The
+maps considered used standard and well known measures including credible
+intervals and standard deviation, statistical significance, box plots
+and distributions. These maps ranged from static pdfs or infographics,
+to interactive online resources.
 
-\begin{figure}
+The interactivity of the more modern maps enabled supplementary
+information to be incorporated without cluttering the screen, such as in
+a tool tip feature.
 
-{\centering \includegraphics[width=0.8\linewidth]{figures/Alberta_CI_viz}
+\hypertarget{publicly-available-atlases-using-choropleth-methods}{%
+\subsection{Publicly available atlases using choropleth
+methods}\label{publicly-available-atlases-using-choropleth-methods}}
 
-}
+Cancer maps are effective communication tools for a general or
+non-expert audience. They are commonly used in the public domain to
+communicate results of sophisticated statistical analyses. The heavy use
+of chloropleth maps within the research literature is reflected in the
+types of maps that are found in the public domain. A grey literature
+review conducted by (ref) identified 33 cancer atlases published on the
+internet between January 2010 and November 2015. These chloropleth maps
+were mostly published by non-commercial organisations, including
+not-for-profits (NFPs), government, research organisations, advocacy
+groups or a partnership between an NFP \& government. Only one map was
+published by a commercial entity.
 
-\caption{Example of CI visualisation for uncertainty representation in cancer mapping (1/3). Source: Alberta Health IHDA Geographic. (2012) }\label{fig:ci-viz2}
-\end{figure}
+The cancer atlases covered geographies from all around the world, four
+were global atlases. Most focussed on single nations, the United States
+was considered by eleven atlases, the United Kingdom by seven, followed
+by three of Australia, two of Canada, and one of each from Switzerland,
+Germany, Norway. One atlas covered the European Union. Not all maps had
+a national focus and ten covered a region or state rather than an entire
+nation. The states or counties/regions covered were South Australia
+(AUS), Queensland (AUS), Ontario (CAN), Valencia (Spain), Pennsylvania
+county Massachusetts (US), New Hampshire (US), Cape Cod (US), Missouri
+(US), Florida (US), New York State (US) and Arizona (US).
 
-\begin{figure}
-
-{\centering \includegraphics[width=0.55\linewidth]{figures/Pennsylvania_CI_viz}
-
-}
-
-\caption{Example of CI visualisation for uncertainty representation in cancer mapping (2/3). Source: Pensylvania Cancer Atlas }\label{fig:ci-viz3}
-\end{figure}
+\hypertarget{examples-of-supporting-statistics-in-public-facing-atlases}{%
+\subsubsection{Examples of supporting statistics in public facing
+atlases}\label{examples-of-supporting-statistics-in-public-facing-atlases}}
 
@@
 measure of uncertainty. The most common measure used to represent
-uncertainty were credible or confidence intervals (CIs). CIs were either
-visualised by including their bounds in a scatterplot or graph of
-estimates vs region (see Figures @ref(fig:ci-viz2)\footnote{Alberta
-  Health IHDA Geographic. (2012). Age-Standardised Incidence Rate of
-  COPD, 2011. Retrieved from:
-  \url{http://www.health.alberta.ca/health-info/IHDA-geographic/COPD/incidence-agestandard/atlas.html?epik=0GJSpE_IW34lx}},
-and @ref(fig:ci-viz3)\footnote{Centres for Disease Control and
-  Prevention (CDC). (n.d). United States Cancer Statistics: An
-  Interactive Cancer Atlas (InCA). Retrieved
-  from:\url{https://nccd.cdc.gov/DCPC_INCA/}} positioned next to the
-map, or reported numerically through the CI upper and lower bounds
-listed in a data table (see Figure 3.4\footnote{Pennsylvania Cancer
-  Atlas. (n.d). Retrieved from:
-  \url{https://www.geovista.psu.edu/grants/CDC/?epik=0lJSpE_IW34lx}}).
-Of those that visualised the CIs, 30\% (n=10) embedded the visualisation
-within a tool tip function which visualised the CI when the mouse
-hovered over the relevant area (see Figure 3.5\footnote{International
-  Agency for Research on Cancer. (2017). Atlas of Cancer Mortality in
-  the European Union and European Economic Area 1993-1997, Annex 4 -
-  Cancer mortality maps by site. Retrieved from:
-  \url{http://www.iarc.fr/en/publications/pdfs-online/epi/sp159/}}).
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.6\linewidth]{figures/CDC_CI_viz_tooltip}
-
-}
-
-\caption{Example of CI visualisation for uncertainty representation in cancer mapping (3/3).
-Source: Centres for Disease Control and Prevention (CDC).
-United States Cancer Statistics: An Interactive Cancer Atlas (InCA) }\label{fig:ci-tooltip}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.5\linewidth]{figures/uncert_table}
-
-}
-
-\caption{Example of an interactive data table with CI upper and lower bounds used in cancer mapping. Source: Pennsylvania Cancer Atlas}\label{fig:ci-table}
-\end{figure}
+uncertainty were credible or confidence intervals (CIs).
 
@@
 \caption{Example of standard deviation visualised in cancer mapping}\label{fig:stndvn-example}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.4\linewidth]{figures/Boxplot2}
-
-}
-
-\caption{Example of boxplot used in cancer mapping. Source: Source: Pensylvania Cancer Atlas }\label{fig:boxplot-example}
 \end{figure}
@@
 
-\hypertarget{cartograms}{%
-\subsection{Cartograms}\label{cartograms}}
+\hypertarget{cartograms-as-an-alternative-display}{%
+\subsection{Cartograms as an alternative
+display}\label{cartograms-as-an-alternative-display}}
+
+A cartogram alters the map base with the intention of improving the
+presentation of the statistic of interest. For a single variable of
+interest, each map area is changed to emphasise the distribution by
+representing the corresponding value, in comparison to the value of the
+other areas (Dougenik, Chrisman, and Niemeyer
+\protect\hyperlink{ref-ACCAC}{1985}).
 
 Choropleths may be considered true topological maps, however, if the
-land mass presented covers enough of the globe, there must be a
-transformation or distortion to display in 2D. The amount of distortion
-is related to the distance displayed Tobler
-(\protect\hyperlink{ref-GAMP}{1963}). The world projections reflect the
-frequent distortions seen from altering perspectives. Choropleth maps
-will always be distorted if they cover enough of the globe, as will
+land mass displayed covers enough of the globe, there must be a
+transformation or distortion to display the land in 2D. The amount of
+distortion is related to the distance covered by the landmass displayed
+Tobler (\protect\hyperlink{ref-GAMP}{1963}). World map projections
+reflect the frequent perspectives used to view the earth. Choropleth
+maps will always be distorted if they cover enough of the globe, just as
 photographs of the globe from space. Choropleth creation requires
@@
 projection.
+
+If the statistic presented on the map base relies on physical distance
+and is influenced by the topology there is no transformation needed
+beyond choosing a projection. The purposeful distortion of the map space
+is beneficial when a uniform density of the map base is desired. When a
+map base is transformed according to population density, population
+becomes a uniformly distributed background for the statistic presented
+(Berry, Morrill, and Tobler, \protect\hyperlink{ref-GOINO}{n.d.}).
 
@@
 
-Waldo Tobler (\protect\hyperlink{ref-TVSSS}{2012}) explores many
-graphical techniques, and suggests there are particular methods for
-particular purposes. Cartograms provide an alternative visualisation
-method for statistical and geographical information. The key difference
-between a choropleth and a cartogram is the desirable augmentation of
-the size, shape or distance of geographical areas (Dorling
+If the statistic presented on the map base relies on physical distance
+and is influenced by the topology there is no transformation needed
+beyond choosing a projection. The purposeful distortion of the map space
+is beneficial when a uniform density of the map base is desired, Dorling
+(\protect\hyperlink{ref-ACTUC}{2011}) suggests `population distribution
+is often extremely uneven in former British colonies', this makes the
+distortion necessary. When a map base is transformed according to
+population density, population becomes a uniformly distributed
+background for the statistic presented (Berry, Morrill, and Tobler,
+\protect\hyperlink{ref-GOINO}{n.d.}). Using choropleth maps for
+population characteristics requires graphic distortions when the
+population concerned varies greatly in density (Griffin
+\protect\hyperlink{ref-CTTMB}{1980}). When implementing a distortion of
+the geographical shape according to population, an area cartogram (Olson
+\protect\hyperlink{ref-NAC}{1976}), population-by-area cartograms
+(Levison and Haddon Jr \protect\hyperlink{ref-TAAM}{1965}), or
+iso-demographic map is the result.
+
+Cartograms provide an alternative visualisation method for statistical
+and geographical information. The key difference between a choropleth
+and a cartogram is the desirable augmentation of the size, shape or
+distance of geographical areas (Dorling
 \protect\hyperlink{ref-ACTUC}{2011}). Cartograms may be seen as an
@@
 distortion is proportional to a value other other than actual earth size
-area (Olson \protect\hyperlink{ref-NAC}{1976}).
-
-Using choropleth maps for population characteristics requires graphic
-distortions when the population concerned varies greatly in density
-(Griffin \protect\hyperlink{ref-CTTMB}{1980}). When implementing a
-distortion of the geographical shape according to population, an area
-cartogram (Olson \protect\hyperlink{ref-NAC}{1976}), or iso-demographic
-map is the result. When visualising population statistics Dorling
-(\protect\hyperlink{ref-ACTUC}{2011}) considers this equitable
-representation design `more socially just, giving due attention to all
+area (Olson \protect\hyperlink{ref-NAC}{1976}). A disadvantage of the
+conventional map is that sparsely populated rural areas may be
+emphasized, whereas the areas representing cities are very small, making
+interpretation of spatial patterns very difficult. The distortion of a
+cartogram prevents the density obscuring the spatial patterns (Levison
+and Haddon Jr \protect\hyperlink{ref-TAAM}{1965}). The spatial
+transformation of map regions relative to the data emphasises the data
+distribution instead of land size (Kocmoud and House
+\protect\hyperlink{ref-CBATCC}{1998}). When visualising population
+statistics Dorling (\protect\hyperlink{ref-ACTUC}{2011}) considers this
+equitable representation design `more socially just', or honest (Dent
+\protect\hyperlink{ref-NISCC}{1972}), giving due attention to all
 members of the population and reducing the visual impact of large areas
@@
 socio-economic data is best presented on a cartogram for urban areas.
-Howe (\protect\hyperlink{ref-HEDP}{1989}) agrees that 'cancer occurs in
+Howe (\protect\hyperlink{ref-HEDP}{1989}) agrees that `cancer occurs in
 people, not in geographical areas' and the map bases of population
@@
 cartograms to highlight small areas and uncover local-level
-inequalities. Using health estimates from large areas can prevent
-drawing attention to the inequalities.
+inequalities.
 
-The area on the map space can also be used represent a value. There have
-been many algorithms presented, Nusrat and Kobourov
+The creation of cartograms was largely in the hands of professional
+cartographers (Kraak \protect\hyperlink{ref-CD}{2017}). Dorling
+(\protect\hyperlink{ref-ACTUC}{2011}) discusses early approaches
+including John Hunter and Jonathan Young (1968) and Durham's wooden tile
+method, Skoda and Robertson's (1972) steel ball bearing approach and
+Tobler's (1973) computer programs. Howe
+(\protect\hyperlink{ref-HEDP}{1989}) discusses the impact of electronic
+computer-assisted techniques. Geographical information systems allowed
+map users, and researchers to implement their own cartograms, but these
+systems are utilised depending on `the effectiveness, efficiency, and
+satisfaction of the map products (Nielsen 1994)'(Kraak
+\protect\hyperlink{ref-CD}{2017}).
+
+There have been many algorithms presented, Nusrat and Kobourov
 (\protect\hyperlink{ref-SAIC}{2016}) provided a framework to investigate
 implementations and the ``statistical accuracy, geographical accuracy,
-and topological accuracy''. The spatial transformation of map regions
-relative to the data emphasises the data distribution instead of land
-size (Kocmoud and House \protect\hyperlink{ref-CBATCC}{1998}).
-
-Cancer statistics and atlases
-
-The creation of cartograms was largely in the hands of cartographers.
-Dorling (\protect\hyperlink{ref-ACTUC}{2011}) discusses early approaches
-including John Hunter and Jonathan Young (1968) and Durham's wooden tile
-method, Skoda and Robertson's (1972) steel ball bearing approach and
-Tobler's (1973) computer programs.
+and topological accuracy''.
 
@@
 ``best'' map' (Monmonier and Schnell, 1988).
-
-`Firstly, that no such transformation has a unique solution, leaving the
-subjective choice in the hands of the individual map designer.'
-
-Design goals: - to maintain as much as possible of the intricate detail
-- to maximize the simplicity of the transformation - maintenance of
-directional accuracy - attempt to influence and direct the map user's
-attention,
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\tightlist
-\item
-  that the size of the electoral subdivisions should be made
-  proportional to their respective electoral populations,
-\item
-  that the contiguity relationships between the DCU's should be
-  maintained,
-\item
-  that true directions should be retained from the centre of the
-  transformation,
-\item
-  that the main shape characteristics of the DCU's should be preserved,
-  and
-\item
-  that the map should possess the overall shape of the built-up area.
-\end{enumerate}
 
@@
 
-A contiguous cartogram allows the map space to highlight the
-distribution of the variable. These cartograms maintain connectivity of
-the map regions while areas are resized according to a statistic, this
-often must occur at the expense of the shape (Kocmoud and House
-\protect\hyperlink{ref-CBATCC}{1998},@NAC, @TAAM). From a computer
+A contiguous cartogram maintains connectivity of the map regions while
+areas are resized according to a statistic. This transformation often
+occurs at the expense of the shape of areas (Kocmoud and House
+\protect\hyperlink{ref-CBATCC}{1998}, @NAC, @TAAM). From a computer
 graphics perspective, Min Ouyang and Revesz
@@
 
-describes the process of creating a population cartogram for Canada,
-using the 1966 Census. An intentional goal of this project was to
-maintain contiguity, while attempting to keep the actual shape of
-places. The end result was a `very accurate isodemo-graphic map of
-Canada'. This intentional design goal coincided with the rising interest
-in urban geography.
+An intentional goal when creating the 1966 Census population cartogram
+for Canada was to maintain contiguity, while attempting to keep the
+actual shape of places. The end result was a `very accurate
+isodemo-graphic map of Canada'. This intentional design goal coincided
+with the rising interest in urban geography.
 
@@
 \begin{quote}
-If, for instance, it is desirable that areas on a map have boundaries
+``If, for instance, it is desirable that areas on a map have boundaries
 which are as simple as possible, why not draw the areas as simple shapes
-in the first place?
+in the first place?''
 \end{quote}
@@
 thousand' areas.
-
-Dorling (\protect\hyperlink{ref-ACTUC}{2011}) suggests `population
-distribution is often extremely uneven in former British colonies'.
 
@@
 
-\hypertarget{centroid-displays}{%
-\subsection{Centroid displays}\label{centroid-displays}}
-
-Dot plot: one dot for each region, coloured, and placed at centroid.
-
-Olson (\protect\hyperlink{ref-NAC}{1976}) gives an example of
-
-Plotting centroids on top of geographies. (Size is kept constant)
-
-Replies on the idea that every area is important, no matter the size.
-This gives equal emphasis to every area, allows distributions and
-relationships between neighbours to become more clear.
-
 \hypertarget{a-critique-of-mapping-methods}{%
@@
 \begin{quote}
-designing a map tailored to precise goals {[}is{]} easier than forcing a
-single map to accommodate diverse objectives - Bell et al.
+``designing a map tailored to precise goals {[}is{]} easier than forcing
+a single map to accommodate diverse objectives'' Bell et al.
 (\protect\hyperlink{ref-CPISACA}{2006})
@@
 
-With that in mind, the intended user and message to communicate should
-drive map selection. first is the nature or inherent properties of the
-device itself second is the ease and accuracy with which map users are
-able to extract information from it
+Waldo Tobler (\protect\hyperlink{ref-TVSSS}{2012}) explores many
+graphical techniques, and suggests there are particular methods for
+particular purposes. To choose an approporiate map display, the map
+creator must consider the intended user, and message the map will
+communicate. It is the objectives of the investigator that will drive
+the choice of representation (Bell et al.
+\protect\hyperlink{ref-CPISACA}{2006}).
 
-Moore and Carpenter (\protect\hyperlink{ref-SAMGIS}{1999}) suggests it
-is the ``investigators' objectives'' that drive the ``representation of
-diseases on maps''. -infectious diseases may require choropleth methods
+There are two keys presented by Moore and Carpenter
+(\protect\hyperlink{ref-SAMGIS}{1999}) to drive the choice of display: -
+the properties of the visualisation, and - the ease or accuracy of
+information extraction for map users
 
-`Where control of the message is important, static maps will continue to
-be the most effective, although good tables, graphs, and explanatory
+\hypertarget{animation-and-interactivity}{%
+\section{Animation and
+Interactivity}\label{animation-and-interactivity}}
+
+Recent developments in technology allowed interactive web atlases.
+
+\begin{quote}
+``Where control of the message is important, static maps will continue
+to be the most effective, although good tables, graphs, and explanatory
 text are still needed in order to ensure that different people will see
-the same thing in the maps' Bell et al.
+the same thing in the maps'' Bell et al.
 (\protect\hyperlink{ref-CPISACA}{2006})
+\end{quote}
 
-(only worth including if it is possible for us to implement) Tabular
-form comparing and contrasting - Relationship to geography
-
-\begin{itemize}
-\tightlist
-\item
-  Show using cancer examples
-\end{itemize}
-
-\hypertarget{animation}{%
-\section{Animation}\label{animation}}
-
-lends to temporal pattern exploration
-
-Keim et al. (\protect\hyperlink{ref-ECGC}{2002}) ?? Highlight the value
-of animating contiguous to see changes over time, US can be recognisable
-but animation aides interpretation
+The intention of interactivity and animation is to allow users to answer
+their own questions.
 
@@
 
-What software did we use, for e.g.
-
 \hypertarget{references}{%
@@
 
+Wickham (\protect\hyperlink{ref-tidyerse}{2017}) for data analysis,
+Bivand, Nowosad, and Lovelace (\protect\hyperlink{ref-spData}{2019}) and
+Pebesma (\protect\hyperlink{ref-sf}{2018}) to implement plotting of
+spatial data. Kobakian and Cook (\protect\hyperlink{ref-sugarbag}{2019})
+to create hexagon tesselation. Arnold
+(\protect\hyperlink{ref-ggthemes}{2019}) to enhance plot displays.
+Jeworutzki (\protect\hyperlink{ref-cartogram}{2018}) for contiguous and
+non-contiguous cartogram displays.
+
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-ggthemes}{}%
+Arnold, Jeffrey B. 2019. \emph{Ggthemes: Extra Themes, Scales and Geoms
+for 'Ggplot2'}. \url{https://CRAN.R-project.org/package=ggthemes}.
+
 \leavevmode\hypertarget{ref-CPISACA}{}%
@@
 Berry, Brian J. L., Richard L. Morrill, and Waldo R. Tobler. n.d.
-``GEOGRAPHIC Ordering of Information: NEW Opportunities'' 16 (4).
-Association of American Geographers: 39--44.
-\url{https://doi.org/10.1111/j.0033-0124.1964.039_q.x}.
+``GEOGRAPHIC Ordering of Information: NEW Opportunities'' 16 (4):
+39--44. \url{https://doi.org/10.1111/j.0033-0124.1964.039_q.x}.
+
+\leavevmode\hypertarget{ref-spData}{}%
+Bivand, Roger, Jakub Nowosad, and Robin Lovelace. 2019. \emph{SpData:
+Datasets for Spatial Analysis}.
+\url{https://CRAN.R-project.org/package=spData}.
 
@@
 
+\leavevmode\hypertarget{ref-NISCC}{}%
+Dent, Borden D. 1972. ``A Note on the Importance of Shape in Cartogram
+Communication.'' \emph{Journal of Geography} 71 (7): 393--401.
+\url{https://doi.org/10.1080/00221347208981697}.
+
+\leavevmode\hypertarget{ref-MACM}{}%
+d'Onofrio, Alberto, Chiara Mazzetta, Chris Robertson, Michel Smans,
+Peter Boyle, and Mathieu Boniol. 2016. ``Maps and Atlases of Cancer
+Mortality: A Review of a Useful Tool to Trigger New Questions.'' Journal
+Article. \emph{Ecancermedicalscience} 10: 670--70.
+\url{https://doi.org/10.3332/ecancer.2016.670}.
+
 \leavevmode\hypertarget{ref-ACTUC}{}%
@@
 
+\leavevmode\hypertarget{ref-ACCAC}{}%
+Dougenik, James A., Nicholas R. Chrisman, and Duane R. Niemeyer. 1985.
+``AN Algorithm to Construct Continuous Area Cartograms.'' \emph{The
+Professional Geographer} 37 (1): 75--81.
+\url{https://doi.org/10.1111/j.0033-0124.1985.00075.x}.
+
 \leavevmode\hypertarget{ref-SE}{}%
@@
 Griffin, T.L.C. 1980. ``Cartographic Transformation of the Thematic Map
-Base.'' \emph{Cartography} 11 (3). Taylor \& Francis: 163--74.
+Base.'' \emph{Cartography} 11 (3): 163--74.
 \url{https://doi.org/10.1080/00690805.1980.10438102}.
@@
 
+\leavevmode\hypertarget{ref-cartogram}{}%
+Jeworutzki, Sebastian. 2018. \emph{Cartogram: Create Cartograms with R}.
+\url{https://CRAN.R-project.org/package=cartogram}.
+
 \leavevmode\hypertarget{ref-ECGC}{}%
@@
 
+\leavevmode\hypertarget{ref-sugarbag}{}%
+Kobakian, Stephanie, and Dianne Cook. 2019. \emph{Sugarbag: Create
+Tessellated Hexagon Maps}.
+\url{https://CRAN.R-project.org/package=sugarbag}.
+
 \leavevmode\hypertarget{ref-CBATCC}{}%
@@
 Approach to Constructing Continuous Cartograms.'' In.
+
+\leavevmode\hypertarget{ref-CD}{}%
+Kraak, M. J. 2017. ``Cartographic Design.'' In \emph{The International
+Encyclopedia of Geography : People, the Earth, Environment, and
+Technology}, edited by D. Richardson, 1--16. United States: Wiley.
 
@@
 Olson, Judy M. 1976. ``NONCONTIGUOUS Area Cartograms.'' \emph{The
-Professional Geographer} 28 (4). Routledge: 371--80.
+Professional Geographer} 28 (4): 371--80.
 \url{https://doi.org/10.1111/j.0033-0124.1976.00371.x}.
 
+\leavevmode\hypertarget{ref-sf}{}%
+Pebesma, Edzer. 2018. ``Simple Features for R: Standardized Support for
+Spatial Vector Data.'' \emph{The R Journal} 10 (1): 439--46.
+\url{https://doi.org/10.32614/RJ-2018-009}.
+
 \leavevmode\hypertarget{ref-GAMP}{}%
-Tobler, Waldo. 1963. ``Geographic Area and Map Projections,'' January.
-American Geographical Society, 59--78.
-\url{https://www.jstor.org/stable/212809}.
+Tobler, Waldo. 1963. ``Geographic Area and Map Projections,'' January,
+59--78. \url{https://www.jstor.org/stable/212809}.
 
@@
 ---------. 2004. ``Thirty Five Years of Computer Cartograms.''
-\emph{Annals of the Association of American Geographers} 94 (1). Taylor
-\& Francis Group: 58--73.
+\emph{Annals of the Association of American Geographers} 94 (1): 58--73.
 
@@
 
+\leavevmode\hypertarget{ref-tidyerse}{}%
+Wickham, Hadley. 2017. \emph{Tidyverse: Easily Install and Load the
+'Tidyverse'}. \url{https://CRAN.R-project.org/package=tidyverse}.
+
 \end{document}

### 2019-08-13 review_article.Rmd
save choropleth grid

@@
     toc: yes
-    toc_depth: 2
-  html_document:
-    df_print: paged
-    toc: yes
-    toc_depth: '2'
+    toc_depth: 3
 fontsize: 11pt
@@
 knitr::opts_chunk$set(echo = FALSE)
+library(ggplot2)
 ```
@@
 <!-- Displaying of cancer data on choropleths -->
-
-<!-- Displaying of cancer data on choropleths -->
 Choropleth maps can be useful devices for communicating information to public on a familiar map base.
+A cancer atlas is a choropleth map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries. 
+Choropleths consider the spatial structures and present information regarding the individuals, their communities and organization into larger units @SAMGIS.
 In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality. 
-A cancer atlas is a choropleth map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
-@MACM provide the definition and reports that cancer atlases began with Haviland's maps in 1875.
 The data collection methods of cancer mortality rates across regions, and the administrative control within regions lends itself to choropleth visualisation. 
+@MACM provides the definition of a cancer atlas, beginning with Haviland's maps in 1875, they attribute UK cancer atlases to Howe (1963), and early work in US cancer atlases can be attributed to Burbank (1971).
 The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
-Early work in US cancer atlases can be attributed to Burbank ???, and UK cancer atlases to Howe 1963 ???.
 The choropleth maps presented levels via hatchings or dots on a black and white scale.
@@
 
-<!--
-A particular spatial structure includes
-the individuals affected and how they are connected in
-communities, as well as the dynamics of these communities
-and their organization into larger units. @SAMGIS
--->
-
-<!-- Measures reported in cancer atlases -->
-
 <!-- Measures reported in cancer atlases -->
 Almost 100 years of cancer mapping in the United States and the United Kingdom has seen increased effectiveness in the presentation cancer statistics.
-Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the the higher prevalence of cancers in older populations.
+Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the higher prevalence of cancers in older populations.
 Howe [-@HEDP] describes Stock's development of the standardised mortality ratios through the 1930s.
@@
 
-```{r chloropleth-grid, echo=FALSE, message=FALSE}
-
+```{r chloropleth-grid-create, echo=FALSE, message=FALSE, eval = FALSE}
 library(grid)
@@
 
-gridExtra::grid.arrange(grid::rasterGrob(plot1),grid::rasterGrob(plot2), grid::rasterGrob(plot3), 
+chloropleth_grid <- gridExtra::grid.arrange(grid::rasterGrob(plot1),grid::rasterGrob(plot2), grid::rasterGrob(plot3), 
                         grid::rasterGrob(plot4), grid::rasterGrob(plot5), grid::rasterGrob(plot6),
@@
 
+ggsave(choropleth_grid, filename = "figures/choropleth_grid.png", device = "png", width = 12, height = 6)
+```
+
+```{r chloropleth-grid-print, echo = FALSE}
+knitr::include_graphics("figures/choropleth_grid.png")
 ```
@@
 
-```{r ci-viz2, fig.cap = "Example of CI visualisation for uncertainty representation in cancer mapping (1/3).
-Source: Alberta Health IHDA Geographic.
-(2012) ", echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width='50%', eval = FALSE}
-
-knitr::include_graphics("figures/Alberta_CI_viz.png")
-```
-
-```{r ci-viz3, fig.cap = "Example of CI visualisation for uncertainty representation in cancer mapping (2/3).
-Source: Pensylvania Cancer Atlas ", echo=FALSE, message=FALSE, warning=FALSE, out.width='50%', fig.align='center', eval = FALSE}
-
-knitr::include_graphics("figures/Pennsylvania_CI_viz.png")
-
-```
-
-```{r ci-tooltip, fig.cap= "Example of CI visualisation for uncertainty representation in cancer mapping (3/3).
-Source: Centres for Disease Control and Prevention (CDC).
-United States Cancer Statistics: An Interactive Cancer Atlas (InCA) ", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '50%', fig.align = 'center', eval = FALSE}
-
-knitr::include_graphics("figures/CDC_CI_viz_tooltip.png")
-```
-
-```{r ci-table, fig.cap= "Example of an interactive data table with CI upper and lower bounds used in cancer mapping.
-Source: Pennsylvania Cancer Atlas", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '50%', fig.align = 'center', eval = FALSE}
-
-knitr::include_graphics("figures/uncert_table.png")
-```
-
 Methods for representing sources of uncertainty information can be visualised or communicated in different ways, examples identified through this grey literature review are listed below.
-
-```{r stndvn-example, fig.cap= "Example of standard deviation visualised in cancer mapping", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '90%', fig.align = 'center'}
-
-knitr::include_graphics("figures/standard_deviation_large.png")
-```
-
-```{r boxplot-example, fig.cap= "Example of boxplot used in cancer mapping.
-Source: Source: Pensylvania Cancer Atlas ", echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width= '40%', fig.align = 'center', eval = FALSE}
-
-knitr::include_graphics("figures/Boxplot2.png")
-```
 

### 2019-08-13 review_article.Rmd
Referencing tables

@@
 Howe [-@HEDP] describes Stock's development of the standardised mortality ratios through the 1930s.
-Table 1 presents summarises the measures presented in published cancer atlases, and provides a definition of each measure.
+Table \ref{measures} presents summarises the measures presented in published cancer atlases, and provides a definition of each measure.
 
-<!-- ??? Table tags-->
-Table 1: Measures used to report cancer statistics
+Table: \label{measures} Measures used to report cancer statistics
 
@@
 Close to half of the atlases identified (42%, n=14) included some measure of uncertainty. The most common measure used to represent uncertainty were credible or confidence intervals (CIs).
+Methods for representing sources of uncertainty information can be visualised or communicated in different ways, examples identified through this grey literature review are listed below.
 
@@
 -->
-
-Methods for representing sources of uncertainty information can be visualised or communicated in different ways, examples identified through this grey literature review are listed below.
 

### 2019-08-14 usa_plots.R
spell check on choropleth

@@ -148,3 +148,40 @@ ggchoroCRS <- gridExtra::grid.arrange(ggchoro1, ggchoro2, ggchoro3, ggchoro4)
 
 ggsave(filename = "figures/ggchoroCRS.png", plot = ggchoroCRS,
   device = "png", width = 12, height = 6)
+
+
+#####
+# Choropleth grid
+
+library(grid)
+library(png)
+library(gridExtra)
+
+plot1 <- png::readPNG('cancer_map_images/chloropleths/1.png')
+plot2 <- png::readPNG('cancer_map_images/chloropleths/14.chloropleth1.png')
+plot3 <- png::readPNG('cancer_map_images/chloropleths/15.chloropleth.png')
+plot4 <- png::readPNG('cancer_map_images/chloropleths/3.chloropleth.png')
+plot5 <- png::readPNG('cancer_map_images/chloropleths/5.chloropleth.png')
+plot6 <- png::readPNG('cancer_map_images/chloropleths/6.chloropleth.png')
+plot6 <- png::readPNG('cancer_map_images/chloropleths/7.chloropleth.png')
+plot7 <- png::readPNG('cancer_map_images/chloropleths/8.chloropleth.png')
+plot8 <- png::readPNG('cancer_map_images/chloropleths/20.chloropleth1.png')
+plot9 <- png::readPNG('cancer_map_images/chloropleths/21.chloropleth2.png')
+plot10 <- png::readPNG('cancer_map_images/chloropleths/23.chloropleth.png')
+plot11 <- png::readPNG('cancer_map_images/chloropleths/24.chloropleth.png')
+plot12 <- png::readPNG('cancer_map_images/chloropleths/25.chloropleth.png')
+plot13 <- png::readPNG('cancer_map_images/chloropleths/26.chloropleth.png')
+plot14 <- png::readPNG('cancer_map_images/chloropleths/27.chloropleth.png')
+plot15 <- png::readPNG('cancer_map_images/chloropleths/29.chloropleth.png')
+plot16 <- png::readPNG('cancer_map_images/chloropleths/30.chloropleth.png')
+
+
+chloropleth_grid <- gridExtra::grid.arrange(grid::rasterGrob(plot1),grid::rasterGrob(plot2), grid::rasterGrob(plot3), 
+  grid::rasterGrob(plot4), grid::rasterGrob(plot5), grid::rasterGrob(plot6),
[…]

### 2019-08-15 review_article.Rmd
correct spelling of choropleth in images

@@
 
-```{r chloropleth-grid-create, echo=FALSE, message=FALSE, eval = FALSE}
+```{r choropleth-grid-create, echo=FALSE, message=FALSE, eval = FALSE}
 library(grid)
@@
 
-plot1 <- png::readPNG('cancer_map_images/chloropleths/1.png')
-plot2 <- png::readPNG('cancer_map_images/chloropleths/14.chloropleth1.png')
-plot3 <- png::readPNG('cancer_map_images/chloropleths/15.chloropleth.png')
-plot4 <- png::readPNG('cancer_map_images/chloropleths/3.chloropleth.png')
-plot5 <- png::readPNG('cancer_map_images/chloropleths/5.chloropleth.png')
-plot6 <- png::readPNG('cancer_map_images/chloropleths/6.chloropleth.png')
-plot6 <- png::readPNG('cancer_map_images/chloropleths/7.chloropleth.png')
-plot7 <- png::readPNG('cancer_map_images/chloropleths/8.chloropleth.png')
-plot8 <- png::readPNG('cancer_map_images/chloropleths/20.chloropleth1.png')
-plot9 <- png::readPNG('cancer_map_images/chloropleths/21.chloropleth2.png')
-plot10 <- png::readPNG('cancer_map_images/chloropleths/23.chloropleth.png')
-plot11 <- png::readPNG('cancer_map_images/chloropleths/24.chloropleth.png')
-plot12 <- png::readPNG('cancer_map_images/chloropleths/25.chloropleth.png')
-plot13 <- png::readPNG('cancer_map_images/chloropleths/26.chloropleth.png')
-plot14 <- png::readPNG('cancer_map_images/chloropleths/27.chloropleth.png')
-plot15 <- png::readPNG('cancer_map_images/chloropleths/29.chloropleth.png')
-plot16 <- png::readPNG('cancer_map_images/chloropleths/30.chloropleth.png')
+plot1 <- png::readPNG('cancer_map_images/choropleths/1.png')
+plot2 <- png::readPNG('cancer_map_images/choropleths/14.choropleth1.png')
+plot3 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
+plot4 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
+plot5 <- png::readPNG('cancer_map_images/choropleths/5.choropleth.png')
+plot6 <- png::readPNG('cancer_map_images/choropleths/6.choropleth.png')
+plot6 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
+plot7 <- png::readPNG('cancer_map_images/choropleths/8.choropleth.png')
+plot8 <- png::readPNG('cancer_map_images/choropleths/20.choropleth1.png')
+plot9 <- png::readPNG('cancer_map_images/choropleths/21.choropleth2.png')
+plot10 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
+plot11 <- png::readPNG('cancer_map_images/choropleths/24.choropleth.png')
+plot12 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
+plot13 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
+plot14 <- png::readPNG('cancer_map_images/choropleths/27.choropleth.png')
+plot15 <- png::readPNG('cancer_map_images/choropleths/29.choropleth.png')
+plot16 <- png::readPNG('cancer_map_images/choropleths/30.choropleth.png')
 
-chloropleth_grid <- gridExtra::grid.arrange(grid::rasterGrob(plot1),grid::rasterGrob(plot2), grid::rasterGrob(plot3), 
+choropleth_grid <- gridExtra::grid.arrange(grid::rasterGrob(plot1),grid::rasterGrob(plot2), grid::rasterGrob(plot3), 
                         grid::rasterGrob(plot4), grid::rasterGrob(plot5), grid::rasterGrob(plot6),
@@
 
-```{r chloropleth-grid-print, echo = FALSE}
+```{r choropleth-grid-print, echo = FALSE}
 knitr::include_graphics("figures/choropleth_grid.png")
+
 ```

### 2019-08-15 review_article.Rmd
change people to populations in the intro.
I'm not 100% sure of this change. discuss with Steph. But the statistics on cancer maps are population statistics rather than individual statistics. it's one of the challenges of cancer mapping - how do we make population statistics relevant to an individual who is viewing the map.

@@
 However, it is not enough for areas on maps to be recognisable, if information regarding the distribution of the statistic cannot be gleaned.
-In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer statistics.
+In situations where the population and their health is of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer statistics.
 Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS]. 
 This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes, and health outcomes more broadly.
-Cancer statistics directly represent and relate to the people living within individual geographical areas.
+Cancer statistics directly represent and relate to the populations living within individual geographical areas.
 

### 2019-08-15 review_article.Rmd
#current best practices - a few small edits
nothing major. There is one doubled up sentence. wasn't sure which one to delete.

@@
 Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
+Now many examples of choropleth maps can be found on the websites of government, media, health and other public facing organisations (insert ref for Jessie's grey lit review).
 A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form. 
@@
 Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
-While the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
-The different population and geographical sizes of administrative areas can attract attention to the shades of the unpopulated but large areas [@EI]. 
+__(repeated sentence)__
+While the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of both non-uniform size, and population densities.
+In addition these boundaries are often subject to change over time.
+The different population and geographical sizes of administrative areas can unduly attract focus to the shades of the unpopulated but large areas [@EI]. 
 Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].
@@
 <!-- Measures reported in cancer atlases -->
-Almost 100 years of cancer mapping in the United States and the United Kingdom has seen increased effectiveness in the presentation cancer statistics.
+Almost 100 years of cancer mapping in the United States and the United Kingdom has seen increased effectiveness in the presentation of cancer statistics.
 Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the higher prevalence of cancers in older populations.
 Howe [-@HEDP] describes Stock's development of the standardised mortality ratios through the 1930s.
-Table \ref{measures} presents summarises the measures presented in published cancer atlases, and provides a definition of each measure.
+Table \ref{measures} summarises the measures presented in published cancer atlases, and provides a definition of each measure.
 

### 2019-08-16 review_article.Rmd
Editing

--- 2019-08-15 review_article.Rmd
@@
 
-Maps have been adopted to present geospatial statistics for centuries.
-Maps connect data and statistics to the geographical representation of areas that are generally familiar to the audience.
-However, it is not enough for areas on maps to be recognisable, if information regarding the distribution of the statistic cannot be gleaned.
-In situations where the people are of interest, not the land they live on, it is reasonable to explore views that enhance the communication of the cancer statistics.
-Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS]. 
-This has spurred innovations over the previous centuries to enhance the shapes and maps presented to effectively communicate cancer outcomes, and health outcomes more broadly.
+Geospatial statistics have been presented and stored on maps for centuries.
+Maps connect data for areas to the geographical representation which is usually familiar to the audience.
+However, it is not enough for areas on maps to be recognisable, the distribution of the statistic must also be visible.
 Cancer statistics directly represent and relate to the people living within individual geographical areas.
+It is reasonable to explore views that enhance the communication of the cancer statistics, especially as in these situations the people are of interest, not the land they live on. 
+This has spurred innovations over the previous centuries to enhance maps to effectively communicate cancer outcomes, and health outcomes more broadly.
 
 The visualisation methods used to present cancer statistics will depend on the intended message and users of the map. 
-Transforming statistics to visualisations considers individual observations or observation points aggregated into communities and geographical units.
+Presenting statistics as visualisations requires transforming individual observations into aggregations of communities as geographical units.
 Visualising diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH]. 
@@
 <!-- what is a choropleth map -->
-A choropleth map is used to display the spatial characteritics of a relationship, or variability, of measurements. 
-As an alternative storage device to a table, it preserves locations for geographically ordered data, with usage dating back to the 1800s [@GOINO].
+A choropleth map is used to display the characteristics of a spatial relationship of measurements. 
+A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns on a familiar geographic form. 
+Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
+Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS]. 
 A choropleth is constructed by drawing the geographic or political boundaries, and filling the shapes with colours to represent values of a measured variable [@EI].
 Early versions of choropleth maps used symbols or patterns instead of colour.
-@CPISACA discusses their use in visualising cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
-Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
-A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form. 
-Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
+As an alternative storage device to a table, it preserves locations for geographically ordered data, with usage dating back to the 1800s [@GOINO].
+@CPISACA discuss the use of choropleths to visualise cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
 
-```{r choro, fig.cap = "A choropleth map of the United States of America.", echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+```{r choro, fig.cap = "A choropleth map of the United States of America.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
 knitr::include_graphics("figures/ggchoro.png")
@@
 A cancer atlas is a choropleth map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries. 
-Choropleths consider the spatial structures and present information regarding the individuals, their communities and organization into larger units @SAMGIS.
 In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality. 
@@
 
-```{r choropleth-grid-create, echo=FALSE, message=FALSE, eval = FALSE}
+```{r choropleth-grid-create, message=FALSE, eval = FALSE}
 library(grid)
@@
 
-ggsave(choropleth_grid, filename = "figures/choropleth_grid.png", device = "png", width = 12, height = 6)
+ggsave(filename = "figures/choropleth_grid.png", plot = choropleth_grid, device = "png", width = 12, height = 6, dpi = 300)
 ```
 
-```{r choropleth-grid-print, echo = FALSE}
+```{r choropleth-grid-print, echo = FALSE, results = "asis"}
 knitr::include_graphics("figures/choropleth_grid.png")
-
 ```
@@
 Cruickshank's (1947) as cited by @DMAHP, discusses using visuals as 'formal statistical assessment of the spatial pattern'.
-However, when presenting cancer atlases, @MACM believes the intuition derived must be 'validated by rigorous statistical analyses.'
+However, when presenting cancer atlases, @MACM believes the intuition derived must be 'validated by rigorous statistical analyses.' @CPISACA suggests 'good tables, graphs, and explanatory text' support understanding and inference derived from maps will be consistent across a range of viewers.
+The interactivity of the more modern maps enabled supplementary information to be incorporated without cluttering the screen, such as in a tool tip feature.
 
-<!--
-'Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps' @CPISACA
+<!-- geographic hierarchies  -->
 
--->
-
-<!-- The supplementary statistics ??? -->
+<!-- statistical uncertainty -->
 These additional statistics often include a measure of the statistical uncertainty of the values of the statistics presented in a choropleth.
@@
 The maps considered used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
-These maps ranged from static pdfs or infographics, to interactive online resources.
+The maps employing uncertainty ranged from static pdfs or infographics, to interactive online resources.
 
-<!-- Presentation methods of supplementary statistics ??? -->
-The interactivity of the more modern maps enabled supplementary information to be incorporated without cluttering the screen, such as in a tool tip feature.
+<!-- demographics -->
+
+<!-- sex, race,  -->
 
@@
 Cancer maps are effective communication tools for a general or non-expert audience. They are commonly used in the public domain to communicate results of sophisticated statistical analyses.
-The heavy use of chloropleth maps within the research literature is reflected in the types of maps that are found in the public domain.
+The heavy use of choropleth maps within the research literature is reflected in the types of maps that are found in the public domain.
 A grey literature review conducted by (ref) identified 33 cancer atlases published on the internet between January 2010 and November 2015.
-These chloropleth maps were mostly published by non-commercial organisations, including not-for-profits (NFPs), government, research organisations, advocacy groups or a partnership between an NFP & government.
+These choropleth maps were mostly published by non-commercial organisations, including not-for-profits (NFPs), government, research organisations, advocacy groups or a partnership between an NFP & government.
 Only one map was published by a commercial entity.
@@
 For a single variable of interest, each map area is changed to emphasise the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].
+The changes in the map base occur by altering individual areas, by altering the shapes or boundaries.
 
@@
 Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
-To choose an approporiate map display, the map creator must consider the intended user, and message the map will communicate.
+To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
 It is the objectives of the investigator that will drive the choice of representation [@CPISACA].
@@
 @tidyerse for data analysis, @spData and @sf to implement plotting of spatial data.
-@sugarbag to create hexagon tesselation. @ggthemes to enhance plot displays.
+@sugarbag to create hexagon tessellation. @ggthemes to enhance plot displays.
 @cartogram for contiguous and non-contiguous cartogram displays.

### 2019-08-16 review_article.Rmd
size of figure

@@
 
-<!--
-They are untrustworthy when interpretations of spatial association are needed; there are both severe physiological and psychological constraints on our abilities to assess correlations of spatial distributions visually.
--->
-
 <!-- Displaying of cancer data on choropleths -->
@@
 The choropleth maps presented levels via hatchings or dots on a black and white scale.
-These atlases were key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets @MACM.
+These atlases were key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets [@MACM].
 
@@
 Howe [-@HEDP] describes Stock's development of the standardised mortality ratios through the 1930s.
-Table \ref{measures} presents summarises the measures presented in published cancer atlases, and provides a definition of each measure.
+Table \ref{tab:measures} presents summarises the measures presented in published cancer atlases, and provides a definition of each measure.
 
-Table: \label{measures} Measures used to report cancer statistics
+Table: \label{tab:measures} Measures used to report cancer statistics
 
@@
 
-ggsave(filename = "figures/choropleth_grid.png", plot = choropleth_grid, device = "png", width = 12, height = 6, dpi = 300)
+ggsave(filename = "figures/choropleth_grid.png", plot = choropleth_grid, device = "png", width = 6, height = 5, dpi = 300)
 ```

### 2019-08-19 review_article.Rmd
Supporting Statistics section

@@
 
-A map communicates quickly and draws attention to prominent geographies, but an atlas is often supplemented with supporting statistics.
-Cruickshank's (1947) as cited by @DMAHP, discusses using visuals as 'formal statistical assessment of the spatial pattern'.
-However, when presenting cancer atlases, @MACM believes the intuition derived must be 'validated by rigorous statistical analyses.' @CPISACA suggests 'good tables, graphs, and explanatory text' support understanding and inference derived from maps will be consistent across a range of viewers.
+The presentation of indicators in atlases can also be in found in bars, tables and scatterplots.
+This is a useful alternative display, it is not the only additional display as an atlas is often supplemented with supporting statistics.
+When presenting cancer atlases, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses.' This opinion contradicts Cruickshank's (1947) as cited by @DMAHP, which discusses using visuals as 'formal statistical assessment of the spatial pattern'.
+@CPISACA suggests additional materials such as 'good tables, graphs, and explanatory text' support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers.
 The interactivity of the more modern maps enabled supplementary information to be incorporated without cluttering the screen, such as in a tool tip feature.
@@
 <!-- geographic hierarchies  -->
+While atlases are often used to describe differences between areas, the may allow statistics to be displayed at different levels of aggregation.
+World atlases can allow for displays of data aggregated into continents, country, states or provinces [@IARC_3].
+
+<!-- population distribution -->
+It is likely that each population area will have a different amount of people, information regarding the distribution of population levels may be provided in a table or histogram display [@NICR_1].
+Atlases could also connect the population to the land available to them by communicating population density.
+
+<!-- demographics: sex, race,  -->
 
@@
 These additional statistics often include a measure of the statistical uncertainty of the values of the statistics presented in a choropleth.
-In the review of atlases in the public domain, atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map.
-Maps that only reported this information within the supplementary material were not considered to have directly attempted to report uncertainty.
-The maps considered used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
-The maps employing uncertainty ranged from static pdfs or infographics, to interactive online resources.
 
-<!-- demographics -->
-
-<!-- sex, race,  -->
+<!-- Socioeconomic Indicators -->
+There may also be demographic information provided regarding each population area drawn on the maps.
+This information could include a socioeconomic indicators such as unemployment, education level attained.
+The areas may be ranked and the statistics could be presented as categories describing the ranking.
 
@@
 
+The atlases are not only maps.
+Many provide additional information to help users understand and interpret the spatial distribution.
+Table \ref{tab:supporting-stats} explores the use of these supports in a variety of online map sources.
+Demographics include information regarding the age and sex distribution of the areas presented.
+Socioeconomic indicators include Unemployment rates of regions, Remoteness and Education levels achieved.
+Statistical uncertainty is communicated through a Confidence (CI) interval, statistical significance levels, Boxplots, Distribution plots, and reporting the Sample Size and Standard Deviations.
+These additional pieces of information can help to validate, explain or explore relationships between cancer incidence and survival rates.
+
+Table: \label{tab:supporting-stats} Supporting statistics in public facing atlases
+| Atlas                    | Geographic hierarchies | Population Distributions | Demographics | Statistical Uncertainty | Socioeconomic Indicators |
+|--------------------------|------------------------|--------------------------|--------------|-------------------------|--------------------------|
+| All Ireland Cancer Atlas |                        |                          |              |                         |                          |
+|                          |                        |                          |              |                         |                          |
+|                          |                        |                          |              |                         |                          |
+
+<!-- geographic hierarchies  -->
+While atlases are often used to describe differences between areas, the may allow statistics to be displayed at different levels of aggregation.
+World atlases can allow for displays of data aggregated into continents, country, states or provinces [@IARC_3].
+
+<!-- population distribution -->
+It is likely that each population area will have a different amount of people, information regarding the distribution of population levels may be provided in a table or histogram display [@NICR_1].
+Atlases could also connect the population to the land available to them by communicating population density.
+
+<!-- statistical uncertainty -->
+These additional statistics often include a measure of the statistical uncertainty of the values of the statistics presented in a choropleth.
+In the review of atlases in the public domain, atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map.
+The maps considered used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
+The maps employing uncertainty ranged from static pdfs or infographics, to interactive online resources.
 Close to half of the atlases identified (42%, n=14) included some measure of uncertainty. The most common measure used to represent uncertainty were credible or confidence intervals (CIs).
-Methods for representing sources of uncertainty information can be visualised or communicated in different ways, examples identified through this grey literature review are listed below.
 
-<!--
-Create a table:
-cofactors: Greater capital city regions, states, socio demographic status, Remoteness
+<!-- Demographics -->
+Age groups,
 
-Supplementary stats
-stat        | Literature examples | Public atlas examples
-uncertainty |      ...            |    ...
-co factor(1)|      ...            |    ...
-co factor(2)|      ...            |    ...
--->
-
-Table: (\#tab:method-exp) Implicit and explicit measures of uncertainty.
-
-|Measure | Example | 
-|:-------------------------|:-------------------------------------------|
-|CI Interval |Figures 3.1, 3.2, 3.3|
-|Statistical Significance|Textured overlay on top of coloured regions used to indicate statistical significance|
-|Distribution|Figure 3.4|
-|Boxplots|Figures 3.4, 3.5|
-|Sample Size|Textured overlay or lack of colour on a region, was used to show regions with small sample size|
-|Standard deviation |Figure 3.6 - the second map in the bottom right corner shows standard deviation|
+<!-- Socioeconomic Indicators -->
 

### 2019-08-19 review_article.Rmd
captions for USA figures

@@
 
-```{r choro, fig.cap = "A choropleth map of the United States of America.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+```{r choro, fig.cap = "A choropleth map of the United States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+The state of Utah has an unusually low rate of lung and bronchus incidence.
+The state of Kentucky is high, the surrounding states also have high rates.
+There is a pattern of average higher rates toward the East Coast and lower rates toward the West Coast.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
 knitr::include_graphics("figures/ggchoro.png")
@@
 
+<!--
+
+'In Australia the urban federal constituencies occupy only a tenth of the land, but contain nine tenths of the people.
+It would be almost unthinkable to show election results for that country on a conventional equal land area map.'
+This 1966 cartogram uses mostly straight lines, and the result looks very little like the geographical shape of Australia.
+
+'Given the increasingly uneven population distribution of the United States and the growing social divides between the populations of neighbourhoods living at different densities, the need for cartograms like this is greater now than ever.'
+
+-->
+
 <!-- There are many alternative views of the world-->
@@
 
-!["Choropleth maps of the United States of America using four coordinate reference systems."](figures/ggchoroCRS.png)
+```{r choroCRS, fig.cap = "Four choropleth maps of the United States of America using various coordinate reference systems.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+The map projections alter the shapes and angles of the boundaries of each state.
+Maps a and b are similar in their straight edges, unlike maps c and d which curve on the northern United States border.
+", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+knitr::include_graphics("figures/ggchoroCRS.png")
+```
 
@@
 
-!["A contiguous cartogram of the Unites States of America"](figures/ggcont.png)
+```{r ggcont, fig.cap = "A contiguous cartogram map of the United States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each stated shape has been distorted according to the population of the state in 2015.
+The state of California has become much larger due to it's large population density.
+This draws attention to the densely populated North East region, and detracts from the Mid West.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+knitr::include_graphics("figures/ggcont.png")
+```
 
@@
 
-!["A Non - contiguous cartogram of the Unites States of America"](figures/ggncont.png)
+```{r ggncont, fig.cap = "A Non - contiguous cartogram map of the United States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state shape has been maintained, but the size has altered according to the population of the state in 2015.
+The state of California has remained closer to it's original size than it's surrounding states.
+The North East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut.
+This draws attention to the densely populated North East region, and the sparse Mid West.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+knitr::include_graphics("figures/ggncont.png")
+```
 
-!["A dorling cartogram of the Unites States of America"](figures/ggdorl.png)
+```{r ggdorl, fig.cap = "A dorling cartogram of the Unites States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been represented by a circle, but the size was determined by the population of the state in 2015.
+The North East states remain closer to their neighbours, and may be displaced from their geographic location.
+The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+knitr::include_graphics("figures/ggdorl.png")
+```
 
@@
 While contiguous cartograms may be a 'more sophisticated' method, they produce 'very complex shapes'.
-Circular cartograms use the same shape for every region represented, and size them according to the statistic represented or the population for a base map.
+Circular cartograms use the same simple shapes for every region represented, and size them according to the statistic represented or the population for a base map.
+This familar shape may be more effective that contiguous cartograms, as the shapes used have 'no meaning' after distortions are applied @NISCC.
 To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighbouring areas over many iterations.
@@
 <!--
--simple shapes of the circle, square, and triangle are relatively easily recognized.
 -sacrifice communication efficiency;
@@
 @NISCC
--->
 
-'In Australia the urban federal constituencies occupy only a tenth of the land, but contain nine tenths of the people.
-It would be almost unthinkable to show election results for that country on a conventional equal land area map.'
-This 1966 cartogram uses mostly straight lines, and the result looks very little like the geographical shape of Australia.
-
-'Given the increasingly uneven population distribution of the United States and the growing social divides between the populations of neighbourhoods living at different densities, the need for cartograms like this is greater now than ever.'
-
-Used in displays of the UK by Howe in 1986 cited by @HEDP
-<!--
-# Temporal displays
 -->

### 2019-08-19 usa_plots.R

@@ -123,25 +123,25 @@ ggdot
 ggchoro1 <- ggplot(st_transform(cancer, 3857)) + 
   geom_sf(aes(fill = AgeAdjustedRate)) +
   scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
-  ggtitle("The United States using EPSG: 3857") +
+  ggtitle("a. The United States using EPSG: 3857") +
   theme_void()+ guides(fill = FALSE)
 
 ggchoro2 <- ggplot(st_transform(cancer, 2163)) + 
   geom_sf(aes(fill = AgeAdjustedRate)) +
   scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
-  ggtitle("The United States using EPSG: 2163") +
+  ggtitle("b. The United States using EPSG: 2163") +
   theme_void()+ guides(fill = FALSE)
 
 ggchoro3 <- ggplot(st_transform(cancer, 4326)) + 
   geom_sf(aes(fill = AgeAdjustedRate)) +
   scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
-  ggtitle("The United States using EPSG: 4326") +
+  ggtitle("c. The United States using EPSG: 4326") +
   theme_void()+ guides(fill = FALSE)
 
 ggchoro4 <- ggplot(st_transform(cancer, 2955)) + 
   geom_sf(aes(fill = AgeAdjustedRate)) +
   scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
-  ggtitle("The United States using EPSG: 2955") +
+  ggtitle("d. The United States using EPSG: 2955") +
   theme_void() + guides(fill = FALSE)
 
 ggchoroCRS <- gridExtra::grid.arrange(ggchoro1, ggchoro2, ggchoro3, ggchoro4)

### 2019-08-19 review_article.Rmd
edits to supporting statistics

@@
 
-The presentation of indicators in atlases can also be in found in bars, tables and scatterplots.
-This is a useful alternative display, it is not the only additional display as an atlas is often supplemented with supporting statistics.
+The primary display of an atlas is often the choropleth map.
+The presentation of materials in atlases can also utilise bar charts, tables and scatterplots, these are useful alternative displays.
+They are not only used as an additional display for the cancer incidence and mortality rates, as an atlas is often supplemented with supporting statistics.
 When presenting cancer atlases, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses.' This opinion contradicts Cruickshank's (1947) as cited by @DMAHP, which discusses using visuals as 'formal statistical assessment of the spatial pattern'.
@@
 <!-- geographic hierarchies  -->
-While atlases are often used to describe differences between areas, the may allow statistics to be displayed at different levels of aggregation.
+Cancer atlases are often used to describe differences between areas, they may allow statistics to be displayed at different levels of aggregation.
 World atlases can allow for displays of data aggregated into continents, country, states or provinces [@IARC_3].
@@
 It is likely that each population area will have a different amount of people, information regarding the distribution of population levels may be provided in a table or histogram display [@NICR_1].
-Atlases could also connect the population to the land available to them by communicating population density.
+Atlases could also connect the population to the land available to them by communicating population density as is shown in the United States map examples.
 
@@
 <!-- statistical uncertainty -->
-These additional statistics often include a measure of the statistical uncertainty of the values of the statistics presented in a choropleth.
+The additional statistics often also include a measure of the statistical uncertainty of the cancer statistics presented in a choropleths. <!-- reference Aus cancer atlas DMAQ?-->
 
@@
 This information could include a socioeconomic indicators such as unemployment, education level attained.
-The areas may be ranked and the statistics could be presented as categories describing the ranking.
+The areas are usually ranked and allocated to quintiles to be presented as categories describing the ranking.
 
@@
 Table: \label{tab:supporting-stats} Supporting statistics in public facing atlases
+
 | Atlas                    | Geographic hierarchies | Population Distributions | Demographics | Statistical Uncertainty | Socioeconomic Indicators |

### 2019-08-19 review_article.Rmd
contiguous noncontiguous editing

@@
 
-Cancer maps are effective communication tools for a general or non-expert audience. They are commonly used in the public domain to communicate results of sophisticated statistical analyses.
-The heavy use of choropleth maps within the research literature is reflected in the types of maps that are found in the public domain.
-A grey literature review conducted by (ref) identified 33 cancer atlases published on the internet between January 2010 and November 2015.
-These choropleth maps were mostly published by non-commercial organisations, including not-for-profits (NFPs), government, research organisations, advocacy groups or a partnership between an NFP & government.
+Cancer maps are effective communication tools for a general public or non-expert audience. They are commonly used in the public domain to communicate results of sophisticated statistical analyses.
+The heavy use of choropleth maps within the research literature is also reflected in the types of maps that are found in the public domain.
+A grey literature review of cancer atlases available in the public domain<!--(ref Jessie)--> identified 33 cancer atlases published on the internet between January 2010 and November 2015.
+These choropleth maps were mostly published by non-commercial organisations, including not-for-profits, government, research organisations, advocacy groups or government funded partnerships.
 Only one map was published by a commercial entity.
@@
 <!-- geographic coverage-->
-The cancer atlases covered geographies from all around the world, four were global atlases.
+The cancer atlases covered geographies from all around the world, but only four were global atlases.
 Most focussed on single nations, the United States was considered by eleven atlases, the United Kingdom by seven, followed by three of Australia, two of Canada, and one of each from Switzerland, Germany, Norway.
@@
 
-The atlases are not only maps.
-Many provide additional information to help users understand and interpret the spatial distribution.
+The atlases provided additional information to help users understand and interpret the spatial distribution presented.
 Table \ref{tab:supporting-stats} explores the use of these supports in a variety of online map sources.
@@
 <!-- Demographics -->
-Age groups,
-
+Atlases often specified their use of gender in the diplays.
+Atlases that allowed for inteaction also allowed users to select between displays for males, females or both depending on the type of cancers explored.
 <!-- Socioeconomic Indicators -->
+Few atlases also explored the impact of rurality on cancer rates.
 
@@
 
-<!--
-
-'In Australia the urban federal constituencies occupy only a tenth of the land, but contain nine tenths of the people.
-It would be almost unthinkable to show election results for that country on a conventional equal land area map.'
-This 1966 cartogram uses mostly straight lines, and the result looks very little like the geographical shape of Australia.
-
-'Given the increasingly uneven population distribution of the United States and the growing social divides between the populations of neighbourhoods living at different densities, the need for cartograms like this is greater now than ever.'
-
--->
+Australia presents an extreme case of an urban rural divide.
+The land mass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas.
+To present election results on a choropleth map should be 'unthinkable', as it means diminishing the visual impact of majority of the electorates.
+A 1966 cartogram was presented an alternative which uses mostly straight lines, and the result looks very little like the geographical shape of Australia.
+This issue is felt in any nation which experiences an uneven population distribution.
+The United States is affect by the different densities of neighbourhoods or states.
+As this population distribution continues to change the need for cartograms as an alternative to a choropleth map should only increase.
 
@@
 World map projections reflect the frequent perspectives used to view the earth.
-Choropleth maps will always be distorted if they cover enough of the globe, just as photographs of the globe from space.
+Choropleth maps will always be distorted if they cover enough of the globe, just like photographs of the globe from space.
 Choropleth creation requires choosing a map projection that shows a favourable distortion of the geography for presenting the set of spatial information.
-Diagrams that do not specify a projection can be considered to have some unknown projection. <!-- ??? cite tobler.-->
-
-<!-- Denominators -->
-If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed beyond choosing a projection. 
-The purposeful distortion of the map space is beneficial when a uniform density of the map base is desired.
-When a map base is transformed according to population density, population becomes a uniformly distributed background for the statistic presented [@GOINO].
+Selecting a display can prevent misinterpretation of global statistics, as global maps face the challenge of equitable displays of land mass on maps [@RSCW].
+Maps that do not specify a projection can be considered to have some unknown projection. <!-- ??? cite tobler.-->
+If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed, beyond choosing a reasonable projection.
 
@@
 <!-- Denominators -->
-If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed beyond choosing a projection. 
-The purposeful distortion of the map space is beneficial when a uniform density of the map base is desired, @ACTUC suggests 'population distribution is often extremely uneven in former British colonies', this makes the distortion necessary.
-When a map base is transformed according to population density, population becomes a uniformly distributed background for the statistic presented [@GOINO]. 
-Using choropleth maps for population characteristics requires graphic distortions when the population concerned varies greatly in density [@CTTMB].
-When implementing a distortion of the geographical shape according to population, an area cartogram [@NAC], population-by-area cartograms [@TAAM], or iso-demographic map is the result.
+
+The purposeful distortion of the map space, transformed according to population density, is beneficial when a uniform density of the map base is desired.
+Population then becomes a uniformly distributed background for the statistic presented [@GOINO].
+@ACTUC suggests 'population distribution is often extremely uneven in former British colonies', this makes the distortion necessary [@CTTMB].
+When implementing a distortion of the geographical shape according to population, the resulting display is an area cartogram [@NAC], population-by-area cartogram [@TAAM], or iso-demographic map.
 
@@
 A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, making interpretation of spatial patterns very difficult.
-The distortion of a cartogram prevents the density obscuring the spatial patterns [@TAAM]. 
+The distortion of a cartogram accounts for the population density, preventing it from obscuring the spatial patterns [@TAAM]. 
 The spatial transformation of map regions relative to the data emphasises the data distribution instead of land size [@CBATCC].
 When visualising population statistics @ACTUC considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
-@CTTMB suggest that spatial socio-economic data is best presented on a cartogram for urban areas.
-@HEDP agrees that 'cancer occurs in people, not in geographical areas' and the map bases of population reflect this and avoid allocating 'undue prominence' to rural areas. 
+@HEDP suggests that 'cancer occurs in people, not in geographical areas' and @CTTMB believe that spatial socio-economic data, like cancer rates, are best presented on a cartogram for urban areas as the population map base avoids allocating 'undue prominence' to rural areas.
 @MTMSIH encourage the use of cartograms to highlight small areas and uncover local-level inequalities.
-
-<!--
- Disease atlases were implemented using both geographic (Howe 1963) and demographic bases (Howe 1970).
-
-Design goals:
-- to maintain as much as possible of the intricate detail
-- to maximize the simplicity of the transformation
-- maintenance of directional accuracy
-- attempt to influence and direct the map user's attention,
-
-2) that the contiguity relationships between the DCU's should be maintained, 
-3) that true directions should be retained from the centre of the transformation,
-4) that the main shape characteristics of the DCU's should be preserved, and 
-5) that the map should possess the overall shape of the built-up area. 
--->
-
-<!--
-@NISCC
-two advantages of the cartogram: (1) new relationships are often revealed visually, and ( Finally, and most important for the present discussion, they suggest that the ability to preserve shape of the statistical units drops with an increase in the number of these units, and in situations where the difference is great between the smallest and largest values mapped.
-Until we know how the map reader compares it to an ordinary thematic map, we are experimenting with a graphic device that may be reducing (or at least hindering) the flow of information from the map to the map reader.
--->
 
@@
 
-The creation of cartograms was largely in the hands of professional cartographers [@CD]. 
-@ACTUC discusses early approaches including John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball bearing approach and Tobler's (1973) computer programs.
+The creation of cartograms was historically in the hands of professional cartographers [@CD]. 
+Early approaches including John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball bearing approach and Tobler's (1973) computer programs [@ACTUC].
+Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’[@CD].
 @HEDP discusses the impact of electronic computer-assisted techniques.
-Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’[@CD].
-
-There have been many algorithms presented, @SAIC provided a framework to investigate implementations and the "statistical accuracy, geographical accuracy, and topological accuracy".
 
@@
 @ACTUC reiterates: 'There is no "best" cartogram or method of creating cartograms just as there is no "best" map' (Monmonier and Schnell, 1988). 
+There have been many algorithms presented, @SAIC provided a framework to investigate implementations and the "statistical accuracy, geographical accuracy, and topological accuracy".
 
@@
 This transformation often occurs at the expense of the shape of areas [@CBATCC, @NAC, @TAAM].
-From a computer graphics perspective, @ACA believe it is a problem of 'map deformation' to account for the value assigned to each area, they provide three methods for creating value-by-area cartograms.
+From a computer graphics perspective, @ACA explain the application of 'map deformation' to account for the value assigned to each area, they provide three methods for creating value-by-area cartograms.
 Examples include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [-@ACTUC], Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and Tikunov’s Line Integral Method, Constraint-Based Method (Kocmoud and House) [-@CBATCC].
@@
 The end result was a 'very accurate isodemo-graphic map of Canada'.
-This intentional design goal coincided with the rising interest in urban geography.
-
-<!--
-presents 7 desirable characteristics
-1. Independent of region traversal order
-2. Independent of coordinate axes
-3. Conformal mapping
-4. Global displacements per iteration
-5. Intersection prevention
-6. Ability to fix (pin down) points
-7. User controls on area vs. shape
-
-Australia (McGlashan 1977),
-
--->
+This intentional design goal coincided with the rising interest in urban geography and presentation of social statistics.
 
@@
 Tobler's Conformal mapping means to preserve angles
-locally so that the shapes of very small areas on a traditional map and a
-cartogram would be similar.
+locally so that the shapes of very small areas on a traditional map and a cartogram would be similar.
 @CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes.
@@
 
-### Non-Contiguous
-
 <!--
-[-@NAC] observation units stay in more or less their correct location and maintain correct shape, but at least some units are separated by gaps from their neighbours.
-Each area in a non-contiguous cartogram is scaled down or up according to value for the specific area.
-A scaling factor specific to
-
-the empty area between units is meaningful. 
-the degree of difference from the original map that is the real message of a cartogram, @NAC
-
-A cartogram using the unit with the mean or median density as the anchor would also show relative discrepancies, but it is more difficult to produce by computer because many units increase in size and overlap.
-@NAC
-
-The success of the
-visual representation (contiguous or noncontiguous)
-depends, in fact, on the readers’
-recognition of the units shown and
-such recognizable units are generally fairly
-large for the area represented. @NAC
-
-constant shapes are easily recognized regardless of scale;
-
-Presenting an alternative view can help to correct interpretation of global statistics when some areas loom large on maps [@RSCW].
--->
-
-```{r ggncont, fig.cap = "A Non - contiguous cartogram map of the United States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each state shape has been maintained, but the size has altered according to the population of the state in 2015.
-The state of California has remained closer to it's original size than it's surrounding states.
-The North East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut.
-This draws attention to the densely populated North East region, and the sparse Mid West.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
-knitr::include_graphics("figures/ggncont.png")
-```
-
-```{r ggdorl, fig.cap = "A dorling cartogram of the Unites States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each state has been represented by a circle, but the size was determined by the population of the state in 2015.
-The North East states remain closer to their neighbours, and may be displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
-knitr::include_graphics("figures/ggdorl.png")
-```
-
-Dorling -@ACTUC puts forward a simple question:
-
->"If, for instance, it is desirable that areas on a map have boundaries which are as simple as possible, why not draw the areas as simple shapes in the first place?"
-
-He answers this with his implementation of maps created with 'the simplest of all shapes'.
-While contiguous cartograms may be a 'more sophisticated' method, they produce 'very complex shapes'.
-Circular cartograms use the same simple shapes for every region represented, and size them according to the statistic represented or the population for a base map.
-This familar shape may be more effective that contiguous cartograms, as the shapes used have 'no meaning' after distortions are applied @NISCC.
-To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighbouring areas over many iterations.
-This implementation can work for up to 'one hundred thousand' areas.
-
-<!--
--sacrifice communication efficiency;
--There is evidence, too, that angular shapes are more easily seen than ones with smooth outlines.
-Shapes that contain the qualities of simplicity and regularity are more readily perceived than others that do not possess these qualities.
--Finally, psychologists tell us that shapes having familiarity are more quickly
-recognized than nonsense shapes or those that have no meaning.
--The cartographer can efficiently communicate with this kind of map if he has preserved particular elements of the original outline and incorporated them in the new shape.
--
-@NISCC
-
--->
 
@@
 1987), the Soviet Union (Levin 1980), Japan (Shigematsu 1977), the Federal Republic of Germany
+-->
 
-<!--
-@CVBGSD presents bivariate maps, using size and colour for different variables. 
--->
+### Non-Contiguous
+
+Non-Contiguous cartograms succeed in maintaing the shape of the areas presented.
+Each area stays in a similar position to their location on a choropleth map.
+The choropleth map base is often also plotted as a comparison point to highlight the change in area.
+The addition is the gap between areas, created as each individual area shrinks or grows according the associated value of the statistic.
+@NAC discusses creation of these maps, the significance of the empty areas left between the geographic boundaries and the new shape, and the 'degree of difference from the original map that is the real message' of these displays.
+
+The shapes are valuable for recognition and allows users to orient themselves on the display.
+Map creators can efficiently communicate with this kind of map by keeping the outlines or particular elements of the original in the new shape [@NISCC].
+The scale of the areas does not impact on the shape recognition. However it may impact on the visibility of all areas if small areas expand beyond their boundaries.
+
+```{r ggncont, fig.cap = "A Non - contiguous cartogram map of the United States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state shape has been maintained, but the size has altered according to the population of the state in 2015.
+The state of California has remained closer to it's original size than it's surrounding states.
+The North East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut.
+This draws attention to the densely populated North East region, and the sparse Mid West.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+knitr::include_graphics("figures/ggncont.png")
+```
+
+### Dorling
+
+Daniel Dorling presented an alternative to highlight the spatial distribution and neighbourhood relationships without complex distortions of borders and boundaries.
+This approach opposes preserving the intricate shape details.
+This concept is founded in the simple question put forward by Daniel Dorling -@ACTUC:
+
+>"If, for instance, it is desirable that areas on a map have boundaries which are as simple as possible, why not draw the areas as simple shapes in the first place?"
+
+He answers this with his implementation of maps created with 'the simplest of all shapes'. He acknowledged the sophistication of contiguous cartograms but critiqued their 'very complex shapes'.
+Circular cartograms use the same simple shapes for every region represented, and size the shapes according to the statistic represented or the population for a base map.
+This familar shape may be more effective for understanding the spatial distribution than contiguous cartograms, as the 'nonsense' shapes used have 'no meaning' after distortions are applied @NISCC.
+To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighbouring areas over many iterations.
+This implementation can work for up to 'one hundred thousand' areas.
+
+```{r ggdorl, fig.cap = "A dorling cartogram of the Unites States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been represented by a circle, but the size was determined by the population of the state in 2015.
+The North East states remain closer to their neighbours, and may be displaced from their geographic location.
+The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+knitr::include_graphics("figures/ggdorl.png")
+```
+
+### Alternative displays
+
+Original value by area cartograms, especially the tile method, laid the foundation for the modern alternative displays.
 

### 2019-08-19 review_article.Rmd
Add figure reference

@@
 By providing a visual representation of cancer outcomes, geographic patterns of disease are able to be identified and effectively addressed.
+@SE recognises one of the key challenges with mapping spatial patterns of disease is the design of visualisations.
+This paper addresses the visualisation techniques and their applications to cancer statistics. Highlighting the differences and historic use of these displays.
 
@@
 A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns on a familiar geographic form. 
-Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
+Figure \ref{fig:choro} shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
 Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS]. 
@@
 
+The atlas of Cancer in Queensland [@DACQ] focussed on highlighting the difference in experience for those living in rural and disadvantaged areas.
+They used Standardised Incidence Ratios modelled via Bayesian methods, and explored the information presented in a range of selected atlases.
+These atlases provided Incidence, Survival and Mortality rates to the public on Areal or Isopleth map bases.
+
+The presentation of these rates considered not only the map base, but also appropriate choices of colour blind friendly colour schemes, and categories of the values to highlight signifcantly different areas.
+
 <!--
@@
 Bertin 1967
+Coleman MP, Gatta G, Verdecchia A, Esteve J, Sant M, Storm H, Allemani C,Ciccolallo L, Santaquilani M, Berrino F:EUROCARE-3 summary: cancersurvival in Europe at the end of the 20th century.Ann Oncol2003,14(Suppl 5):v128-149
 
-Current atlases utilising choropleth methods 
-- include Australia:
+National Cancer Institute:Cancer Trends Progress Report - 2009/2010 UpdateBethesda: NCI, NIH, DHHS; 2010
 
-cites: Cancer Mortality Atlas of EU and of the European Economic Area last published 8 years ago
-@MTMSIH Bayesian disease mapping methods for areal data - explains key concepts applied to create an atlas
-Explores Australia as a case study, utilising QLD SA2 areas
+Pickle LW:A history and critique of U.S. mortality atlases.Spatial andSpatio-temporal Epidemiology2009,1:3-17.
 
+Mason TJ, McKay FW, Hoover R, Blot WJ, Fraumeni JF:Atlas of cancermortality for U.S. counties: 1950-1969Washington: U.S. Govt. Printing Office;1975.
+howmick T, Robinson AC, Gruver A, MacEachren AM, Lengerich EJ:Distributed usability evaluation of the Pennsylvania Cancer Atlas.Int JHealth Geogr2008,7:36.
+
+Queensland Cancer Registry:Cancer in Queensland: Incidence, Mortality,Survival and Prevalence,1982 to 2007Brisbane: QCR, Cancer CouncilQueensland and Queensland Health; 2010.
+
+New York State Department of Health Maps of Cancer Incidence byCounty.[http://www.health.state.ny.us/statistics/cancer/registry/cntymaps/index.htm].
+Quinn M, Wood H, Cooper N, Rowan S, Eds:Cancer Atlas of the UnitedKingdom and Ireland 1991-2000.London: Office for National Statistics;2005.
 -->
@@
 Tobler's Conformal mapping means to preserve angles
-locally so that the shapes of very small areas on a traditional map and a cartogram would be similar.
+locally so that the shapes of small areas on a traditional map and a cartogram would be similar.
 @CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes. 
-Distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the noncontiguous cartogram but is replaced by the meaningful empty-space property [@NAC, @ECGC].
+@MDAC define the term 'mosaic cartograms' for hexagonal tile displays, where the amount of tiles for each area can be used to communicate the statistic of regions. 
+The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off be made between boundary complexity and simplicity.
 
 <!--
-
 Tobler's method and the many implementations that 'elaborated' on it are derived from 'numerical approximations to a pair of equations'[@ACTUC].
 They all operate through incremental adjustments, and can produce wildly different outcomes from small changes in the inputs.
-
-@TFYCC Value-Area Cartograms.
-In these cartograms a region,country, or continent is subdivided into small regions, each of which is represented by a rectangle.
-This rectangle is proportionate in area to the value which it represents in certain statistical distributions.
-The regions are grouped in approximately the same positions as they are on the map.
 
@@
 
+As the trade-off regarding boundaries approaches simplicity, the distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the noncontiguous cartogram but is replaced by the meaningful empty-space property [@NAC, @ECGC].
 The shapes are valuable for recognition and allows users to orient themselves on the display.
@@
 
-### Dorling
+### Alternative Displays
 
@@
 
-### Alternative displays
-
-Original value by area cartograms, especially the tile method, laid the foundation for the modern alternative displays.
-
-@MDAC define the term 'mosaic cartograms'. 
-Compare amount of tiles to contrast population of regions.
-'Cartograms show a data value per input region by scaling each region such that its area is proportional to its data value.
-Mosaic cartograms show data in multiples of tiles, hence the input data must consist of, or be cast into, small integer units.'
+The groundwork for the simplistic approach had been laid in the mid 1930's be Raisz.
+@TFYCC quotes the official definition of Value-Area Cartograms, the simplistic displays which represent each area as a single rectangle, sized according to the value of the statistic.
+This rectangluar display also allows for tiling, where geographic neighbours also share rectangle borders.
 
@@
 <!--
-@SE Researchers mapping spatial patterns of disease
-for small areas are confronted with a number of
-challenges, including cartographic design, the
-modifiable area unit problem, and also the “small
-number” problem, where some places have no,
-or few, disease events and/or a low population at
-risk of contracting the disease.
-
 -->

### 2019-08-19 review_article.Rmd
Cancer atlases in literature

@@
 A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns on a familiar geographic form. 
-Figure \ref{fig:choro} shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
+Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
 Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS]. 
@@
 The presentation of these rates considered not only the map base, but also appropriate choices of colour blind friendly colour schemes, and categories of the values to highlight signifcantly different areas.
-
 <!--

### 2019-08-20 review_article.Rmd
change order of grid and small edits in text..
Need to check these as I had some conflicts that I had to resolve..

@@
 The visualisation methods used to present cancer statistics will depend on the intended message and users of the map. 
-Presenting statistics as visualisations requires transforming individual observations into aggregations of communities as geographical units.
+Transforming statistics to visualisations takes advantage of the human visual system, which has been demonstrated to an effective communication channel for a wide range of audiences (insert munzer reference).
+Visualising cancer statistics enables the non-expert audience to access the outputs of complicated spatial epidemeaological analyses.
+These visualisation show individual observations, or observation points, aggregated into communities and geographical units.
+Visualising diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH]. 
+Disease maps help to present geographic patterns that may be overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
+By providing a visual representation of cancer statistics, geographic patterns of disease are able to be identified, investigated further and/or effectively addressed.
+@SE recognises one of the key challenges with mapping spatial patterns of disease is the design of visualisations.
+This paper addresses the visualisation techniques and their applications to cancer statistics. Highlighting the differences and historic use of these displays.
+Transforming statistics to visualisations takes advantage of the human visual system, which has been demonstrated to an effective communication channel for a wide range of audiences (insert munzer reference).
+Visualising cancer statistics enables the non-expert audience to access the outputs of complicated spatial epidemeological analyses.
+These visualisation show individual observations, or observation points, aggregated into communities and geographical units.
 Visualising diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH]. 
 Disease maps help to present geographic patterns that may have been overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
-By providing a visual representation of cancer outcomes, geographic patterns of disease are able to be identified and effectively addressed.
-@SE recognises one of the key challenges with mapping spatial patterns of disease is the design of visualisations.
-This paper addresses the visualisation techniques and their applications to cancer statistics. Highlighting the differences and historic use of these displays.
+By providing a visual representation of cancer statistics, geographic patterns of disease are able to be identified, nvestigated further and/or effectively addressed.
 
 # Disease mapping methods
+<!-- insert paragraph summarising/ stating the different types of maps used for cancer mapping. --->
 
@@
 Early versions of choropleth maps used symbols or patterns instead of colour.
+
 As an alternative storage device to a table, it preserves locations for geographically ordered data, with usage dating back to the 1800s [@GOINO].
 @CPISACA discuss the use of choropleths to visualise cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
+@CPISACA discusses their use in visualising cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
+Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
+As such, many examples of choropleth maps are found on the websites of government, media, health and other public facing organisations (insert ref for Jessie's grey lit review).
+A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form. 
+Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
 
@@
 Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the higher prevalence of cancers in older populations.
+
 Howe [-@HEDP] describes Stock's development of the standardised mortality ratios through the 1930s.
-Table \ref{tab:measures} presents summarises the measures presented in published cancer atlases, and provides a definition of each measure.
+Table \ref{measures} summarises the measures presented in published cancer atlases, between 2005 and 2015 and provides a definition of each measure.
 
@@
 
-```{r choropleth-grid-create, message=FALSE, eval = FALSE}
+```{r choropleth-grid-create, echo=FALSE, message=FALSE, eval = FALSE,}
 library(grid)
@@
 
-plot1 <- png::readPNG('cancer_map_images/choropleths/1.png')
-plot2 <- png::readPNG('cancer_map_images/choropleths/14.choropleth1.png')
-plot3 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
-plot4 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
-plot5 <- png::readPNG('cancer_map_images/choropleths/5.choropleth.png')
+plot1 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
+plot2 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
+plot3 <- png::readPNG('cancer_map_images/choropleths/1.png')
+plot4 <- png::readPNG('cancer_map_images/choropleths/29.choropleth.png')
+plot5 <- png::readPNG('cancer_map_images/choropleths/30.choropleth.png')
 plot6 <- png::readPNG('cancer_map_images/choropleths/6.choropleth.png')
-plot6 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
-plot7 <- png::readPNG('cancer_map_images/choropleths/8.choropleth.png')
+plot7 <- png::readPNG('cancer_map_images/choropleths/8.choropleth2.png')
 plot8 <- png::readPNG('cancer_map_images/choropleths/20.choropleth1.png')
-plot9 <- png::readPNG('cancer_map_images/choropleths/21.choropleth2.png')
+plot9 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
 plot10 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
 plot11 <- png::readPNG('cancer_map_images/choropleths/24.choropleth.png')
-plot12 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
+plot12 <- png::readPNG('cancer_map_images/choropleths/14.choropleth1.png')
 plot13 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
 plot14 <- png::readPNG('cancer_map_images/choropleths/27.choropleth.png')
-plot15 <- png::readPNG('cancer_map_images/choropleths/29.choropleth.png')
-plot16 <- png::readPNG('cancer_map_images/choropleths/30.choropleth.png')
+plot15 <- png::readPNG('cancer_map_images/choropleths/21.choropleth2.png')
+plot16 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
+plot17 <- png::readPNG('cancer_map_images/choropleths/5.choropleth.png')
 
-choropleth_grid <- gridExtra::grid.arrange(grid::rasterGrob(plot1),grid::rasterGrob(plot2), grid::rasterGrob(plot3), 
-                        grid::rasterGrob(plot4), grid::rasterGrob(plot5), grid::rasterGrob(plot6),
-                        grid::rasterGrob(plot7), grid::rasterGrob(plot8), grid::rasterGrob(plot9),
-                        grid::rasterGrob(plot10),grid::rasterGrob(plot11), grid::rasterGrob(plot12), 
-                        grid::rasterGrob(plot13), grid::rasterGrob(plot14), grid::rasterGrob(plot15), 
-                        grid::rasterGrob(plot16), nrow=4)
+choropleth_grid <- gridExtra::grid.arrange(
+  grid::rasterGrob(plot1), 
+  grid::rasterGrob(plot2), 
+  grid::rasterGrob(plot3), 
+  grid::rasterGrob(plot4), 
+  grid::rasterGrob(plot5), 
+  grid::rasterGrob(plot6),
+  grid::rasterGrob(plot7), 
+  grid::rasterGrob(plot8), 
+  grid::rasterGrob(plot9),
+  grid::rasterGrob(plot10),
+  grid::rasterGrob(plot11), 
+  grid::rasterGrob(plot12),
+  grid::rasterGrob(plot13), 
+  grid::rasterGrob(plot14), 
+  grid::rasterGrob(plot15),
+  grid::rasterGrob(plot16),
+  grid::rasterGrob(plot17))
 
-ggsave(filename = "figures/choropleth_grid.png", plot = choropleth_grid, device = "png", width = 6, height = 5, dpi = 300)
+ggsave(choropleth_grid, filename = "figures/choropleth_grid.png", device = "png", width = 12, height = 6)
+
 ```
 
-```{r choropleth-grid-print, echo = FALSE, results = "asis"}
+```{r choropleth-grid-print, echo = FALSE, out.width= '100%', fig.cap = "Insert grid title and references to individual tiles (ref)"}
 knitr::include_graphics("figures/choropleth_grid.png")

### 2019-08-20 review_article.Rmd
delete chloropleth folder

@@
   grid::rasterGrob(plot12),
-  grid::rasterGrob(plot13), 
+  #grid::rasterGrob(plot13), 
   grid::rasterGrob(plot14), 

### 2019-08-20 review_article.Rmd
part arrange tiles by country

@@
 
-plot1 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
-plot2 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
-plot3 <- png::readPNG('cancer_map_images/choropleths/1.png')
-plot4 <- png::readPNG('cancer_map_images/choropleths/29.choropleth.png')
-plot5 <- png::readPNG('cancer_map_images/choropleths/30.choropleth.png')
-plot6 <- png::readPNG('cancer_map_images/choropleths/6.choropleth.png')
+plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth2.png')
+plot2 <- png::readPNG('cancer_map_images/choropleths/29.choropleth.png')
+plot3 <- png::readPNG('cancer_map_images/choropleths/30.choropleth.png')
+plot4 <- png::readPNG('cancer_map_images/choropleths/1.png')
+plot5 <- png::readPNG('cancer_map_images/choropleths.2/6.choropleth.png')
+plot6 <- png::readPNG('cancer_map_images/choropleths/27.choropleth.png')
 plot7 <- png::readPNG('cancer_map_images/choropleths/8.choropleth2.png')
 plot8 <- png::readPNG('cancer_map_images/choropleths/20.choropleth1.png')
-plot9 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
+plot9 <- png::readPNG('cancer_map_images/choropleths.2/7.choropleth.png')
 plot10 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
-plot11 <- png::readPNG('cancer_map_images/choropleths/24.choropleth.png')
+plot11 <- png::readPNG('cancer_map_images/choropleths.2/24.choropleth.png')
 plot12 <- png::readPNG('cancer_map_images/choropleths/14.choropleth1.png')
 plot13 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
-plot14 <- png::readPNG('cancer_map_images/choropleths/27.choropleth.png')
-plot15 <- png::readPNG('cancer_map_images/choropleths/21.choropleth2.png')
+plot14 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
+plot15 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
 plot16 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')

### 2019-08-22 review_article.Rmd
Supporting statistics

--- 2019-08-20 review_article.Rmd
@@
 The visualisation methods used to present cancer statistics will depend on the intended message and users of the map. 
-Transforming statistics to visualisations takes advantage of the human visual system, which has been demonstrated to an effective communication channel for a wide range of audiences (insert munzer reference).
-Visualising cancer statistics enables the non-expert audience to access the outputs of complicated spatial epidemeaological analyses.
-These visualisation show individual observations, or observation points, aggregated into communities and geographical units.
+Presenting statistics as visualisations requires transforming individual observations into aggregations of communities as geographical units.
 Visualising diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH]. 
-Disease maps help to present geographic patterns that may be overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
-By providing a visual representation of cancer statistics, geographic patterns of disease are able to be identified, investigated further and/or effectively addressed.
+Disease maps help to present geographic patterns that may have been overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
+By providing a visual representation of cancer outcomes, geographic patterns of disease are able to be identified and effectively addressed.
 @SE recognises one of the key challenges with mapping spatial patterns of disease is the design of visualisations.
 This paper addresses the visualisation techniques and their applications to cancer statistics. Highlighting the differences and historic use of these displays.
-Transforming statistics to visualisations takes advantage of the human visual system, which has been demonstrated to an effective communication channel for a wide range of audiences (insert munzer reference).
-Visualising cancer statistics enables the non-expert audience to access the outputs of complicated spatial epidemeological analyses.
-These visualisation show individual observations, or observation points, aggregated into communities and geographical units.
-Visualising diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH]. 
-Disease maps help to present geographic patterns that may have been overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
-By providing a visual representation of cancer statistics, geographic patterns of disease are able to be identified, nvestigated further and/or effectively addressed.
 
 # Disease mapping methods
-<!-- insert paragraph summarising/ stating the different types of maps used for cancer mapping. --->
 
@@
 A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns on a familiar geographic form. 
-Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
+\ref{fig:choro} 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
 Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS]. 
@@
 Early versions of choropleth maps used symbols or patterns instead of colour.
-
 As an alternative storage device to a table, it preserves locations for geographically ordered data, with usage dating back to the 1800s [@GOINO].
 @CPISACA discuss the use of choropleths to visualise cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
-@CPISACA discusses their use in visualising cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
-Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
-As such, many examples of choropleth maps are found on the websites of government, media, health and other public facing organisations (insert ref for Jessie's grey lit review).
-A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns across a familiar geographic form. 
-Figure 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
 
@@
 Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the higher prevalence of cancers in older populations.
-
 Howe [-@HEDP] describes Stock's development of the standardised mortality ratios through the 1930s.
-Table \ref{measures} summarises the measures presented in published cancer atlases, between 2005 and 2015 and provides a definition of each measure.
+Table \ref{tab:measures} presents summarises the measures presented in published cancer atlases, and provides a definition of each measure.
 
@@
 
-```{r choropleth-grid-create, echo=FALSE, message=FALSE, eval = FALSE,}
+```{r choropleth-grid-create, message=FALSE, eval = FALSE}
 library(grid)
@@
 
-plot1 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
-plot2 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
-plot3 <- png::readPNG('cancer_map_images/choropleths/1.png')
-plot4 <- png::readPNG('cancer_map_images/choropleths/29.choropleth.png')
-plot5 <- png::readPNG('cancer_map_images/choropleths/30.choropleth.png')
+plot1 <- png::readPNG('cancer_map_images/choropleths/1.png')
+plot2 <- png::readPNG('cancer_map_images/choropleths/14.choropleth1.png')
+plot3 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
+plot4 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
+plot5 <- png::readPNG('cancer_map_images/choropleths/5.choropleth.png')
 plot6 <- png::readPNG('cancer_map_images/choropleths/6.choropleth.png')
-plot7 <- png::readPNG('cancer_map_images/choropleths/8.choropleth2.png')
+plot6 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
+plot7 <- png::readPNG('cancer_map_images/choropleths/8.choropleth.png')
 plot8 <- png::readPNG('cancer_map_images/choropleths/20.choropleth1.png')
-plot9 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
+plot9 <- png::readPNG('cancer_map_images/choropleths/21.choropleth2.png')
 plot10 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
 plot11 <- png::readPNG('cancer_map_images/choropleths/24.choropleth.png')
-plot12 <- png::readPNG('cancer_map_images/choropleths/14.choropleth1.png')
+plot12 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
 plot13 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
 plot14 <- png::readPNG('cancer_map_images/choropleths/27.choropleth.png')
-plot15 <- png::readPNG('cancer_map_images/choropleths/21.choropleth2.png')
-plot16 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
-plot17 <- png::readPNG('cancer_map_images/choropleths/5.choropleth.png')
+plot15 <- png::readPNG('cancer_map_images/choropleths/29.choropleth.png')
+plot16 <- png::readPNG('cancer_map_images/choropleths/30.choropleth.png')
 
-choropleth_grid <- gridExtra::grid.arrange(
-  grid::rasterGrob(plot1), 
-  grid::rasterGrob(plot2), 
-  grid::rasterGrob(plot3), 
-  grid::rasterGrob(plot4), 
-  grid::rasterGrob(plot5), 
-  grid::rasterGrob(plot6),
-  grid::rasterGrob(plot7), 
-  grid::rasterGrob(plot8), 
-  grid::rasterGrob(plot9),
-  grid::rasterGrob(plot10),
-  grid::rasterGrob(plot11), 
-  grid::rasterGrob(plot12),
-  grid::rasterGrob(plot13), 
-  grid::rasterGrob(plot14), 
-  grid::rasterGrob(plot15),
-  grid::rasterGrob(plot16),
-  grid::rasterGrob(plot17))
+choropleth_grid <- gridExtra::grid.arrange(grid::rasterGrob(plot1),grid::rasterGrob(plot2), grid::rasterGrob(plot3), 
+                        grid::rasterGrob(plot4), grid::rasterGrob(plot5), grid::rasterGrob(plot6),
+                        grid::rasterGrob(plot7), grid::rasterGrob(plot8), grid::rasterGrob(plot9),
+                        grid::rasterGrob(plot10),grid::rasterGrob(plot11), grid::rasterGrob(plot12), 
+                        grid::rasterGrob(plot13), grid::rasterGrob(plot14), grid::rasterGrob(plot15), 
+                        grid::rasterGrob(plot16), nrow=4)
 
-ggsave(choropleth_grid, filename = "figures/choropleth_grid.png", device = "png", width = 12, height = 6)
-
+ggsave(filename = "figures/choropleth_grid.png", plot = choropleth_grid, device = "png", width = 6, height = 5, dpi = 300)
 ```
 
-```{r choropleth-grid-print, echo = FALSE, out.width= '100%', fig.cap = "Insert grid title and references to individual tiles (ref)"}
+```{r choropleth-grid-print, echo = FALSE, results = "asis"}
 knitr::include_graphics("figures/choropleth_grid.png")
@@
 
-The atlases provided additional information to help users understand and interpret the spatial distribution presented.
-Table \ref{tab:supporting-stats} explores the use of these supports in a variety of online map sources.
-Demographics include information regarding the age and sex distribution of the areas presented.
-Socioeconomic indicators include Unemployment rates of regions, Remoteness and Education levels achieved.
-Statistical uncertainty is communicated through a Confidence (CI) interval, statistical significance levels, Boxplots, Distribution plots, and reporting the Sample Size and Standard Deviations.
+The atlases provided additional information to help users understand and interpret the spatial distribution presented. The use of these supports were found in a variety of online map sources.
 These additional pieces of information can help to validate, explain or explore relationships between cancer incidence and survival rates.
 
-Table: \label{tab:supporting-stats} Supporting statistics in public facing atlases
+#### Geographic hierarchies
+While atlases are often used to describe differences between areas, statistics may be displayed at different levels of aggregation.
+The World Health Organisation and UN Regions can also be used to aggregate areas [@IARC_3].
+World atlases can allow for displays of data aggregated into continents, countries, states, provinces and congressional districts [@USCS].
 
-| Atlas                    | Geographic hierarchies | Population Distributions | Demographics | Statistical Uncertainty | Socioeconomic Indicators |
-|--------------------------|------------------------|--------------------------|--------------|-------------------------|--------------------------|
-| All Ireland Cancer Atlas |                        |                          |              |                         |                          |
-|                          |                        |                          |              |                         |                          |
-|                          |                        |                          |              |                         |                          |
+#### Population distribution
+It is likely that each population area will have a different amount of people, information regarding the distribution of population levels may be provided in a table or histogram display [@NICR_1]. 
+Atlases have the opportunity to connect the population to the land available to them by communicating population density.
 
-<!-- geographic hierarchies  -->
-While atlases are often used to describe differences between areas, the may allow statistics to be displayed at different levels of aggregation.
-World atlases can allow for displays of data aggregated into continents, country, states or provinces [@IARC_3].
-
-<!-- population distribution -->
-It is likely that each population area will have a different amount of people, information regarding the distribution of population levels may be provided in a table or histogram display [@NICR_1].
-Atlases could also connect the population to the land available to them by communicating population density.
-
-<!-- statistical uncertainty -->
-These additional statistics often include a measure of the statistical uncertainty of the values of the statistics presented in a choropleth.
+#### Statistical uncertainty
+Statistical uncertainty is communicated through a Confidence (CI) interval, statistical significance levels, Boxplots, Distribution plots, and reporting the Sample Size and Standard Deviations.
+The additional statistics often include a measure of the statistical uncertainty of the values of the statistics presented in a choropleth.
 In the review of atlases in the public domain, atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map.
@@
 
-<!-- Demographics -->
-Atlases often specified their use of gender in the diplays.
-Atlases that allowed for inteaction also allowed users to select between displays for males, females or both depending on the type of cancers explored.
-<!-- Socioeconomic Indicators -->
-Few atlases also explored the impact of rurality on cancer rates.
+#### Demographics
+Demographics include information regarding the age and sex distribution of the areas presented. 
+<!-- sex -->
+Sex is an important cofactor for cancer atlases. 
+As some cancers are sex specific, and others may be found in both males and females, atlases often specify their use of gender in the diplays. 
+Atlases can allow for interaction as users can select between displays for males, females or both depending on the type of cancers explored.
+
+#### Socioeconomic indicators
+Socioeconomic indicators can explain how the experience of cancer prevalence varies for various members of a society.
+These indicators include unemployment rates, poverty rates and percentages, remoteness and education levels achieved though only a few atlases also explored the impact of rurality on cancer rates. 
+The Human Development Index can be used to understand the socioeconomic experience of a community, as can the Income levels as measured by World bank list of economies[@IARC_3].
 
@@
 To be able to recognise the significant changes, a reader will usually have to know the initial geography to find the differences in the new cartogram layout [@NAC].
-Tobler's Conformal mapping means to preserve angles
-locally so that the shapes of small areas on a traditional map and a cartogram would be similar.
+Tobler's Conformal mapping means to preserve angles locally so that the shapes of small areas on a traditional map and a cartogram would be similar.
 @CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes. 
@@
 
-<!-- @CPISACA contrasted several online atlases in a 2006 paper, consider for structure?
-
-(only worth including if it is possible for us to implement)
-Tabular form comparing and contrasting
-- Relationship to geography
-- Show using cancer examples
--->
-
 # Animation and Interactivity

### 2019-08-26 review_article.Rmd
condensed supporting materials to one section

@@
 
-Geospatial statistics have been presented and stored on maps for centuries.
-Maps connect data for areas to the geographical representation which is usually familiar to the audience.
-However, it is not enough for areas on maps to be recognisable, the distribution of the statistic must also be visible.
-Cancer statistics directly represent and relate to the people living within individual geographical areas.
-It is reasonable to explore views that enhance the communication of the cancer statistics, especially as in these situations the people are of interest, not the land they live on. 
+Cancer statistics directly represent and relate to the people living within geographical areas.
+Geospatial statistics have been presented on maps to communicate information to the general public for centuries as they utilise familiar geographic shapes.
+The mapping of diseases summarises spatial variations and can be considered a descriptive statistic of the relationship between a disease and the underlying geography [@SEMA].
+The distribution of the statistic must be visible, it is not enough for areas on maps to be recognisable.
+It is reasonable to explore views that enhance the communication of the cancer statistics, especially as it is the people that are of interest, not the land they live on. 
 This has spurred innovations over the previous centuries to enhance maps to effectively communicate cancer outcomes, and health outcomes more broadly.
@@
 
-### Supporting material in cancer atlases
-
-The primary display of an atlas is often the choropleth map.
-The presentation of materials in atlases can also utilise bar charts, tables and scatterplots, these are useful alternative displays.
-They are not only used as an additional display for the cancer incidence and mortality rates, as an atlas is often supplemented with supporting statistics.
-When presenting cancer atlases, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses.' This opinion contradicts Cruickshank's (1947) as cited by @DMAHP, which discusses using visuals as 'formal statistical assessment of the spatial pattern'.
-@CPISACA suggests additional materials such as 'good tables, graphs, and explanatory text' support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers.
-The interactivity of the more modern maps enabled supplementary information to be incorporated without cluttering the screen, such as in a tool tip feature.
-
-<!-- geographic hierarchies  -->
-Cancer atlases are often used to describe differences between areas, they may allow statistics to be displayed at different levels of aggregation.
-World atlases can allow for displays of data aggregated into continents, country, states or provinces [@IARC_3].
-
-<!-- population distribution -->
-It is likely that each population area will have a different amount of people, information regarding the distribution of population levels may be provided in a table or histogram display [@NICR_1].
-Atlases could also connect the population to the land available to them by communicating population density as is shown in the United States map examples.
-
-<!-- demographics: sex, race,  -->
-
-<!-- statistical uncertainty -->
-The additional statistics often also include a measure of the statistical uncertainty of the cancer statistics presented in a choropleths. <!-- reference Aus cancer atlas DMAQ?-->
-
-<!-- Socioeconomic Indicators -->
-There may also be demographic information provided regarding each population area drawn on the maps.
-This information could include a socioeconomic indicators such as unemployment, education level attained.
-The areas are usually ranked and allocated to quintiles to be presented as categories describing the ranking.
-
 ## Publicly available atlases using choropleth methods
@@
 
-### Examples of supporting statistics in public facing atlases
+## Supporting statistics in public facing atlases
 
+The primary display of an atlas is often the choropleth map.
+Cruickshank's (1947) as cited by @DMAHP, discusses using visuals as 'formal statistical assessment of the spatial pattern'. 
+Atlases can also utilise bar charts, tables and scatterplots as useful alternative displays. 
+These displays can present the cancer incidence and mortality rates, as well as the supporting statistics of the atlas. 
+When presenting cancer atlases, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses.'
+@CPISACA suggests additional materials such as 'good tables, graphs, and explanatory text' support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers.
+The interactivity of modern maps enables supplementary information to be incorporated without cluttering the screen, such as in a tool tip feature. 
 The atlases provided additional information to help users understand and interpret the spatial distribution presented. The use of these supports were found in a variety of online map sources.
@@
 
-#### Geographic hierarchies
+### Geographic hierarchies
 While atlases are often used to describe differences between areas, statistics may be displayed at different levels of aggregation.
@@
 
-#### Population distribution
+### Population distribution
 It is likely that each population area will have a different amount of people, information regarding the distribution of population levels may be provided in a table or histogram display [@NICR_1]. 
 Atlases have the opportunity to connect the population to the land available to them by communicating population density.
+Atlases can also connect the population to the land available to them by communicating population density as is shown in the United States map examples.
 
-#### Statistical uncertainty
+### Statistical uncertainty
 Statistical uncertainty is communicated through a Confidence (CI) interval, statistical significance levels, Boxplots, Distribution plots, and reporting the Sample Size and Standard Deviations.
@@
 
-#### Demographics
+### Demographics
 Demographics include information regarding the age and sex distribution of the areas presented. 
@@
 
-#### Socioeconomic indicators
+### Socioeconomic indicators
 Socioeconomic indicators can explain how the experience of cancer prevalence varies for various members of a society.
@@
 The Human Development Index can be used to understand the socioeconomic experience of a community, as can the Income levels as measured by World bank list of economies[@IARC_3]. 
+The areas are usually ranked and allocated to quintiles to be presented as categories describing the ranking.
 
@@
 -->
-
-<!--
--->

### 2019-08-26 review_article.Rmd
scaffold for discussion

@@
 
-The presentation of these rates considered not only the map base, but also appropriate choices of colour blind friendly colour schemes, and categories of the values to highlight signifcantly different areas.
+The presentation of these rates considered not only the map base, but also appropriate choices of colour blind friendly colour schemes, and categories of the values to highlight significantly different areas.
 <!--
@@
 Cruickshank's (1947) as cited by @DMAHP, discusses using visuals as 'formal statistical assessment of the spatial pattern'. 
-Atlases can also utilise bar charts, tables and scatterplots as useful alternative displays. 
+Atlases can also utilise bar charts, tables and scatter-plots as useful alternative displays. 
 These displays can present the cancer incidence and mortality rates, as well as the supporting statistics of the atlas. 
@@
 Sex is an important cofactor for cancer atlases. 
-As some cancers are sex specific, and others may be found in both males and females, atlases often specify their use of gender in the diplays. 
+As some cancers are sex specific, and others may be found in both males and females, atlases often specify their use of gender in the displays. 
 Atlases can allow for interaction as users can select between displays for males, females or both depending on the type of cancers explored.
 
-### Socioeconomic indicators
-Socioeconomic indicators can explain how the experience of cancer prevalence varies for various members of a society.
+### Socio-economic indicators
+Socio-economic indicators can explain how the experience of cancer prevalence varies for various members of a society.
 These indicators include unemployment rates, poverty rates and percentages, remoteness and education levels achieved though only a few atlases also explored the impact of rurality on cancer rates. 
-The Human Development Index can be used to understand the socioeconomic experience of a community, as can the Income levels as measured by World bank list of economies[@IARC_3]. 
+The Human Development Index can be used to understand the socio-economic experience of a community, as can the Income levels as measured by World bank list of economies[@IARC_3]. 
 The areas are usually ranked and allocated to quintiles to be presented as categories describing the ranking.
@@
 
-Non-Contiguous cartograms succeed in maintaing the shape of the areas presented.
+Non-Contiguous cartograms succeed in maintaining the shape of the areas presented.
 Each area stays in a similar position to their location on a choropleth map.
@@
 Circular cartograms use the same simple shapes for every region represented, and size the shapes according to the statistic represented or the population for a base map.
-This familar shape may be more effective for understanding the spatial distribution than contiguous cartograms, as the 'nonsense' shapes used have 'no meaning' after distortions are applied @NISCC.
+This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms, as the 'nonsense' shapes used have 'no meaning' after distortions are applied @NISCC.
 To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighbouring areas over many iterations.
@@
 @TFYCC quotes the official definition of Value-Area Cartograms, the simplistic displays which represent each area as a single rectangle, sized according to the value of the statistic.
-This rectangluar display also allows for tiling, where geographic neighbours also share rectangle borders.
+This rectangular display also allows for tiling, where geographic neighbours also share rectangle borders.
 
-# A critique of mapping methods
+# A discussion of mapping methods
 
@@
  - the ease or accuracy of information extraction for map users
+
+<!-- Choropleths, true topology, geographic distance, good for showing geographic importance -->
+Visualisation properties have been explored across several disease mapping displays.
+Choropleth displays utilise more traditional cartographic methods, this means they are usually true to the topography.
+Tobler's work highlights the need for consideration of the transformation of the globe to a 2D space.
+When there is a relationship between the cancer type being mapped and geographical features, a choropleth map would pair this information visually.
+These maps are best used when the geography is of importance to the data in the display.
+Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbours.
+
+<!-- Cartograms, sacrifice geography, focus on the outliers/ relationships in data space -->
+Pairing geography and cancer data was historically the work of epidemiologists, however it did not require them to be the map designer.
+This was expected as cartographers were the creators of geographic displays, and epidemiologists utilised the maps that had been prepared in advance.
+This disconnect did not allow the map to be prepared with the data in mind.
+Cartogram methods began exploring the concept of diagrammatic cartograms, displays that highlighted the importance of data, as well as the geography.
+This approach began a shift in the focus and responsibility of map creators.
+Cartogram displays began incorporating the statistics and population into the design of the display. 
+Shifting and sometimes sacrificing familiar boundaries to draw attention to outliers in the data space or population density.
+The difference between the familiar map display helps highlight the impact of the disease on communities.
+
+<!-- Alternatives, remove geographic shapes, keep spatial relationship, focus on the outliers/ relationships in population space -->
+By removing the unexpected boundaries and shapes, the spatial distribution is the primary concept communicated by the display.
+Areas that are geographically close will maintain connectedness in some way, but the population and the statistic will take precedence.
+These alternatives are especially helpful for data aggregations where administrative boundaries break populations into groups.
+The experience of each collective may be worth considering as their experiences may be similar in terms of the services and facilities they share.
+Population affects are highlighted by these displays, as this display allows a more equitable view of each community, and does not minimise those that operate on a smaller geographic scale.
 

### 2019-08-26 review_article.Rmd
Animation and Interactivity thoughts

@@
 To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
-It is the objectives of the investigator that will drive the choice of representation [@CPISACA].
-
-There are two keys presented by @SAMGIS to drive the choice of display: 
- - the properties of the visualisation, and 
- - the ease or accuracy of information extraction for map users
+Techonology has played an enormous role in increasing the opportunities for map makers.
+The computation and graphics power have made creation, alteration and interactivity possible.
+The internet has made communication of scientific data and results increasingly accessible.
+Creating maps of dieases now involves more decisions as these options have expanded and it is the objectives of the investigator that will drive the choices.
+@CPISACA and @SAMGIS have provided suggestions and comments to help map creators best communicate their health data and spatial analyses.
 
@@
 
-Recent developments in technology allowed interactive web atlases.
+Traditionally, epidemiologists as disease mappers would have understood the characteristics of a disease lifespan and movements.
+they would have also considered the interactions between geography, the environment and the people captured by a display.
+It is important for map creators to develop and then test the message being communicated, epidemiologists are able to do this with their expert knowledge.
+Static visualisations can help to aid the written results provided by experts, and pairing an explanation with a map or diagram can help to draw attention to a point, group or area of interest.
+They have the caveat of a need to provide enough information for the static displays to be understandable out of their original context.
 
->"Where control of the message is important, static maps will continue to be the most effective, although good tables, graphs, and explanatory text are still needed in order to ensure that different people will see the same thing in the maps" @CPISACA
+Recent developments in technology have created an opportunity to migrate atlases from paper displays in books to interactive web atlases.
+The medium chosen to communicate health data should always be driven by the message to communicate and the needs of the audience.
+This can be seen as a continuum, where the message is most important, static or animated graphics allw control over the display and interpretation.
+Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore.
 
-The intention of interactivity and animation is to allow users to answer their own questions.
+<!-- Animations -->
+@CPISACA provide weather maps as well developed and appropriate examples of spatial displays to communicate to the general public and encourage those creating disease maps to leverage the success of weather displays.
+The movement of a weather system will follow a forecasted path, the animation will communicate this as all map users can follow the path of the system across the geography over a specified period of time.
 
-<!--
-Using e-Delphi to Evaluate the Pennsylvania Cancer Atlas - explains PCA and reasons for interactivity.
--->
+<!-- Interactivity-->
+The move from static, print media displays toward interactivity allow users to develop maps to ask and answer their own questions.
+Developmets and adoptions of this technology has allowed map users more control of the message displayed.
+They are able to see the familiar shapes, zoom in to explore densely populated communities, or change the display completely to highlight the population space.
+@CPISACA provide three questions that are generally when using a disease map:
 
-<!--
-@ECGC ?? Highlight the value of animating contiguous to see changes over time, US can be recognisable but animation aides interpretation 
--->
+- What is the mortality rate in a certain area?
+- Are there geographic trends in the data, or regions of unusually high or low rates?
+- Is the lung cancer mortality pattern similar to the pattern of smoking prevalence shown in a companion map?
+
+A single map display historically created by epidemiologists, may not effectively answer all three at the same time.
+Especially when areas are small and hard to identify in a display.
+However, with the ability to animate, map creators can specify and answer these questions by directing the attention of a map user to the areas of interest.
+Animations can be used to control the message being communicated by capturing and directing attention of users, it will also ensure all users follow the same interpretive path, as their eyes are drawn across the map to the same comparison points.
+
+This oppportunity provides room to move away from the traditional approach of a single map for a single use, into a more explorative territory.
+Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
 
 # Acknowledgements
+
+We thank Dr Susanna Cramb (Spatial Modeller, Cancer Council Queensland) and Dr Peter Baade (Senior Research Fellow, Cancer Council Queensland) for providing the opportunity to cover this exploration of disease mapping methods in writing.
+This literature review would not be possible without the opportunity provided by Queesnland University of Technology, and Cancer Council Queensland, and the roles of Professor Kerrie Mengersen (Professor of Statistics, Science and Engineering Faculty, QUT) and Dr Earl Duncan (Research Associate at ARC Centre of Excellence for Mathematical & Statistical Frontiers, QUT) in contributing to the Australian Cancer Atlas and supervision of Stephanie's Research Masters.
+
+It was in the development of this online cancer atlas that methods for disease map displays, and visual communication strategies were explored.
+We are thankful for the opportunity to write about these visualisation displays and the situations in which they are appropriate.
 

### 2019-08-27 review_article.Rmd
Small edits

@@
 A choropleth map is used to display the characteristics of a spatial relationship of measurements. 
-A choropleth map is a true map of the topology, constructed for visual inspection of spatial patterns on a familiar geographic form. 
+They are true maps of the topology, constructed for visual inspection of spatial patterns on a familiar geographic form.
+Classed choropleth maps break a continuous variable into a small set of categories, often quintiles, and use a gradual colour scheme to shade all areas that fall in the same category [@CPISACA].
 \ref{fig:choro} 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
@@
 
->"designing a map tailored to precise goals [is] easier than forcing a single map to accommodate diverse objectives" @CPISACA
-
-Waldo Tobler [-@TVSSS] explores many graphical techniques, and suggests there are particular methods for particular purposes.
 To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
+To discover whether the display is appropriate, consider the  
 Techonology has played an enormous role in increasing the opportunities for map makers.

### 2019-08-27 review_article.Rmd
rearrangement of sections
allow changes to be made to the figures

@@
 
-# Disease mapping methods
-
-## Current best practice map displays for cancer data
+# Map displays for disease data
 
@@
 
-```{r choropleth-grid-create, message=FALSE, eval = FALSE}
+```{r choropleth-grid-create, message=FALSE}
 library(grid)
@@
 
-## Publicly available atlases using choropleth methods
+# Publicly available atlases
 
@@
 
-## Cartograms as an alternative display
+# Alternative displays
+
+## Cartograms
 
@@
 
-### Alternative Displays
+### Alternatives
 
@@
 
-# A discussion of mapping methods
+# Comparison of mapping methods
 

### 2019-08-29 review_article.Rmd
Reduce cancer map grid to 9 tiles
included examples of different resolution (global, national, state), low counts, divergent vs continuous scales, and rates vs relative risk.

@@
 
-plot1 <- png::readPNG('cancer_map_images/choropleths/1.png')
-plot2 <- png::readPNG('cancer_map_images/choropleths/14.choropleth1.png')
-plot3 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
-plot4 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
-plot5 <- png::readPNG('cancer_map_images/choropleths/5.choropleth.png')
-plot6 <- png::readPNG('cancer_map_images/choropleths/6.choropleth.png')
-plot6 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
-plot7 <- png::readPNG('cancer_map_images/choropleths/8.choropleth.png')
-plot8 <- png::readPNG('cancer_map_images/choropleths/20.choropleth1.png')
-plot9 <- png::readPNG('cancer_map_images/choropleths/21.choropleth2.png')
-plot10 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
-plot11 <- png::readPNG('cancer_map_images/choropleths/24.choropleth.png')
-plot12 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
-plot13 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
-plot14 <- png::readPNG('cancer_map_images/choropleths/27.choropleth.png')
-plot15 <- png::readPNG('cancer_map_images/choropleths/29.choropleth.png')
-plot16 <- png::readPNG('cancer_map_images/choropleths/30.choropleth.png')
+plot1 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
+plot2 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
+plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
+plot4 <- png::readPNG('cancer_map_images/choropleths/20.choropleth1.png')
+plot5 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
+plot6 <- png::readPNG('cancer_map_images/choropleths/14.choropleth2.png')
+plot7 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
+plot8 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
+plot9 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
+plot10 <- png::readPNG('cancer_map_images/choropleths/30.choropleth.png')
+plot11 <- png::readPNG('cancer_map_images/choropleths/29.choropleth.png')
+plot12 <- png::readPNG('cancer_map_images/choropleths/1.png')
+plot13 <- png::readPNG('cancer_map_images/choropleths/24.choropleth.png')
+plot14 <- png::readPNG('cancer_map_images/choropleths/6.choropleth.png')
+plot15 <- png::readPNG('cancer_map_images/choropleths/27.choropleth.png')
+plot16 <- png::readPNG('cancer_map_images/choropleths/8.choropleth.png')
+plot17 <- png::readPNG('cancer_map_images/choropleths/5.choropleth.png')
 
-choropleth_grid <- gridExtra::grid.arrange(grid::rasterGrob(plot1),grid::rasterGrob(plot2), grid::rasterGrob(plot3), 
-                        grid::rasterGrob(plot4), grid::rasterGrob(plot5), grid::rasterGrob(plot6),
-                        grid::rasterGrob(plot7), grid::rasterGrob(plot8), grid::rasterGrob(plot9),
-                        grid::rasterGrob(plot10),grid::rasterGrob(plot11), grid::rasterGrob(plot12), 
-                        grid::rasterGrob(plot13), grid::rasterGrob(plot14), grid::rasterGrob(plot15), 
-                        grid::rasterGrob(plot16), nrow=4)
+choropleth_grid <- gridExtra::grid.arrange(grid::rasterGrob(plot1),
+                  grid::rasterGrob(plot2), 
+                  grid::rasterGrob(plot3),                
+                  grid::rasterGrob(plot4), 
+                  grid::rasterGrob(plot5), 
+                  grid::rasterGrob(plot6),
+                  grid::rasterGrob(plot7), 
+                  grid::rasterGrob(plot8), 
+                  grid::rasterGrob(plot9), ncol=3)
 
@@
 ```{r choropleth-grid-print, echo = FALSE, results = "asis"}
-knitr::include_graphics("figures/choropleth_grid.png")
+#knitr::include_graphics("figures/choropleth_grid.png")
 ```

### 2019-08-29 review_article.Rmd
begin adding captions to figure 2

@@
 
-```{r choropleth-grid-create, message=FALSE}
+```{r choropleth-grid-create, message=FALSE, fig.cap = "Public facing choropleth cancer maps.
+[Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx), "}
 library(grid)
@@
 
-choropleth_grid <- gridExtra::grid.arrange(grid::rasterGrob(plot1),
+choropleth_grid <- gridExtra::grid.arrange(
+                  grid::rasterGrob(plot1),
                   grid::rasterGrob(plot2), 

### 2019-08-30 review_article.Rmd
add fig caption to grid image. won't render in pdf

@@
 
-```{r choropleth-grid-create, message=FALSE, fig.cap = "Public facing choropleth cancer maps.
-[Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx), "}
+```{r choropleth-grid-create, message=FALSE, fig.caption = "Public facing choropleth cancer maps.
+From right to left and top to bottom: [1.
+Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx),  [2.
+Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), [3.
+Atlas of Cancer in QLD - Lung cancer in males](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/)"}
+
+# fig.cap = "Public facing choropleth cancer maps.
+From right to left and top to bottom: [1.
+Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx),  [2.
+Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), [3.
+Atlas of Cancer in QLD - Lung cancer in males](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/),(http://www.cancer-rates.info/naaccr/), [5.
+Atlas of Childhood Cancer in Ontario](http://www.pogo.ca/research-data/pogo-atlas/), 6.
+An Atlas of Cancer in South Australia](https://www.cancersa.org.au/assets/images/pdfs/An%20Atlas%20of%20Cancer%20in%20South%20Australia%20-%20Full%20Report.pdf ), [7.
+The Environment and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/), [8.
+Map of Cancer Mortality Rates in Spain]() "}
+
 library(grid)
@@
 ```{r choropleth-grid-print, echo = FALSE, results = "asis"}
+
 #knitr::include_graphics("figures/choropleth_grid.png")

### 2019-08-30 review_article.Rmd
add links for all maps to figure caption

--- 2019-08-29 review_article.Rmd
@@
 ```{r choropleth-grid-create, message=FALSE, fig.cap = "Public facing choropleth cancer maps.
-[Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx), "}
+From right to left and top to bottom: [a.
+Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx)  (Incidence rates per 100,000 for all cancers, male and female),  [b.
+Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), [c.
+Atlas of Cancer in QLD - Relative incidence ration, Lung cancer in males](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/), [d.
+CINA+ Online Cancer in North America](http://www.cancer-rates.info/naaccr/) (Incidence rates of all invasive cancers per 100,000, male and female), [e.
+Atlas of Childhood Cancer in Ontario](http://www.pogo.ca/research-data/pogo-atlas/) (Incidence rate of childhood cancer per 100,000, men and women), [f.
+An Atlas of Cancer in South Australia](https://www.cancersa.org.au/assets/images/pdfs/An%20Atlas%20of%20Cancer%20in%20South%20Australia%20-%20Full%20Report.pdf) (Incidence rate per 100,00, lung cancer in females), [g.
+The Environment and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/) (Lung cancer in females), [h.
+Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) (Relative risk, Lung cancer, men vs women), [i.
+United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/) (Incidence rate per 100,000, all cancer types, men and women"}
+
 library(grid)

### 2019-08-30 review_article.Rmd
edited one 2 words in the intro

@@
 The distribution of the statistic must be visible, it is not enough for areas on maps to be recognisable.
-It is reasonable to explore views that enhance the communication of the cancer statistics, especially as it is the people that are of interest, not the land they live on. 
-This has spurred innovations over the previous centuries to enhance maps to effectively communicate cancer outcomes, and health outcomes more broadly.
+It is important to explore map views and designs that enhance the communication of the cancer statistics, especially as it is the populations that live on the land that are of interest, not the land itself.
+This has spurred innovations over the previous centuries to enhance maps to effectively communicate both cancer outcomes and health outcomes more broadly.
 

### 2019-08-30 review_article.Rmd
small edits to first paragraph of "Supplementary statistics" section

@@
 
-Cancer maps are effective communication tools for a general public or non-expert audience. They are commonly used in the public domain to communicate results of sophisticated statistical analyses.
-The heavy use of choropleth maps within the research literature is also reflected in the types of maps that are found in the public domain.
-A grey literature review of cancer atlases available in the public domain<!--(ref Jessie)--> identified 33 cancer atlases published on the internet between January 2010 and November 2015.
-These choropleth maps were mostly published by non-commercial organisations, including not-for-profits, government, research organisations, advocacy groups or government funded partnerships.
-Only one map was published by a commercial entity.
+Cancer maps are effective tools for communicating to wide range of audiences, including the general public and others not trained in statistical analyses.
+These visualisation enable these non-expert audiences to access the outputs of sophisticated statistical analyses.
+Cruickshank's (1947) as cited by @DMAHP, discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
+Overwhelmingly cancer maps utilised to communicate to the public and other non-expert audiences are choropleths.
+A grey literature review of cancer atlases available in the public domain [@roberts2019communication] identified 33 cancer atlases published on the internet between January 2010 and November 2015, all of which were choropleths.
+These maps were published by non-commercial organisations, including not-for-profits, government, research organisations, advocacy groups or government funded partnerships.Only one map was published by a commercial entity.
+The use of choropleths within the public domain mirrors the heavy use of choropleth maps within the research literature, discussed above.
 
 <!-- geographic coverage-->
+<!-- is this that interesting to the reader-->
 The cancer atlases covered geographies from all around the world, but only four were global atlases.
-Most focussed on single nations, the United States was considered by eleven atlases, the United Kingdom by seven, followed by three of Australia, two of Canada, and one of each from Switzerland, Germany, Norway.
+Most focused on single nations, the United States was considered by eleven atlases, the United Kingdom by seven, followed by three of Australia, two of Canada, and one of each from Switzerland, Germany, Norway.
 One atlas covered the European Union.
@@
 The states or counties/regions covered were South Australia (AUS), Queensland (AUS), Ontario (CAN), Valencia (Spain), Pennsylvania county Massachusetts (US), New Hampshire (US), Cape Cod (US), Missouri (US), Florida (US), New York State (US) and Arizona (US).
+Figure 2.a.b and c show a global, national and state choropleth respectively.
 
-## Supporting statistics in public facing atlases
+<!-- While there is significant variation across these maps in terms of the measure represented, the scale of the map, the colour choices, the legend design and supplementary/supporting statistics, the mapping approach has not changes and almost exclusively only choropleth maps are used.
+Unlike innovation in spatial statistical methods, interactity, data collection and record keeping little work has been done to implement different mapping approaches -->
 
-The primary display of an atlas is often the choropleth map.
-Cruickshank's (1947) as cited by @DMAHP, discusses using visuals as 'formal statistical assessment of the spatial pattern'. 
-Atlases can also utilise bar charts, tables and scatter-plots as useful alternative displays. 
-These displays can present the cancer incidence and mortality rates, as well as the supporting statistics of the atlas. 
-When presenting cancer atlases, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses.'
+## Supplementary statistics in public facing atlases
+
+While cancer maps are powerful visualisations that summarise complex statistical analyses, the statistics represented in these maps cannot tell the entire story.
+Supplementary graphs and plots are often included that add more depth and more information to the map.
+Maps can also utilise bar charts, tables and scatter-plots as useful additional displays.
+These displays can provide alternative views of the cancer statistics, as well as the supporting statistical outputs such as error, confidence intervals, distributions, sample/population sizes, standard deviation, as well as others..
+When presenting cancer maps, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses.' These supplementary statistics aim to enable this validation.
 @CPISACA suggests additional materials such as 'good tables, graphs, and explanatory text' support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers.
+
+<!-- insert figure with 3 examples of supplementary graphs-->
+
+**Interactivity**
+
 The interactivity of modern maps enables supplementary information to be incorporated without cluttering the screen, such as in a tool tip feature. 

### 2019-08-30 review_article.Rmd
end of day commit

@@
 The interactivity of modern maps enables supplementary information to be incorporated without cluttering the screen, such as in a tool tip feature.
-The atlases provided additional information to help users understand and interpret the spatial distribution presented. The use of these supports were found in a variety of online map sources.
-These additional pieces of information can help to validate, explain or explore relationships between cancer incidence and survival rates.
+The use of these supports were found in a variety of online map sources.
+These interactive features provide additional information to help users understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
+The interactive features of the publically avaialble maps identified by @roberts2019communication were.......
 

### 2019-09-02 review_article.Rmd
Map methods table

--- 2019-08-30 review_article.Rmd
@@
 
-To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
-To discover whether the display is appropriate, consider the  
-Techonology has played an enormous role in increasing the opportunities for map makers.
-The computation and graphics power have made creation, alteration and interactivity possible.
-The internet has made communication of scientific data and results increasingly accessible.
-Creating maps of dieases now involves more decisions as these options have expanded and it is the objectives of the investigator that will drive the choices.
-@CPISACA and @SAMGIS have provided suggestions and comments to help map creators best communicate their health data and spatial analyses.
-
 <!-- Choropleths, true topology, geographic distance, good for showing geographic importance -->
 Visualisation properties have been explored across several disease mapping displays.
-Choropleth displays utilise more traditional cartographic methods, this means they are usually true to the topography.
-Tobler's work highlights the need for consideration of the transformation of the globe to a 2D space.
+To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
+Choropleth displays utilise more traditional cartographic methods, they are usually true to the topography, displaying familiar boundaries of countries, states or administrative areas.
+Choropleth methods require a decision to be made about the projection of the display, in this case aim to find a map projection that gives the least distorted representation of the geography [@UMPT].
 When there is a relationship between the cancer type being mapped and geographical features, a choropleth map would pair this information visually.
@@
 Population affects are highlighted by these displays, as this display allows a more equitable view of each community, and does not minimise those that operate on a smaller geographic scale.
+
+Table: \label{tab:methods} Mapping methods used to display cancer statistics
+
+|  | Choropleth | Contiguous Cartogram | Non-contiguous Cartogram | Dorling | Tilegrams |
+|:---------------------------------------------------:|------------|----------------------|--------------------------|---------|-----------|
+| Feature Preservation: |  |  |  |  |  |
+| Preserves shapes | Y | N | Y | N | N |
+| Preserves neighbours | Y | Y | Y | S | S |
+| Uniform use of shape | N | N | N | Y | Y |
+| Statistical transformation of the base map: |  |  |  |  |  |
+| Used for Bi-variate maps eg. population & statistic | N | Y | Y | Y | Y |
+|  |  |  |  |  |  |
+
+Creating maps of dieases now involves more decisions to be made by map makers, rather than cartographers. 
+Techonology has played an enormous role in increasing the opportunities for map makers.
+The computation and graphics power have made creation, alteration and interactivity possible.
+as these options have expanded and it is the objectives of the investigator that will drive the choices.
+@CPISACA and @SAMGIS have provided suggestions and comments to help map creators best communicate their health data and spatial analyses.
 

### 2019-09-02 review_article.Rmd
Add tilegram

@@
 
-Daniel Dorling presented an alternative to highlight the spatial distribution and neighbourhood relationships without complex distortions of borders and boundaries.
-This approach opposes preserving the intricate shape details.
-This concept is founded in the simple question put forward by Daniel Dorling -@ACTUC:
+Daniel Dorling presented an alternative display engineere to highlight the spatial distribution and neighbourhood relationships without complex distortions of borders and boundaries.
+This approach opposes preserving the intricate shape details and is founded in the simple question put forward by Daniel Dorling -@ACTUC:
 
@@
 
-He answers this with his implementation of maps created with 'the simplest of all shapes'. He acknowledged the sophistication of contiguous cartograms but critiqued their 'very complex shapes'.
+He acknowledged the sophistication of contiguous cartograms but critiqued their 'very complex shapes', he answers this with his implementation of maps created using 'the simplest of all shapes'. 
 Circular cartograms use the same simple shapes for every region represented, and size the shapes according to the statistic represented or the population for a base map.
@@
 This rectangular display also allows for tiling, where geographic neighbours also share rectangle borders.
+
+Square grids have been used by FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR. Where each area is represented by a square of the same dimensions.
+An example is provided in \ref{fig:ggtilegram}
+
+```{r ggtilegram, fig.cap = "A tilegram of the Unites States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been represented by a square, each square is the same size.
+The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.
+The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar..", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+knitr::include_graphics("figures/ggtilegram.png")
+```
 
@@
 |:---------------------------------------------------:|------------|----------------------|--------------------------|---------|-----------|
-| Feature Preservation: |  |  |  |  |  |
+| Feature preservation: |  |  |  |  |  |
 | Preserves shapes | Y | N | Y | N | N |
@@
 | Statistical transformation of the base map: |  |  |  |  |  |
-| Used for Bi-variate maps eg. population & statistic | N | Y | Y | Y | Y |
-|  |  |  |  |  |  |
+| Used for bi-variate maps eg. population & statistic | N | Y | Y | Y | Y |
 

### 2019-09-02 usa_plots.R

@@ -39,6 +39,7 @@ ggplot(cancer, aes(x = total_pop_15 )) + geom_density()
 title <- ""
   #"Average rate of incidence for lung and bronchus for females and males in the United States 2012-2016, \n where the size of each state has been adjusted for population"
 
+###############################################################################
 # Create a choropleth
 ggchoro <- ggplot(cancer) + 
   geom_sf(aes(fill = AgeAdjustedRate)) +
@@ -47,9 +48,10 @@ ggchoro <- ggplot(cancer) +
   coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
   theme_void() +
   theme(legend.position ="bottom")
-ggsave(filename = "figures/ggchoro.png", device = "png", width = 12, height = 6)
+ggsave(filename = "figures/ggchoro.png", device = "png", dpi = 300,  width = 12, height = 6)
 ggchoro
 
+###############################################################################
 # Cartograms
 #install.packages("cartogram")
 library(cartogram)
@@ -63,9 +65,10 @@ ggcont <- ggplot(cont) +
   scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
   coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
   theme_void() +   theme(legend.position ="bottom")
-ggsave(filename = "figures/ggcont.png", device = "png", width = 12, height = 6)
+ggsave(filename = "figures/ggcont.png", device = "png", dpi = 300,  width = 12, height = 6)
 ggcont
 
+###############################################################################
 # Non - Contiguous Cartograms
 # Needs a scaling factor
 
@@ -85,10 +88,11 @@ ggncont <- ggplot(ncont) +
   scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
   coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
   theme_void() + theme(legend.position ="bottom")
[…]

### 2019-09-03 review_article.Rmd
Figure captions

@@
     toc_depth: 3
+    fig_caption: true
 fontsize: 11pt
 link-citations: yes
-fig_caption: yes
 biblio-style: vancouver
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
 library(ggplot2)
+library(grid)
+library(png)
+library(gridExtra)
 ```
@@
 The state of Kentucky is high, the surrounding states also have high rates.
-There is a pattern of average higher rates toward the East Coast and lower rates toward the West Coast.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+There is a pattern of average higher rates toward the East Coast and lower rates toward the West Coast.", message=FALSE, warning=FALSE, out.width='80%'}
 knitr::include_graphics("figures/ggchoro.png")
@@
 
-library(grid)
-library(png)
-library(gridExtra)
-
 plot1 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
@@
 plot9 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
-plot10 <- png::readPNG('cancer_map_images/choropleths/30.choropleth.png')
-plot11 <- png::readPNG('cancer_map_images/choropleths/29.choropleth.png')
-plot12 <- png::readPNG('cancer_map_images/choropleths/1.png')
-plot13 <- png::readPNG('cancer_map_images/choropleths/24.choropleth.png')
-plot14 <- png::readPNG('cancer_map_images/choropleths/6.choropleth.png')
-plot15 <- png::readPNG('cancer_map_images/choropleths/27.choropleth.png')
-plot16 <- png::readPNG('cancer_map_images/choropleths/8.choropleth.png')
-plot17 <- png::readPNG('cancer_map_images/choropleths/5.choropleth.png')
 
@@
 
-```{r choropleth-grid-print, echo = FALSE, results = "asis"}
-
-#knitr::include_graphics("figures/choropleth_grid.png")
-```
-
 # Publicly available atlases
@@
 Cancer maps are effective tools for communicating to wide range of audiences, including the general public and others not trained in statistical analyses.
-These visualisation enable these non-expert audiences to access the outputs of sophisticated statistical analyses.
+These visualisation enable thse non-expert audiences to access the outputs of sophisticated statistical analyses.
 Cruickshank's (1947) as cited by @DMAHP, discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
@@
 The use of choropleths within the public domain mirrors the heavy use of choropleth maps within the research literature, discussed above.
-
-<!-- geographic coverage-->
-<!-- is this that interesting to the reader-->
-The cancer atlases covered geographies from all around the world, but only four were global atlases.
+The cancer atlases covered geographies from all around the world.
+One atlas covered the European Union.
 Most focused on single nations, the United States was considered by eleven atlases, the United Kingdom by seven, followed by three of Australia, two of Canada, and one of each from Switzerland, Germany, Norway.
-One atlas covered the European Union.
+Figure 2.a.b and c show a global, national and state choropleth respectively. 
 Not all maps had a national focus and ten covered a region or state rather than an entire nation.
 The states or counties/regions covered were South Australia (AUS), Queensland (AUS), Ontario (CAN), Valencia (Spain), Pennsylvania county Massachusetts (US), New Hampshire (US), Cape Cod (US), Missouri (US), Florida (US), New York State (US) and Arizona (US).
-Figure 2.a.b and c show a global, national and state choropleth respectively.
 
@@
 
-<!-- insert figure with 3 examples of supplementary graphs-->
+```{r supp-grid-create, fig.cap = "Interactive controls of displays in public facing choropleth cancer maps.
+From left to right and top to bottom: [a.
+Globocan 2018: Cancer Today](http://globocan.iarc.fr/Pages/Map.aspx)(Controls for indicator, sex, age groups, continents, and cancer types) , [b.
+Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/) (Controls for indicator, age, year, and geographical areas), [c.
+Global Cancer Map](http://globalcancermap.com/) (Controls for indicator, and cancer type), [d.
+United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html) (Controls for demographics, trends, geography, indicators, risk factors, sex, cancer type, and year)."}
+i3 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png')
+i15 <- png::readPNG('cancer_map_images/interactivity/15.interactivity.png')
+i4 <- png::readPNG('cancer_map_images/interactivity/4.interactivity.png')
+i5 <- png::readPNG('cancer_map_images/interactivity/5.interactivity.png')
+i7 <- png::readPNG('cancer_map_images/interactivity/7.interactivity.png')
+
+lay <- rbind(c(1,1,2,2,2,2),
+             c(1,1,2,2,2,2),
+             c(1,1,2,2,2,2),
+             c(1,1,3,3,3,3),
+             c(1,1,4,4,4,4),
+             c(1,1,5,5,5,5))
+
+supp_grid <- gridExtra::grid.arrange(
+  grid::rasterGrob(i3), grid::rasterGrob(i15),
+  grid::rasterGrob(i5), grid::rasterGrob(i7), grid::rasterGrob(i4), layout_matrix = lay)
+
+ggsave(filename = "figures/supp_grid.png", plot = supp_grid, device = "png", width = 6, height = 5, dpi = 300)
+```
 
@@
 The map projections alter the shapes and angles of the boundaries of each state.
-Maps a and b are similar in their straight edges, unlike maps c and d which curve on the northern United States border.
-", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+Maps a and b are similar in their straight edges, unlike maps c and d which curve on the northern United States border.", message=FALSE, warning=FALSE, out.width='80%'}
 knitr::include_graphics("figures/ggchoroCRS.png")
@@
 The state of California has become much larger due to it's large population density.
-This draws attention to the densely populated North East region, and detracts from the Mid West.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+This draws attention to the densely populated North East region, and detracts from the Mid West.", message=FALSE, warning=FALSE, out.width='80%'}
 knitr::include_graphics("figures/ggcont.png")
@@
 The North East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut.
-This draws attention to the densely populated North East region, and the sparse Mid West.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+This draws attention to the densely populated North East region, and the sparse Mid West.", message=FALSE, warning=FALSE, out.width='80%'}
 knitr::include_graphics("figures/ggncont.png")
@@
 The North East states remain closer to their neighbours, and may be displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.", message=FALSE, warning=FALSE, out.width='80%'}
 knitr::include_graphics("figures/ggdorl.png")
@@
 Square grids have been used by FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR. Where each area is represented by a square of the same dimensions.
-An example is provided in \ref{fig:ggtilegram}
+Figure \ref{fig:ggtilegram} shows a square tilegram representing each state using a single square.
 
@@
 The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar..", message=FALSE, warning=FALSE, fig.align='center', out.width='80%'}
+The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.", message=FALSE, warning=FALSE, out.width='80%'}
 knitr::include_graphics("figures/ggtilegram.png")
@@
 
-|  | Choropleth | Contiguous Cartogram | Non-contiguous Cartogram | Dorling | Tilegrams |
-|:---------------------------------------------------:|------------|----------------------|--------------------------|---------|-----------|
+|  | Choropleth | Contiguous | Non-contiguous | Dorling | Tilegrams |
+|:-----------:|------------|-----------|---------------|---------|-----------|
 | Feature preservation: |  |  |  |  |  |
@@
 | Uniform use of shape | N | N | N | Y | Y |
-| Statistical transformation of the base map: |  |  |  |  |  |
-| Used for bi-variate maps eg. population & statistic | N | Y | Y | Y | Y |
+| Bivariate display* | N | Y | Y | Y | Y |
+* A bivariate display refers to a map that allows two variables to be plotted. One variable via the size of the area, and another by filling the area with a colour.
 
@@
 It was in the development of this online cancer atlas that methods for disease map displays, and visual communication strategies were explored.
-We are thankful for the opportunity to write about these visualisation displays and the situations in which they are appropriate.
+We are thankful for the opportunity to write about these visualisations and the situations in which they are appropriate.
 

### 2019-09-09 review_article.Rmd
Figures using cowplots

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE, fig.width = 6, fig.height = 5)
 library(ggplot2)
@@
 
-```{r choropleth-grid-create, message=FALSE, fig.cap = "Public facing choropleth cancer maps.
-From right to left and top to bottom: [a.
-Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx)  (Incidence rates per 100,000 for all cancers, male and female),  [b.
+```{r choropleth-grid-create, message=FALSE, fig.cap = "Public facing choropleth cancer maps.[a.
+The Environment and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/) (Lung cancer in females)  (Incidence rates per 100,000 for all cancers, male and female),  [b.
 Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), [c.
 Atlas of Cancer in QLD - Relative incidence ration, Lung cancer in males](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/), [d.
-CINA+ Online Cancer in North America](http://www.cancer-rates.info/naaccr/) (Incidence rates of all invasive cancers per 100,000, male and female), [e.
-Atlas of Childhood Cancer in Ontario](http://www.pogo.ca/research-data/pogo-atlas/) (Incidence rate of childhood cancer per 100,000, men and women), [f.
-An Atlas of Cancer in South Australia](https://www.cancersa.org.au/assets/images/pdfs/An%20Atlas%20of%20Cancer%20in%20South%20Australia%20-%20Full%20Report.pdf) (Incidence rate per 100,00, lung cancer in females), [g.
-The Environment and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/) (Lung cancer in females), [h.
-Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) (Relative risk, Lung cancer, men vs women), [i.
-United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/) (Incidence rate per 100,000, all cancer types, men and women"}
+Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx), [e.
+United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/) (Incidence rate per 100,000, all cancer types, men and women), [f.
+Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) (Relative risk, Lung cancer, men vs women), [g.
+An Atlas of Cancer in South Australia](https://www.cancersa.org.au/assets/images/pdfs/An%20Atlas%20of%20Cancer%20in%20South%20Australia%20-%20Full%20Report.pdf) (Incidence rate per 100,000, lung cancer in females) "}
+library(cowplot)
 
-plot1 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
+# Import the images
+plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
 plot2 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
 plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
-plot4 <- png::readPNG('cancer_map_images/choropleths/20.choropleth1.png')
-plot5 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
-plot6 <- png::readPNG('cancer_map_images/choropleths/14.choropleth2.png')
-plot7 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
-plot8 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
-plot9 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
+plot4 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png') 
+plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
+plot6 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
+plot7 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
 
-choropleth_grid <- gridExtra::grid.arrange(
-                  grid::rasterGrob(plot1),
-                  grid::rasterGrob(plot2), 
-                  grid::rasterGrob(plot3),                
-                  grid::rasterGrob(plot4), 
-                  grid::rasterGrob(plot5), 
-                  grid::rasterGrob(plot6),
-                  grid::rasterGrob(plot7), 
-                  grid::rasterGrob(plot8), 
-                  grid::rasterGrob(plot9), ncol=3)
+ggdraw() +
+  draw_plot(rasterGrob(plot1), 0, .75, 0.33, .25) +
+  draw_plot(rasterGrob(plot2), 0.33, 0.75, 0.33, .25) +
+  draw_plot(rasterGrob(plot3), 0.66, 0.5, .33, .5) +
+  draw_plot(rasterGrob(plot4), 0, 0.5, .33, .25) +
+  draw_plot(rasterGrob(plot5), 0.33, 0.5, .33, .25) +
+  draw_plot(rasterGrob(plot6), 0, 0, .66, .5) +
+  draw_plot(rasterGrob(plot7), .66, 0, .33, .5) +
+  draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
+    c(0.015, 0.33, 0.63, 0.015, 0.33, 0.015, 0.63), 
+    c(0.99, 0.99, 0.99, 0.74, 0.74, 0.49, 0.49), size = 15)
 
-ggsave(filename = "figures/choropleth_grid.png", plot = choropleth_grid, device = "png", width = 6, height = 5, dpi = 300)
+ggsave(filename = "figures/choropleth_grid.png", 
+  device = "png", width = 6, height = 5, dpi = 300)
 ```
@@
 
-While cancer maps are powerful visualisations that summarise complex statistical analyses, the statistics represented in these maps cannot tell the entire story.
-Supplementary graphs and plots are often included that add more depth and more information to the map.
-Maps can also utilise bar charts, tables and scatter-plots as useful additional displays.
-These displays can provide alternative views of the cancer statistics, as well as the supporting statistical outputs such as error, confidence intervals, distributions, sample/population sizes, standard deviation, as well as others..
-When presenting cancer maps, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses.' These supplementary statistics aim to enable this validation.
+Cancer maps are powerful visualisations that summarise complex statistical analyses, however the statistics represented in these maps cannot tell the entire story.
+Supplementary graphs and plots are often included to add more depth and information to the map.
 @CPISACA suggests additional materials such as 'good tables, graphs, and explanatory text' support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers.
+There are many visualisations used for displays of statistical summaries, these may be dot plots, bar plots, box plots, timesseries plots, cumulative distribution plots, scatter plots, Q-Q plots. 
+These additional displays of the cancer distribution can provide alternative views of the cancer statistics, as well as the supporting statistics including error, confidence intervals, distributions, sample or population sizes, standard deviation and other measures.
+When presenting cancer maps, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses', the supplementary statistics help for this validation.
 
 ```{r supp-grid-create, fig.cap = "Interactive controls of displays in public facing choropleth cancer maps.
-From left to right and top to bottom: [a.
+[a.
 Globocan 2018: Cancer Today](http://globocan.iarc.fr/Pages/Map.aspx)(Controls for indicator, sex, age groups, continents, and cancer types) , [b.
 Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/) (Controls for indicator, age, year, and geographical areas), [c.
-Global Cancer Map](http://globalcancermap.com/) (Controls for indicator, and cancer type), [d.
+The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M) (Controls for predetermined combinations of cancer type or risk factors, mortality or incidence, and gender) [d.
+Global Cancer Map](http://globalcancermap.com/) (Controls for indicator, and cancer type), [e.
 United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html) (Controls for demographics, trends, geography, indicators, risk factors, sex, cancer type, and year)."}
-i3 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png')
-i15 <- png::readPNG('cancer_map_images/interactivity/15.interactivity.png')
-i4 <- png::readPNG('cancer_map_images/interactivity/4.interactivity.png')
-i5 <- png::readPNG('cancer_map_images/interactivity/5.interactivity.png')
-i7 <- png::readPNG('cancer_map_images/interactivity/7.interactivity.png')
+i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png')
+i2 <- png::readPNG('cancer_map_images/interactivity/15.interactivity.png')
+i3 <- png::readPNG('cancer_map_images/interactivity/4.interactivity.png')
+i4 <- png::readPNG('cancer_map_images/interactivity/5.interactivity.png')
+i5 <- png::readPNG('cancer_map_images/interactivity/7.interactivity.png')
 
-lay <- rbind(c(1,1,2,2,2,2),
-             c(1,1,2,2,2,2),
-             c(1,1,2,2,2,2),
-             c(1,1,3,3,3,3),
-             c(1,1,4,4,4,4),
-             c(1,1,5,5,5,5))
+ggdraw() +
+  draw_plot(rasterGrob(i1), 0,  0.13,   .28,  .83) + #a
+  draw_plot(rasterGrob(i2), 0.32, 0.56, .68, .4) + #b
+  draw_plot(rasterGrob(i3), 0.32, 0.41, .68, .13) + #c
+  draw_plot(rasterGrob(i4), 0.32, 0.14, .68, .25) + #d
+  draw_plot(rasterGrob(i5), 0, 0, 1, .13) + #e
+  draw_plot_label(c("a", "b", "c", "d", "e"), 
+    c(0.01, 0.3, 0.3, 0.3, 0.01), 
+    c(0.99, 0.99, 0.59, 0.40, 0.19), size = 15)
 
-supp_grid <- gridExtra::grid.arrange(
-  grid::rasterGrob(i3), grid::rasterGrob(i15),
-  grid::rasterGrob(i5), grid::rasterGrob(i7), grid::rasterGrob(i4), layout_matrix = lay)
-
-ggsave(filename = "figures/supp_grid.png", plot = supp_grid, device = "png", width = 6, height = 5, dpi = 300)
+ggsave(filename = "figures/supp_grid.png", device = "png", 
+  width = 6, height = 5, dpi = 300)
 ```
 
-**Interactivity**
-
-The interactivity of modern maps enables supplementary information to be incorporated without cluttering the screen, such as in a tool tip feature.
-The use of these supports were found in a variety of online map sources.
-These interactive features provide additional information to help users understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-The interactive features of the publically avaialble maps identified by @roberts2019communication were.......
+The interactivity of modern mapping methods enable supplementary information to be incorporated without cluttering the screen.
+Tool tip features and drop down menus allow for user exploration, small screens are not a problem with the flexibility provided by zooming and panning [@HTLWM].
+This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accesible' ways [@TNTEA].
+The use of these supports were found in a many online diease mapmap sources.
+These interactive features provide an opportunity for users to explore additional information to help users understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
+This directly enables epidemiologists to explore spatial distributions of cancer statistics.
+The interactive features of the publically available maps identified by @roberts2019communication include exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
 
@@
 Socio-economic indicators can explain how the experience of cancer prevalence varies for various members of a society.
-These indicators include unemployment rates, poverty rates and percentages, remoteness and education levels achieved though only a few atlases also explored the impact of rurality on cancer rates. 
+These indicators include unemployment rates, poverty rates, remoteness and education levels achieved though, only a few atlases also explored the impact of rurality on cancer rates. 
+These rates may also be explored as percentages above or below the mean or median value for the set of spatial areas.
 The Human Development Index can be used to understand the socio-economic experience of a community, as can the Income levels as measured by World bank list of economies[@IARC_3]. 
@@
 <!-- Tobler map projections-->
-Choropleths may be considered true topological maps, however, if the land mass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D.
+Choropleths may be considered true topological maps, however, if the land mass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D [@HTLWM].
 The amount of distortion is related to the distance covered by the landmass displayed @GAMP.
@@
 Selecting a display can prevent misinterpretation of global statistics, as global maps face the challenge of equitable displays of land mass on maps [@RSCW].
-Maps that do not specify a projection can be considered to have some unknown projection. <!-- ??? cite tobler.-->
 If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed, beyond choosing a reasonable projection.
@@
 The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC]. 
+@HTLWM suggests that white lies may be employed to create useful displays and map creators have the ability to draw lines that may distort the geometry and suppress features and it is easy for the average person to disregard the impact of transformations used to create cartograms.
 Cartograms may be seen as an extension of map transformations and projections. The favourable distortion is proportional to a value other other than actual earth size area [@NAC].
@@
 
-Square grids have been used by FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR. Where each area is represented by a square of the same dimensions.
+A similar method, where each geographic area is represented using a square, tesselated to create a square grid.
+This method has been used by FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
+Each area is represented by a square of the same dimensions.
 Figure \ref{fig:ggtilegram} shows a square tilegram representing each state using a single square.
@@
 
-Traditionally, epidemiologists as disease mappers would have understood the characteristics of a disease lifespan and movements.
-they would have also considered the interactions between geography, the environment and the people captured by a display.
+Traditionally, epidemiologists as disease mappers would have considered the interactions between the spatial distribution of a diesease, the environment and the people captured by a display.
+Recent developments in technology have created an opportunity to migrate atlases from paper map collections to web atlases.
+The medium chosen to communicate health data should always be driven by the message to communicate and the needs of the audience.
+Where the message is most important, static or animated graphics allow control over the display and interpretation.
 It is important for map creators to develop and then test the message being communicated, epidemiologists are able to do this with their expert knowledge.
 Static visualisations can help to aid the written results provided by experts, and pairing an explanation with a map or diagram can help to draw attention to a point, group or area of interest.
-They have the caveat of a need to provide enough information for the static displays to be understandable out of their original context.
-
-Recent developments in technology have created an opportunity to migrate atlases from paper displays in books to interactive web atlases.
-The medium chosen to communicate health data should always be driven by the message to communicate and the needs of the audience.
-This can be seen as a continuum, where the message is most important, static or animated graphics allw control over the display and interpretation.
+Static maps do have the caveat of needing to provide enough information for the display to be understandable out of the original context.
+Animations do not tolerate or require user input, this makes them useful for directed descriptions of processes, such as directing attention to a particular region or time period [@HTLWM].
 Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore.
@@
 <!-- Animations -->
-@CPISACA provide weather maps as well developed and appropriate examples of spatial displays to communicate to the general public and encourage those creating disease maps to leverage the success of weather displays.
-The movement of a weather system will follow a forecasted path, the animation will communicate this as all map users can follow the path of the system across the geography over a specified period of time.
+@CPISACA provide weather maps as well developed and appropriate examples of spatial displays to communicate to the general public, and encourage those creating disease maps to leverage the success of weather displays.
+The movement of a weather system will follow a forecasted path, the animation will communicate this so that all map users can follow the path of the weather system across the geography over a specified period of time.
 
-<!-- Interactivity-->
 The move from static, print media displays toward interactivity allow users to develop maps to ask and answer their own questions.
 Developmets and adoptions of this technology has allowed map users more control of the message displayed.
-They are able to see the familiar shapes, zoom in to explore densely populated communities, or change the display completely to highlight the population space.
+Users are able to see the familiar shapes, zoom in to explore densely populated communities, or change the display completely to highlight the population space.
 @CPISACA provide three questions that are generally when using a disease map:
@@
 A single map display historically created by epidemiologists, may not effectively answer all three at the same time.
-Especially when areas are small and hard to identify in a display.
+This is especially true when areas are small and difficult to identify in a display.
 However, with the ability to animate, map creators can specify and answer these questions by directing the attention of a map user to the areas of interest.
@@
 
+<!-- Interactivity-->
 This oppportunity provides room to move away from the traditional approach of a single map for a single use, into a more explorative territory.

### 2019-09-09 usa_plots.R

@@ -20,6 +20,7 @@ library(spData)
 library(ggthemes)
 library(maptools)
 library(sf)
+library(cartogram)
 
 # read in data
 cancer <- read_csv("data/USCSlung.csv") %>% rename(NAME = Area)
@@ -53,8 +54,6 @@ ggchoro
 
 ###############################################################################
 # Cartograms
-#install.packages("cartogram")
-library(cartogram)
 
 # Contiguous Cartograms
 cont <- cartogram_cont(cancer,

### 2019-09-09 review_article.Rmd
Rectangular cartogram sources

@@
 ### Statistical uncertainty
-Statistical uncertainty is communicated through a Confidence (CI) interval, statistical significance levels, Boxplots, Distribution plots, and reporting the Sample Size and Standard Deviations.
 The additional statistics often include a measure of the statistical uncertainty of the values of the statistics presented in a choropleth.
-In the review of atlases in the public domain, atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map.
+In the review of atlases in the public domain, atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map [@@roberts2019communication].
 The maps considered used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
+Other methods involve providing adjacent maps or overlapping maps with symbols [@VSSDCUC].
 The maps employing uncertainty ranged from static pdfs or infographics, to interactive online resources.
+Communicating the uncertainty associated with the estimates through a confidence (CI) interval, statistical significance levels, boxplots, distribution plots, and reporting sample size and standard deviations.
 Close to half of the atlases identified (42%, n=14) included some measure of uncertainty. The most common measure used to represent uncertainty were credible or confidence intervals (CIs).
@@
 <!-- Denominators -->
-
+Event cartograms change the area of regions on a map depending on the amount of disease related events, but this does not consider the effects of land area and population [@VSSDCUC].
 The purposeful distortion of the map space, transformed according to population density, is beneficial when a uniform density of the map base is desired.
@@
 
-The groundwork for the simplistic approach had been laid in the mid 1930's be Raisz. 
+The groundwork for this approach had been laid in the mid 1930's by Raisz, and rectangular cartograms provide dramatic comparisons and are especially useful for correcting misconceptions communicated by geographic maps.
 @TFYCC quotes the official definition of Value-Area Cartograms, the simplistic displays which represent each area as a single rectangle, sized according to the value of the statistic.
-This rectangular display also allows for tiling, where geographic neighbours also share rectangle borders.
+This rectangular display also allows for tiling, where geographic neighbours placed in suitable relative positions also share borders, however contiguity may be sacrificed [@CDWCS].
+Rectangular cartograms allow for bivariate displays, population can be effectively communicated by the size of each rectangular, and a second variable can be communicated using colour [@ORC].
 
@@
 Figure \ref{fig:ggtilegram} shows a square tilegram representing each state using a single square.
+
+Recommended criteria to contrast mapping methods include average cartographic error, and maximum cartographic error, correct adjacencies, maximum aspect ratio, and suitable relative positions [@ORC].
+However, this does not consider the issues with actually producing rectangular cartograms.
+Algorithms for the creation of rectangular cartograms
 
@@
 This was expected as cartographers were the creators of geographic displays, and epidemiologists utilised the maps that had been prepared in advance.
-This disconnect did not allow the map to be prepared with the data in mind.
-Cartogram methods began exploring the concept of diagrammatic cartograms, displays that highlighted the importance of data, as well as the geography.
+Cartograms with the population variable used for base map distortion can be seen as density equalising maps, that reduce the visual prominence of low-population areas [@VSSDCUC].
 This approach began a shift in the focus and responsibility of map creators.

### 2019-09-10 review_article.Rmd
introduction revision

@@
 
-Cancer statistics directly represent and relate to the people living within geographical areas.
-Geospatial statistics have been presented on maps to communicate information to the general public for centuries as they utilise familiar geographic shapes.
+<!-- What data are we working with, why, what are the purposes -->
+Cancer statistics represent information on people, usually delivered as an aggregated value for a geopolitical area.
+Presenting statistics requires transforming individual observations into aggregations of communities as geographical units, in large part for privacy protection or political and policy purposes.
+The information could be as simple as counts per area (e.g.
+state, province, local government area, post/zip code).
+Counts alone is not sufficient on which to compare different areas, because the populations of different areas are likely different.
+In this case, the counts data needs to be merged with population data to appropriately calibrate it to incidence, for example, rate per 100000 people.
+This type of data is collected on a routine basis for public health purposes, and may be made available to the general public as a service to the community.
+The task, then, is to examine what are the usual ways to communicate cancer statistics, to the public, are there alternative approaches, and what are the pros and cons of these choices.
+
+<!-- Geospatial statistics have been presented on maps to communicate information to the general public for centuries as they utilise familiar geographic shapes.
 The mapping of diseases summarises spatial variations and can be considered a descriptive statistic of the relationship between a disease and the underlying geography [@SEMA].
@@
 It is important to explore map views and designs that enhance the communication of the cancer statistics, especially as it is the populations that live on the land that are of interest, not the land itself.
-This has spurred innovations over the previous centuries to enhance maps to effectively communicate both cancer outcomes and health outcomes more broadly.
+This has spurred innovations over the previous centuries to enhance maps to effectively communicate both cancer outcomes and health outcomes more broadly.-->
 
-The visualisation methods used to present cancer statistics will depend on the intended message and users of the map. 
-Presenting statistics as visualisations requires transforming individual observations into aggregations of communities as geographical units.
+<!-- Ways to visualise, whats our hypothesis -->
+A common approach to communicate cancer statistics, is to display statistics on a map.
+This is a disease map: the statistic is mapped to colour and the geographic region is filled with this colour.
+The viewer would then be able to examine the spatial distribution of the disease incidence, where there is a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hotspots of the disease.
 Visualising diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH].
 Disease maps help to present geographic patterns that may have been overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
-By providing a visual representation of cancer outcomes, geographic patterns of disease are able to be identified and effectively addressed.
+By providing a visual representation of cancer outcomes, geographic patterns of disease are able to be identified and effectively addressed with public healh policy and actions.
 @SE recognises one of the key challenges with mapping spatial patterns of disease is the design of visualisations.
 This paper addresses the visualisation techniques and their applications to cancer statistics. Highlighting the differences and historic use of these displays.
+
+<!-- Structure of the paper -->
+The paper is organised as follows.
+The next section describes the choropleth map which is the common approach to disease maps.
+Section XXX surveys atlases in use today.
+Section XXX describes an alternative display, the cartogram what may be useful for countries that have heterogeneously sized geographic units.
+The pros and cons of these approaches is discussed in Section XXX.
+Interactivity options are described in Section XXX, and the paper finishes with a discussion.
 

### 2019-09-10 review_article.Rmd
introduction revision tweaks

@@
 A common approach to communicate cancer statistics, is to display statistics on a map.
-This is a disease map: the statistic is mapped to colour and the geographic region is filled with this colour.
+This is a choropleth map: the statistic is mapped to colour and the geographic region is filled with this colour.
 The viewer would then be able to examine the spatial distribution of the disease incidence, where there is a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hotspots of the disease.
@@
 The pros and cons of these approaches is discussed in Section XXX.
-Interactivity options are described in Section XXX, and the paper finishes with a discussion.
+Disease maps are ore useful when made interactive, and common options are described in Section XXX, along with a discussion of benefits and disadvantages.
+The last section summarises the paper and discusses future directions.
 

### 2019-09-10 review_article.Rmd
acknowledgements

--- 2019-09-09 review_article.Rmd
@@
 @ACTUC suggests 'population distribution is often extremely uneven in former British colonies', this makes the distortion necessary [@CTTMB].
-When implementing a distortion of the geographical shape according to population, the resulting display is an area cartogram [@NAC], population-by-area cartogram [@TAAM], or iso-demographic map.
+When implementing a distortion of the geographical shape according to population, the resulting display is an area cartogram [@NAC], or population-by-area cartogram [@TAAM].
 
@@
 
+We would like to acknowledge Dr Earl Duncan (Research Associate at ARC Centre of Excellence for Mathematical & Statistical Frontiers, QUT).
+His time and effort given to edit and comment on this paper was invaluable.
+
 We thank Dr Susanna Cramb (Spatial Modeller, Cancer Council Queensland) and Dr Peter Baade (Senior Research Fellow, Cancer Council Queensland) for providing the opportunity to cover this exploration of disease mapping methods in writing.
-This literature review would not be possible without the opportunity provided by Queesnland University of Technology, and Cancer Council Queensland, and the roles of Professor Kerrie Mengersen (Professor of Statistics, Science and Engineering Faculty, QUT) and Dr Earl Duncan (Research Associate at ARC Centre of Excellence for Mathematical & Statistical Frontiers, QUT) in contributing to the Australian Cancer Atlas and supervision of Stephanie's Research Masters.
+This literature review would not be possible without the opportunity provided by Queesnland University of Technology, and Cancer Council Queensland, and the roles of Professor Kerrie Mengersen (Professor of Statistics, Science and Engineering Faculty, QUT) and Dr Earl Duncan in contributing to the Australian Cancer Atlas and supervision of Stephanie's Research Masters.
 

### 2019-09-10 review_article.Rmd
Australian cancer atlas animation

@@
 The Environment and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/) (Lung cancer in females)  (Incidence rates per 100,000 for all cancers, male and female), [b.
-Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), [c.
+Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx),  [c.
 Atlas of Cancer in QLD - Relative incidence ration, Lung cancer in males](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/), [d.
-Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx), [e.
+Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), [e.
 United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/) (Incidence rate per 100,000, all cancer types, men and women), [f.
@@
 plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
-plot2 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
+plot2 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png') 
 plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
-plot4 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png') 
+plot4 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
 plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
@@
   draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
-    c(0.015, 0.33, 0.63, 0.015, 0.33, 0.015, 0.63), 
+    c(0.015, 0.33, 0.63, 0.015, 0.33, 0.014, 0.63), 
     c(0.99, 0.99, 0.99, 0.74, 0.74, 0.49, 0.49), size = 15)
@@
 |:-----------:|------------|-----------|---------------|---------|-----------|
-| Feature preservation: |  |  |  |  |  |
 | Preserves shapes | Y | N | Y | N | N |
@@
 | Uniform use of shape | N | N | N | Y | Y |
-| Bivariate display* | N | Y | Y | Y | Y |
-* A bivariate display refers to a map that allows two variables to be plotted. One variable via the size of the area, and another by filling the area with a colour.
+| Bivariate display^[A bivariate display refers to a map that allows two variables to be plotted.
+One variable via the size of the area, and another by filling the area with a colour.] | N | Y | Y | Y | Y |
 
@@
 The movement of a weather system will follow a forecasted path, the animation will communicate this so that all map users can follow the path of the weather system across the geography over a specified period of time.
+In the disease mapping space, the Australian Cancer Atlas [-@TACA] provides [tours](
+https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
+This directs user's attention and gives them tools to plan their own exploration.
 
-The move from static, print media displays toward interactivity allow users to develop maps to ask and answer their own questions.
-Developmets and adoptions of this technology has allowed map users more control of the message displayed.
-Users are able to see the familiar shapes, zoom in to explore densely populated communities, or change the display completely to highlight the population space.
 @CPISACA provide three questions that are generally when using a disease map:
@@
 <!-- Interactivity-->
+The move from static, print media displays toward interactivity allow users to develop maps to ask and answer their own questions.
+Developmets and adoptions of this technology has allowed map users more control of the message displayed.
+Users are able to see the familiar shapes, zoom in to explore densely populated communities, or change the display completely to highlight the population space.
+
 This oppportunity provides room to move away from the traditional approach of a single map for a single use, into a more explorative territory.

### 2019-09-10 review_article.Rmd
first choropleth map

--- 2019-09-10 review_article.Rmd
@@
 Classed choropleth maps break a continuous variable into a small set of categories, often quintiles, and use a gradual colour scheme to shade all areas that fall in the same category [@CPISACA].
-\ref{fig:choro} 1 shows a choropleth map where each state of the United States of America is coloured by the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016.
+Figure \ref{fig:choro} shows a choropleth map of the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016 in the USA.
+Data is extracted from XXX.
+XXX What do the values mean?
+What is the rate?
+Is is 40 per 100,000 or 1 million or 10,000?
 Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS]. 
@@
 
-```{r choro, fig.cap = "A choropleth map of the United States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-The state of Utah has an unusually low rate of lung and bronchus incidence.
-The state of Kentucky is high, the surrounding states also have high rates.
-There is a pattern of average higher rates toward the East Coast and lower rates toward the West Coast.", message=FALSE, warning=FALSE, out.width='80%'}
-knitr::include_graphics("figures/ggchoro.png")
+```{r choro, fig.cap = "Choropleth map of the average age-adjusted rate of new cases of lung and bronchus cancer 2012-2016 in the USA.
+XXX What is the rate of?
+Per how many people?
+Utah has an unusually low rate and Kentucky is high.
+There is generally a west to east trend, with higher rates towards the easern states.", message=FALSE, warning=FALSE, out.width='100%', fig.align='center', fig.width=12}
+library(sugarbag)
+library(tidyverse)
+library(spData)
+library(ggthemes)
+library(maptools)
+library(sf)
+library(cartogram)
+#knitr::include_graphics("figures/ggchoro.png")
+# read in data
+cancer <- read_csv("data/USCSlung.csv") %>% rename(NAME = Area)
+
+# Join polygons to data
+data(us_states)
+cancer <- left_join(us_states, cancer, by = c("NAME"))
+cancer <- st_transform(cancer, 3857)
+
+b <- st_bbox(cancer)
+ggplot(cancer) + 
+  geom_sf(aes(fill = AgeAdjustedRate)) +
+  scale_fill_distiller("Age-adjusted rate (per 1000)", type = "seq", palette = "YlOrRd",  direction = 1) + 
+  #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
+  theme_void() +
+  theme(legend.position ="bottom", legend.key.width = unit(2, "cm"))
 ```
@@
 @ACTUC suggests 'population distribution is often extremely uneven in former British colonies', this makes the distortion necessary [@CTTMB].
-When implementing a distortion of the geographical shape according to population, the resulting display is an area cartogram [@NAC], or population-by-area cartogram [@TAAM].
+When implementing a distortion of the geographical shape according to population, the resulting display is an area cartogram [@NAC], population-by-area cartogram [@TAAM], or iso-demographic map.
 
@@
 
-We would like to acknowledge Dr Earl Duncan (Research Associate at ARC Centre of Excellence for Mathematical & Statistical Frontiers, QUT).
-His time and effort given to edit and comment on this paper was invaluable.
-
 We thank Dr Susanna Cramb (Spatial Modeller, Cancer Council Queensland) and Dr Peter Baade (Senior Research Fellow, Cancer Council Queensland) for providing the opportunity to cover this exploration of disease mapping methods in writing.
-This literature review would not be possible without the opportunity provided by Queesnland University of Technology, and Cancer Council Queensland, and the roles of Professor Kerrie Mengersen (Professor of Statistics, Science and Engineering Faculty, QUT) and Dr Earl Duncan in contributing to the Australian Cancer Atlas and supervision of Stephanie's Research Masters.
+This literature review would not be possible without the opportunity provided by Queesnland University of Technology, and Cancer Council Queensland, and the roles of Professor Kerrie Mengersen (Professor of Statistics, Science and Engineering Faculty, QUT) and Dr Earl Duncan (Research Associate at ARC Centre of Excellence for Mathematical & Statistical Frontiers, QUT) in contributing to the Australian Cancer Atlas and supervision of Stephanie's Research Masters.
 

### 2019-09-10 review_article.Rmd
improved captions for choropleths

@@
 Classed choropleth maps break a continuous variable into a small set of categories, often quintiles, and use a gradual colour scheme to shade all areas that fall in the same category [@CPISACA].
-Figure \ref{fig:choro} shows a choropleth map of the average annual rate of new cases of lung and bronchus cancer from years 2012 to 2016 in the USA.
-Data is extracted from XXX.
-XXX What do the values mean?
-What is the rate?
-Is is 40 per 100,000 or 1 million or 10,000?
+Figure \ref{fig:choro} shows a classed choropleth map of the United States.
+Each state has been coloured according to the average age-adjusted rate per 100,000 people, of new cases of lung and bronchus for females and males from 2012 to 2016.
+The data was extracted from the [official federal statistics](https://gis.cdc.gov/cancer/USCS/DataViz.html) on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
+
 Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS]. 
@@
 
-```{r choro, fig.cap = "Choropleth map of the average age-adjusted rate of new cases of lung and bronchus cancer 2012-2016 in the USA.
-XXX What is the rate of?
-Per how many people?
+```{r choro, fig.cap = "Choropleth map of the average age-adjusted rate per 100,000 people, of new cases of lung and bronchus cancer 2012-2016 in the USA.
 Utah has an unusually low rate and Kentucky is high.

### 2019-09-10 review_article.Rmd
Section links

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE, fig.width = 6, fig.height = 5)
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE, fig.height = 5)
 library(ggplot2)
@@
 The paper is organised as follows.
-The next section describes the choropleth map which is the common approach to disease maps.
-Section XXX surveys atlases in use today.
-Section XXX describes an alternative display, the cartogram what may be useful for countries that have heterogeneously sized geographic units.
-The pros and cons of these approaches is discussed in Section XXX.
-Disease maps are ore useful when made interactive, and common options are described in Section XXX, along with a discussion of benefits and disadvantages.
+The next section, [Map displays for disease data](#choropleths), describes the choropleth map which is the common approach to disease maps.
+Section [Publicly available atlases](#public) surveys atlases in use today.
+The [Alternative displays](#alternative) describes an alternative display, the cartogram what may be useful for countries that have heterogeneously sized geographic units.
+The pros and cons of these approaches is discussed in Section [Comparison of mapping methods](#compare).
+Disease maps are more useful when made interactive, and common options are described in Section [Animation and Interactivity](#interacting), along with a discussion of benefits and disadvantages.
 The last section summarises the paper and discusses future directions.
 
-# Map displays for disease data
+# Map displays for disease data {#choropleths}
 
@@
 Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) (Relative risk, Lung cancer, men vs women), [g.
-An Atlas of Cancer in South Australia](https://www.cancersa.org.au/assets/images/pdfs/An%20Atlas%20of%20Cancer%20in%20South%20Australia%20-%20Full%20Report.pdf) (Incidence rate per 100,000, lung cancer in females) "}
+An Atlas of Cancer in South Australia](https://www.cancersa.org.au/assets/images/pdfs/An%20Atlas%20of%20Cancer%20in%20South%20Australia%20-%20Full%20Report.pdf) (Incidence rate per 100,000, lung cancer in females)"}
 library(cowplot)
@@
     c(0.015, 0.33, 0.63, 0.015, 0.33, 0.014, 0.63), 
-    c(0.99, 0.99, 0.99, 0.74, 0.74, 0.49, 0.49), size = 15)
+    c(0.98, 0.98, 0.98, 0.74, 0.74, 0.47, 0.47), size = 15)
 
@@
 The use of choropleths within the public domain mirrors the heavy use of choropleth maps within the research literature, discussed above.
-The cancer atlases covered geographies from all around the world.
-One atlas covered the European Union.
-Most focused on single nations, the United States was considered by eleven atlases, the United Kingdom by seven, followed by three of Australia, two of Canada, and one of each from Switzerland, Germany, Norway.
-Figure 2.a.b and c show a global, national and state choropleth respectively. 
-Not all maps had a national focus and ten covered a region or state rather than an entire nation.
-The states or counties/regions covered were South Australia (AUS), Queensland (AUS), Ontario (CAN), Valencia (Spain), Pennsylvania county Massachusetts (US), New Hampshire (US), Cape Cod (US), Missouri (US), Florida (US), New York State (US) and Arizona (US).
+The cancer atlases covered geographies from all around the world, most focused on single nations.
+Figure \@ref(choropleth-grid-create) displays a global (b), national (a,d,e,f) and state (c,g) choropleths.
 
@@
 
-# Alternative displays
+# Alternative displays {#alternative}
 
@@
 The map projections alter the shapes and angles of the boundaries of each state.
-Maps a and b are similar in their straight edges, unlike maps c and d which curve on the northern United States border.", message=FALSE, warning=FALSE, out.width='80%'}
+Maps a and b are similar in their straight edges, unlike maps c and d which curve on the northern United States border.", message=FALSE, warning=FALSE, out.width='80%', fig.align='center', fig.width=12}
 knitr::include_graphics("figures/ggchoroCRS.png")
@@
 The North East states remain closer to their neighbours, and may be displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.", message=FALSE, warning=FALSE, out.width='80%'}
+The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.", message=FALSE, warning=FALSE, out.width='80%', fig.align='center', fig.width=12}
 knitr::include_graphics("figures/ggdorl.png")
@@
 
-# Comparison of mapping methods
+# Comparison of mapping methods (#compare)
 
@@
 
-# Animation and Interactivity
+# Animation and Interactivity {#interacting}
 

### 2019-09-10 review_article.Rmd
atlases section

--- 2019-09-10 review_article.Rmd
@@
 Overwhelmingly cancer maps utilised to communicate to the public and other non-expert audiences are choropleths.
-A grey literature review of cancer atlases available in the public domain [@roberts2019communication] identified 33 cancer atlases published on the internet between January 2010 and November 2015, all of which were choropleths.
-These maps were published by non-commercial organisations, including not-for-profits, government, research organisations, advocacy groups or government funded partnerships.Only one map was published by a commercial entity.
+
+## Overview
+
+@roberts2019communication identified 33 publicly available cancer atlases, published between January 2010 and November 2015.
+All of these use choropleth maps.
+All but one of these were published by non-commercial organisations, including not-for-profits, government, research organisations, advocacy groups or government funded partnerships.
 The use of choropleths within the public domain mirrors the heavy use of choropleth maps within the research literature, discussed above.

### 2019-09-10 review_article.Rmd
done revising the choropleth map section, material on atlases shifted to the next section

@@
 The paper is organised as follows.
-The next section, [Map displays for disease data](#choropleths), describes the choropleth map which is the common approach to disease maps.
-Section [Publicly available atlases](#public) surveys atlases in use today.
-The [Alternative displays](#alternative) describes an alternative display, the cartogram what may be useful for countries that have heterogeneously sized geographic units.
+The next section describes the choropleth map which is the common approach to disease maps.
+Section \ref{public} surveys atlases in use today.
+The (#alternative) describes an alternative display, the cartogram what may be useful for countries that have heterogeneously sized geographic units.
 The pros and cons of these approaches is discussed in Section [Comparison of mapping methods](#compare).
@@
 <!-- what is a choropleth map -->
-A choropleth map is used to display the characteristics of a spatial relationship of measurements. 
-They are true maps of the topology, constructed for visual inspection of spatial patterns on a familiar geographic form.
-Classed choropleth maps break a continuous variable into a small set of categories, often quintiles, and use a gradual colour scheme to shade all areas that fall in the same category [@CPISACA].
-Figure \ref{fig:choro} shows a classed choropleth map of the United States.
-Each state has been coloured according to the average age-adjusted rate per 100,000 people, of new cases of lung and bronchus for females and males from 2012 to 2016.
+A choropleth map is used to display the spatial relationship of measurements.
+The geography  is faithfully rendered, and the colour rendering is designed to reveal spatial patterns among data values.
+A choropleth is constructed by drawing the geographic or political boundaries, and filling the shapes with colours to represent values of a measured variable [@EI].
+Figure \ref{fig:choro} shows a choropleth of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged over 2012 through 2016.
 The data was extracted from the [official federal statistics](https://gis.cdc.gov/cancer/USCS/DataViz.html) on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
 
+<!-- history and types of choropleths, variations -->
+Early versions of choropleth maps used symbols or patterns instead of colour.
+As an alternative storage device to a table, it preserves locations for geographically ordered data, with usage dating back to the 1800s [@GOINO] XXX Is this the right reference here?
+It doesn't seem to match the statement of early examples.
+Maybe you want to reference Dupin's map?
+@CPISACA discuss the use of choropleths to visualise cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
+
+<!-- psychology -->
+Utilising the state boundaries can make a map familiar to read [@CIBMUK], and allows viewers to visually infer the spatial relationships in the data, i.e.
+how cancer rate differs across the states.
+The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
+Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
+
+<!-- how do we use choropleth's -->
 Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS].
-A choropleth is constructed by drawing the geographic or political boundaries, and filling the shapes with colours to represent values of a measured variable [@EI].
-Early versions of choropleth maps used symbols or patterns instead of colour.
-As an alternative storage device to a table, it preserves locations for geographically ordered data, with usage dating back to the 1800s [@GOINO].
-@CPISACA discuss the use of choropleths to visualise cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
+In Figure \ref{fig:choro}, a west to east spatial trend  of increasing rates, can be seen.
+There is also a spatial outlier -- Utah has a noticeably lower rate that its neighbours.
+Also Kentucky has a noticeably high rate, and Maine also has a higher rate than its neighbours.
+There is something of a cluster of higher rates around the tobacco states.
 
@@
   geom_sf(aes(fill = AgeAdjustedRate)) +
-  scale_fill_distiller("Age-adjusted rate (per 1000)", type = "seq", palette = "YlOrRd",  direction = 1) + 
+  scale_fill_distiller("Age-adjusted rate \n (per 100,000)", type = "seq", palette = "YlOrRd",  direction = 1) + 
   #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
@@
 
-<!-- psychology -->
-Utilising the familiar state boundaries can make a map intuitive to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data, i.e.
-how cancer rate differs across the states.
-The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
-Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
+<!-- potential problems -->
 While the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
@@
 Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].
+
+```{r choropleth-grid-create, message=FALSE, fig.cap = "Public facing choropleth cancer maps.[a.
+The Environment and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/) (Lung cancer in females)  (Incidence rates per 100,000 for all cancers, male and female), [b.
+Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx),  [c.
+Atlas of Cancer in QLD - Relative incidence ration, Lung cancer in males](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/), [d.
+Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), [e.
+United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/) (Incidence rate per 100,000, all cancer types, men and women), [f.
+Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) (Relative risk, Lung cancer, men vs women), [g.
+An Atlas of Cancer in South Australia](https://www.cancersa.org.au/assets/images/pdfs/An%20Atlas%20of%20Cancer%20in%20South%20Australia%20-%20Full%20Report.pdf) (Incidence rate per 100,000, lung cancer in females)"}
+library(cowplot)
+
+# Import the images
+plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
+plot2 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png') 
+plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
+plot4 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
+plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
+plot6 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
+plot7 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
+
+ggdraw() +
+  draw_plot(rasterGrob(plot1), 0, .75, 0.33, .25) +
+  draw_plot(rasterGrob(plot2), 0.33, 0.75, 0.33, .25) +
+  draw_plot(rasterGrob(plot3), 0.66, 0.5, .33, .5) +
+  draw_plot(rasterGrob(plot4), 0, 0.5, .33, .25) +
+  draw_plot(rasterGrob(plot5), 0.33, 0.5, .33, .25) +
+  draw_plot(rasterGrob(plot6), 0, 0, .66, .5) +
+  draw_plot(rasterGrob(plot7), .66, 0, .33, .5) +
+  draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
+    c(0.015, 0.33, 0.63, 0.015, 0.33, 0.014, 0.63), 
+    c(0.98, 0.98, 0.98, 0.74, 0.74, 0.47, 0.47), size = 15)
+
+ggsave(filename = "figures/choropleth_grid.png", 
+  device = "png", width = 6, height = 5, dpi = 300)
+```
+
+# Cancer atlases
 
@@
 
-```{r choropleth-grid-create, message=FALSE, fig.cap = "Public facing choropleth cancer maps.[a.
-The Environment and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/) (Lung cancer in females)  (Incidence rates per 100,000 for all cancers, male and female), [b.
-Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx),  [c.
-Atlas of Cancer in QLD - Relative incidence ration, Lung cancer in males](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/), [d.
-Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), [e.
-United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/) (Incidence rate per 100,000, all cancer types, men and women), [f.
-Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) (Relative risk, Lung cancer, men vs women), [g.
-An Atlas of Cancer in South Australia](https://www.cancersa.org.au/assets/images/pdfs/An%20Atlas%20of%20Cancer%20in%20South%20Australia%20-%20Full%20Report.pdf) (Incidence rate per 100,000, lung cancer in females)"}
-library(cowplot)
-
-# Import the images
-plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
-plot2 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png') 
-plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
-plot4 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
-plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
-plot6 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
-plot7 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
-
-ggdraw() +
-  draw_plot(rasterGrob(plot1), 0, .75, 0.33, .25) +
-  draw_plot(rasterGrob(plot2), 0.33, 0.75, 0.33, .25) +
-  draw_plot(rasterGrob(plot3), 0.66, 0.5, .33, .5) +
-  draw_plot(rasterGrob(plot4), 0, 0.5, .33, .25) +
-  draw_plot(rasterGrob(plot5), 0.33, 0.5, .33, .25) +
-  draw_plot(rasterGrob(plot6), 0, 0, .66, .5) +
-  draw_plot(rasterGrob(plot7), .66, 0, .33, .5) +
-  draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
-    c(0.015, 0.33, 0.63, 0.015, 0.33, 0.014, 0.63), 
-    c(0.98, 0.98, 0.98, 0.74, 0.74, 0.47, 0.47), size = 15)
-
-ggsave(filename = "figures/choropleth_grid.png", 
-  device = "png", width = 6, height = 5, dpi = 300)
-```
-
-# Publicly available atlases
-
 Cancer maps are effective tools for communicating to wide range of audiences, including the general public and others not trained in statistical analyses.
@@
 
-## Overview
+## Overview of publicly available atlases
 
@@
 
-## Supplementary statistics in public facing atlases
+XXX more details on the examples in the next figure
+
+## Common statistics displayed
 
@@
 
-# Alternative displays {#alternative}
+## Interaction with atlases
+
+# Cartograms and other spatially warped displays {#alternative}
 
@@
 
-### Alternatives
+### Dorling
 
@@
 
+## Tilegram
+
 ```{r ggtilegram, fig.cap = "A tilegram of the Unites States of America.
@@
 
-# Comparison of mapping methods (#compare)
+Bob Rudis, geogrid
+
+# Comparison of maps and cartograms {#compare}
 

### 2019-09-11 review_article.Rmd
Interactivity section restored

--- 2019-09-10 review_article.Rmd
@@
 Overwhelmingly cancer maps utilised to communicate to the public and other non-expert audiences are choropleths.
-
-## Overview
-
-@roberts2019communication identified 33 publicly available cancer atlases, published between January 2010 and November 2015.
-All of these use choropleth maps.
-All but one of these were published by non-commercial organisations, including not-for-profits, government, research organisations, advocacy groups or government funded partnerships.
+A grey literature review of cancer atlases available in the public domain [@roberts2019communication] identified 33 cancer atlases published on the internet between January 2010 and November 2015, all of which were choropleths.
+These maps were published by non-commercial organisations, including not-for-profits, government, research organisations, advocacy groups or government funded partnerships.Only one map was published by a commercial entity.
 The use of choropleths within the public domain mirrors the heavy use of choropleth maps within the research literature, discussed above.
@@
 Unlike innovation in spatial statistical methods, interactity, data collection and record keeping little work has been done to implement different mapping approaches -->
+
+The interactivity of modern mapping methods enable supplementary information to be incorporated without cluttering the screen.
+Tool tip features and drop down menus allow for user exploration, small screens are not a problem with the flexibility provided by zooming and panning [@HTLWM].
+This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accesible' ways [@TNTEA].
+The use of these supports were found in a many online diease mapmap sources.
+These interactive features provide an opportunity for users to explore additional information to help users understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
+This directly enables epidemiologists to explore spatial distributions of cancer statistics.
+The interactive features of the publically available maps identified by @roberts2019communication include exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
 

### 2019-09-11 review_article.Rmd
figure and section labelling

--- 2019-09-10 review_article.Rmd
@@
 The next section describes the choropleth map which is the common approach to disease maps.
-Section \ref{public} surveys atlases in use today.
-The (#alternative) describes an alternative display, the cartogram what may be useful for countries that have heterogeneously sized geographic units.
-The pros and cons of these approaches is discussed in Section [Comparison of mapping methods](#compare).
-Disease maps are more useful when made interactive, and common options are described in Section [Animation and Interactivity](#interacting), along with a discussion of benefits and disadvantages.
+Section XXX surveys atlases in use today.
+Section XXX describes an alternative display, the cartogram what may be useful for countries that have heterogeneously sized geographic units.
+The pros and cons of these approaches is discussed in Section  XXX.
+Disease maps are more useful when made interactive, and common options are described in Section  XXX, along with a discussion of benefits and disadvantages.
 The last section summarises the paper and discusses future directions.
@@
 
+The interactivity of modern mapping methods enable supplementary information to be incorporated without cluttering the screen.
+Tool tip features and drop down menus allow for user exploration, small screens are not a problem with the flexibility provided by zooming and panning [@HTLWM].
+This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accesible' ways [@TNTEA].
+The use of these supports were found in a many online diease mapmap sources.
+These interactive features provide an opportunity for users to explore additional information to help users understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
+This directly enables epidemiologists to explore spatial distributions of cancer statistics.
+The interactive features of the publically available maps identified by @roberts2019communication include exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
+
 XXX more details on the examples in the next figure
@@
 
-```{r supp-grid-create, fig.cap = "Interactive controls of displays in public facing choropleth cancer maps.
-[a.
-Globocan 2018: Cancer Today](http://globocan.iarc.fr/Pages/Map.aspx)(Controls for indicator, sex, age groups, continents, and cancer types) , [b.
-Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/) (Controls for indicator, age, year, and geographical areas), [c.
-The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M) (Controls for predetermined combinations of cancer type or risk factors, mortality or incidence, and gender) [d.
-Global Cancer Map](http://globalcancermap.com/) (Controls for indicator, and cancer type), [e.
-United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html) (Controls for demographics, trends, geography, indicators, risk factors, sex, cancer type, and year)."}
-i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png')
-i2 <- png::readPNG('cancer_map_images/interactivity/15.interactivity.png')
-i3 <- png::readPNG('cancer_map_images/interactivity/4.interactivity.png')
-i4 <- png::readPNG('cancer_map_images/interactivity/5.interactivity.png')
-i5 <- png::readPNG('cancer_map_images/interactivity/7.interactivity.png')
-
-ggdraw() +
-  draw_plot(rasterGrob(i1), 0,  0.13,   .28,  .83) + #a
-  draw_plot(rasterGrob(i2), 0.32, 0.56, .68, .4) + #b
-  draw_plot(rasterGrob(i3), 0.32, 0.41, .68, .13) + #c
-  draw_plot(rasterGrob(i4), 0.32, 0.14, .68, .25) + #d
-  draw_plot(rasterGrob(i5), 0, 0, 1, .13) + #e
-  draw_plot_label(c("a", "b", "c", "d", "e"), 
-    c(0.01, 0.3, 0.3, 0.3, 0.01), 
-    c(0.99, 0.99, 0.59, 0.40, 0.19), size = 15)
-
-ggsave(filename = "figures/supp_grid.png", device = "png", 
-  width = 6, height = 5, dpi = 300)
-```
-
 The interactivity of modern mapping methods enable supplementary information to be incorporated without cluttering the screen.
@@
 @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
+
+```{r supp-grid-create, fig.cap = "Interactive controls of displays in public facing choropleth cancer maps.
+[a.
+Globocan 2018: Cancer Today](http://globocan.iarc.fr/Pages/Map.aspx)(Controls for indicator, sex, age groups, continents, and cancer types) , [b.
+Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/) (Controls for indicator, age, year, and geographical areas), [c.
+The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M) (Controls for predetermined combinations of cancer type or risk factors, mortality or incidence, and gender) [d.
+Global Cancer Map](http://globalcancermap.com/) (Controls for indicator, and cancer type), [e.
+United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html) (Controls for demographics, trends, geography, indicators, risk factors, sex, cancer type, and year)."}
+i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png') # b in choro grid
+i2 <- png::readPNG('cancer_map_images/interactivity/15.interactivity.png') # d in choro grid
+i3 <- png::readPNG('cancer_map_images/interactivity/4.interactivity.png')
+i4 <- png::readPNG('cancer_map_images/interactivity/5.interactivity.png')
+i5 <- png::readPNG('cancer_map_images/interactivity/7.interactivity.png') # e in choro grid
+
+ggdraw() +
+  draw_plot(rasterGrob(i1), 0,  0.13,   .28,  .83) + #a
+  draw_plot(rasterGrob(i2), 0.32, 0.56, .68, .4) + #b
+  draw_plot(rasterGrob(i3), 0.32, 0.41, .68, .13) + #c
+  draw_plot(rasterGrob(i4), 0.32, 0.14, .68, .25) + #d
+  draw_plot(rasterGrob(i5), 0, 0, 1, .13) + #e
+  draw_plot_label(c("i", "ii", "iii", "iv", "v"), 
+    c(0.01, 0.3, 0.3, 0.3, 0.01), 
+    c(0.99, 0.99, 0.57, 0.40, 0.15), size = 15)
+
+ggsave(filename = "figures/supp_grid.png", device = "png", 
+  width = 6, height = 5, dpi = 300)
+```
 
@@
 The additional statistics often include a measure of the statistical uncertainty of the values of the statistics presented in a choropleth.
-In the review of atlases in the public domain, atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map [@@roberts2019communication].
+In the review of atlases in the public domain, atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map [@roberts2019communication].
 The maps considered used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.

### 2019-09-11 review_article.Rmd
Earl's suggested small edits

@@
 Visualising diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH].
-Disease maps help to present geographic patterns that may have been overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
+Disease maps help to present geographic patterns that may be overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
 By providing a visual representation of cancer outcomes, geographic patterns of disease are able to be identified and effectively addressed with public healh policy and actions.
@@
 The next section describes the choropleth map which is the common approach to disease maps.
-Section XXX surveys atlases in use today.
-Section XXX describes an alternative display, the cartogram what may be useful for countries that have heterogeneously sized geographic units.
-The pros and cons of these approaches is discussed in Section  XXX.
-Disease maps are more useful when made interactive, and common options are described in Section  XXX, along with a discussion of benefits and disadvantages.
+Section \ref{ch:public} surveys atlases in use today.
+Section \ref{ch:alternatives} describes an alternative display, the cartogram what may be useful for countries that have heterogeneously sized geographic units.
+The pros and cons of these approaches is discussed in Section \ref{ch:compare}.
+Disease maps are more useful when made interactive, and common options are described in Section \ref{ch:interacting}, along with a discussion of benefits and disadvantages.
 The last section summarises the paper and discusses future directions.
 
-# Map displays for disease data {#choropleths}
+# Map displays for disease data {#ch:choropleths}
 
 <!-- what is a choropleth map -->
-A choropleth map is used to display the spatial relationship of measurements.
+A choropleth map is used to display the spatial relationship of measurements by shading areas of a map.
 The geography is faithfully rendered, and the colour rendering is designed to reveal spatial patterns among data values.
@@
 Early versions of choropleth maps used symbols or patterns instead of colour.
-As an alternative storage device to a table, it preserves locations for geographically ordered data, with usage dating back to the 1800s [@GOINO] XXX Is this the right reference here?
-It doesn't seem to match the statement of early examples.
-Maybe you want to reference Dupin's map?
 @CPISACA discuss the use of choropleths to visualise cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
@@
 <!-- psychology -->
-Utilising the state boundaries can make a map familiar to read [@CIBMUK], and allows viewers to visually infer the spatial relationships in the data, i.e.
-how cancer rate differs across the states.
+Utilising the state boundaries can make a map familiar to read [@CIBMUK], and allows viewers to visually infer the spatial relationships in the data. 
 The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
@@
 The different population and geographical sizes of administrative areas can attract attention to the shades of the unpopulated but large areas [@EI]. 
-Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].
+Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention away from the 'potentially more important results in the more populous communities' that are geographically smaller [@SE].
 
@@
 
-# Cancer atlases
+# Cancer atlases {#ch:public}
 
@@
 <!-- Measures reported in cancer atlases -->
-Almost 100 years of cancer mapping in the United States and the United Kingdom has seen increased effectiveness in the presentation cancer statistics.
+The presentation of cancer statistics has changed over time with greater access to computational power.
 Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the higher prevalence of cancers in older populations.
 Howe [-@HEDP] describes Stock's development of the standardised mortality ratios through the 1930s.
-Table \ref{tab:measures} presents summarises the measures presented in published cancer atlases, and provides a definition of each measure.
+Table \ref{tab:measures} summarises the measures presented in published cancer atlases, and provides a definition of each measure.
 
@@
 
-The presentation of these rates considered not only the map base, but also appropriate choices of colour blind friendly colour schemes, and categories of the values to highlight significantly different areas.
+The presentation of these rates considers not only the shapes of the areas, but also appropriate choices of colour blind friendly colour schemes, and categories of the values to highlight significantly different areas.
 <!--
@@
 Cancer maps are effective tools for communicating to wide range of audiences, including the general public and others not trained in statistical analyses.
-These visualisation enable thse non-expert audiences to access the outputs of sophisticated statistical analyses.
-Cruickshank's (1947) as cited by @DMAHP, discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
-Overwhelmingly cancer maps utilised to communicate to the public and other non-expert audiences are choropleths.
+These visualisations enable non-expert audiences to interpret the outputs of sophisticated statistical analyses.
+Cruickshank (1947) as cited by @DMAHP, discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
+Overwhelmingly, cancer maps utilised to communicate to the public and other non-expert audiences are choropleths.
 
@@
 Supplementary graphs and plots are often included to add more depth and information to the map.
-@CPISACA suggests additional materials such as 'good tables, graphs, and explanatory text' support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers.
+@CPISACA suggests additional materials such as tables, graphs, and text explanations support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers.
 There are many visualisations used for displays of statistical summaries, these may be dot plots, bar plots, box plots, timesseries plots, cumulative distribution plots, scatter plots, Q-Q plots. 
@@
 ### Geographic hierarchies
+
 While atlases are often used to describe differences between areas, statistics may be displayed at different levels of aggregation.
-The World Health Organisation and UN Regions can also be used to aggregate areas [@IARC_3].
+Global health statistics can be aggregated to administrative and arbitrarily defined regions, such as those used by the World Health Organisation and the United Nations [@IARC_3].
 World atlases can allow for displays of data aggregated into continents, countries, states, provinces and congressional districts [@USCS].
@@
 ### Population distribution
-It is likely that each population area will have a different amount of people, information regarding the distribution of population levels may be provided in a table or histogram display [@NICR_1]. 
+
+It is extremely likely that each population area will have a different number of people.
+The distribution of the population residing in all areas may also be communicated in a table or histogram display [@NICR_1].
 Atlases have the opportunity to connect the population to the land available to them by communicating population density.
-Atlases can also connect the population to the land available to them by communicating population density as is shown in the United States map examples.
+Atlases can also connect the population to the land available to them by communicating population density.
 
 ### Statistical uncertainty
-The additional statistics often include a measure of the statistical uncertainty of the values of the statistics presented in a choropleth.
+Additional statistics that accompany an atlas often include a measure of the statistical uncertainty surrounding the statistics presented in a choropleth.
 In the review of atlases in the public domain, atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map [@roberts2019communication].
@@
 Other methods involve providing adjacent maps or overlapping maps with symbols [@VSSDCUC].
-The maps employing uncertainty ranged from static pdfs or infographics, to interactive online resources.
-Communicating the uncertainty associated with the estimates through a confidence (CI) interval, statistical significance levels, boxplots, distribution plots, and reporting sample size and standard deviations.
+The maps employing uncertainty ranged from static documents or infographics, to interactive online visualisations.
+Communicating the statistical uncertainty associated with the estimates occurs using confidence intervals (CI), credible intervals (CrI), statistical significance levels, boxplots, distribution plots, and reporting sample size and standard deviations.
 Close to half of the atlases identified (42%, n=14) included some measure of uncertainty. The most common measure used to represent uncertainty were credible or confidence intervals (CIs).
@@
 Sex is an important cofactor for cancer atlases. 
-As some cancers are sex specific, and others may be found in both males and females, atlases often specify their use of gender in the displays. 
-Atlases can allow for interaction as users can select between displays for males, females or both depending on the type of cancers explored.
+As some cancers are sex specific, and others may be found in both males and females, atlases often specify the relevant sex as part of the visual output in the displays. 
+Digital atlases allow for users to interact with the controls of the displays, they can select males, females or both depending on the type of cancers explored.
 
@@
 Socio-economic indicators can explain how the experience of cancer prevalence varies for various members of a society.
-These indicators include unemployment rates, poverty rates, remoteness and education levels achieved though, only a few atlases also explored the impact of rurality on cancer rates. 
+These indicators include unemployment rates, poverty rates, remoteness, and education levels achieved though, only a few atlases also explored the impact of rurality on cancer rates. 
 These rates may also be explored as percentages above or below the mean or median value for the set of spatial areas.
-The Human Development Index can be used to understand the socio-economic experience of a community, as can the Income levels as measured by World bank list of economies[@IARC_3]. 
-The areas are usually ranked and allocated to quintiles to be presented as categories describing the ranking.
+The Human Development Index can be used to understand the socio-economic experience of a community, as well as Income levels which can be sourced from the World Bank list of economies [@IARC_3]. 
+The areas are often ranked and allocated to quintiles, each quintile can be presented as categories describing the ranking.
 
@@
 
-# Cartograms and other spatially warped displays {#alternative}
+# Cartograms and other spatially warped displays {#ch:alternatives}
 
@@
 For a single variable of interest, each map area is changed to emphasise the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].
-The changes in the map base occur by altering individual areas, by altering the shapes or boundaries.
+Changes in the map base are implemented by altering the boundaries, and therefore shapes, of individual areas.
 
@@
 To present election results on a choropleth map should be 'unthinkable', as it means diminishing the visual impact of majority of the electorates.
-A 1966 cartogram was presented an alternative which uses mostly straight lines, and the result looks very little like the geographical shape of Australia.
-This issue is felt in any nation which experiences an uneven population distribution.
-The United States is affect by the different densities of neighbourhoods or states.
-As this population distribution continues to change the need for cartograms as an alternative to a choropleth map should only increase.
+A 1966 cartogram presented an alternative where boundary lines were largely straight line, and the result looked very little like the geographical shape of Australia.
+This issue is felt in any nation which experiences a spatially heterogeneous population distribution.
+As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
 
@@
 The spatial transformation of map regions relative to the data emphasises the data distribution instead of land size [@CBATCC].
-When visualising population statistics @ACTUC considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
+When visualising population statistics, @ACTUC considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
 @HEDP suggests that 'cancer occurs in people, not in geographical areas' and @CTTMB believe that spatial socio-economic data, like cancer rates, are best presented on a cartogram for urban areas as the population map base avoids allocating 'undue prominence' to rural areas.
@@
 Early approaches including John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball bearing approach and Tobler's (1973) computer programs [@ACTUC].
-Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’[@CD].
+Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994), [@CD].
 @HEDP discusses the impact of electronic computer-assisted techniques.
@@
 @ACTUC reiterates: 'There is no "best" cartogram or method of creating cartograms just as there is no "best" map' (Monmonier and Schnell, 1988). 
-There have been many algorithms presented, @SAIC provided a framework to investigate implementations and the "statistical accuracy, geographical accuracy, and topological accuracy".
+@SAIC provided a framework to investigate implementations of the many algorithms presented, and the "statistical accuracy, geographical accuracy, and topological accuracy".
 
@@
 The state of California has become much larger due to it's large population density.
-This draws attention to the densely populated North East region, and detracts from the Mid West.", message=FALSE, warning=FALSE, out.width='80%'}
+This draws attention to the densely populated North East region, and detracts from the less populated Mid West.", message=FALSE, warning=FALSE, out.width='80%'}
 knitr::include_graphics("figures/ggcont.png")
@@
 
-A contiguous cartogram maintains connectivity of the map regions while areas are resized according to a statistic.
+A contiguous cartogram maintains connectivity of the map regions while areas are altered according to a statistic.
 This transformation often occurs at the expense of the shape of areas [@CBATCC, @NAC, @TAAM].
@@
 @CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes. 
-@MDAC define the term 'mosaic cartograms' for hexagonal tile displays, where the amount of tiles for each area can be used to communicate the statistic of regions. 
+@MDAC define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area can be used to communicate the statistic of regions. 
 The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off be made between boundary complexity and simplicity.
@@
 
-As the trade-off regarding boundaries approaches simplicity, the distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the noncontiguous cartogram but is replaced by the meaningful empty-space property [@NAC, @ECGC].
+As the trade-off regarding boundaries approaches simplicity, the distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the non-contiguous cartogram but is replaced by the meaningful empty-space property [@NAC, @ECGC].
 The shapes are valuable for recognition and allows users to orient themselves on the display.
@@
 
-Daniel Dorling presented an alternative display engineere to highlight the spatial distribution and neighbourhood relationships without complex distortions of borders and boundaries.
+Daniel Dorling presented an alternative display engineers to highlight the spatial distribution and neighbourhood relationships without complex distortions of borders and boundaries.
 This approach opposes preserving the intricate shape details and is founded in the simple question put forward by Daniel Dorling -@ACTUC:
@@
 
-He acknowledged the sophistication of contiguous cartograms but critiqued their 'very complex shapes', he answers this with his implementation of maps created using 'the simplest of all shapes'. 
-Circular cartograms use the same simple shapes for every region represented, and size the shapes according to the statistic represented or the population for a base map.
+He acknowledged the sophistication of contiguous cartograms but critiqued their 'very complex shapes,' he answers this with his implementation of maps created using 'the simplest of all shapes'. 
+Circular cartograms use the same simple shapes for every region represented, and resizes the shapes according to the statistic represented or the population for a base map.
 This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms, as the 'nonsense' shapes used have 'no meaning' after distortions are applied @NISCC.
@@
 
-# Comparison of maps and cartograms {#compare}
+# Comparison of maps and cartograms {#ch:compare}
 
@@
 
-# Animation and Interactivity {#interacting}
+# Animation and Interactivity {#ch:interacting}
 

### 2019-09-15 aus_plots.R
Fix figure colour schemes

@@ -2,19 +2,19 @@
 library(sugarbag)
 library(tidyverse)
 
-# Find the shape file
-aus <- file.choose()
+# 
+# # Filter islands
+# aus_sf <- aus_sf %>% 
+#   filter(!(SA2_NAME11 %in% 
+#       c("Christmas Island", "Cocos (Keeling) Islands", "Lord Howe Island")))
 
-# Read in shape file, and simplify
-aus_sf <- read_shape(shp_path = aus, simplify = TRUE, keep = 0.05)
+# Join with cancer data from AIHW
+#https://www.aihw.gov.au/getmedia/df6732ee-5246-4c9f-a458-df85c88fc512/aihw-can-108-cancer-SA3-incidence-mortality.xls.aspx
 
-# Filter islands
-aus_sf <- aus_sf %>% 
-  filter(!(SA2_NAME11 %in% 
-      c("Christmas Island", "Cocos (Keeling) Islands", "Lord Howe Island")))
+library(readxl)
+cimar_sa3 <- read_excel("data/cimar-sa3.xls", 
+  sheet = "Incidence Persons", range = "A3:Z337")
 
-# Join with cancer data
-SIR <- read_csv("data/SIR Downloadable Data.csv") 
 
 SIR <- SIR %>%
   filter(Year == "2010-2014") 

### 2019-09-15 sa3_plots.R

@@ -1,0 +1,168 @@
+#sa3_plots
+
+# Create a choropleth of Australia
+library(sugarbag)
+library(tidyverse)
+library(readxl)
+library(spData)
+library(ggthemes)
+library(maptools)
+library(sf)
+library(cartogram)
+
+# Join with cancer data from AIHW
+#https://www.aihw.gov.au/getmedia/df6732ee-5246-4c9f-a458-df85c88fc512/aihw-can-108-cancer-SA3-incidence-mortality.xls.aspx
+
+cimar_sa3 <- read_excel("data/cimar-sa3.xls", 
+  sheet = "Incidence Persons", range = "A3:Z337") %>%  
+  # select lung cancer variables
+  select(Code = `...1`,
+    Name = `...2`,
+    `Total incidence` = `...22`,
+    Population = `...23`,
+    `Crude rate (per 100,000)` = `...24`,
+    `Age-standardised rate (per 100,000)` = `...25`,
+    `Rate ratio (relative to all of Australia)` = `...26`
+  ) %>% 
+  filter(!is.na(Name)) %>% 
+  filter(!(Name %in% c("Name", "Australia"))) %>% 
+  mutate(Population = as.numeric(Population))
+
+
+# Filter for desired areas
+cimar_sa3 <- cimar_sa3 %>% 
+  filter(!(Name %in% 
+    c("Christmas Island", "Cocos (Keeling) Islands", "Lord Howe Island", "Jervis Bay")))
+
+# Join with sa3 sf object
+sa3lung <- absmapsdata::sa32016 %>% 
+  left_join(cimar_sa3, by = c("sa3_name_2016"= "Name")) %>%
+  st_as_sf() %>%
+  filter(!is.na(Population)) %>%
+  filter(!st_is_empty(geometry))
+
+sa3lung <- st_transform(sa3lung, 3112)
+
+b <- st_bbox(sa3lung)
+
+###############################################################################
+
+sa3lung <- sa3lung %>% 
+  mutate(`Age-standardised rate (per 100,000)` = as.numeric(`Age-standardised rate (per 100,000)`))
+  
+# Create a choropleth
+aus_ggchoro <- ggplot(sa3lung) + 
+  geom_sf(aes(fill = `Age-standardised rate (per 100,000)`), colour = NA) +
+  scale_fill_distiller(type = "seq", palette = "Purples",  direction = 1, na.value = "light grey") + 
+  ggtitle(title) +
+  theme_void() +
[…]

### 2019-09-15 usa_plots.R

@@ -21,13 +21,17 @@ library(ggthemes)
 library(maptools)
 library(sf)
 library(cartogram)
+library(grid)
+library(png)
+library(gridExtra)
 
 # read in data
 cancer <- read_csv("data/USCSlung.csv") %>% rename(NAME = Area)
 
 
 # Join polygons to data
-data(us_states)
+# devtools::install_github("hrbrmstr/albersusa")
+library(albersusa)
 cancer <- left_join(us_states, cancer, by = c("NAME"))
 cancer <- st_transform(cancer, 3857)
 
@@ -43,10 +47,10 @@ title <- ""
 ###############################################################################
 # Create a choropleth
 ggchoro <- ggplot(cancer) + 
-  geom_sf(aes(fill = AgeAdjustedRate)) +
-  scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
+  geom_sf(aes(fill = AgeAdjustedRate), colour = NA) +
+  scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
   ggtitle(title) +
-  coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
+  #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
   theme_void() +
   theme(legend.position ="bottom")
 ggsave(filename = "figures/ggchoro.png", device = "png", dpi = 300,  width = 12, height = 6)
@@ -59,10 +63,10 @@ ggchoro
 cont <- cartogram_cont(cancer,
   weight = "total_pop_15") %>% st_as_sf()
 ggcont <- ggplot(cont) + 
-  geom_sf(aes(fill = AgeAdjustedRate)) + 
+  geom_sf(aes(fill = AgeAdjustedRate), colour = NA) + 
   ggtitle(title) +
-  scale_fill_distiller(type = "seq", palette = "YlOrRd",  direction = 1) + 
-  coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
+  scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
[…]

### 2019-09-15 review_article.Rmd
Table of caption information for USA plots

@@
 Figure \ref{fig:choro} shows a choropleth of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged over 2012 through 2016.
-The data was extracted from the [official federal statistics](https://gis.cdc.gov/cancer/USCS/DataViz.html) on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
+The data was extracted from the official federal statistics by @usa_lung on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
 
@@
 ggplot(cancer) + 
-  geom_sf(aes(fill = AgeAdjustedRate)) +
-  scale_fill_distiller("Age-adjusted rate \n (per 100,000)", type = "seq", palette = "YlOrRd",  direction = 1) + 
+  geom_sf(aes(fill = AgeAdjustedRate), colour = NA) +
+  scale_fill_distiller("Age-adjusted rate \n (per 100,000)", type = "seq", palette = "RdPu",  direction = 1) + 
   #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
@@
 
-### Contiguous
-
-<!-- What is a contiguous cartogram -->
-
 ```{r ggcont, fig.cap = "A contiguous cartogram map of the United States of America.
@@
 
+```{r ggncont, fig.cap = "A Non - contiguous cartogram map of the United States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state shape has been maintained, but the size has altered according to the population of the state in 2015.
+The state of California has remained closer to it's original size than it's surrounding states.
+The North East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut.
+This draws attention to the densely populated North East region, and the sparse Mid West.", message=FALSE, warning=FALSE, out.width='80%'}
+knitr::include_graphics("figures/ggncont.png")
+```
+
+```{r ggdorl, fig.cap = "A dorling cartogram of the Unites States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been represented by a circle, but the size was determined by the population of the state in 2015.
+The North East states remain closer to their neighbours, and may be displaced from their geographic location.
+The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.", message=FALSE, warning=FALSE, out.width='80%', fig.align='center', fig.width=12}
+knitr::include_graphics("figures/ggdorl.png")
+```
+
+```{r gghexmap, fig.cap = "A hexagon tesselation map of the Unites States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been represented by a square, each square is the same size.
+The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.
+The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.", message=FALSE, warning=FALSE, out.width='80%'}
+knitr::include_graphics("figures/gghexmap.png")
+```
+
+Table: \label{tab:usa} Maps used to present statistics for the Uited States of America
+
+|Map display |Details| 
+|:--------------------|:------------------------------------------------|  
+|a.
+contiguous cartogram|A contiguous cartogram map of the United States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each stated shape has been distorted according to the population of the state in 2015.
+The state of California has become much larger due to it's large population density.
+This draws attention to the densely populated North East region, and detracts from the less populated Mid West.|
+|b.
+Non - Contiguous|A Non - contiguous cartogram map of the United States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state shape has been maintained, but the size has altered according to the population of the state in 2015.
+The state of California has remained closer to it's original size than it's surrounding states.
+The North East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut.
+This draws attention to the densely populated North East region, and the sparse Mid West.|
+|Dorling|A dorling cartogram of the Unites States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been represented by a circle, but the size was determined by the population of the state in 2015.
+The North East states remain closer to their neighbours, and may be displaced from their geographic location.
+The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.|
+|Hexagon Tesselation|A hexagon tesselation map of the Unites States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been represented by a square, each square is the same size.
+The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.
+The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.|
+
+### Contiguous
+
+<!-- What is a contiguous cartogram -->
 <!-- Intentionally preserve neighbours -->
@@
 
-```{r ggncont, fig.cap = "A Non - contiguous cartogram map of the United States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each state shape has been maintained, but the size has altered according to the population of the state in 2015.
-The state of California has remained closer to it's original size than it's surrounding states.
-The North East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut.
-This draws attention to the densely populated North East region, and the sparse Mid West.", message=FALSE, warning=FALSE, out.width='80%'}
-knitr::include_graphics("figures/ggncont.png")
-```
-
 ### Dorling
@@
 This implementation can work for up to 'one hundred thousand' areas.
-
-```{r ggdorl, fig.cap = "A dorling cartogram of the Unites States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each state has been represented by a circle, but the size was determined by the population of the state in 2015.
-The North East states remain closer to their neighbours, and may be displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.", message=FALSE, warning=FALSE, out.width='80%', fig.align='center', fig.width=12}
-knitr::include_graphics("figures/ggdorl.png")
-```
 

### 2019-09-15 sa3_plots.R

@@ -54,7 +54,6 @@ sa3lung <- sa3lung %>%
 aus_ggchoro <- ggplot(sa3lung) + 
   geom_sf(aes(fill = `Age-standardised rate (per 100,000)`), colour = NA) +
   scale_fill_distiller(type = "seq", palette = "Purples",  direction = 1, na.value = "light grey") + 
-  ggtitle(title) +
   theme_void() +
   coord_sf(crs = CRS("+init=epsg:3112"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
   theme(legend.position ="bottom") + 
@@ -112,8 +111,7 @@ dorl <- sa3lung %>% cartogram_dorling(.,
 aus_ggdorl <- ggplot(dorl) + 
   geom_sf(aes(fill = `Age-standardised rate (per 100,000)`)) + 
   scale_fill_distiller(type = "seq", palette = "Purples",  direction = 1) + 
-  coord_sf(crs = CRS("+init=epsg:3112"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
-  theme_void() + theme(legend.position ="bottom")
+ theme_void() + theme(legend.position ="bottom")
 aus_ggdorl
 ggsave(filename = "figures/aus_ggdorl.png", device = "png", dpi = 300,  width = 12, height = 6)
 
@@ -132,28 +130,11 @@ hexmap <- allocate(
 hexagons <- fortify_hexagon(hexmap, sf_id = "sa3_name_2016", hex_size = 0.7) %>% 
   left_join(st_drop_geometry(sa3lung))
 
-create_polygons <- function(x){
-  polygon <- st_multipoint(as.matrix(bind_rows(x)), dim = "XY") %>% st_cast("POLYGON")
-  return(polygon)
-}
-
-hexpolygons <- hexagons %>% 
-  st_as_sf(., coords = c("long", "lat")) %>% 
-  group_by(sa3_name_2016) %>%
-  st_cast("POLYGON")
-  
-hexagons <- sa3lung %>% 
-  st_drop_geometry() %>% 
-  arrange(sa3_name_2016) %>% 
-  mutate(geometry = st_geometry(hexpolygons$geometry)) %>% 
-  st_as_sf() %>% st_set_crs(., 3112) %>% 
-  ungroup()
-  
-
 aus_gghexmap <- ggplot(hexagons) + 
[…]

### 2019-09-15 usa_plots.R

@@ -53,8 +53,8 @@ ggchoro <- ggplot(cancer) +
   #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
   theme_void() +
   theme(legend.position ="bottom")
-ggsave(filename = "figures/ggchoro.png", device = "png", dpi = 300,  width = 12, height = 6)
 ggchoro
+ggsave(filename = "figures/ggchoro.png", device = "png", dpi = 300,  width = 12, height = 6)
 
 ###############################################################################
 # Cartograms
@@ -181,37 +181,22 @@ ggsave(filename = "figures/ggtilegram.png", plot = ggtilegram,
 cancermap <- st_transform(cancer, "+proj=longlat +datum=WGS84 +no_defs")
 centroids <- create_centroids(cancermap, "NAME")
 grid <- create_grid(centroids = centroids, 
-  hex_size = 2,buffer_dist = 5)
+  hex_size = 2.2 ,buffer_dist = 5)
 hexmap <- allocate(
   centroids = centroids, hex_grid = grid, 
-  sf_id = "NAME", hex_size = 2, hex_filter = 5, 
-  #focal_points = tibble(point = "mean", longitude = mean(centroids$longitude), latitude = mean(centroids$latitude)), 
+  sf_id = "NAME", hex_size = 2.2, hex_filter = 5, 
+  focal_points = tibble(point = "mean", longitude = mean(centroids$longitude), latitude = mean(centroids$latitude)), 
   verbose = TRUE, width = 30
   )
 
-hexagons <- fortify_hexagon(hexmap, sf_id = "NAME", hex_size = 2) %>%  
-  left_join(st_drop_geometry(sa3lung))
-
-create_polygons <- function(x){
-  polygon <- st_multipoint(as.matrix(bind_rows(x)), dim = "XY") %>% st_cast("POLYGON")
-  return(polygon)
-}
-
-hexpolygons <- hexagons %>% 
-  st_as_sf(., coords = c("long", "lat")) %>% 
-  group_by(NAME) %>%
-  st_cast("POLYGON")
-
-hexagons <- cancer %>% 
-  st_drop_geometry() %>% 
-  arrange(NAME) %>% 
-  mutate(geometry = st_geometry(hexpolygons$geometry)) %>% 
-  st_as_sf() %>% st_set_crs(., 3112)
[…]

### 2019-09-15 review_article.Rmd
Grid for all alternative maps, to do: add in code to make these in the Rmd

@@
 
-```{r ggcont, fig.cap = "A contiguous cartogram map of the United States of America.
+```{r ggcartograms, fig.cap = "Alternative maps of the United States of America.
 Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
@@
 The state of California has become much larger due to it's large population density.
-This draws attention to the densely populated North East region, and detracts from the less populated Mid West.", message=FALSE, warning=FALSE, out.width='80%'}
-knitr::include_graphics("figures/ggcont.png")
-```
+This draws attention to the densely populated North East region, and detracts from the less populated Mid West.", message=FALSE, warning=FALSE, out.width='100%'}
+g1 <- png::readPNG("figures/ggcont.png")
+g2 <- png::readPNG("figures/ggncont.png")
+g3 <- png::readPNG("figures/ggdorl.png")
+g4 <- png::readPNG("figures/gghexmap.png")
+#knitr::include_graphics("figures/usa_grid.png")
 
-```{r ggncont, fig.cap = "A Non - contiguous cartogram map of the United States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each state shape has been maintained, but the size has altered according to the population of the state in 2015.
-The state of California has remained closer to it's original size than it's surrounding states.
-The North East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut.
-This draws attention to the densely populated North East region, and the sparse Mid West.", message=FALSE, warning=FALSE, out.width='80%'}
-knitr::include_graphics("figures/ggncont.png")
-```
-
-```{r ggdorl, fig.cap = "A dorling cartogram of the Unites States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each state has been represented by a circle, but the size was determined by the population of the state in 2015.
-The North East states remain closer to their neighbours, and may be displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.", message=FALSE, warning=FALSE, out.width='80%', fig.align='center', fig.width=12}
-knitr::include_graphics("figures/ggdorl.png")
-```
-
-```{r gghexmap, fig.cap = "A hexagon tesselation map of the Unites States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each state has been represented by a square, each square is the same size.
-The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.", message=FALSE, warning=FALSE, out.width='80%'}
-knitr::include_graphics("figures/gghexmap.png")
+library(cowplot)
+ggdraw() +
+  draw_plot(rasterGrob(g1), 0, .5, 0.5, 0.5) +
+  draw_plot(rasterGrob(g2), 0.5, 0.5, 0.5, 0.5) +
+  draw_plot(rasterGrob(g3), 0.0, 0.0, 0.5, 0.5) +
+  draw_plot(rasterGrob(g4), 0.5, 0, 0.5, 0.5)+
+  draw_plot_label(c("a", "b", "c", "d"), 
+    c(0, .5, 0, 0.5), 
+    c(1, 1, .5, .5), size = 15)
 ```

### 2019-09-15 usa_plots.R

@@ -196,7 +196,7 @@ gghexmap <- ggplot(hexagons) +
   geom_polygon(aes(x= long, y = lat, group = NAME, 
                    fill = AgeAdjustedRate), colour = NA) + 
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) +
-  theme_map() + theme(legend.position ="bottom") + coord_equal()
+  theme_void() + theme(legend.position ="bottom") + coord_equal()
 gghexmap
 ggsave(filename = "figures/gghexmap.png", plot = gghexmap,
   device = "png", dpi = 300,  width = 12, height = 6)

### 2019-09-16 review_article.Rmd
add sections 3.1.1 to 3.1.7 with descriptions of maps in Figure 2.

@@
 
-```{r choropleth-grid-create, message=FALSE, fig.cap = "Public facing choropleth cancer maps.[a.
-The Environment and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/) (Lung cancer in females)  (Incidence rates per 100,000 for all cancers, male and female), [b.
-Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx),  [c.
-Atlas of Cancer in QLD - Relative incidence ration, Lung cancer in males](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/), [d.
-Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), [e.
-United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/) (Incidence rate per 100,000, all cancer types, men and women), [f.
-Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) (Relative risk, Lung cancer, men vs women), [g.
-An Atlas of Cancer in South Australia](https://www.cancersa.org.au/assets/images/pdfs/An%20Atlas%20of%20Cancer%20in%20South%20Australia%20-%20Full%20Report.pdf) (Incidence rate per 100,000, lung cancer in females)"}
+```{r choropleth-grid-create, message=FALSE, fig.cap = "Publicly available choropleth cancer maps, published between 2010 and 2015. Sourced from: @roberts2019communication"}
 library(cowplot)
@@
 The use of choropleths within the public domain mirrors the heavy use of choropleth maps within the research literature, discussed above.
-The cancer atlases covered geographies from all around the world, most focused on single nations.
-Figure \@ref(choropleth-grid-create) displays a global (b), national (a,d,e,f) and state (c,g) choropleths.
+The cancer atlases identified by @roberts2019communication covered geographies from all around the world, most focused on single nations.
+Figure \ref{fig:choropleth-grid-create} displays a global (b), national (a,d,e,f) and state (c,g) choropleths.
+The sections below provide details on each of the maps within Figure \ref{fig:choropleth-grid-create}.
 
@@
 Unlike innovation in spatial statistical methods, interactity, data collection and record keeping little work has been done to implement different mapping approaches -->
+
+### The Environment and Health Atlas of England and Wales
+
+Figure 2.a contains [The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/).
+This map shows the relative risk for women developing lunch cancer across England and Wales in 2010.
+The cancer data used to generate this map came from Office for National Statistics (ONS) (England) and from the Welsh Cancer Intelligence and Surveillance Unit (WCISU).
+
+### Globocan 2012: Estimated Cancer Incidence, Mortality and Prevalence Worldwide
+
+The map seen in Figure 2.b is from the [Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx).
+This global map shows age standardized incidence rates (per 100,000) of all cancers for both men and women at national level, for 2018.
+This map is published by the [World Health Organisation's International Agency for Research on Cancer](https://www.iarc.fr/).
+Data was sourced from cancer registries in each country, contributing registries can be seen in the supplementary material on the cancer atlas website.
+
+### Atlas of Cancer in Queensland
+
+Figure 2.c shows the [The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/).
+This map was published by the Queensland Cancer Council and shows the relative incidence ratio of lung cancer in males in the state of QLD within Australia (1998 - 2007).
+Data to generate this map was sourced from the [Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/).
+
+### Bowel Cancer Australia Atlas 
+Figure 2.d shows the [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/).
+Published by [Bowel Cancer Australia](https://www.bowelcanceraustralia.org/) (Australia).
+This map shows the percentage of males between 50 - 54 years of age that diagnosed with bowel cancer in 2016 in Australia.
+The source of the data is not provided.
+
+### United States Cancer Statistics: An Interactive Cancer Statistics Website
+The [United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/) can be seen in Figure 2.e.
+This map contains  the Incidence rate per 100,000, of all cancer types for man and women in the United States for 2016, aggregated at the state level.
+The map was published by the [Centers for Disease Control and Prevention](https://www.cdc.gov/).
+Incidence data seen in this map were compiled from cancer registries meeting U.S.
+Cancer Statistics data quality criteria covering 100% of the U.S.
+population.
+
+<!-- suggested citations: U.S.
+Cancer Statistics Working Group.
+U.S.
+Cancer Statistics Data Visualizations Tool, based on November 2018 submission data (1999-2016): U.S.
+Department of Health and Human Services, Centers for Disease Control and Prevention and National Cancer Institute; www.cdc.gov/cancer/dataviz, June 2019.-->
+
+### Map of Cancer Mortality Rates in Spain
+The [Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) can be seen in Figure 2.f.
+The side by side maps show relative risk of lung cancer for men vs women based on data from 2004 - 2008.
+The source of the data and statistical methods are unknown.
+
+### Atlas of Childhood Cancer in Ontario 
+Figure 2.g shows the [incidence rate of childhood cancers](https://www.pogo.ca/wp-content/uploads/2015/02/POGO_CC-Atlas-3-Incidence_Feb-2015.pdf) per 100,000, by census division, age 0-14, in Ontario, from 1995 to 2004.
+
+<!-- map is on pg 41 or pdf report-->
 
@@
 
-```{r supp-grid-create, fig.cap = "Interactive controls of displays in public facing choropleth cancer maps.
+```{r supp-grid-create, fig.cap = "Interactive controls of displays in publicly available choropleth cancer maps.
 [a.

### 2019-09-18 sa3_plots.R
projected hexagons

@@ -130,6 +130,21 @@ hexmap <- allocate(
 hexagons <- fortify_hexagon(hexmap, sf_id = "sa3_name_2016", hex_size = 0.7) %>% 
   left_join(st_drop_geometry(sa3lung))
 
+hexagons_sf <- hexagons %>% 
+  select(sa3_name_2016, long, lat) %>% 
+  sf::st_as_sf(coords = c("long", "lat"), crs = 4283) %>%
+  group_by(sa3_name_2016) %>% 
+  summarise(do_union = FALSE) %>%
+  st_cast("POLYGON")
+
+# Remove geometry of sa3 geographic areas
+sa3_ng <- sf::st_drop_geometry(sa3lung)
+
+hex <- sa3lung %>% 
+  # ensure correct order by ordering alphabetically
+  arrange(sa3_name_2016) %>% 
+  mutate(geometry = hexagons_sf$geometry)
+
 aus_gghexmap <- ggplot(hexagons) + 
   geom_polygon(aes(x= long, y = lat, group = sa3_name_2016, 
                    fill = `Age-standardised rate (per 100,000)`), colour = NA) + 

### 2019-09-18 review_article.Rmd
Animation paragraph

@@
 
-```{r choropleth-grid-create, message=FALSE, fig.cap = "Publicly available choropleth cancer maps, published between 2010 and 2015. Sourced from: @roberts2019communication"}
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Publicly available choropleth cancer maps, published between 2010 and 2015.
+Sourced from: @roberts2019communication", out.width='100%', fig.align='center', fig.width=12}
 library(cowplot)
@@
 
-The interactivity of modern mapping methods enable supplementary information to be incorporated without cluttering the screen.
-Tool tip features and drop down menus allow for user exploration, small screens are not a problem with the flexibility provided by zooming and panning [@HTLWM].
+Modern mapping methods have the opportunity to incorporate interactive features.
+This enables supplementary information to be incorporated into online atlases without cluttering the screen.
+Tool tip features and drop down menus allow for user exploration with zooming and panning features to allow flexibility in the display [@HTLWM].
+The use of these supports were found in a many online disease map sources.
+This directly enables epidemiologists to explore spatial distributions of cancer statistics.
+Basic interactivity is often implemented by controls outside of the plot space [@TGA],
+this exchanges an image or set of data for another.
+Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve complex selection interactivity.
+
+These interactive features provide an opportunity for users to explore the additional information available.
+This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
 This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accesible' ways [@TNTEA].
-The use of these supports were found in a many online diease mapmap sources.
-These interactive features provide an opportunity for users to explore additional information to help users understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-This directly enables epidemiologists to explore spatial distributions of cancer statistics.
-The interactive features of the publically available maps identified by @roberts2019communication include exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+The interactive features of the publically available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
 @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
@@
 
-Traditionally, epidemiologists as disease mappers would have considered the interactions between the spatial distribution of a diesease, the environment and the people captured by a display.
-Recent developments in technology have created an opportunity to migrate atlases from paper map collections to web atlases.
-The medium chosen to communicate health data should always be driven by the message to communicate and the needs of the audience.
-Where the message is most important, static or animated graphics allow control over the display and interpretation.
-It is important for map creators to develop and then test the message being communicated, epidemiologists are able to do this with their expert knowledge.
-Static visualisations can help to aid the written results provided by experts, and pairing an explanation with a map or diagram can help to draw attention to a point, group or area of interest.
-Static maps do have the caveat of needing to provide enough information for the display to be understandable out of the original context.
-Animations do not tolerate or require user input, this makes them useful for directed descriptions of processes, such as directing attention to a particular region or time period [@HTLWM].
-Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore.
-
-<!-- Animations -->
-@CPISACA provide weather maps as well developed and appropriate examples of spatial displays to communicate to the general public, and encourage those creating disease maps to leverage the success of weather displays.
-The movement of a weather system will follow a forecasted path, the animation will communicate this so that all map users can follow the path of the weather system across the geography over a specified period of time.
-In the disease mapping space, the Australian Cancer Atlas [-@TACA] provides [tours](
-https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
-This directs user's attention and gives them tools to plan their own exploration.
+Epidemiologists consider the interactions between the spatial distribution of a disease, the environment and the people affected.
+Recent developments in technology have created an opportunity to migrate the presentation of epidological results from atlases in paper map collections to web atlases.
+When the map is supplementary to the discussion, static visualisations can help to aid the written results provided by experts, and pairing an explanation with a map or diagram can help to draw attention to a point, group or area of interest.
+Static maps often need to include enough information for the display to be understandable out of the original context.
+There are now many suitable methods and mediums to communicate health data, the choice between them should always be driven by the intended message to communicate and the needs of the audience.
 
@@
 However, with the ability to animate, map creators can specify and answer these questions by directing the attention of a map user to the areas of interest.
-Animations can be used to control the message being communicated by capturing and directing attention of users, it will also ensure all users follow the same interpretive path, as their eyes are drawn across the map to the same comparison points.
 
-<!-- Interactivity-->
+<!-- Animations -->
+Animations provide a linear, fixed narrative to passive users, communicating a message by capturing and directing attention [@TGA].
+@TGA suggests the use of moving objects demands attention by taking advantage of the behaviour of the visual cognition system.
+They can be produced to not tolerate or require user input, creating useful, directed descriptions of processes, such as highlighting a particular region or time period [@HTLWM].
+Animations ensure all users follow the same interpretive path but creators cannnot control the level of engagement.
+
+Animations can be implemented to allow users varying degrees of control. 
+A basic form of interactivty with animations is the play, pause and rewind of films, where the film is intended to be watched in one direction but the position in the animation can be controlled.
+@CPISACA provide weather maps as well developed and appropriate examples of spatial displays to communicate to the general public, and encourage those creating disease maps to leverage the success of weather displays.
+The movement of a weather system will follow a forecasted path, the animation will communicate this so that all map users can follow the path of the weather system across the geography over a specified period of time.
+More complex animations may include steps, when user interaction via clicks animates the display between stages. 
+The Australian Cancer Atlas [-@TACA] provides [tours](
+https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
+This directs user's attention and gives them tools to plan their own exploration.
+
+<!-- Interactivity (Compare movies or tv to video games) -->
 The move from static, print media displays toward interactivity allow users to develop maps to ask and answer their own questions.
+Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore.
+
 Developmets and adoptions of this technology has allowed map users more control of the message displayed.
@@
 Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
+
+Where the message is most important, static or animated graphics allow control over the display and interpretation.
+However before publishing, it is important for map creators to develop and then test the message being communicated, this task requires expert knowledge.
+
+<!--
+A complex directed animation approach is implemented in steps by Google Earth, in their use of stories.
+Within these directed steps users have the opportunity to use interactive controls to change the map directions or watch videos.
+One popular implementation communicates Dr.
+Jane Goodall's life journey using Google Earth displays of significant locations such as her birthplace Bournemouth, England and Tanzania's Gombe National Park.
+-->
 

### 2019-09-19 review_article.Rmd
add first draft of interactivity section

@@
 
-Figure 2.a contains [The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/).
-This map shows the relative risk for women developing lunch cancer across England and Wales in 2010.
+Figure 2.a contains an image from the *[The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/)*.
+This map shows the relative risk for women developing lung cancer in England and Wales in 2010.
 The cancer data used to generate this map came from Office for National Statistics (ONS) (England) and from the Welsh Cancer Intelligence and Surveillance Unit (WCISU).
@@
 
-The map seen in Figure 2.b is from the [Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx).
-This global map shows age standardized incidence rates (per 100,000) of all cancers for both men and women at national level, for 2018.
+The map seen in Figure 2.b is from the *[Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx)*.
+This global map shows age standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018.
 This map is published by the [World Health Organisation's International Agency for Research on Cancer](https://www.iarc.fr/).
@@
 
-Figure 2.c shows the [The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/).
-This map was published by the Queensland Cancer Council and shows the relative incidence ratio of lung cancer in males in the state of QLD within Australia (1998 - 2007).
-Data to generate this map was sourced from the [Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/).
+Figure 2.c shows an extract from the *[The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/)*.
+This map was published by the Queensland Cancer Council and shows the relative incidence ratio of lung cancer in males in the state of QLD within Australia based on data from 1998 to 2007.
+Data to generate this map was sourced from the *[Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/)*.
 
 ### Bowel Cancer Australia Atlas 
-Figure 2.d shows the [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/).
-Published by [Bowel Cancer Australia](https://www.bowelcanceraustralia.org/) (Australia).
-This map shows the percentage of males between 50 - 54 years of age that diagnosed with bowel cancer in 2016 in Australia.
+Figure 2.d shows the *[Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/)*.
+Published by *[Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)* (Australia).
+This map shows the percentage of Australian males between 50 - 54 years of age that diagnosed with bowel cancer in 2016 in Australia.
 The source of the data is not provided.
@@
 ### United States Cancer Statistics: An Interactive Cancer Statistics Website
-The [United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/) can be seen in Figure 2.e.
-This map contains  the Incidence rate per 100,000, of all cancer types for man and women in the United States for 2016, aggregated at the state level.
-The map was published by the [Centers for Disease Control and Prevention](https://www.cdc.gov/).
+The *[United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/)* can be seen in Figure 2.e.
+This map contains  the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level.
+The map was published by the *[Centers for Disease Control and Prevention](https://www.cdc.gov/)*.
 Incidence data seen in this map were compiled from cancer registries meeting U.S.
@@
 ### Map of Cancer Mortality Rates in Spain
-The [Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) can be seen in Figure 2.f.
-The side by side maps show relative risk of lung cancer for men vs women based on data from 2004 - 2008.
+The *[Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html)* can be seen in Figure 2.f.
+The side by side maps show relative risk of lung cancer for men vs women based on data from 2004 to 2008.
 The source of the data and statistical methods are unknown.
@@
 ### Atlas of Childhood Cancer in Ontario 
-Figure 2.g shows the [incidence rate of childhood cancers](https://www.pogo.ca/wp-content/uploads/2015/02/POGO_CC-Atlas-3-Incidence_Feb-2015.pdf) per 100,000, by census division, age 0-14, in Ontario, from 1995 to 2004.
+Figure 2.g shows an extract from the Atlas of Childhood Cancer in Ontario and specifically displays the *[incidence rate of childhood cancers](https://www.pogo.ca/wp-content/uploads/2015/02/POGO_CC-Atlas-3-Incidence_Feb-2015.pdf)* per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004.
 
@@
 
-Modern mapping methods have the opportunity to incorporate interactive features.
-This enables supplementary information to be incorporated into online atlases without cluttering the screen.
-Tool tip features and drop down menus allow for user exploration with zooming and panning features to allow flexibility in the display [@HTLWM].
-The use of these supports were found in a many online disease map sources.
-This directly enables epidemiologists to explore spatial distributions of cancer statistics.
-Basic interactivity is often implemented by controls outside of the plot space [@TGA],
-this exchanges an image or set of data for another.
-Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve complex selection interactivity.
+## Interactivety
+
+One of the concerns of  uncertainty communication to the non-expert is fear of cognitive overload [@mcgranaghan1993cartographic] in which the the user reaches an information threshold, beyond which will not be able to make sense of the information.
+These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
+Interactivety is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
+
+Interactivety enables supplementary information to be incorporated into online atlases without cluttering the display.
+Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
+The use of these supports were found in various online cancer maps identified by @roberts2019communication.
+The controls for basic interactive features are often placed outside of the plot space [@TGA], thus the map image is updated/replaced as the user interacts with the controls.
+For example, changing the population age or other demographic variables.
+Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactivities.
+
+(insert figure of both types of interactivity)
+
+Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
+Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).
 
@@
 This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accesible' ways [@TNTEA].
+This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accesible' ways [@TNTEA].
 The interactive features of the publically available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
 @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
+
+(insert figures showing different interactivity)
+
+<!--
+
+Technological advances may provide a solution to this through the use of interactivity, animation, and sound as additional channels for representing uncertainty information without interfering with the ability to see any features that are present in the display [@maceachren1992visualizing].
+However, if uncertainty information is perceived as integral and inseparable from the data it is associated with, assumptions of uncertainty creating cognitive overload may not be as valid.
+
+-->
 

### 2019-09-19 sa3_plots.R
fix usa and aus figure

@@ -69,6 +69,7 @@ aus_ggchoro
 cont <- sa3lung %>% 
   cartogram_cont(.,
   weight = "Population") %>% st_as_sf()
+save(cont, "data/auscont.rda")
 aus_ggcont <- ggplot(cont) + 
   geom_sf(aes(fill = `Age-standardised rate (per 100,000)`)) + 
   scale_fill_distiller(type = "seq", palette = "Purples",  direction = 1) + 
@@ -76,7 +77,6 @@ aus_ggcont <- ggplot(cont) +
   theme_void() +   theme(legend.position ="bottom")
 aus_ggcont
 ggsave(filename = "figures/aus_ggcont.png", device = "png", dpi = 300,  width = 12, height = 6)
-aus_ggcont
 
 
 ###############################################################################
@@ -120,14 +120,14 @@ ggsave(filename = "figures/aus_ggdorl.png", device = "png", dpi = 300,  width = 
 sa3lungmap <- st_transform(sa3lung, "+proj=longlat +datum=WGS84 +no_defs")
 centroids <- create_centroids(sa3lungmap, "sa3_name_2016")
 grid <- create_grid(centroids = centroids, 
-  hex_size = 0.7,buffer_dist = 5)
+  hex_size = 0.8,buffer_dist = 5)
 hexmap <- allocate(
   centroids = centroids, hex_grid = grid, 
-  sf_id = "sa3_name_2016", hex_size = 0.7, hex_filter = 8, 
+  sf_id = "sa3_name_2016", hex_size = 0.8, hex_filter = 8, 
   focal_points = capital_cities, verbose = TRUE, width = 30
 )
 
-hexagons <- fortify_hexagon(hexmap, sf_id = "sa3_name_2016", hex_size = 0.7) %>% 
+hexagons <- fortify_hexagon(hexmap, sf_id = "sa3_name_2016", hex_size = 0.8) %>% 
   left_join(st_drop_geometry(sa3lung))
 
 hexagons_sf <- hexagons %>% 

### 2019-09-19 usa_plots.R

@@ -68,8 +68,10 @@ ggcont <- ggplot(cont) +
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
   #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
   theme_void() +   theme(legend.position ="bottom")
-ggsave(filename = "figures/ggcont.png", device = "png", dpi = 300,  width = 12, height = 6)
+
+save(cont, file = "data/cont.rda")
 ggcont
+ggsave(filename = "figures/ggcont.png", device = "png", dpi = 300,  width = 12, height = 6)
 
 ###############################################################################
 # Non - Contiguous Cartograms
@@ -91,6 +93,8 @@ ggncont <- ggplot(ncont) +
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
   #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
   theme_void() + theme(legend.position ="bottom")
+
+save(ncont, "data/ncont.rda")
 ggncont
 ggsave(filename = "figures/ggncont.png", device = "png", dpi = 300,  width = 12, height = 6)
 
@@ -192,15 +196,30 @@ hexmap <- allocate(
 hexagons <- fortify_hexagon(hexmap, sf_id = "NAME", hex_size = 2.5) %>%  
   left_join(st_drop_geometry(cancer))
 
-gghexmap <- ggplot(hexagons) + 
-  geom_polygon(aes(x= long, y = lat, group = NAME, 
-                   fill = AgeAdjustedRate), colour = NA) + 
+hexagons_sf <- hexagons %>% 
+  select(NAME, long, lat) %>% 
+  sf::st_as_sf(coords = c("long", "lat"), crs = 4283) %>%
+  group_by(NAME) %>% 
+  summarise(do_union = FALSE) %>%
+  st_cast("POLYGON")
+
+# Remove geometry of sa3 geographic areas
+cancer_ng <- sf::st_drop_geometry(cancer)
+
+hex <- cancer_ng %>% 
+  # ensure correct order by ordering alphabetically
+  arrange(NAME) %>% 
+  mutate(geometry = hexagons_sf$geometry) %>% st_as_sf()
+
+
[…]

### 2019-09-19 review_article.Rmd
cartograms caption

@@
 The map projections alter the shapes and angles of the boundaries of each state.
-Maps a and b are similar in their straight edges, unlike maps c and d which curve on the northern United States border.", message=FALSE, warning=FALSE, out.width='80%', fig.align='center', fig.width=12}
+Maps a and b are similar in their straight edges, unlike maps c and d which curve on the northern United States border.", message=FALSE, warning=FALSE, out.width='100%', fig.align='center', fig.width=12}
 knitr::include_graphics("figures/ggchoroCRS.png")
@@
 
-```{r ggcartograms, fig.cap = "Alternative maps of the United States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each stated shape has been distorted according to the population of the state in 2015.
-The state of California has become much larger due to it's large population density.
-This draws attention to the densely populated North East region, and detracts from the less populated Mid West.", message=FALSE, warning=FALSE, out.width='100%'}
+```{r ggcartograms, fig.cap = "Common alternatives to maps, showing the same information as in Figure \ref{choroCRS} for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexmap (non-contiguous).
+In (a) - (c) the state has been resized, and reshaped, to match the 2015 population of the state.
+This provides a better sense of the extent of disease relative to the population in the country, and can help alleviate losing information about physically small but population dense states.
+In the hexmap (d) each state is given equal size, and thus equal emphasis.", message=FALSE, warning=FALSE, out.width='100%'}
 g1 <- png::readPNG("figures/ggcont.png")

### 2019-09-19 review_article.Rmd
add references and remove word uncertainty

@@
 
-## Interactivety
+## Interactivity
 
-One of the concerns of  uncertainty communication to the non-expert is fear of cognitive overload [@mcgranaghan1993cartographic] in which the the user reaches an information threshold, beyond which will not be able to make sense of the information.
+One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the the user reaches an information threshold, beyond which will not be able to make sense of the information.
 These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
-Interactivety is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
+Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
 
-Interactivety enables supplementary information to be incorporated into online atlases without cluttering the display.
+Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display.
 Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].

### 2019-09-19 review_article.Rmd
interacting and animating switch

--- 2019-09-19 review_article.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE, fig.height = 5)
+knitr::opts_chunk$set(echo = FALSE, fig.height = 5)
 library(ggplot2)
+library(png)
 library(grid)
-library(png)
-library(gridExtra)
 ```
@@
 library(cartogram)
-#knitr::include_graphics("figures/ggchoro.png")
 # read in data
@@
     c(0.015, 0.33, 0.63, 0.015, 0.33, 0.014, 0.63), 
-    c(0.98, 0.98, 0.98, 0.74, 0.74, 0.47, 0.47), size = 15)
+    c(0.98, 0.98, 0.98, 0.74, 0.74, 0.47, 0.47), size = 15) +
+  theme_set(theme_cowplot())
 
@@
 
-Figure 2.a contains an image from the *[The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/)*.
-This map shows the relative risk for women developing lung cancer in England and Wales in 2010.
+Figure 2.a contains [The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/).
+This map shows the relative risk for women developing lunch cancer across England and Wales in 2010.
 The cancer data used to generate this map came from Office for National Statistics (ONS) (England) and from the Welsh Cancer Intelligence and Surveillance Unit (WCISU).
@@
 
-The map seen in Figure 2.b is from the *[Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx)*.
-This global map shows age standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018.
+The map seen in Figure 2.b is from the [Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx).
+This global map shows age standardized incidence rates (per 100,000) of all cancers for both men and women at national level, for 2018.
 This map is published by the [World Health Organisation's International Agency for Research on Cancer](https://www.iarc.fr/).
@@
 
-Figure 2.c shows an extract from the *[The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/)*.
-This map was published by the Queensland Cancer Council and shows the relative incidence ratio of lung cancer in males in the state of QLD within Australia based on data from 1998 to 2007.
-Data to generate this map was sourced from the *[Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/)*.
+Figure 2.c shows the [The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/).
+This map was published by the Queensland Cancer Council and shows the relative incidence ratio of lung cancer in males in the state of QLD within Australia (1998 - 2007).
+Data to generate this map was sourced from the [Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/).
 
 ### Bowel Cancer Australia Atlas 
-Figure 2.d shows the *[Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/)*.
-Published by *[Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)* (Australia).
-This map shows the percentage of Australian males between 50 - 54 years of age that diagnosed with bowel cancer in 2016 in Australia.
+Figure 2.d shows the [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/).
+Published by [Bowel Cancer Australia](https://www.bowelcanceraustralia.org/) (Australia).
+This map shows the percentage of males between 50 - 54 years of age that diagnosed with bowel cancer in 2016 in Australia.
 The source of the data is not provided.
@@
 ### United States Cancer Statistics: An Interactive Cancer Statistics Website
-The *[United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/)* can be seen in Figure 2.e.
-This map contains  the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level.
-The map was published by the *[Centers for Disease Control and Prevention](https://www.cdc.gov/)*.
+The [United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/) can be seen in Figure 2.e.
+This map contains  the Incidence rate per 100,000, of all cancer types for man and women in the United States for 2016, aggregated at the state level.
+The map was published by the [Centers for Disease Control and Prevention](https://www.cdc.gov/).
 Incidence data seen in this map were compiled from cancer registries meeting U.S.
@@
 ### Map of Cancer Mortality Rates in Spain
-The *[Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html)* can be seen in Figure 2.f.
-The side by side maps show relative risk of lung cancer for men vs women based on data from 2004 to 2008.
+The [Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) can be seen in Figure 2.f.
+The side by side maps show relative risk of lung cancer for men vs women based on data from 2004 - 2008.
 The source of the data and statistical methods are unknown.
@@
 ### Atlas of Childhood Cancer in Ontario 
-Figure 2.g shows an extract from the Atlas of Childhood Cancer in Ontario and specifically displays the *[incidence rate of childhood cancers](https://www.pogo.ca/wp-content/uploads/2015/02/POGO_CC-Atlas-3-Incidence_Feb-2015.pdf)* per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004.
+Figure 2.g shows the [incidence rate of childhood cancers](https://www.pogo.ca/wp-content/uploads/2015/02/POGO_CC-Atlas-3-Incidence_Feb-2015.pdf) per 100,000, by census division, age 0-14, in Ontario, from 1995 to 2004.
 
@@
 
-## Interactivety
-
-One of the concerns of  uncertainty communication to the non-expert is fear of cognitive overload [@mcgranaghan1993cartographic] in which the the user reaches an information threshold, beyond which will not be able to make sense of the information.
-These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
-Interactivety is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
-
-Interactivety enables supplementary information to be incorporated into online atlases without cluttering the display.
-Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
-The use of these supports were found in various online cancer maps identified by @roberts2019communication.
-The controls for basic interactive features are often placed outside of the plot space [@TGA], thus the map image is updated/replaced as the user interacts with the controls.
-For example, changing the population age or other demographic variables.
-Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactivities.
-
-(insert figure of both types of interactivity)
-
-Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
-Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).
+Modern mapping methods have the opportunity to incorporate interactive features.
+This enables supplementary information to be incorporated into online atlases without cluttering the screen.
+Tool tip features and drop down menus allow for user exploration with zooming and panning features to allow flexibility in the display [@HTLWM].
+The use of these supports were found in a many online disease map sources.
+This directly enables epidemiologists to explore spatial distributions of cancer statistics.
+Basic interactivity is often implemented by controls outside of the plot space [@TGA],
+this exchanges an image or set of data for another.
+Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve complex selection interactivity.
 
@@
 This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accesible' ways [@TNTEA].
+This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accesible' ways [@TNTEA].
 The interactive features of the publically available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
 @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
-
-(insert figures showing different interactivity)
-
-<!--
-
-Technological advances may provide a solution to this through the use of interactivity, animation, and sound as additional channels for representing uncertainty information without interfering with the ability to see any features that are present in the display [@maceachren1992visualizing].
-However, if uncertainty information is perceived as integral and inseparable from the data it is associated with, assumptions of uncertainty creating cognitive overload may not be as valid.
-
--->
 
@@
     c(0.01, 0.3, 0.3, 0.3, 0.01), 
-    c(0.99, 0.99, 0.57, 0.40, 0.15), size = 15)
+    c(0.99, 0.99, 0.57, 0.40, 0.15), size = 15) + 
+  theme_set(theme_cowplot())
 
@@
 This draws attention to the densely populated North East region, and detracts from the less populated Mid West.", message=FALSE, warning=FALSE, out.width='100%'}
-g1 <- png::readPNG("figures/ggcont.png")
-g2 <- png::readPNG("figures/ggncont.png")
-g3 <- png::readPNG("figures/ggdorl.png")
-g4 <- png::readPNG("figures/gghexmap.png")
-#knitr::include_graphics("figures/usa_grid.png")
+usa1 <- png::readPNG("figures/ggcont.png")
+usa2 <- png::readPNG("figures/ggncont.png")
+usa3 <- png::readPNG("figures/ggdorl.png")
+usa4 <- png::readPNG("figures/gghexmap.png")
 
@@
 ggdraw() +
-  draw_plot(rasterGrob(g1), 0, .5, 0.5, 0.5) +
-  draw_plot(rasterGrob(g2), 0.5, 0.5, 0.5, 0.5) +
-  draw_plot(rasterGrob(g3), 0.0, 0.0, 0.5, 0.5) +
-  draw_plot(rasterGrob(g4), 0.5, 0, 0.5, 0.5)+
+  draw_plot(rasterGrob(usa1), 0, .5, 0.5, 0.5) +
+  draw_plot(rasterGrob(usa2), 0.5, 0.5, 0.5, 0.5) +
+  draw_plot(rasterGrob(usa3), 0.0, 0.0, 0.5, 0.5) +
+  draw_plot(rasterGrob(usa4), 0.5, 0, 0.5, 0.5)+
   draw_plot_label(c("a", "b", "c", "d"), 
     c(0, .5, 0, 0.5), 
-    c(1, 1, .5, .5), size = 15)
+    c(1, 1, .5, .5), size = 15) +
+  theme_set(theme_cowplot())
 ```
 
-Table: \label{tab:usa} Maps used to present statistics for the Uited States of America
+Table: \label{tab:usa} Maps used to present statistics for the United States of America
 
@@
 The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.|
+
+```{r auscartograms, fig.cap = "Alternative maps of the United States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each stated shape has been distorted according to the population of the state in 2015.
+The state of California has become much larger due to it's large population density.
+This draws attention to the densely populated North East region, and detracts from the less populated Mid West.", message=FALSE, warning=FALSE, out.width='100%'}
+aus1 <- png::readPNG("figures/aus_ggcont.png")
+aus2 <- png::readPNG("figures/aus_ggncont.png")
+aus3 <- png::readPNG("figures/aus_ggdorl.png")
+aus4 <- png::readPNG("figures/aus_gghexmap.png")
+
+library(cowplot)
+ggdraw() +
+  draw_plot(rasterGrob(aus1), 0, .5, 0.5, 0.5) +
+  draw_plot(rasterGrob(aus2), 0.5, 0.5, 0.5, 0.5) +
+  draw_plot(rasterGrob(aus3), 0.0, 0.0, 0.5, 0.5) +
+  draw_plot(rasterGrob(aus4), 0.5, 0, 0.5, 0.5)+
+  draw_plot_label(c("a", "b", "c", "d"), 
+    c(0, .5, 0, 0.5), 
+    c(1, 1, .5, .5), size = 15) +
+  theme_set(theme_cowplot())
+```
 
@@
 
-# Animation and Interactivity {#ch:interacting}
+# Interactivity and Animation  {#ch:interacting}
+<!--
+Recent developments in technology have created an opportunity to migrate the presentation of epidological results from atlases in paper map collections to web atlases.
+Pairing an explanation with a map or diagram can help to draw attention to a point, group or area of interest.
+Maps may form supplementary material to the discussion, in this case static visualisations can help to aid written results. 
+Static maps often need to include enough information for the display to be understandable out of the original context.
+-->
 
-Epidemiologists consider the interactions between the spatial distribution of a disease, the environment and the people affected.
-Recent developments in technology have created an opportunity to migrate the presentation of epidological results from atlases in paper map collections to web atlases.
-When the map is supplementary to the discussion, static visualisations can help to aid the written results provided by experts, and pairing an explanation with a map or diagram can help to draw attention to a point, group or area of interest.
-Static maps often need to include enough information for the display to be understandable out of the original context.
-There are now many suitable methods and mediums to communicate health data, the choice between them should always be driven by the intended message to communicate and the needs of the audience.
+There are now many suitable methods and mediums to communicate health data.
+The choice between them should always be driven by the intended message to be communicated and the needs of the audience.
 
-@CPISACA provide three questions that are generally when using a disease map:
+@CPISACA provide three questions that are generally asked when using a disease map:
 
@@
 
-A single map display historically created by epidemiologists, may not effectively answer all three at the same time.
+A single map display may not effectively answer all three at the same time.
 This is especially true when areas are small and difficult to identify in a display.
-However, with the ability to animate, map creators can specify and answer these questions by directing the attention of a map user to the areas of interest.
+However, with the ability to interact with a map, or animate the display, map creators can specify and answer these questions by directing the attention of a map user to the areas of interest.
+
+<!-- Interactivity (Compare movies or tv to video games) -->
+The progress from static print media displays to interactive methods has allowed users to develop maps to ask and answer their own questions.
+Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore. 
+This oppportunity provides an alternative to the traditional approach of a single map for a single use, into a more explorative territory.
+Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
+
+Developmets and adoptions of this technology has allowed map users more control of the message displayed.
+Figure \ref{fig:supp-grid-create} presents the control menus for several online cancer atlases.
+Users can select combinations of cancer type, gender, and other factors to answers questions relevant to them and their work.
+Users can use these controls to change the layer of geography shown, such as continents or countries, states or counties.
+Thess display can show familiar shapes for users to orient themselves on the map, then zoom in to explore densely populated communities.
 
@@
 More complex animations may include steps, when user interaction via clicks animates the display between stages.
+This approach is implemented by Google Earth 
+Where the message is most important, static or animated graphics allow control over the display and interpretation.
 The Australian Cancer Atlas [-@TACA] provides [tours](
@@
 This directs user's attention and gives them tools to plan their own exploration.
-
-<!-- Interactivity (Compare movies or tv to video games) -->
-The move from static, print media displays toward interactivity allow users to develop maps to ask and answer their own questions.
-Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore.
-
-Developmets and adoptions of this technology has allowed map users more control of the message displayed.
-Users are able to see the familiar shapes, zoom in to explore densely populated communities, or change the display completely to highlight the population space.
-
-This oppportunity provides room to move away from the traditional approach of a single map for a single use, into a more explorative territory.
-Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
-
-Where the message is most important, static or animated graphics allow control over the display and interpretation.
-However before publishing, it is important for map creators to develop and then test the message being communicated, this task requires expert knowledge.
 
@@
 
+Several R [@R] packages were used to produce this paper.
+For data analysis the tidyverse package [@tidyverse] provided many useful functions.
+ggplot2 package [@ggplot2] was used to create maps, with RColorBrewer package [@RColorBrewer] providing additional colour palettes and ggthemes [@ggthemes] package providing themes.
+The png package [@png] was used to access png images taken from online web atlases, cowplot package [@cowplot] was used to arrange these plots into grouped images.
+
+The following packages were used to create tranformations from the sf [@sf] geographical shapes of the states of America [@spData] and the XXX Australian Statistical Areas (Level 3):
+The cartogram [@cartogram] package creates contiguous, non-contiguous and Dorling cartgorams.
+Hexagon tilegram displays were created using the sugarbag package [@sugarbag].
+
 # References
-
-<!--
-# Software
-Commonly available tools to implement.
--->
-@tidyerse for data analysis, @spData and @sf to implement plotting of spatial data.
-@sugarbag to create hexagon tessellation. @ggthemes to enhance plot displays.
-@cartogram for contiguous and non-contiguous cartogram displays.
 

### 2019-09-19 review_article.Rmd
move interactivity 3.2 to 3.4
steff

@@
 usa3 <- png::readPNG("figures/ggdorl.png")
-usa4 <- png::readPNG("figures/gghexmap.png")
+#usa4 <- png::readPNG("figures/gghexmap.png")
 
@@
   draw_plot(rasterGrob(usa3), 0.0, 0.0, 0.5, 0.5) +
-  draw_plot(rasterGrob(usa4), 0.5, 0, 0.5, 0.5)+
+  #draw_plot(rasterGrob(usa4), 0.5, 0, 0.5, 0.5)+
   draw_plot_label(c("a", "b", "c", "d"), 

### 2019-09-19 review_article.Rmd

--- 2019-09-19 review_article.Rmd
@@
     c(0, .5, 0, 0.5), 
-    c(1, 1, .5, .5), size = 15)
+    c(1, 1, .5, .5), size = 15) +
+  theme_set(theme_cowplot())
 ```
 
-Table: \label{tab:usa} Maps used to present statistics for the Uited States of America
+Table: \label{tab:usa} Maps used to present statistics for the United States of America
 

### 2019-09-19 review_article.Rmd
no hexmap fig

@@
 g3 <- png::readPNG("figures/ggdorl.png")
-g4 <- png::readPNG("figures/gghexmap.png")
+#g4 <- png::readPNG("figures/gghexmap.png")
 #knitr::include_graphics("figures/usa_grid.png")
@@
   draw_plot(rasterGrob(g3), 0.0, 0.0, 0.5, 0.5) +
-  draw_plot(rasterGrob(g4), 0.5, 0, 0.5, 0.5)+
+  #draw_plot(rasterGrob(g4), 0.5, 0, 0.5, 0.5)+
   draw_plot_label(c("a", "b", "c", "d"), 

### 2019-09-19 review_article.Rmd
removed tilegram figure

@@
 
+<!--
 ```{r ggtilegram, fig.cap = "A tilegram of the Unites States of America.
@@
 ```
+-->
 

### 2019-09-19 review_article.Rmd
inetractive graphics references

@@
 
-# Cartograms and other spatially warped displays {#ch:alternatives}
+# Alternatives to choropleth maps {#ch:alternatives}
+
+## Micromaps
+XXX Something needs to be said about micromaps https://cran.r-project.org/web/packages/micromapST/vignettes/JSSmicromapSTv63i03.pdf
 
@@
 
+XXX Define interactivity, and also animation
+
 There are now many suitable methods and mediums to communicate health data.
@@
 
+- the arcGIS standard https://pro.arcgis.com/en/pro-app/help/mapping/animation/overview-of-animation.htm
+- very general infovis reference that refers to Buja, Cook, Swayne distinguishing between direct manipulation and interactive graphics https://hal.inria.fr/tel-01096366/file/VA2_PERIN_CHARLES_17112014.pdf
+- explanation of inetractivity and direct manipulation: https://books.google.com.au/books?id=A6TMqxkW-EoC&pg=PA6&lpg=PA6&dq=direct+manipulation+%22swayne%22+%22klinke%22&source=bl&ots=tXYKP8JbbV&sig=ACfU3U3KhN-_YNt_UV5SmThsl4D2gY9A4g&hl=en&sa=X&ved=2ahUKEwj6u9vgqtzkAhVKeysKHQ6iByUQ6AEwA3oECAgQAQ#v=onepage&q=direct%20manipulation%20%22swayne%22%20%22klinke%22&f=false
+- might be useful to look at https://towardsdatascience.com/a-complete-guide-to-an-interactive-geographical-map-using-python-f4c5197e23e0
+- a classic http://www.cs.umd.edu/hcil/trs/2001-08/2001-08.html
+- this is one reference that maybe yoou have missed: https://alsino.io/static/papers/BeyondChoropleths_AlsinoSkowronnek.pdf
+
 # Acknowledgements

### 2019-09-21 review_article.Rmd
figures

--- 2019-09-19 review_article.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, fig.height = 5)
+knitr::opts_chunk$set(echo = FALSE)
 library(ggplot2)
@@
 A choropleth is constructed by drawing the geographic or political boundaries, and filling the shapes with colours to represent values of a measured variable [@EI].
-Figure \ref{fig:choro} shows a choropleth of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged over 2012 through 2016.
+Figure \ref{fig:choroCRS} shows a choropleth of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged over 2012 through 2016.
 The data was extracted from the official federal statistics by @usa_lung on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
@@
 Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS].
-In Figure \ref{fig:choro}, a west to east spatial trend of increasing rates, can be seen.
+In Figure \ref{fig:choroCRS}, a west to east spatial trend of increasing rates, can be seen.
 There is also a spatial outlier -- Utah has a noticeably lower rate that its neighbours.
@@
 There is something of a cluster of higher rates around the tobacco states.
-
-```{r choro, fig.cap = "Choropleth map of the average age-adjusted rate per 100,000 people, of new cases of lung and bronchus cancer 2012-2016 in the USA.
-Utah has an unusually low rate and Kentucky is high.
-There is generally a west to east trend, with higher rates towards the easern states.", message=FALSE, warning=FALSE, out.width='100%', fig.align='center', fig.width=12}
-library(sugarbag)
-library(tidyverse)
-library(spData)
-library(ggthemes)
-library(maptools)
-library(sf)
-library(cartogram)
-# read in data
-cancer <- read_csv("data/USCSlung.csv") %>% rename(NAME = Area)
-
-# Join polygons to data
-data(us_states)
-cancer <- left_join(us_states, cancer, by = c("NAME"))
-cancer <- st_transform(cancer, 3857)
-
-b <- st_bbox(cancer)
-ggplot(cancer) + 
-  geom_sf(aes(fill = AgeAdjustedRate), colour = NA) +
-  scale_fill_distiller("Age-adjusted rate \n (per 100,000)", type = "seq", palette = "RdPu",  direction = 1) + 
-  #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
-  theme_void() +
-  theme(legend.position ="bottom", legend.key.width = unit(2, "cm"))
-```
 
@@
 
-```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Publicly available choropleth cancer maps, published between 2010 and 2015.
-Sourced from: @roberts2019communication", out.width='100%', fig.align='center', fig.width=12}
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Publicly available choropleth cancer maps, published between 2010 and 2015.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
 library(cowplot)
@@
     c(0.015, 0.33, 0.63, 0.015, 0.33, 0.014, 0.63), 
-    c(0.98, 0.98, 0.98, 0.74, 0.74, 0.47, 0.47), size = 15) +
-  theme_set(theme_cowplot())
+    c(0.99, 0.99, 0.99, 0.74, 0.74, 0.47, 0.47), size = 20)
 
@@
 
-(insert figure of both types of interactivity)
-
-Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
-Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).
-
-These interactive features provide an opportunity for users to explore the additional information available.
-This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accesible' ways [@TNTEA].
-The interactive features of the publically available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
-@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
-
-(insert figures showing different interactivity)
-
-<!--
-
-Technological advances may provide a solution to this through the use of interactivity, animation, and sound as additional channels for representing uncertainty information without interfering with the ability to see any features that are present in the display [@maceachren1992visualizing].
-However, if uncertainty information is perceived as integral and inseparable from the data it is associated with, assumptions of uncertainty creating cognitive overload may not be as valid.
-
--->
-
-XXX more details on the examples in the next figure
-
-## Common statistics displayed
-
-Cancer maps are powerful visualisations that summarise complex statistical analyses, however the statistics represented in these maps cannot tell the entire story.
-Supplementary graphs and plots are often included to add more depth and information to the map.
-@CPISACA suggests additional materials such as tables, graphs, and text explanations support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers.
-There are many visualisations used for displays of statistical summaries, these may be dot plots, bar plots, box plots, timesseries plots, cumulative distribution plots, scatter plots, Q-Q plots. 
-These additional displays of the cancer distribution can provide alternative views of the cancer statistics, as well as the supporting statistics including error, confidence intervals, distributions, sample or population sizes, standard deviation and other measures.
-When presenting cancer maps, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses', the supplementary statistics help for this validation.
-
-The interactivity of modern mapping methods enable supplementary information to be incorporated without cluttering the screen.
-Tool tip features and drop down menus allow for user exploration, small screens are not a problem with the flexibility provided by zooming and panning [@HTLWM].
-This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accesible' ways [@TNTEA].
-The use of these supports were found in a many online diease mapmap sources.
-These interactive features provide an opportunity for users to explore additional information to help users understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-This directly enables epidemiologists to explore spatial distributions of cancer statistics.
-The interactive features of the publically available maps identified by @roberts2019communication include exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
-@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
-
 ```{r supp-grid-create, fig.cap = "Interactive controls of displays in publicly available choropleth cancer maps.
@@
 Global Cancer Map](http://globalcancermap.com/) (Controls for indicator, and cancer type), [e.
-United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html) (Controls for demographics, trends, geography, indicators, risk factors, sex, cancer type, and year)."}
+United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html) (Controls for demographics, trends, geography, indicators, risk factors, sex, cancer type, and year).", fig.width = 12, fig.height = 9}
 i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png') # b in choro grid
@@
     c(0.01, 0.3, 0.3, 0.3, 0.01), 
-    c(0.99, 0.99, 0.57, 0.40, 0.15), size = 15) + 
-  theme_set(theme_cowplot())
+    c(0.99, 0.99, 0.57, 0.40, 0.15), size = 20)
 
 ggsave(filename = "figures/supp_grid.png", device = "png", 
-  width = 6, height = 5, dpi = 300)
+  width = 6, height = 6, dpi = 300)
 ```
+
+The interactivity of modern mapping methods enable supplementary information to be incorporated without cluttering the screen.
+Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
+Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).
+
+These interactive features provide an opportunity for users to explore the additional information available.
+Tool tip features and drop down menus allow for user exploration, small screens are not a problem with the flexibility provided by zooming and panning [@HTLWM].
+This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accesible' ways [@TNTEA].This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
+This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accesible' ways [@TNTEA].
+The interactive features of the publically available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
+
+<!--
+
+Technological advances may provide a solution to this through the use of interactivity, animation, and sound as additional channels for representing uncertainty information without interfering with the ability to see any features that are present in the display [@maceachren1992visualizing].
+However, if uncertainty information is perceived as integral and inseparable from the data it is associated with, assumptions of uncertainty creating cognitive overload may not be as valid.
+
+-->
+
+## Common statistics displayed
+
+Cancer maps are powerful visualisations that summarise complex statistical analyses, however the statistics represented in these maps cannot tell the entire story.
+Supplementary graphs and plots are often included to add more depth and information to the map.
+@CPISACA suggests additional materials such as tables, graphs, and text explanations support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers.
+There are many visualisations used for displays of statistical summaries, these may be dot plots, bar plots, box plots, timesseries plots, cumulative distribution plots, scatter plots, Q-Q plots. 
+These additional displays of the cancer distribution can provide alternative views of the cancer statistics, as well as the supporting statistics including error, confidence intervals, distributions, sample or population sizes, standard deviation and other measures.
+When presenting cancer maps, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses', the supplementary statistics help for this validation.
 
@@
 The areas are often ranked and allocated to quintiles, each quintile can be presented as categories describing the ranking.
-
-## Interaction with atlases
 
@@
 g3 <- png::readPNG("figures/ggdorl.png")
-#g4 <- png::readPNG("figures/gghexmap.png")
+g4 <- png::readPNG("figures/gghexmap.png")
 #knitr::include_graphics("figures/usa_grid.png")
+
+load("usa_legend.rda")
 
@@
 ggdraw() +
-  draw_plot(rasterGrob(g1), 0, .5, 0.5, 0.5) +
-  draw_plot(rasterGrob(g2), 0.5, 0.5, 0.5, 0.5) +
-  draw_plot(rasterGrob(g3), 0.0, 0.0, 0.5, 0.5) +
-  #draw_plot(rasterGrob(g4), 0.5, 0, 0.5, 0.5)+
+  draw_plot(rasterGrob(g1), 0, .5, 0.5, 0.45) +
+  draw_plot(rasterGrob(g2), 0.5, 0.5, 0.5, 0.45) +
+  draw_plot(rasterGrob(g3), 0.0, 0.0, 0.5, 0.45) +
+  draw_plot(rasterGrob(g4), 0.5, 0, 0.5, 0.45) +
+  draw_plot(usa_legend, 0, 0, 1, 0.1) +
   draw_plot_label(c("a", "b", "c", "d"), 
     c(0, .5, 0, 0.5), 
-    c(1, 1, .5, .5), size = 15) +
-  theme_set(theme_cowplot())
+    c(1, 1, .5, .5), size = 20)
 ```
@@
     c(0, .5, 0, 0.5), 
-    c(1, 1, .5, .5), size = 15) +
-  theme_set(theme_cowplot())
+    c(1, 1, .5, .5), size = 20)
 ```
@@
 Each area is represented by a square of the same dimensions.
-Figure \ref{fig:ggtilegram} shows a square tilegram representing each state using a single square.
+Figure \ref{fig:ggcartograms} d shows a hexagon tilegram representing each state using a single square.
 
@@
 ## Tilegram
+
+A tilegram provides a tesselated display of consistent shapes.
+Each tile is usally one unit of measurement, this could be geographic regions such as states, or population based where one tile is used for a consistent measure of population.
+<!-- cite mosiac cartograms --> 
+<!--XXX cite geogrid here?-->
+A simple tilegram may present the states in a tesselated grid display, where geographic neighbours are found next to each other, with some neccessary displacement employed for regions with more than four neighbours.
+These tiles may be labelled or coloured to represent a value.
+A data visualisation can be geographically faceted, so that a smaller visualisation of equal size is created for each geographic subset of the data.
+This requires grid displays that are often created once by hand and reused.
 
@@
 When there is a relationship between the cancer type being mapped and geographical features, a choropleth map would pair this information visually.
+However these displays are limited to univariate displays, usually presented by filling each area with a colour.
 These maps are best used when the geography is of importance to the data in the display.
+They favor geographically large regions and small regions are often represented by inset displays that zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choroleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
 Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbours.
@@
 <!-- Cartograms, sacrifice geography, focus on the outliers/ relationships in data space -->
-Pairing geography and cancer data was historically the work of epidemiologists, however it did not require them to be the map designer.
-This was expected as cartographers were the creators of geographic displays, and epidemiologists utilised the maps that had been prepared in advance.
-Cartograms with the population variable used for base map distortion can be seen as density equalising maps, that reduce the visual prominence of low-population areas [@VSSDCUC].
-This approach began a shift in the focus and responsibility of map creators.
-Cartogram displays began incorporating the statistics and population into the design of the display. 
-Shifting and sometimes sacrificing familiar boundaries to draw attention to outliers in the data space or population density.
+Cartographers were historically the creators of geographic displays, and epidemiologists utilised the maps that had been prepared in advance.
+Cartograms with the population variable used to distortion the geographic map can be used as density equalising maps, that reduce the visual prominence of low-population areas [@VSSDCUC]. 
+Cartogram displays incorporate the statistics and population into the design of the display,
+shifting and sometimes sacrificing familiar boundaries to draw attention to outliers in the data space or population density.
+The unusual shapes are difficult to compare and can prevent comparison of the statistic.
+Cartograms also favour extreme statistics in the positive direction.
+Small extreme values are often lost due to the resulting small sized areas the do not draw user's attention.
+The change in the map can also provide opportunities for bivariate displays, using the change in area for one variable, and colour for another.
 The difference between the familiar map display helps highlight the impact of the disease on communities.
@@
 <!-- Alternatives, remove geographic shapes, keep spatial relationship, focus on the outliers/ relationships in population space -->
-By removing the unexpected boundaries and shapes, the spatial distribution is the primary concept communicated by the display.
-Areas that are geographically close will maintain connectedness in some way, but the population and the statistic will take precedence.
+Alternative maps may replace the unexpected boundaries and shapes with familiar, simple shapes. This makes the spatial distribution the primary concept communicated by the display.
+Areas that are geographically close will maintain connectedness in some way, but the population and the statistic will dominate.
+Tilegram displays may only allow one variable to be visualised if each region is given consistent map space, this occurs in geofaceting. 
+Dorlings provide bivariate displays as they use the size of the shape for one variable, and colour for another.
 These alternatives are especially helpful for data aggregations where administrative boundaries break populations into groups.
-The experience of each collective may be worth considering as their experiences may be similar in terms of the services and facilities they share.
-Population affects are highlighted by these displays, as this display allows a more equitable view of each community, and does not minimise those that operate on a smaller geographic scale.
+The experience of each community may be worth considering, as the experiences shared by the population within them may be similar due to the services and facilities they share.
+This display allows a more equitable view of each community, and does not minimise those that operate on a smaller geographic scale.
+
+Faceting provides an opportunity to contrast the overall distribution of variables across a geographic space or collection of geographic areas.
+This display makes is difficult to compare specific regions.
 

### 2019-09-21 sa3_plots.R

@@ -36,12 +36,14 @@ cimar_sa3 <- cimar_sa3 %>%
 
 # Join with sa3 sf object
 sa3lung <- absmapsdata::sa32016 %>% 
-  left_join(cimar_sa3, by = c("sa3_name_2016"= "Name")) %>%
+  left_join(cimar_sa3, by = c("sa3_name_2016"= "Name")) %>% 
   st_as_sf() %>%
   filter(!is.na(Population)) %>%
+  #mutate(Population = ifelse(is.na(Population), 1, Population)) %>% 
   filter(!st_is_empty(geometry))
 
 sa3lung <- st_transform(sa3lung, 3112)
+#%>%  st_simplify(., preserveTopology = TRUE,  dTolerance = 0) 
 
 b <- st_bbox(sa3lung)
 
@@ -58,26 +60,30 @@ aus_ggchoro <- ggplot(sa3lung) +
   coord_sf(crs = CRS("+init=epsg:3112"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
   theme(legend.position ="bottom") + 
   labs(fill = "Age-standardised rate\n(per 100,000)")
-ggsave(filename = "figures/aus_ggchoro.png", device = "png", dpi = 300,  width = 12, height = 6)
 aus_ggchoro
+ggsave(filename = "figures/aus_ggchoro.png", device = "png", dpi = 300,  width = 7, height = 6)
 
 
 ###############################################################################
 # Cartograms
-
+  
 # Contiguous Cartograms
 cont <- sa3lung %>% 
   cartogram_cont(.,
-  weight = "Population") %>% st_as_sf()
+  weight = "Population", itermax = 15) %>%
+  st_as_sf()
+
 save(cont, "data/auscont.rda")
 aus_ggcont <- ggplot(cont) + 
   geom_sf(aes(fill = `Age-standardised rate (per 100,000)`)) + 
   scale_fill_distiller(type = "seq", palette = "Purples",  direction = 1) + 
   coord_sf(crs = CRS("+init=epsg:3112"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
-  theme_void() +   theme(legend.position ="bottom")
+  theme_void()+guides(fill=FALSE)
 aus_ggcont
-ggsave(filename = "figures/aus_ggcont.png", device = "png", dpi = 300,  width = 12, height = 6)
[…]

### 2019-09-21 usa_plots.R

@@ -30,8 +30,6 @@ cancer <- read_csv("data/USCSlung.csv") %>% rename(NAME = Area)
 
 
 # Join polygons to data
-# devtools::install_github("hrbrmstr/albersusa")
-library(albersusa)
 cancer <- left_join(us_states, cancer, by = c("NAME"))
 cancer <- st_transform(cancer, 3857)
 
@@ -54,7 +52,10 @@ ggchoro <- ggplot(cancer) +
   theme_void() +
   theme(legend.position ="bottom")
 ggchoro
-ggsave(filename = "figures/ggchoro.png", device = "png", dpi = 300,  width = 12, height = 6)
+#ggsave(filename = "figures/ggchoro.png", device = "png", dpi = 300, width = 7, height = 6)
+
+usa_legend <- get_legend(ggchoro)
+save(usa_legend, file = "usa_legend.rda")
 
 ###############################################################################
 # Cartograms
@@ -67,11 +68,11 @@ ggcont <- ggplot(cont) +
   ggtitle(title) +
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
   #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
-  theme_void() +   theme(legend.position ="bottom")
+  theme_void() + guides(fill = FALSE)
 
 save(cont, file = "data/cont.rda")
 ggcont
-ggsave(filename = "figures/ggcont.png", device = "png", dpi = 300,  width = 12, height = 6)
+ggsave(filename = "figures/ggcont.png", device = "png", dpi = 300, width = 7, height = 6)
 
 ###############################################################################
 # Non - Contiguous Cartograms
@@ -92,11 +93,11 @@ ggncont <- ggplot(ncont) +
   ggtitle(title) + 
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
   #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
-  theme_void() + theme(legend.position ="bottom")
+  theme_void() + guides(fill = FALSE)
 
 save(ncont, "data/ncont.rda")
 ggncont
[…]

### 2019-09-21 review_article.Rmd

@@
 
+One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the the user reaches an information threshold, beyond which will not be able to make sense of the information.
+These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
+Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
+
+Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display.
+Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
+The use of these supports were found in various online cancer maps identified by @roberts2019communication.
+The controls for basic interactive features are often placed outside of the plot space [@TGA], thus the map image is updated/replaced as the user interacts with the controls.
+For example, changing the population age or other demographic variables.
+Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactivities.
+
+(insert figure of both types of interactivity)
+
+Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
+Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).
+
+These interactive features provide an opportunity for users to explore the additional information available.
+This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
+This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accesible' ways [@TNTEA].
+The interactive features of the publically available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
+
+(insert figures showing different interactivity)
+
+<!--
+
+Technological advances may provide a solution to this through the use of interactivity, animation, and sound as additional channels for representing uncertainty information without interfering with the ability to see any features that are present in the display [@maceachren1992visualizing].
+However, if uncertainty information is perceived as integral and inseparable from the data it is associated with, assumptions of uncertainty creating cognitive overload may not be as valid.
+
+-->
+
 # Cartograms and other spatially warped displays {#ch:alternatives}
+
+## Micromaps
+XXX Something needs to be said about micromaps https://cran.r-project.org/web/packages/micromapST/vignettes/JSSmicromapSTv63i03.pdf
 

### 2019-09-22 review_article.Rmd
improved plots and interactivity sections

--- 2019-09-21 review_article.Rmd
@@
 
-```{r supp-grid-create, fig.cap = "Interactive controls of displays in publicly available choropleth cancer maps.
+```{r interacting, fig.cap = "Interactive controls of displays in publicly available choropleth cancer maps.
 [a.
@@
 
+One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the the user reaches an information threshold, beyond which will not be able to make sense of the information.
+These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
+Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
+
+Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display.
+Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
+The use of these supports were found in various online cancer maps identified by @roberts2019communication.
+The controls for basic interactive features are often placed outside of the plot space [@TGA], thus the map image is updated/replaced as the user interacts with the controls.
+For example, changing the population age or other demographic variables.
+Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactivities.
+
+(insert figure of both types of interactivity)
+
+Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
+Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).
+
+These interactive features provide an opportunity for users to explore the additional information available.
+This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
+This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accesible' ways [@TNTEA].
+The interactive features of the publically available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
+
+<!--
+
+Technological advances may provide a solution to this through the use of interactivity, animation, and sound as additional channels for representing uncertainty information without interfering with the ability to see any features that are present in the display [@maceachren1992visualizing].
+However, if uncertainty information is perceived as integral and inseparable from the data it is associated with, assumptions of uncertainty creating cognitive overload may not be as valid.
+
+-->
+
 # Cartograms and other spatially warped displays {#ch:alternatives}
+
+## Micromaps
+
+Choropleths are univariate displays, they do not inherently support multi-dimensionality. This makes it difficult to pair demographic, environmental and other factors with spatial distributions.
+@MMST present linked micromap plots as a solution to the issue of multi-dimensionality when visually linking geographic and statistical data.
+This design is widely used in online atlas displays, it groups areas of similar values for one variable, and uses the additional columns to provide displays that contrast the states in each group by other variables.
+The display juxtaposes of choropleth maps and statistical plots.
+Showing one map per group of the key separating variable, in a row with each additional statistical plot.
+
+<!--
+Visualizing Data Patterns with Micromaps.
+Two new templates for epidemiology applications: linkedmicromap plots and conditioned choropleth maps-->
+
+Linked micromaps predominantly utilise the choropleth map for displays of spatial distributions.
+It is one of several alternative displays that allow maps to become bi-variate displays.
 
@@
 The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.|
-|Hexagon Tesselation|A hexagon tesselation map of the Unites States of America.
+|Hexagon Tessellation|A hexagon tessellation map of the Unites States of America.
 Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
@@
 aus4 <- png::readPNG("figures/aus_gghexmap.png")
+load("aus_legend.rda")
 
@@
 ggdraw() +
-  draw_plot(rasterGrob(aus1), 0, .5, 0.5, 0.5) +
-  draw_plot(rasterGrob(aus2), 0.5, 0.5, 0.5, 0.5) +
-  draw_plot(rasterGrob(aus3), 0.0, 0.0, 0.5, 0.5) +
-  draw_plot(rasterGrob(aus4), 0.5, 0, 0.5, 0.5)+
+  draw_plot(rasterGrob(aus1), 0, .55, 0.5, 0.45) +
+  draw_plot(rasterGrob(aus2), 0.5, 0.55, 0.5, 0.45) +
+  draw_plot(rasterGrob(aus3), 0.0, 0.1, 0.5, 0.45) +
+  draw_plot(rasterGrob(aus4), 0.5, 0.1, 0.5, 0.45) +
+  draw_plot(aus_legend, 0, 0, 1, .1) +
   draw_plot_label(c("a", "b", "c", "d"), 
@@
 This rectangular display also allows for tiling, where geographic neighbours placed in suitable relative positions also share borders, however contiguity may be sacrificed [@CDWCS].
-Rectangular cartograms allow for bivariate displays, population can be effectively communicated by the size of each rectangular, and a second variable can be communicated using colour [@ORC].
+Rectangular cartograms allow for bi-variate displays, population can be effectively communicated by the size of each rectangular, and a second variable can be communicated using colour [@ORC].
 
-A similar method, where each geographic area is represented using a square, tesselated to create a square grid.
+A similar method, where each geographic area is represented using a square, tessellated to create a square grid.
 This method has been used by FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
@@
 
-Recommended criteria to contrast mapping methods include average cartographic error, and maximum cartographic error, correct adjacencies, maximum aspect ratio, and suitable relative positions [@ORC].
+Recommended criteria to contrast mapping methods include average cartographic error, and maximum cartographic error, correct adjacency, maximum aspect ratio, and suitable relative positions [@ORC].
 However, this does not consider the issues with actually producing rectangular cartograms.
@@
 
-A tilegram provides a tesselated display of consistent shapes.
-Each tile is usally one unit of measurement, this could be geographic regions such as states, or population based where one tile is used for a consistent measure of population.
-<!-- cite mosiac cartograms --> 
+A tilegram provides a tessellated display of consistent shapes.
+Each tile is usually one unit of measurement, this could be geographic regions such as states, or population based where one tile is used for a consistent measure of population.
+<!-- cite mosaic cartograms --> 
 <!--XXX cite geogrid here?-->
-A simple tilegram may present the states in a tesselated grid display, where geographic neighbours are found next to each other, with some neccessary displacement employed for regions with more than four neighbours.
+A simple tilegram may present the states in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
 These tiles may be labelled or coloured to represent a value.
@@
 When there is a relationship between the cancer type being mapped and geographical features, a choropleth map would pair this information visually.
-However these displays are limited to univariate displays, usually presented by filling each area with a colour.
+However these displays are limited to uni-variate displays, usually presented by filling each area with a colour.
 These maps are best used when the geography is of importance to the data in the display.
-They favor geographically large regions and small regions are often represented by inset displays that zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choroleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
+They favor geographically large regions and small regions are often represented by inset displays that zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
 Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbours.
@@
 Small extreme values are often lost due to the resulting small sized areas the do not draw user's attention.
-The change in the map can also provide opportunities for bivariate displays, using the change in area for one variable, and colour for another.
+The change in the map can also provide opportunities for bi-variate displays, using the change in area for one variable, and colour for another.
 The difference between the familiar map display helps highlight the impact of the disease on communities.
@@
 Tilegram displays may only allow one variable to be visualised if each region is given consistent map space, this occurs in geofaceting. 
-Dorlings provide bivariate displays as they use the size of the shape for one variable, and colour for another.
+Dorlings provide bi-variate displays as they use the size of the shape for one variable, and colour for another.
 These alternatives are especially helpful for data aggregations where administrative boundaries break populations into groups.
@@
 | Uniform use of shape | N | N | N | Y | Y |
-| Bivariate display^[A bivariate display refers to a map that allows two variables to be plotted.
+| Bi-variate display^[A bi-variate display refers to a map that allows two variables to be plotted.
 One variable via the size of the area, and another by filling the area with a colour.] | N | Y | Y | Y | Y |
 
-Creating maps of dieases now involves more decisions to be made by map makers, rather than cartographers. 
-Techonology has played an enormous role in increasing the opportunities for map makers.
+Creating maps of diseases now involves more decisions to be made by map makers, rather than cartographers. 
+Technology has played an enormous role in increasing the opportunities for map makers.
 The computation and graphics power have made creation, alteration and interactivity possible.
@@
 <!--
-Recent developments in technology have created an opportunity to migrate the presentation of epidological results from atlases in paper map collections to web atlases.
+Recent developments in technology have created an opportunity to migrate the presentation of epidemiological results from atlases in paper map collections to web atlases.
 Pairing an explanation with a map or diagram can help to draw attention to a point, group or area of interest.
@@
 -->
+
+XXX Define interactivity, and also animation
 
@@
 Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore. 
-This oppportunity provides an alternative to the traditional approach of a single map for a single use, into a more explorative territory.
+This opportunity provides an alternative to the traditional approach of a single map for a single use, into a more exploratory territory.
 Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
+Direct manipulation of web displays allow continuous view of the map, which change due to the physical actions of users. The affects of these actions are rapid, incremental, and reversible @DMIV.
 
-Developmets and adoptions of this technology has allowed map users more control of the message displayed.
-Figure \ref{fig:supp-grid-create} presents the control menus for several online cancer atlases.
+Developments and adoptions of this technology has allowed map users more control of the message displayed. Figure \ref{fig:interacting} presents the control menus for several online cancer atlases.
 Users can select combinations of cancer type, gender, and other factors to answers questions relevant to them and their work.
 Users can use these controls to change the layer of geography shown, such as continents or countries, states or counties.
-Thess display can show familiar shapes for users to orient themselves on the map, then zoom in to explore densely populated communities.
+These display can show familiar shapes for users to orient themselves on the map, then zoom in to explore densely populated communities.
 
 <!-- Animations -->
-Animations provide a linear, fixed narrative to passive users, communicating a message by capturing and directing attention [@TGA].
+Animations communicate a story, showcasing a critical region by defining and progressing through a small set of keyframes, or waypoints @ESRI.
+The view the transitions between frames in a linear, fixed narrative to passive users, communicating a message by capturing and directing their attention [@TGA]. 
 @TGA suggests the use of moving objects demands attention by taking advantage of the behaviour of the visual cognition system.
 They can be produced to not tolerate or require user input, creating useful, directed descriptions of processes, such as highlighting a particular region or time period [@HTLWM].
-Animations ensure all users follow the same interpretive path but creators cannnot control the level of engagement.
+Animations ensure all users follow the same interpretive path but creators cannot control the level of engagement.
 
 Animations can be implemented to allow users varying degrees of control. 
-A basic form of interactivty with animations is the play, pause and rewind of films, where the film is intended to be watched in one direction but the position in the animation can be controlled.
+A basic form of interactivity with animations is the play, pause and rewind of films, where the film is intended to be watched in one direction but the position in the animation can be controlled.
 @CPISACA provide weather maps as well developed and appropriate examples of spatial displays to communicate to the general public, and encourage those creating disease maps to leverage the success of weather displays.
@@
 
+- very general infovis reference that refers to Buja, Cook, Swayne distinguishing between direct manipulation and interactive graphics https://hal.inria.fr/tel-01096366/file/VA2_PERIN_CHARLES_17112014.pdf
+- explanation of inetractivity and direct manipulation: https://books.google.com.au/books?id=A6TMqxkW-EoC&pg=PA6&lpg=PA6&dq=direct+manipulation+%22swayne%22+%22klinke%22&source=bl&ots=tXYKP8JbbV&sig=ACfU3U3KhN-_YNt_UV5SmThsl4D2gY9A4g&hl=en&sa=X&ved=2ahUKEwj6u9vgqtzkAhVKeysKHQ6iByUQ6AEwA3oECAgQAQ#v=onepage&q=direct%20manipulation%20%22swayne%22%20%22klinke%22&f=false
+- might be useful to look at https://towardsdatascience.com/a-complete-guide-to-an-interactive-geographical-map-using-python-f4c5197e23e0
+- a classic http://www.cs.umd.edu/hcil/trs/2001-08/2001-08.html
+- this is one reference that maybe yoou have missed: https://alsino.io/static/papers/BeyondChoropleths_AlsinoSkowronnek.pdf
+
 # Acknowledgements
@@
 We thank Dr Susanna Cramb (Spatial Modeller, Cancer Council Queensland) and Dr Peter Baade (Senior Research Fellow, Cancer Council Queensland) for providing the opportunity to cover this exploration of disease mapping methods in writing.
-This literature review would not be possible without the opportunity provided by Queesnland University of Technology, and Cancer Council Queensland, and the roles of Professor Kerrie Mengersen (Professor of Statistics, Science and Engineering Faculty, QUT) and Dr Earl Duncan in contributing to the Australian Cancer Atlas and supervision of Stephanie's Research Masters.
+This literature review would not be possible without the opportunity provided by Queensland University of Technology, and Cancer Council Queensland, and the roles of Professor Kerrie Mengersen (Professor of Statistics, Science and Engineering Faculty, QUT) and Dr Earl Duncan in contributing to the Australian Cancer Atlas and supervision of Stephanie's Research Masters.
 
@@
 
-The following packages were used to create tranformations from the sf [@sf] geographical shapes of the states of America [@spData] and the XXX Australian Statistical Areas (Level 3):
-The cartogram [@cartogram] package creates contiguous, non-contiguous and Dorling cartgorams.
+The following packages were used to create transformations from the sf [@sf] geographical shapes of the states of America [@spData] and the XXX Australian Statistical Areas (Level 3):
+The cartogram [@cartogram] package creates contiguous, non-contiguous and Dorling cartograms.
 Hexagon tilegram displays were created using the sugarbag package [@sugarbag].

### 2019-09-22 sa3_plots.R

@@ -10,6 +10,20 @@ library(maptools)
 library(sf)
 library(cartogram)
 
+
+# Filter for centroids only within longitudes
+checkgeom <- function(x){
+  xval <- st_coordinates(x)[[2]]
+  
+  if (xval < 109){return(FALSE)}
+  if (xval > 155){return(FALSE)}
+  return(TRUE)
+}
+
+sa3lung <- absmapsdata::sa32016 %>% 
+  filter(!st_is_empty(geometry)) %>%
+  filter(checkgeom(x = geometry))
+
 # Join with cancer data from AIHW
 #https://www.aihw.gov.au/getmedia/df6732ee-5246-4c9f-a458-df85c88fc512/aihw-can-108-cancer-SA3-incidence-mortality.xls.aspx
 
@@ -28,30 +42,31 @@ cimar_sa3 <- read_excel("data/cimar-sa3.xls",
   filter(!(Name %in% c("Name", "Australia"))) %>% 
   mutate(Population = as.numeric(Population))
 
-
 # Filter for desired areas
 cimar_sa3 <- cimar_sa3 %>% 
-  filter(!(Name %in% 
-    c("Christmas Island", "Cocos (Keeling) Islands", "Lord Howe Island", "Jervis Bay")))
+  filter((Name %in% sa3lung$sa3_name_2016))
 
 # Join with sa3 sf object
-sa3lung <- absmapsdata::sa32016 %>% 
-  left_join(cimar_sa3, by = c("sa3_name_2016"= "Name")) %>% 
+sa3lung <- sa3lung %>% 
+  left_join(cimar_sa3, by = c("sa3_name_2016" = "Name")) %>% 
   st_as_sf() %>%
-  filter(!is.na(Population)) %>%
-  #mutate(Population = ifelse(is.na(Population), 1, Population)) %>% 
+  #filter(!is.na(Population)) %>%
+  mutate(Population = as.numeric(ifelse(is.na(Population), 1, Population)),
+         `Age-standardised rate (per 100,000)` = as.numeric(ifelse(is.na(`Age-standardised rate (per 100,000)`),
+                                                        1, `Age-standardised rate (per 100,000)`))) %>% 
   filter(!st_is_empty(geometry))
 
 sa3lung <- st_transform(sa3lung, 3112)
-#%>%  st_simplify(., preserveTopology = TRUE,  dTolerance = 0) 
 
 b <- st_bbox(sa3lung)
+b["xmin"]<- -2181807
[…]

### 2019-09-22 usa_plots.R

@@ -54,8 +54,6 @@ ggchoro <- ggplot(cancer) +
 ggchoro
 #ggsave(filename = "figures/ggchoro.png", device = "png", dpi = 300, width = 7, height = 6)
 
-usa_legend <- get_legend(ggchoro)
-save(usa_legend, file = "usa_legend.rda")
 
 ###############################################################################
 # Cartograms
@@ -73,6 +71,10 @@ ggcont <- ggplot(cont) +
 save(cont, file = "data/cont.rda")
 ggcont
 ggsave(filename = "figures/ggcont.png", device = "png", dpi = 300, width = 7, height = 6)
+
+
+usa_legend <- get_legend(ggchoro)
+save(usa_legend, file = "usa_legend.rda")
 
 ###############################################################################
 # Non - Contiguous Cartograms
@@ -223,41 +225,7 @@ ggsave(filename = "figures/gghexmap.png", plot = gghexmap,
   device = "png", dpi = 300, width = 7, height = 6)
 
 
+
 usa_grid <- gridExtra::grid.arrange(ggcont, ggncont, ggdorl, gghexmap, nrow = 2)
 ggsave(filename = "figures/usa_grid.png", plot = usa_grid,
   device = "png", dpi = 300, width = 7, height = 6)
-
-
-#####
-# Choropleth grid
-
-
-plot1 <- png::readPNG('cancer_map_images/chloropleths/1.png')
-plot2 <- png::readPNG('cancer_map_images/chloropleths/14.chloropleth1.png')
-plot3 <- png::readPNG('cancer_map_images/chloropleths/15.chloropleth.png')
-plot4 <- png::readPNG('cancer_map_images/chloropleths/3.chloropleth.png')
-plot5 <- png::readPNG('cancer_map_images/chloropleths/5.chloropleth.png')
-plot6 <- png::readPNG('cancer_map_images/chloropleths/6.chloropleth.png')
-plot6 <- png::readPNG('cancer_map_images/chloropleths/7.chloropleth.png')
-plot7 <- png::readPNG('cancer_map_images/chloropleths/8.chloropleth.png')
-plot8 <- png::readPNG('cancer_map_images/chloropleths/20.chloropleth1.png')
-plot9 <- png::readPNG('cancer_map_images/chloropleths/21.chloropleth2.png')
[…]

### 2019-09-23 review_article.Rmd
Cartogram with single legend

--- 2019-09-21 review_article.Rmd
@@
 
-load("usa_legend.rda")
+load("figures/usa_legend.rda")
 
@@
 
+load("figures/aus_legend.rda")
+
 library(cowplot)
 ggdraw() +
-  draw_plot(rasterGrob(aus1), 0, .5, 0.5, 0.5) +
-  draw_plot(rasterGrob(aus2), 0.5, 0.5, 0.5, 0.5) +
-  draw_plot(rasterGrob(aus3), 0.0, 0.0, 0.5, 0.5) +
-  draw_plot(rasterGrob(aus4), 0.5, 0, 0.5, 0.5)+
+  draw_plot(rasterGrob(aus1), 0, .5, 0.5, 0.45) +
+  draw_plot(rasterGrob(aus2), 0.5, 0.5, 0.5, 0.45) +
+  draw_plot(rasterGrob(aus3), 0.0, 0.0, 0.5, 0.45) +
+  draw_plot(rasterGrob(aus4), 0.5, 0, 0.5, 0.45) +
+  draw_plot(aus_legend, 0, 0, 1, 0.1) +
   draw_plot_label(c("a", "b", "c", "d"), 

### 2019-09-23 sa3_plots.R

@@ -64,6 +64,9 @@ aus_ggchoro
 ggsave(filename = "figures/aus_ggchoro.png", device = "png", dpi = 300,  width = 7, height = 6)
 
 
+aus_legend <- get_legend(aus_ggchoro)
+save(aus_legend, file = "figures/aus_legend.rda")
+
 ###############################################################################
 # Cartograms
   
@@ -82,8 +85,6 @@ aus_ggcont <- ggplot(cont) +
 aus_ggcont
 ggsave(filename = "figures/aus_ggcont.png", device = "png", dpi = 300,  width = 7, height = 6)
 
-aus_legend <- get_legend(aus_ggchoro)
-save(aus_legend, file = "aus_legend.rda")
 
 ###############################################################################
 # Non - Contiguous Cartograms

### 2019-09-23 usa_plots.R

@@ -13,7 +13,6 @@
 # 2012-2016
 
 # Export
-
 library(sugarbag)
 library(tidyverse)
 library(spData)
@@ -21,6 +20,7 @@ library(ggthemes)
 library(maptools)
 library(sf)
 library(cartogram)
+library(cowplot)
 library(grid)
 library(png)
 library(gridExtra)
@@ -55,7 +55,7 @@ ggchoro
 #ggsave(filename = "figures/ggchoro.png", device = "png", dpi = 300, width = 7, height = 6)
 
 usa_legend <- get_legend(ggchoro)
-save(usa_legend, file = "usa_legend.rda")
+save(usa_legend, file = "figures/usa_legend.rda")
 
 ###############################################################################
 # Cartograms

### 2019-09-23 review_article.Rmd
tilegrams

--- 2019-09-22 review_article.Rmd
@@
 load("aus_legend.rda")
+
+load("figures/aus_legend.rda")
 
@@
 <!-- cite mosaic cartograms -->
-<!--XXX cite geogrid here?-->
+
 A simple tilegram may present the states in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
@@
 <!-- Choropleths, true topology, geographic distance, good for showing geographic importance -->
-Visualisation properties have been explored across several disease mapping displays.
 To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
@@
 Choropleth methods require a decision to be made about the projection of the display, in this case aim to find a map projection that gives the least distorted representation of the geography [@UMPT].
-When there is a relationship between the cancer type being mapped and geographical features, a choropleth map would pair this information visually.
-However these displays are limited to uni-variate displays, usually presented by filling each area with a colour.
-These maps are best used when the geography is of importance to the data in the display.
-They favor geographically large regions and small regions are often represented by inset displays that zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
-Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbours.
+When there is a relationship between the variable being mapped and geographical features, a choropleth map would pair this information visually.
+However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a colour.
+These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
+Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbours. The map is reusable and fit for many purposes.
 
@@
 Cartographers were historically the creators of geographic displays, and epidemiologists utilised the maps that had been prepared in advance.
-Cartograms with the population variable used to distortion the geographic map can be used as density equalising maps, that reduce the visual prominence of low-population areas [@VSSDCUC]. 
+Cartograms with the population variable used to distortion the geographic map regions and boundaries are density equalising maps, that reduce the visual prominence of large, but low-population areas [@VSSDCUC].
 Cartogram displays incorporate the statistics and population into the design of the display,
-shifting and sometimes sacrificing familiar boundaries to draw attention to outliers in the data space or population density.
-The unusual shapes are difficult to compare and can prevent comparison of the statistic.
+shifting and sometimes sacrificing familiar boundaries to draw attention to large outliers in the data space or population density.
+The unusual shapes are not longer familiar and may be difficult to compare.
+This can prevent effective comparison of the statistic.
 Cartograms also favour extreme statistics in the positive direction.
 Small extreme values are often lost due to the resulting small sized areas the do not draw user's attention.
-The change in the map can also provide opportunities for bi-variate displays, using the change in area for one variable, and colour for another.
+The change in the map based on one variable provides the opportunity for bi-variate displays, using the change in area for one variable, and colour for another.
 The difference between the familiar map display helps highlight the impact of the disease on communities.

### 2019-09-23 review_article.Rmd
changes to table 1, added tilegram and geofacet

@@
 |:--------------------|:------------------------------------------------|  
-|1. IR (Incidence Ratio)|$(IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}$, |
+|1. Count|Crude cancer counts|
+|2. Rate per 100,000|Cancer incidence per 100,000 population|  
+|3. New cancer cases per 100,000|Specific methods could not be found|   
+|4. IR (Incidence Ratio)|$(IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}$, |
 ||Cancer incidence rate in region $i$ over the average cancer incidence rate for the total region|
-|2. SIR (Standardised Incidence Ratio)| IR standardised by age structure in each region $i$|
-|3. RER |$RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}$|
+|5. Age Adjusted Rate per 100,000| #5 standardised by age structure or region|   
+|6. Age Adjusted Relative Risk| RR standardised by age structure in each region $i$|  
+|7. SIR (Standardised Incidence Ratio)| IR standardised by age structure in each region $i$|
+|8. Below or above Expected|Alternative expression of the SIR|
+|9. RER |$RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}$|
 |(Relative Excess Risk)|Represents the estimate of cancer related mortality within five years of diagnosis|
 ||Also referred to as 'excess hazard ratio'|
-|4. Age Adjusted Relative Risk| RR standardised by age structure in each region $i$|   
-|5. Rate per 100,000|Cancer incidence per 100,000 population|  
-|6. Age Adjusted Rate per 100,000| #5 standardised by age structure or region|   
-|7. New cancer cases per 100,000|Specific methods could not be found|   
-|8. Count|Crude cancer counts|  
-|9. Below or above Expected|Alternative expression of the SIR|
 
@@
 ggsave(filename = "figures/supp_grid.png", device = "png", 
-  width = 6, height = 6, dpi = 300)
+  width = 12, height = 9, dpi = 300)
 ```
@@
 
-# Cartograms and other spatially warped displays {#ch:alternatives}
+# Alternative to choropleths {#ch:alternatives}
 
@@
 
-Choropleths are univariate displays, they do not inherently support multi-dimensionality. This makes it difficult to pair demographic, environmental and other factors with spatial distributions.
-@MMST present linked micromap plots as a solution to the issue of multi-dimensionality when visually linking geographic and statistical data.
-This design is widely used in online atlas displays, it groups areas of similar values for one variable, and uses the additional columns to provide displays that contrast the states in each group by other variables.
-The display juxtaposes of choropleth maps and statistical plots.
-Showing one map per group of the key separating variable, in a row with each additional statistical plot.
+Choropleths are univariate displays, they do not inherently support multi-dimensionality
+This makes it difficult to pair demographic, environmental and other factors with spatial distributions.
+@MMST present linked micromap plots as a solution to the issue of multi-dimensionality to visually link geographic and statistical data.
+This design is widely used in online atlas displays, it groups areas based on their value for one variable, and uses  additional columns to provide displays that contrast the areas in each group by other variables.
+The display juxtaposes of choropleth maps and statistical plots;
+it shows one map per group of the key separating variable, in a row with each additional statistical plot.
 
@@
 
-A simple tilegram may present the states in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
+A simple tilegram presents the areas in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
 These tiles may be labelled or coloured to represent a value.
-A data visualisation can be geographically faceted, so that a smaller visualisation of equal size is created for each geographic subset of the data.
-This requires grid displays that are often created once by hand and reused.
 
@@
 
-Bob Rudis, geogrid
+Tilegrams can be extended to allow more than just simple colouring of tiles to present data.
+Geofaceting utilises a grid display, the arrangement of tiles creates a grid that mimics the geographic topology of the set of areas @IGF. 
+Like tilegrams the arrangement of the areas can be reused for other spatial distribution visulisations.
+Unlike tilegrams geofaceting replicates visualisations for each geographic subset of the data.
+The amount of information able to be communicated has increased from one value per region to one visualisation, this is a more flexible display, especially as it provides a simple way to communicate time series data.
 

### 2019-09-23 usa_plots.R

@@ -167,22 +167,6 @@ ggsave(filename = "figures/ggchoroCRS.png", plot = ggchoroCRS,
 
 
 ###############################################################################
-#Tilegram
-library(statebins)
-
-## US example
-ggtilegram <- left_join(adat, cancer %>% rename(state = NAME)) %>%
-  statebins(
-    value_col = "AgeAdjustedRate", 
-    name = "AgeAdjustedRate"
-  ) +
-  labs(title = "") + 
-  theme_void()+ guides(fill = FALSE) +
-  scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1)
-
-ggsave(filename = "figures/ggtilegram.png", plot = ggtilegram,
-  device = "png", dpi = 300, width = 7, height = 6)
-
 
 cancermap <- st_transform(cancer, "+proj=longlat +datum=WGS84 +no_defs")
 centroids <- create_centroids(cancermap, "NAME")
@@ -228,3 +212,43 @@ ggsave(filename = "figures/gghexmap.png", plot = gghexmap,
 usa_grid <- gridExtra::grid.arrange(ggcont, ggncont, ggdorl, gghexmap, nrow = 2)
 ggsave(filename = "figures/usa_grid.png", plot = usa_grid,
   device = "png", dpi = 300, width = 7, height = 6)
+
+###############################################################################
+
+
+# Tilegram
+library(statebins)
+
+## US example
+ggtilegram <- statebins(cancer %>% rename(state = NAME),
+    value_col = "AgeAdjustedRate", 
+    name = "AgeAdjustedRate",type = "seq", palette = "RdPu",  direction = 1
+  ) +
+  labs(title = "") + 
+  theme_void() + guides(fill = FALSE)
+ggtilegram
+ggsave(filename = "figures/ggtilegram.png", plot = ggtilegram,
+  device = "png", dpi = 300, width = 7, height = 6)
+
+
+# Geofacet
+library(geofacet)
+# read in data
+cancer_f <- read_csv("data/USCS_lung_f.csv") %>% mutate(Sex = "F")
+cancer_m <- read_csv("data/USCS_lung_m.csv") %>% mutate(Sex = "M")
+
+ggfacet <- bind_rows(cancer_f, cancer_m) %>% 
+    mutate(state = gsub("'", "", Area),
+    AgeAdjustedRate = parse_number(AgeAdjustedRate)) %>% 
[…]

### 2019-09-23 review_article.Rmd
added geofacet section

@@
 it shows one map per group of the key separating variable, in a row with each additional statistical plot.
+Linked micromaps predominantly utilise the choropleth map for displays of spatial distributions.
+It is one of several alternative displays that allow maps to become bi-variate displays.
 
@@
 Two new templates for epidemiology applications: linkedmicromap plots and conditioned choropleth maps-->
-
-Linked micromaps predominantly utilise the choropleth map for displays of spatial distributions.
-It is one of several alternative displays that allow maps to become bi-variate displays.
 
@@
 
-## Tilegram
+## Tile map
 
-A tilegram provides a tessellated display of consistent shapes.
+A Tile map provides a tessellated display of consistent shapes.
 Each tile is usually one unit of measurement, this could be geographic regions such as states, or population based where one tile is used for a consistent measure of population.
@@
 
-A simple tilegram presents the areas in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
+A simple Tile map presents the areas in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
 These tiles may be labelled or coloured to represent a value.
+Tile maps may be difficult to create, they are best created manually, with additional time and care required as the amount of geographic areas to include increases.
 
-<!--
-```{r ggtilegram, fig.cap = "A tilegram of the Unites States of America.
+```{r ggtilemap, fig.cap = "A Tile map and geofaceted map of the Unites States of America.
 Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
@@
 The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.", message=FALSE, warning=FALSE, out.width='80%'}
-knitr::include_graphics("figures/ggtilegram.png")
+The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.", message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
+
+# Import the images
+t1 <- png::readPNG('figures/ggtilegram.png')
+t2 <- png::readPNG('figures/ggfacet.png')
+
+ggdraw() +
+  draw_plot(rasterGrob(t1), 0, 0, 1, 1) +
+  draw_plot(rasterGrob(t2), 1, 0, 1, 1) +
+  draw_plot_label(c("a", "b"), 
+    c(0.01, 0.51), 
+    c(0.48, 0.48), size = 20)
+
+ggsave(filename = "figures/choropleth_grid.png", 
+  device = "png", width = 12, height = 6, dpi = 300)
+
 ```
--->
 
+## Geofaceting
 Tilegrams can be extended to allow more than just simple colouring of tiles to present data.
-Geofaceting utilises a grid display, the arrangement of tiles creates a grid that mimics the geographic topology of the set of areas @IGF. 
-Like tilegrams the arrangement of the areas can be reused for other spatial distribution visulisations.
-Unlike tilegrams geofaceting replicates visualisations for each geographic subset of the data.
-The amount of information able to be communicated has increased from one value per region to one visualisation, this is a more flexible display, especially as it provides a simple way to communicate time series data.
+@IGF formalises the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas. 
+Like tilegrams, the arrangement of the areas can be reused for other spatial distribution visulisations.
+Geofaceting have the functionality of facets, often used to replicate visualisations for each subset of the data, in this case the data subsets are geographic.
+The amount of information able to be communicated has increased from one value per region in a tilegram to one visualisation, this is a more flexible display.
+Virtually any plot display can be fit into the tile representing the areas, allowing displays of multiple variables or values per geographic entity.
+As the amount of areas increase care must be taken to ensure legibility of the displays.
 
@@
 The progress from static print media displays to interactive methods has allowed users to develop maps to ask and answer their own questions.
-Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore. 
+Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions that allow inspection of the data from many views [@DQBCM].
+Interaction covers manipulation of the graph directly, via the graph controls or command-line [@RPIG].
+
 This opportunity provides an alternative to the traditional approach of a single map for a single use, into a more exploratory territory.
 Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
+Tooltips allow users to obtain additional information regarding specific geographic areas.
 Direct manipulation of web displays allow continuous view of the map, which change due to the physical actions of users. The affects of these actions are rapid, incremental, and reversible @DMIV.
@@
 Developments and adoptions of this technology has allowed map users more control of the message displayed. Figure \ref{fig:interacting} presents the control menus for several online cancer atlases.
-Users can select combinations of cancer type, gender, and other factors to answers questions relevant to them and their work.
-Users can use these controls to change the layer of geography shown, such as continents or countries, states or counties.
+The user interface for these online atlases allow combinations of cancer type, gender, and other factors to be shown on the map display.
+These controls may also change the geographies shown, such as continents or countries, states or counties.
 These display can show familiar shapes for users to orient themselves on the map, then zoom in to explore densely populated communities.
+Interactions with online atlases can also change the amount of detail shown regarding the display.
+It is uncommon to see linked displays utilised in online atlases, but there are several frameworks that support interactive displays of multivariate data.
+Linked displays allow for direct manipulation, @RPIG uses Swayne and Klinke's definition of high level interaction that occurs via data selection or brushing tools that highlight subsets of graphical elements such as polygons or points.
+Queries form a low level interaction, returning values or information used in the graphical display without referring to the original data source.
 
@@
 More complex animations may include steps, when user interaction via clicks animates the display between stages.
-This approach is implemented by Google Earth 
 Where the message is most important, static or animated graphics allow control over the display and interpretation.
@@
 - a classic http://www.cs.umd.edu/hcil/trs/2001-08/2001-08.html
-- this is one reference that maybe yoou have missed: https://alsino.io/static/papers/BeyondChoropleths_AlsinoSkowronnek.pdf
+- this is one reference that maybe you have missed: https://alsino.io/static/papers/BeyondChoropleths_AlsinoSkowronnek.pdf
 

### 2019-09-24 review_article.Rmd
put tilegram and ggfacet together

@@
 Current best practices: Choropleths
-Current exploratory practices: Cartograms: cont, non cont, (tilegrams, dorlings)
+Current exploratory practices: Cartograms: cont, non cont, (tilegrams, Dorlings)
 Suggested uses of 
@@
 <!-- what is a choropleth map -->
-A choropleth map is used to display the spatial relationship of measurements by shading areas of a map.
+A choropleth map is used to display differences in the geographical distribution of data by spatial unit by shading areas of a map.
 The geography is faithfully rendered, and the colour rendering is designed to reveal spatial patterns among data values.
@@
 While the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
-The different population and geographical sizes of administrative areas can attract attention to the shades of the unpopulated but large areas [@EI]. 
+The different population and geographical sizes of administrative areas can attract attention to the shades of the unpopulated but large areas [@EI], Skowronnek [-@BCM] calls this an area-size bias. 
 Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention away from the 'potentially more important results in the more populous communities' that are geographically smaller [@SE].
@@
 
-(insert figure of both types of interactivity)
-
 Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
@@
 @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
-
-<!--
-
-Technological advances may provide a solution to this through the use of interactivity, animation, and sound as additional channels for representing uncertainty information without interfering with the ability to see any features that are present in the display [@maceachren1992visualizing].
-However, if uncertainty information is perceived as integral and inseparable from the data it is associated with, assumptions of uncertainty creating cognitive overload may not be as valid.
-
--->
 
@@
 <!-- What is a cartogram? -->
-A cartogram alters the map base with the intention of improving the presentation of the statistic of interest.
+Choropleths imply uniformity of data across the geographic space, when considering population densities it is extremely unlikely to be uniform [@BCM].
+A cartogram displays abstracted geographic space, with the intention of improving the presentation of the statistic of interest.
 For a single variable of interest, each map area is changed to emphasise the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].
@@
 
-Table: \label{tab:usa} Maps used to present statistics for the United States of America
+Table: \label{tab:usa} Maps used to present statistics for the United States of America.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
 
@@
 |a.
-contiguous cartogram|A contiguous cartogram map of the United States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each stated shape has been distorted according to the population of the state in 2015.
+contiguous cartogram| Each stated shape has been distorted according to the population of the state in 2015.
 The state of California has become much larger due to it's large population density.
@@
 |b.
-Non - Contiguous|A Non - contiguous cartogram map of the United States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each state shape has been maintained, but the size has altered according to the population of the state in 2015.
+Non - Contiguous| The geographic shape of the states has been maintained, but the size has altered according to the population of the state in 2015.
 The state of California has remained closer to it's original size than it's surrounding states.
@@
 This draws attention to the densely populated North East region, and the sparse Mid West.|
-|Dorling|A dorling cartogram of the Unites States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each state has been represented by a circle, but the size was determined by the population of the state in 2015.
+|Dorling| The states have been represented by a circle, but the size was determined by the population of the state in 2015.
 The North East states remain closer to their neighbours, and may be displaced from their geographic location.
 The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.|
-|Hexagon Tessellation|A hexagon tessellation map of the Unites States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each state has been represented by a square, each square is the same size.
+|Hexagon Tessellation| Each state is represented by a hexagon of equal size.
 The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.
@@
 
-A Tile map provides a tessellated display of consistent shapes.
+A tile map provides a tessellated display of consistent shapes.
 Each tile is usually one unit of measurement, this could be geographic regions such as states, or population based where one tile is used for a consistent measure of population.
@@
 
-A simple Tile map presents the areas in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
+A simple tile map presents the areas in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
 These tiles may be labelled or coloured to represent a value.
@@
 
-```{r ggtilemap, fig.cap = "A Tile map and geofaceted map of the Unites States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+```{r ggtilemap, fig.cap = "A tile map and geofaceted map of the Unites States of America.
+Each state in the tilegram has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
 Each state has been represented by a square, each square is the same size.
-The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.", message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
-
-# Import the images
-t1 <- png::readPNG('figures/ggtilegram.png')
-t2 <- png::readPNG('figures/ggfacet.png')
-
-ggdraw() +
-  draw_plot(rasterGrob(t1), 0, 0, 1, 1) +
-  draw_plot(rasterGrob(t2), 1, 0, 1, 1) +
-  draw_plot_label(c("a", "b"), 
-    c(0.01, 0.51), 
-    c(0.48, 0.48), size = 20)
-
-ggsave(filename = "figures/choropleth_grid.png", 
-  device = "png", width = 12, height = 6, dpi = 300)
-
+The geofaceted map allows the rates for males and females to be contrasted.
+It has the same layout as the tilegram, but shows more information per state than ", message=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width = "100%"}
+knitr::include_graphics("figures/gggrids.png")
 ```
@@
 As the amount of areas increase care must be taken to ensure legibility of the displays.
+@RPIG acknowledges that this display suffers when the areas have very irregular shapes and large size disparities, as manual creation of the grid is still required.
 
@@
 Choropleth displays utilise more traditional cartographic methods, they are usually true to the topography, displaying familiar boundaries of countries, states or administrative areas.
+Highlighting the geographic distribution, and differences between equal spatial units.
+When there is a relationship between the variable being mapped and geographical features, a choropleth map would pair this information visually.
 Choropleth methods require a decision to be made about the projection of the display, in this case aim to find a map projection that gives the least distorted representation of the geography [@UMPT].
-When there is a relationship between the variable being mapped and geographical features, a choropleth map would pair this information visually.
 However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a colour.
@@
 # Interactivity and Animation  {#ch:interacting}
-<!--
-Recent developments in technology have created an opportunity to migrate the presentation of epidemiological results from atlases in paper map collections to web atlases.
-Pairing an explanation with a map or diagram can help to draw attention to a point, group or area of interest.
-Maps may form supplementary material to the discussion, in this case static visualisations can help to aid written results. 
-Static maps often need to include enough information for the display to be understandable out of the original context.
--->
-
-XXX Define interactivity, and also animation
 
@@
 <!-- Interactivity (Compare movies or tv to video games) -->
-The progress from static print media displays to interactive methods has allowed users to develop maps to ask and answer their own questions.
+The progress from static print media displays to interactive methods allows users to develop maps to ask and answer their own questions.
+Interactive methods involve direct manipulation of web displays, keeping a continuous view of the map, which changes due to the physical actions of users. 
+The manipulation of the graph may occur directly, via the graph controls or the command-line [@RPIG].
+The affects of these actions are rapid, incremental, and reversible @DMIV.
 Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions that allow inspection of the data from many views [@DQBCM].
-Interaction covers manipulation of the graph directly, via the graph controls or command-line [@RPIG].
-
 This opportunity provides an alternative to the traditional approach of a single map for a single use, into a more exploratory territory.
@@
 Tooltips allow users to obtain additional information regarding specific geographic areas.
-Direct manipulation of web displays allow continuous view of the map, which change due to the physical actions of users. The affects of these actions are rapid, incremental, and reversible @DMIV.
 
-Developments and adoptions of this technology has allowed map users more control of the message displayed. Figure \ref{fig:interacting} presents the control menus for several online cancer atlases.
+Developments and adoptions of this technology has allowed map users more control of the message displayed.
+Figure \ref{fig:interacting} presents forms of low level interactions with a user interface, these are the control menus for several online cancer atlases.
 The user interface for these online atlases allow combinations of cancer type, gender, and other factors to be shown on the map display.
 These controls may also change the geographies shown, such as continents or countries, states or counties.
-These display can show familiar shapes for users to orient themselves on the map, then zoom in to explore densely populated communities.
-Interactions with online atlases can also change the amount of detail shown regarding the display.
+These displays can show familiar shapes for users to orient themselves on the map, then zoom in to explore densely populated communities.
+Interactions with online atlases can also change the amount of detail shown.
 It is uncommon to see linked displays utilised in online atlases, but there are several frameworks that support interactive displays of multivariate data.
 Linked displays allow for direct manipulation, @RPIG uses Swayne and Klinke's definition of high level interaction that occurs via data selection or brushing tools that highlight subsets of graphical elements such as polygons or points.
-Queries form a low level interaction, returning values or information used in the graphical display without referring to the original data source.
+Low level interaction is more common, this occurs via the user interface or the map.
+Interacting with the map returns values or information used in the graphical display without referring to, of changing, the original data source of the map.
 
@@
 A basic form of interactivity with animations is the play, pause and rewind of films, where the film is intended to be watched in one direction but the position in the animation can be controlled.
-@CPISACA provide weather maps as well developed and appropriate examples of spatial displays to communicate to the general public, and encourage those creating disease maps to leverage the success of weather displays.
-The movement of a weather system will follow a forecasted path, the animation will communicate this so that all map users can follow the path of the weather system across the geography over a specified period of time.
-More complex animations may include steps, when user interaction via clicks animates the display between stages.
+@CPISACA provide weather maps as a thoroughly developed example of animation of spatial displays to communicate information to the general public.
+The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period of time.
 Where the message is most important, static or animated graphics allow control over the display and interpretation.
+User interaction more allow clicks to progress between stages of more complex animations.
 The Australian Cancer Atlas [-@TACA] provides [tours](
 https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
-This directs user's attention and gives them tools to plan their own exploration.
-
-<!--
-A complex directed animation approach is implemented in steps by Google Earth, in their use of stories.
-Within these directed steps users have the opportunity to use interactive controls to change the map directions or watch videos.
-One popular implementation communicates Dr.
-Jane Goodall's life journey using Google Earth displays of significant locations such as her birthplace Bournemouth, England and Tanzania's Gombe National Park.
--->
-
-- very general infovis reference that refers to Buja, Cook, Swayne distinguishing between direct manipulation and interactive graphics https://hal.inria.fr/tel-01096366/file/VA2_PERIN_CHARLES_17112014.pdf
-- explanation of inetractivity and direct manipulation: https://books.google.com.au/books?id=A6TMqxkW-EoC&pg=PA6&lpg=PA6&dq=direct+manipulation+%22swayne%22+%22klinke%22&source=bl&ots=tXYKP8JbbV&sig=ACfU3U3KhN-_YNt_UV5SmThsl4D2gY9A4g&hl=en&sa=X&ved=2ahUKEwj6u9vgqtzkAhVKeysKHQ6iByUQ6AEwA3oECAgQAQ#v=onepage&q=direct%20manipulation%20%22swayne%22%20%22klinke%22&f=false
-- might be useful to look at https://towardsdatascience.com/a-complete-guide-to-an-interactive-geographical-map-using-python-f4c5197e23e0
-- a classic http://www.cs.umd.edu/hcil/trs/2001-08/2001-08.html
-- this is one reference that maybe you have missed: https://alsino.io/static/papers/BeyondChoropleths_AlsinoSkowronnek.pdf
+This implementation of animation gives users tools to plan their own exploration.
 

### 2019-09-24 usa_plots.R

@@ -225,7 +225,8 @@ ggtilegram <- statebins(cancer %>% rename(state = NAME),
     name = "AgeAdjustedRate",type = "seq", palette = "RdPu",  direction = 1
   ) +
   labs(title = "") + 
-  theme_void() + guides(fill = FALSE)
+  theme_void() +
+  theme(legend.position ="bottom")
 ggtilegram
 ggsave(filename = "figures/ggtilegram.png", plot = ggtilegram,
   device = "png", dpi = 300, width = 7, height = 6)
@@ -233,22 +234,30 @@ ggsave(filename = "figures/ggtilegram.png", plot = ggtilegram,
 
 # Geofacet
 library(geofacet)
+us_grid <- read_csv("us_grid.csv")
+
 # read in data
 cancer_f <- read_csv("data/USCS_lung_f.csv") %>% mutate(Sex = "F")
 cancer_m <- read_csv("data/USCS_lung_m.csv") %>% mutate(Sex = "M")
 
-ggfacet <- bind_rows(cancer_f, cancer_m) %>% 
+cancer_mf <- bind_rows(cancer_f, cancer_m) %>% 
     mutate(state = gsub("'", "", Area),
     AgeAdjustedRate = parse_number(AgeAdjustedRate)) %>% 
-  select(state, Sex, AgeAdjustedRate) %>% 
-  ggplot() + 
+  select(state, Sex, AgeAdjustedRate)
+
+ggfacet <- ggplot(cancer_mf) + 
   geom_col(aes(x = Sex, y = AgeAdjustedRate, fill = Sex)) + 
   scale_fill_brewer(type = "qual", palette = "Pastel1",  direction = 1) + 
-  theme_minimal() + 
-  facet_geo(~ state, grid = "us_state_grid2",label = "code")+
+  theme_void() + 
+  facet_geo(~ state, grid = us_grid, label = "code") +
   theme(legend.position ="bottom")
 
 ggfacet
 ggsave(filename = "figures/ggfacet.png", plot = ggfacet,
   device = "png", dpi = 300, width = 7, height = 6)
 
+
+p <- plot_grid(ggtilegram, ggfacet, labels = c('A', 'B'), label_size = 12, align = v, )
+
+ggsave(filename = "figures/gggrids.png", plot = p,
+  device = "png", dpi = 300, width = 12, height = 6)

### 2019-09-24 review_article.Rmd
produces pdf but gives permission errors

@@
 # Introduction
-<!--
-Current best practices: Choropleths
-Current exploratory practices: Cartograms: cont, non cont, (tilegrams, Dorlings)
-Suggested uses of 
--->
-
 <!-- What data are we working with, why, what are the purposes -->
@@
 Each area is represented by a square of the same dimensions.
-Figure \ref{fig:ggcartograms} d shows a hexagon tilegram representing each state using a single square.
+Figure \ref{fig:ggcartograms} d shows a hexagon tile map representing each state using a single square.
 
@@
 ```{r ggtilemap, fig.cap = "A tile map and geofaceted map of the Unites States of America.
-Each state in the tilegram has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state in the tile map has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
 Each state has been represented by a square, each square is the same size.
 The geofaceted map allows the rates for males and females to be contrasted.
-It has the same layout as the tilegram, but shows more information per state than ", message=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width = "100%"}
+It has the same layout as the tile map, but shows more information per state than ", message=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width = "100%"}
 knitr::include_graphics("figures/gggrids.png")
@@
 ## Geofaceting
-Tilegrams can be extended to allow more than just simple colouring of tiles to present data.
+Tile maps can be extended to allow more than just simple colouring of tiles to present data.
 @IGF formalises the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas. 
-Like tilegrams, the arrangement of the areas can be reused for other spatial distribution visulisations.
+Like tile maps, the arrangement of the areas can be reused for other spatial distribution visulisations.
 Geofaceting have the functionality of facets, often used to replicate visualisations for each subset of the data, in this case the data subsets are geographic.
-The amount of information able to be communicated has increased from one value per region in a tilegram to one visualisation, this is a more flexible display.
+The amount of information able to be communicated has increased from one value per region in a tile map to one visualisation, this is a more flexible display.
 Virtually any plot display can be fit into the tile representing the areas, allowing displays of multiple variables or values per geographic entity.
@@
 Areas that are geographically close will maintain connectedness in some way, but the population and the statistic will dominate.
-Tilegram displays may only allow one variable to be visualised if each region is given consistent map space, this occurs in geofaceting. 
+Tile map displays may only allow one variable to be visualised if each region is given consistent map space, this occurs in geofaceting. 
 Dorlings provide bi-variate displays as they use the size of the shape for one variable, and colour for another.
@@
 
-|  | Choropleth | Contiguous | Non-contiguous | Dorling | Tilegrams |
+|  | Choropleth | Contiguous | Non-contiguous | Dorling | Tile maps |
 |:-----------:|------------|-----------|---------------|---------|-----------|
@@
 
-The following packages were used to create transformations from the sf [@sf] geographical shapes of the states of America [@spData] and the XXX Australian Statistical Areas (Level 3):
+The following packages were used to create transformations from the sf [@sf] geographical shapes of the states of America [@spData] and the Australian Statistical Areas (Level 3):
 The cartogram [@cartogram] package creates contiguous, non-contiguous and Dorling cartograms.
-Hexagon tilegram displays were created using the sugarbag package [@sugarbag].
+Hexagon tile map displays were created using the sugarbag package [@sugarbag].
 

### 2019-09-24 review_article.Rmd
animation figure

@@
 This intentional design goal coincided with the rising interest in urban geography and presentation of social statistics.
+Figure \ref{fig:ggcartograms} a shows a population contiguous cartogram of the United States.
+All states are visible and the shape of the United States overall is still recognisable.
+In contrast, Figure \ref{fig:auscartograms} a shows an Australian contiguous cartogram also based on population.
+The south east is completely distorted, areas with low popuation are still large very the map, as their initial area was used to determine the area allocated in the contiguous cartogram.
 
@@
 @CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes.
-@MDAC define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area can be used to communicate the statistic of regions. 
-The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off be made between boundary complexity and simplicity.
-
-<!--
-Tobler's method and the many implementations that 'elaborated' on it are derived from 'numerical approximations to a pair of equations'[@ACTUC].
-They all operate through incremental adjustments, and can produce wildly different outcomes from small changes in the inputs.
-
-Computer generated map examples:  @HEDP
-(Hopps et al. 1968; Armstrong 1972). There
-has followed a flood of disease atlases, mainly concentrating on the modem problems of cancer and degenerative diseases from countries as scattered as the United States (Burbank 1971; Mason et al.
-1975, 1976; Pickle et al.
-1987), the Soviet Union (Levin 1980), Japan (Shigematsu 1977), the Federal Republic of Germany
--->
 
@@
 Map creators can efficiently communicate with this kind of map by keeping the outlines or particular elements of the original in the new shape [@NISCC].
-The scale of the areas does not impact on the shape recognition. However it may impact on the visibility of all areas if small areas expand beyond their boundaries.
+The scale of the areas does not impact on the shape recognition.
+However it may impact on the visibility of all areas if small areas expand beyond their boundaries. 
+Figure \ref{fig:ggcartograms} b shows small overlaps in the north east region of the United States.
+These overlaps are also shown in the non-contiguous cartogram of Australia in Figure \ref{fig:auscartograms} b.
 
@@
 This implementation can work for up to 'one hundred thousand' areas.
+Figure \ref{fig:ggcartograms} c shows the displacement of the north east region of the United States in the Dorling cartogram, the rest of the country is still recognisable.
+This displacement is shown to an extreme in \ref{fig:auscartograms} c the Dorling cartogram of Australia.
 
@@
 Each area is represented by a square of the same dimensions.
-Figure \ref{fig:ggcartograms} d shows a hexagon tile map representing each state using a single square.
+Figure \ref{fig:ggcartograms} d shows a hexagon tile map representing each state using a single hexagon.
 
@@
 Tile maps may be difficult to create, they are best created manually, with additional time and care required as the amount of geographic areas to include increases.
+@MDAC define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area can be used to communicate the statistic of regions. 
+The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off be made between boundary complexity and simplicity.
 
@@
 
+```{r animating, fig.cap = "Animations in publicly available choropleth cancer maps.
+[a.
+Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC)(Controls for indicator, sex, age groups, continents, and cancer types) , [b.
+Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality) (Controls for indicator, age, year, and geographical areas)", fig.width = 12, fig.height = 9}
+a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
+a2 <- png::readPNG("cancer_map_images/animation/2.animation.png")
+
+ggdraw() +
+  draw_plot(rasterGrob(a1), 0, 0, 0.3, 1) +
+  draw_plot(rasterGrob(a2), 0.3, 0, 0.7, 1) +
+  draw_plot_label(c("I", "II"), 
+    c(0, .3), c(1, 1), size = 20)
+```
+
 # Acknowledgements

### 2019-09-24 review_article.Rmd
change spelling to US format

@@
 ---
-title: "Cartogram Mapping and its Application to Cancer Data Visualisation"
+title: "Cartogram Mapping and its Application to Cancer Data Visualization"
 author: "Stephanie Kobakian, Jessie Roberts and Dianne Cook"
@@
 output:
+  word_document:
+    toc: yes
+    toc_depth: '3'
   pdf_document:
+    fig_caption: yes
     number_sections: yes
@@
     toc_depth: 3
-    fig_caption: true
+link-citations: yes
 fontsize: 11pt
-link-citations: yes
 biblio-style: vancouver
@@
 - needs to have a clear relevance to spatial patterns in cancer epidemiology
-- presents alternative ways of visualising those spatial patterns, or the uncertainty in those patterns
-- a more applied-focussed paper (e.g. lots of cool images), rather than theoretical (e.g. lots of formula).
+- presents alternative ways of visualizing those spatial patterns, or the uncertainty in those patterns
+- a more applied-focused paper (e.g. lots of cool images), rather than theoretical (e.g. lots of formula).
 -->
@@
 <!-- What data are we working with, why, what are the purposes -->
-Cancer statistics represent information on people, usually delivered as an aggregated value for a geopolitical area.
-Presenting statistics requires transforming individual observations into aggregations of communities as geographical units, in large part for privacy protection or political and policy purposes.
+Cancer statistics are usually delivered as an aggregated value for a geopolitical area.
+Presenting these statistics requires transforming individual observations into aggregations of communities as geographical units, in large part for privacy protection or political and policy purposes.
 The information could be as simple as counts per area (e.g.
 state, province, local government area, post/zip code).
-Counts alone is not sufficient on which to compare different areas, because the populations of different areas are likely different.
+Counts alone is not sufficient to compare different areas, because the populations of different areas are likely different.
 In this case, the counts data needs to be merged with population data to appropriately calibrate it to incidence, for example, rate per 100000 people.
@@
 
-<!-- Geospatial statistics have been presented on maps to communicate information to the general public for centuries as they utilise familiar geographic shapes.
-The mapping of diseases summarises spatial variations and can be considered a descriptive statistic of the relationship between a disease and the underlying geography [@SEMA].
-The distribution of the statistic must be visible, it is not enough for areas on maps to be recognisable.
-It is important to explore map views and designs that enhance the communication of the cancer statistics, especially as it is the populations that live on the land that are of interest, not the land itself.
-This has spurred innovations over the previous centuries to enhance maps to effectively communicate both cancer outcomes and health outcomes more broadly.-->
-
-<!-- Ways to visualise, whats our hypothesis -->
+<!-- Ways to visualize, whats our hypothesis -->
 A common approach to communicate cancer statistics, is to display statistics on a map.
-This is a choropleth map: the statistic is mapped to colour and the geographic region is filled with this colour.
-The viewer would then be able to examine the spatial distribution of the disease incidence, where there is a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hotspots of the disease.
-Visualising diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH].
+Using a choropleth map: the statistic is mapped to color and the geographic region is filled with this color.
+The viewer would then be able to examine the spatial distribution of the disease incidence, where there is a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease.
+Visualizing diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH].
 Disease maps help to present geographic patterns that may be overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
-By providing a visual representation of cancer outcomes, geographic patterns of disease are able to be identified and effectively addressed with public healh policy and actions.
-@SE recognises one of the key challenges with mapping spatial patterns of disease is the design of visualisations.
-This paper addresses the visualisation techniques and their applications to cancer statistics. Highlighting the differences and historic use of these displays.
+By providing a visual representation of cancer outcomes, geographic patterns of disease are able to be identified and effectively addressed with public health policy and actions.
+@SE recognizes one of the key challenges with mapping spatial patterns of disease is the design of visualizations.
+This paper addresses the visualization techniques and their applications to cancer statistics. Highlighting the differences and historic use of these displays.
 
@@
 Disease maps are more useful when made interactive, and common options are described in Section \ref{ch:interacting}, along with a discussion of benefits and disadvantages.
-The last section summarises the paper and discusses future directions.
+The last section summarizes the paper and discusses future directions.
 
@@
 A choropleth map is used to display differences in the geographical distribution of data by spatial unit by shading areas of a map.
-The geography is faithfully rendered, and the colour rendering is designed to reveal spatial patterns among data values.
-A choropleth is constructed by drawing the geographic or political boundaries, and filling the shapes with colours to represent values of a measured variable [@EI].
+The geography is faithfully rendered, and the color rendering is designed to reveal spatial patterns among data values.
+A choropleth is constructed by drawing the geographic or political boundaries, and filling the shapes with colors to represent values of a measured variable [@EI].
 Figure \ref{fig:choroCRS} shows a choropleth of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged over 2012 through 2016.
@@
 <!-- history and types of choropleths, variations -->
-Early versions of choropleth maps used symbols or patterns instead of colour.
-@CPISACA discuss the use of choropleths to visualise cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
+Early versions of choropleth maps used symbols or patterns instead of color.
+@CPISACA discuss the use of choropleths to visualize cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
 
 <!-- psychology -->
-Utilising the state boundaries can make a map familiar to read [@CIBMUK], and allows viewers to visually infer the spatial relationships in the data. 
+Utilizing the state boundaries can make a map familiar to read [@CIBMUK], and allows viewers to visually infer the spatial relationships in the data. 
 The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
@@
 <!-- how do we use choropleth's -->
-Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS].
+Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organizing communities into representable units [@SAMGIS].
 In Figure \ref{fig:choroCRS}, a west to east spatial trend of increasing rates, can be seen.
@@
 <!-- potential problems -->
-While the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
-The different population and geographical sizes of administrative areas can attract attention to the shades of the unpopulated but large areas [@EI], Skowronnek [-@BCM] calls this an area-size bias. 
+While the areas are recognizable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
+The different population and geographical sizes of administrative areas can attract attention to the shades of the underpopulated but large areas [@EI], Skowronnek [-@BCM] calls this an area-size bias.
 Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention away from the 'potentially more important results in the more populous communities' that are geographically smaller [@SE].
@@
 In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality. 
-The data collection methods of cancer mortality rates across regions, and the administrative control within regions lends itself to choropleth visualisation. 
+The data collection methods of cancer mortality rates across regions, and the administrative control within regions lends itself to choropleth visualization. 
 @MACM provides the definition of a cancer atlas, beginning with Haviland's maps in 1875, they attribute UK cancer atlases to Howe (1963), and early work in US cancer atlases can be attributed to Burbank (1971).
 The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
-The choropleth maps presented levels via hatchings or dots on a black and white scale.
+The choropleth maps presented levels via hatching or dots on a black and white scale.
 These atlases were key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets [@MACM].
@@
 Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the higher prevalence of cancers in older populations.
-Howe [-@HEDP] describes Stock's development of the standardised mortality ratios through the 1930s.
-Table \ref{tab:measures} summarises the measures presented in published cancer atlases, and provides a definition of each measure.
+Howe [-@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s.
+Table \ref{tab:measures} summarizes the measures presented in published cancer atlases, and provides a definition of each measure.
 
@@
 ||Cancer incidence rate in region $i$ over the average cancer incidence rate for the total region|
-|5. Age Adjusted Rate per 100,000| #5 standardised by age structure or region|   
-|6. Age Adjusted Relative Risk| RR standardised by age structure in each region $i$|  
-|7. SIR (Standardised Incidence Ratio)| IR standardised by age structure in each region $i$|
+|5. Age Adjusted Rate per 100,000| #5 standardized by age structure or region|   
+|6. Age Adjusted Relative Risk| RR standardized by age structure in each region $i$|  
+|7. SIR (Standardized Incidence Ratio)| IR standardized by age structure in each region $i$|
 |8. Below or above Expected|Alternative expression of the SIR|
@@
 
-The atlas of Cancer in Queensland [@DACQ] focussed on highlighting the difference in experience for those living in rural and disadvantaged areas.
-They used Standardised Incidence Ratios modelled via Bayesian methods, and explored the information presented in a range of selected atlases.
+The atlas of Cancer in Queensland [@DACQ] focused on highlighting the difference in experience for those living in rural and disadvantaged areas.
+They used Standardized Incidence Ratios modeled via Bayesian methods, and explored the information presented in a range of selected atlases.
 These atlases provided Incidence, Survival and Mortality rates to the public on Areal or Isopleth map bases.
-
-The presentation of these rates considers not only the shapes of the areas, but also appropriate choices of colour blind friendly colour schemes, and categories of the values to highlight significantly different areas.
-<!--
-
-Examples of atlases:
-Kraak 1998, Kraak and Ormeling 2011
-Bertin 1967
-Coleman MP, Gatta G, Verdecchia A, Esteve J, Sant M, Storm H, Allemani C,Ciccolallo L, Santaquilani M, Berrino F:EUROCARE-3 summary: cancersurvival in Europe at the end of the 20th century.Ann Oncol2003,14(Suppl 5):v128-149
-
-National Cancer Institute:Cancer Trends Progress Report - 2009/2010 UpdateBethesda: NCI, NIH, DHHS; 2010
-
-Pickle LW:A history and critique of U.S. mortality atlases.Spatial andSpatio-temporal Epidemiology2009,1:3-17.
-
-Mason TJ, McKay FW, Hoover R, Blot WJ, Fraumeni JF:Atlas of cancermortality for U.S. counties: 1950-1969Washington: U.S. Govt. Printing Office;1975.
-howmick T, Robinson AC, Gruver A, MacEachren AM, Lengerich EJ:Distributed usability evaluation of the Pennsylvania Cancer Atlas.Int JHealth Geogr2008,7:36.
-
-Queensland Cancer Registry:Cancer in Queensland: Incidence, Mortality,Survival and Prevalence,1982 to 2007Brisbane: QCR, Cancer CouncilQueensland and Queensland Health; 2010.
-
-New York State Department of Health Maps of Cancer Incidence byCounty.[http://www.health.state.ny.us/statistics/cancer/registry/cntymaps/index.htm].
-Quinn M, Wood H, Cooper N, Rowan S, Eds:Cancer Atlas of the UnitedKingdom and Ireland 1991-2000.London: Office for National Statistics;2005.
--->
+The presentation of these rates considers not only the shapes of the areas, but also appropriate choices of color blind friendly color schemes, and categories of the values to highlight significantly different areas.
 
 Cancer maps are effective tools for communicating to wide range of audiences, including the general public and others not trained in statistical analyses.
-These visualisations enable non-expert audiences to interpret the outputs of sophisticated statistical analyses.
+These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analyses.
 Cruickshank (1947) as cited by @DMAHP, discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
@@
 
-<!-- While there is significant variation across these maps in terms of the measure represented, the scale of the map, the colour choices, the legend design and supplementary/supporting statistics, the mapping approach has not changes and almost exclusively only choropleth maps are used.
-Unlike innovation in spatial statistical methods, interactity, data collection and record keeping little work has been done to implement different mapping approaches -->
+<!-- While there is significant variation across these maps in terms of the measure represented, the scale of the map, the color choices, the legend design and supplementary/supporting statistics, the mapping approach has not changes and almost exclusively only choropleth maps are used.
+Unlike innovation in spatial statistical methods, interactivity, data collection and record keeping little work has been done to implement different mapping approaches -->
 
@@
 For example, changing the population age or other demographic variables.
-Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactivities.
+Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactive capabilities.
 
@@
 Tool tip features and drop down menus allow for user exploration, small screens are not a problem with the flexibility provided by zooming and panning [@HTLWM].
-This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accesible' ways [@TNTEA].This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accesible' ways [@TNTEA].
-The interactive features of the publically available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accessible' ways [@TNTEA].This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
+This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accessible' ways [@TNTEA].
+The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
 @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
@@
 
-Cancer maps are powerful visualisations that summarise complex statistical analyses, however the statistics represented in these maps cannot tell the entire story.
+Cancer maps are powerful visualizations that summarize complex statistical analyses, however the statistics represented in these maps cannot tell the entire story.
 Supplementary graphs and plots are often included to add more depth and information to the map.
 @CPISACA suggests additional materials such as tables, graphs, and text explanations support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers.
-There are many visualisations used for displays of statistical summaries, these may be dot plots, bar plots, box plots, timesseries plots, cumulative distribution plots, scatter plots, Q-Q plots. 
+There are many visualizations used for displays of statistical summaries, these may be dot plots, bar plots, box plots, times series plots, cumulative distribution plots, scatter plots, Q-Q plots. 
 These additional displays of the cancer distribution can provide alternative views of the cancer statistics, as well as the supporting statistics including error, confidence intervals, distributions, sample or population sizes, standard deviation and other measures.
@@
 Other methods involve providing adjacent maps or overlapping maps with symbols [@VSSDCUC].
-The maps employing uncertainty ranged from static documents or infographics, to interactive online visualisations.
-Communicating the statistical uncertainty associated with the estimates occurs using confidence intervals (CI), credible intervals (CrI), statistical significance levels, boxplots, distribution plots, and reporting sample size and standard deviations.
+The maps employing uncertainty ranged from static documents or info-graphics, to interactive online visualizations.
+Communicating the statistical uncertainty associated with the estimates occurs using confidence intervals (CI), credible intervals (CrI), statistical significance levels, box plots, distribution plots, and reporting sample size and standard deviations.
 Close to half of the atlases identified (42%, n=14) included some measure of uncertainty. The most common measure used to represent uncertainty were credible or confidence intervals (CIs).
@@
 For example, changing the population age or other demographic variables.
-Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactivities.
+Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection tools.
 
@@
 This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accesible' ways [@TNTEA].
-The interactive features of the publically available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accessible' ways [@TNTEA].
+The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
 @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
@@
 
-Choropleths are univariate displays, they do not inherently support multi-dimensionality
+Choropleths are uni-variate displays, they do not inherently support multi-dimensionality.
 This makes it difficult to pair demographic, environmental and other factors with spatial distributions.
@@
 ```{r choroCRS, fig.cap = "Four choropleth maps of the United States of America using various coordinate reference systems.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been colored according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
 The map projections alter the shapes and angles of the boundaries of each state.
@@
 <!-- Common variables used to create cartograms, e.g. population, mortality -->
-Cartograms provide an alternative visualisation method for statistical and geographical information.
+Cartograms provide an alternative visualization method for statistical and geographical information.
 The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC]. 
@@
 The spatial transformation of map regions relative to the data emphasises the data distribution instead of land size [@CBATCC].
-When visualising population statistics, @ACTUC considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
+When visualizing population statistics, @ACTUC considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
 @HEDP suggests that 'cancer occurs in people, not in geographical areas' and @CTTMB believe that spatial socio-economic data, like cancer rates, are best presented on a cartogram for urban areas as the population map base avoids allocating 'undue prominence' to rural areas.
@@
 Table: \label{tab:usa} Maps used to present statistics for the United States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been colored according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
 
@@
 The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.|
+The sparsity of the population in the Mid West is highlight by the light yellow color, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.|
 
 ```{r auscartograms, fig.cap = "Alternative maps of the United States of America.
-Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been colored according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
 Each stated shape has been distorted according to the population of the state in 2015.
@@
 Figure \ref{fig:ggcartograms} a shows a population contiguous cartogram of the United States.
-All states are visible and the shape of the United States overall is still recognisable.
+All states are visible and the shape of the United States overall is still recognizable.
 In contrast, Figure \ref{fig:auscartograms} a shows an Australian contiguous cartogram also based on population.
-The south east is completely distorted, areas with low popuation are still large very the map, as their initial area was used to determine the area allocated in the contiguous cartogram.
+The south east is completely distorted, areas with low population are still large very the map, as their initial area was used to determine the area allocated in the contiguous cartogram.
 
@@
 This implementation can work for up to 'one hundred thousand' areas.
-Figure \ref{fig:ggcartograms} c shows the displacement of the north east region of the United States in the Dorling cartogram, the rest of the country is still recognisable.
+Figure \ref{fig:ggcartograms} c shows the displacement of the north east region of the United States in the Dorling cartogram, the rest of the country is still recognizable.
 This displacement is shown to an extreme in \ref{fig:auscartograms} c the Dorling cartogram of Australia.
@@
 This rectangular display also allows for tiling, where geographic neighbours placed in suitable relative positions also share borders, however contiguity may be sacrificed [@CDWCS].
-Rectangular cartograms allow for bi-variate displays, population can be effectively communicated by the size of each rectangular, and a second variable can be communicated using colour [@ORC].
+Rectangular cartograms allow for bi-variate displays, population can be effectively communicated by the size of each rectangular, and a second variable can be communicated using color [@ORC].
 
@@
 A simple tile map presents the areas in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
-These tiles may be labelled or coloured to represent a value.
+These tiles may be labelled or colored to represent a value.
 Tile maps may be difficult to create, they are best created manually, with additional time and care required as the amount of geographic areas to include increases.
@@
 ```{r ggtilemap, fig.cap = "A tile map and geofaceted map of the Unites States of America.
-Each state in the tile map has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state in the tile map has been colored according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
 Each state has been represented by a square, each square is the same size.
@@
 ## Geofaceting
-Tile maps can be extended to allow more than just simple colouring of tiles to present data.
-@IGF formalises the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas. 
-Like tile maps, the arrangement of the areas can be reused for other spatial distribution visulisations.
-Geofaceting have the functionality of facets, often used to replicate visualisations for each subset of the data, in this case the data subsets are geographic.
-The amount of information able to be communicated has increased from one value per region in a tile map to one visualisation, this is a more flexible display.
+Tile maps can be extended to allow more than just simple coloring of tiles to present data.
+@IGF formalizes the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas. 
+Like tile maps, the arrangement of the areas can be reused for other spatial distribution visualizations.
+Geofaceting have the functionality of facets, often used to replicate visualizations for each subset of the data, in this case the data subsets are geographic.
+The amount of information able to be communicated has increased from one value per region in a tile map to one visualization, this is a more flexible display.
 Virtually any plot display can be fit into the tile representing the areas, allowing displays of multiple variables or values per geographic entity.
@@
 Choropleth methods require a decision to be made about the projection of the display, in this case aim to find a map projection that gives the least distorted representation of the geography [@UMPT].
-However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a colour.
+However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a color.
 These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
@@
 Small extreme values are often lost due to the resulting small sized areas the do not draw user's attention.
-The change in the map based on one variable provides the opportunity for bi-variate displays, using the change in area for one variable, and colour for another.
+The change in the map based on one variable provides the opportunity for bi-variate displays, using the change in area for one variable, and color for another.
 The difference between the familiar map display helps highlight the impact of the disease on communities.
@@
 Areas that are geographically close will maintain connectedness in some way, but the population and the statistic will dominate.
-Tile map displays may only allow one variable to be visualised if each region is given consistent map space, this occurs in geofaceting. 
-Dorlings provide bi-variate displays as they use the size of the shape for one variable, and colour for another.
+Tile map displays may only allow one variable to be visualized if each region is given consistent map space, this occurs in geofaceting. 
+Dorlings provide bi-variate displays as they use the size of the shape for one variable, and color for another.
 These alternatives are especially helpful for data aggregations where administrative boundaries break populations into groups.
@@
 | Bi-variate display^[A bi-variate display refers to a map that allows two variables to be plotted.
-One variable via the size of the area, and another by filling the area with a colour.] | N | Y | Y | Y | Y |
+One variable via the size of the area, and another by filling the area with a color.] | N | Y | Y | Y | Y |
 
@@
 Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
-Tooltips allow users to obtain additional information regarding specific geographic areas.
+Tool tips allow users to obtain additional information regarding specific geographic areas.
 
@@
 It was in the development of this online cancer atlas that methods for disease map displays, and visual communication strategies were explored.
-We are thankful for the opportunity to write about these visualisations and the situations in which they are appropriate.
+We are thankful for the opportunity to write about these visualizations and the situations in which they are appropriate.
 
@@
 For data analysis the tidyverse package [@tidyverse] provided many useful functions.
-ggplot2 package [@ggplot2] was used to create maps, with RColorBrewer package [@RColorBrewer] providing additional colour palettes and ggthemes [@ggthemes] package providing themes.
+ggplot2 package [@ggplot2] was used to create maps, with RColorBrewer package [@RColorBrewer] providing additional color palettes and ggthemes [@ggthemes] package providing themes.
 The png package [@png] was used to access png images taken from online web atlases, cowplot package [@cowplot] was used to arrange these plots into grouped images.

### 2019-09-24 review_article.Rmd
spell check for US and missing years added to references

@@
 output:
-  word_document:
-    toc: yes
-    toc_depth: '3'
   pdf_document:
@@
     toc_depth: 3
+  word_document:
+    toc: yes
+    toc_depth: '3'
 link-citations: yes
@@
 
-- Future steps should consider applications not currently utilised for cancer outcome displays, where their use may improve decision making
+- Future steps should consider applications not currently utilized for cancer outcome displays, where their use may improve decision making
 - needs to have a clear relevance to spatial patterns in cancer epidemiology
@@
 state, province, local government area, post/zip code).
-Counts alone is not sufficient to compare different areas, because the populations of different areas are likely different.
+Counts alone are not sufficient to compare areas, because the populations of areas are all different.
 In this case, the counts data needs to be merged with population data to appropriately calibrate it to incidence, for example, rate per 100000 people.
@@
 Disease maps help to present geographic patterns that may be overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
-By providing a visual representation of cancer outcomes, geographic patterns of disease are able to be identified and effectively addressed with public health policy and actions.
+By providing a visual representation of cancer outcomes, geographic patterns of disease may be identified and effectively addressed with public health policy and actions.
 @SE recognizes one of the key challenges with mapping spatial patterns of disease is the design of visualizations.
@@
 <!-- Structure of the paper -->
-The paper is organised as follows.
+The paper is organized as follows.
 The next section describes the choropleth map which is the common approach to disease maps.
@@
 In Figure \ref{fig:choroCRS}, a west to east spatial trend of increasing rates, can be seen.
-There is also a spatial outlier -- Utah has a noticeably lower rate that its neighbours.
-Also Kentucky has a noticeably high rate, and Maine also has a higher rate than its neighbours.
+There is also a spatial outlier -- Utah has a noticeably lower rate that its neighbors.
+Also Kentucky has a noticeably high rate, and Maine also has a higher rate than its neighbors.
 There is something of a cluster of higher rates around the tobacco states.
@@
 
-```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Publicly available choropleth cancer maps, published between 2010 and 2015.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Publicly available choropleth cancer maps published between 2010 and 2015.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
 library(cowplot)
@@
 
-The atlas of Cancer in Queensland [@DACQ] focused on highlighting the difference in experience for those living in rural and disadvantaged areas.
-They used Standardized Incidence Ratios modeled via Bayesian methods, and explored the information presented in a range of selected atlases.
-These atlases provided Incidence, Survival and Mortality rates to the public on Areal or Isopleth map bases.
-The presentation of these rates considers not only the shapes of the areas, but also appropriate choices of color blind friendly color schemes, and categories of the values to highlight significantly different areas.
-
 Cancer maps are effective tools for communicating to wide range of audiences, including the general public and others not trained in statistical analyses.
@@
 Cruickshank (1947) as cited by @DMAHP, discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
-Overwhelmingly, cancer maps utilised to communicate to the public and other non-expert audiences are choropleths.
+Overwhelmingly, cancer maps utilized to communicate to the public and other non-expert audiences are choropleths.
+These atlases provided Incidence, Survival and Mortality rates to the public on areal or isopleth map bases.
+
+The Atlas of Cancer in Queensland [@DACQ] focused on highlighting the difference in experience for those living in rural and disadvantaged areas, the Standardized Incidence Ratios (SIR) were modeled via Bayesian methods.
+The presentation of these rates considers not only the shapes of the areas, but also appropriate choices of color blind friendly color schemes, and categories of the values to highlight significantly different areas.
 
@@
 All of these use choropleth maps.
-All but one of these were published by non-commercial organisations, including not-for-profits, government, research organisations, advocacy groups or government funded partnerships.
+All but one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government funded partnerships.
 The use of choropleths within the public domain mirrors the heavy use of choropleth maps within the research literature, discussed above.
 The cancer atlases identified by @roberts2019communication covered geographies from all around the world, most focused on single nations.
-Figure \ref{fig:choropleth-grid-create} displays a global (b), national (a,d,e,f) and state (c,g) choropleths.
+Figure \ref{fig:choropleth-grid-create} displays a global (b), national (a, d, e, f) and state (c, g) choropleths.
 The sections below provide details on each of the maps within Figure \ref{fig:choropleth-grid-create}.
@@
 
-Figure 2.a contains an image from the *[The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/)*.
+Figure 2.a contains an image from *[The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/)*.
 This map shows the relative risk for women developing lung cancer in England and Wales in 2010.
@@
 This global map shows age standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018.
-This map is published by the [World Health Organisation's International Agency for Research on Cancer](https://www.iarc.fr/).
+This map is published by the [World Health organization's International Agency for Research on Cancer](https://www.iarc.fr/).
 Data was sourced from cancer registries in each country, contributing registries can be seen in the supplementary material on the cancer atlas website.
@@
 
-Figure 2.c shows an extract from the *[The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/)*.
+Figure 2.c shows an extract from  *[The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/)*.
 This map was published by the Queensland Cancer Council and shows the relative incidence ratio of lung cancer in males in the state of QLD within Australia based on data from 1998 to 2007.
@@
 
-One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the the user reaches an information threshold, beyond which will not be able to make sense of the information.
+One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
 These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
@@
 
-The interactivity of modern mapping methods enable supplementary information to be incorporated without cluttering the screen.
+The interactivity of modern mapping methods enables supplementary information to be incorporated without cluttering the screen.
 Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
@@
 While atlases are often used to describe differences between areas, statistics may be displayed at different levels of aggregation.
-Global health statistics can be aggregated to administrative and arbitrarily defined regions, such as those used by the World Health Organisation and the United Nations [@IARC_3].
+Global health statistics can be aggregated to administrative and arbitrarily defined regions, such as those used by the World Health organization and the United Nations [@IARC_3].
 World atlases can allow for displays of data aggregated into continents, countries, states, provinces and congressional districts [@USCS].
@@
 The distribution of the population residing in all areas may also be communicated in a table or histogram display [@NICR_1].
-Atlases have the opportunity to connect the population to the land available to them by communicating population density.
+Atlases can connect the population to the land available to them by communicating population density.
 Atlases can also connect the population to the land available to them by communicating population density.
@@
 In the review of atlases in the public domain, atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map [@roberts2019communication].
-The maps considered used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
+The maps considered used standard and well-known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
 Other methods involve providing adjacent maps or overlapping maps with symbols [@VSSDCUC].
-The maps employing uncertainty ranged from static documents or info-graphics, to interactive online visualizations.
+The maps employing uncertainty ranged from static documents or infographics, to interactive online visualizations.
 Communicating the statistical uncertainty associated with the estimates occurs using confidence intervals (CI), credible intervals (CrI), statistical significance levels, box plots, distribution plots, and reporting sample size and standard deviations.
@@
 The Human Development Index can be used to understand the socio-economic experience of a community, as well as Income levels which can be sourced from the World Bank list of economies [@IARC_3]. 
-The areas are often ranked and allocated to quintiles, each quintile can be presented as categories describing the ranking.
+The areas are often ranked and allocated to quintiles; each quintile can be presented as categories describing the ranking.
 
-One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the the user reaches an information threshold, beyond which will not be able to make sense of the information.
+One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
 These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
@@
 it shows one map per group of the key separating variable, in a row with each additional statistical plot.
-Linked micromaps predominantly utilise the choropleth map for displays of spatial distributions.
+Linked micromaps predominantly utilize the choropleth map for displays of spatial distributions.
 It is one of several alternative displays that allow maps to become bi-variate displays.
@@
 <!-- What is a cartogram? -->
-Choropleths imply uniformity of data across the geographic space, when considering population densities it is extremely unlikely to be uniform [@BCM].
+Choropleths imply uniformity of data across the geographic space however population densities are extremely unlikely to be uniform [@BCM].
 A cartogram displays abstracted geographic space, with the intention of improving the presentation of the statistic of interest.
-For a single variable of interest, each map area is changed to emphasise the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].
+For a single variable of interest, each map area is changed to emphasize the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].
 Changes in the map base are implemented by altering the boundaries, and therefore shapes, of individual areas.
@@
 Choropleth maps will always be distorted if they cover enough of the globe, just like photographs of the globe from space.
-Choropleth creation requires choosing a map projection that shows a favourable distortion of the geography for presenting the set of spatial information.
+Choropleth creation requires choosing a map projection that shows a favorable distortion of the geography for presenting the set of spatial information.
 Selecting a display can prevent misinterpretation of global statistics, as global maps face the challenge of equitable displays of land mass on maps [@RSCW].
@@
 @HTLWM suggests that white lies may be employed to create useful displays and map creators have the ability to draw lines that may distort the geometry and suppress features and it is easy for the average person to disregard the impact of transformations used to create cartograms.
-Cartograms may be seen as an extension of map transformations and projections. The favourable distortion is proportional to a value other other than actual earth size area [@NAC].
+Cartograms may be seen as an extension of map transformations and projections. The favorable distortion is proportional to a value other other than actual earth size area [@NAC].
 A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, making interpretation of spatial patterns very difficult.
 The distortion of a cartogram accounts for the population density, preventing it from obscuring the spatial patterns [@TAAM]. 
-The spatial transformation of map regions relative to the data emphasises the data distribution instead of land size [@CBATCC].
+The spatial transformation of map regions relative to the data emphasizes the data distribution instead of land size [@CBATCC].
 When visualizing population statistics, @ACTUC considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
@@
 Early approaches including John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball bearing approach and Tobler's (1973) computer programs [@ACTUC].
-Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994), [@CD].
+Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilized depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994), [@CD].
 @HEDP discusses the impact of electronic computer-assisted techniques.
@@
 |Dorling| The states have been represented by a circle, but the size was determined by the population of the state in 2015.
-The North East states remain closer to their neighbours, and may be displaced from their geographic location.
+The North East states remain closer to their neighbors, and may be displaced from their geographic location.
 The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.|
@@
 The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the light yellow color, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.|
+The sparsity of the population in the Mid West is highlight by the light yellow color, the age adjusted rate in Kentucky is the darkest and its neighbors are similar.|
 
@@
 <!-- What is a contiguous cartogram -->
-<!-- Intentionally preserve neighbours -->
+<!-- Intentionally preserve neighbors -->
 
@@
 @TFYCC quotes the official definition of Value-Area Cartograms, the simplistic displays which represent each area as a single rectangle, sized according to the value of the statistic.
-This rectangular display also allows for tiling, where geographic neighbours placed in suitable relative positions also share borders, however contiguity may be sacrificed [@CDWCS].
+This rectangular display also allows for tiling, where geographic neighbors placed in suitable relative positions also share borders, however contiguity may be sacrificed [@CDWCS].
 Rectangular cartograms allow for bi-variate displays, population can be effectively communicated by the size of each rectangular, and a second variable can be communicated using color [@ORC].
@@
 
-A simple tile map presents the areas in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
+A simple tile map presents the areas in a tessellated grid display, where geographic neighbors are found next to each other, with some necessary displacement employed for regions with more than four neighbors.
 These tiles may be labelled or colored to represent a value.
@@
 To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
-Choropleth displays utilise more traditional cartographic methods, they are usually true to the topography, displaying familiar boundaries of countries, states or administrative areas.
+Choropleth displays utilize more traditional cartographic methods, they are usually true to the topography, displaying familiar boundaries of countries, states or administrative areas.
 Highlighting the geographic distribution, and differences between equal spatial units.
@@
 These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
-Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbours. The map is reusable and fit for many purposes.
+Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbors. The map is reusable and fit for many purposes.
 
 <!-- Cartograms, sacrifice geography, focus on the outliers/ relationships in data space -->
-Cartographers were historically the creators of geographic displays, and epidemiologists utilised the maps that had been prepared in advance.
+Cartographers were historically the creators of geographic displays, and epidemiologists utilized the maps that had been prepared in advance.
 Cartograms with the population variable used to distortion the geographic map regions and boundaries are density equalising maps, that reduce the visual prominence of large, but low-population areas [@VSSDCUC].
@@
 This can prevent effective comparison of the statistic.
-Cartograms also favour extreme statistics in the positive direction.
+Cartograms also favor extreme statistics in the positive direction.
 Small extreme values are often lost due to the resulting small sized areas the do not draw user's attention.
@@
 | Preserves shapes | Y | N | Y | N | N |
-| Preserves neighbours | Y | Y | Y | S | S |
+| Preserves neighbors | Y | Y | Y | S | S |
 | Uniform use of shape | N | N | N | Y | Y |
@@
 Interactions with online atlases can also change the amount of detail shown.
-It is uncommon to see linked displays utilised in online atlases, but there are several frameworks that support interactive displays of multivariate data.
+It is uncommon to see linked displays utilized in online atlases, but there are several frameworks that support interactive displays of multivariate data.
 Linked displays allow for direct manipulation, @RPIG uses Swayne and Klinke's definition of high level interaction that occurs via data selection or brushing tools that highlight subsets of graphical elements such as polygons or points.
@@
 <!-- Animations -->
-Animations communicate a story, showcasing a critical region by defining and progressing through a small set of keyframes, or waypoints @ESRI.
-The view the transitions between frames in a linear, fixed narrative to passive users, communicating a message by capturing and directing their attention [@TGA]. 
+Animations communicate a story, showcasing a critical region by defining and progressing through a small set of keyframes, or waypoints [@ESRI].
+@TGA explains that animations show the transitions between frames in a linear, fixed narrative to passive users, communicating a message by capturing and directing their attention. 
 @TGA suggests the use of moving objects demands attention by taking advantage of the behaviour of the visual cognition system.

### 2019-09-24 review_article.Rmd
Table corrections

@@
 ||Cancer incidence rate in region $i$ over the average cancer incidence rate for the total region|
-|5. Age Adjusted Rate per 100,000| #5 standardized by age structure or region|   
+|5. Age Adjusted Rate per 100,000| Standardized by age structure or region|   
 |6. Age Adjusted Relative Risk| RR standardized by age structure in each region $i$|  
@@
 This draws attention to the densely populated North East region, and the sparse Mid West.|
-|Dorling| The states have been represented by a circle, but the size was determined by the population of the state in 2015.
+|c.
+Dorling| The states have been represented by a circle, but the size was determined by the population of the state in 2015.
 The North East states remain closer to their neighbors, and may be displaced from their geographic location.
 The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.|
-|Hexagon Tessellation| Each state is represented by a hexagon of equal size.
+|d.
+Hexagon Tessellation| Each state is represented by a hexagon of equal size.
 The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.

### 2019-09-24 usa_plots.R
clean up geofacet plot

@@ -220,11 +220,12 @@ ggsave(filename = "figures/usa_grid.png", plot = usa_grid,
 library(statebins)
 
 ## US example
-ggtilegram <- statebins(cancer %>% rename(state = NAME),
+ggtilegram <- cancer %>% rename(state = NAME) %>% 
+  statebins(., state_col = "state",
     value_col = "AgeAdjustedRate", 
-    name = "AgeAdjustedRate",type = "seq", palette = "RdPu",  direction = 1
+    name = "AgeAdjustedRate",
+    type = "seq", palette = "RdPu",  direction = 1
   ) +
-  labs(title = "") + 
   theme_void() +
   theme(legend.position ="bottom")
 ggtilegram
@@ -234,7 +235,8 @@ ggsave(filename = "figures/ggtilegram.png", plot = ggtilegram,
 
 # Geofacet
 library(geofacet)
-us_grid <- read_csv("us_grid.csv")
+library(geogrid)
+us_grid <- read_csv("us_grid.csv") %>% filter(col>1)
 
 # read in data
 cancer_f <- read_csv("data/USCS_lung_f.csv") %>% mutate(Sex = "F")
@@ -248,16 +250,16 @@ cancer_mf <- bind_rows(cancer_f, cancer_m) %>%
 ggfacet <- ggplot(cancer_mf) + 
   geom_col(aes(x = Sex, y = AgeAdjustedRate, fill = Sex)) + 
   scale_fill_brewer(type = "qual", palette = "Pastel1",  direction = 1) + 
-  theme_void() + 
   facet_geo(~ state, grid = us_grid, label = "code") +
-  theme(legend.position ="bottom")
+  theme_minimal() + 
+  theme(legend.position ="bottom",axis.title.y=element_blank() )
 
 ggfacet
 ggsave(filename = "figures/ggfacet.png", plot = ggfacet,
-  device = "png", dpi = 300, width = 7, height = 6)
+  device = "png", dpi = 300, width = 10, height = 8)
 
-
-p <- plot_grid(ggtilegram, ggfacet, labels = c('A', 'B'), label_size = 12, align = v, )
+p <- plot_grid(rasterGrob(png::readPNG("figures/ggtilegram.png")),
+              rasterGrob(png::readPNG("figures/ggfacet.png")), labels = c('A', 'B'), label_size = 20)
 
 ggsave(filename = "figures/gggrids.png", plot = p,
[…]

### 2019-09-24 review_article.Rmd
cannot fit tilegrid on single page

@@
 
-As the trade-off regarding boundaries approaches simplicity, the distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the non-contiguous cartogram but is replaced by the meaningful empty-space property [@NAC, @ECGC].
+As the trade-off regarding boundaries approaches simplicity, the distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the non-contiguous cartogram but is replaced by the meaningful empty-space property [@ECGC, @NAC].
 The shapes are valuable for recognition and allows users to orient themselves on the display.
@@
 The geofaceted map allows the rates for males and females to be contrasted.
-It has the same layout as the tile map, but shows more information per state than ", message=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width = "100%"}
+It has the same layout as the tile map, but shows more information per state.",fig.height = 8, fig.width = 5}
 knitr::include_graphics("figures/gggrids.png")

### 2019-09-24 usa_plots.R

@@ -230,7 +230,7 @@ ggtilegram <- cancer %>% rename(state = NAME) %>%
   theme(legend.position ="bottom")
 ggtilegram
 ggsave(filename = "figures/ggtilegram.png", plot = ggtilegram,
-  device = "png", dpi = 300, width = 7, height = 6)
+  device = "png", dpi = 300, width =10, height = 8)
 
 
 # Geofacet
@@ -251,15 +251,16 @@ ggfacet <- ggplot(cancer_mf) +
   geom_col(aes(x = Sex, y = AgeAdjustedRate, fill = Sex)) + 
   scale_fill_brewer(type = "qual", palette = "Pastel1",  direction = 1) + 
   facet_geo(~ state, grid = us_grid, label = "code") +
-  theme_minimal() + 
-  theme(legend.position ="bottom",axis.title.y=element_blank() )
+  theme_classic() + 
+  theme(legend.position ="bottom",axis.title.y=element_blank(), axis.text.y = element_text(colour = "black"))
 
 ggfacet
 ggsave(filename = "figures/ggfacet.png", plot = ggfacet,
   device = "png", dpi = 300, width = 10, height = 8)
 
 p <- plot_grid(rasterGrob(png::readPNG("figures/ggtilegram.png")),
-              rasterGrob(png::readPNG("figures/ggfacet.png")), labels = c('A', 'B'), label_size = 20)
+          rasterGrob(png::readPNG("figures/ggfacet.png")), 
+          labels = c('A', 'B'), label_size = 20, ncol = 1)
 
 ggsave(filename = "figures/gggrids.png", plot = p,
-  device = "png", dpi = 300, width = 18, height = 9)
+  device = "png", dpi = 300, width = 8, height = 16)

### 2019-09-24 review_article.Rmd
working on conclusion

--- 2019-09-24 review_article.Rmd
@@
 ---
-title: "Cartogram Mapping and its Application to Cancer Data Visualization"
+title: "Cartogram Mapping and its Application to Cancer Data Visualisation"
 author: "Stephanie Kobakian, Jessie Roberts and Dianne Cook"
@@
   pdf_document:
-    fig_caption: yes
     number_sections: yes
@@
     toc_depth: 3
-  word_document:
-    toc: yes
-    toc_depth: '3'
+    fig_caption: true
+fontsize: 11pt
 link-citations: yes
-fontsize: 11pt
 biblio-style: vancouver
@@
 
-- Future steps should consider applications not currently utilized for cancer outcome displays, where their use may improve decision making
+- Future steps should consider applications not currently utilised for cancer outcome displays, where their use may improve decision making
 - needs to have a clear relevance to spatial patterns in cancer epidemiology
-- presents alternative ways of visualizing those spatial patterns, or the uncertainty in those patterns
-- a more applied-focused paper (e.g. lots of cool images), rather than theoretical (e.g. lots of formula).
+- presents alternative ways of visualising those spatial patterns, or the uncertainty in those patterns
+- a more applied-focussed paper (e.g. lots of cool images), rather than theoretical (e.g. lots of formula).
 -->
@@
 <!-- What data are we working with, why, what are the purposes -->
-Cancer statistics are usually delivered as an aggregated value for a geopolitical area.
-Presenting these statistics requires transforming individual observations into aggregations of communities as geographical units, in large part for privacy protection or political and policy purposes.
+Cancer statistics represent information on people, usually delivered as an aggregated value for a geopolitical area.
+Presenting statistics requires transforming individual observations into aggregations of communities as geographical units, in large part for privacy protection or political and policy purposes.
 The information could be as simple as counts per area (e.g.
 state, province, local government area, post/zip code).
-Counts alone are not sufficient to compare areas, because the populations of areas are all different.
+Counts alone is not sufficient on which to compare different areas, because the populations of different areas are likely different.
 In this case, the counts data needs to be merged with population data to appropriately calibrate it to incidence, for example, rate per 100000 people.
@@
 
-<!-- Ways to visualize, whats our hypothesis -->
+<!-- Geospatial statistics have been presented on maps to communicate information to the general public for centuries as they utilise familiar geographic shapes.
+The mapping of diseases summarises spatial variations and can be considered a descriptive statistic of the relationship between a disease and the underlying geography [@SEMA].
+The distribution of the statistic must be visible, it is not enough for areas on maps to be recognisable.
+It is important to explore map views and designs that enhance the communication of the cancer statistics, especially as it is the populations that live on the land that are of interest, not the land itself.
+This has spurred innovations over the previous centuries to enhance maps to effectively communicate both cancer outcomes and health outcomes more broadly.-->
+
+<!-- Ways to visualise, whats our hypothesis -->
 A common approach to communicate cancer statistics, is to display statistics on a map.
-Using a choropleth map: the statistic is mapped to color and the geographic region is filled with this color.
-The viewer would then be able to examine the spatial distribution of the disease incidence, where there is a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease.
-Visualizing diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH].
+This is a choropleth map: the statistic is mapped to colour and the geographic region is filled with this colour.
+The viewer would then be able to examine the spatial distribution of the disease incidence, where there is a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hotspots of the disease.
+Visualising diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH].
 Disease maps help to present geographic patterns that may be overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
-By providing a visual representation of cancer outcomes, geographic patterns of disease may be identified and effectively addressed with public health policy and actions.
-@SE recognizes one of the key challenges with mapping spatial patterns of disease is the design of visualizations.
-This paper addresses the visualization techniques and their applications to cancer statistics. Highlighting the differences and historic use of these displays.
+By providing a visual representation of cancer outcomes, geographic patterns of disease are able to be identified and effectively addressed with public healh policy and actions.
+@SE recognises one of the key challenges with mapping spatial patterns of disease is the design of visualisations.
+This paper addresses the visualisation techniques and their applications to cancer statistics. Highlighting the differences and historic use of these displays.
 
 <!-- Structure of the paper -->
-The paper is organized as follows.
+The paper is organised as follows.
 The next section describes the choropleth map which is the common approach to disease maps.
@@
 Disease maps are more useful when made interactive, and common options are described in Section \ref{ch:interacting}, along with a discussion of benefits and disadvantages.
-The last section summarizes the paper and discusses future directions.
+The last section summarises the paper and discusses future directions.
 
@@
 A choropleth map is used to display differences in the geographical distribution of data by spatial unit by shading areas of a map.
-The geography is faithfully rendered, and the color rendering is designed to reveal spatial patterns among data values.
-A choropleth is constructed by drawing the geographic or political boundaries, and filling the shapes with colors to represent values of a measured variable [@EI].
+The geography is faithfully rendered, and the colour rendering is designed to reveal spatial patterns among data values.
+A choropleth is constructed by drawing the geographic or political boundaries, and filling the shapes with colours to represent values of a measured variable [@EI].
 Figure \ref{fig:choroCRS} shows a choropleth of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged over 2012 through 2016.
@@
 <!-- history and types of choropleths, variations -->
-Early versions of choropleth maps used symbols or patterns instead of color.
-@CPISACA discuss the use of choropleths to visualize cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
+Early versions of choropleth maps used symbols or patterns instead of colour.
+@CPISACA discuss the use of choropleths to visualise cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
 
 <!-- psychology -->
-Utilizing the state boundaries can make a map familiar to read [@CIBMUK], and allows viewers to visually infer the spatial relationships in the data. 
+Utilising the state boundaries can make a map familiar to read [@CIBMUK], and allows viewers to visually infer the spatial relationships in the data. 
 The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
@@
 <!-- how do we use choropleth's -->
-Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organizing communities into representable units [@SAMGIS].
+Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organising communities into representable units [@SAMGIS].
 In Figure \ref{fig:choroCRS}, a west to east spatial trend of increasing rates, can be seen.
-There is also a spatial outlier -- Utah has a noticeably lower rate that its neighbors.
-Also Kentucky has a noticeably high rate, and Maine also has a higher rate than its neighbors.
+There is also a spatial outlier -- Utah has a noticeably lower rate that its neighbours.
+Also Kentucky has a noticeably high rate, and Maine also has a higher rate than its neighbours.
 There is something of a cluster of higher rates around the tobacco states.
@@
 <!-- potential problems -->
-While the areas are recognizable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
-The different population and geographical sizes of administrative areas can attract attention to the shades of the underpopulated but large areas [@EI], Skowronnek [-@BCM] calls this an area-size bias.
+While the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
+The different population and geographical sizes of administrative areas can attract attention to the shades of the unpopulated but large areas [@EI], Skowronnek [-@BCM] calls this an area-size bias. 
 Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention away from the 'potentially more important results in the more populous communities' that are geographically smaller [@SE].
 
-```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Publicly available choropleth cancer maps published between 2010 and 2015.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Publicly available choropleth cancer maps, published between 2010 and 2015.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
 library(cowplot)
@@
 In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality. 
-The data collection methods of cancer mortality rates across regions, and the administrative control within regions lends itself to choropleth visualization. 
+The data collection methods of cancer mortality rates across regions, and the administrative control within regions lends itself to choropleth visualisation. 
 @MACM provides the definition of a cancer atlas, beginning with Haviland's maps in 1875, they attribute UK cancer atlases to Howe (1963), and early work in US cancer atlases can be attributed to Burbank (1971).
 The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
-The choropleth maps presented levels via hatching or dots on a black and white scale.
+The choropleth maps presented levels via hatchings or dots on a black and white scale.
 These atlases were key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets [@MACM].
@@
 Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the higher prevalence of cancers in older populations.
-Howe [-@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s.
-Table \ref{tab:measures} summarizes the measures presented in published cancer atlases, and provides a definition of each measure.
+Howe [-@HEDP] describes Stock's development of the standardised mortality ratios through the 1930s.
+Table \ref{tab:measures} summarises the measures presented in published cancer atlases, and provides a definition of each measure.
 
@@
 ||Cancer incidence rate in region $i$ over the average cancer incidence rate for the total region|
-|5. Age Adjusted Rate per 100,000| #5 standardized by age structure or region|   
-|6. Age Adjusted Relative Risk| RR standardized by age structure in each region $i$|  
-|7. SIR (Standardized Incidence Ratio)| IR standardized by age structure in each region $i$|
+|5. Age Adjusted Rate per 100,000| #5 standardised by age structure or region|   
+|6. Age Adjusted Relative Risk| RR standardised by age structure in each region $i$|  
+|7. SIR (Standardised Incidence Ratio)| IR standardised by age structure in each region $i$|
 |8. Below or above Expected|Alternative expression of the SIR|
@@
 
+The atlas of Cancer in Queensland [@DACQ] focussed on highlighting the difference in experience for those living in rural and disadvantaged areas.
+They used Standardised Incidence Ratios modelled via Bayesian methods, and explored the information presented in a range of selected atlases.
+These atlases provided Incidence, Survival and Mortality rates to the public on Areal or Isopleth map bases.
+
+The presentation of these rates considers not only the shapes of the areas, but also appropriate choices of colour blind friendly colour schemes, and categories of the values to highlight significantly different areas.
+<!--
+
+Examples of atlases:
+Kraak 1998, Kraak and Ormeling 2011
+Bertin 1967
+Coleman MP, Gatta G, Verdecchia A, Esteve J, Sant M, Storm H, Allemani C,Ciccolallo L, Santaquilani M, Berrino F:EUROCARE-3 summary: cancersurvival in Europe at the end of the 20th century.Ann Oncol2003,14(Suppl 5):v128-149
+
+National Cancer Institute:Cancer Trends Progress Report - 2009/2010 UpdateBethesda: NCI, NIH, DHHS; 2010
+
+Pickle LW:A history and critique of U.S. mortality atlases.Spatial andSpatio-temporal Epidemiology2009,1:3-17.
+
+Mason TJ, McKay FW, Hoover R, Blot WJ, Fraumeni JF:Atlas of cancermortality for U.S. counties: 1950-1969Washington: U.S. Govt. Printing Office;1975.
+howmick T, Robinson AC, Gruver A, MacEachren AM, Lengerich EJ:Distributed usability evaluation of the Pennsylvania Cancer Atlas.Int JHealth Geogr2008,7:36.
+
+Queensland Cancer Registry:Cancer in Queensland: Incidence, Mortality,Survival and Prevalence,1982 to 2007Brisbane: QCR, Cancer CouncilQueensland and Queensland Health; 2010.
+
+New York State Department of Health Maps of Cancer Incidence byCounty.[http://www.health.state.ny.us/statistics/cancer/registry/cntymaps/index.htm].
+Quinn M, Wood H, Cooper N, Rowan S, Eds:Cancer Atlas of the UnitedKingdom and Ireland 1991-2000.London: Office for National Statistics;2005.
+-->
+
 Cancer maps are effective tools for communicating to wide range of audiences, including the general public and others not trained in statistical analyses.
-These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analyses.
+These visualisations enable non-expert audiences to interpret the outputs of sophisticated statistical analyses.
 Cruickshank (1947) as cited by @DMAHP, discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
-Overwhelmingly, cancer maps utilized to communicate to the public and other non-expert audiences are choropleths.
-These atlases provided Incidence, Survival and Mortality rates to the public on areal or isopleth map bases.
-
-The Atlas of Cancer in Queensland [@DACQ] focused on highlighting the difference in experience for those living in rural and disadvantaged areas, the Standardized Incidence Ratios (SIR) were modeled via Bayesian methods.
-The presentation of these rates considers not only the shapes of the areas, but also appropriate choices of color blind friendly color schemes, and categories of the values to highlight significantly different areas.
+Overwhelmingly, cancer maps utilised to communicate to the public and other non-expert audiences are choropleths.
 
@@
 All of these use choropleth maps.
-All but one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government funded partnerships.
+All but one of these were published by non-commercial organisations, including not-for-profits, government, research organisations, advocacy groups or government funded partnerships.
 The use of choropleths within the public domain mirrors the heavy use of choropleth maps within the research literature, discussed above.
 The cancer atlases identified by @roberts2019communication covered geographies from all around the world, most focused on single nations.
-Figure \ref{fig:choropleth-grid-create} displays a global (b), national (a, d, e, f) and state (c, g) choropleths.
+Figure \ref{fig:choropleth-grid-create} displays a global (b), national (a,d,e,f) and state (c,g) choropleths.
 The sections below provide details on each of the maps within Figure \ref{fig:choropleth-grid-create}.
 
-<!-- While there is significant variation across these maps in terms of the measure represented, the scale of the map, the color choices, the legend design and supplementary/supporting statistics, the mapping approach has not changes and almost exclusively only choropleth maps are used.
-Unlike innovation in spatial statistical methods, interactivity, data collection and record keeping little work has been done to implement different mapping approaches -->
+<!-- While there is significant variation across these maps in terms of the measure represented, the scale of the map, the colour choices, the legend design and supplementary/supporting statistics, the mapping approach has not changes and almost exclusively only choropleth maps are used.
+Unlike innovation in spatial statistical methods, interactity, data collection and record keeping little work has been done to implement different mapping approaches -->
 
@@
 
-Figure 2.a contains an image from *[The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/)*.
+Figure 2.a contains an image from the *[The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/)*.
 This map shows the relative risk for women developing lung cancer in England and Wales in 2010.
@@
 This global map shows age standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018.
-This map is published by the [World Health organization's International Agency for Research on Cancer](https://www.iarc.fr/).
+This map is published by the [World Health Organisation's International Agency for Research on Cancer](https://www.iarc.fr/).
 Data was sourced from cancer registries in each country, contributing registries can be seen in the supplementary material on the cancer atlas website.
@@
 
-Figure 2.c shows an extract from  *[The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/)*.
+Figure 2.c shows an extract from the *[The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/)*.
 This map was published by the Queensland Cancer Council and shows the relative incidence ratio of lung cancer in males in the state of QLD within Australia based on data from 1998 to 2007.
@@
 
-One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
+One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the the user reaches an information threshold, beyond which will not be able to make sense of the information.
 These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
@@
 For example, changing the population age or other demographic variables.
-Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactive capabilities.
+Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactivities.
 
@@
 
-The interactivity of modern mapping methods enables supplementary information to be incorporated without cluttering the screen.
+The interactivity of modern mapping methods enable supplementary information to be incorporated without cluttering the screen.
 Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
@@
 Tool tip features and drop down menus allow for user exploration, small screens are not a problem with the flexibility provided by zooming and panning [@HTLWM].
-This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accessible' ways [@TNTEA].This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accessible' ways [@TNTEA].
-The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accesible' ways [@TNTEA].This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
+This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accesible' ways [@TNTEA].
+The interactive features of the publically available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
 @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
@@
 
-Cancer maps are powerful visualizations that summarize complex statistical analyses, however the statistics represented in these maps cannot tell the entire story.
+Cancer maps are powerful visualisations that summarise complex statistical analyses, however the statistics represented in these maps cannot tell the entire story.
 Supplementary graphs and plots are often included to add more depth and information to the map.
 @CPISACA suggests additional materials such as tables, graphs, and text explanations support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers.
-There are many visualizations used for displays of statistical summaries, these may be dot plots, bar plots, box plots, times series plots, cumulative distribution plots, scatter plots, Q-Q plots. 
+There are many visualisations used for displays of statistical summaries, these may be dot plots, bar plots, box plots, timesseries plots, cumulative distribution plots, scatter plots, Q-Q plots. 
 These additional displays of the cancer distribution can provide alternative views of the cancer statistics, as well as the supporting statistics including error, confidence intervals, distributions, sample or population sizes, standard deviation and other measures.
@@
 While atlases are often used to describe differences between areas, statistics may be displayed at different levels of aggregation.
-Global health statistics can be aggregated to administrative and arbitrarily defined regions, such as those used by the World Health organization and the United Nations [@IARC_3].
+Global health statistics can be aggregated to administrative and arbitrarily defined regions, such as those used by the World Health Organisation and the United Nations [@IARC_3].
 World atlases can allow for displays of data aggregated into continents, countries, states, provinces and congressional districts [@USCS].
@@
 The distribution of the population residing in all areas may also be communicated in a table or histogram display [@NICR_1].
-Atlases can connect the population to the land available to them by communicating population density.
+Atlases have the opportunity to connect the population to the land available to them by communicating population density.
 Atlases can also connect the population to the land available to them by communicating population density.
@@
 In the review of atlases in the public domain, atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map [@roberts2019communication].
-The maps considered used standard and well-known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
+The maps considered used standard and well known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
 Other methods involve providing adjacent maps or overlapping maps with symbols [@VSSDCUC].
-The maps employing uncertainty ranged from static documents or infographics, to interactive online visualizations.
-Communicating the statistical uncertainty associated with the estimates occurs using confidence intervals (CI), credible intervals (CrI), statistical significance levels, box plots, distribution plots, and reporting sample size and standard deviations.
+The maps employing uncertainty ranged from static documents or infographics, to interactive online visualisations.
+Communicating the statistical uncertainty associated with the estimates occurs using confidence intervals (CI), credible intervals (CrI), statistical significance levels, boxplots, distribution plots, and reporting sample size and standard deviations.
 Close to half of the atlases identified (42%, n=14) included some measure of uncertainty. The most common measure used to represent uncertainty were credible or confidence intervals (CIs).
@@
 The Human Development Index can be used to understand the socio-economic experience of a community, as well as Income levels which can be sourced from the World Bank list of economies [@IARC_3]. 
-The areas are often ranked and allocated to quintiles; each quintile can be presented as categories describing the ranking.
+The areas are often ranked and allocated to quintiles, each quintile can be presented as categories describing the ranking.
 
-One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
+One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the the user reaches an information threshold, beyond which will not be able to make sense of the information.
 These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
@@
 For example, changing the population age or other demographic variables.
-Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection tools.
+Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactivities.
 
@@
 This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accessible' ways [@TNTEA].
-The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accesible' ways [@TNTEA].
+The interactive features of the publically available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
 @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
@@
 
-Choropleths are uni-variate displays, they do not inherently support multi-dimensionality.
+Choropleths are univariate displays, they do not inherently support multi-dimensionality
 This makes it difficult to pair demographic, environmental and other factors with spatial distributions.
@@
 it shows one map per group of the key separating variable, in a row with each additional statistical plot.
-Linked micromaps predominantly utilize the choropleth map for displays of spatial distributions.
+Linked micromaps predominantly utilise the choropleth map for displays of spatial distributions.
 It is one of several alternative displays that allow maps to become bi-variate displays.
@@
 <!-- What is a cartogram? -->
-Choropleths imply uniformity of data across the geographic space however population densities are extremely unlikely to be uniform [@BCM].
+Choropleths imply uniformity of data across the geographic space, when considering population densities it is extremely unlikely to be uniform [@BCM].
 A cartogram displays abstracted geographic space, with the intention of improving the presentation of the statistic of interest.
-For a single variable of interest, each map area is changed to emphasize the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].
+For a single variable of interest, each map area is changed to emphasise the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].
 Changes in the map base are implemented by altering the boundaries, and therefore shapes, of individual areas.
@@
 Choropleth maps will always be distorted if they cover enough of the globe, just like photographs of the globe from space.
-Choropleth creation requires choosing a map projection that shows a favorable distortion of the geography for presenting the set of spatial information.
+Choropleth creation requires choosing a map projection that shows a favourable distortion of the geography for presenting the set of spatial information.
 Selecting a display can prevent misinterpretation of global statistics, as global maps face the challenge of equitable displays of land mass on maps [@RSCW].
@@
 ```{r choroCRS, fig.cap = "Four choropleth maps of the United States of America using various coordinate reference systems.
-Each state has been colored according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
 The map projections alter the shapes and angles of the boundaries of each state.
@@
 <!-- Common variables used to create cartograms, e.g. population, mortality -->
-Cartograms provide an alternative visualization method for statistical and geographical information.
+Cartograms provide an alternative visualisation method for statistical and geographical information.
 The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC]. 
 @HTLWM suggests that white lies may be employed to create useful displays and map creators have the ability to draw lines that may distort the geometry and suppress features and it is easy for the average person to disregard the impact of transformations used to create cartograms.
-Cartograms may be seen as an extension of map transformations and projections. The favorable distortion is proportional to a value other other than actual earth size area [@NAC].
+Cartograms may be seen as an extension of map transformations and projections. The favourable distortion is proportional to a value other other than actual earth size area [@NAC].
 A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, making interpretation of spatial patterns very difficult.
 The distortion of a cartogram accounts for the population density, preventing it from obscuring the spatial patterns [@TAAM]. 
-The spatial transformation of map regions relative to the data emphasizes the data distribution instead of land size [@CBATCC].
-When visualizing population statistics, @ACTUC considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
+The spatial transformation of map regions relative to the data emphasises the data distribution instead of land size [@CBATCC].
+When visualising population statistics, @ACTUC considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
 @HEDP suggests that 'cancer occurs in people, not in geographical areas' and @CTTMB believe that spatial socio-economic data, like cancer rates, are best presented on a cartogram for urban areas as the population map base avoids allocating 'undue prominence' to rural areas.
@@
 Early approaches including John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball bearing approach and Tobler's (1973) computer programs [@ACTUC].
-Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilized depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994), [@CD].
+Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994), [@CD].
 @HEDP discusses the impact of electronic computer-assisted techniques.
@@
 Table: \label{tab:usa} Maps used to present statistics for the United States of America.
-Each state has been colored according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
 
@@
 |Dorling| The states have been represented by a circle, but the size was determined by the population of the state in 2015.
-The North East states remain closer to their neighbors, and may be displaced from their geographic location.
+The North East states remain closer to their neighbours, and may be displaced from their geographic location.
 The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.|
@@
 The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the light yellow color, the age adjusted rate in Kentucky is the darkest and its neighbors are similar.|
+The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.|
 
 ```{r auscartograms, fig.cap = "Alternative maps of the United States of America.
-Each state has been colored according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
 Each stated shape has been distorted according to the population of the state in 2015.
@@
 <!-- What is a contiguous cartogram -->
-<!-- Intentionally preserve neighbors -->
+<!-- Intentionally preserve neighbours -->
 
@@
 This intentional design goal coincided with the rising interest in urban geography and presentation of social statistics.
-Figure \ref{fig:ggcartograms} a shows a population contiguous cartogram of the United States.
-All states are visible and the shape of the United States overall is still recognizable.
-In contrast, Figure \ref{fig:auscartograms} a shows an Australian contiguous cartogram also based on population.
-The south east is completely distorted, areas with low population are still large very the map, as their initial area was used to determine the area allocated in the contiguous cartogram.
 
@@
 @CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes. 
+@MDAC define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area can be used to communicate the statistic of regions. 
+The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off be made between boundary complexity and simplicity.
+
+<!--
+Tobler's method and the many implementations that 'elaborated' on it are derived from 'numerical approximations to a pair of equations'[@ACTUC].
+They all operate through incremental adjustments, and can produce wildly different outcomes from small changes in the inputs.
+
+Computer generated map examples:  @HEDP
+(Hopps et al. 1968; Armstrong 1972). There
+has followed a flood of disease atlases, mainly concentrating on the modem problems of cancer and degenerative diseases from countries as scattered as the United States (Burbank 1971; Mason et al.
+1975, 1976; Pickle et al.
+1987), the Soviet Union (Levin 1980), Japan (Shigematsu 1977), the Federal Republic of Germany
+-->
 
@@
 Map creators can efficiently communicate with this kind of map by keeping the outlines or particular elements of the original in the new shape [@NISCC].
-The scale of the areas does not impact on the shape recognition.
-However it may impact on the visibility of all areas if small areas expand beyond their boundaries. 
-Figure \ref{fig:ggcartograms} b shows small overlaps in the north east region of the United States.
-These overlaps are also shown in the non-contiguous cartogram of Australia in Figure \ref{fig:auscartograms} b.
+The scale of the areas does not impact on the shape recognition. However it may impact on the visibility of all areas if small areas expand beyond their boundaries.
 
@@
 This implementation can work for up to 'one hundred thousand' areas.
-Figure \ref{fig:ggcartograms} c shows the displacement of the north east region of the United States in the Dorling cartogram, the rest of the country is still recognizable.
-This displacement is shown to an extreme in \ref{fig:auscartograms} c the Dorling cartogram of Australia.
 
@@
 @TFYCC quotes the official definition of Value-Area Cartograms, the simplistic displays which represent each area as a single rectangle, sized according to the value of the statistic.
-This rectangular display also allows for tiling, where geographic neighbors placed in suitable relative positions also share borders, however contiguity may be sacrificed [@CDWCS].
-Rectangular cartograms allow for bi-variate displays, population can be effectively communicated by the size of each rectangular, and a second variable can be communicated using color [@ORC].
+This rectangular display also allows for tiling, where geographic neighbours placed in suitable relative positions also share borders, however contiguity may be sacrificed [@CDWCS].
+Rectangular cartograms allow for bi-variate displays, population can be effectively communicated by the size of each rectangular, and a second variable can be communicated using colour [@ORC].
 
@@
 Each area is represented by a square of the same dimensions.
-Figure \ref{fig:ggcartograms} d shows a hexagon tile map representing each state using a single hexagon.
+Figure \ref{fig:ggcartograms} d shows a hexagon tile map representing each state using a single square.
 
@@
 
-A simple tile map presents the areas in a tessellated grid display, where geographic neighbors are found next to each other, with some necessary displacement employed for regions with more than four neighbors.
-These tiles may be labelled or colored to represent a value.
+A simple tile map presents the areas in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
+These tiles may be labelled or coloured to represent a value.
 Tile maps may be difficult to create, they are best created manually, with additional time and care required as the amount of geographic areas to include increases.
-@MDAC define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area can be used to communicate the statistic of regions. 
-The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off be made between boundary complexity and simplicity.
 
 ```{r ggtilemap, fig.cap = "A tile map and geofaceted map of the Unites States of America.
-Each state in the tile map has been colored according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each state in the tile map has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
 Each state has been represented by a square, each square is the same size.
@@
 ## Geofaceting
-Tile maps can be extended to allow more than just simple coloring of tiles to present data.
-@IGF formalizes the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas. 
-Like tile maps, the arrangement of the areas can be reused for other spatial distribution visualizations.
-Geofaceting have the functionality of facets, often used to replicate visualizations for each subset of the data, in this case the data subsets are geographic.
-The amount of information able to be communicated has increased from one value per region in a tile map to one visualization, this is a more flexible display.
+Tile maps can be extended to allow more than just simple colouring of tiles to present data.
+@IGF formalises the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas. 
+Like tile maps, the arrangement of the areas can be reused for other spatial distribution visulisations.
+Geofaceting have the functionality of facets, often used to replicate visualisations for each subset of the data, in this case the data subsets are geographic.
+The amount of information able to be communicated has increased from one value per region in a tile map to one visualisation, this is a more flexible display.
 Virtually any plot display can be fit into the tile representing the areas, allowing displays of multiple variables or values per geographic entity.
@@
 To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
-Choropleth displays utilize more traditional cartographic methods, they are usually true to the topography, displaying familiar boundaries of countries, states or administrative areas.
+Choropleth displays utilise more traditional cartographic methods, they are usually true to the topography, displaying familiar boundaries of countries, states or administrative areas.
 Highlighting the geographic distribution, and differences between equal spatial units.
@@
 Choropleth methods require a decision to be made about the projection of the display, in this case aim to find a map projection that gives the least distorted representation of the geography [@UMPT].
-However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a color.
+However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a colour.
 These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
-Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbors. The map is reusable and fit for many purposes.
+Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbours. The map is reusable and fit for many purposes.
 
 <!-- Cartograms, sacrifice geography, focus on the outliers/ relationships in data space -->
-Cartographers were historically the creators of geographic displays, and epidemiologists utilized the maps that had been prepared in advance.
+Cartographers were historically the creators of geographic displays, and epidemiologists utilised the maps that had been prepared in advance.
 Cartograms with the population variable used to distortion the geographic map regions and boundaries are density equalising maps, that reduce the visual prominence of large, but low-population areas [@VSSDCUC].
@@
 This can prevent effective comparison of the statistic.
-Cartograms also favor extreme statistics in the positive direction.
+Cartograms also favour extreme statistics in the positive direction.
 Small extreme values are often lost due to the resulting small sized areas the do not draw user's attention.
-The change in the map based on one variable provides the opportunity for bi-variate displays, using the change in area for one variable, and color for another.
+The change in the map based on one variable provides the opportunity for bi-variate displays, using the change in area for one variable, and colour for another.
 The difference between the familiar map display helps highlight the impact of the disease on communities.
@@
 Areas that are geographically close will maintain connectedness in some way, but the population and the statistic will dominate.
-Tile map displays may only allow one variable to be visualized if each region is given consistent map space, this occurs in geofaceting. 
-Dorlings provide bi-variate displays as they use the size of the shape for one variable, and color for another.
+Tile map displays may only allow one variable to be visualised if each region is given consistent map space, this occurs in geofaceting. 
+Dorlings provide bi-variate displays as they use the size of the shape for one variable, and colour for another.
 These alternatives are especially helpful for data aggregations where administrative boundaries break populations into groups.
@@
 | Preserves shapes | Y | N | Y | N | N |
-| Preserves neighbors | Y | Y | Y | S | S |
+| Preserves neighbours | Y | Y | Y | S | S |
 | Uniform use of shape | N | N | N | Y | Y |
 | Bi-variate display^[A bi-variate display refers to a map that allows two variables to be plotted.
-One variable via the size of the area, and another by filling the area with a color.] | N | Y | Y | Y | Y |
+One variable via the size of the area, and another by filling the area with a colour.] | N | Y | Y | Y | Y |
 
@@
 Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
-Tool tips allow users to obtain additional information regarding specific geographic areas.
+Tooltips allow users to obtain additional information regarding specific geographic areas.
 
@@
 Interactions with online atlases can also change the amount of detail shown.
-It is uncommon to see linked displays utilized in online atlases, but there are several frameworks that support interactive displays of multivariate data.
+It is uncommon to see linked displays utilised in online atlases, but there are several frameworks that support interactive displays of multivariate data.
 Linked displays allow for direct manipulation, @RPIG uses Swayne and Klinke's definition of high level interaction that occurs via data selection or brushing tools that highlight subsets of graphical elements such as polygons or points.
@@
 <!-- Animations -->
-Animations communicate a story, showcasing a critical region by defining and progressing through a small set of keyframes, or waypoints [@ESRI].
-@TGA explains that animations show the transitions between frames in a linear, fixed narrative to passive users, communicating a message by capturing and directing their attention. 
+Animations communicate a story, showcasing a critical region by defining and progressing through a small set of keyframes, or waypoints @ESRI.
+The view the transitions between frames in a linear, fixed narrative to passive users, communicating a message by capturing and directing their attention [@TGA]. 
 @TGA suggests the use of moving objects demands attention by taking advantage of the behaviour of the visual cognition system.
@@
 
-```{r animating, fig.cap = "Animations in publicly available choropleth cancer maps.
-[a.
-Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC)(Controls for indicator, sex, age groups, continents, and cancer types) , [b.
-Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality) (Controls for indicator, age, year, and geographical areas)", fig.width = 12, fig.height = 9}
-a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
-a2 <- png::readPNG("cancer_map_images/animation/2.animation.png")
+# Conclusions
 
-ggdraw() +
-  draw_plot(rasterGrob(a1), 0, 0, 0.3, 1) +
-  draw_plot(rasterGrob(a2), 0.3, 0, 0.7, 1) +
-  draw_plot_label(c("I", "II"), 
-    c(0, .3), c(1, 1), size = 20)
-```
+<!-- summary of mapping -->
+This paper provides an overview of mapping practices as used for cancer atlases, and new approaches that could be adopted.
+The conventional approach is the choropleth map, and it is widely used.
+When there are small areas, as occurs in Australia where the population is concentrated on the coast, the information about cancer can be lost, and alternatives are needed.
+Making an inset can clarify congested regions but this breaks the viewers attention, because they need to shift focus from the map to the inset, and if there are many congested areas, many insets would be needed.
+The map alternatives, like cartograms and their variations, can be useful to allow the spatial distribution of cancer data to be digested.
+
+<!-- statistics -->
+There are many different statistics that are commonly used for display.
+The most basic is incidence rate, which is easy to understand.
+It is common to see relative rates in many maps which measure how far a region is above or below the average.
+The purpose of using a relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates.
+However, we lose the incidence rate information and thus interpretability.
+A region might be much higher than average, but it may not be close to a health concern, because all regions have low incidence.
+
+<!-- interaction -->
 
@@
 It was in the development of this online cancer atlas that methods for disease map displays, and visual communication strategies were explored.
-We are thankful for the opportunity to write about these visualizations and the situations in which they are appropriate.
+We are thankful for the opportunity to write about these visualisations and the situations in which they are appropriate.
 
@@
 For data analysis the tidyverse package [@tidyverse] provided many useful functions.
-ggplot2 package [@ggplot2] was used to create maps, with RColorBrewer package [@RColorBrewer] providing additional color palettes and ggthemes [@ggthemes] package providing themes.
+ggplot2 package [@ggplot2] was used to create maps, with RColorBrewer package [@RColorBrewer] providing additional colour palettes and ggthemes [@ggthemes] package providing themes.
 The png package [@png] was used to access png images taken from online web atlases, cowplot package [@cowplot] was used to arrange these plots into grouped images.

### 2019-09-24 review_article.Rmd

--- 2019-09-24 review_article.Rmd
@@
 
+# Conclusions
+
+<!-- summary of mapping -->
+This paper provides an overview of mapping practices as used for cancer atlases, and new approaches that could be adopted.
+The conventional approach is the choropleth map, and it is widely used.
+When there are small areas, as occurs in Australia where the population is concentrated on the coast, the information about cancer can be lost, and alternatives are needed.
+Making an inset can clarify congested regions but this breaks the viewers attention, because they need to shift focus from the map to the inset, and if there are many congested areas, many insets would be needed.
+The map alternatives, like cartograms and their variations, can be useful to allow the spatial distribution of cancer data to be digested.
+
+<!-- statistics -->
+There are many different statistics that are commonly used for display.
+The most basic is incidence rate, which is easy to understand.
+It is common to see relative rates in many maps which measure how far a region is above or below the average.
+The purpose of using a relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates.
+However, we lose the incidence rate information and thus interpretability.
+A region might be much higher than average, but it may not be close to a health concern, because all regions have low incidence.
+
+<!-- interaction -->
+
 # Acknowledgements

### 2019-09-24 review_article.Rmd
re-organised some material and caption for interactivity examples

@@
 This makes it difficult to pair demographic, environmental and other factors with spatial distributions.
-@MMST present linked micromap plots as a solution to the issue of multi-dimensionality to visually link geographic and statistical data.
-This design is widely used in online atlas displays, it groups areas based on their value for one variable, and uses  additional columns to provide displays that contrast the areas in each group by other variables.
-The display juxtaposes of choropleth maps and statistical plots;
-it shows one map per group of the key separating variable, in a row with each additional statistical plot.
-Linked micromaps predominantly utilize the choropleth map for displays of spatial distributions.
-It is one of several alternative displays that allow maps to become bi-variate displays.
 
@@
 
+## Micromaps
+
+@MMST present linked micromap plots as a solution to the issue of multi-dimensionality to visually link geographic and statistical data.
+This design is widely used in online atlas displays, it groups areas based on their value for one variable, and uses  additional columns to provide displays that contrast the areas in each group by other variables.
+The display juxtaposes of choropleth maps and statistical plots;
+it shows one map per group of the key separating variable, in a row with each additional statistical plot.
+Linked micromaps predominantly utilize the choropleth map for displays of spatial distributions.
+It is one of several alternative displays that allow maps to become bivariate displays. (XXX Maybe also refer to Petra Kunert's visumap package since bivariate map displays has been raised here)
+
 # Comparison of maps and cartograms {#ch:compare}
@@
 
+<!-- this is vague and uninformative 
 There are now many suitable methods and mediums to communicate health data.
-The choice between them should always be driven by the intended message to be communicated and the needs of the audience.
+The choice between them should always be driven by the intended message to be communicated and the needs of the audience.-->
 
+Interactivity and animation add the ability to display more information in the same display, without overloading a single display. It is also a way to invite the viewer to engage with the display.
+
+XXX Define interactivity, define animation
+
+XXX summarise literature on interactivity with maps, and also animation.
+
+Thinking about interactivity in terms of atlas purpose, it is useful to identify the primary questions that might be asked (or communicated).
 @CPISACA provide three questions that are generally asked when using a disease map:
@@
 
-```{r animating, fig.cap = "Animations in publicly available choropleth cancer maps.
+```{r animating, fig.cap = "Interactivity and animations in publicly available choropleth cancer maps.
 [a.

### 2019-09-25 review_article.Rmd
Interaction & animation introductory paragraph

--- 2019-09-24 review_article.Rmd
@@
 
+## Micromaps
+
+@MMST present linked micromap plots to visually link geographic and statistical data, this serves as a solution to  multi-dimensionality issues.
+This design is widely used in online atlas displays.
+Areas are grouped based on their value for one variable, and additional columns provide displays that contrast the areas in each group by other variables.
+The display juxtaposes of choropleth maps and statistical plots;
+it shows one map per group of the key separating variable, in a row with each additional statistical plot.
+Linked micromaps predominantly utilize the choropleth map for displays of spatial distributions.
+It is one of several alternative displays that allow maps to become bivariate displays. (XXX Maybe also refer to Petra Kunert's visumap package since bivariate map displays has been raised here)
+
 # Comparison of maps and cartograms {#ch:compare}
@@
 
-There are now many suitable methods and mediums to communicate health data.
-The choice between them should always be driven by the intended message to be communicated and the needs of the audience.
+<!-- XXX Define interactivity, define animation -->
+Interactivity and animation enhance online atlases by enabling additional information to be presented in the same display.
+Implementing movement and allowing for changes in the display invites the viewer to engage with the display, and the information presented.
+Interactive methods involve direct manipulation of web displays, keeping a continuous view of the map as it changes due to the physical actions of users. 
+The manipulation of the web display may occur by directly altering the display, clicking the controls or command-line prompts [@RPIG].
+@TGA explains that animations show the transitions between frames in a linear, fixed narrative to passive users, communicating a message by capturing and directing their attention. 
+This does not require direct user interaction on the display, but may allow users to start, pause or end the narrative displayed.
 
+Web atlases have employed interactivity to expand the communicative ability of an atlas.
+It is useful to identify the primary questions that might be asked by atlases users, and allow the display to change to best answer these questions.
 @CPISACA provide three questions that are generally asked when using a disease map:
@@
 A single map display may not effectively answer all three at the same time.
-This is especially true when areas are small and difficult to identify in a display.
+This is especially true when some areas are small and difficult to identify in a display.
 However, with the ability to interact with a map, or animate the display, map creators can specify and answer these questions by directing the attention of a map user to the areas of interest.
@@
 The progress from static print media displays to interactive methods allows users to develop maps to ask and answer their own questions.
-Interactive methods involve direct manipulation of web displays, keeping a continuous view of the map, which changes due to the physical actions of users. 
-The manipulation of the graph may occur directly, via the graph controls or the command-line [@RPIG].
 The affects of these actions are rapid, incremental, and reversible @DMIV.
@@
 Animations communicate a story, showcasing a critical region by defining and progressing through a small set of keyframes, or waypoints [@ESRI].
-@TGA explains that animations show the transitions between frames in a linear, fixed narrative to passive users, communicating a message by capturing and directing their attention. 
 @TGA suggests the use of moving objects demands attention by taking advantage of the behaviour of the visual cognition system.

### 2019-09-25 review_article.Rmd
add citations to map refrences

@@
 
-Figure 2.a contains an image from *[The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/)*.
+Figure 2.a contains an image from *[The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/)*[@EnvEnglandWales2].
 This map shows the relative risk for women developing lung cancer in England and Wales in 2010.
@@
 
-The map seen in Figure 2.b is from the *[Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx)*.
+The map seen in Figure 2.b is from the *[Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx)* [@Globocan].
 This global map shows age standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018.
@@
 
-Figure 2.c shows an extract from  *[The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/)*.
+Figure 2.c shows an extract from  *[The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/)* [@QLDcancerAtlas].
 This map was published by the Queensland Cancer Council and shows the relative incidence ratio of lung cancer in males in the state of QLD within Australia based on data from 1998 to 2007.
@@
 ### Bowel Cancer Australia Atlas 
-Figure 2.d shows the *[Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/)*.
+Figure 2.d shows the *[Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/)* [@Bowel].
 Published by *[Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)* (Australia).
@@
 ### United States Cancer Statistics: An Interactive Cancer Statistics Website
-The *[United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/)* can be seen in Figure 2.e.
+The *[United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/)* [@USInteractive] can be seen in Figure 2.e.
 This map contains  the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level.
@@
 
-<!-- suggested citations: U.S.
-Cancer Statistics Working Group.
-U.S.
-Cancer Statistics Data Visualizations Tool, based on November 2018 submission data (1999-2016): U.S.
-Department of Health and Human Services, Centers for Disease Control and Prevention and National Cancer Institute; www.cdc.gov/cancer/dataviz, June 2019.-->
-
 ### Map of Cancer Mortality Rates in Spain
-The *[Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html)* can be seen in Figure 2.f.
+The *[Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html)* [@cancerSpain] can be seen in Figure 2.f.
 The side by side maps show relative risk of lung cancer for men vs women based on data from 2004 to 2008.
@@
 ### Atlas of Childhood Cancer in Ontario 
-Figure 2.g shows an extract from the Atlas of Childhood Cancer in Ontario and specifically displays the *[incidence rate of childhood cancers](https://www.pogo.ca/wp-content/uploads/2015/02/POGO_CC-Atlas-3-Incidence_Feb-2015.pdf)* per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004.
-
-<!-- map is on pg 41 or pdf report-->
+Figure 2.g shows an extract from the Atlas of Childhood Cancer in Ontario and specifically displays the *[incidence rate of childhood cancers](https://www.pogo.ca/wp-content/uploads/2015/02/POGO_CC-Atlas-3-Incidence_Feb-2015.pdf)* [@OntarioPediatric] per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004.
 
@@
 Implementing movement and allowing for changes in the display invites the viewer to engage with the display, and the information presented.
+
 Interactive methods involve direct manipulation of web displays, keeping a continuous view of the map as it changes due to the physical actions of users. 
@@
 
-Web atlases have employed interactivity to expand the communicative ability of an atlas.
-It is useful to identify the primary questions that might be asked by atlases users, and allow the display to change to best answer these questions.
+Web atlases have employed interactivity and animation to expand the communicative ability of a disease map.
+It is important to identify the primary questions that might be asked by the users of a map, and allow the display to change to best answer these questions.
 @CPISACA provide three questions that are generally asked when using a disease map:
@@
 A single map display may not effectively answer all three at the same time.
-This is especially true when some areas are small and difficult to identify in a display.
+This is especially true when some areas are small and difficult to identify clearly in a single display or resolution.
 However, with the ability to interact with a map, or animate the display, map creators can specify and answer these questions by directing the attention of a map user to the areas of interest.
@@
 Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions that allow inspection of the data from many views [@DQBCM].
-This opportunity provides an alternative to the traditional approach of a single map for a single use, into a more exploratory territory.
+
+Interactivity and animation provide an alternative to the traditional approach of a single map for a single use, and enable the map to become more exploratory.
 Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
@@
 
-Developments and adoptions of this technology has allowed map users more control of the message displayed.
+Development and adoption of this technology has allowed map users more control of the displayed message.
 Figure \ref{fig:interacting} presents forms of low level interactions with a user interface, these are the control menus for several online cancer atlases.
-The user interface for these online atlases allow combinations of cancer type, gender, and other factors to be shown on the map display.
+The user interface for these online atlases allow combinations of cancer type, gender, and other factors to be shown on the map display and the user to move between and/or compare different views of these variables.
 These controls may also change the geographies shown, such as continents or countries, states or counties.

### 2019-09-25 review_article.Rmd
Types of linked plots, and interacting

--- 2019-09-25 review_article.Rmd
@@
 
-## Micromaps
+## Multivariate displays
 
@@
 it shows one map per group of the key separating variable, in a row with each additional statistical plot.
-Linked micromaps predominantly utilize the choropleth map for displays of spatial distributions.
-It is one of several alternative displays that allow maps to become bivariate displays. (XXX Maybe also refer to Petra Kunert's visumap package since bivariate map displays has been raised here)
+Linked micromaps predominantly utilize the choropleth map for displays of spatial relationships, that are seen when spatial neighbours are alotted to the same group.
+It is one of several alternative displays that allow maps to become bivariate displays, commonly used to present both an estimate and the associated uncertainty.
+
+Bivariate choropleth maps [@VUADBC] blend colour schemes to convey the intersection of categorised levels of an estimate and the associated uncertainty for each spatial area.
+@VUADBC also suggests map pixelation, each region is broken into small pixels, the individual pixels are allocated values that reflect the uncertainty around the area's estimate by creating texture.
+This display can also be animated, with each frame produced by resampling the pixels.
+Areas with uncertain values will flicker more dramatically than areas with more certain values.
+
+Solutions have been provided for bi-variate displays for movement data, containing Origin - Destination matrices and flow map representations [@MMTVCG]. 
+This method makes the maps and matrix connections clear, and easy to link geographic areas with link lines not crossing. Bar charts can also be added to these to show group totals.
 
@@
 This opportunity provides an alternative to the traditional approach of a single map for a single use, into a more exploratory territory.
+Dynamic visual interfaces allow users to interaction with spatial information, they respond when users interrogate them for information [@MCRG].
 Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
 Tool tips allow users to obtain additional information regarding specific geographic areas.
+Interaction allows for the focus of the display to change in spatial scale, or location, through zooming and panning functionality.
 
@@
 Interactions with online atlases can also change the amount of detail shown.
+
 It is uncommon to see linked displays utilized in online atlases, but there are several frameworks that support interactive displays of multivariate data.
 Linked displays allow for direct manipulation, @RPIG uses Swayne and Klinke's definition of high level interaction that occurs via data selection or brushing tools that highlight subsets of graphical elements such as polygons or points.
+Selections can be made statistically through interactions which allow subsets to be selected from statistical displays, such as tables of data.
+Monmonier implemented linking choropleths with statistical plots, adding a geographic component to scatter plots [@MCRG], the use of link allows spatial dimensions to be reserved for geographic locations of areas.
+The statistical variation of additional variables can be presented in other appropriate displays.
 Low level interaction is more common, this occurs via the user interface or the map.

### 2019-09-25 review_article.Rmd

--- 2019-09-25 review_article.Rmd
@@
 Figure 2.g shows an extract from the Atlas of Childhood Cancer in Ontario and specifically displays the *[incidence rate of childhood cancers](https://www.pogo.ca/wp-content/uploads/2015/02/POGO_CC-Atlas-3-Incidence_Feb-2015.pdf)* [@OntarioPediatric] per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004.
-
-## Interactivity
-
-One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
-These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
-Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
-
-Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display.
-Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
-The use of these supports were found in various online cancer maps identified by @roberts2019communication.
-The controls for basic interactive features are often placed outside of the plot space [@TGA], thus the map image is updated/replaced as the user interacts with the controls.
-For example, changing the population age or other demographic variables.
-Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactive capabilities.
-
-```{r interacting, fig.cap = "Interactive controls of displays in publicly available choropleth cancer maps.
-[a.
-Globocan 2018: Cancer Today](http://globocan.iarc.fr/Pages/Map.aspx)(Controls for indicator, sex, age groups, continents, and cancer types) , [b.
-Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/) (Controls for indicator, age, year, and geographical areas), [c.
-The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M) (Controls for predetermined combinations of cancer type or risk factors, mortality or incidence, and gender) [d.
-Global Cancer Map](http://globalcancermap.com/) (Controls for indicator, and cancer type), [e.
-United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html) (Controls for demographics, trends, geography, indicators, risk factors, sex, cancer type, and year).", fig.width = 12, fig.height = 9}
-i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png') # b in choro grid
-i2 <- png::readPNG('cancer_map_images/interactivity/15.interactivity.png') # d in choro grid
-i3 <- png::readPNG('cancer_map_images/interactivity/4.interactivity.png')
-i4 <- png::readPNG('cancer_map_images/interactivity/5.interactivity.png')
-i5 <- png::readPNG('cancer_map_images/interactivity/7.interactivity.png') # e in choro grid
-
-ggdraw() +
-  draw_plot(rasterGrob(i1), 0,  0.13,   .28,  .83) + #a
-  draw_plot(rasterGrob(i2), 0.32, 0.56, .68, .4) + #b
-  draw_plot(rasterGrob(i3), 0.32, 0.41, .68, .13) + #c
-  draw_plot(rasterGrob(i4), 0.32, 0.14, .68, .25) + #d
-  draw_plot(rasterGrob(i5), 0, 0, 1, .13) + #e
-  draw_plot_label(c("i", "ii", "iii", "iv", "v"), 
-    c(0.01, 0.3, 0.3, 0.3, 0.01), 
-    c(0.99, 0.99, 0.57, 0.40, 0.15), size = 20)
-
-ggsave(filename = "figures/supp_grid.png", device = "png", 
-  width = 12, height = 9, dpi = 300)
-```
-
-The interactivity of modern mapping methods enables supplementary information to be incorporated without cluttering the screen.
-Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
-Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).
-
-These interactive features provide an opportunity for users to explore the additional information available.
-Tool tip features and drop down menus allow for user exploration, small screens are not a problem with the flexibility provided by zooming and panning [@HTLWM].
-This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accessible' ways [@TNTEA].This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accessible' ways [@TNTEA].
-The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
-@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
-
-<!--
-
-Technological advances may provide a solution to this through the use of interactivity, animation, and sound as additional channels for representing uncertainty information without interfering with the ability to see any features that are present in the display [@maceachren1992visualizing].
-However, if uncertainty information is perceived as integral and inseparable from the data it is associated with, assumptions of uncertainty creating cognitive overload may not be as valid.
-
--->
 
@@
 @TGA explains that animations show the transitions between frames in a linear, fixed narrative to passive users, communicating a message by capturing and directing their attention. 
-This does not require direct user interaction on the display, but may allow users to start, pause or end the narrative displayed.
+This may require direct user interaction on the display.
 
@@
 
-<!-- Interactivity (Compare movies or tv to video games) -->
-The progress from static print media displays to interactive methods allows users to develop maps to ask and answer their own questions.
+## Interactivity
+
+Interactivity can be implemented within modern mapping methods to incorporate additional information and complexity without overloading the user.
+Adding too much information to a map may create cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
+Complexity and density of visualization methods can overwhelm novice decision makers, however experts are able to use the detail more readily when making-decisions [@cliburn2002design]. 
+Supplementary information can be incorporated into online atlases without cluttering the display.
+Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning.
+Allowing users to explore the map for more information and providing flexibility in the display [@HTLWM].
+The use of these supports were found in various online cancer maps identified by @roberts2019communication.
+Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
+Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).
+For example, changing the population, age or other demographic variables.
+The controls for basic interactive features are often placed outside of the plot space [@TGA], the map image is updated/replaced as the user interacts with the controls. 
+Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactive capabilities.
+
+Relationships between spatial areas and diseases can be explored with sophistication in nontraditional but still 'cognitively accessible' ways [@TNTEA].
+This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
+The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
 The affects of these actions are rapid, incremental, and reversible @DMIV.
 Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions that allow inspection of the data from many views [@DQBCM].
-
 Interactivity and animation provide an alternative to the traditional approach of a single map for a single use, and enable the map to become more exploratory.
@@
 Interactions with online atlases can also change the amount of detail shown.
+
+```{r interacting, fig.cap = "Interactive controls of displays in publicly available choropleth cancer maps.
+[a.
+Globocan 2018: Cancer Today](http://globocan.iarc.fr/Pages/Map.aspx)(Controls for indicator, sex, age groups, continents, and cancer types) , [b.
+Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/) (Controls for indicator, age, year, and geographical areas), [c.
+The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M) (Controls for predetermined combinations of cancer type or risk factors, mortality or incidence, and gender) [d.
+Global Cancer Map](http://globalcancermap.com/) (Controls for indicator, and cancer type), [e.
+United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html) (Controls for demographics, trends, geography, indicators, risk factors, sex, cancer type, and year).", fig.width = 12, fig.height = 9}
+i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png') # b in choro grid
+i2 <- png::readPNG('cancer_map_images/interactivity/15.interactivity.png') # d in choro grid
+i3 <- png::readPNG('cancer_map_images/interactivity/4.interactivity.png')
+i4 <- png::readPNG('cancer_map_images/interactivity/5.interactivity.png')
+i5 <- png::readPNG('cancer_map_images/interactivity/7.interactivity.png') # e in choro grid
+
+ggdraw() +
+  draw_plot(rasterGrob(i1), 0,  0.13,   .28,  .83) + #a
+  draw_plot(rasterGrob(i2), 0.32, 0.56, .68, .4) + #b
+  draw_plot(rasterGrob(i3), 0.32, 0.41, .68, .13) + #c
+  draw_plot(rasterGrob(i4), 0.32, 0.14, .68, .25) + #d
+  draw_plot(rasterGrob(i5), 0, 0, 1, .13) + #e
+  draw_plot_label(c("i", "ii", "iii", "iv", "v"), 
+    c(0.01, 0.3, 0.3, 0.3, 0.01), 
+    c(0.99, 0.99, 0.57, 0.40, 0.15), size = 20)
+
+ggsave(filename = "figures/supp_grid.png", device = "png", 
+  width = 12, height = 9, dpi = 300)
+```
+
 It is uncommon to see linked displays utilized in online atlases, but there are several frameworks that support interactive displays of multivariate data.
+@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
 Linked displays allow for direct manipulation, @RPIG uses Swayne and Klinke's definition of high level interaction that occurs via data selection or brushing tools that highlight subsets of graphical elements such as polygons or points.
+Selections can be made statistically through interactions which allow subsets to be selected from statistical displays, such as tables of data.
+Monmonier implemented linking choropleths with statistical plots, adding a geographic component to scatter plots [@MCRG], the use of link allows spatial dimensions to be reserved for geographic locations of areas.
+The statistical variation of additional variables can be presented in other appropriate displays.
 Low level interaction is more common, this occurs via the user interface or the map.
@@
 Animations ensure all users follow the same interpretive path but creators cannot control the level of engagement.
-
-Animations can be implemented to allow users varying degrees of control. 
-A basic form of interactivity with animations is the play, pause and rewind of films, where the film is intended to be watched in one direction but the position in the animation can be controlled.
-@CPISACA provide weather maps as a thoroughly developed example of animation of spatial displays to communicate information to the general public.
-The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period of time.
-Where the message is most important, static or animated graphics allow control over the display and interpretation.
-User interaction more allow clicks to progress between stages of more complex animations.
-The Australian Cancer Atlas [-@TACA] provides [tours](
-https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
-This implementation of animation gives users tools to plan their own exploration.
 
@@
 ```
+
+Animations can be implemented to allow users varying degrees of control.
+Figure \ref{fig:animating} a gives an example of how time can be included through animating displays.
+This is a basic form of interactivity with animation allowing users to play, pause and rewind films, where the film is intended to be watched in one direction but the position in the animation can be controlled.
+@CPISACA provide weather maps as a thoroughly developed example of animation of spatial displays to communicate information to the general public.
+The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period of time.
+Where the message is most important, static or animated graphics allow control over the display and interpretation.
+User interaction more allow clicks to progress between stages of more complex animations.
+The Australian Cancer Atlas [-@TACA] provides [tours](
+https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
+This implementation of animation gives users tools to plan their own exploration.
 

### 2019-09-26 review_article.Rmd
update references

@@
 
-Figure 2.a contains an image from *[The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/)*[@EnvEnglandWales2].
+Figure \ref{fig:choropleth-grid-create}a contains an image from *[The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/)* [@EnvEnglandWales2].
 This map shows the relative risk for women developing lung cancer in England and Wales in 2010.
@@
 
-The map seen in Figure 2.b is from the *[Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx)* [@Globocan].
+The map seen in Figure \ref{fig:choropleth-grid-create}b is from the *[Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx)* [@Globocan].
 This global map shows age standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018.
-This map is published by the [World Health organization's International Agency for Research on Cancer](https://www.iarc.fr/).
+This map is published by the [World Health Organization's International Agency for Research on Cancer](https://www.iarc.fr/).
 Data was sourced from cancer registries in each country, contributing registries can be seen in the supplementary material on the cancer atlas website.
@@
 
-Figure 2.c shows an extract from  *[The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/)* [@QLDcancerAtlas].
+Figure \ref{fig:choropleth-grid-create}c shows an extract from  *[The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/)* [@QLDcancerAtlas].
 This map was published by the Queensland Cancer Council and shows the relative incidence ratio of lung cancer in males in the state of QLD within Australia based on data from 1998 to 2007.
@@
 ### Bowel Cancer Australia Atlas 
-Figure 2.d shows the *[Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/)* [@Bowel].
+Figure \ref{fig:choropleth-grid-create}d shows the *[Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/)* [@Bowel].
 Published by *[Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)* (Australia).
@@
 ### United States Cancer Statistics: An Interactive Cancer Statistics Website
-The *[United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/)* [@USInteractive] can be seen in Figure 2.e.
+The *[United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/)* [@USInteractive] can be seen in Figure \ref{fig:choropleth-grid-create}e.
 This map contains  the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level.
@@
 ### Map of Cancer Mortality Rates in Spain
-The *[Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html)* [@cancerSpain] can be seen in Figure 2.f.
+The *[Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html)* [@cancerSpain] can be seen in Figure \ref{fig:choropleth-grid-create}f.
 The side by side maps show relative risk of lung cancer for men vs women based on data from 2004 to 2008.
@@
 ### Atlas of Childhood Cancer in Ontario 
-Figure 2.g shows an extract from the Atlas of Childhood Cancer in Ontario and specifically displays the *[incidence rate of childhood cancers](https://www.pogo.ca/wp-content/uploads/2015/02/POGO_CC-Atlas-3-Incidence_Feb-2015.pdf)* [@OntarioPediatric] per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004.
+Figure \ref{fig:choropleth-grid-create}g shows an extract from the Atlas of Childhood Cancer in Ontario and specifically displays the *[incidence rate of childhood cancers](https://www.pogo.ca/wp-content/uploads/2015/02/POGO_CC-Atlas-3-Incidence_Feb-2015.pdf)* [@OntarioPediatric] per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004.
 

### 2019-09-27 review_article.Rmd
Included abstract

@@
 \section*{Abstract}
+Choropleth maps are a traditional form of presenting spatial data.
+They are widely used in cancer atlases, leveraging the familiar geographical regions to communicate cancer statistics around the world.
+The presentation of this data has improved with the developments in the statistics used to measure the impact of cancer on a population.
+Web atlases have utilized developments in interaction and animation techonolgies to empower individual users to explore cancer statistics for their own purposes.
+There has also been progress and contributions to alternative displays that present spatial data with varying levels of focus on the spatial distribution of a statistic.
+This work highlights the impact of the varying sizes of areas created by the administrative boundaries that define communities and affects the way that spatial impact of cancer on Australian communities is understood.
+These displays are not often utilized in cancer atlases, this provides an opportunity to improve the communication of cancer statistics.
+Including alternative visualizations will provide users a new perspective on the spatial distribution of cancer statistics.
 
@@
 Overwhelmingly, cancer maps utilized to communicate to the public and other non-expert audiences are choropleths.
-These atlases provided Incidence, Survival and Mortality rates to the public on areal or isopleth map bases.
+These atlases provided Incidence, Survival and Mortality rates to the public on areal map bases.
 

### 2019-09-27 review_article.Rmd
interactivity section headings

@@
 ggdraw() +
-  draw_plot(rasterGrob(g1), 0, .5, 0.5, 0.45) +
-  draw_plot(rasterGrob(g2), 0.5, 0.5, 0.5, 0.45) +
-  draw_plot(rasterGrob(g3), 0.0, 0.0, 0.5, 0.45) +
-  draw_plot(rasterGrob(g4), 0.5, 0, 0.5, 0.45) +
+  draw_plot(rasterGrob(g1), 0, .5, 0.5, 0.55) +
+  draw_plot(rasterGrob(g2), 0.5, 0.5, 0.5, 0.55) +
+  draw_plot(rasterGrob(g3), 0.0, 0.0, 0.5, 0.55) +
+  draw_plot(rasterGrob(g4), 0.5, 0, 0.5, 0.55) +
   draw_plot(usa_legend, 0, 0, 1, 0.1) +
@@
 ggdraw() +
-  draw_plot(rasterGrob(aus1), 0, .55, 0.5, 0.45) +
-  draw_plot(rasterGrob(aus2), 0.5, 0.55, 0.5, 0.45) +
-  draw_plot(rasterGrob(aus3), 0.0, 0.1, 0.5, 0.45) +
-  draw_plot(rasterGrob(aus4), 0.5, 0.1, 0.5, 0.45) +
+  draw_plot(rasterGrob(aus1), 0, .55, 0.5, 0.55) +
+  draw_plot(rasterGrob(aus2), 0.5, 0.55, 0.5, 0.55) +
+  draw_plot(rasterGrob(aus3), 0.0, 0.1, 0.5, 0.55) +
+  draw_plot(rasterGrob(aus4), 0.5, 0.1, 0.5, 0.55) +
   draw_plot(aus_legend, 0, 0, 1, .1) +
@@
 
-# Interactivity and Animation  {#ch:interacting}
+# User interaction with maps  {#ch:interacting}
 
@@
 
-## Interactivity
+## Interaction and animation in publicly available atlases
 
@@
 ```
+
+## Under-utilised techniques
 

### 2019-09-27 review_article.Rmd
Features table

--- 2019-09-27 review_article.Rmd
@@
 
-|  | Choropleth | Contiguous | Non-contiguous | Dorling | Tile maps |
-|:-----------:|------------|-----------|---------------|---------|-----------|
-| Preserves shapes | Y | N | Y | N | N |
-| Preserves neighbors | Y | Y | Y | S | S |
-| Uniform use of shape | N | N | N | Y | Y |
-| Bi-variate display^[A bi-variate display refers to a map that allows two variables to be plotted.
-One variable via the size of the area, and another by filling the area with a color.] | N | Y | Y | Y | Y |
+|  | Choropleth | Contiguous | Non-contiguous | Dorling | Tile maps | Geofacets |
+|:-----------:|------------|-----------|---------------|---------|-----------|----------|
+| Spatial distortion | N | Y | Y | Y | Y | Y |
+| Preserves neighbors | Y | Y | Y | S | S | S |
+| Conceals small areas | Y | S | N | N | N | N |
+| Uniform shape | N | N | N | Y | Y | Y |
+| Univariate only | Y | Y | Y | S | S | N |
+| Manual construction | N | N | N | N | Y | Y |
 

### 2019-09-27 review_article.Rmd
Interactivity sections

@@
 
-# Interactivity and Animation  {#ch:interacting}
+# User interactions with maps {#ch:interacting}
 
-<!-- XXX Define interactivity, define animation -->
 Interactivity and animation enhance online atlases by enabling additional information to be presented in the same display.
@@
 
-## Interactivity
+## Interaction in publicly available atlases
 
@@
 Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
-Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).
+Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information [@maceachren1992visualizing, @van1994visualization].
 For example, changing the population, age or other demographic variables.
@@
 ```
+
+## Underutilized techniques
 

### 2019-09-27 review_article.Rmd
table

@@
 |:-----------:|------------|-----------|---------------|---------|-----------|----------|
-| Spatial distortion | N | Y | Y | Y | Y | Y |
-| Preserves neighbors | Y | Y | Y | S | S | S |
-| Conceals small areas | Y | S | N | N | N | N |
-| Uniform shape | N | N | N | Y | Y | Y |
-| Univariate only | Y | Y | Y | S | S | N |
-| Manual construction | N | N | N | N | Y | Y |
+|Spatial distortion   |  N  |  Y  |  Y  |  Y  |  Y  |  Y  |
+|Preserves neighbors  |  Y  |  Y  |  Y  |  S  |  S  |  S  |
+|Conceals small areas |  Y  |  S  |  N  |  N  |  N  |  N  |
+|Uniform shape        |  N  |  N  |  N  |  Y  |  Y  |  Y  |
+|Univariate only      |  Y  |  Y  |  Y  |  S  |  S  |  N  |
+|Manual construction  |  N  |  N  |  N  |  N  |  Y  |  Y  |
 

### 2019-09-27 review_article.Rmd

@@
 
-## Interaction in publicly available atlases
+## Interaction and animation in publicly available atlases
 
@@
 
-## Underutilized techniques
+## Under-utilised techniques
 

### 2019-09-27 usa_plots.R
fixed USA choropleths

@@ -139,32 +139,28 @@ ggdorl
 ggchoro1 <- ggplot(st_transform(cancer, 3857)) + 
   geom_sf(aes(fill = AgeAdjustedRate), colour = NA) +
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
-  ggtitle("a. The United States using EPSG: 3857") +
   theme_void()+ guides(fill = FALSE)
 
 ggchoro2 <- ggplot(st_transform(cancer, 2163)) + 
   geom_sf(aes(fill = AgeAdjustedRate), colour = NA) +
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
-  ggtitle("b. The United States using EPSG: 2163") +
   theme_void()+ guides(fill = FALSE)
 
 ggchoro3 <- ggplot(st_transform(cancer, 4326)) + 
   geom_sf(aes(fill = AgeAdjustedRate), colour = NA) +
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
-  ggtitle("c. The United States using EPSG: 4326") +
   theme_void()+ guides(fill = FALSE)
 
 ggchoro4 <- ggplot(st_transform(cancer, 2955)) + 
   geom_sf(aes(fill = AgeAdjustedRate), colour = NA) +
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
-  ggtitle("d. The United States using EPSG: 2955") +
   theme_void() + guides(fill = FALSE)
 
-ggchoroCRS <- gridExtra::grid.arrange(ggchoro1, ggchoro2, ggchoro3, ggchoro4)
+ggchoroCRS <- plot_grid(ggchoro1, ggchoro2, ggchoro3, ggchoro4,
+          labels = "auto")
 
 ggsave(filename = "figures/ggchoroCRS.png", plot = ggchoroCRS,
-  device = "png", dpi = 300, width = 7, height = 6)
-
+  device = "png", dpi = 300, width = 10, height = 7)
 
 ###############################################################################
 
@@ -260,7 +256,7 @@ ggsave(filename = "figures/ggfacet.png", plot = ggfacet,
 
 p <- plot_grid(rasterGrob(png::readPNG("figures/ggtilegram.png")),
           rasterGrob(png::readPNG("figures/ggfacet.png")), 
-          labels = c('A', 'B'), label_size = 20, ncol = 1)
[…]

### 2019-09-27 review_article.Rmd
plot updates for sizing

@@
 library(ggplot2)
+library(cowplot)
 library(png)
@@
 The map projections alter the shapes and angles of the boundaries of each state.
-Maps a and b are similar in their straight edges, unlike maps c and d which curve on the northern United States border.", message=FALSE, warning=FALSE, out.width='100%', fig.align='center', fig.width=12}
-knitr::include_graphics("figures/ggchoroCRS.png")
+Maps a and b are similar in their straight edges, unlike maps c and d which curve on the northern United States border.", message=FALSE, warning=FALSE, out.width='100%', fig.align='center'}
+#knitr::include_graphics("figures/ggchoroCRS.png")
+c1 <- png::readPNG("figures/gg3857.png")
+c2 <- png::readPNG("figures/gg2163.png")
+c3 <- png::readPNG("figures/gg4326.png")
+c4 <- png::readPNG("figures/gg2955.png")
+#knitr::include_graphics("figures/usa_grid.png")
+load("figures/usa_legend.rda")
+
+ggdraw() +
+  draw_plot(rasterGrob(c1), 0, .55, 0.5,    0.45) +
+  draw_plot(rasterGrob(c2), 0.5, 0.55, 0.5, 0.45) +
+  draw_plot(rasterGrob(c3), 0.0, 0.1, 0.5, 0.45) +
+  draw_plot(rasterGrob(c4), 0.5, 0.1, 0.5,   0.45) +
+  draw_plot(usa_legend, 0, 0, 1, 0.1, scale = 2) +
+  draw_plot_label(c("a", "b", "c", "d"), 
+    c(0, .5, 0, 0.5), 
+    c(1, 1, .5, .5), size = 20)
 ```
@@
 
-library(cowplot)
 ggdraw() +
-  draw_plot(rasterGrob(g1), 0, .5, 0.5, 0.55) +
-  draw_plot(rasterGrob(g2), 0.5, 0.5, 0.5, 0.55) +
-  draw_plot(rasterGrob(g3), 0.0, 0.0, 0.5, 0.55) +
-  draw_plot(rasterGrob(g4), 0.5, 0, 0.5, 0.55) +
-  draw_plot(usa_legend, 0, 0, 1, 0.1) +
+  draw_plot(rasterGrob(g1), 0, .55, 0.5,  0.45) +
+  draw_plot(rasterGrob(g2), 0.5, 0.55, 0.5, 0.45) +
+  draw_plot(rasterGrob(g3), 0.0, 0.1, 0.5, 0.45) +
+  draw_plot(rasterGrob(g4), 0.5, 0.1, 0.5, 0.45) +
+  draw_plot(usa_legend, 0, 0, 1, 0.1, scale = 2) +
   draw_plot_label(c("a", "b", "c", "d"), 

### 2019-09-27 sa3_plots.R

@@ -111,13 +111,11 @@ sa3lung %>%
   ggplot(.) +
   geom_density(aes(x = sva)) + geom_vline(aes(xintercept = 7))
 
-# The sa3 of Albury is used as the anchor unit
-
-ncont <- cartogram_ncont(sa3lung, k = 1/5,
+ncont <- cartogram_ncont(sa3lung, k = 1/4,
   weight = "Population") %>% st_as_sf() %>% 
   rename(`Age-standardised rate (per 100,000)` = `Age.standardised.rate..per.100.000.`)
 aus_ggncont <- ggplot(ncont) + 
-  geom_sf(data=aus, fill = NA, colour = "grey", size = 0.001) +
+  geom_sf(data=aus, fill = NA, colour = "grey", size = 0.01) +
   geom_sf(aes(fill = `Age-standardised rate (per 100,000)`), colour = NA) + 
   coord_sf(crs = CRS("+init=epsg:3112"), xlim =
              c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +

### 2019-09-27 usa_plots.R

@@ -26,7 +26,8 @@ library(png)
 library(gridExtra)
 
 # read in data
-cancer <- read_csv("data/USCSlung.csv") %>% rename(NAME = Area)
+cancer <- read_csv("data/USCSlung.csv") %>% 
+  rename(NAME = Area)
 
 
 # Join polygons to data
@@ -44,7 +45,9 @@ title <- ""
 
 ###############################################################################
 # Create a choropleth
-ggchoro <- ggplot(cancer) + 
+ggchoro <- cancer %>% 
+  rename(`Age adjusted rate (per 100,000)` = AgeAdjustedRate) %>% 
+  ggplot() + 
   geom_sf(aes(fill = AgeAdjustedRate), colour = NA) +
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
   ggtitle(title) +
@@ -96,7 +99,7 @@ ggncont <- ggplot(ncont) +
   #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
   theme_void() + guides(fill = FALSE)
 
-save(ncont, "data/ncont.rda")
+save(ncont, file = "data/ncont.rda")
 ggncont
 ggsave(filename = "figures/ggncont.png", device = "png", dpi = 300, width = 7, height = 6)
 
@@ -111,8 +114,8 @@ ggdorl <- ggplot(dorl) +
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
   #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
   theme_void()+ guides(fill = FALSE)
-ggsave(filename = "figures/ggdorl.png", device = "png", dpi = 300, width = 7, height = 6)
 ggdorl
+ggsave(filename = "figures/ggdorl.png", device = "png", dpi = 300, width = 7, height = 6)
 
 
 ###############################################################################
@@ -140,27 +143,49 @@ ggchoro1 <- ggplot(st_transform(cancer, 3857)) +
   geom_sf(aes(fill = AgeAdjustedRate), colour = NA) +
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
   theme_void()+ guides(fill = FALSE)
[…]

### 2019-09-27 review_article.Rmd
vancouver referencing style, will need to fix up text in some spots

@@
 fontsize: 11pt
-biblio-style: vancouver
+csl: vancouver.csl
 yolo: yes

### 2019-09-27 sa3_plots.R
perth on grid of aus maps

@@ -111,7 +111,7 @@ sa3lung %>%
   ggplot(.) +
   geom_density(aes(x = sva)) + geom_vline(aes(xintercept = 7))
 
-ncont <- cartogram_ncont(sa3lung, k = 1/4,
+ncont <- cartogram_ncont(sa3lung, k = 1/2,
   weight = "Population") %>% st_as_sf() %>% 
   rename(`Age-standardised rate (per 100,000)` = `Age.standardised.rate..per.100.000.`)
 aus_ggncont <- ggplot(ncont) + 
@@ -200,6 +200,6 @@ ggsave(filename = "figures/aus_gghexmap.png", plot = aus_gghexmap,
 ###############################################################################
 # Aus grid
 
-aus_grid <- gridExtra::grid.arrange(aus_ggcont, aus_ggncont, aus_ggdorl, aus_gghexmap, nrow = 2)
+aus_grid <- gridExtra::grid.arrange(aus_ggcont, full_ggncont, aus_ggdorl, aus_gghexmap, nrow = 2)
 ggsave(filename = "figures/aus_grid.png", plot = aus_grid,
   device = "png", dpi = 300,  width = 7, height = 6)

### 2019-09-28 review_article.Rmd
update references

@@
 Figure \ref{fig:choroCRS} shows a choropleth of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged over 2012 through 2016.
-The data was extracted from the official federal statistics by @usa_lung on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
+The data was extracted from the official federal statistics by @USInteractive on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
 
@@
 While atlases are often used to describe differences between areas, statistics may be displayed at different levels of aggregation.
-Global health statistics can be aggregated to administrative and arbitrarily defined regions, such as those used by the World Health organization and the United Nations [@IARC_3].
-World atlases can allow for displays of data aggregated into continents, countries, states, provinces and congressional districts [@USCS].
+Global health statistics can be aggregated to administrative and arbitrarily defined regions, such as those used by the World Health Organization and the United Nations [@IARC_3].
+World atlases can allow for displays of data aggregated into continents, countries, states, provinces and congressional districts [@USInteractive].
 

### 2019-09-29 review_article.Rmd
trimmed down interactivity section, need to check missing refs, and need urls to show up in bib

@@
 
-Table: \label{tab:measures} Measures used to report cancer statistics
+Table: \label{tab:measures} Common statistics for reporting cancer information.
 
@@
 
-## Overview of publicly available atlases
+## Overview of publicly available atlases \label{sec:publicatlases}
 
@@
 
-```{r ggtilemap, fig.cap = "A tile map and geofaceted map of the Unites States of America.
-Each state in the tile map has been colored according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each state has been represented by a square, each square is the same size.
-The geofaceted map allows the rates for males and females to be contrasted.
-It has the same layout as the tile map, but shows more information per state.",fig.height = 8, fig.width = 5}
+```{r ggtilemap, fig.cap = "Two more alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
+In the tile map, the layout approximates spatial location, with each state being an equal box filled with colour representing cancer incidence.
+The geofaceted map shows bar charts laid out in a grid approximating the spatial location of the state.
+The age-adjusted rates for males and females is shown.
+This type of display allows multiple variables to be displayed for each geographic area.", out.width="100%"}
 knitr::include_graphics("figures/gggrids.png")
@@
 
-Table: \label{tab:methods} Mapping methods used to display cancer statistics
+Table: \label{tab:methods} Summary of features and constraints on common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
 
-|                     | Choropleth | Contiguous | Non-contiguous | Dorling | Tile maps | Geofacets |
-|:-----------:|------------|-----------|---------------|---------|-----------|----------|
+|                     | Choropleth | Contiguous | Non-contig | Dorling | Tile maps | Geofacets |
+|:-----------|:------------:|:-----------:|:---------------:|:---------:|:-----------:|:----------:|
 |Spatial distortion   |  N  |  Y  |  Y  |  Y  |  Y  |  Y  |
@@
 
-Interactivity and animation enhance online atlases by enabling additional information to be presented in the same display.
-Implementing movement and allowing for changes in the display invites the viewer to engage with the display, and the information presented.
+Interactivity and animation can add value to online atlases, by enabling additional information to be presented on-the-fly with one display.
+It also invites the viewer to engage with the display, and the information.
 
-Interactive methods involve direct manipulation of web displays, keeping a continuous view of the map as it changes due to the physical actions of users. 
-The manipulation of the web display may occur by directly altering the display, clicking the controls or command-line prompts [@RPIG].
-@TGA explains that animations show the transitions between frames in a linear, fixed narrative to passive users, communicating a message by capturing and directing their attention. 
-This may require direct user interaction on the display.
+The word "interactive" is used, and perhaps, confused for many different graphics actions.
+At one extreme, it is synonymous with direct manipulation of the display, such as selecting an area, mousing over, or panning and zooming, using mouse action on the plot.
+At the other extreme, it refers to graphical user interfaces (GUI) such as selecting an item on a menu to change the variable being displayed, or checking a toggle box, or dragging a scrollbar.
+@RPIG provides a good overview of types of interactivity with plots.
+a key difference from an animation is that interactivity typically involves on-the-fly computations, and changes to the plot.
 
-Web atlases have employed interactivity and animation to expand the communicative ability of a disease map.
-It is important to identify the primary questions that might be asked by the users of a map, and allow the display to change to best answer these questions.
-@CPISACA provide three questions that are generally asked when using a disease map:
+The primary interactive tools found in online cancer maps include
 
-- What is the mortality rate in a certain area?
-- Are there geographic trends in the data, or regions of unusually high or low rates?
-- Is the lung cancer mortality pattern similar to the pattern of smoking prevalence shown in a companion map?
+- drop down menus to select features,
+- variable selection, 
+- tool tips, 
+- zooming and panning.
 
-A single map display may not effectively answer all three at the same time.
-This is especially true when some areas are small and difficult to identify clearly in a single display or resolution.
-However, with the ability to interact with a map, or animate the display, map creators can specify and answer these questions by directing the attention of a map user to the areas of interest.
+The latter two would be considered to be direct manipulation on the map display, and the former two are GUI interfaces. The effects of these actions are rapid, incremental, and reversible [@DMIV].
+
+There is a long and extensive literature about interactivity with maps.
+[@mcgranaghan1993cartographic] discuss the way adding interactive elements can alleviate cognitive overload caused when too much information is loaded onto a map.
+[@cliburn2002design] discuss the trade-offs required for expert vs novice audiences.
+The expert might desire more complexity for decision-making but this could overwhelm a novice reader.
+
+@HTLWM says that interactivity can be used to allow users to explore the map for more information and provides flexibility for the display.
+The user can toggle between different variables, map views or even multiple realizations of future scenarios [@goodchild1994introduction].
+This provides additional mechanisms for the users to digest the uncertainty of the available information [@maceachren1992visualizing, @van1994visualization].
+Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions that allow inspection of the data from many views [@DQBCM].
+@TNTEA says user interaction with maps helps them to understand and interpret the spatial distribution of disease, to validate, explain or explore the presented statistics and their relationships to each other.
+
+Animation, in contrast to interactivity, usually involves pre-computing views, and showing these in a sequence.
+@TGA provided an overview of animation in association with the R package `gganimate` [@gganimate], with some focus on maps, explaining transitions between frames in a linear, to provide a fixed narrative to passive users.
+Animations are used to communicate a message by capturing and directing users attention.
+It is most often employed to show changes over time.
+The controls for basic animation are usually placed outside of the plot space [@TGA], the map image is updated/replaced as the animation progresses.
 
@@
 
-Interactivity can be implemented within modern mapping methods to incorporate additional information and complexity without overloading the user.
-Adding too much information to a map may create cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
-Complexity and density of visualization methods can overwhelm novice decision makers, however experts are able to use the detail more readily when making-decisions [@cliburn2002design]. 
-Supplementary information can be incorporated into online atlases without cluttering the display.
-Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning.
-Allowing users to explore the map for more information and providing flexibility in the display [@HTLWM].
-The use of these supports were found in various online cancer maps identified by @roberts2019communication.
-Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
-Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information [@maceachren1992visualizing, @van1994visualization].
-For example, changing the population, age or other demographic variables.
-The controls for basic interactive features are often placed outside of the plot space [@TGA], the map image is updated/replaced as the user interacts with the controls. 
-Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactive capabilities.
+Many publicly available cancer atlases employ interactivity and/or animation.
+Figure \ref{fig:interacting} shows the interactive GUI tools provided with the public atlases discussed in Section \ref{sec:publicatlases}.
+Mostly these are menus, and toggles, for choosing indicators, and subsets of the population.
 
-Relationships between spatial areas and diseases can be explored with sophistication in nontraditional but still 'cognitively accessible' ways [@TNTEA].
-This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
-The affects of these actions are rapid, incremental, and reversible @DMIV.
-Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions that allow inspection of the data from many views [@DQBCM].
-Interactivity and animation provide an alternative to the traditional approach of a single map for a single use, and enable the map to become more exploratory.
-Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
-Tool tips allow users to obtain additional information regarding specific geographic areas.
-
-Development and adoption of this technology has allowed map users more control of the displayed message.
-Figure \ref{fig:interacting} presents forms of low level interactions with a user interface, these are the control menus for several online cancer atlases.
-The user interface for these online atlases allow combinations of cancer type, gender, and other factors to be shown on the map display and the user to move between and/or compare different views of these variables.
-These controls may also change the geographies shown, such as continents or countries, states or counties.
-These displays can show familiar shapes for users to orient themselves on the map, then zoom in to explore densely populated communities.
-Interactions with online atlases can also change the amount of detail shown.
-
-```{r interacting, fig.cap = "Interactive controls of displays in publicly available choropleth cancer maps.
-[a.
-Globocan 2018: Cancer Today](http://globocan.iarc.fr/Pages/Map.aspx)(Controls for indicator, sex, age groups, continents, and cancer types) , [b.
-Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/) (Controls for indicator, age, year, and geographical areas), [c.
-The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M) (Controls for predetermined combinations of cancer type or risk factors, mortality or incidence, and gender) [d.
-Global Cancer Map](http://globalcancermap.com/) (Controls for indicator, and cancer type), [e.
-United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html) (Controls for demographics, trends, geography, indicators, risk factors, sex, cancer type, and year).", fig.width = 12, fig.height = 9}
+```{r interacting, fig.cap = "Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
 i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png') # b in choro grid
@@
 
-It is uncommon to see linked displays utilized in online atlases, but there are several frameworks that support interactive displays of multivariate data.
+Figure \ref{fig:animating} shows two examples of more sophisticated interactive maps.
+The Spanish Cancer map (left) contains a linked display between a choropleth map, and time series plots of cancer change.
+In linked plots, changing values in one display will trigger changes of corresponding elements in another display.
+Here, the temporal change in the choropleth map can be played out as an animation.
+Mousing over the time series plots will highlight the line for a particular region.
+The Canadian Breast Cancer Mortality map (right) has a magnifying glass which allows the user to zoom into small areas.
+It is easy to control and shows precise details in the small areas.
+
+<!-- It is uncommon to see linked displays utilized in online atlases, but there are several frameworks that support interactive displays of multivariate data.
 @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
@@
 Interacting with the map returns values or information used in the graphical display without referring to, of changing, the original data source of the map.
+-->
 
 <!-- Animations -->
-Animations communicate a story, showcasing a critical region by defining and progressing through a small set of keyframes, or waypoints [@ESRI].
+<!--Animations communicate a story, showcasing a critical region by defining and progressing through a small set of keyframes, or waypoints [@ESRI].
 @TGA suggests the use of moving objects demands attention by taking advantage of the behaviour of the visual cognition system.
@@
 Animations ensure all users follow the same interpretive path but creators cannot control the level of engagement.
+-->
 
-```{r animating, fig.cap = "Interactivity and animations in publicly available choropleth cancer maps.
-[a.
-Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC)(Controls for indicator, sex, age groups, continents, and cancer types) , [b.
-Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality) (Controls for indicator, age, year, and geographical areas)", fig.width = 12, fig.height = 9}
+```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a.
+Linked maps and time series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b.
+A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 9}
 a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
@@
 
-Animations can be implemented to allow users varying degrees of control.
+<!-- Animations can be implemented to allow users varying degrees of control.
 Figure \ref{fig:animating} a gives an example of how time can be included through animating displays.
@@
 This implementation of animation gives users tools to plan their own exploration.
+-->
 
@@
 -->
+
+<!-- presents forms of low level interactions with a user interface, these are the control menus for several online cancer atlases.
+The user interface for these online atlases allow combinations of cancer type, gender, and other factors to be shown on the map display and the user to move between and/or compare different views of these variables.
+These controls may also change the geographies shown, such as continents or countries, states or counties.
+These displays can show familiar shapes for users to orient themselves on the map, then zoom in to explore densely populated communities.
+Interactions with online atlases can also change the amount of detail shown.-->
+
+<!-- Interactivity can be implemented within modern mapping methods to incorporate additional information and complexity without overloading the user.
+Adding too much information to a map may create cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
+Complexity and density of visualization methods can overwhelm novice decision makers, however experts are able to use the detail more readily when making-decisions [@cliburn2002design]. 
+Supplementary information can be incorporated into online atlases without cluttering the display.
+Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning.
+Allowing users to explore the map for more information and providing flexibility in the display [@HTLWM].
+The use of these supports were found in various online cancer maps identified by @roberts2019communication.
+Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
+Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information [@maceachren1992visualizing, @van1994visualization].
+For example, changing the population, age or other demographic variables.
+The controls for basic interactive features are often placed outside of the plot space [@TGA], the map image is updated/replaced as the user interacts with the controls. 
+Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactive capabilities.
+
+Relationships between spatial areas and diseases can be explored with sophistication in nontraditional but still 'cognitively accessible' ways [@TNTEA].
+This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
+
+The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+
+Interactivity and animation provide an alternative to the traditional approach of a single map for a single use, and enable the map to become more exploratory.
+Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
+Tool tips allow users to obtain additional information regarding specific geographic areas.
+
+Development and adoption of this technology has allowed map users more control of the displayed message. 
+-->
+
+<!--to expand the communicative ability of a disease map.
+It is important to identify the primary questions that might be asked by the users of a map, and allow the display to change to best answer these questions.
+@CPISACA provide three questions that are generally asked when using a disease map:
+
+- What is the mortality rate in a certain area?
+- Are there geographic trends in the data, or regions of unusually high or low rates?
+- Is the lung cancer mortality pattern similar to the pattern of smoking prevalence shown in a companion map?
+
+A single map display may not effectively answer all three at the same time.
+This is especially true when some areas are small and difficult to identify clearly in a single display or resolution.
+However, with the ability to interact with a map, or animate the display, map creators can specify and answer these questions by directing the attention of a map user to the areas of interest.
+-->

### 2019-09-29 cover letter/cover letter.Rmd

---
from: Stephanie Kobakian
fromqualifications: B.Comm. and B.Eco.
fromtitle: Research Masters Student
fromWWW: srk.netlify.com
fromemail: Stephanie.Kobakian@monash.edu
fromphone: +61 433 699 797
fromsignature: sigfile.png
toname: Editors-in-Chief
opening: Dear Editors-in-Chief Dr Freddie Bray and Dr Wanqing Chen,
closing: Warm regards
spacing: 1.4
output: MonashEBSTemplates::letter
---

Please consider our review paper titled "Cartogram Mapping and its Application to Cancer Data Visualization", for inclusion in the Focused Issue on Spatial Patterns in Cancer Epidemiology in the Annals of Cancer Epidemiology.

We were honoured to receive the request from Drs. Baade and Cramb, and Editor Kathryn Yan, to explore the use of mapping methods to present cancer statistics.
This review paper presents an exploration of cancer mapping methods, considering the use of choropleths in many current online cancer atlases.
It highlights the current uses of interactivity and how this allows users to engage with cancer statistics in an online atlas.

This manuscript encourages those who produce atlases to consider cartograms and other alternative displays that effectively communicate cancer outcomes.
We believe that this manuscript is a good fit for the Annals of Cancer Epidemiology, as it encourages careful consideration of the mapping methods used to present spatial information, we expect that this exploration will be of interest to readers of the journal. We hope this will inspire the creation and improvement of cancer atlases worldwide.

We thank you for your consideration.

### 2019-09-30 review_article.Rmd
hopefully fixed references

@@
 
+@CPISACA provide weather maps as a thoroughly developed example of animation of spatial displays to communicate information to the general public.
+The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period of time.
+
+The Australian Cancer Atlas [-@TACA] provides [tours](
+https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
+This implementation of animation gives users tools to plan their own exploration.
+
 ## Interaction and animation in publicly available atlases
@@
 <!-- interaction -->
+Interaction with maps is an important component to public atlases.
+A purpose is to provide access to more information than is possible to display in a single map, without overwhelming the viewer.
+Too many choices though can similarly overwhelm a viewer, and thus decisions do need to be made about content to provide, or hide, for accurate and comprehensive communication of information.
+Similarly, providing ways for users to interact with the display encourages engagement, and creative, efficient, elegant, interactive tools elicits curiosity about the data.
 
@@
 
-We would like to acknowledge Dr Earl Duncan (Research Associate at ARC Centre of Excellence for Mathematical & Statistical Frontiers, QUT).
-His time and effort given to edit and comment on this paper was invaluable.
+The authors would like thank Dr Earl Duncan, Professor Kerrie Mengersen, Dr Susanna Cramb and Dr Peter Baade for conversations on the content of this article.
 
-We thank Dr Susanna Cramb (Spatial Modeller, Cancer Council Queensland) and Dr Peter Baade (Senior Research Fellow, Cancer Council Queensland) for providing the opportunity to cover this exploration of disease mapping methods in writing.
-This literature review would not be possible without the opportunity provided by Queensland University of Technology, and Cancer Council Queensland, and the roles of Professor Kerrie Mengersen (Professor of Statistics, Science and Engineering Faculty, QUT) and Dr Earl Duncan in contributing to the Australian Cancer Atlas and supervision of Stephanie's Research Masters.
+The following R [@R] packages were used to produce this paper: tidyverse [@tidyverse], RColorBrewer [@RColorBrewer], ggthemes [@ggthemes], png [@png], cowplot [@cowplot], sf [@sf], spData [@spData], cartogram [@cartogram], sugarbag [@sugarbag], knitr [@knitr], rmarkdown [@rmarkdown].
 
-It was in the development of this online cancer atlas that methods for disease map displays, and visual communication strategies were explored.
-We are thankful for the opportunity to write about these visualizations and the situations in which they are appropriate.
-
-Several R [@R] packages were used to produce this paper.
-For data analysis the tidyverse package [@tidyverse] provided many useful functions.
-ggplot2 package [@ggplot2] was used to create maps, with RColorBrewer package [@RColorBrewer] providing additional color palettes and ggthemes [@ggthemes] package providing themes.
-The png package [@png] was used to access png images taken from online web atlases, cowplot package [@cowplot] was used to arrange these plots into grouped images.
-
-The following packages were used to create transformations from the sf [@sf] geographical shapes of the states of America [@spData] and the Australian Statistical Areas (Level 3):
-The cartogram [@cartogram] package creates contiguous, non-contiguous and Dorling cartograms.
-Hexagon tile map displays were created using the sugarbag package [@sugarbag].
+Files to reproduce the paper, and code to reproduce the plots, are available at https://github.com/srkobakian/review.
 

### 2019-09-30 review_article.Rmd
abstract

@@
 \section*{Abstract}
-Choropleth maps are a traditional form of presenting spatial data.
-They are widely used in cancer atlases, leveraging the familiar geographical regions to communicate cancer statistics around the world.
-The presentation of this data has improved with the developments in the statistics used to measure the impact of cancer on a population.
-Web atlases have utilized developments in interaction and animation techonolgies to empower individual users to explore cancer statistics for their own purposes.
-There has also been progress and contributions to alternative displays that present spatial data with varying levels of focus on the spatial distribution of a statistic.
-This work highlights the impact of the varying sizes of areas created by the administrative boundaries that define communities and affects the way that spatial impact of cancer on Australian communities is understood.
-These displays are not often utilized in cancer atlases, this provides an opportunity to improve the communication of cancer statistics.
-Including alternative visualizations will provide users a new perspective on the spatial distribution of cancer statistics.
+Choropleth maps are widely used in cancer atlases, and traditionally communicate cancer statistics over geographic domains.
+The type of statistic displayed varies immensely, and effective choices can assist in the understanding of impact of cancer on a population.
+Atlases have moved from print media to web format, and the availability of graphical user interfaces, and user interaction and animation, greatly enhance the ability to communicate more information and positively engage readers.
+More use of alternatives to choropleth maps, such as cartograms, could be employed in online atlases, to better representation of spatial distributions, and show statistics on small areas.
+This is especially true for the geography of Australia, where the population is concentrated is small areas along the coastlines.
 

### 2019-09-30 review_article.Rmd
typo in bib
typo in abstract

@@
 Atlases have moved from print media to web format, and the availability of graphical user interfaces, and user interaction and animation, greatly enhance the ability to communicate more information and positively engage readers.
-More use of alternatives to choropleth maps, such as cartograms, could be employed in online atlases, to better representation of spatial distributions, and show statistics on small areas.
+More use of alternatives to choropleth maps, such as cartograms, could be employed in online atlases, for better representation of spatial distributions, and show statistics on small areas.
 This is especially true for the geography of Australia, where the population is concentrated is small areas along the coastlines.
@@
 
-# Alternative to choropleths {#ch:alternatives}
-
-## Micromaps
-
-Choropleths are uni-variate displays, they do not inherently support multi-dimensionality.
-This makes it difficult to pair demographic, environmental and other factors with spatial distributions.
-
-<!--
-Visualizing Data Patterns with Micromaps.
-Two new templates for epidemiology applications: linkedmicromap plots and conditioned choropleth maps-->
+# Alternatives to choropleths {#ch:alternatives}
 
@@
 
-# User interactions with maps {#ch:interacting}
+# User interaction {#ch:interacting}
 

### 2019-09-30 review_article.Rmd
cartograms done

@@
 \section*{Abstract}
+<!-- 119 words -->
 Choropleth maps are widely used in cancer atlases, and traditionally communicate cancer statistics over geographic domains.
@@
 # Introduction
+<!-- 436 words -->
 <!-- What data are we working with, why, what are the purposes -->
@@
 The viewer would then be able to examine the spatial distribution of the disease incidence, where there is a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease.
-Visualizing diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses [@MTMSIH].
+Visualizing diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses.
 Disease maps help to present geographic patterns that may be overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
@@
 Section \ref{ch:public} surveys atlases in use today.
-Section \ref{ch:alternatives} describes an alternative display, the cartogram what may be useful for countries that have heterogeneously sized geographic units.
+Section \ref{ch:alternatives} describes  alternative displays, including the cartogram which is useful when the map has heterogeneously sized geographic units.
 The pros and cons of these approaches is discussed in Section \ref{ch:compare}.
 Disease maps are more useful when made interactive, and common options are described in Section \ref{ch:interacting}, along with a discussion of benefits and disadvantages.
-The last section summarizes the paper and discusses future directions.
+The last section summarizes the paper.
 
 # Map displays for disease data {#ch:choropleths}
-
+<!-- 381 words -->
 <!-- what is a choropleth map -->
-A choropleth map is used to display differences in the geographical distribution of data by spatial unit by shading areas of a map.
+A choropleth map is used to display differences in the geographical distribution of data by spatial unit by shading areas of a map [@EI].
 The geography is faithfully rendered, and the color rendering is designed to reveal spatial patterns among data values.
-A choropleth is constructed by drawing the geographic or political boundaries, and filling the shapes with colors to represent values of a measured variable [@EI].
-Figure \ref{fig:choroCRS} shows a choropleth of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged over 2012 through 2016.
-The data was extracted from the official federal statistics by @USInteractive on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
-
-<!-- history and types of choropleths, variations -->
+Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organizing communities into representable units [@SAMGIS].
 Early versions of choropleth maps used symbols or patterns instead of color.
 @CPISACA discuss the use of choropleths to visualize cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
+In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality.
 
-<!-- psychology -->
-Utilizing the state boundaries can make a map familiar to read [@CIBMUK], and allows viewers to visually infer the spatial relationships in the data. 
-The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
-Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
-
-<!-- how do we use choropleth's -->
-Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organizing communities into representable units [@SAMGIS].
-In Figure \ref{fig:choroCRS}, a west to east spatial trend of increasing rates, can be seen.
+Figure \ref{fig:choroCRS} shows various geographic projections of choropleth maps of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged over 2012 through 2016.
+The data was extracted from the official federal statistics by @USInteractive on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
+A west to east spatial trend of increasing rates, can be seen.
 There is also a spatial outlier -- Utah has a noticeably lower rate that its neighbors.
@@
 
+<!-- psychology -->
+Utilizing the state boundaries can make a map familiar to read [@CIBMUK], and allows viewers to visually infer the spatial relationships in the data. 
+The familiarity of the geography is a worthy consideration when presenting results of spatial analysis, to consumers that extend beyond researchers, to the public, policymakers and the media [@CPISACA].
+
 <!-- potential problems -->
@@
 The different population and geographical sizes of administrative areas can attract attention to the shades of the underpopulated but large areas [@EI], Skowronnek [-@BCM] calls this an area-size bias.
-Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention away from the 'potentially more important results in the more populous communities' that are geographically smaller [@SE].
+Choropleths can inhibit visual inference when presenting human-related statistics as the display may draw attention away from the 'potentially more important results in the more populous communities' that are geographically smaller [@SE].
 
@@
 # Cancer atlases {#ch:public}
-
+<!-- 360 words -->
 <!-- Displaying of cancer data on choropleths -->
-Choropleth maps can be useful devices for communicating information to public on a familiar map base.
-A cancer atlas is a choropleth map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries. 
-In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality. 
-The data collection methods of cancer mortality rates across regions, and the administrative control within regions lends itself to choropleth visualization. 
-@MACM provides the definition of a cancer atlas, beginning with Haviland's maps in 1875, they attribute UK cancer atlases to Howe (1963), and early work in US cancer atlases can be attributed to Burbank (1971).
-The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
-The choropleth maps presented levels via hatching or dots on a black and white scale.
-These atlases were key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets [@MACM].
+A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
+Atlases are key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets [@MACM].
+The data collection of statistics across regions, and the administrative control within regions, lends itself to choropleth visualization.
+Atlases date to Haviland's maps in 1875, with UK cancer atlases appearing in 1961 [@HEDP], and early work in US cancer atlases appearing in 1971 [@burbank].
+The presentation of cancer statistics has increased with greater access to computational power, and the availability of geographic information systems software [@SE].
 
 <!-- Measures reported in cancer atlases -->
-The presentation of cancer statistics has changed over time with greater access to computational power.
 Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the higher prevalence of cancers in older populations.
-Howe [-@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s.
-Table \ref{tab:measures} summarizes the measures presented in published cancer atlases, and provides a definition of each measure.
+Howe [-@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s. Table \ref{tab:measures} summarizes the measures commonly presented in published cancer atlases.
 
-Table: \label{tab:measures} Common statistics for reporting cancer information.
+Table: \label{tab:measures} Commonly used measures for reporting cancer information.
 
@@
 |2. Rate per 100,000|Cancer incidence per 100,000 population|  
-|3. New cancer cases per 100,000|Specific methods could not be found|   
-|4. IR (Incidence Ratio)|$(IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}$, |
+|3. IR (Incidence Ratio)|$(IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}$, |
 ||Cancer incidence rate in region $i$ over the average cancer incidence rate for the total region|
-|5. Age Adjusted Rate per 100,000| Standardized by age structure or region|   
-|6. Age Adjusted Relative Risk| RR standardized by age structure in each region $i$|  
-|7. SIR (Standardized Incidence Ratio)| IR standardized by age structure in each region $i$|
-|8. Below or above Expected|Alternative expression of the SIR|
-|9. RER |$RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}$|
-|(Relative Excess Risk)|Represents the estimate of cancer related mortality within five years of diagnosis|
-||Also referred to as 'excess hazard ratio'|
+|4. Age Adjusted Rate per 100,000| Standardized by age structure or region|   
+|5. Age Adjusted Relative Risk| Standardized by age structure in each region $i$|  
+|6. SIR (Standardized Incidence Ratio)| Standardized by incidence in each region $i$|
+|7. Below or above Expected|Alternative expression of the SIR|
+|8. RER |$RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}$|
+|(Relative Excess Risk)|Represents the estimate of cancer related mortality within five years of diagnosis. Also referred to as 'excess hazard ratio'|
 
@@
 
-Cancer maps are effective tools for communicating to wide range of audiences, including the general public and others not trained in statistical analyses.
+Cancer maps are effective tools for communicating incidence, survival and mortality to wide range of audiences, including the general public and others not trained in statistical analyses.
 These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analyses.
 Cruickshank (1947) as cited by @DMAHP, discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
-Overwhelmingly, cancer maps utilized to communicate to the public and other non-expert audiences are choropleths.
-These atlases provided Incidence, Survival and Mortality rates to the public on areal map bases.
+Overwhelmingly, the cancer maps utilized to communicate to the public and other non-expert audiences are choropleths.
 
-The Atlas of Cancer in Queensland [@DACQ] focused on highlighting the difference in experience for those living in rural and disadvantaged areas, the Standardized Incidence Ratios (SIR) were modeled via Bayesian methods.
+<!-- written below The Atlas of Cancer in Queensland [@DACQ] focused on highlighting the difference in experience for those living in rural and disadvantaged areas, the Standardized Incidence Ratios (SIR) were modeled via Bayesian methods.
 The presentation of these rates considers not only the shapes of the areas, but also appropriate choices of color blind friendly color schemes, and categories of the values to highlight significantly different areas.
+-->
 
 ## Overview of publicly available atlases \label{sec:publicatlases}
-
-@roberts2019communication identified 33 publicly available cancer atlases, published between January 2010 and November 2015.
+<!-- 416 words -->
+Roberts [@roberts2019communication] identified 33 publicly available cancer atlases, published between January 2010 and November 2015, from different countries around the globe.
 All of these use choropleth maps.
 All but one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government funded partnerships.
-The use of choropleths within the public domain mirrors the heavy use of choropleth maps within the research literature, discussed above.
-The cancer atlases identified by @roberts2019communication covered geographies from all around the world, most focused on single nations.
-Figure \ref{fig:choropleth-grid-create} displays a global (b), national (a, d, e, f) and state (c, g) choropleths.
-The sections below provide details on each of the maps within Figure \ref{fig:choropleth-grid-create}.
+Figure \ref{fig:choropleth-grid-create} displays a subset of maps from these atlases.
 
@@
 
-Figure \ref{fig:choropleth-grid-create}a contains an image from *[The Environmental and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/)* [@EnvEnglandWales2].
-This map shows the relative risk for women developing lung cancer in England and Wales in 2010.
-The cancer data used to generate this map came from Office for National Statistics (ONS) (England) and from the Welsh Cancer Intelligence and Surveillance Unit (WCISU).
+Figure \ref{fig:choropleth-grid-create}a shows the relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2].
+The cancer data was from Office for National Statistics (ONS) (England) and from the Welsh Cancer Intelligence and Surveillance Unit (WCISU).
 
@@
 
-The map seen in Figure \ref{fig:choropleth-grid-create}b is from the *[Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx)* [@Globocan].
-This global map shows age standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018.
-This map is published by the [World Health Organization's International Agency for Research on Cancer](https://www.iarc.fr/).
-Data was sourced from cancer registries in each country, contributing registries can be seen in the supplementary material on the cancer atlas website.
+The map seen in Figure \ref{fig:choropleth-grid-create}b shows age standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018 [@Globocan].
+It is published by the [World Health Organization's International Agency for Research on Cancer](https://www.iarc.fr/).
+Data was sourced from cancer registries of each country.
 
@@
 
-Figure \ref{fig:choropleth-grid-create}c shows an extract from  *[The Atlas of Cancer in QLD](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/)* [@QLDcancerAtlas].
-This map was published by the Queensland Cancer Council and shows the relative incidence ratio of lung cancer in males in the state of QLD within Australia based on data from 1998 to 2007.
-Data to generate this map was sourced from the *[Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/)*.
+Figure \ref{fig:choropleth-grid-create}c shows the relative incidence ratio of lung cancer in males in the state of QLD within Australia based on data from 1998 to 2007 [@QLDcancerAtlas].
+It is published by the Queensland Cancer Council, using data from the *[Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/)*.
 
 ### Bowel Cancer Australia Atlas 
-Figure \ref{fig:choropleth-grid-create}d shows the *[Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/)* [@Bowel].
-Published by *[Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)* (Australia).
-This map shows the percentage of Australian males between 50 - 54 years of age that diagnosed with bowel cancer in 2016 in Australia.
-The source of the data is not provided.
+Figure \ref{fig:choropleth-grid-create}d shows the percentage of Australian males between 50 - 54 years of age diagnosed with bowel cancer in 2016 in Australia [@Bowel].
+It is published by *[Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)*.
 
 ### United States Cancer Statistics: An Interactive Cancer Statistics Website
+
 The *[United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/)* [@USInteractive] can be seen in Figure \ref{fig:choropleth-grid-create}e.
 This map contains  the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level.
-The map was published by the *[Centers for Disease Control and Prevention](https://www.cdc.gov/)*.
-Incidence data seen in this map were compiled from cancer registries meeting U.S.
-Cancer Statistics data quality criteria covering 100% of the U.S.
-population.
+The map was published by the *[Centers for Disease Control and Prevention](https://www.cdc.gov/)*, with data from state cancer registries.
 
 ### Map of Cancer Mortality Rates in Spain
-The *[Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html)* [@cancerSpain] can be seen in Figure \ref{fig:choropleth-grid-create}f.
-The side by side maps show relative risk of lung cancer for men vs women based on data from 2004 to 2008.
-The source of the data and statistical methods are unknown.
+
+Figure \ref{fig:choropleth-grid-create}f shows side by side maps of relative risk of lung cancer for men vs women for 2004 to 2008 [@cancerSpain].
 
 ### Atlas of Childhood Cancer in Ontario
-Figure \ref{fig:choropleth-grid-create}g shows an extract from the Atlas of Childhood Cancer in Ontario and specifically displays the *[incidence rate of childhood cancers](https://www.pogo.ca/wp-content/uploads/2015/02/POGO_CC-Atlas-3-Incidence_Feb-2015.pdf)* [@OntarioPediatric] per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004.
 
-## Common statistics displayed
+Figure \ref{fig:choropleth-grid-create}g  displays the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
 
-Cancer maps are powerful visualizations that summarize complex statistical analyses, however the statistics represented in these maps cannot tell the entire story.
-Supplementary graphs and plots are often included to add more depth and information to the map.
-@CPISACA suggests additional materials such as tables, graphs, and text explanations support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers.
-There are many visualizations used for displays of statistical summaries, these may be dot plots, bar plots, box plots, times series plots, cumulative distribution plots, scatter plots, Q-Q plots. 
-These additional displays of the cancer distribution can provide alternative views of the cancer statistics, as well as the supporting statistics including error, confidence intervals, distributions, sample or population sizes, standard deviation and other measures.
-When presenting cancer maps, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses', the supplementary statistics help for this validation.
+## Additional considerations
+<!-- 479 words -->
+Maps cannot tell the entire story.
+Supplementary graphs and plots are often included to add more depth and information.
+Additional materials such as tables, graphs, and text explanations, support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers [@CPISACA].
+The many displays of statistical summaries, including dot plots, bar plots, box plots, cumulative distribution plots, scatter plots, normal probability plots, can provide alternative views of the cancer statistics.
+These can also display supporting statistics such as error, confidence intervals, distributions, sample or population sizes, standard deviation.
+When presenting cancer maps, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses', and the display of the supplementary statistics helps in this regard.
 
@@
 The distribution of the population residing in all areas may also be communicated in a table or histogram display [@NICR_1].
-Atlases can connect the population to the land available to them by communicating population density.
-Atlases can also connect the population to the land available to them by communicating population density.
+Atlases can connect the population to the land available to them by communicating population density. Alternatives to choropleth to address this are described in Section \ref{ch:alternatives} .
 
 ### Statistical uncertainty
-Additional statistics that accompany an atlas often include a measure of the statistical uncertainty surrounding the statistics presented in a choropleth.
-In the review of atlases in the public domain, atlases were considered to report uncertainty to the non-expert user if they included a measure of statistical uncertainty either within or alongside the map [@roberts2019communication].
-The maps considered used standard and well-known measures including credible intervals and standard deviation, statistical significance, box plots and distributions.
-Other methods involve providing adjacent maps or overlapping maps with symbols [@VSSDCUC].
-The maps employing uncertainty ranged from static documents or infographics, to interactive online visualizations.
-Communicating the statistical uncertainty associated with the estimates occurs using confidence intervals (CI), credible intervals (CrI), statistical significance levels, box plots, distribution plots, and reporting sample size and standard deviations.
-Close to half of the atlases identified (42%, n=14) included some measure of uncertainty. The most common measure used to represent uncertainty were credible or confidence intervals (CIs).
+
+With any number comes some uncertainty about its value.
+This may occur for a number of reasons: numbers are drawn from samples, errors occur in the disease reporting, values are simulated from a model to maintain privacy.
+Close to half of the atlases identified (42%, n=14) included some measure of uncertainty.
+Uncertainty is often recorded as sample or population size, standard deviation, confidence intervals, statistical significance, quantiles, credible intervals.
+The most common measure used to represent uncertainty were credible or confidence intervals (CIs).
+
+Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it can be difficult to display effectively.
+The map focuses on displaying the statistic, and lacks additional space to represent the uncertainty.
+Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two approaches.
 
 ### Demographics
+
 Demographics include information regarding the age and sex distribution of the areas presented.
-<!-- sex -->
-Sex is an important cofactor for cancer atlases. 
-As some cancers are sex specific, and others may be found in both males and females, atlases often specify the relevant sex as part of the visual output in the displays. 
-Digital atlases allow for users to interact with the controls of the displays, they can select males, females or both depending on the type of cancers explored.
+Digital atlases allow for users to control the display, to select subsets such as males, females or those aged over 65.
 
 ### Socio-economic indicators
-Socio-economic indicators can explain how the experience of cancer prevalence varies for various members of a society.
-These indicators include unemployment rates, poverty rates, remoteness, and education levels achieved though, only a few atlases also explored the impact of rurality on cancer rates. 
-These rates may also be explored as percentages above or below the mean or median value for the set of spatial areas.
-The Human Development Index can be used to understand the socio-economic experience of a community, as well as Income levels which can be sourced from the World Bank list of economies [@IARC_3]. 
-The areas are often ranked and allocated to quintiles; each quintile can be presented as categories describing the ranking.
 
-One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
-These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
-Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
-
-Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display.
-Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
-The use of these supports were found in various online cancer maps identified by @roberts2019communication.
-The controls for basic interactive features are often placed outside of the plot space [@TGA], thus the map image is updated/replaced as the user interacts with the controls.
-For example, changing the population age or other demographic variables.
-Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection tools.
-
-Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
-Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).
-
-These interactive features provide an opportunity for users to explore the additional information available.
-This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accessible' ways [@TNTEA].
-The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
-@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
+Socio-economic indicators, such asunemployment rates, poverty rates, remoteness, and education levels,  can explain how the experience of cancer prevalence varies for various members of a society.
+Few atlases provide this level of detail.
+Data is available from the United Nations Human Development Index and financial indicators are available from the World Bank [@IARC_3].
 
@@
 ## Cartograms
+<!-- 1496 words inc table-->
 
@@
 Choropleths imply uniformity of data across the geographic space however population densities are extremely unlikely to be uniform [@BCM].
-A cartogram displays abstracted geographic space, with the intention of improving the presentation of the statistic of interest.
-For a single variable of interest, each map area is changed to emphasize the distribution by representing the corresponding value, in comparison to the value of the other areas [@ACCAC].
-Changes in the map base are implemented by altering the boundaries, and therefore shapes, of individual areas.
-
-Australia presents an extreme case of an urban rural divide.
-The land mass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas.
-To present election results on a choropleth map should be 'unthinkable', as it means diminishing the visual impact of majority of the electorates.
-A 1966 cartogram presented an alternative where boundary lines were largely straight line, and the result looked very little like the geographical shape of Australia.
-This issue is felt in any nation which experiences a spatially heterogeneous population distribution.
-As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
+The cartogram was developed to address this, by transforming the map to better indicate population [@ACCAC].
+The resulting display more accurately communicates the impact of the disease across the population, as recorded by the statistic, at the sacrifice of geographic accuracy.
 
 <!-- There are many alternative views of the world-->
-<!-- Tobler map projections-->
+<!-- Tobler map projections - I think this is not relevant here
 Choropleths may be considered true topological maps, however, if the land mass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D [@HTLWM].
@@
 If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed, beyond choosing a reasonable projection. 
+-->
 
-```{r choroCRS, fig.cap = "Four choropleth maps of the United States of America using various coordinate reference systems.
-Each state has been colored according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-The map projections alter the shapes and angles of the boundaries of each state.
-Maps a and b are similar in their straight edges, unlike maps c and d which curve on the northern United States border.", message=FALSE, warning=FALSE, out.width='100%', fig.align='center'}
+```{r choroCRS, fig.cap = "Four choropleth maps average age-adjusted rate of incidence for lung and bronchus in the United States averaged over sex and years 2012-2016, using four different map projections.
+The map projections alter the shapes and angles of the boundaries of each state.", message=FALSE, warning=FALSE, out.width='100%', fig.align='center'}
 #knitr::include_graphics("figures/ggchoroCRS.png")
@@
 <!-- Denominators -->
-Event cartograms change the area of regions on a map depending on the amount of disease related events, but this does not consider the effects of land area and population [@VSSDCUC].
 The purposeful distortion of the map space, transformed according to population density, is beneficial when a uniform density of the map base is desired.
 Population then becomes a uniformly distributed background for the statistic presented [@GOINO].
-@ACTUC suggests 'population distribution is often extremely uneven in former British colonies', this makes the distortion necessary [@CTTMB].
+"population distribution is often extremely uneven in former British colonies" [@ACTUC ], and this makes the distortion necessary [@CTTMB].
 When implementing a distortion of the geographical shape according to population, the resulting display is an area cartogram [@NAC], or population-by-area cartogram [@TAAM].
+Event cartograms change the area of regions on a map depending on the amount of disease related events, but this does not consider the effects of land area and population [@VSSDCUC].
 
@@
 The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC]. 
-@HTLWM suggests that white lies may be employed to create useful displays and map creators have the ability to draw lines that may distort the geometry and suppress features and it is easy for the average person to disregard the impact of transformations used to create cartograms.
+Monmonier [@HTLWM] suggests that white lies may be employed to create useful displays and map creators have the ability to draw lines that may distort the geometry and suppress features and it is easy for the average person to disregard the impact of transformations used to create cartograms.
 Cartograms may be seen as an extension of map transformations and projections. The favorable distortion is proportional to a value other other than actual earth size area [@NAC].
@@
 The spatial transformation of map regions relative to the data emphasizes the data distribution instead of land size [@CBATCC].
-When visualizing population statistics, @ACTUC considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
-@HEDP suggests that 'cancer occurs in people, not in geographical areas' and @CTTMB believe that spatial socio-economic data, like cancer rates, are best presented on a cartogram for urban areas as the population map base avoids allocating 'undue prominence' to rural areas.
-@MTMSIH encourage the use of cartograms to highlight small areas and uncover local-level inequalities.
+When visualizing population statistics, Dorling [@ACTUC] considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
+Howe [@HEDP] suggests that 'cancer occurs in people, not in geographical areas' and @CTTMB believe that spatial socio-economic data, like cancer rates, are best presented on a cartogram for urban areas as the population map base avoids allocating 'undue prominence' to rural areas.
+<!-- @MTMSIH encourage the use of cartograms to highlight small areas and uncover local-level inequalities. DI SAYS I think that focuses on bayesian analysis not on cartograms-->
 
@@
 Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilized depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994), [@CD].
-@HEDP discusses the impact of electronic computer-assisted techniques.
+Howe [@HEDP] discusses the impact of electronic computer-assisted techniques.
 
 There are many alternatives to consider, the intended audience of the map, and its purpose are key points in cartogram use and creation.
-@ACTUC reiterates: 'There is no "best" cartogram or method of creating cartograms just as there is no "best" map' (Monmonier and Schnell, 1988). 
-@SAIC provided a framework to investigate implementations of the many algorithms presented, and the "statistical accuracy, geographical accuracy, and topological accuracy".
+ 'There is no "best" cartogram or method of creating cartograms just as there is no "best" map' (Monmonier and Schnell, 1988) [@ACTUC]. 
+Nusrat and Kobourov [@SAIC] provided a framework to investigate implementations of the many algorithms presented, and the "statistical accuracy, geographical accuracy, and topological accuracy".
 
@@
 |a.
-contiguous cartogram| Each stated shape has been distorted according to the population of the state in 2015.
+Contiguous| Each stated shape has been distorted according to the population of the state in 2015.
 The state of California has become much larger due to it's large population density.
@@
 |b.
-Non - Contiguous| The geographic shape of the states has been maintained, but the size has altered according to the population of the state in 2015.
+Non-contiguous| The geographic shape of the states has been maintained, but the size has altered according to the population of the state in 2015.
 The state of California has remained closer to it's original size than it's surrounding states.
@@
 
-```{r auscartograms, fig.cap = "Alternative maps of the United States of America.
-Each state has been colored according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
-Each stated shape has been distorted according to the population of the state in 2015.
-The state of California has become much larger due to it's large population density.
-This draws attention to the densely populated North East region, and detracts from the less populated Mid West.", message=FALSE, warning=FALSE, out.width='100%'}
-aus1 <- png::readPNG("figures/aus_ggcont.png")
-aus2 <- png::readPNG("figures/aus_ggncont.png")
-aus3 <- png::readPNG("figures/aus_ggdorl.png")
-aus4 <- png::readPNG("figures/aus_gghexmap.png")
-load("aus_legend.rda")
-
-load("figures/aus_legend.rda")
-
-library(cowplot)
-ggdraw() +
-  draw_plot(rasterGrob(aus1), 0, .55, 0.5, 0.55) +
-  draw_plot(rasterGrob(aus2), 0.5, 0.55, 0.5, 0.55) +
-  draw_plot(rasterGrob(aus3), 0.0, 0.1, 0.5, 0.55) +
-  draw_plot(rasterGrob(aus4), 0.5, 0.1, 0.5, 0.55) +
-  draw_plot(aus_legend, 0, 0, 1, .1) +
-  draw_plot_label(c("a", "b", "c", "d"), 
-    c(0, .5, 0, 0.5), 
-    c(1, 1, .5, .5), size = 20)
-```
+Figure \ref{fig:ggcartograms} shows four different cartograms for the same data displayed in Figure \ref{fig:choroCRS}.
+Table {tab:usa} summarizes information that can be observed in the four types of cartograms.
 
@@
 
+<!-- 
 An intentional goal when creating the 1966 Census population cartogram for Canada was to maintain contiguity, while attempting to keep the actual shape of places.
@@
 @CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes. 
+-->
 
-### Non-Contiguous
+### Non-contiguous
 
-Non-Contiguous cartograms succeed in maintaining the shape of the areas presented.
+Non-contiguous cartograms preserve shape of the area over connectivity.
 Each area stays in a similar position to their location on a choropleth map.
@@
 The addition is the gap between areas, created as each individual area shrinks or grows according the associated value of the statistic.
-@NAC discusses creation of these maps, the significance of the empty areas left between the geographic boundaries and the new shape, and the 'degree of difference from the original map that is the real message' of these displays.
+Olson [@NAC] discusses creation of these maps, the significance of the empty areas left between the geographic boundaries and the new shape, and the 'degree of difference from the original map that is the real message' of these displays.
 
 As the trade-off regarding boundaries approaches simplicity, the distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the non-contiguous cartogram but is replaced by the meaningful empty-space property [@ECGC, @NAC].
-The shapes are valuable for recognition and allows users to orient themselves on the display.
+
+<!-- The shapes are valuable for recognition and allows users to orient themselves on the display.
 Map creators can efficiently communicate with this kind of map by keeping the outlines or particular elements of the original in the new shape [@NISCC].
@@
 However it may impact on the visibility of all areas if small areas expand beyond their boundaries. 
-Figure \ref{fig:ggcartograms} b shows small overlaps in the north east region of the United States.
+Figure \ref{fig:ggcartograms}b shows small overlaps in the north east region of the United States.
 These overlaps are also shown in the non-contiguous cartogram of Australia in Figure \ref{fig:auscartograms} b.
+-->
 
@@
 A similar method, where each geographic area is represented using a square, tessellated to create a square grid.
-This method has been used by FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
+This method has been used broadly in the media, e.g.
+FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
 Each area is represented by a square of the same dimensions.
-Figure \ref{fig:ggcartograms} d shows a hexagon tile map representing each state using a single hexagon.
+<!--
+Figure \ref{fig:ggcartograms}d shows a hexagon tile map representing each state using a single hexagon.
 
@@
 Algorithms for the creation of rectangular cartograms
+-->
 
@@
 
-## Geofaceting
+## Geofacet
 Tile maps can be extended to allow more than just simple coloring of tiles to present data.
@@
 # Comparison of maps and cartograms {#ch:compare}
+<!-- 640 words -->
 
@@
 
+```{r auscartograms, fig.cap = "Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape presaerved, (c) Dorling, (d) hexmap.
+The contiguous cartogram has expanded the highly populated areas, while preserving the full shapes of rural areas.
+If it accurately sized areas by population the country would be unrecognisable.
+The shape-preserved is unreadable.
+Dorling has lost all geographic context.
+The hexmap provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%'}
+aus1 <- png::readPNG("figures/aus_ggcont.png")
+aus2 <- png::readPNG("figures/aus_ggncont.png")
+aus3 <- png::readPNG("figures/aus_ggdorl.png")
+aus4 <- png::readPNG("figures/aus_gghexmap.png")
+load("aus_legend.rda")
+
+load("figures/aus_legend.rda")
+
+library(cowplot)
+ggdraw() +
+  draw_plot(rasterGrob(aus1), 0, .55, 0.5, 0.50) +
+  draw_plot(rasterGrob(aus2), 0.5, 0.55, 0.5, 0.50) +
+  draw_plot(rasterGrob(aus3), 0.0, 0.1, 0.5, 0.55) +
+  draw_plot(rasterGrob(aus4), 0.5, 0.1, 0.5, 0.55) +
+  draw_plot(aus_legend, 0, 0, 1, .1) +
+  draw_plot_label(c("a", "b", "c", "d"), 
+    c(0, .5, 0, 0.5), 
+    c(1, 1, .5, .5), size = 20)
+```
+
+## Neither choropleths or cartograms perform well for Australia
+
+Australia presents an extreme case of an urban rural divide.
+The land mass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas.
+To present election results on a choropleth map should be 'unthinkable', as it means diminishing the visual impact of majority of the electorates.
+A 1966 cartogram presented an alternative where boundary lines were largely straight line, and the result looked very little like the geographical shape of Australia.
+This issue is felt in any nation which experiences a spatially heterogeneous population distribution.
+As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
+
 # User interaction {#ch:interacting}
+<!-- 634 words -->
 
-Interactivity and animation can add value to online atlases, by enabling additional information to be presented on-the-fly with one display.
+<!-- Interactivity and animation can add value to online atlases, by enabling additional information to be presented on-the-fly with one display.
 It also invites the viewer to engage with the display, and the information.
@@
 The latter two would be considered to be direct manipulation on the map display, and the former two are GUI interfaces. The effects of these actions are rapid, incremental, and reversible [@DMIV]. 
+-->
 
-There is a long and extensive literature about interactivity with maps.
-[@mcgranaghan1993cartographic] discuss the way adding interactive elements can alleviate cognitive overload caused when too much information is loaded onto a map.
-[@cliburn2002design] discuss the trade-offs required for expert vs novice audiences.
-The expert might desire more complexity for decision-making but this could overwhelm a novice reader.
+One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
+It can be a juggling act for a diverse audience, with experts probably preferring more detail [@cliburn2002design].
+Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
+The effect of  user-centered interactive actions needs to produce rapid, incremental, and reversible changes to the display [@DMIV].
 
@@
 @TNTEA says user interaction with maps helps them to understand and interpret the spatial distribution of disease, to validate, explain or explore the presented statistics and their relationships to each other.
+
+Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display.
+Interactive design features, found in online cancer maps, include tool tips, drop down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
+The use of these supports were found in various online cancer maps identified by @roberts2019communication.
 
@@
 
-## Interaction and animation in publicly available atlases
+<!-- 
+XXX This part needs to go to interactivity!
+One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
+These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
+Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
 
-Many publicly available cancer atlases employ interactivity and/or animation.
-Figure \ref{fig:interacting} shows the interactive GUI tools provided with the public atlases discussed in Section \ref{sec:publicatlases}.
-Mostly these are menus, and toggles, for choosing indicators, and subsets of the population.
+Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display.
+Interactive design features, found in online cancer maps, include tool tips, drop down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
+The use of these supports were found in various online cancer maps identified by @roberts2019communication.
+The controls for basic interactive features are often placed outside of the plot space [@TGA], thus the map image is updated/replaced as the user interacts with the controls.
+For example, changing the population age or other demographic variables.
+Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection tools.
+
+Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
+Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).
+
+These interactive features provide an opportunity for users to explore the additional information available.
+This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
+This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accessible' ways [@TNTEA].
+The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
+@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
+-->
 
@@
 ```
+
+```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a.
+Linked maps and time series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b.
+A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 9}
+a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
+a2 <- png::readPNG("cancer_map_images/animation/2.animation.png")
+
+ggdraw() +
+  draw_plot(rasterGrob(a1), 0, 0, 0.3, 1) +
+  draw_plot(rasterGrob(a2), 0.3, 0, 0.7, 1) +
+  draw_plot_label(c("I", "II"), 
+    c(0, .3), c(1, 1), size = 20)
+```
+
+## Interaction and animation in publicly available atlases
+
+Many publicly available cancer atlases employ interactivity and/or animation.
+Figure \ref{fig:interacting} shows the interactive GUI tools provided with the public atlases discussed in Section \ref{sec:publicatlases}.
+Mostly these are menus, and toggles, for choosing indicators, and subsets of the population.
 
@@
 
-```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a.
-Linked maps and time series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b.
-A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 9}
-a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
-a2 <- png::readPNG("cancer_map_images/animation/2.animation.png")
-
-ggdraw() +
-  draw_plot(rasterGrob(a1), 0, 0, 0.3, 1) +
-  draw_plot(rasterGrob(a2), 0.3, 0, 0.7, 1) +
-  draw_plot_label(c("I", "II"), 
-    c(0, .3), c(1, 1), size = 20)
-```
-
 <!-- Animations can be implemented to allow users varying degrees of control.
@@
 # Conclusions
+<!-- 307 words -->
 
@@
 # Acknowledgements
-
+<!-- 77 words -->
 The authors would like thank Dr Earl Duncan, Professor Kerrie Mengersen, Dr Susanna Cramb and Dr Peter Baade for conversations on the content of this article.

### 2019-09-30 review_article.Rmd
done with alternative displays

@@
 ## Tile map
-
+<!-- 173 words -->
 A tile map provides a tessellated display of consistent shapes.
@@
 Tile maps may be difficult to create, they are best created manually, with additional time and care required as the amount of geographic areas to include increases.
-@MDAC define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area can be used to communicate the statistic of regions. 
+Cano and others [@MDAC] define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area can be used to communicate the statistic of regions. 
 The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off be made between boundary complexity and simplicity.
@@
 ## Geofacet
-Tile maps can be extended to allow more than just simple coloring of tiles to present data.
-@IGF formalizes the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas. 
+<!-- 166 words -->
+Hafen [@IGF] formalizes the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas. 
 Like tile maps, the arrangement of the areas can be reused for other spatial distribution visualizations.
@@
 Virtually any plot display can be fit into the tile representing the areas, allowing displays of multiple variables or values per geographic entity.
-As the amount of areas increase care must be taken to ensure legibility of the displays.
-@RPIG acknowledges that this display suffers when the areas have very irregular shapes and large size disparities, as manual creation of the grid is still required.
+<!-- As the amount of areas increase care must be taken to ensure legibility of the displays.
+@RPIG acknowledges that this display suffers when the areas have very irregular shapes and large size disparities CORRECT reference is Cheng's thesis-->
+Manual creation of the grid is required.
 
 ## Multivariate displays
+<!-- 222 words -->
 
-@MMST present linked micromap plots to visually link geographic and statistical data, this serves as a solution to  multi-dimensionality issues.
-This design is widely used in online atlas displays.
+Pickle and others [@MMST] present linked micromap plots to visually link geographic and statistical data, this serves as a solution to  multi-dimensionality issues.
+This design has been used in online atlas displays of US states.
 Areas are grouped based on their value for one variable, and additional columns provide displays that contrast the areas in each group by other variables.
@@
 Bivariate choropleth maps [@VUADBC] blend colour schemes to convey the intersection of categorised levels of an estimate and the associated uncertainty for each spatial area.
-@VUADBC also suggests map pixelation, each region is broken into small pixels, the individual pixels are allocated values that reflect the uncertainty around the area's estimate by creating texture.
+Lucchesi and Wikle [@VUADBC] also suggests map pixelation, each region is broken into small pixels, the individual pixels are allocated values that reflect the uncertainty around the area's estimate by creating texture.
 This display can also be animated, with each frame produced by resampling the pixels.
@@
 
-Solutions have been provided for bi-variate displays for movement data, containing Origin - Destination matrices and flow map representations [@MMTVCG]. 
+<!-- Solutions have been provided for bi-variate displays for movement data, containing Origin - Destination matrices and flow map representations [@MMTVCG]. 
 This method makes the maps and matrix connections clear, and easy to link geographic areas with link lines not crossing. Bar charts can also be added to these to show group totals.
+-->
 

### 2019-11-12 sa3_plots.R
plots for symposium

@@ -9,6 +9,7 @@ library(ggthemes)
 library(maptools)
 library(sf)
 library(cartogram)
+library(cowplot)
 
 
 # Filter for centroids only within longitudes
@@ -67,16 +68,25 @@ aus <- absmapsdata::state2016 %>%
 
 ###############################################################################
 
+invthm <- theme_void() + theme(
+  panel.background = element_rect(fill = "transparent", colour = NA), 
+  plot.background = element_rect(fill = "transparent", colour = NA),
+  legend.background = element_rect(fill = "transparent", colour = NA),
+  legend.key = element_rect(fill = "transparent", colour = NA),
+  text = element_text(colour = "white", size = 20),
+)
+
 # Create a choropleth
 aus_ggchoro <- ggplot(sa3lung) + 
   geom_sf(aes(fill = `Age-standardised rate (per 100,000)`), colour = NA) +
   scale_fill_distiller(type = "seq", palette = "Purples",  direction = 1, na.value = "light grey") + 
-  theme_void() +
+  
   coord_sf(crs = CRS("+init=epsg:3112"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
-  theme(legend.position ="bottom") + 
+  invthm + theme(legend.position ="bottom") +
   labs(fill = "Age-standardised rate\n(per 100,000)")
 aus_ggchoro
-ggsave(filename = "figures/aus_ggchoro.png", device = "png", dpi = 300,  width = 7, height = 6)
+ggsave(filename = "figures/aus_ggchoro.png", device = "png",
+  bg = "transparent",  dpi = 300,  width = 7, height = 6)
 
 
 aus_legend <- get_legend(aus_ggchoro)
@@ -92,13 +102,14 @@ cont <- sa3lung %>%
   st_as_sf()
 
 save(cont, "data/auscont.rda")
+load("data/auscont.rda")
 aus_ggcont <- ggplot(cont) + 
   geom_sf(aes(fill = `Age-standardised rate (per 100,000)`)) + 
   scale_fill_distiller(type = "seq", palette = "Purples",  direction = 1) + 
   coord_sf(crs = CRS("+init=epsg:3112"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
-  theme_void()+guides(fill=FALSE)
+  invthm +  guides(fill=FALSE)
 aus_ggcont
[…]

### 2019-11-12 cover letter/cover letter.Rmd

@@
 fromsignature: sigfile.png
-toname: Editors-in-Chief
-opening: Dear Editors-in-Chief Dr Freddie Bray and Dr Wanqing Chen,
-closing: Warm regards
+toname: Editors-in-Chief Dr Freddie Bray and Dr Wanqing Chen
+opening: Dear Dr Bray and Dr Chen,
+closing: Sincerely,
 spacing: 1.4
@@
 
-We were honoured to receive the request from Drs. Baade and Cramb, and Editor Kathryn Yan, to explore the use of mapping methods to present cancer statistics.
-This review paper presents an exploration of cancer mapping methods, considering the use of choropleths in many current online cancer atlases.
+We were honoured to receive an invitation to contribute this paper from Drs. Baade and Cramb, and Editor Kathryn Yan.
+This review paper presents an exploration of cancer mapping methods, focusing on variations on choropleths for online cancer atlases.
 It highlights the current uses of interactivity and how this allows users to engage with cancer statistics in an online atlas.

### 2019-11-12 review_article.Rmd

@@
     number_sections: yes
-    toc: yes
+    toc: no
     toc_depth: 3
   word_document:
-    toc: yes
+    toc: no
     toc_depth: '3'
@@
 Hexagon Tessellation| Each state is represented by a hexagon of equal size.
-The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location.
+The neighboring states are easily contrasted, however the north east regions has been displaced from their geographic location.
 The sparsity of the population in the Mid West is highlight by the light yellow color, the age adjusted rate in Kentucky is the darkest and its neighbors are similar.|
@@
 
-Daniel Dorling presented an alternative display engineers to highlight the spatial distribution and neighbourhood relationships without complex distortions of borders and boundaries.
+Daniel Dorling presented an alternative display engineers to highlight the spatial distribution and neighborhood relationships without complex distortions of borders and boundaries.
 This approach opposes preserving the intricate shape details and is founded in the simple question put forward by Daniel Dorling -@ACTUC:
@@
 This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms, as the 'nonsense' shapes used have 'no meaning' after distortions are applied @NISCC.
-To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighbouring areas over many iterations.
+To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighboring areas over many iterations.
 This implementation can work for up to 'one hundred thousand' areas.
@@
 it shows one map per group of the key separating variable, in a row with each additional statistical plot.
-Linked micromaps predominantly utilize the choropleth map for displays of spatial relationships, that are seen when spatial neighbours are alotted to the same group.
+Linked micromaps predominantly utilize the choropleth map for displays of spatial relationships, that are seen when spatial neighbors are alotted to the same group.
 It is one of several alternative displays that allow maps to become bivariate displays, commonly used to present both an estimate and the associated uncertainty.
 
-Bivariate choropleth maps [@VUADBC] blend colour schemes to convey the intersection of categorised levels of an estimate and the associated uncertainty for each spatial area.
+Bivariate choropleth maps [@VUADBC] blend color schemes to convey the intersection of categorized levels of an estimate and the associated uncertainty for each spatial area.
 Lucchesi and Wikle [@VUADBC] also suggests map pixelation, each region is broken into small pixels, the individual pixels are allocated values that reflect the uncertainty around the area's estimate by creating texture.
@@
 
-# Comparison of maps and cartograms {#ch:compare}
-<!-- 640 words -->
-
-<!-- Choropleths, true topology, geographic distance, good for showing geographic importance -->
-To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
-Choropleth displays utilize more traditional cartographic methods, they are usually true to the topography, displaying familiar boundaries of countries, states or administrative areas.
-Highlighting the geographic distribution, and differences between equal spatial units.
-When there is a relationship between the variable being mapped and geographical features, a choropleth map would pair this information visually.
-Choropleth methods require a decision to be made about the projection of the display, in this case aim to find a map projection that gives the least distorted representation of the geography [@UMPT].
-However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a color.
-These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
-Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbors. The map is reusable and fit for many purposes.
-
-<!-- Cartograms, sacrifice geography, focus on the outliers/ relationships in data space -->
-Cartographers were historically the creators of geographic displays, and epidemiologists utilized the maps that had been prepared in advance.
-Cartograms with the population variable used to distortion the geographic map regions and boundaries are density equalising maps, that reduce the visual prominence of large, but low-population areas [@VSSDCUC].
-Cartogram displays incorporate the statistics and population into the design of the display,
-shifting and sometimes sacrificing familiar boundaries to draw attention to large outliers in the data space or population density.
-The unusual shapes are not longer familiar and may be difficult to compare.
-This can prevent effective comparison of the statistic.
-Cartograms also favor extreme statistics in the positive direction.
-Small extreme values are often lost due to the resulting small sized areas the do not draw user's attention.
-The change in the map based on one variable provides the opportunity for bi-variate displays, using the change in area for one variable, and color for another.
-The difference between the familiar map display helps highlight the impact of the disease on communities.
-
-<!-- Alternatives, remove geographic shapes, keep spatial relationship, focus on the outliers/ relationships in population space -->
-Alternative maps may replace the unexpected boundaries and shapes with familiar, simple shapes. This makes the spatial distribution the primary concept communicated by the display.
-Areas that are geographically close will maintain connectedness in some way, but the population and the statistic will dominate.
-Tile map displays may only allow one variable to be visualized if each region is given consistent map space, this occurs in geofaceting. 
-Dorlings provide bi-variate displays as they use the size of the shape for one variable, and color for another.
-These alternatives are especially helpful for data aggregations where administrative boundaries break populations into groups.
-The experience of each community may be worth considering, as the experiences shared by the population within them may be similar due to the services and facilities they share.
-This display allows a more equitable view of each community, and does not minimise those that operate on a smaller geographic scale.
-
-Faceting provides an opportunity to contrast the overall distribution of variables across a geographic space or collection of geographic areas.
-This display makes is difficult to compare specific regions.
-
-Table: \label{tab:methods} Summary of features and constraints on common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
-
-|                     | Choropleth | Contiguous | Non-contig | Dorling | Tile maps | Geofacets |
-|:-----------|:------------:|:-----------:|:---------------:|:---------:|:-----------:|:----------:|
-|Spatial distortion   |  N  |  Y  |  Y  |  Y  |  Y  |  Y  |
-|Preserves neighbors  |  Y  |  Y  |  Y  |  S  |  S  |  S  |
-|Conceals small areas |  Y  |  S  |  N  |  N  |  N  |  N  |
-|Uniform shape        |  N  |  N  |  N  |  Y  |  Y  |  Y  |
-|Univariate only      |  Y  |  Y  |  Y  |  S  |  S  |  N  |
-|Manual construction  |  N  |  N  |  N  |  N  |  Y  |  Y  |
-
-Creating maps of diseases now involves more decisions to be made by map makers, rather than cartographers. 
-Technology has played an enormous role in increasing the opportunities for map makers.
-The computation and graphics power have made creation, alteration and interactivity possible.
-as these options have expanded and it is the objectives of the investigator that will drive the choices.
-@CPISACA and @SAMGIS have provided suggestions and comments to help map creators best communicate their health data and spatial analyses.
+# Comparison and critique {#ch:compare}
+<!-- 408 words -->
 
@@
 To present election results on a choropleth map should be 'unthinkable', as it means diminishing the visual impact of majority of the electorates.
-A 1966 cartogram presented an alternative where boundary lines were largely straight line, and the result looked very little like the geographical shape of Australia.
+Dorling [@ACTUC] has a cartogram showing the 1966 general election results, which looked very little like the geographical shape of Australia.
 This issue is felt in any nation which experiences a spatially heterogeneous population distribution.
 As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
+
+Figure \ref{fig:auscartograms} shows four main types of cartograms displaying melanoma incidence on SA3 areas in Australia.
+The version of a contiguous cartogram (a) has expanded the highly populated areas, while preserving the full shapes of rural areas.
+It has not fully resolved the population transformation of areas, and if it had accurately sized areas by population the country would be unrecognizable.
+The shape-preserved cartogram is unreadable, and all areas have been reduced to tiny spots on the map.
+Zooming in on a high-resolution output shows that it does indeed preserve the areas though.
+The Dorling cartogram has lost all geographic context.
+The hexmap provides a reasonable spatial distribution despite having too much white-space in the outback areas.
+
+## Pros and cons of different choices
+
+Choropleths provide a familiar display.
+This may be less threatening to an audience that could be frightened by statistics, but they suffer primarily from unfaithfully representing the disease information by hiding small geographic areas.
+Inset maps like in Brisbane city in Figure \ref{fig:choropleth-grid-create}c of the state of Queensland, are commonly used to reduce distorted interpretations, but it is a bandaid remedy.
+
+Cartograms provide the spatial distortion to more accurately convey the statistical distribution, focuses on the human impact of the disease.
+However, if the population density is highly dissonant with geographic density the cartogram will lose all spatial context, and thus futile.
+
+Some mix of tiling, faceting or even micromaps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies.
+Table \ref{tab:methods} summarizes key criteria for evaluating map and alternative displays.
+Suggestions and comments to help map creators best communicate their health data and spatial analyses a re provided in @CPISACA and @SAMGIS.
+
+<!-- Choropleths, true topology, geographic distance, good for showing geographic importance 
+To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
+Choropleth displays utilize more traditional cartographic methods, they are usually true to the topography, displaying familiar boundaries of countries, states or administrative areas.
+Highlighting the geographic distribution, and differences between equal spatial units.
+When there is a relationship between the variable being mapped and geographical features, a choropleth map would pair this information visually.
+Choropleth methods require a decision to be made about the projection of the display, in this case aim to find a map projection that gives the least distorted representation of the geography [@UMPT].
+However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a color.
+These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
+Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbors. The map is reusable and fit for many purposes.
+-->
+
+<!-- Cartograms, sacrifice geography, focus on the outliers/ relationships in data space 
+Cartographers were historically the creators of geographic displays, and epidemiologists utilized the maps that had been prepared in advance.
+Cartograms with the population variable used to distortion the geographic map regions and boundaries are density equalizing maps, that reduce the visual prominence of large, but low-population areas [@VSSDCUC].
+Cartogram displays incorporate the statistics and population into the design of the display,
+shifting and sometimes sacrificing familiar boundaries to draw attention to large outliers in the data space or population density.
+The unusual shapes are not longer familiar and may be difficult to compare.
+This can prevent effective comparison of the statistic.
+Cartograms also favor extreme statistics in the positive direction.
+Small extreme values are often lost due to the resulting small sized areas the do not draw user's attention.
+The change in the map based on one variable provides the opportunity for bi-variate displays, using the change in area for one variable, and color for another.
+The difference between the familiar map display helps highlight the impact of the disease on communities.
+-->
+
+<!-- Alternatives, remove geographic shapes, keep spatial relationship, focus on the outliers/ relationships in population space 
+Alternative maps may replace the unexpected boundaries and shapes with familiar, simple shapes. This makes the spatial distribution the primary concept communicated by the display.
+Areas that are geographically close will maintain connectedness in some way, but the population and the statistic will dominate.
+Tile map displays may only allow one variable to be visualized if each region is given consistent map space, this occurs in geofaceting. 
+Dorlings provide bi-variate displays as they use the size of the shape for one variable, and color for another.
+These alternatives are especially helpful for data aggregations where administrative boundaries break populations into groups.
+The experience of each community may be worth considering, as the experiences shared by the population within them may be similar due to the services and facilities they share.
+This display allows a more equitable view of each community, and does not minimise those that operate on a smaller geographic scale.
+
+Faceting provides an opportunity to contrast the overall distribution of variables across a geographic space or collection of geographic areas.
+This display makes is difficult to compare specific regions.-->
+
+Table: \label{tab:methods} Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
+
+|                     | Choropleth | Contiguous | Non-contig | Dorling | Tile maps | Geofacets |
+|:-----------|:------------:|:-----------:|:---------------:|:---------:|:-----------:|:----------:|
+|Spatial distortion   |  N  |  Y  |  Y  |  Y  |  Y  |  Y  |
+|Preserves neighbors  |  Y  |  Y  |  Y  |  S  |  S  |  S  |
+|Conceals small areas |  Y  |  S  |  N  |  N  |  N  |  N  |
+|Uniform shape        |  N  |  N  |  N  |  Y  |  Y  |  Y  |
+|Univariate only      |  Y  |  Y  |  Y  |  S  |  S  |  N  |
+|Manual construction  |  N  |  N  |  N  |  N  |  Y  |  Y  |
+
+<!-- Creating maps of diseases now involves more decisions to be made by map makers, rather than cartographers. 
+Technology has played an enormous role in increasing the opportunities for map makers.
+The computation and graphics power have made creation, alteration and interactivity possible.
+as these options have expanded and it is the objectives of the investigator that will drive the choices.
+@CPISACA and @SAMGIS have provided suggestions and comments to help map creators best communicate their health data and spatial analyses.-->
 
@@
 Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
-The effect of  user-centered interactive actions needs to produce rapid, incremental, and reversible changes to the display [@DMIV].
+Effective  user-centered interactive actions produce rapid, incremental, and reversible changes to the display [@DMIV].
 
-@HTLWM says that interactivity can be used to allow users to explore the map for more information and provides flexibility for the display.
+Monmonier [@HTLWM] says that interactivity can be used to allow users to explore the map for more information and provides flexibility for the display.
 The user can toggle between different variables, map views or even multiple realizations of future scenarios [@goodchild1994introduction].
@@
 
-```{r interacting, fig.cap = "Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
-i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png') # b in choro grid
-i2 <- png::readPNG('cancer_map_images/interactivity/15.interactivity.png') # d in choro grid
-i3 <- png::readPNG('cancer_map_images/interactivity/4.interactivity.png')
-i4 <- png::readPNG('cancer_map_images/interactivity/5.interactivity.png')
-i5 <- png::readPNG('cancer_map_images/interactivity/7.interactivity.png') # e in choro grid
-
-ggdraw() +
-  draw_plot(rasterGrob(i1), 0,  0.13,   .28,  .83) + #a
-  draw_plot(rasterGrob(i2), 0.32, 0.56, .68, .4) + #b
-  draw_plot(rasterGrob(i3), 0.32, 0.41, .68, .13) + #c
-  draw_plot(rasterGrob(i4), 0.32, 0.14, .68, .25) + #d
-  draw_plot(rasterGrob(i5), 0, 0, 1, .13) + #e
-  draw_plot_label(c("i", "ii", "iii", "iv", "v"), 
-    c(0.01, 0.3, 0.3, 0.3, 0.01), 
-    c(0.99, 0.99, 0.57, 0.40, 0.15), size = 20)
-
-ggsave(filename = "figures/supp_grid.png", device = "png", 
-  width = 12, height = 9, dpi = 300)
-```
-
-```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a.
-Linked maps and time series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b.
-A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 9}
-a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
-a2 <- png::readPNG("cancer_map_images/animation/2.animation.png")
-
-ggdraw() +
-  draw_plot(rasterGrob(a1), 0, 0, 0.3, 1) +
-  draw_plot(rasterGrob(a2), 0.3, 0, 0.7, 1) +
-  draw_plot_label(c("I", "II"), 
-    c(0, .3), c(1, 1), size = 20)
-```
-
 ## Interaction and animation in publicly available atlases
@@
 Many publicly available cancer atlases employ interactivity and/or animation.
-Figure \ref{fig:interacting} shows the interactive GUI tools provided with the public atlases discussed in Section \ref{sec:publicatlases}.
+Figure 6 <!--\ref{fig:interacting}--> shows the interactive GUI tools provided with the public atlases discussed in Section \ref{sec:publicatlases}.
 Mostly these are menus, and toggles, for choosing indicators, and subsets of the population.
 
-## Under-utilised techniques
+## Under-utilized techniques
 
-Figure \ref{fig:animating} shows two examples of more sophisticated interactive maps.
+Figure 7 <!--\ref{fig:animating}--> shows two examples of more sophisticated interactive maps.
 The Spanish Cancer map (left) contains a linked display between a choropleth map, and time series plots of cancer change.
@@
 -->
+
+```{r interacting, fig.cap = "Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
+i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png') # b in choro grid
+i2 <- png::readPNG('cancer_map_images/interactivity/15.interactivity.png') # d in choro grid
+i3 <- png::readPNG('cancer_map_images/interactivity/4.interactivity.png')
+i4 <- png::readPNG('cancer_map_images/interactivity/5.interactivity.png')
+i5 <- png::readPNG('cancer_map_images/interactivity/7.interactivity.png') # e in choro grid
+
+ggdraw() +
+  draw_plot(rasterGrob(i1), 0,  0.13,   .28,  .83) + #a
+  draw_plot(rasterGrob(i2), 0.32, 0.56, .68, .4) + #b
+  draw_plot(rasterGrob(i3), 0.32, 0.41, .68, .13) + #c
+  draw_plot(rasterGrob(i4), 0.32, 0.14, .68, .25) + #d
+  draw_plot(rasterGrob(i5), 0, 0, 1, .13) + #e
+  draw_plot_label(c("i", "ii", "iii", "iv", "v"), 
+    c(0.01, 0.3, 0.3, 0.3, 0.01), 
+    c(0.99, 0.99, 0.57, 0.40, 0.15), size = 20)
+
+ggsave(filename = "figures/supp_grid.png", device = "png", 
+  width = 12, height = 9, dpi = 300)
+```
+
+```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a.
+Linked maps and time series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b.
+A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 9}
+a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
+a2 <- png::readPNG("cancer_map_images/animation/2.animation.png")
+
+ggdraw() +
+  draw_plot(rasterGrob(a1), 0, 0, 0.3, 1) +
+  draw_plot(rasterGrob(a2), 0.3, 0, 0.7, 1) +
+  draw_plot_label(c("I", "II"), 
+    c(0, .3), c(1, 1), size = 20)
+```
 

### 2019-12-08 version1/Response/Response.Rmd
begin response

---
from: Stephanie Kobakian
fromqualifications: BA
fromtitle: Research Masters Student
fromemail: stephanie.kobakian@monash.edu
fromphone: +61 433 699 797
toname: Professor Peter Baade and Dr Susanna Cramb
opening: Dear Professor Peter Baade and Dr Susanna Cramb
closing: Warm regards
spacing: 1.4
output: MonashEBSTemplates::letter
---

Thank you for your letter containing your feedback regarding our submision "Cartogram Mapping and its Application to Cancer Data Visualization", for inclusion in the Focused Issue on Spatial Patterns in Cancer Epidemiology in the Annals of Cancer Epidemiology.

This review paper presented an exploration of cancer mapping methods, focusing on variations on choropleths for online cancer atlases.

**Editor suggestion:**
**We would suggest that the aim of the manuscript could be phrased to highlight the current approaches and limitations of typical visualisation methods for cancer maps, along with outlining possible alternative methods and challenges in implementing these?**

We have restructed our original submission to provide a clear path from current approaches used, to the alternative approaches that should be explored. We highlight the limitations of using current display methods, to encourage the use of  alternative display methods. We also present the challenges that may be faced when implementing the alternative designs.

The manuscript now contains the following sections:

Current approaches
Limitations of current approaches
Alternative visualisation methods
Challenges of alternative approaches

We have inclued changes to reflect your minor notes:

**Editor suggestion:**
**The authors mention a focus on Atlases published between 2010 and 2015, but then include a bowel cancer atlas from 2016. **

**Also, within the bowel cancer atlas data, the reporting of percent of males aged 50-54 with bowel cancer in 2016 is incorrect; it should be the % of the ERP in each area that are males aged 50-54 years. **

We have corrected the description of the statistic used in the bowel cancer atlas of Australia:

Bowel Cancer Australia Atlas:
...shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia  . It is published by *[Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)*.

**Reviewer A: **
**Major**
*Comment 1: *
The Abstract should be more focused- perhaps give an example of how cartogram mapping has been applied to data visualisation. What are the benefits of this approach and suggestions for the future.

*Reply 1:* 
We agree with the reviewer that alternative approaches have been fantastic when used in political contexts. Choropleth maps have often been used.

*Changes in the text:*

*Comment 2: *

2. The key message seems to get lost in the Abstract. The Conclusion is vague.

*Reply :*

*Changes in the text:*

*Comment 3: *

3. Entire paper needs substantial English language editing

*Reply :*

*Changes in the text:*

*Comment 4: *

4. Please avoid the use of sentences which have been used throughout the paper where you start with a reference number: such as “[4] discusses the use of (Map displays for disease data, Line 15-16) They do not make any sense and must be modified.

*Reply :*

*Changes in the text:*

*Comment 5: *

5. Please provide numbers for Figures and Tables

*Reply :*

*Changes in the text:*

*Comment 6: *

6. Please clearly state the primary aim of this paper and structure it accordingly.

*Reply :*

*Changes in the text:*

*Comment 7: *

7. Conclusions must discuss the key focus and conclusions of this paper.

*Reply :*

*Changes in the text:*

*Minor Comments*

- Please check the use of past and present tense
- There are many minor grammatical errors that need to checked.
- Please edit the last section of the Introduction. Either number each section and then refer to them or find another way to appropriately refer to each section.

We thank you for your consideration.

### 2019-12-08 version1/Response/Response.tex
begin response

\documentclass[11pt,a4paper,]{letter}
\usepackage{bera}

\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{geometry}
\geometry{a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color,hyperref,url,fontawesome}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[absolute,overlay]{textpos}
\setlength{\TPHorizModule}{1cm}
\setlength{\TPVertModule}{1cm}

\newsavebox{\lethdone}
\newsavebox{\lethdtwo}
\newsavebox{\headname}

\sbox{\lethdone}{\hbox{\includegraphics[width=6.5cm]{MBSportrait}}}

\newlength{\junkw}

\makeatletter
\def\ps@monash{\begin{textblock}{8}(1.2,1)
\usebox{\lethdone}
\end{textblock}
\begin{textblock}{6}(13.5,1)
\usebox{\headname}
\end{textblock}
\begin{textblock}{10}(1.2,27.2)\usebox{\lethdtwo}\end{textblock}\def\thepage{}}

\@ifundefined{opening}{}{%
\renewcommand*{\opening}[1]{\thispagestyle{monash}%
   {\@date\par}%
  \vspace{2\parskip}%
  {\raggedright \toname \\ \toaddress \par}%
  \vspace{2\parskip}%
  #1\par\nobreak}}
\makeatother

\def\section#1{\vspace{0.3cm}\par{\textsf{\bfseries\Large #1}}\vspace*{0.02cm}\par}
\def\subsection#1{\vspace{0.3cm}\par{\textsf{\bfseries\large #1}}\vspace*{0.02cm}\par} %}

% Date
\def\Date{\number\day}
\def\Month{\ifcase\month\or
 January\or February\or March\or April\or May\or June\or
 July\or August\or September\or October\or November\or December\fi}
\def\Year{\number\year}

% Spacing
\RequirePackage{setspace}

% Fix href problems
\def\href#1{}

\sbox{\headname}{\parbox{10cm}{\fontsize{10}{12}\sffamily\color[gray]{0.4}
{\large\textbf{Stephanie Kobakian}}\\
BA\\
Research Masters Student\\[0.2cm]
\begin{tabular}{@{}ll@{}}
      \faicon{home}       & \\
      \faicon{envelope}   & \href{mailto:stephanie.kobakian@monash.edu}{\nolinkurl{stephanie.kobakian@monash.edu}}\\
      \faicon{phone}      & +61 433 699 797
\end{tabular}}}

\sbox{\lethdtwo}{\hbox{\fontsize{9}{11}\sffamily\color[gray]{0.4}\begin{tabular}{@{}ll@{}}
\multicolumn{2}{@{}l}{Department of Econometrics \& Business Statistics}\\
\multicolumn{2}{@{}l}{Monash University}\\
\multicolumn{2}{@{}l}{Victoria 3800, Australia.}\\[0.2cm]
\multicolumn{2}{@{}l}{ABN: 12 377 614 012\quad CRICOS Provider Number: 00008C}
\end{tabular}}}

\date{\Date~\Month~\Year}

\begin{document}
\begin{letter}{Professor Peter Baade and Dr Susanna Cramb\\}
\setstretch{1}
\vspace*{2cm}
\opening{Dear Professor Peter Baade and Dr Susanna Cramb}
\setstretch{1.4}
Thank you for your letter containing your feedback regarding our submision ``Cartogram Mapping and its Application to Cancer Data Visualization'', for inclusion in the Focused Issue on Spatial Patterns in Cancer Epidemiology in the Annals of Cancer Epidemiology.

This review paper presented an exploration of cancer mapping methods, focusing on variations on choropleths for online cancer atlases.

\textbf{Editor suggestion:}
\textbf{We would suggest that the aim of the manuscript could be phrased to highlight the current approaches and limitations of typical visualisation methods for cancer maps, along with outlining possible alternative methods and challenges in implementing these?}

We have restructed our original submission to provide a clear path from current approaches used, to the alternative approaches that should be explored. We highlight the limitations of using current display methods, to encourage the use of alternative display methods. We also present the challenges that may be faced when implementing the alternative designs.

The manuscript now contains the following sections:

Current approaches
Limitations of current approaches
Alternative visualisation methods
Challenges of alternative approaches

We have inclued changes to reflect your minor notes:

\textbf{Editor suggestion:}
\textbf{The authors mention a focus on Atlases published between 2010 and 2015, but then include a bowel cancer atlas from 2016. }

\textbf{Also, within the bowel cancer atlas data, the reporting of percent of males aged 50-54 with bowel cancer in 2016 is incorrect; it should be the \% of the ERP in each area that are males aged 50-54 years. }

We have corrected the description of the statistic used in the bowel cancer atlas of Australia:

Bowel Cancer Australia Atlas:
\ldots{}shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia . It is published by \emph{\href{https://www.bowelcanceraustralia.org/}{Bowel Cancer Australia}}.

\textbf{Reviewer A: }
\textbf{Major}
\emph{Comment 1: }
The Abstract should be more focused- perhaps give an example of how cartogram mapping has been applied to data visualisation. What are the benefits of this approach and suggestions for the future.

\emph{Reply 1:}
We agree with the reviewer that alternative approaches have been fantastic when used in political contexts. Choropleth maps have often been used.

\emph{Changes in the text:}

\emph{Comment 2: }

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  The key message seems to get lost in the Abstract. The Conclusion is vague.
\end{enumerate}

\emph{Reply :}

\emph{Changes in the text:}

\emph{Comment 3: }

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Entire paper needs substantial English language editing
\end{enumerate}

\emph{Reply :}

\emph{Changes in the text:}

\emph{Comment 4: }

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Please avoid the use of sentences which have been used throughout the paper where you start with a reference number: such as ``{[}4{]} discusses the use of (Map displays for disease data, Line 15-16) They do not make any sense and must be modified.
\end{enumerate}

\emph{Reply :}

\emph{Changes in the text:}

\emph{Comment 5: }

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Please provide numbers for Figures and Tables
\end{enumerate}

\emph{Reply :}

\emph{Changes in the text:}

\emph{Comment 6: }

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{5}
\tightlist
\item
  Please clearly state the primary aim of this paper and structure it accordingly.
\end{enumerate}

\emph{Reply :}

\emph{Changes in the text:}

\emph{Comment 7: }

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{6}
\tightlist
\item
  Conclusions must discuss the key focus and conclusions of this paper.
\end{enumerate}

\emph{Reply :}

\emph{Changes in the text:}

\emph{Minor Comments}

\begin{itemize}
\tightlist
\item
  Please check the use of past and present tense
\item
  There are many minor grammatical errors that need to checked.
\item
  Please edit the last section of the Introduction. Either number each section and then refer to them or find another way to appropriately refer to each section.
\end{itemize}

We thank you for your consideration.
\closing{Warm regards\\[0.2cm]\hspace*{0.5cm}\includegraphics[height=1.5cm]{}}
\end{letter}

\end{document}

### 2019-12-08 review_article.Rmd
begin response

@@
 <!-- 119 words -->
-Choropleth maps are widely used in cancer atlases, and traditionally communicate cancer statistics over geographic domains.
+<!--
+Choropleth maps are widely used in cancer atlases to communicate cancer statistics over geographic domains.
 The type of statistic displayed varies immensely, and effective choices can assist in the understanding of impact of cancer on a population.
@@
 This is especially true for the geography of Australia, where the population is concentrated is small areas along the coastlines.
+-->
+
+Cancer atlases communicate cancer statistics over geographic domains. These domains are often subdivided by administrative areas such as countries, states or suburbs.
+When communicating human-related statistics in Australia, the geographic map base may highlight the sparsely populated rural areas.
+The smaller geographic areas may not draw the attention of readers but they may be important to consider if they are more densely populated.
+To
 
@@
 Counts alone are not sufficient to compare areas, because the populations of areas are all different.
-In this case, the counts data needs to be merged with population data to appropriately calibrate it to incidence, for example, rate per 100000 people.
-This type of data is collected on a routine basis for public health purposes, and may be made available to the general public as a service to the community.
+In this case, the count data needs to be merged with population data to appropriately calibrate it to incidence, for example, the rate per 100000 people.
+This type of data is collected on a routine basis for public health purposes and may be made available to the general public as a service to the community.
 The task, then, is to examine what are the usual ways to communicate cancer statistics, to the public, are there alternative approaches, and what are the pros and cons of these choices.
@@
 <!-- Ways to visualize, whats our hypothesis -->
-A common approach to communicate cancer statistics, is to display statistics on a map.
+A common approach to communicate cancer statistics is to display statistics on a map.
 Using a choropleth map: the statistic is mapped to color and the geographic region is filled with this color.
@@
 Disease maps help to present geographic patterns that may be overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
-By providing a visual representation of cancer outcomes, geographic patterns of disease may be identified and effectively addressed with public health policy and actions.
+By providing a visual representation of cancer outcomes, geographic patterns of the disease may be identified and effectively addressed with public health policy and actions.
 @SE recognizes one of the key challenges with mapping spatial patterns of disease is the design of visualizations.
@@
 Section \ref{ch:alternatives} describes  alternative displays, including the cartogram which is useful when the map has heterogeneously sized geographic units.
-The pros and cons of these approaches is discussed in Section \ref{ch:compare}.
+The pros and cons of these approaches are discussed in Section \ref{ch:compare}.
 Disease maps are more useful when made interactive, and common options are described in Section \ref{ch:interacting}, along with a discussion of benefits and disadvantages.
@@
 The geography is faithfully rendered, and the color rendering is designed to reveal spatial patterns among data values.
-Identifying and explaining spatial structures, patterns, and processes involves considering the individuals in communities and organizing communities into representable units [@SAMGIS].
+Identifying and explaining spatial structures, patterns, and processes involve considering the individuals in communities and organizing communities into representable units [@SAMGIS].
 Early versions of choropleth maps used symbols or patterns instead of color.
@@
 
-Figure \ref{fig:choroCRS} shows various geographic projections of choropleth maps of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged over 2012 through 2016.
+Figure \ref{fig:choroCRS} shows various geographic projections of choropleth maps of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged from 2012 to 2016.
 The data was extracted from the official federal statistics by @USInteractive on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
-A west to east spatial trend of increasing rates, can be seen.
-There is also a spatial outlier -- Utah has a noticeably lower rate that its neighbors.
-Also Kentucky has a noticeably high rate, and Maine also has a higher rate than its neighbors.
+The rates exhibit a spatial trend, increasing from west to east.
+There is also a spatial outlier -- Utah has a noticeably lower rate than its neighbors.
+Also, Kentucky has a noticeably high rate, and Maine also has a higher rate than its neighbors.
 There is something of a cluster of higher rates around the tobacco states.
@@
 A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
-Atlases are key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets [@MACM].
-The data collection of statistics across regions, and the administrative control within regions, lends itself to choropleth visualization.
+Atlases are key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work-related exposures, and high-risk diets [@MACM].
+The data collection of statistics across regions and the administrative control within regions lends itself to choropleth visualization.
 Atlases date to Haviland's maps in 1875, with UK cancer atlases appearing in 1961 [@HEDP], and early work in US cancer atlases appearing in 1971 [@burbank].
@@
 <!-- Measures reported in cancer atlases -->
-Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the higher prevalence of cancers in older populations.
+Mortality rates are now often presented as relative rates of risk across the population, and age-adjusted to correct for the higher prevalence of cancers in older populations.
 Howe [-@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s. Table \ref{tab:measures} summarizes the measures commonly presented in published cancer atlases.
@@
 |:--------------------|:------------------------------------------------|  
-|1. Count|Crude cancer counts|
-|2. Rate per 100,000|Cancer incidence per 100,000 population|  
-|3. IR (Incidence Ratio)|$(IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}$, |
-||Cancer incidence rate in region $i$ over the average cancer incidence rate for the total region|
-|4. Age Adjusted Rate per 100,000| Standardized by age structure or region|   
-|5. Age Adjusted Relative Risk| Standardized by age structure in each region $i$|  
+|1. Count| Crude cancer counts|
+|2. Rate per 100,000| Cancer incidence per 100,000 population|  
+|3. IR (Incidence Ratio)| $(IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}$, |
+|| The cancer incidence rate in region $i$ over the average cancer incidence rate for all of the regions|
+|4. Age-Adjusted Rate per 100,000| Standardized by age structure or region|   
+|5. Age-Adjusted Relative Risk| Standardized by age structure in each region $i$|  
 |6. SIR (Standardized Incidence Ratio)| Standardized by incidence in each region $i$|
-|7. Below or above Expected|Alternative expression of the SIR|
-|8. RER |$RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}$|
-|(Relative Excess Risk)|Represents the estimate of cancer related mortality within five years of diagnosis. Also referred to as 'excess hazard ratio'|
+|7. Below or above Expected| An alternative expression of the SIR|
+|8. RER | $RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}$|
+|(Relative Excess Risk)| Represents the estimate of cancer-related mortality within five years of diagnosis. Also referred to as 'excess hazard ratio'|
 
@@
 
-Cancer maps are effective tools for communicating incidence, survival and mortality to wide range of audiences, including the general public and others not trained in statistical analyses.
+Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the general public and others not trained in statistical analyses.
 These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analyses.
@@
 All of these use choropleth maps.
-All but one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government funded partnerships.
+All but one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships.
 Figure \ref{fig:choropleth-grid-create} displays a subset of maps from these atlases.
@@
 Figure \ref{fig:choropleth-grid-create}a shows the relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2].
-The cancer data was from Office for National Statistics (ONS) (England) and from the Welsh Cancer Intelligence and Surveillance Unit (WCISU).
+The cancer data was from the Office for National Statistics (ONS) (England) and the Welsh Cancer Intelligence and Surveillance Unit (WCISU).
 
@@
 
-The map seen in Figure \ref{fig:choropleth-grid-create}b shows age standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018 [@Globocan].
+The map seen in Figure \ref{fig:choropleth-grid-create}b shows age-standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018 [@Globocan].
 It is published by the [World Health Organization's International Agency for Research on Cancer](https://www.iarc.fr/).
@@
 ### Bowel Cancer Australia Atlas 
-Figure \ref{fig:choropleth-grid-create}d shows the percentage of Australian males between 50 - 54 years of age diagnosed with bowel cancer in 2016 in Australia [@Bowel].
+Figure \ref{fig:choropleth-grid-create}d shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia [@Bowel].
 It is published by *[Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)*.
@@
 
-Figure \ref{fig:choropleth-grid-create}f shows side by side maps of relative risk of lung cancer for men vs women for 2004 to 2008 [@cancerSpain].
+Figure \ref{fig:choropleth-grid-create}f shows side by side maps of the relative risk of lung cancer separately for men and women from 2004 to 2008 [@cancerSpain].
 
@@
 
-It is extremely likely that each population area will have a different number of people.
-The distribution of the population residing in all areas may also be communicated in a table or histogram display [@NICR_1].
+Each population area will likely have a different number of people. The distribution of the population residing in all areas may also be communicated in a table or histogram display [@NICR_1].
 Atlases can connect the population to the land available to them by communicating population density. Alternatives to choropleth to address this are described in Section \ref{ch:alternatives} .
@@
 With any number comes some uncertainty about its value.
-This may occur for a number of reasons: numbers are drawn from samples, errors occur in the disease reporting, values are simulated from a model to maintain privacy.
+This can occur for several reasons: numbers are drawn from samples, errors occur in the disease reporting, values are simulated from a model to maintain privacy.
 Close to half of the atlases identified (42%, n=14) included some measure of uncertainty.
@@
 Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it can be difficult to display effectively.
-The map focuses on displaying the statistic, and lacks additional space to represent the uncertainty.
+The map focuses on displaying the statistic and lacks additional space to represent the uncertainty.
 Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two approaches.
@@
 
-Socio-economic indicators, such asunemployment rates, poverty rates, remoteness, and education levels,  can explain how the experience of cancer prevalence varies for various members of a society.
+Socio-economic indicators, such as unemployment rates, poverty rates, remoteness, and education levels,  can explain how the experience of cancer prevalence varies for various members of society.
 Few atlases provide this level of detail.
@@
 <!-- Tobler map projections - I think this is not relevant here
-Choropleths may be considered true topological maps, however, if the land mass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D [@HTLWM].
+Choropleths may be considered true topological maps however,, if the landmass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D [@HTLWM].
 The amount of distortion is related to the distance covered by the landmass displayed @GAMP.
@@
 Choropleth creation requires choosing a map projection that shows a favorable distortion of the geography for presenting the set of spatial information.
-Selecting a display can prevent misinterpretation of global statistics, as global maps face the challenge of equitable displays of land mass on maps [@RSCW].
+Selecting a display can prevent misinterpretation of global statistics, as global maps face the challenge of equitable displays of landmass on maps [@RSCW].
 If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed, beyond choosing a reasonable projection. 
@@
 
-```{r choroCRS, fig.cap = "Four choropleth maps average age-adjusted rate of incidence for lung and bronchus in the United States averaged over sex and years 2012-2016, using four different map projections.
+```{r choroCRS, fig.cap = "Four choropleth maps average age-adjusted rate of incidence for lung and bronchus in the United States averaged from sex and years 2012-2016, using four different map projections.
 The map projections alter the shapes and angles of the boundaries of each state.", message=FALSE, warning=FALSE, out.width='100%', fig.align='center'}
@@
 The purposeful distortion of the map space, transformed according to population density, is beneficial when a uniform density of the map base is desired.
-Population then becomes a uniformly distributed background for the statistic presented [@GOINO].
+The population then becomes a uniformly distributed background for the statistic presented [@GOINO].
 "population distribution is often extremely uneven in former British colonies" [@ACTUC ], and this makes the distortion necessary [@CTTMB].
 When implementing a distortion of the geographical shape according to population, the resulting display is an area cartogram [@NAC], or population-by-area cartogram [@TAAM].
-Event cartograms change the area of regions on a map depending on the amount of disease related events, but this does not consider the effects of land area and population [@VSSDCUC].
+Event cartograms change the area of regions on a map depending on the amount of disease-related events, but this does not consider the effects of land area and population [@VSSDCUC].
 
@@
 The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC]. 
-Monmonier [@HTLWM] suggests that white lies may be employed to create useful displays and map creators have the ability to draw lines that may distort the geometry and suppress features and it is easy for the average person to disregard the impact of transformations used to create cartograms.
-Cartograms may be seen as an extension of map transformations and projections. The favorable distortion is proportional to a value other other than actual earth size area [@NAC].
-A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, making interpretation of spatial patterns very difficult.
+Monmonier [@HTLWM] suggests that white lies may be employed to create useful displays and map creators can draw lines that may distort the geometry and suppress features and it is easy for the average person to disregard the impact of transformations used to create cartograms.
+Cartograms may be seen as an extension of map transformations and projections. The favorable distortion is proportional to a valueother than the actual earth size area [@NAC].
+A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, interpreting spatial patterns very difficult.
 The distortion of a cartogram accounts for the population density, preventing it from obscuring the spatial patterns [@TAAM]. 
@@
 The creation of cartograms was historically in the hands of professional cartographers [@CD]. 
-Early approaches including John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball bearing approach and Tobler's (1973) computer programs [@ACTUC].
-Geographical information systems allowed map users, and researchers to implement their own cartograms, but these systems are utilized depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994), [@CD].
+Early approaches including John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball-bearing approach and Tobler's (1973) computer programs [@ACTUC].
+Geographical information systems allowed map users, and researchers to create cartograms, but these systems are utilized depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994), [@CD].
 Howe [@HEDP] discusses the impact of electronic computer-assisted techniques.
@@
 
-```{r ggcartograms, fig.cap = "Common alternatives to maps, showing the same information as in Figure \ref{choroCRS} for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexmap (non-contiguous).
+```{r ggcartograms, fig.cap = "Common alternatives to maps, showing the same information as in Figure \ref{choroCRS} for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous).
 In (a) - (c) the state has been resized, and reshaped, to match the 2015 population of the state.
-This provides a better sense of the extent of disease relative to the population in the country, and can help alleviate losing information about physically small but population dense states.
-In the hexmap (d) each state is given equal size, and thus equal emphasis.", message=FALSE, warning=FALSE, out.width='100%'}
+This provides a better sense of the extent of disease relative to the population in the country and can help alleviate losing information about physically small but population-dense states.
+In the hexagon tilemap (d) each state is given equal size and thus equal emphasis.", message=FALSE, warning=FALSE, out.width='100%'}
 g1 <- png::readPNG("figures/ggcont.png")
@@
 Contiguous| Each stated shape has been distorted according to the population of the state in 2015.
-The state of California has become much larger due to it's large population density.
-This draws attention to the densely populated North East region, and detracts from the less populated Mid West.|
+The state of California has become much larger due to its large population density.
+This draws attention to the densely populated North-East region and detracts from the less populated Mid West.|
 |b.
 Non-contiguous| The geographic shape of the states has been maintained, but the size has altered according to the population of the state in 2015.
-The state of California has remained closer to it's original size than it's surrounding states.
-The North East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut.
-This draws attention to the densely populated North East region, and the sparse Mid West.|
+The state of California has remained closer to its original size than it's surrounding states.
+The North-East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut.
+This draws attention to the densely populated North-East region and the sparse Mid West.|
 |c.
 Dorling| The states have been represented by a circle, but the size was determined by the population of the state in 2015.
-The North East states remain closer to their neighbors, and may be displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.|
+The North-East states remain closer to their neighbors and could be displaced from their geographic location.
+The sparsity of the population in the Mid West is highlighted by the distance between the circles, located at the geographic centroids.|
 |d.
 Hexagon Tessellation| Each state is represented by a hexagon of equal size.
-The neighboring states are easily contrasted, however the north east regions has been displaced from their geographic location.
-The sparsity of the population in the Mid West is highlight by the light yellow color, the age adjusted rate in Kentucky is the darkest and its neighbors are similar.|
+The neighboring states are easily contrasted however the North-East regions have been displaced from their geographic location.
+The sparsity of the population in the Mid West is highlighteded by the light yellow color, the Age-Adjusted rate in Kentucky is the darkest and its neighbors are similar.|
 
@@
 
-Non-contiguous cartograms preserve shape of the area over connectivity.
-Each area stays in a similar position to their location on a choropleth map.
-The choropleth map base is often also plotted as a comparison point to highlight the change in area.
-The addition is the gap between areas, created as each individual area shrinks or grows according the associated value of the statistic.
-Olson [@NAC] discusses creation of these maps, the significance of the empty areas left between the geographic boundaries and the new shape, and the 'degree of difference from the original map that is the real message' of these displays.
+Non-contiguous cartograms prioritize the shapes of the areas instead of connectivity.
+Each area stays in a similar position to its location on a choropleth map.
+The choropleth map base is often also displayed to allow comparisons to be made regarding the change in the area displayed.
+The addition is the gap between areas, created as each area shrinks or grows according to the associated value of the statistic.
+Olson [@NAC] discusses the creation of these maps, the significance of the empty areas left between the geographic boundaries and the new shape, and the 'degree of difference from the original map that is the real message' of these displays.
 
@@
 However it may impact on the visibility of all areas if small areas expand beyond their boundaries. 
-Figure \ref{fig:ggcartograms}b shows small overlaps in the north east region of the United States.
+Figure \ref{fig:ggcartograms}b shows small overlaps in the North-East region of the United States.
 These overlaps are also shown in the non-contiguous cartogram of Australia in Figure \ref{fig:auscartograms} b.
@@
 
-Daniel Dorling presented an alternative display engineers to highlight the spatial distribution and neighborhood relationships without complex distortions of borders and boundaries.
+Daniel Dorling presents an alternative display engineered to highlight the spatial distribution and neighborhood relationships without complex distortions of borders and boundaries.
 This approach opposes preserving the intricate shape details and is founded in the simple question put forward by Daniel Dorling -@ACTUC:
@@
 He acknowledged the sophistication of contiguous cartograms but critiqued their 'very complex shapes,' he answers this with his implementation of maps created using 'the simplest of all shapes'. 
-Circular cartograms use the same simple shapes for every region represented, and resizes the shapes according to the statistic represented or the population for a base map.
-This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms, as the 'nonsense' shapes used have 'no meaning' after distortions are applied @NISCC.
-To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighboring areas over many iterations.
+Circular cartograms use the same simple shapes for every region represented and resize the shapes according to the statistic represented or the population for a base map.
+This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms, as the 'nonsense' shapes shown in contiguous cartograms have 'no meaning' after distortions are applied @NISCC.
+To produce a compelling map, a gravity model is applied to avoid overlaps and keep spatial relationships with neighboring areas over many iterations.
 This implementation can work for up to 'one hundred thousand' areas.
 
-The groundwork for this approach had been laid in the mid 1930's by Raisz, and rectangular cartograms provide dramatic comparisons and are especially useful for correcting misconceptions communicated by geographic maps.
+The groundwork for this approach had been laid in the mid-1930s by Raisz, and rectangular cartograms provide dramatic comparisons and are especially useful for correcting misconceptions communicated by geographic maps.
 @TFYCC quotes the official definition of Value-Area Cartograms, the simplistic displays which represent each area as a single rectangle, sized according to the value of the statistic.
-This rectangular display also allows for tiling, where geographic neighbors placed in suitable relative positions also share borders, however contiguity may be sacrificed [@CDWCS].
-Rectangular cartograms allow for bi-variate displays, population can be effectively communicated by the size of each rectangular, and a second variable can be communicated using color [@ORC].
+This rectangular display also allows for tiling, where geographic neighbors placed in suitable relative positions also share borders however, contiguity may be sacrificed [@CDWCS].
+Rectangular cartograms allow for bi-variate displays as the population can be effectively communicated by the size of each rectangular, and a second variable can be communicated using color [@ORC].
 
@@
 <!--
-Figure \ref{fig:ggcartograms}d shows a hexagon tile map representing each state using a single hexagon.
+Figure \ref{fig:ggcartograms}d shows a hexagon tilemap representing each state using a single hexagon.
 
@@
 
-## Tile map
+## tilemap
 <!-- 173 words -->
-A tile map provides a tessellated display of consistent shapes.
-Each tile is usually one unit of measurement, this could be geographic regions such as states, or population based where one tile is used for a consistent measure of population.
+A tilemap provides a tessellated display of consistent shapes.
+Each tile is usually one unit of measurement, this could be geographic regions such as states, or population-based where one tile is used for a consistent measure of population.
 <!-- cite mosaic cartograms -->
 
-A simple tile map presents the areas in a tessellated grid display, where geographic neighbors are found next to each other, with some necessary displacement employed for regions with more than four neighbors.
-These tiles may be labelled or colored to represent a value.
-Tile maps may be difficult to create, they are best created manually, with additional time and care required as the amount of geographic areas to include increases.
+A simple tilemap presents the areas in a tessellated grid display, where geographic neighbors are found next to each other, with some necessary displacement employed for regions with more than four neighbors.
+These tiles may be labeled or colored to represent a value.
+tilemaps may be difficult to create, they are best created manually, with additional time and care required as the number of geographic areas to include increases.
 Cano and others [@MDAC] define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area can be used to communicate the statistic of regions. 
-The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off be made between boundary complexity and simplicity.
+The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off to be made between boundary complexity and simplicity.
 
-```{r ggtilemap, fig.cap = "Two more alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
-In the tile map, the layout approximates spatial location, with each state being an equal box filled with colour representing cancer incidence.
+```{r ggtilemap, fig.cap = "Two more alternative displays, tilemap (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
+In the tilemap, the layout approximates spatial location, with each state being an equal box filled with color representing cancer incidence.
 The geofaceted map shows bar charts laid out in a grid approximating the spatial location of the state.
-The age-adjusted rates for males and females is shown.
+The age-adjusted rates for males and females are shown.
 This type of display allows multiple variables to be displayed for each geographic area.", out.width="100%"}
@@
 Hafen [@IGF] formalizes the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas. 
-Like tile maps, the arrangement of the areas can be reused for other spatial distribution visualizations.
-Geofaceting have the functionality of facets, often used to replicate visualizations for each subset of the data, in this case the data subsets are geographic.
-The amount of information able to be communicated has increased from one value per region in a tile map to one visualization, this is a more flexible display.
+Like tilemaps, the arrangement of the areas can be reused for other spatial distribution visualizations.
+Geofaceting has the functionality of facets, often used to replicate visualizations for each subset of the data. In this case, the data subsets are geographic.
+The amount of information able to be communicated has increased from one value per region in a tilemap to one visualization, this is a more flexible display.
 Virtually any plot display can be fit into the tile representing the areas, allowing displays of multiple variables or values per geographic entity.
@@
 
-```{r auscartograms, fig.cap = "Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape presaerved, (c) Dorling, (d) hexmap.
-The contiguous cartogram has expanded the highly populated areas, while preserving the full shapes of rural areas.
-If it accurately sized areas by population the country would be unrecognisable.
+```{r auscartograms, fig.cap = "Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tilemap.
+The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas.
+If it accurately sized areas by population the country would be unrecognizable.
 The shape-preserved is unreadable.
 Dorling has lost all geographic context.
-The hexmap provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%'}
+The hexagon tilemap provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%'}
+
 aus1 <- png::readPNG("figures/aus_ggcont.png")
@@
 
-Australia presents an extreme case of an urban rural divide.
-The land mass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas.
-To present election results on a choropleth map should be 'unthinkable', as it means diminishing the visual impact of majority of the electorates.
+Australia presents an extreme case of an urban-rural divide.
+The landmass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas.
+To present election results on a choropleth map should be 'unthinkable', as it means diminishing the visual impact of the majority of the electorates.
 Dorling [@ACTUC] has a cartogram showing the 1966 general election results, which looked very little like the geographical shape of Australia.
@@
 Figure \ref{fig:auscartograms} shows four main types of cartograms displaying melanoma incidence on SA3 areas in Australia.
-The version of a contiguous cartogram (a) has expanded the highly populated areas, while preserving the full shapes of rural areas.
+The version of a contiguous cartogram (a) has expanded the highly populated areas while preserving the full shapes of rural areas.
 It has not fully resolved the population transformation of areas, and if it had accurately sized areas by population the country would be unrecognizable.
@@
 The Dorling cartogram has lost all geographic context.
-The hexmap provides a reasonable spatial distribution despite having too much white-space in the outback areas.
+The hexagon tilemap provides a reasonable spatial distribution despite having too much white-space in the outback areas.
 
@@
 Some mix of tiling, faceting or even micromaps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies.
-Table \ref{tab:methods} summarizes key criteria for evaluating map and alternative displays.
-Suggestions and comments to help map creators best communicate their health data and spatial analyses a re provided in @CPISACA and @SAMGIS.
+Table \ref{tab:methods} summarizes the key criteria for evaluating maps and alternative displays.
+Suggestions and comments to help map creators best communicate their health data and spatial analyses are provided in @CPISACA and @SAMGIS.
 
@@
 However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a color.
-These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
+These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighteded by the inset map.
 Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbors. The map is reusable and fit for many purposes.
@@
 Areas that are geographically close will maintain connectedness in some way, but the population and the statistic will dominate.
-Tile map displays may only allow one variable to be visualized if each region is given consistent map space, this occurs in geofaceting. 
+tilemap displays may only allow one variable to be visualized if each region is given consistent map space, this occurs in geofaceting. 
 Dorlings provide bi-variate displays as they use the size of the shape for one variable, and color for another.
@@
 
-|                     | Choropleth | Contiguous | Non-contig | Dorling | Tile maps | Geofacets |
+|                     | Choropleth | Contiguous | Non-contig | Dorling | tilemaps | Geofacets |
 |:-----------|:------------:|:-----------:|:---------------:|:---------:|:-----------:|:----------:|
@@
 
-- drop down menus to select features,
+- drop-down menus to select features,
 - variable selection, 
-- tool tips, 
+- tooltips, 
 - zooming and panning.
@@
 This provides additional mechanisms for the users to digest the uncertainty of the available information [@maceachren1992visualizing, @van1994visualization].
-Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions that allow inspection of the data from many views [@DQBCM].
+When the needs of the audience are changeable and are also the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions.
+This can allow inspection of the data from many views [@DQBCM].
 @TNTEA says user interaction with maps helps them to understand and interpret the spatial distribution of disease, to validate, explain or explore the presented statistics and their relationships to each other.
@@
 Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display.
-Interactive design features, found in online cancer maps, include tool tips, drop down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
-The use of these supports were found in various online cancer maps identified by @roberts2019communication.
+Interactive design features, found in online cancer maps, include tooltips, drop-down menus, data selection, zooming, and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
+The use of these supports can be found in various online cancer maps identified by @roberts2019communication.
 
-Animation, in contrast to interactivity, usually involves pre-computing views, and showing these in a sequence.
+Animation, in contrast to interactivity, usually involves pre-computing views and showing these in a sequence.
 @TGA provided an overview of animation in association with the R package `gganimate` [@gganimate], with some focus on maps, explaining transitions between frames in a linear, to provide a fixed narrative to passive users.
-Animations are used to communicate a message by capturing and directing users attention.
+Animations are used to communicate a message by capturing and directing users' attention.
 It is most often employed to show changes over time.
@@
 @CPISACA provide weather maps as a thoroughly developed example of animation of spatial displays to communicate information to the general public.
-The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period of time.
+The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period.
 
 The Australian Cancer Atlas [-@TACA] provides [tours](
-https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
-This implementation of animation gives users tools to plan their own exploration.
+https://atlas.cancer.org.au/app/tour/lungcancer) that change the display to the draw users' attention to areas on the map that are relevant to the story.
+This implementation of animation gives users tools to plan their exploration.
 
@@
 Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display.
-Interactive design features, found in online cancer maps, include tool tips, drop down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
+Interactive design features, found in online cancer maps, include tooltips, drop-down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
 The use of these supports were found in various online cancer maps identified by @roberts2019communication.
@@
 Figure 7 <!--\ref{fig:animating}--> shows two examples of more sophisticated interactive maps.
-The Spanish Cancer map (left) contains a linked display between a choropleth map, and time series plots of cancer change.
+The Spanish Cancer map (left) contains a linked display between a choropleth map and time series plots of cancer change.
 In linked plots, changing values in one display will trigger changes of corresponding elements in another display.
@@
 Mousing over the time series plots will highlight the line for a particular region.
-The Canadian Breast Cancer Mortality map (right) has a magnifying glass which allows the user to zoom into small areas.
-It is easy to control and shows precise details in the small areas.
+The Canadian Breast Cancer Mortality map (right) has a magnifying glass that allows the user to zoom into small areas.
+It is easy to control and shows precise details in small areas.
 
@@
 ```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a.
-Linked maps and time series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b.
+Linked maps and time-series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b.
 A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 9}
@@
 <!-- summary of mapping -->
-This paper provides an overview of mapping practices as used for cancer atlases, and new approaches that could be adopted.
+This paper provides an overview of mapping practices as used for cancer atlases and new approaches that could be adopted.
 The conventional approach is the choropleth map, and it is widely used.
 When there are small areas, as occurs in Australia where the population is concentrated on the coast, the information about cancer can be lost, and alternatives are needed.
-Making an inset can clarify congested regions but this breaks the viewers attention, because they need to shift focus from the map to the inset, and if there are many congested areas, many insets would be needed.
+Making an inset can clarify congested regions but this breaks the viewers' attention because they need to shift focus from the map to the inset, and if there are many congested areas, many insets would be needed.
 The map alternatives, like cartograms and their variations, can be useful to allow the spatial distribution of cancer data to be digested.
@@
 <!-- statistics -->
-There are many different statistics that are commonly used for display.
-The most basic is incidence rate, which is easy to understand.
+Many different statistics are commonly used for cancer displays.
+The most basic is the incidence rate, which is easy to understand.
 It is common to see relative rates in many maps which measure how far a region is above or below the average.
@@
 However, we lose the incidence rate information and thus interpretability.
-A region might be much higher than average, but it may not be close to a health concern, because all regions have low incidence.
+A region might be much higher than average, but it may not be close to a health concern, because all regions have a low incidence.
 
 <!-- interaction -->
-Interaction with maps is an important component to public atlases.
+Interaction with maps is an important component of public atlases.
 A purpose is to provide access to more information than is possible to display in a single map, without overwhelming the viewer.
 Too many choices though can similarly overwhelm a viewer, and thus decisions do need to be made about content to provide, or hide, for accurate and comprehensive communication of information.
-Similarly, providing ways for users to interact with the display encourages engagement, and creative, efficient, elegant, interactive tools elicits curiosity about the data.
+Similarly, providing ways for users to interact with the display encourages engagement, and creative, efficient, elegant, interactive tools elicit curiosity about the data.
 
@@
 <!-- 77 words -->
-The authors would like thank Dr Earl Duncan, Professor Kerrie Mengersen, Dr Susanna Cramb and Dr Peter Baade for conversations on the content of this article.
+The authors would like to thank Dr. Earl Duncan, Professor Kerrie Mengersen, Dr. Susanna Cramb and Dr. Peter Baade for conversations on the content of this article.
 
@@
 Adding too much information to a map may create cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
-Complexity and density of visualization methods can overwhelm novice decision makers, however experts are able to use the detail more readily when making-decisions [@cliburn2002design]. 
+Complexity and density of visualization methods can overwhelm novice decision makers however, experts are able to use the detail more readily when making-decisions [@cliburn2002design]. 
 Supplementary information can be incorporated into online atlases without cluttering the display.
-Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning.
+Interactive design features found in online cancer maps include tool tip features, drop-down menus, data selection, zooming and panning.
 Allowing users to explore the map for more information and providing flexibility in the display [@HTLWM].
@@
 Interactivity and animation provide an alternative to the traditional approach of a single map for a single use, and enable the map to become more exploratory.
-Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
-Tool tips allow users to obtain additional information regarding specific geographic areas.
+Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tooltips, and data selection tools [@TGA].
+tooltips allow users to obtain additional information regarding specific geographic areas.
 

### 2019-12-08 version1/version2.Rmd
begin response

--- 2019-12-08 review_article.Rmd
@@
 output:
+  word_document:
+    toc: no
+    toc_depth: '3'
   pdf_document:
@@
     toc_depth: 3
-  word_document:
-    toc: no
-    toc_depth: '3'
 link-citations: yes
@@
 ```
-<!--
-
-Please note that the submission deadline is September 30, 2019.
-
-Author notes:
-6000 words maximum.
-focused  application  of  literature  to address  a  relevant  clinical  issue:
-Mapping cancer outcomes
-Cartogram applications should point to examples of cancer outcomes, with the main cartogram section discussing sources considered in grey literature review.
-Examples of political areas may be considered if cancer data is aggregated at these levels.
-
-- Future steps should consider applications not currently utilized for cancer outcome displays, where their use may improve decision making
-- needs to have a clear relevance to spatial patterns in cancer epidemiology
-- presents alternative ways of visualizing those spatial patterns, or the uncertainty in those patterns
-- a more applied-focused paper (e.g. lots of cool images), rather than theoretical (e.g. lots of formula).
--->
 
@@
 <!-- 119 words -->
-<!--
-Choropleth maps are widely used in cancer atlases to communicate cancer statistics over geographic domains.
-The type of statistic displayed varies immensely, and effective choices can assist in the understanding of impact of cancer on a population.
-Atlases have moved from print media to web format, and the availability of graphical user interfaces, and user interaction and animation, greatly enhance the ability to communicate more information and positively engage readers.
-More use of alternatives to choropleth maps, such as cartograms, could be employed in online atlases, for better representation of spatial distributions, and show statistics on small areas.
-This is especially true for the geography of Australia, where the population is concentrated is small areas along the coastlines.
--->
 
@@
 The smaller geographic areas may not draw the attention of readers but they may be important to consider if they are more densely populated.
-To
 
@@
 
-<!-- written below The Atlas of Cancer in Queensland [@DACQ] focused on highlighting the difference in experience for those living in rural and disadvantaged areas, the Standardized Incidence Ratios (SIR) were modeled via Bayesian methods.
-The presentation of these rates considers not only the shapes of the areas, but also appropriate choices of color blind friendly color schemes, and categories of the values to highlight significantly different areas.
--->
-
 ## Overview of publicly available atlases \label{sec:publicatlases}
@@
 Figure \ref{fig:choropleth-grid-create} displays a subset of maps from these atlases.
-
-<!-- While there is significant variation across these maps in terms of the measure represented, the scale of the map, the color choices, the legend design and supplementary/supporting statistics, the mapping approach has not changes and almost exclusively only choropleth maps are used.
-Unlike innovation in spatial statistical methods, interactivity, data collection and record keeping little work has been done to implement different mapping approaches -->
 
@@
 
+While there is significant variation across these maps in terms of the measure represented, the scale of the map, the color choices, the legend design and supplementary/supporting statistics, the mapping approach has not changes and almost exclusively only choropleth maps are used.
+Unlike innovation in spatial statistical methods, interactivity, data collection and record keeping little work has been done to implement different mapping approaches.
+
 ## Cartograms
@@
 The resulting display more accurately communicates the impact of the disease across the population, as recorded by the statistic, at the sacrifice of geographic accuracy.
-
-<!-- There are many alternative views of the world-->
-<!-- Tobler map projections - I think this is not relevant here
-Choropleths may be considered true topological maps however,, if the landmass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D [@HTLWM].
-The amount of distortion is related to the distance covered by the landmass displayed @GAMP.
-World map projections reflect the frequent perspectives used to view the earth.
-Choropleth maps will always be distorted if they cover enough of the globe, just like photographs of the globe from space.
-Choropleth creation requires choosing a map projection that shows a favorable distortion of the geography for presenting the set of spatial information.
-Selecting a display can prevent misinterpretation of global statistics, as global maps face the challenge of equitable displays of landmass on maps [@RSCW].
-If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed, beyond choosing a reasonable projection. 
--->
 
@@
 
-<!-- 
-An intentional goal when creating the 1966 Census population cartogram for Canada was to maintain contiguity, while attempting to keep the actual shape of places.
-The end result was a 'very accurate isodemo-graphic map of Canada'.
-This intentional design goal coincided with the rising interest in urban geography and presentation of social statistics.
-Figure \ref{fig:ggcartograms} a shows a population contiguous cartogram of the United States.
-All states are visible and the shape of the United States overall is still recognizable.
-In contrast, Figure \ref{fig:auscartograms} a shows an Australian contiguous cartogram also based on population.
-The south east is completely distorted, areas with low population are still large very the map, as their initial area was used to determine the area allocated in the contiguous cartogram.
-
-To be able to recognise the significant changes, a reader will usually have to know the initial geography to find the differences in the new cartogram layout [@NAC].
-Tobler's Conformal mapping means to preserve angles locally so that the shapes of small areas on a traditional map and a cartogram would be similar.
-@CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes. 
--->
-
 ### Non-contiguous
@@
 
-<!-- The shapes are valuable for recognition and allows users to orient themselves on the display.
+The shapes are valuable for recognition and allows users to orient themselves on the display.
 Map creators can efficiently communicate with this kind of map by keeping the outlines or particular elements of the original in the new shape [@NISCC].
@@
 These overlaps are also shown in the non-contiguous cartogram of Australia in Figure \ref{fig:auscartograms} b.
--->
 
@@
 FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
-Each area is represented by a square of the same dimensions.
-<!--
-Figure \ref{fig:ggcartograms}d shows a hexagon tilemap representing each state using a single hexagon.
+Each area is represented by a s
 
-Recommended criteria to contrast mapping methods include average cartographic error, and maximum cartographic error, correct adjacency, maximum aspect ratio, and suitable relative positions [@ORC].
-However, this does not consider the issues with actually producing rectangular cartograms.
-Algorithms for the creation of rectangular cartograms
--->
-
-## tilemap
+## Hexagon tilemap
 <!-- 173 words -->
@@
 The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off to be made between boundary complexity and simplicity.
+quare of the same dimensions.
+
+Figure \ref{fig:ggcartograms}d shows a hexagon tilemap representing each state using a single hexagon.
 
@@
 Virtually any plot display can be fit into the tile representing the areas, allowing displays of multiple variables or values per geographic entity.
-<!-- As the amount of areas increase care must be taken to ensure legibility of the displays.
-@RPIG acknowledges that this display suffers when the areas have very irregular shapes and large size disparities CORRECT reference is Cheng's thesis-->
+As the amount of areas increase care must be taken to ensure legibility of the displays.
+@IVMVTSAD acknowledges that this display suffers when the areas have very irregular shapes and large size disparities.
 Manual creation of the grid is required.
@@
 Areas with uncertain values will flicker more dramatically than areas with more certain values.
-
-<!-- Solutions have been provided for bi-variate displays for movement data, containing Origin - Destination matrices and flow map representations [@MMTVCG]. 
-This method makes the maps and matrix connections clear, and easy to link geographic areas with link lines not crossing. Bar charts can also be added to these to show group totals.
--->
 

### 2019-12-08 Response/Response.Rmd
Response in top level

--- 2019-12-08 version1/Response/Response.Rmd
@@
 *Changes in the text:* 
+Cancer atlases communicate cancer statistics over geographic domains.
+These domains are subdivided by administrative areas such as countries, states or suburbs.
+When communicating human-related statistics in Australia, the geographic map base highlights the sparsely populated rural areas.
+The smaller geographic areas may not draw the attention of readers but they are important to consider if they are more densely populated.
+Alternative map displays can reveal patterns that are not obvious when the same data is shown on a choropleth map.
+Alternative displays have proven to be successful on online news sites, especially for the communication of election data, where votes are aggregated at the state level.
 
@@
 
-*Reply :*
+*Reply 2:* 
+See *Reply 1* for changes to the abstract. Additionally, the conclusion has also been edited to be more direct in concluding the alternatives presented.
 

### 2019-12-08 Response/Response.tex
Response in top level

--- 2019-12-08 version1/Response/Response.tex
@@
 \emph{Changes in the text:}
+Cancer atlases communicate cancer statistics over geographic domains.
+These domains are subdivided by administrative areas such as countries, states or suburbs.
+When communicating human-related statistics in Australia, the geographic map base highlights the sparsely populated rural areas.
+The smaller geographic areas may not draw the attention of readers but they are important to consider if they are more densely populated.
+Alternative map displays can reveal patterns that are not obvious when the same data is shown on a choropleth map.
+Alternative displays have proven to be successful on online news sites, especially for the communication of election data, where votes are aggregated at the state level.
 
@@
 
-\emph{Reply :}
+\emph{Reply 2:}
+See \emph{Reply 1} for changes to the abstract. Additionally, the conclusion has also been edited to be more direct in concluding the alternatives presented.
 

### 2019-12-08 review_article.Rmd
fix referencing

@@
 ```
-<!--
-
-Please note that the submission deadline is September 30, 2019.
-
-Author notes:
-6000 words maximum.
-focused  application  of  literature  to address  a  relevant  clinical  issue:
-Mapping cancer outcomes
-Cartogram applications should point to examples of cancer outcomes, with the main cartogram section discussing sources considered in grey literature review.
-Examples of political areas may be considered if cancer data is aggregated at these levels.
-
-- Future steps should consider applications not currently utilized for cancer outcome displays, where their use may improve decision making
-- needs to have a clear relevance to spatial patterns in cancer epidemiology
-- presents alternative ways of visualizing those spatial patterns, or the uncertainty in those patterns
-- a more applied-focused paper (e.g. lots of cool images), rather than theoretical (e.g. lots of formula).
--->
 
 \section*{Abstract}
-<!-- 119 words -->
-<!--
-Choropleth maps are widely used in cancer atlases to communicate cancer statistics over geographic domains.
-The type of statistic displayed varies immensely, and effective choices can assist in the understanding of impact of cancer on a population.
-Atlases have moved from print media to web format, and the availability of graphical user interfaces, and user interaction and animation, greatly enhance the ability to communicate more information and positively engage readers.
-More use of alternatives to choropleth maps, such as cartograms, could be employed in online atlases, for better representation of spatial distributions, and show statistics on small areas.
-This is especially true for the geography of Australia, where the population is concentrated is small areas along the coastlines.
--->
+<!-- 105 words -->
 
-Cancer atlases communicate cancer statistics over geographic domains. These domains are often subdivided by administrative areas such as countries, states or suburbs.
-When communicating human-related statistics in Australia, the geographic map base may highlight the sparsely populated rural areas.
-The smaller geographic areas may not draw the attention of readers but they may be important to consider if they are more densely populated.
-To
+Cancer atlases communicate cancer statistics over geographic domains.
+These domains are subdivided by administrative areas such as countries, states or suburbs.
+When communicating human-related statistics in Australia, the geographic map base highlights the sparsely populated rural areas.
+The smaller geographic areas may not draw the attention of readers but they are important to consider if they are more densely populated.
+Alternative map displays can reveal patterns that are not obvious when the same data is shown on a choropleth map.
+Alternative displays have proven to be successful on online news sites, especially for the communication of election data, where votes are aggregated at state level.
 
@@
 <!-- What data are we working with, why, what are the purposes -->
-Cancer statistics are usually delivered as an aggregated value for a geopolitical area.
+Cancer statistics are often delivered as an aggregated value for a geopolitical area.
 Presenting these statistics requires transforming individual observations into aggregations of communities as geographical units, in large part for privacy protection or political and policy purposes.
@@
 By providing a visual representation of cancer outcomes, geographic patterns of the disease may be identified and effectively addressed with public health policy and actions.
-@SE recognizes one of the key challenges with mapping spatial patterns of disease is the design of visualizations.
+Daniel Exeter [@SE] recognizes one of the key challenges with mapping spatial patterns of disease is the design of visualizations.
 This paper addresses the visualization techniques and their applications to cancer statistics. Highlighting the differences and historic use of these displays.
@@
 
-# Map displays for disease data {#ch:choropleths}
+# Traditional approaches for cancer map displays {#ch:choropleths}
 <!-- 381 words -->
@@
 Early versions of choropleth maps used symbols or patterns instead of color.
-@CPISACA discuss the use of choropleths to visualize cancer data, and @DMAHP gives an overview of the development of these maps for displaying disease data.
+The use of choropleths to visualize cancer data [@CPISACA], and the development of these maps for displaying disease data [@DMAHP].
 In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality.
@@
 Figure \ref{fig:choroCRS} shows various geographic projections of choropleth maps of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged from 2012 to 2016.
-The data was extracted from the official federal statistics by @USInteractive on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
+The data was extracted from the official federal statistics on cancer incidence and deaths [@USInteractive], produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
 The rates exhibit a spatial trend, increasing from west to east.
@@
 
-# Cancer atlases {#ch:public}
+## Cancer atlases {#ch:public}
 <!-- 360 words -->
@@
 These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analyses.
-Cruickshank (1947) as cited by @DMAHP, discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
+Cruickshank (1947) as cited by S.
+D.
+Walter [@DMAHP], discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
 Overwhelmingly, the cancer maps utilized to communicate to the public and other non-expert audiences are choropleths.
@@
 These can also display supporting statistics such as error, confidence intervals, distributions, sample or population sizes, standard deviation.
-When presenting cancer maps, @MACM believes the intuition derived from maps must be 'validated by rigorous statistical analyses', and the display of the supplementary statistics helps in this regard.
+When presenting cancer maps, the intuition derived from maps must be 'validated by rigorous statistical analyses' [@MACM], and the display of the supplementary statistics helps in this regard.
 
@@
 
-# Alternatives to choropleths {#ch:alternatives}
+# Limitations of choropleth displays
+
+Australia presents an extreme case of an urban-rural divide.
+The landmass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas.
+To present election results on a choropleth map should be 'unthinkable', as it means diminishing the visual impact of the majority of the electorates.
+Dorling [@ACTUC] has a cartogram showing the 1966 general election results, which looked very little like the geographical shape of Australia.
+This issue is felt in any nation which experiences a spatially heterogeneous population distribution.
+As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
+
+Choropleths provide a familiar display.
+This may be less threatening to an audience that could be frightened by statistics, but they suffer primarily from unfaithfully representing the disease information by hiding small geographic areas.
+Inset maps like in Brisbane city in Figure \ref{fig:choropleth-grid-create}c of the state of Queensland, are commonly used to reduce distorted interpretations, but it is a bandaid remedy.
+
+# Contemporary alternatives to choropleths {#ch:alternatives}
 
@@
 When visualizing population statistics, Dorling [@ACTUC] considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
-Howe [@HEDP] suggests that 'cancer occurs in people, not in geographical areas' and @CTTMB believe that spatial socio-economic data, like cancer rates, are best presented on a cartogram for urban areas as the population map base avoids allocating 'undue prominence' to rural areas.
+Howe [@HEDP] suggests that 'cancer occurs in people, not in geographical areas' and that spatial socio-economic data, like cancer rates, are best presented on a cartogram for urban areas as the population map base avoids allocating 'undue prominence' to rural areas [@CTTMB].
 <!-- @MTMSIH encourage the use of cartograms to highlight small areas and uncover local-level inequalities. DI SAYS I think that focuses on bayesian analysis not on cartograms-->
@@
 This transformation often occurs at the expense of the shape of areas [@CBATCC, @NAC, @TAAM].
-From a computer graphics perspective, @ACA explain the application of 'map deformation' to account for the value assigned to each area, they provide three methods for creating value-by-area cartograms.
-Examples include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [-@ACTUC], Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and Tikunov’s Line Integral Method, Constraint-Based Method (Kocmoud and House) [-@CBATCC].
+Three methods are provided for creating value-by-area cartograms [@ACA]. They implement 'map deformation' to account for the value assigned to each area.
+Other methods include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [-@ACTUC], Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and Tikunov’s Line Integral Method, Constraint-Based Method (Kocmoud and House) [-@CBATCC].
 
@@
 Circular cartograms use the same simple shapes for every region represented and resize the shapes according to the statistic represented or the population for a base map.
-This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms, as the 'nonsense' shapes shown in contiguous cartograms have 'no meaning' after distortions are applied @NISCC.
+This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms, as the 'nonsense' shapes shown in contiguous cartograms have 'no meaning' after distortions are applied [@NISCC].
 To produce a compelling map, a gravity model is applied to avoid overlaps and keep spatial relationships with neighboring areas over many iterations.
@@
 
-# Comparison and critique {#ch:compare}
+# Limitations of alternative displays {#ch:compare}
 <!-- 408 words -->
@@
 
-Australia presents an extreme case of an urban-rural divide.
-The landmass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas.
-To present election results on a choropleth map should be 'unthinkable', as it means diminishing the visual impact of the majority of the electorates.
-Dorling [@ACTUC] has a cartogram showing the 1966 general election results, which looked very little like the geographical shape of Australia.
-This issue is felt in any nation which experiences a spatially heterogeneous population distribution.
-As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
-
 Figure \ref{fig:auscartograms} shows four main types of cartograms displaying melanoma incidence on SA3 areas in Australia.
@@
 
-Choropleths provide a familiar display.
-This may be less threatening to an audience that could be frightened by statistics, but they suffer primarily from unfaithfully representing the disease information by hiding small geographic areas.
-Inset maps like in Brisbane city in Figure \ref{fig:choropleth-grid-create}c of the state of Queensland, are commonly used to reduce distorted interpretations, but it is a bandaid remedy.
-
 Cartograms provide the spatial distortion to more accurately convey the statistical distribution, focuses on the human impact of the disease.
@@
 Table \ref{tab:methods} summarizes the key criteria for evaluating maps and alternative displays.
-Suggestions and comments to help map creators best communicate their health data and spatial analyses are provided in @CPISACA and @SAMGIS.
+Suggestions and comments to help map creators best communicate their health data and spatial analyses are provided by Moore and Carpenter [@SAMGIS], Bell et al.
+[@CPISACA].
 
@@
 This can allow inspection of the data from many views [@DQBCM].
-@TNTEA says user interaction with maps helps them to understand and interpret the spatial distribution of disease, to validate, explain or explore the presented statistics and their relationships to each other.
+User interaction with maps helps to understand and interpret the spatial distribution of disease, to validate, explain or explore the presented statistics and their relationships to each other [@TNTEA].
 
@@
 Interactive design features, found in online cancer maps, include tooltips, drop-down menus, data selection, zooming, and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
-The use of these supports can be found in various online cancer maps identified by @roberts2019communication.
+The use of these supports can be found in various online cancer maps [@roberts2019communication].
 
 Animation, in contrast to interactivity, usually involves pre-computing views and showing these in a sequence.
-@TGA provided an overview of animation in association with the R package `gganimate` [@gganimate], with some focus on maps, explaining transitions between frames in a linear, to provide a fixed narrative to passive users.
+An overview of animation in association with the R package `gganimate` [@gganimate], with some focus on maps, explaining transitions between frames in a linear, to provide a fixed narrative to passive users [@TGA].
 Animations are used to communicate a message by capturing and directing users' attention.
@@
 
-@CPISACA provide weather maps as a thoroughly developed example of animation of spatial displays to communicate information to the general public.
+Weather maps are a thoroughly developed example of animation of spatial displays to communicate information to the general public [@CPISACA].
 The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period.

### 2019-12-09 review_article.Rmd
language editing

@@
 Cancer atlases communicate cancer statistics over geographic domains.
-These domains are subdivided by administrative areas such as countries, states or suburbs.
-When communicating human-related statistics in Australia, the geographic map base highlights the sparsely populated rural areas.
+These domains are subdivided into administrative regions such as countries, states or suburbs.
+When communicating human-related statistics in Australia, the geographic map base draws attention to the sparsely populated rural areas.
 The smaller geographic areas may not draw the attention of readers but they are important to consider if they are more densely populated.
@@
 <!-- What data are we working with, why, what are the purposes -->
-Cancer statistics are often delivered as an aggregated value for a geopolitical area.
-Presenting these statistics requires transforming individual observations into aggregations of communities as geographical units, in large part for privacy protection or political and policy purposes.
-The information could be as simple as counts per area (e.g.
+Cancer statistics are often delivered as aggregated values for geopolitical areas.
+Presenting these statistics requires aggregating individual observations into statistics for the geographical units, especially for privacy protection but also for political and policy purposes.
+Counts per area require minimal transformation (e.g.
 state, province, local government area, post/zip code).
-Counts alone are not sufficient to compare areas, because the populations of areas are all different.
-In this case, the count data needs to be merged with population data to appropriately calibrate it to incidence, for example, the rate per 100000 people.
+However, counts alone are not sufficient for the comparison areas, because the populations of areas are all different.
+The count data needs to be merged with population data to present incidence rates, for example, the rate per 100000 people.
 This type of data is collected on a routine basis for public health purposes and may be made available to the general public as a service to the community.
-The task, then, is to examine what are the usual ways to communicate cancer statistics, to the public, are there alternative approaches, and what are the pros and cons of these choices.
 
 <!-- Ways to visualize, whats our hypothesis -->
-A common approach to communicate cancer statistics is to display statistics on a map.
-Using a choropleth map: the statistic is mapped to color and the geographic region is filled with this color.
-The viewer would then be able to examine the spatial distribution of the disease incidence, where there is a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease.
+Map displays are commonly required for the communication of cancer statistics over geographic domains.
+When using a choropleth map: the statistic is mapped to color and the geographic region is filled with this color.
+Disease maps help to present geographic patterns that may be overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
+Providing a visual representation of cancer outcomes allows geographic patterns of the disease to be identified and effectively addressed with public health policy and actions.
+The spatial distribution of the disease incidence can be examined using a choropleth and may reveal a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease.
 Visualizing diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses.
-Disease maps help to present geographic patterns that may be overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
-By providing a visual representation of cancer outcomes, geographic patterns of the disease may be identified and effectively addressed with public health policy and actions.
 Daniel Exeter [@SE] recognizes one of the key challenges with mapping spatial patterns of disease is the design of visualizations.
-This paper addresses the visualization techniques and their applications to cancer statistics. Highlighting the differences and historic use of these displays.
+This paper considers the current visualization techniques to communicate statistics to the public, and their applications to cancer statistics.
+Alternative approaches are also considered, and the limitations of the visualization methods, highlighting the differences and historic use of these displays.
 
@@
 The paper is organized as follows.
-The next section describes the choropleth map which is the common approach to disease maps.
-Section \ref{ch:public} surveys atlases in use today.
+The next section describes the choropleth map which is the common approach to disease maps and presents examples of atlases in use today.
+The limitations of the choropleth map are discussed in Section \ref{ch:chorolimit}.
 Section \ref{ch:alternatives} describes alternative displays, including the cartogram which is useful when the map has heterogeneously sized geographic units.
-The pros and cons of these approaches are discussed in Section \ref{ch:compare}.
+Section \ref{ch:altlimits} presents the limitaions of producing and using alternative displays.
 Disease maps are more useful when made interactive, and common options are described in Section \ref{ch:interacting}, along with a discussion of benefits and disadvantages.
-The last section summarizes the paper.
 
@@
 <!-- what is a choropleth map -->
-A choropleth map is used to display differences in the geographical distribution of data by spatial unit by shading areas of a map [@EI].
-The geography is faithfully rendered, and the color rendering is designed to reveal spatial patterns among data values.
-Identifying and explaining spatial structures, patterns, and processes involve considering the individuals in communities and organizing communities into representable units [@SAMGIS].
+A choropleth map displays the geographical distribution of data over a set of spatial units by shading areas of a map [@EI].
+The geography of a choropleth is faithfully rendered, and the color rendering is designed to reveal spatial patterns among data values.
+Identifying and explaining spatial structures, patterns, and processes involve considering the individuals and organizing them into representable units of communities [@SAMGIS].
 Early versions of choropleth maps used symbols or patterns instead of color.
-The use of choropleths to visualize cancer data [@CPISACA], and the development of these maps for displaying disease data [@DMAHP].
+Choropleths can be used for displaying disease data [@DMAHP], including cancer data [@CPISACA].
 In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality.
@@
 There is also a spatial outlier -- Utah has a noticeably lower rate than its neighbors.
-Also, Kentucky has a noticeably high rate, and Maine also has a higher rate than its neighbors.
-There is something of a cluster of higher rates around the tobacco states.
+Also, Kentucky and Maine have higher rates than their neighbors.
+There is possibly a cluster of higher rates around the tobacco states.
 
 <!-- psychology -->
-Utilizing the state boundaries can make a map familiar to read [@CIBMUK], and allows viewers to visually infer the spatial relationships in the data. 
-The familiarity of the geography is a worthy consideration when presenting results of spatial analysis, to consumers that extend beyond researchers, to the public, policymakers and the media [@CPISACA].
+Displaying familiar state boundaries can make a map simpler to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data using their mental model of the geography.
+Consumers of disease displays may include researchers, the general public, policymakers and the media [@CPISACA], the familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
 
 <!-- potential problems -->
-While the areas are recognizable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time.
-The different population and geographical sizes of administrative areas can attract attention to the shades of the underpopulated but large areas [@EI], Skowronnek [-@BCM] calls this an area-size bias.
+The familiar shapes are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time. 
+The different population and geographical sizes of administrative areas can attract attention to the shades of the underpopulated but large areas [@EI], Skowronnek [@BCM] calls this an area-size bias.
 Choropleths can inhibit visual inference when presenting human-related statistics as the display may draw attention away from the 'potentially more important results in the more populous communities' that are geographically smaller [@SE].
@@
 Atlases are key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work-related exposures, and high-risk diets [@MACM].
-The data collection of statistics across regions and the administrative control within regions lends itself to choropleth visualization.
-Atlases date to Haviland's maps in 1875, with UK cancer atlases appearing in 1961 [@HEDP], and early work in US cancer atlases appearing in 1971 [@burbank].
+The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization.
+Cancer maps and atlases date back to Haviland's maps in 1875, and early work in US cancer atlases appearing in 1971 [@burbank].
 The presentation of cancer statistics has increased with greater access to computational power, and the availability of geographic information systems software [@SE].
@@
 <!-- Measures reported in cancer atlases -->
+Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the general public and others not trained in statistical analyses.
+These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analyses.
+Cruickshank (1947) as cited by S.
+D.
+Walter [@DMAHP], discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
+Overwhelmingly, the cancer maps utilized to communicate to the public and other non-expert audiences are choropleths.
+
+The statistics presented have been developed over time.
 Mortality rates are now often presented as relative rates of risk across the population, and age-adjusted to correct for the higher prevalence of cancers in older populations.
-Howe [-@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s. Table \ref{tab:measures} summarizes the measures commonly presented in published cancer atlases.
+Howe [@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s. Table \ref{tab:measures} summarizes the measures commonly presented in published cancer atlases.
 
@@
 
-<!-- Cancer atlas literature -->
-
-Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the general public and others not trained in statistical analyses.
-These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analyses.
-Cruickshank (1947) as cited by S.
-D.
-Walter [@DMAHP], discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
-Overwhelmingly, the cancer maps utilized to communicate to the public and other non-expert audiences are choropleths.
-
-<!-- written below The Atlas of Cancer in Queensland [@DACQ] focused on highlighting the difference in experience for those living in rural and disadvantaged areas, the Standardized Incidence Ratios (SIR) were modeled via Bayesian methods.
-The presentation of these rates considers not only the shapes of the areas, but also appropriate choices of color blind friendly color schemes, and categories of the values to highlight significantly different areas.
--->
-
 ## Overview of publicly available atlases \label{sec:publicatlases}
 <!-- 416 words -->
-Roberts [@roberts2019communication] identified 33 publicly available cancer atlases, published between January 2010 and November 2015, from different countries around the globe.
-All of these use choropleth maps.
+Roberts [@roberts2019communication] identified 33 publicly available cancer atlases, published between 2010 and 2018, from different countries around the globe.
+All of these online atlases used choropleth maps.
 All but one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships.
 Figure \ref{fig:choropleth-grid-create} displays a subset of maps from these atlases.
-
-<!-- While there is significant variation across these maps in terms of the measure represented, the scale of the map, the color choices, the legend design and supplementary/supporting statistics, the mapping approach has not changes and almost exclusively only choropleth maps are used.
-Unlike innovation in spatial statistical methods, interactivity, data collection and record keeping little work has been done to implement different mapping approaches -->
 
@@
 
-With any number comes some uncertainty about its value.
-This can occur for several reasons: numbers are drawn from samples, errors occur in the disease reporting, values are simulated from a model to maintain privacy.
-Close to half of the atlases identified (42%, n=14) included some measure of uncertainty.
+Uncertainty about the value of a statistic is often communicated in cancer atlases.
+Uncertainty may occur for several reasons: numbers are drawn from samples, errors occur in the disease reporting, or values are simulated from a model to maintain privacy.
 Uncertainty is often recorded as sample or population size, standard deviation, confidence intervals, statistical significance, quantiles, credible intervals.
@@
 
-# Limitations of choropleth displays
+# Limitations of choropleth displays {#ch:chorolimit}
 
-Australia presents an extreme case of an urban-rural divide.
+A choropleth display may be limited by the administrative boundaries used to define regions.
+This issue is often discussed in the context of election mapping, where each electorate decides the outcome of one parliamentary seat in the house or representatives.
+While all states are visible in a map display of Australia, smaller regions such as electorates present an extreme case of an urban-rural divide. 
 The landmass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas.
 To present election results on a choropleth map should be 'unthinkable', as it means diminishing the visual impact of the majority of the electorates.
-Dorling [@ACTUC] has a cartogram showing the 1966 general election results, which looked very little like the geographical shape of Australia.
-This issue is felt in any nation which experiences a spatially heterogeneous population distribution.
-As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
+Using a map of Australia as a visual statistic would communicate an incorrect distrbution of votes across the country.
+The issue of spatially heterogeneous population distribution across a country's landmass is common to many countries, especially British colonies [@ACTUC].
 
@@
 
+ As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
 # Contemporary alternatives to choropleths {#ch:alternatives}
@@
 The purposeful distortion of the map space, transformed according to population density, is beneficial when a uniform density of the map base is desired.
-The population then becomes a uniformly distributed background for the statistic presented [@GOINO].
-"population distribution is often extremely uneven in former British colonies" [@ACTUC ], and this makes the distortion necessary [@CTTMB].
+The population then becomes a uniformly distributed background for the statistic presented [@GOINO], "population distribution is often extremely uneven in former British colonies" [@ACTUC], and this makes the distortion necessary [@CTTMB].
 When implementing a distortion of the geographical shape according to population, the resulting display is an area cartogram [@NAC], or population-by-area cartogram [@TAAM].
@@
 Three methods are provided for creating value-by-area cartograms [@ACA]. They implement 'map deformation' to account for the value assigned to each area.
-Other methods include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [-@ACTUC], Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and Tikunov’s Line Integral Method, Constraint-Based Method (Kocmoud and House) [-@CBATCC].
+Other methods include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [@ACTUC], Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and Tikunov’s Line Integral Method, Constraint-Based Method (Kocmoud and House) [@CBATCC].
 
@@
 However, if the population density is highly dissonant with geographic density the cartogram will lose all spatial context, and thus futile.
+Dorling [@ACTUC] has a cartogram showing the 1966 general election results, which looked very little like the geographical shape of Australia.
 
@@
 
-The Australian Cancer Atlas [-@TACA] provides [tours](
+The Australian Cancer Atlas [@TACA] provides [tours](
 https://atlas.cancer.org.au/app/tour/lungcancer) that change the display to the draw users' attention to areas on the map that are relevant to the story.
@@
 User interaction more allow clicks to progress between stages of more complex animations.
-The Australian Cancer Atlas [-@TACA] provides [tours](
+The Australian Cancer Atlas [@TACA] provides [tours](
 https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.

### 2019-12-09 Response/Response.Rmd
response letter

@@
 fromphone: +61 433 699 797
+fromsignature: sigfile.png
 toname: Professor Peter Baade and Dr Susanna Cramb
-opening: Dear Professor Peter Baade and Dr Susanna Cramb
+opening: Dear Professor Peter Baade and Dr Susanna Cramb,
 closing: Warm regards
@@
 
-Current approaches
-Limitations of current approaches
-Alternative visualisation methods
-Challenges of alternative approaches
+**Traditional approaches for cancer map displays:** This section discusses the choropleth map as the common map display used in cancer maps and atlases.
+*Limitations of current approaches:* This section presents the possible issues with displaying cancer statistics on a choropleth map.
+*Alternative visualisation methods:* We have shown that there are many alternative displays that highlight population dense regions.
+*Challenges of alternative approaches:* Alternative displays also have limitations when displaying spatial distributions of cancer statistics.
 
@@
 
-**Editor suggestion:**
-**The authors mention a focus on Atlases published between 2010 and 2015, but then include a bowel cancer atlas from 2016. **
+*Editor suggestion:*
+*The authors mention a focus on Atlases published between 2010 and 2015, but then include a bowel cancer atlas from 2016. *
 
-**Also, within the bowel cancer atlas data, the reporting of percent of males aged 50-54 with bowel cancer in 2016 is incorrect; it should be the % of the ERP in each area that are males aged 50-54 years.
-**
+*Also, within the bowel cancer atlas data, the reporting of percent of males aged 50-54 with bowel cancer in 2016 is incorrect; it should be the % of the ERP in each area that are males aged 50-54 years.
+*
 
-We have corrected the description of the statistic used in the bowel cancer atlas of Australia:
+We have changed the image used from the Bowel Cancer Atlas. To reflect this change the description of the statistic used in the bowel cancer atlas of Australia ahs been corrected:
 
 Bowel Cancer Australia Atlas:
-...shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia  .
-It is published by *[Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)*.
+...shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia.
+It is published by Bowel Cancer Australia https://www.bowelcanceraustralia.org/.
 
-**Reviewer A: **
-**Major**
+*Reviewer A: *
+*Major*
 *Comment 1: *
@@
 *Reply 1:* 
-We agree with the reviewer that alternative approaches have been fantastic when used in political contexts. Choropleth maps have often been used.
+We agree with the reviewer. Alternative mapping displays have been implemented in political contexts. Choropleth maps have often been used.
 

### 2019-12-09 review_article.Rmd
response letter

@@
 Section \ref{ch:alternatives} describes alternative displays, including the cartogram which is useful when the map has heterogeneously sized geographic units.
-Section \ref{ch:altlimits} presents the limitaions of producing and using alternative displays.
+Section \ref{sec:altlimits} presents the limitaions of producing and using alternative displays.
 Disease maps are more useful when made interactive, and common options are described in Section \ref{ch:interacting}, along with a discussion of benefits and disadvantages.
@@
 
-## Overview of publicly available atlases \label{sec:publicatlases}
 <!-- 416 words -->
@@
 <!-- 479 words -->
-Maps cannot tell the entire story.
 Supplementary graphs and plots are often included to add more depth and information.
@@
  As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
+
 # Contemporary alternatives to choropleths {#ch:alternatives}
@@
 
-# Limitations of alternative displays {#ch:compare}
+# Limitations of alternative displays {#sec:altlimits}
 <!-- 408 words -->
@@
 Many publicly available cancer atlases employ interactivity and/or animation.
-Figure 6 <!--\ref{fig:interacting}--> shows the interactive GUI tools provided with the public atlases discussed in Section \ref{sec:publicatlases}.
+Figure 6 <!--\ref{fig:interacting}--> shows the interactive GUI tools provided with the public atlases.
 Mostly these are menus, and toggles, for choosing indicators, and subsets of the population.

### 2019-12-11 review_article.Rmd
Selection of Atlases section restructured

@@
 ---
-title: "Cartogram Mapping and its Application to Cancer Data Visualization"
+title: "Cancer Applications of Choropleth maps, and the Potential of Cartograms and Alternative Map Displays"
 author: "Stephanie Kobakian, Jessie Roberts and Dianne Cook"
@@
 Section \ref{ch:alternatives} describes alternative displays, including the cartogram which is useful when the map has heterogeneously sized geographic units.
-Section \ref{sec:altlimits} presents the limitaions of producing and using alternative displays.
+Section \ref{sec:compare} presents the limitaions of producing and using alternative displays.
 Disease maps are more useful when made interactive, and common options are described in Section \ref{ch:interacting}, along with a discussion of benefits and disadvantages.
@@
 
-```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Publicly available choropleth cancer maps published between 2010 and 2015.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Publicly available choropleth cancer maps..", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
 library(cowplot)
@@
 
-### The Environment and Health Atlas of England and Wales
+The selection varies in the geographies explored, Figure \ref{fig:choropleth-grid-create}b, Globocan 2012 [@Globocan] explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide utilising data sourced from cancer registries of each country.
+The Bowel Cancer Australia Atlas presents an example of a cancer specific atlas, Figure \ref{fig:choropleth-grid-create}d shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia [@Bowel].
+The choice of gender can be selected in the Bowel Cancer Atlas, an alternative approach in employed in the Map of Cancer Mortality Rates in Spain, Figure \ref{fig:choropleth-grid-create}f shows side by side maps of the relative risk of lung cancer separately for men and women [@cancerSpain].
 
-Figure \ref{fig:choropleth-grid-create}a shows the relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2].
-The cancer data was from the Office for National Statistics (ONS) (England) and the Welsh Cancer Intelligence and Surveillance Unit (WCISU).
+There are vast differences in the geographies considered by online atlases. 
+Figure \ref{fig:choropleth-grid-create}b shows global age-standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018.
+The United States Cancer Statistics [@USInteractive] can be seen in Figure \ref{fig:choropleth-grid-create}e.
+This map contains the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level.
+The Environment and Health Atlas of England and Wales, Figure \ref{fig:choropleth-grid-create}a shows the relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2] at a neighbourhood (small-area) scale.
+The Atlas of Cancer in Queensland, Figure \ref{fig:choropleth-grid-create}c shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 in the state of Queensland within Australia [@QLDcancerAtlas].
 
-### Globocan 2012: Estimated Cancer Incidence, Mortality and Prevalence Worldwide
+Age specific atlases are less common, Figure \ref{fig:choropleth-grid-create}g displays Atlas of Childhood Cancer in Ontario, this communicates the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
+This atlas communicates information regarding a very small geographic region.
 
-The map seen in Figure \ref{fig:choropleth-grid-create}b shows age-standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018 [@Globocan].
-It is published by the [World Health Organization's International Agency for Research on Cancer](https://www.iarc.fr/).
-Data was sourced from cancer registries of each country.
-
-### Atlas of Cancer in Queensland
-
-Figure \ref{fig:choropleth-grid-create}c shows the relative incidence ratio of lung cancer in males in the state of QLD within Australia based on data from 1998 to 2007 [@QLDcancerAtlas].
-It is published by the Queensland Cancer Council, using data from the *[Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/)*.
-
-### Bowel Cancer Australia Atlas 
-Figure \ref{fig:choropleth-grid-create}d shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia [@Bowel].
-It is published by *[Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)*.
-
-### United States Cancer Statistics: An Interactive Cancer Statistics Website
-
-The *[United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/)* [@USInteractive] can be seen in Figure \ref{fig:choropleth-grid-create}e.
-This map contains  the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level.
-The map was published by the *[Centers for Disease Control and Prevention](https://www.cdc.gov/)*, with data from state cancer registries.
-
-### Map of Cancer Mortality Rates in Spain
-
-Figure \ref{fig:choropleth-grid-create}f shows side by side maps of the relative risk of lung cancer separately for men and women from 2004 to 2008 [@cancerSpain].
-
-### Atlas of Childhood Cancer in Ontario
-
-Figure \ref{fig:choropleth-grid-create}g  displays the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
-
-## Additional considerations
+## Additional materials
 <!-- 479 words -->
@@
 To present election results on a choropleth map should be 'unthinkable', as it means diminishing the visual impact of the majority of the electorates.
-Using a map of Australia as a visual statistic would communicate an incorrect distrbution of votes across the country.
+Using a map of Australia as a visual statistic would communicate an incorrect distribution of votes across the country.
 The issue of spatially heterogeneous population distribution across a country's landmass is common to many countries, especially British colonies [@ACTUC].
@@
 This may be less threatening to an audience that could be frightened by statistics, but they suffer primarily from unfaithfully representing the disease information by hiding small geographic areas.
-Inset maps like in Brisbane city in Figure \ref{fig:choropleth-grid-create}c of the state of Queensland, are commonly used to reduce distorted interpretations, but it is a bandaid remedy.
+Inset maps like in Brisbane city in Figure \ref{fig:choropleth-grid-create}c of the state of Queensland, are commonly used to reduce distorted interpretations, but it is a band aid remedy.
 
@@
 Monmonier [@HTLWM] suggests that white lies may be employed to create useful displays and map creators can draw lines that may distort the geometry and suppress features and it is easy for the average person to disregard the impact of transformations used to create cartograms.
-Cartograms may be seen as an extension of map transformations and projections. The favorable distortion is proportional to a valueother than the actual earth size area [@NAC].
+Cartograms may be seen as an extension of map transformations and projections. The favorable distortion is proportional to a value other than the actual earth size area [@NAC].
 A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, interpreting spatial patterns very difficult.
@@
 Howe [@HEDP] suggests that 'cancer occurs in people, not in geographical areas' and that spatial socio-economic data, like cancer rates, are best presented on a cartogram for urban areas as the population map base avoids allocating 'undue prominence' to rural areas [@CTTMB].
-<!-- @MTMSIH encourage the use of cartograms to highlight small areas and uncover local-level inequalities. DI SAYS I think that focuses on bayesian analysis not on cartograms-->
+<!-- @MTMSIH encourage the use of cartograms to highlight small areas and uncover local-level inequalities. DI SAYS I think that focuses on Bayesian analysis not on cartograms-->
 
@@
 The neighboring states are easily contrasted however the North-East regions have been displaced from their geographic location.
-The sparsity of the population in the Mid West is highlighteded by the light yellow color, the Age-Adjusted rate in Kentucky is the darkest and its neighbors are similar.|
+The sparsity of the population in the Mid West is highlighted by the light yellow color, the Age-Adjusted rate in Kentucky is the darkest and its neighbors are similar.|
 
@@
 
-## tilemap
+## tile map
 <!-- 173 words -->
@@
 A simple tilemap presents the areas in a tessellated grid display, where geographic neighbors are found next to each other, with some necessary displacement employed for regions with more than four neighbors.
-These tiles may be labeled or colored to represent a value.
-tilemaps may be difficult to create, they are best created manually, with additional time and care required as the number of geographic areas to include increases.
+These tiles may be labelled or colored to represent a value.
+tile maps may be difficult to create, they are best created manually, with additional time and care required as the number of geographic areas to include increases.
 Cano and others [@MDAC] define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area can be used to communicate the statistic of regions. 
@@
 Hafen [@IGF] formalizes the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas. 
-Like tilemaps, the arrangement of the areas can be reused for other spatial distribution visualizations.
+Like tile maps, the arrangement of the areas can be reused for other spatial distribution visualizations.
 Geofaceting has the functionality of facets, often used to replicate visualizations for each subset of the data. In this case, the data subsets are geographic.
@@
 
-Pickle and others [@MMST] present linked micromap plots to visually link geographic and statistical data, this serves as a solution to  multi-dimensionality issues.
+Pickle and others [@MMST] present linked micro map plots to visually link geographic and statistical data, this serves as a solution to  multi-dimensionality issues.
 This design has been used in online atlas displays of US states.
@@
 it shows one map per group of the key separating variable, in a row with each additional statistical plot.
-Linked micromaps predominantly utilize the choropleth map for displays of spatial relationships, that are seen when spatial neighbors are alotted to the same group.
+Linked micro maps predominantly utilize the choropleth map for displays of spatial relationships, that are seen when spatial neighbors are allotted to the same group.
 It is one of several alternative displays that allow maps to become bivariate displays, commonly used to present both an estimate and the associated uncertainty.
@@
 
-# Limitations of alternative displays {#sec:altlimits}
+# Comparison and critique of alternative displays {#sec:compare}
 <!-- 408 words -->
@@
 However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a color.
-These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighteded by the inset map.
+These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
 Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbors. The map is reusable and fit for many purposes.
@@
 
-|                     | Choropleth | Contiguous | Non-contig | Dorling | tilemaps | Geofacets |
+|                     | Choropleth | Contiguous | Non-contig | Dorling | tile maps | Geofacets |
 |:-----------|:------------:|:-----------:|:---------------:|:---------:|:-----------:|:----------:|
@@
 - variable selection, 
-- tooltips, 
+- tool tips, 
 - zooming and panning.
@@
 Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
-Effective  user-centered interactive actions produce rapid, incremental, and reversible changes to the display [@DMIV].
+Effective  user-centred interactive actions produce rapid, incremental, and reversible changes to the display [@DMIV].
 
@@
 Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display.
-Interactive design features, found in online cancer maps, include tooltips, drop-down menus, data selection, zooming, and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
+Interactive design features, found in online cancer maps, include tool tips, drop-down menus, data selection, zooming, and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
 The use of these supports can be found in various online cancer maps [@roberts2019communication].

### 2019-12-12 review_article.Rmd
editing to camparisons of alternative displays

@@
 fontsize: 11pt
-csl: vancouver.csl
+csl: springer-vancouver.csl
 yolo: yes
@@
 Cancer atlases communicate cancer statistics over geographic domains.
-These domains are subdivided into administrative regions such as countries, states or suburbs.
-When communicating human-related statistics in Australia, the geographic map base draws attention to the sparsely populated rural areas.
-The smaller geographic areas may not draw the attention of readers but they are important to consider if they are more densely populated.
-Alternative map displays can reveal patterns that are not obvious when the same data is shown on a choropleth map.
-Alternative displays have proven to be successful on online news sites, especially for the communication of election data, where votes are aggregated at state level.
+They subdivide these domains into administrative regions such as countries, states, or suburbs.
+When communicating human-related statistics in Australia, the geographic map base draws attention to sparsely populated rural areas.
+The smaller geographic areas may not draw the attention of readers, but they are important to consider if they are more densely populated.
+Alternative map displays can be more effective than a choropleth map at revealing spatial patterns.
+Alternative displays are effective for the communication of election data aggregated at the state or electorate level on online news sites.
 
@@
 <!-- What data are we working with, why, what are the purposes -->
-Cancer statistics are often delivered as aggregated values for geopolitical areas.
-Presenting these statistics requires aggregating individual observations into statistics for the geographical units, especially for privacy protection but also for political and policy purposes.
-Counts per area require minimal transformation (e.g.
-state, province, local government area, post/zip code).
-However, counts alone are not sufficient for the comparison areas, because the populations of areas are all different.
-The count data needs to be merged with population data to present incidence rates, for example, the rate per 100000 people.
-This type of data is collected on a routine basis for public health purposes and may be made available to the general public as a service to the community.
+Researchers, health authorities, governments and not-for-profits are common communicators of cancer statistics.
+They often present statistics to the public as aggregated values for geopolitical areas.
+Presenting these statistics requires aggregating individual observations into the statistics for the geographical units, especially for privacy protection, but also for political and policy purposes.
+Data collection or aggregation regions for the amount of cancer diagnoses or excess deaths per area include states, provinces, local government areas, and post/zip codes.
+To provide the amount, or count, of the diagnoses or excess deaths with these areas, it needs minimal transformation.
+However, the amount of people living in each of the areas in a set will be different.
+Those who wish to calculate rates can use the amount of people living in an area as a denominator and the diagnoses can become a numerator to create rates per amount of people.
+Health registries routinely collect data for public health reasons and may make this available to the public as a service to the community.
 
 <!-- Ways to visualize, whats our hypothesis -->
-Map displays are commonly required for the communication of cancer statistics over geographic domains.
-When using a choropleth map: the statistic is mapped to color and the geographic region is filled with this color.
-Disease maps help to present geographic patterns that may be overlooked in a table, obscuring the geospatially related statistics [@SAMGIS].
-Providing a visual representation of cancer outcomes allows geographic patterns of the disease to be identified and effectively addressed with public health policy and actions.
-The spatial distribution of the disease incidence can be examined using a choropleth and may reveal a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease.
-Visualizing diseases on maps is often the first step in exploratory spatial data analysis and effectively helps in the formulation of hypotheses.
-Daniel Exeter [@SE] recognizes one of the key challenges with mapping spatial patterns of disease is the design of visualizations.
-This paper considers the current visualization techniques to communicate statistics to the public, and their applications to cancer statistics.
-Alternative approaches are also considered, and the limitations of the visualization methods, highlighting the differences and historic use of these displays.
+Choropleth map displays are commonly used for the communication of cancer statistics over geographic domains.
+In this display, map creators map a statistic to color and apply the colour to the polygons representing geographic regions.
+Map users may overlook geographic patterns of disease when obscured in a table [@SAMGIS].
+Visualizing diseases on maps is often the first step in exploratory spatial data analysis and helps in the formulation of hypotheses.
+Providing a visual representation of cancer outcomes allows identification, and public health policy and actions can address these patterns.
+choropleth maps allow the examination of the spatial distribution of the disease incidence, this may reveal a trend in longitude or latitude, differences between rural areas and urban, or coastal and inland areas, or even specific hot spots of the disease.
+Daniel Exeter [@SE] recognizes that one of the key challenges with mapping spatial patterns of disease is the design of visualizations.
+This paper considers the current visualization techniques to communicate statistics to members of the public and their applications to cancer statistics.
+It also considers alternative approaches and compares the visualization methods, highlighting the differences and historic use of these displays.
 
 <!-- Structure of the paper -->
-The paper is organized as follows.
-The next section describes the choropleth map which is the common approach to disease maps and presents examples of atlases in use today.
-The limitations of the choropleth map are discussed in Section \ref{ch:chorolimit}.
-Section \ref{ch:alternatives} describes alternative displays, including the cartogram which is useful when the map has heterogeneously sized geographic units.
-Section \ref{sec:compare} presents the limitaions of producing and using alternative displays.
-Disease maps are more useful when made interactive, and common options are described in Section \ref{ch:interacting}, along with a discussion of benefits and disadvantages.
+The paper is structured as follows.
+The next section describes the choropleth map, which is the common approach to disease maps and presents examples of atlases in use today.
+Section 3 discusses the limitations of the choropleth map.
+Section 4 describes alternative displays, including the cartogram, which is useful when the map has heterogeneously sized geographic units.
+Section 5 presents the limitations of the production and use of alternative displays.
+Disease maps are more useful when made interactive, Section 6 describes common options, along with a discussion of benefits and disadvantages.
 
-# Traditional approaches for cancer map displays {#ch:choropleths}
+# Traditional approaches for cancer map displays {#ch:choropleth maps}
 <!-- 381 words -->
 <!-- what is a choropleth map -->
-A choropleth map displays the geographical distribution of data over a set of spatial units by shading areas of a map [@EI].
-The geography of a choropleth is faithfully rendered, and the color rendering is designed to reveal spatial patterns among data values.
+A choropleth is traditionally used to make a cancer map.
+It is used to show the geographical distribution of data values by shading areas of a map [@EI].
+A choropleth map involves faithfully rendering the geography, and the design of the color scheme and rendering reveal spatial patterns among data values.
 Identifying and explaining spatial structures, patterns, and processes involve considering the individuals and organizing them into representable units of communities [@SAMGIS].
 Early versions of choropleth maps used symbols or patterns instead of color.
-Choropleths can be used for displaying disease data [@DMAHP], including cancer data [@CPISACA].
-In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality.
+choropleth maps can be used for displaying disease data [@DMAHP], including cancer data [@CPISACA].
+In epidemiology, choropleth maps are often used as a tool to study the spatial distribution of cancer incidence and mortality.
 
-Figure \ref{fig:choroCRS} shows various geographic projections of choropleth maps of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged from 2012 to 2016.
-The data was extracted from the official federal statistics on cancer incidence and deaths [@USInteractive], produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
+Figure 2<!--\ref{fig:choroCRS}--> shows various geographic projections of choropleth maps of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged from 2012 to 2016.
+The official federal statistics provide cancer incidence and deaths [5], produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
 The rates exhibit a spatial trend, increasing from west to east.
-There is also a spatial outlier -- Utah has a noticeably lower rate than its neighbors.
+Utah is a spatial outlier, it has a noticeably lower rate than its neighbors.
 Also, Kentucky and Maine have higher rates than their neighbors.
@@
 <!-- psychology -->
-Displaying familiar state boundaries can make a map simpler to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data using their mental model of the geography.
-Consumers of disease displays may include researchers, the general public, policymakers and the media [@CPISACA], the familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
+Displaying familiar state boundaries can make a map easier to read [@CIBMUK] and allow viewers to infer the spatial relationships visually in the data using their mental model of the geography.
+Consumers of disease displays may include researchers, members of the public, policymakers, and the media  [@CPISACA].
+For these consumers, the familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
 
 <!-- potential problems -->
-The familiar shapes are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time. 
+Politically driven boundaries produced the now familiar shapes or regions, this has created individual areas of non-uniform size.
+These areas have different population densities and are subject to change.
 The different population and geographical sizes of administrative areas can attract attention to the shades of the underpopulated but large areas [@EI], Skowronnek [@BCM] calls this an area-size bias.
-Choropleths can inhibit visual inference when presenting human-related statistics as the display may draw attention away from the 'potentially more important results in the more populous communities' that are geographically smaller [@SE].
+choropleth maps can inhibit visual inference when presenting human-related statistics as the display may draw attention away from the 'potentially more important results in the more populous communities' that are geographically smaller [@SE].
 
-```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Publicly available choropleth cancer maps..", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Examples of publicly available choropleth cancer maps.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
 library(cowplot)
@@
 <!-- 360 words -->
-<!-- Displaying of cancer data on choropleths -->
-A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
+<!-- Displaying of cancer data on choropleth maps -->
+A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country or group of countries.
 Atlases are key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work-related exposures, and high-risk diets [@MACM].
 The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization.
-Cancer maps and atlases date back to Haviland's maps in 1875, and early work in US cancer atlases appearing in 1971 [@burbank].
-The presentation of cancer statistics has increased with greater access to computational power, and the availability of geographic information systems software [@SE].
+Cancer maps and atlases dating back to Haviland's maps in 1875, and early work in US cancer atlases appearing in 1971 [@burbank].
+The presentation of cancer statistics has increased with greater access to computational power and the availability of geographic information systems software [@SE].
 
 <!-- Measures reported in cancer atlases -->
-Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the general public and others not trained in statistical analyses.
-These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analyses.
+Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the general public and others not trained in statistical analysis.
+These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analysis.
 Cruickshank (1947) as cited by S.
@@
 Walter [@DMAHP], discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
-Overwhelmingly, the cancer maps utilized to communicate to the public and other non-expert audiences are choropleths.
+Overwhelmingly, the cancer maps used to communicate to members of the public and other non-expert audiences are choropleth maps.
 
-The statistics presented have been developed over time.
-Mortality rates are now often presented as relative rates of risk across the population, and age-adjusted to correct for the higher prevalence of cancers in older populations.
-Howe [@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s. Table \ref{tab:measures} summarizes the measures commonly presented in published cancer atlases.
+Epidemiologists and statisticians have developed these statistics over several decades.
+Mortality rates commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations.
+Howe [@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s.
+Table 1<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases.
 
-Table: \label{tab:measures} Commonly used measures for reporting cancer information.
+Table: 1<!--\label{tab:measures}--> Common measures for reporting cancer information.
 
@@
 Roberts [@roberts2019communication] identified 33 publicly available cancer atlases, published between 2010 and 2018, from different countries around the globe.
-All of these online atlases used choropleth maps.
-All but one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships.
-Figure \ref{fig:choropleth-grid-create} displays a subset of maps from these atlases.
+These online atlases used choropleth maps.
+Non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships published all but one of these.
+Figure 1<!--\ref{fig:choropleth-grid-create}--> displays a subset of maps from these atlases.
 
-The selection varies in the geographies explored, Figure \ref{fig:choropleth-grid-create}b, Globocan 2012 [@Globocan] explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide utilising data sourced from cancer registries of each country.
-The Bowel Cancer Australia Atlas presents an example of a cancer specific atlas, Figure \ref{fig:choropleth-grid-create}d shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia [@Bowel].
-The choice of gender can be selected in the Bowel Cancer Atlas, an alternative approach in employed in the Map of Cancer Mortality Rates in Spain, Figure \ref{fig:choropleth-grid-create}f shows side by side maps of the relative risk of lung cancer separately for men and women [@cancerSpain].
+The selection varies in the geographies explored, Figure 1<!--\ref{fig:choropleth-grid-create}-->b, Globocan 2012 [@Globocan] explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide utilizing data sourced from cancer registries of each country.
+The Bowel Cancer Australia Atlas presents an example of a cancer specific atlas, Figure 1<!--\ref{fig:choropleth-grid-create}-->d shows the average Standardized Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia [@Bowel].
+Like many of the atlases examined, there is a choice of gender displayed in the Bowel Cancer Atlas.
+An alternative approach used in the Map of Cancer Mortality Rates in Spain, Figure 1<!--\ref{fig:choropleth-grid-create}-->f shows side-by-side maps of the relative risk of lung cancer separately for men and women [@cancerSpain].
 
 There are vast differences in the geographies considered by online atlases. 
-Figure \ref{fig:choropleth-grid-create}b shows global age-standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018.
-The United States Cancer Statistics [@USInteractive] can be seen in Figure \ref{fig:choropleth-grid-create}e.
-This map contains the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level.
-The Environment and Health Atlas of England and Wales, Figure \ref{fig:choropleth-grid-create}a shows the relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2] at a neighbourhood (small-area) scale.
-The Atlas of Cancer in Queensland, Figure \ref{fig:choropleth-grid-create}c shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 in the state of Queensland within Australia [@QLDcancerAtlas].
+Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows global age-standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018.
+The United States Cancer Statistics [@USInteractive] presented in Figure 1<!--\ref{fig:choropleth-grid-create}-->e contains the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level.
+The Environment and Health Atlas of England and Wales, Figure 1<!--\ref{fig:choropleth-grid-create}-->a shows the relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2] at a neighborhood (small-area) scale.
+The Atlas of Cancer in Queensland, Figure 1<!--\ref{fig:choropleth-grid-create}-->c shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 in the state of Queensland within Australia [@QLDcancerAtlas].
 
-Age specific atlases are less common, Figure \ref{fig:choropleth-grid-create}g displays Atlas of Childhood Cancer in Ontario, this communicates the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
-This atlas communicates information regarding a very small geographic region.
+Age-specific atlases are less common, Figure 1<!--\ref{fig:choropleth-grid-create}-->g displays Atlas of Childhood Cancer in Ontario, this communicates the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
+This atlas communicates information regarding a tiny geographic region.
 
@@
 <!-- 479 words -->
-Supplementary graphs and plots are often included to add more depth and information.
-Additional materials such as tables, graphs, and text explanations, support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers [@CPISACA].
+Including supplementary graphs and plots add more depth and information.
+Additional materials such as tables, graphs, and text explanations, support the understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers [@CPISACA].
 The many displays of statistical summaries, including dot plots, bar plots, box plots, cumulative distribution plots, scatter plots, normal probability plots, can provide alternative views of the cancer statistics.
 These can also display supporting statistics such as error, confidence intervals, distributions, sample or population sizes, standard deviation.
-When presenting cancer maps, the intuition derived from maps must be 'validated by rigorous statistical analyses' [@MACM], and the display of the supplementary statistics helps in this regard.
+When presenting cancer maps the display of the supplementary statistics helps.
+The intuition derived from maps must be ‘validated by rigorous statistical analyses’ [@MACM].
 
-### Geographic hierarchies
+The statistics communicated in atlases are often used to describe differences between areas.
+This can occur at different levels of aggregation.
+Aggregation of global health statistics occurs within administrative and arbitrarily defined regions, such as those used by the World Health Organization and the United Nations [@IARC_3].
+World atlases can allow for displays of data aggregated into continents, countries, states, provinces and congressional districts [@USInteractive].
+Each population area will probably have a different number of people.
+Cancer atlases may also communicate the distribution of the population living in all areas in a table or histogram display [@NICR_1].
+Atlases can connect the population to the land available to them by communicating population density.
+Section  4<!--\ref{ch:alternatives}--> considers alternatives to choropleth that introduce population into the visualizations.
 
-While atlases are often used to describe differences between areas, statistics may be displayed at different levels of aggregation.
-Global health statistics can be aggregated to administrative and arbitrarily defined regions, such as those used by the World Health Organization and the United Nations [@IARC_3].
-World atlases can allow for displays of data aggregated into continents, countries, states, provinces and congressional districts [@USInteractive].
-
-### Population distribution
-
-Each population area will likely have a different number of people. The distribution of the population residing in all areas may also be communicated in a table or histogram display [@NICR_1].
-Atlases can connect the population to the land available to them by communicating population density. Alternatives to choropleth to address this are described in Section \ref{ch:alternatives} .
-
-### Statistical uncertainty
-
-Uncertainty about the value of a statistic is often communicated in cancer atlases.
-Uncertainty may occur for several reasons: numbers are drawn from samples, errors occur in the disease reporting, or values are simulated from a model to maintain privacy.
-Uncertainty is often recorded as sample or population size, standard deviation, confidence intervals, statistical significance, quantiles, credible intervals.
-The most common measure used to represent uncertainty were credible or confidence intervals (CIs).
-
-Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it can be difficult to display effectively.
+Map users can define the population further using the demographics of the individuals within each area.
+Demographics that can influence cancer rates include information regarding the age and sex distribution of the areas presented.
+Some digital atlases surveyed allow users to control the display to select subsets such as males, females or those aged over 65.
+Similarly, socioeconomic indicators, such as unemployment rates, poverty rates, remoteness, and education levels,  can explain how the experience of cancer prevalence varies for various members of society.
+Few atlases provide this level of detail.
+Introducing population and demographic information helps to interpret the rates in areas effectively, but there will still be uncertainty around the rates.
+To address this, cancer atlases often communicate uncertainty about the value of a statistic.
+Cancer atlas creators may introduce uncertainty when using samples, or there are errors occurring in the disease reporting, or models are used to simulate values to maintain privacy.
+The statistics that report on uncertainty include sample or population size, standard deviation, confidence intervals, statistical significance, quantiles, and  credible intervals.
+The most common measures used to represent uncertainty were credible or confidence intervals (CIs).
+Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it is difficult to display effectively.
 The map focuses on displaying the statistic and lacks additional space to represent the uncertainty.
@@
 
-### Demographics
-
-Demographics include information regarding the age and sex distribution of the areas presented.
-Digital atlases allow for users to control the display, to select subsets such as males, females or those aged over 65.
-
-### Socio-economic indicators
-
-Socio-economic indicators, such as unemployment rates, poverty rates, remoteness, and education levels,  can explain how the experience of cancer prevalence varies for various members of society.
-Few atlases provide this level of detail.
-Data is available from the United Nations Human Development Index and financial indicators are available from the World Bank [@IARC_3].
-
 # Limitations of choropleth displays {#ch:chorolimit}
 
-A choropleth display may be limited by the administrative boundaries used to define regions.
-This issue is often discussed in the context of election mapping, where each electorate decides the outcome of one parliamentary seat in the house or representatives.
-While all states are visible in a map display of Australia, smaller regions such as electorates present an extreme case of an urban-rural divide. 
+The administrative boundaries used to define regions may limit a choropleth display.
+Those who critique choropleth maps often dicuss election mapping, where each electorate decides the outcome of one parliamentary seat in the house of representatives.
+While all states are visible in a map display of Australia, smaller regions such as electorates present an urban-rural divide, where rural regions are much larger than metropolitan areas of similar or greater populations.
 The landmass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas.
-To present election results on a choropleth map should be 'unthinkable', as it means diminishing the visual impact of the majority of the electorates.
+To present election results on a choropleth map should be 'unthinkable',XXX as it means diminishing the visual impact of most of the electorates.
 Using a map of Australia as a visual statistic would communicate an incorrect distribution of votes across the country.
-The issue of spatially heterogeneous population distribution across a country's landmass is common to many countries, especially British colonies [@ACTUC].
 
-Choropleths provide a familiar display.
-This may be less threatening to an audience that could be frightened by statistics, but they suffer primarily from unfaithfully representing the disease information by hiding small geographic areas.
-Inset maps like in Brisbane city in Figure \ref{fig:choropleth-grid-create}c of the state of Queensland, are commonly used to reduce distorted interpretations, but it is a band aid remedy.
-
+Spatially heterogeneous population distribution across a country's landmass is an issue common to many countries, especially British colonies [@ACTUC].
+However, the boundaries shown in choropleth maps provide a familiar display.
+This may be less threatening to an audience that could be unfamiliar was interpreting spatial statistics, even though they suffer primarily from unfaithfully representing the disease information by hiding small geographic areas.
+Inset maps like in Brisbane city in Figure 1<!--\ref{fig:choropleth-grid-create}-->c of the state of Queensland are commonly used to reduce distorted interpretations, but it is a band aid remedy.
 As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
 
-# Contemporary alternatives to choropleths {#ch:alternatives}
+# Contemporary alternatives to choropleth maps {#ch:alternatives}
 
@@
 <!-- What is a cartogram? -->
-Choropleths imply uniformity of data across the geographic space however population densities are extremely unlikely to be uniform [@BCM].
-The cartogram was developed to address this, by transforming the map to better indicate population [@ACCAC].
-The resulting display more accurately communicates the impact of the disease across the population, as recorded by the statistic, at the sacrifice of geographic accuracy.
+choropleth maps imply uniformity of data across the geographic space but population densities are unlikely to be uniform [@BCM].
+Cartographers developed the cartogram to draw the attention to the population by transforming the map [@ACCAC].
+The resulting display can communicate the impact of the disease more accurately across the population, as recorded by the statistic, at the sacrifice of geographic accuracy.
 
@@
 <!-- Tobler map projections - I think this is not relevant here
-Choropleths may be considered true topological maps however,, if the landmass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D [@HTLWM].
+choropleth maps may be considered true topological maps however,, if the landmass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D [@HTLWM].
 The amount of distortion is related to the distance covered by the landmass displayed @GAMP.
@@
 <!-- Denominators -->
-The purposeful distortion of the map space, transformed according to population density, is beneficial when a uniform density of the map base is desired.
-The population then becomes a uniformly distributed background for the statistic presented [@GOINO], "population distribution is often extremely uneven in former British colonies" [@ACTUC], and this makes the distortion necessary [@CTTMB].
-When implementing a distortion of the geographical shape according to population, the resulting display is an area cartogram [@NAC], or population-by-area cartogram [@TAAM].
+When a map creator desires a uniform population density of the map base, the purposeful distortion of the map space is beneficial.
+The "population distribution is often extremely uneven in former British colonies" [@ACTUC], and this makes the distortion necessary, [@CTTMB] population will then become a uniformly distributed background for the statistic presented [@GOINO].
+An area cartogram [@NAC], or population-by-area cartogram [@TAAM] occurs after the distortion of the geographical shape according to population.
 Event cartograms change the area of regions on a map depending on the amount of disease-related events, but this does not consider the effects of land area and population [@VSSDCUC].
@@
 Cartograms provide an alternative visualization method for statistical and geographical information.
-The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape or distance of geographical areas [@ACTUC]. 
-Monmonier [@HTLWM] suggests that white lies may be employed to create useful displays and map creators can draw lines that may distort the geometry and suppress features and it is easy for the average person to disregard the impact of transformations used to create cartograms.
-Cartograms may be seen as an extension of map transformations and projections. The favorable distortion is proportional to a value other than the actual earth size area [@NAC].
-A disadvantage of the conventional map is that sparsely populated rural areas may be emphasized, whereas the areas representing cities are very small, interpreting spatial patterns very difficult.
-The distortion of a cartogram accounts for the population density, preventing it from obscuring the spatial patterns [@TAAM]. 
+The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape, or distance of geographical areas [@ACTUC]. 
+Monmonier [@HTLWM] suggests that map creators can use white lies to create useful displays by distorting the geometry and suppressing features.
+It is easy for the average person to disregard the impact of transformations used to create cartograms.
+One disadvantage of the conventional map is interpreting spatial patterns is difficult because of the sparsely populated rural areas being emphasized in a choropleth map, and the inner city areas are geographically small.
+The distortion of each area when making a cartogram is proportional to a value other than the actual earth size area [@NAC].
+The distortion in an area cartogram accounts for the population density, preventing it from obscuring the spatial patterns [@TAAM].
 The spatial transformation of map regions relative to the data emphasizes the data distribution instead of land size [@CBATCC].
-When visualizing population statistics, Dorling [@ACTUC] considers this equitable representation design 'more socially just', or honest [@NISCC], giving due attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
+When visualizing population statistics, Dorling [@ACTUC] considers this design 'more socially just', or honest [@NISCC], giving equitable representation and attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
 Howe [@HEDP] suggests that 'cancer occurs in people, not in geographical areas' and that spatial socio-economic data, like cancer rates, are best presented on a cartogram for urban areas as the population map base avoids allocating 'undue prominence' to rural areas [@CTTMB].
-<!-- @MTMSIH encourage the use of cartograms to highlight small areas and uncover local-level inequalities. DI SAYS I think that focuses on Bayesian analysis not on cartograms-->
 
@@
 The creation of cartograms was historically in the hands of professional cartographers [@CD]. 
-Early approaches including John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball-bearing approach and Tobler's (1973) computer programs [@ACTUC].
-Geographical information systems allowed map users, and researchers to create cartograms, but these systems are utilized depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994), [@CD].
+Early approaches by John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball-bearing approach and Tobler's (1973) computer programs [@ACTUC]. 
 Howe [@HEDP] discusses the impact of electronic computer-assisted techniques.
+Geographical information systems allow map creators to produce cartograms and they use these systems depending on ‘the effectiveness, efficiency, and satisfaction of the map products' (Nielsen 1994) [@CD].
 
-There are many alternatives to consider, the intended audience of the map, and its purpose are key points in cartogram use and creation.
- 'There is no "best" cartogram or method of creating cartograms just as there is no "best" map' (Monmonier and Schnell, 1988) [@ACTUC]. 
+There are many issues to consider when creating alternative map displays: the intended audience of the map, and its purpose are key points in cartogram use and creation as 'there is no "best" cartogram or method of creating cartograms just as there is no "best" map' (Monmonier and Schnell, 1988) [@ACTUC].
 Nusrat and Kobourov [@SAIC] provided a framework to investigate implementations of the many algorithms presented, and the "statistical accuracy, geographical accuracy, and topological accuracy".
@@
 ```{r ggcartograms, fig.cap = "Common alternatives to maps, showing the same information as in Figure \ref{choroCRS} for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous).
-In (a) - (c) the state has been resized, and reshaped, to match the 2015 population of the state.
-This provides a better sense of the extent of disease relative to the population in the country and can help alleviate losing information about physically small but population-dense states.
-In the hexagon tilemap (d) each state is given equal size and thus equal emphasis.", message=FALSE, warning=FALSE, out.width='100%'}
+Maps (a) - (c) are created by resizing and reshaping the states of the USA to match the 2015 population of the state.
+This provides a better sense of the extent of disease relative to the population in the country and can help ease losing information about physically small but population-dense states.
+Map creators give each state equal size and thus equal emphasis in (d) the hexagon tile map.", message=FALSE, warning=FALSE, out.width='100%'}
 g1 <- png::readPNG("figures/ggcont.png")
@@
 
-Table: \label{tab:usa} Maps used to present statistics for the United States of America.
-Each state has been colored according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Table: 2<!--\label{tab:usa}--> Maps used to present statistics for the United States of America.
+The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
 
@@
 |a.
-Contiguous| Each stated shape has been distorted according to the population of the state in 2015.
-The state of California has become much larger due to its large population density.
+Contiguous| It has distorted each state's shape according to the population of the state in 2015.
+The state of California has become much larger because of the large population density.
 This draws attention to the densely populated North-East region and detracts from the less populated Mid West.|
 |b.
-Non-contiguous| The geographic shape of the states has been maintained, but the size has altered according to the population of the state in 2015.
-The state of California has remained closer to its original size than it's surrounding states.
-The North-East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut.
+Non-contiguous| It maintains the geographic shape of the states, but the size has altered according to the population of the state in 2015.
+The state of California has remained closer to its original size than its surrounding states.
+The North-East states have remained closer to their geographical size, for Massachusetts and Connecticut.
 This draws attention to the densely populated North-East region and the sparse Mid West.|
 |c.
-Dorling| The states have been represented by a circle, but the size was determined by the population of the state in 2015.
-The North-East states remain closer to their neighbors and could be displaced from their geographic location.
-The sparsity of the population in the Mid West is highlighted by the distance between the circles, located at the geographic centroids.|
+Dorling| Circles are used to represent each state, but the population of the state determines the size in 2015.
+The North-East states remain closer to their neighbors and are slghtly displaced from their geographic location.
+It highlights the sparsity of the population in the Mid West by the distance between the circles at the geographic centroids.|
 |d.
-Hexagon Tessellation| Each state is represented by a hexagon of equal size.
-The neighboring states are easily contrasted however the North-East regions have been displaced from their geographic location.
-The sparsity of the population in the Mid West is highlighted by the light yellow color, the Age-Adjusted rate in Kentucky is the darkest and its neighbors are similar.|
+Hexagon Tessellation| A hexagon of equal size represents each state.
+It is easy to contrast the neighboring states however the North-East regions have been displaced from their geographic location.
+It highlights the sparsity of the population in the Mid West by the light yellow color, the Age-Adjusted rate in Kentucky is the darkest and its neighbors are similar.|
 
-Figure \ref{fig:ggcartograms} shows four different cartograms for the same data displayed in Figure \ref{fig:choroCRS}.
-Table {tab:usa} summarizes information that can be observed in the four types of cartograms.
+Figure 3<!--\ref{fig:ggcartograms}--> shows four different cartograms for the same data displayed in Figure 2<!--\ref{fig:choroCRS}-->.
+The information in Table 2<!--\ref{tab:usa}--> summarizes what can be observed in the four types of cartograms.
 
@@
 
-A contiguous cartogram maintains connectivity of the map regions while areas are altered according to a statistic.
+A contiguous cartogram alters the choropleth according to a statistic and maintains connectivity of the map regions.
 This transformation often occurs at the expense of the shape of areas [@CBATCC, @NAC, @TAAM].
-Three methods are provided for creating value-by-area cartograms [@ACA]. They implement 'map deformation' to account for the value assigned to each area.
+Min Ouyang and Revesz [@ACA] present three algorithms for creating value-by-area cartograms. They implement 'map deformation' to account for the value assigned to each area.
 Other methods include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [@ACTUC], Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and Tikunov’s Line Integral Method, Constraint-Based Method (Kocmoud and House) [@CBATCC].
@@
 This intentional design goal coincided with the rising interest in urban geography and presentation of social statistics.
-Figure \ref{fig:ggcartograms} a shows a population contiguous cartogram of the United States.
+Figure 3<!--\ref{fig:ggcartograms}--> a shows a population contiguous cartogram of the United States.
 All states are visible and the shape of the United States overall is still recognizable.
@@
 
-To be able to recognise the significant changes, a reader will usually have to know the initial geography to find the differences in the new cartogram layout [@NAC].
+To be able to recognize the significant changes, a reader will usually have to know the initial geography to find the differences in the new cartogram layout [@NAC].
 Tobler's Conformal mapping means to preserve angles locally so that the shapes of small areas on a traditional map and a cartogram would be similar.
@@
 Each area stays in a similar position to its location on a choropleth map.
-The choropleth map base is often also displayed to allow comparisons to be made regarding the change in the area displayed.
+Displaying the choropleth map base allows map users to make comparisons regarding the change in the area.
 The addition is the gap between areas, created as each area shrinks or grows according to the associated value of the statistic.
-Olson [@NAC] discusses the creation of these maps, the significance of the empty areas left between the geographic boundaries and the new shape, and the 'degree of difference from the original map that is the real message' of these displays.
+Olson [@NAC] discusses the creation of these maps, the significance of the empty areas left between the geographic boundaries and the new shape, and the 'difference from the original map that is the real message' of these displays.
 
-As the trade-off regarding boundaries approaches simplicity, the distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the non-contiguous cartogram but is replaced by the meaningful empty-space property [@ECGC, @NAC].
-
+The distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition, and the non-contiguous cartogram eliminates this hurdle, and presents the meaningful empty-space property [@ECGC, @NAC].
 <!-- The shapes are valuable for recognition and allows users to orient themselves on the display.
@@
 However it may impact on the visibility of all areas if small areas expand beyond their boundaries. 
-Figure \ref{fig:ggcartograms}b shows small overlaps in the North-East region of the United States.
+Figure 3<!--\ref{fig:ggcartograms}-->b shows small overlaps in the North-East region of the United States.
 These overlaps are also shown in the non-contiguous cartogram of Australia in Figure \ref{fig:auscartograms} b.
@@
 Daniel Dorling presents an alternative display engineered to highlight the spatial distribution and neighborhood relationships without complex distortions of borders and boundaries.
-This approach opposes preserving the intricate shape details and is founded in the simple question put forward by Daniel Dorling -@ACTUC:
+Daniel Dorling puts forward a simple question and his approach opposes preserving the intricate shape details [@ACTUC]:
 
@@
 Circular cartograms use the same simple shapes for every region represented and resize the shapes according to the statistic represented or the population for a base map.
-This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms, as the 'nonsense' shapes shown in contiguous cartograms have 'no meaning' after distortions are applied [@NISCC].
-To produce a compelling map, a gravity model is applied to avoid overlaps and keep spatial relationships with neighboring areas over many iterations.
-This implementation can work for up to 'one hundred thousand' areas.
+This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms.
+Contiguous cartograms use a gravity model to apply distortions to create the 'nonsense' shapes that have 'no meaning' [@NISCC].
+This method applies a gravity model to produce a compelling map, that avoids overlaps and keep spatial relationships with neighboring areas over many iterations.
+This implementation can work for up to ‘one hundred thousand’ areas.
 
-The groundwork for this approach had been laid in the mid-1930s by Raisz, and rectangular cartograms provide dramatic comparisons and are especially useful for correcting misconceptions communicated by geographic maps.
+Raisz laid the groundwork for this approach in the mid-1930s, and his rectangular cartograms provide dramatic comparisons and are especially useful for correcting misconceptions communicated by geographic maps.
 @TFYCC quotes the official definition of Value-Area Cartograms, the simplistic displays which represent each area as a single rectangle, sized according to the value of the statistic.
-This rectangular display also allows for tiling, where geographic neighbors placed in suitable relative positions also share borders however, contiguity may be sacrificed [@CDWCS].
-Rectangular cartograms allow for bi-variate displays as the population can be effectively communicated by the size of each rectangular, and a second variable can be communicated using color [@ORC].
+This rectangular display may sacrifice contiguity but allows for tiling where geographic neighbors placed in suitable relative positions also share borders [@CDWCS].
+Rectangular cartograms communicate bi-variate displays of the population by the size of each rectangular, and they use color to communicate a second variable [@ORC].
 
-A similar method, where each geographic area is represented using a square, tessellated to create a square grid.
+A similar method to a rectangular cartogram, represents each geographic area using a square.
+The squares are tessellated to create a grid.
 This method has been used broadly in the media, e.g.
@@
 <!--
-Figure \ref{fig:ggcartograms}d shows a hexagon tilemap representing each state using a single hexagon.
+Figure 3<!--\ref{fig:ggcartograms}-->d shows a hexagon tilemap representing each state using a single hexagon.
 
@@
 
-## tile map
+## Tile Map
 <!-- 173 words -->
-A tilemap provides a tessellated display of consistent shapes.
-Each tile is usually one unit of measurement, this could be geographic regions such as states, or population-based where one tile is used for a consistent measure of population.
+A tile map provides a tessellated display of consistent shapes.
+Each tile is usually one unit of measurement, this could be geographic regions such as states or population-based that use a consistent measure of population for each tile.
+Regions with over four neighbors require some necessary displacement.
+The tile map uses color to represent a value of a statistic for each area.
+A similar method to a rectangular cartogram represents each geographic area using a square of the same dimensions.
+The squares tessellate to create a grid.
+There are online media sources using this method, such as FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
+Tile maps may be difficult to create as they are best created manually, they require additional time and care as the number of geographic areas to include increases.
 <!-- cite mosaic cartograms -->
 
-A simple tilemap presents the areas in a tessellated grid display, where geographic neighbors are found next to each other, with some necessary displacement employed for regions with more than four neighbors.
-These tiles may be labelled or colored to represent a value.
-tile maps may be difficult to create, they are best created manually, with additional time and care required as the number of geographic areas to include increases.
-Cano and others [@MDAC] define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area can be used to communicate the statistic of regions. 
-The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off to be made between boundary complexity and simplicity.
-
-```{r ggtilemap, fig.cap = "Two more alternative displays, tilemap (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
-In the tilemap, the layout approximates spatial location, with each state being an equal box filled with color representing cancer incidence.
-The geofaceted map shows bar charts laid out in a grid approximating the spatial location of the state.
-The age-adjusted rates for males and females are shown.
-This type of display allows multiple variables to be displayed for each geographic area.", out.width="100%"}
-knitr::include_graphics("figures/gggrids.png")
-```
+Cano and others [@MDAC] define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area or the color of them can communicate the statistic of regions. 
+When using several tile per regions, map makers can adjust the complexity of the boundaries in the resulting display.
+They can also make a trade-off between boundary complexity and simplicity by the size of the tiles used.
 
@@
 Hafen [@IGF] formalizes the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas.
-Like tile maps, the arrangement of the areas can be reused for other spatial distribution visualizations.
-Geofaceting has the functionality of facets, often used to replicate visualizations for each subset of the data. In this case, the data subsets are geographic.
-The amount of information able to be communicated has increased from one value per region in a tilemap to one visualization, this is a more flexible display.
-Virtually any plot display can be fit into the tile representing the areas, allowing displays of multiple variables or values per geographic entity.
-<!-- As the amount of areas increase care must be taken to ensure legibility of the displays.
-@RPIG acknowledges that this display suffers when the areas have very irregular shapes and large size disparities CORRECT reference is Cheng's thesis-->
-Manual creation of the grid is required.
+Like tile maps, they can reuse the arrangement of the areas for other spatial distribution visualizations.
+Geofaceting has the functionality of facets, often used to replicate visualizations for each subset of the data.
+Here, the data subsets are geographic.
+A tile map communicates one value per region in a visualization, geofaceting is a more flexible visualization for communication as it increases the amount of information displayed.
+Virtually any plot display can be shown in the tile representing the areas, allowing displays of multiple variables or values per geographic entity.
+Creating a tile map or a geofacet both require manual creation of the grid, except for the grids that other have already contributed.
+
+```{r ggtilemap, fig.cap = "Two alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
+In the tile map, the layout approximates spatial location, with each state being an equal box filled with color representing cancer incidence.
+The geo-faceted map shows bar charts laid out in a grid approximating the spatial location of the state.
+The maps show age-adjusted rates for males and females.
+This display allows the presentation of multiple variables for each geographic area.", out.width="100%"}
+knitr::include_graphics("figures/gggrids.png")
+```
 
@@
 
-Pickle and others [@MMST] present linked micro map plots to visually link geographic and statistical data, this serves as a solution to  multi-dimensionality issues.
-This design has been used in online atlas displays of US states.
-Areas are grouped based on their value for one variable, and additional columns provide displays that contrast the areas in each group by other variables.
-The display juxtaposes of choropleth maps and statistical plots;
+Pickle and others [@MMST] present linked micro map plots to match geographic and statistical data visually,  this serves as a solution to  multi-dimensionality issues.
+An online atlas of US states uses this display.
+These maps group areas based on their value for one variable, and additional columns provide displays that contrast the areas in each group by other variables.
+The display juxtaposes choropleth maps and statistical plots;
 it shows one map per group of the key separating variable, in a row with each additional statistical plot.
-Linked micro maps predominantly utilize the choropleth map for displays of spatial relationships, that are seen when spatial neighbors are allotted to the same group.
+Linked micro maps predominantly use the choropleth map for displays of spatial relationships. These maps show spatial relationships by allotting spatial neighbors to the same group.
 It is one of several alternative displays that allow maps to become bivariate displays, commonly used to present both an estimate and the associated uncertainty.
@@
 Bivariate choropleth maps [@VUADBC] blend color schemes to convey the intersection of categorized levels of an estimate and the associated uncertainty for each spatial area.
-Lucchesi and Wikle [@VUADBC] also suggests map pixelation, each region is broken into small pixels, the individual pixels are allocated values that reflect the uncertainty around the area's estimate by creating texture.
-This display can also be animated, with each frame produced by resampling the pixels.
+Lucchesi and Wikle [@VUADBC] also suggest map pixilation, it breaks each region into small pixels, and  allocates values to the individual pixels to create texture.
+This reflects the uncertainty around the area's estimate by randomly sampling from the confidence interval of the estimate of the area.
+Animating these displays involves resampling the pixels for each frame.
 Areas with uncertain values will flicker more dramatically than areas with more certain values.
-
-<!-- Solutions have been provided for bi-variate displays for movement data, containing Origin - Destination matrices and flow map representations [@MMTVCG]. 
-This method makes the maps and matrix connections clear, and easy to link geographic areas with link lines not crossing. Bar charts can also be added to these to show group totals.
--->
 
@@
 
-```{r auscartograms, fig.cap = "Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tilemap.
+```{r auscartograms, fig.cap = "Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map.
 The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas.
-If it accurately sized areas by population the country would be unrecognizable.
-The shape-preserved is unreadable.
-Dorling has lost all geographic context.
-The hexagon tilemap provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%'}
+If it accurately sized areas by population, the country would be unrecognizable.
+The shape-preserved is unreadable due to the small area sizes.
+The Dorling cartogram presents all areas but many are difficult to compare.
+The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%'}
 
@@
 
-## Neither choropleths or cartograms perform well for Australia
+## Neither choropleth maps or cartograms perform well for Australia
 
@@
 The version of a contiguous cartogram (a) has expanded the highly populated areas while preserving the full shapes of rural areas.
-It has not fully resolved the population transformation of areas, and if it had accurately sized areas by population the country would be unrecognizable.
-The shape-preserved cartogram is unreadable, and all areas have been reduced to tiny spots on the map.
-Zooming in on a high-resolution output shows that it does indeed preserve the areas though.
+It has not fully resolved the population transformation of areas, and if it had accurately sized areas by population, the country would be unrecognizable.
+The shape-preserved cartogram is unreadable, and it has reduced all areas to tiny spots on the map.
+Zooming in on a high-resolution output shows it does preserve the areas.
 The Dorling cartogram has lost all geographic context.
-The hexagon tilemap provides a reasonable spatial distribution despite having too much white-space in the outback areas.
+The hexagon tile map provides a reasonable spatial distribution despite having too much white-space in the outback areas.
 
@@
 Cartograms provide the spatial distortion to more accurately convey the statistical distribution, focuses on the human impact of the disease.
-However, if the population density is highly dissonant with geographic density the cartogram will lose all spatial context, and thus futile.
+However, if the population density is highly dissonant with geographic density, the cartogram will lose all spatial context, and thus futile.
 Dorling [@ACTUC] has a cartogram showing the 1966 general election results, which looked very little like the geographical shape of Australia.
 
-Some mix of tiling, faceting or even micromaps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies.
-Table \ref{tab:methods} summarizes the key criteria for evaluating maps and alternative displays.
-Suggestions and comments to help map creators best communicate their health data and spatial analyses are provided by Moore and Carpenter [@SAMGIS], Bell et al.
-[@CPISACA].
+Some mix of tiling, faceting or even micro maps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies.
+Table \ref{tab:methods} summarizes the key criteria for testing maps and alternative displays.
+Moore and Carpenter [@SAMGIS] and Bell et al.
+[@CPISACA] provide suggestions and comments to help map creators best communicate their health data and spatial analysis.
 
-<!-- Choropleths, true topology, geographic distance, good for showing geographic importance 
+<!-- choropleth maps, true topology, geographic distance, good for showing geographic importance 
 To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
@@
 However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a color.
-These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure \ref{fig:choropleth-grid-create} c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
+These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure 1<!--\ref{fig:choropleth-grid-create}--> c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
 Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbors. The map is reusable and fit for many purposes.
@@
 The Australian Cancer Atlas [@TACA] provides [tours](
-https://atlas.cancer.org.au/app/tour/lungcancer) that change the display to the draw users' attention to areas on the map that are relevant to the story.
+https://atlas.cancer.org.au/app/tour/lungcancer) that change the display to draw users' attention to areas on the map that are relevant to the story.
 This implementation of animation gives users tools to plan their exploration.
@@
 Selections can be made statistically through interactions which allow subsets to be selected from statistical displays, such as tables of data.
-Monmonier implemented linking choropleths with statistical plots, adding a geographic component to scatter plots [@MCRG], the use of link allows spatial dimensions to be reserved for geographic locations of areas.
+Monmonier implemented linking choropleth maps with statistical plots, adding a geographic component to scatter plots [@MCRG], the use of link allows spatial dimensions to be reserved for geographic locations of areas.
 The statistical variation of additional variables can be presented in other appropriate displays.
@@
 Using prediction:
-assuming that neighbouring μi are correlated leads to a ‘clustering’ prior
+assuming that neighboring μi are correlated leads to a ‘clustering’ prior
 

### 2019-12-16 review_article.Rmd
Figures for AUS and editing

@@
 <!-- What data are we working with, why, what are the purposes -->
-Researchers, health authorities, governments and not-for-profits are common communicators of cancer statistics.
+Researchers, health authorities, governments, not-for-profits and the media are common communicators of cancer statistics.
 They often present statistics to the public as aggregated values for geopolitical areas.
-Presenting these statistics requires aggregating individual observations into the statistics for the geographical units, especially for privacy protection, but also for political and policy purposes.
-Data collection or aggregation regions for the amount of cancer diagnoses or excess deaths per area include states, provinces, local government areas, and post/zip codes.
-To provide the amount, or count, of the diagnoses or excess deaths with these areas, it needs minimal transformation.
-However, the amount of people living in each of the areas in a set will be different.
-Those who wish to calculate rates can use the amount of people living in an area as a denominator and the diagnoses can become a numerator to create rates per amount of people.
-Health registries routinely collect data for public health reasons and may make this available to the public as a service to the community.
+Presenting these statistics requires aggregating individual observations for the geographical units, especially for privacy protection, but also for political and policy purposes.
+Examples of typical geographical units include states, provinces, local government areas, and post/zip codes.
+It is easy to provide counts of the diagnoses or excess deaths within these areas.
+However, the amount of people living in each of the areas in a set of areas will be different.
+Calculating incidence rates involves dividing the count by the population for each area, for example, the rate per 100000 people.
+Counts and population datasets are collected routinely for public health reasons and may be made available to the general public as a service to the community.
 
 <!-- Ways to visualize, whats our hypothesis -->
-Choropleth map displays are commonly used for the communication of cancer statistics over geographic domains.
-In this display, map creators map a statistic to color and apply the colour to the polygons representing geographic regions.
-Map users may overlook geographic patterns of disease when obscured in a table [@SAMGIS].
-Visualizing diseases on maps is often the first step in exploratory spatial data analysis and helps in the formulation of hypotheses.
-Providing a visual representation of cancer outcomes allows identification, and public health policy and actions can address these patterns.
-choropleth maps allow the examination of the spatial distribution of the disease incidence, this may reveal a trend in longitude or latitude, differences between rural areas and urban, or coastal and inland areas, or even specific hot spots of the disease.
-Daniel Exeter [@SE] recognizes that one of the key challenges with mapping spatial patterns of disease is the design of visualizations.
-This paper considers the current visualization techniques to communicate statistics to members of the public and their applications to cancer statistics.
-It also considers alternative approaches and compares the visualization methods, highlighting the differences and historic use of these displays.
+To visualize and communicate geospatial cancer statistics over geographic domain, a choropleth map is the common display.
+Choropleth maps show polygons representing the geographic units, where each polygon is shaded with a color according to the area-specific values of the statistic being conveyed.
+Visualizing this data is helpful as geographic patterns of disease may be obscured when reported in a table [@SAMGIS].
+Providing a visual representation of cancer outcomes allows identification of geographic patterns of the disease that can then be addressed with public health policy and actions.
+The spatial distribution of the disease incidence can be examined using a choropleth and may reveal a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease.
+One of the key challenges with mapping spatial patterns of disease is the design of visualizations [@SE].
+It is important to consider the strengths and weaknesses of designs, as visualizing diseases on maps is often the first step in exploratory spatial data analysis and helps in the formulation of hypotheses.
+This paper considers the current visualization techniques to communicate statistics to the public and their applications to cancer statistics.
+Alternative approaches are also considered, and the limitations of the visualization methods, highlighting the differences and historic use of these displays.
 
@@
 <!-- what is a choropleth map -->
-A choropleth is traditionally used to make a cancer map.
-It is used to show the geographical distribution of data values by shading areas of a map [@EI].
-A choropleth map involves faithfully rendering the geography, and the design of the color scheme and rendering reveal spatial patterns among data values.
+A choropleth map displays the geographical distribution of data over a set of spatial units by shading areas of a map [@EI, @BCM].
+Faithful rendering of the geography, when combined with an appropriate color scheme, can reveal spatial patterns among data values.
 Identifying and explaining spatial structures, patterns, and processes involve considering the individuals and organizing them into representable units of communities [@SAMGIS].
 Early versions of choropleth maps used symbols or patterns instead of color.
-choropleth maps can be used for displaying disease data [@DMAHP], including cancer data [@CPISACA].
+Choropleth maps can be used for displaying disease data [@DMAHP], including cancer data [@CPISACA].
 In epidemiology, choropleth maps are often used as a tool to study the spatial distribution of cancer incidence and mortality.
-
-Figure 2<!--\ref{fig:choroCRS}--> shows various geographic projections of choropleth maps of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged from 2012 to 2016.
-The official federal statistics provide cancer incidence and deaths [5], produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
-The rates exhibit a spatial trend, increasing from west to east.
-Utah is a spatial outlier, it has a noticeably lower rate than its neighbors.
-Also, Kentucky and Maine have higher rates than their neighbors.
-There is possibly a cluster of higher rates around the tobacco states.
 
@@
 Displaying familiar state boundaries can make a map easier to read [@CIBMUK] and allow viewers to infer the spatial relationships visually in the data using their mental model of the geography.
-Consumers of disease displays may include researchers, members of the public, policymakers, and the media  [@CPISACA].
-For these consumers, the familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
-
-<!-- potential problems -->
-Politically driven boundaries produced the now familiar shapes or regions, this has created individual areas of non-uniform size.
-These areas have different population densities and are subject to change.
-The different population and geographical sizes of administrative areas can attract attention to the shades of the underpopulated but large areas [@EI], Skowronnek [@BCM] calls this an area-size bias.
-choropleth maps can inhibit visual inference when presenting human-related statistics as the display may draw attention away from the 'potentially more important results in the more populous communities' that are geographically smaller [@SE].
+The map users of disease displays may include researchers, the public, policymakers, and the media [@CPISACA].
+For these users, the familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
 
@@
 <!-- Displaying of cancer data on choropleth maps -->
-A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country or group of countries.
+A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
 Atlases are key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work-related exposures, and high-risk diets [@MACM].
@@
 <!-- Measures reported in cancer atlases -->
-Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the general public and others not trained in statistical analysis.
+Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the public and others not trained in statistical analysis.
 These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analysis.
@@
 
-Epidemiologists and statisticians have developed these statistics over several decades.
-Mortality rates commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations.
+Epidemiologists and statisticians have developed the statistics used to communicate the burden of cancer over several decades.
+Table 1<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases.
+Mortality rates are commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations.
 Howe [@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s.
-Table 1<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases.
 
@@
 |5. Age-Adjusted Relative Risk| Standardized by age structure in each region $i$|  
-|6. SIR (Standardized Incidence Ratio)| Standardized by incidence in each region $i$|
+|6. SIR (Standardized Incidence Ratio)| Incidence standardized by population at risk in each region $i$|
 |7. Below or above Expected| An alternative expression of the SIR|
@@
 <!-- 416 words -->
-Roberts [@roberts2019communication] identified 33 publicly available cancer atlases, published between 2010 and 2018, from different countries around the globe.
-These online atlases used choropleth maps.
-Non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships published all but one of these.
-Figure 1<!--\ref{fig:choropleth-grid-create}--> displays a subset of maps from these atlases.
-
-The selection varies in the geographies explored, Figure 1<!--\ref{fig:choropleth-grid-create}-->b, Globocan 2012 [@Globocan] explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide utilizing data sourced from cancer registries of each country.
-The Bowel Cancer Australia Atlas presents an example of a cancer specific atlas, Figure 1<!--\ref{fig:choropleth-grid-create}-->d shows the average Standardized Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia [@Bowel].
+Roberts [@roberts2019communication] identified 33 cancer atlases published between 2010 and 2018.
+Each of these online atlases uses choropleth maps.
+All except one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships.
+Figure 1<!--\ref{fig:choropleth-grid-create}--> displays a subset of maps from these atlases, the selection varies in the geographies explored.
+Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows Globocan 2012 [@Globocan] which explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide using data sourced from cancer registries of each country.
+The Bowel Cancer Australia Atlas in Figure 1<!--\ref{fig:choropleth-grid-create}-->d presents an example of a cancer specific atlas, it shows the average Standardized Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia [@Bowel].
 Like many of the atlases examined, there is a choice of gender displayed in the Bowel Cancer Atlas.
@@
 The Environment and Health Atlas of England and Wales, Figure 1<!--\ref{fig:choropleth-grid-create}-->a shows the relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2] at a neighborhood (small-area) scale.
-The Atlas of Cancer in Queensland, Figure 1<!--\ref{fig:choropleth-grid-create}-->c shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 in the state of Queensland within Australia [@QLDcancerAtlas].
+The Atlas of Cancer in Queensland, Figure 1<!--\ref{fig:choropleth-grid-create}-->c shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@SA2] in the state of Queensland within Australia [@QLDcancerAtlas].
 
 Age-specific atlases are less common, Figure 1<!--\ref{fig:choropleth-grid-create}-->g displays Atlas of Childhood Cancer in Ontario, this communicates the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
-This atlas communicates information regarding a tiny geographic region.
+This atlas communicates information regarding a small geographic region.
 
@@
 <!-- 479 words -->
-Including supplementary graphs and plots add more depth and information.
-Additional materials such as tables, graphs, and text explanations, support the understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers [@CPISACA].
-The many displays of statistical summaries, including dot plots, bar plots, box plots, cumulative distribution plots, scatter plots, normal probability plots, can provide alternative views of the cancer statistics.
-These can also display supporting statistics such as error, confidence intervals, distributions, sample or population sizes, standard deviation.
-When presenting cancer maps the display of the supplementary statistics helps.
-The intuition derived from maps must be ‘validated by rigorous statistical analyses’ [@MACM].
+Cancer atlases often display supplementary graphs and plots to add more information.
+Additional materials such as tables, graphs, and text explanations support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers [@CPISACA].
+The many displays of statistical summaries, including dot plots, bar plots, box plots, cumulative distribution plots, scatter plots, and normal probability plots, can provide alternative views of the cancer statistics.
+These can also display supporting statistics such as error, confidence intervals, distributions, sample or population sizes, and standard deviation.
 
@@
 Similarly, socioeconomic indicators, such as unemployment rates, poverty rates, remoteness, and education levels, can explain how the experience of cancer prevalence varies for various members of society.
-Few atlases provide this level of detail.
+Few atlases provide this
+level of detail.
 Introducing population and demographic information helps to interpret the rates in areas effectively, but there will still be uncertainty around the rates.
-To address this, cancer atlases often communicate uncertainty about the value of a statistic.
-Cancer atlas creators may introduce uncertainty when using samples, or there are errors occurring in the disease reporting, or models are used to simulate values to maintain privacy.
-The statistics that report on uncertainty include sample or population size, standard deviation, confidence intervals, statistical significance, quantiles, and  credible intervals.
+To address this, a cancer atlases often communicates uncertainty about the value of a statistic.
+There are several potential sources of uncertainty: sampling error, errors arising from the disease reporting process (or data collection), and errors arising from the statistical modeling or simulation process.
 The most common measures used to represent uncertainty were credible or confidence intervals (CIs).
@@
 
-The administrative boundaries used to define regions may limit a choropleth display.
-Those who critique choropleth maps often dicuss election mapping, where each electorate decides the outcome of one parliamentary seat in the house of representatives.
-While all states are visible in a map display of Australia, smaller regions such as electorates present an urban-rural divide, where rural regions are much larger than metropolitan areas of similar or greater populations.
+Choropleth maps provide a familiar display that may be important for orienting a non-statistical audience.
+The familiar shape boundaries define individual areas of non-uniform size, containing different population densities and may be subject to change.
+A spatially heterogeneous population distribution across a country’s landmass is an issue common to many countries, especially British colonies [@ACTUC].
+
+A disadvantage of the conventional map is that the different population and geographical sizes of administrative areas can attract attention to the shades of the underpopulated but large areas [@EI].
+Skowronnek also [@BCM] discusses how choropleth maps suffer from area-size bias, as they give a 'stronger visual weight' to large  administrative units.
+The administrative boundaries used to define regions may limit a choropleth display, as this display unfaithfully represents the disease distribution across the region by obscuring small geographic areas.
+Sparsely populated rural areas are emphasized, whereas the areas representing inner city communities are very small.
+
+Those who critique choropleth maps often discuss bi-partisan election mapping in the United States of America [@BCM, @NISCC].
+Election mapping in Australia also suffers when presented on choropleth maps, where each electorate decides the outcome of one parliamentary seat in the house of representatives.
+While all states are visible on a map display of Australia, smaller geographic units such as electorates present an extreme case of an urban-rural divide.
 The landmass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas.
-To present election results on a choropleth map should be 'unthinkable',XXX as it means diminishing the visual impact of most of the electorates.
-Using a map of Australia as a visual statistic would communicate an incorrect distribution of votes across the country.
+To present election results on a choropleth map should be ‘unthinkable’, as it means diminishing the visual impact of most of the electorates and incorrectly communicating the distribution of votes across the country.
 
-Spatially heterogeneous population distribution across a country's landmass is an issue common to many countries, especially British colonies [@ACTUC].
-However, the boundaries shown in choropleth maps provide a familiar display.
-This may be less threatening to an audience that could be unfamiliar was interpreting spatial statistics, even though they suffer primarily from unfaithfully representing the disease information by hiding small geographic areas.
-Inset maps like in Brisbane city in Figure 1<!--\ref{fig:choropleth-grid-create}-->c of the state of Queensland are commonly used to reduce distorted interpretations, but it is a band aid remedy.
+Choropleth maps can inhibit visual inference when presenting human-related statistics as the display may draw attention away from the ‘potentially more important results in the more populous communities’ that are geographically smaller [@SE].
+Inset maps like in Brisbane city in Figure Figure 1<!--\ref{fig:choropleth-grid-create}-->c of the state of Queensland are commonly used to reduce distorted interpretations, but it is not always an appropriate solution.
 As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
@@
 <!-- What is a cartogram? -->
-choropleth maps imply uniformity of data across the geographic space but population densities are unlikely to be uniform [@BCM].
+Choropleth maps imply uniformity of data across the geographic space but population densities are unlikely to be uniform [@BCM].
 Cartographers developed the cartogram to draw the attention to the population by transforming the map [@ACCAC].
 The resulting display can communicate the impact of the disease more accurately across the population, as recorded by the statistic, at the sacrifice of geographic accuracy.
-
-<!-- There are many alternative views of the world-->
-<!-- Tobler map projections - I think this is not relevant here
-choropleth maps may be considered true topological maps however,, if the landmass displayed covers enough of the globe, there must be a transformation or distortion to display the land in 2D [@HTLWM].
-The amount of distortion is related to the distance covered by the landmass displayed @GAMP.
-World map projections reflect the frequent perspectives used to view the earth.
-Choropleth maps will always be distorted if they cover enough of the globe, just like photographs of the globe from space.
-Choropleth creation requires choosing a map projection that shows a favorable distortion of the geography for presenting the set of spatial information.
-Selecting a display can prevent misinterpretation of global statistics, as global maps face the challenge of equitable displays of landmass on maps [@RSCW].
-If the statistic presented on the map base relies on physical distance and is influenced by the topology there is no transformation needed, beyond choosing a reasonable projection. 
--->
 

### 2019-12-16 sa3_plots.R

@@ -161,8 +161,10 @@ ggsave(filename = "figures/aus_ggncont.png", plot = full_ggncont,
 
 ###############################################################################
 # Non - Contiguous Dorling Cartograms
-dorl <- sa3lung %>% filter(cent_long >110, cent_long<155) %>% cartogram_dorling(.,
-  weight = "Population", k = 0.1) %>% st_as_sf(crs=CRS("+init=epsg:3112"))
+dorl <- sa3lung %>% filter(cent_long >110, cent_long<155) %>% 
+  mutate(pop = (Population/max(Population))*10) %>% 
+  cartogram_dorling(., k = 0.01, weight = "pop", m_weight = 1) %>%
+  st_as_sf(crs=CRS("+init=epsg:3112"))
 d <- st_bbox(dorl)
 aus_ggdorl <- ggplot(dorl) + 
   geom_sf(aes(fill = `Age-standardised rate (per 100,000)`)) + 

### 2019-12-16 review_article.Rmd
figure ordering

@@
   draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
-    c(0.015, 0.33, 0.63, 0.015, 0.33, 0.014, 0.63), 
-    c(0.99, 0.99, 0.99, 0.74, 0.74, 0.47, 0.47), size = 20)
+    c(0.015, 0.31, 0.66, 0.015, 0.31, 0.014, 0.66), 
+    c(0.99, 0.99, 0.99, 0.74, 0.74, 0.46, 0.46), size = 20)
 
@@
 The Environment and Health Atlas of England and Wales, Figure 1<!--\ref{fig:choropleth-grid-create}-->a shows the relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2] at a neighborhood (small-area) scale.
-The Atlas of Cancer in Queensland, Figure 1<!--\ref{fig:choropleth-grid-create}-->c shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@SA2] in the state of Queensland within Australia [@QLDcancerAtlas].
+The Atlas of Cancer in Queensland, Figure 1<!--\ref{fig:choropleth-grid-create}-->c shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 in the state of Queensland within Australia [@QLDcancerAtlas].
 
@@
 
+<!--
 ```{r choroCRS, fig.cap = "Four choropleth maps average age-adjusted rate of incidence for lung and bronchus in the United States averaged from sex and years 2012-2016, using four different map projections.
@@
 Howe [@HEDP] discusses the impact of electronic computer-assisted techniques.
-Geographical information systems allow map creators to produce cartograms and they use these systems depending on ‘the effectiveness, efficiency, and satisfaction of the map products' (Nielsen 1994) [@CD].
+Geographical information systems allow map creators to produce cartograms and they use these systems depending on ‘the effectiveness, efficiency, and satisfaction of the map products' [@CD].
 
@@
 
-```{r ggcartograms, fig.cap = "Common alternatives to maps, showing the same information as in Figure \ref{choroCRS} for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous).
+```{r ggcartograms, fig.cap = "Common alternatives to maps, showing the same information for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous).
 Maps (a) - (c) are created by resizing and reshaping the states of the USA to match the 2015 population of the state.
@@
 
-Figure 3<!--\ref{fig:ggcartograms}--> shows four different cartograms for the same data displayed in Figure 2<!--\ref{fig:choroCRS}-->.
-The information in Table 2<!--\ref{tab:usa}--> summarizes what can be observed in the four types of cartograms.
+Figure 2<!--\ref{fig:ggcartograms}--> shows four different cartograms for the same data. The information in Table 2<!--\ref{tab:usa}--> summarizes what can be observed in the four types of cartograms.
 
@@
 A contiguous cartogram alters the choropleth according to a statistic and maintains connectivity of the map regions. 
-This transformation often occurs at the expense of the shape of areas [@CBATCC, @NAC, @TAAM].
 Min Ouyang and Revesz [@ACA] present three algorithms for creating value-by-area cartograms. They implement 'map deformation' to account for the value assigned to each area.
@@
 
-<!-- 
-An intentional goal when creating the 1966 Census population cartogram for Canada was to maintain contiguity, while attempting to keep the actual shape of places.
-The end result was a 'very accurate isodemo-graphic map of Canada'.
-This intentional design goal coincided with the rising interest in urban geography and presentation of social statistics.
-Figure 3<!--\ref{fig:ggcartograms}--> a shows a population contiguous cartogram of the United States.
-All states are visible and the shape of the United States overall is still recognizable.
-In contrast, Figure \ref{fig:auscartograms} a shows an Australian contiguous cartogram also based on population.
+Figure 2<!--\ref{fig:ggcartograms}--> a shows a population contiguous cartogram of the United States. All states are visible and the shape of the United States overall is still recognizable.
+In contrast, Figure 3<!--\ref{fig:auscartograms}--> a shows an Australian contiguous cartogram also based on population.
 The south east is completely distorted, areas with low population are still large very the map, as their initial area was used to determine the area allocated in the contiguous cartogram.
@@
 @CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes.
--->
 
@@
 The addition is the gap between areas, created as each area shrinks or grows according to the associated value of the statistic.
-Olson [@NAC] discusses the creation of these maps, the significance of the empty areas left between the geographic boundaries and the new shape, and the 'difference from the original map that is the real message' of these displays.
+Olson [@NAC] discusses the creation of these maps and the significance of the empty areas left between the geographic boundaries and the new shape.
+It is the 'difference from the original map that is the real message' of these displays.
 
 The distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition, and the non-contiguous cartogram eliminates this hurdle, and presents the meaningful empty-space property [@ECGC, @NAC].
-<!-- The shapes are valuable for recognition and allows users to orient themselves on the display.
-Map creators can efficiently communicate with this kind of map by keeping the outlines or particular elements of the original in the new shape [@NISCC].
-The scale of the areas does not impact on the shape recognition.
-However it may impact on the visibility of all areas if small areas expand beyond their boundaries. 
-Figure 3<!--\ref{fig:ggcartograms}-->b shows small overlaps in the North-East region of the United States.
-These overlaps are also shown in the non-contiguous cartogram of Australia in Figure \ref{fig:auscartograms} b.
--->
 
@@
 
-A similar method to a rectangular cartogram, represents each geographic area using a square.
-The squares are tessellated to create a grid.
-This method has been used broadly in the media, e.g.
-FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
-Each area is represented by a square of the same dimensions.
-<!--
-Figure 3<!--\ref{fig:ggcartograms}-->d shows a hexagon tilemap representing each state using a single hexagon.
-
-Recommended criteria to contrast mapping methods include average cartographic error, and maximum cartographic error, correct adjacency, maximum aspect ratio, and suitable relative positions [@ORC].
-However, this does not consider the issues with actually producing rectangular cartograms.
-Algorithms for the creation of rectangular cartograms
--->
-
 ## Tile Map
@@
 A tile map provides a tessellated display of consistent shapes. 
-Each tile is usually one unit of measurement, this could be geographic regions such as states or population-based that use a consistent measure of population for each tile.
+A similar method to a rectangular cartogram, represents each geographic area using a square.
+The squares are tessellated to create a grid.
+This method has been used broadly in the media by FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
+Each area is represented by a square of the same dimensions, each tile is usually one unit of measurement, this could be geographic regions such as states or population-based that use a consistent measure of population for each tile.
 Regions with over four neighbors require some necessary displacement.
@@
 The squares tessellate to create a grid.
-There are online media sources using this method, such as FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
+There are online media sources using this method, these include [NPR](https://www.npr.org/2016/10/18/498406765/npr-battleground-map-hillary-clinton-is-winning-and-its-not-close), [FiveThirtyEight](https://projects.fivethirtyeight.com/2016-election-forecast/), [The Wall Street Journal](http://graphics.wsj.com/elections/2016/2016-electoral-college-map-predictions/), [The Washington Post](https://www.washingtonpost.com/graphics/politics/2016-election/50-state-poll/).
 Tile maps may be difficult to create as they are best created manually, they require additional time and care as the number of geographic areas to include increases.
@@
 The Dorling cartogram presents all areas but many are difficult to compare.
-The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%'}
+The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%', cache=FALSE}
 
@@
 load("aus_legend.rda")
-
 load("figures/aus_legend.rda")
@@
 
-Figure \ref{fig:auscartograms} shows four main types of cartograms displaying melanoma incidence on SA3 areas in Australia.
+Figure 3<!--\ref{fig:auscartograms}--> shows four main types of cartograms displaying melanoma incidence on Statistical Areas at Level 3 [@ABSmaps] areas in Australia.
 The version of a contiguous cartogram (a) has expanded the highly populated areas while preserving the full shapes of rural areas.
@@
 
-## Pros and cons of different choices
+## Limitations of different choices
 
 Cartograms provide the spatial distortion to more accurately convey the statistical distribution, focuses on the human impact of the disease.
-However, if the population density is highly dissonant with geographic density, the cartogram will lose all spatial context, and thus futile.
+However, the transformation of contiguous cartograms often occurs at the expense of the shape of areas [@CBATCC, @NAC, @TAAM].
+When the population density pf the geographic units is highly dissonant with geographic density, the cartogram will lose all spatial context.
 Dorling [@ACTUC] has a cartogram showing the 1966 general election results, which looked very little like the geographical shape of Australia.
@@
 [@CPISACA] provide suggestions and comments to help map creators best communicate their health data and spatial analysis.
-
-<!-- choropleth maps, true topology, geographic distance, good for showing geographic importance 
-To choose an appropriate map display, the map creator must consider the intended user, and message the map will communicate.
-Choropleth displays utilize more traditional cartographic methods, they are usually true to the topography, displaying familiar boundaries of countries, states or administrative areas.
-Highlighting the geographic distribution, and differences between equal spatial units.
-When there is a relationship between the variable being mapped and geographical features, a choropleth map would pair this information visually.
-Choropleth methods require a decision to be made about the projection of the display, in this case aim to find a map projection that gives the least distorted representation of the geography [@UMPT].
-However a choropleth is limited to a uni-variate display, presenting a spatial distribution by filling each area with a color.
-These displays favor geographically large regions over small regions, though inset displays are often used to zoom in on a geographic area containing many small regions, this is seen in Figure 1<!--\ref{fig:choropleth-grid-create}--> c of the state of Queensland, where the capital city Brisbane is highlighted by the inset map.
-Users can identify the areas relevant to them by the familiar boundaries, and can easily contrast the experience of geographic neighbors. The map is reusable and fit for many purposes.
--->
 
@@
 
-## Interaction and animation in publicly available atlases
-
-Many publicly available cancer atlases employ interactivity and/or animation.
-Figure 6 <!--\ref{fig:interacting}--> shows the interactive GUI tools provided with the public atlases.
-Mostly these are menus, and toggles, for choosing indicators, and subsets of the population.
-
 ## Under-utilized techniques
@@
 <!-- 77 words -->
-The authors would like to thank Dr. Earl Duncan, Professor Kerrie Mengersen, Dr. Susanna Cramb and Dr. Peter Baade for conversations on the content of this article.
+The authors would like to thank Dr. Earl Duncan for his contributions in editing and refining the drafts of this article.
+They would also like to thank Professor Kerrie Mengersen, Dr. Susanna Cramb and Dr. Peter Baade for conversations on the content of this article.
 

### 2019-12-16 sa3_plots.R

@@ -89,7 +89,7 @@ aus_ggchoro <- ggplot(sa3lung) +
   scale_fill_distiller(type = "seq", palette = "Purples",  direction = 1, na.value = "light grey") + 
   
   coord_sf(crs = CRS("+init=epsg:3112"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
-  invthm + theme(legend.position ="bottom") +
+ theme(legend.position ="bottom") +
   labs(fill = "Age-standardised rate\n(per 100,000)")
 aus_ggchoro
 ggsave(filename = "figures/aus_ggchoro.png", device = "png",
@@ -133,7 +133,7 @@ ncont <- cartogram_ncont(sa3lung, k = 1/5,
   weight = "Population") %>% st_as_sf() %>% 
   rename(`Age-standardised rate (per 100,000)` = `Age.standardised.rate..per.100.000.`)
 aus_ggncont <- ggplot(ncont) + 
-  geom_sf(data=aus, fill = NA, colour = "grey", size = 0.01) +
+  geom_sf(data=aus, fill = NA, colour = "black", size = 0.1) +
   geom_sf(aes(fill = `Age-standardised rate (per 100,000)`), colour = NA) + 
   coord_sf(crs = CRS("+init=epsg:3112"), xlim =
              c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
@@ -144,7 +144,7 @@ aus_ggncont
 perth_ggncont <- ggplot(ncont %>% filter(gcc_name_2016 == "Greater Perth")) + 
   geom_rect(aes(xmin =-1736000, xmax = -1639000, ymin=-3824000, ymax=-3672300), fill = NA, colour = "black") +
   geom_sf(data= sa3lung %>% filter(gcc_name_2016 == "Greater Perth"),
-          fill = NA, colour = "grey", size = 0.001) +
+          fill = NA, colour = "black", size = 0.1) +
   geom_sf(aes(fill = `Age-standardised rate (per 100,000)`), colour = NA) + 
   scale_fill_distiller(type = "seq", palette = "Purples",  direction = 1) + invthm +
   guides(fill=FALSE)
@@ -156,14 +156,14 @@ full_ggncont <- ggdraw() +
   draw_plot(perth_ggncont, 0.33, 0.10, 0.25, 0.25)
 full_ggncont
[…]

### 2019-12-16 review_article.Rmd
di meeting changes

@@
 ---
-title: "Cancer Applications of Choropleth maps, and the Potential of Cartograms and Alternative Map Displays"
+title: "Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"
 author: "Stephanie Kobakian, Jessie Roberts and Dianne Cook"
@@
 
-# Traditional approaches for cancer map displays {#ch:choropleth maps}
+# Traditional approaches for cancer map displays {#ch:choropleth_maps}
 <!-- 381 words -->
@@
 
-```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Examples of publicly available choropleth cancer maps.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "A selection of choropleth cancer maps available online.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
 library(cowplot)
@@
 
-## Additional materials
+## Additional considerations
 <!-- 479 words -->

### 2019-12-16 review_article.Rmd
critique of choropleths

@@
 
-```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "A selection of choropleth cancer maps available online.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "A selection of choropleth cancer maps from online atlases that are publicly available.
+Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
 library(cowplot)
@@
 
+Australia presents an extreme case of an urban rural divide.
+The land mass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas [@ACTUC].
+This impedes reading information about disease incidence from choropleth maps.
+
+Choropleth maps provide a familiar display, which shows data in a geographically recognisable way.
+A disadvantage is that the different population and geographical sizes of administrative areas can attract attention to the shades of the underpopulated but large areas [@EI].
+Skowronnek also [@BCM] discusses how choropleth maps suffer from area-size bias, as they give a 'stronger visual weight' to large  administrative units.
+The administrative boundaries used to define regions may limit a choropleth display, as this display unfaithfully represents the disease distribution across the region by obscuring small geographic areas.
+Sparsely populated rural areas are emphasized, whereas the areas representing inner city communities are very small.
+This is especially true for Australia.
+
+<!--
 Choropleth maps provide a familiar display that may be important for orienting a non-statistical audience.
@@
 To present election results on a choropleth map should be ‘unthinkable’, as it means diminishing the visual impact of most of the electorates and incorrectly communicating the distribution of votes across the country.
+-->
 
+Inset maps like in Brisbane city in Figure 1<!--\ref{fig:choropleth-grid-create}-->c of the state of Queensland are commonly used to reduce distorted interpretations, but it is a bandaid remedy.
+For Australia, many, many inset maps would be needed.
+
+<!--
 Choropleth maps can inhibit visual inference when presenting human-related statistics as the display may draw attention away from the ‘potentially more important results in the more populous communities’ that are geographically smaller [@SE].
@@
 As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
+-->
 

### 2019-12-16 review_article.Rmd
added package reference

--- 2019-12-16 review_article.Rmd
@@
 
-## Limitations of different choices
+## Limitations of alternative displays
 
@@
 However, the transformation of contiguous cartograms often occurs at the expense of the shape of areas [@CBATCC, @NAC, @TAAM].
-When the population density pf the geographic units is highly dissonant with geographic density, the cartogram will lose all spatial context.
+When the population density of the geographic units is highly dissonant with geographic density, the cartogram will lose all spatial context.
 Dorling [@ACTUC] has a cartogram showing the 1966 general election results, which looked very little like the geographical shape of Australia.
@@
 
-The following R [@R] packages were used to produce this paper: tidyverse [@tidyverse], RColorBrewer [@RColorBrewer], ggthemes [@ggthemes], png [@png], cowplot [@cowplot], sf [@sf], spData [@spData], cartogram [@cartogram], sugarbag [@sugarbag], knitr [@knitr], rmarkdown [@rmarkdown].
+The following R [@R] packages were used to produce this paper: tidyverse [@tidyverse], RColorBrewer [@RColorBrewer], ggthemes [@ggthemes], png [@png], cowplot [@cowplot], sf [@sf], spData [@spData], cartogram [@cartogram], sugarbag [@sugarbag], knitr [@knitr], rmarkdown [@rmarkdown] and absmapsdata [@ABSmaps].
 

### 2019-12-17 review_article.Rmd
updated plots

@@
 
+## Cancer atlases {#ch:public}
+<!-- 360 words -->
+<!-- Displaying of cancer data on choropleth maps -->
+A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
+Atlases are key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work-related exposures, and high-risk diets [@MACM].
+The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization.
+Cancer maps and atlases dating back to Haviland's maps in 1875, and early work in US cancer atlases appearing in 1971 [@burbank].
+The presentation of cancer statistics has increased with greater access to computational power and the availability of geographic information systems software [@SE].
+
+<!-- Measures reported in cancer atlases -->
+Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the public and others not trained in statistical analysis.
+These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analysis.
+Cruickshank (1947) as cited by S.
+D.
+Walter [@DMAHP], discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
+Overwhelmingly, choropleth maps are visulisation chosen to communicate cancer statistics to members of the public and other non-expert audiences.
+
 ```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "A selection of choropleth cancer maps from online atlases that are publicly available.
-Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
+Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends.
+These atllases are described in Table 1.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
 library(cowplot)
@@
 
-## Cancer atlases {#ch:public}
-<!-- 360 words -->
-<!-- Displaying of cancer data on choropleth maps -->
-A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
-Atlases are key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work-related exposures, and high-risk diets [@MACM].
-The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization.
-Cancer maps and atlases dating back to Haviland's maps in 1875, and early work in US cancer atlases appearing in 1971 [@burbank].
-The presentation of cancer statistics has increased with greater access to computational power and the availability of geographic information systems software [@SE].
+Table: 1<!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases.
 
-<!-- Measures reported in cancer atlases -->
-Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the public and others not trained in statistical analysis.
-These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analysis.
-Cruickshank (1947) as cited by S.
-D.
-Walter [@DMAHP], discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
-Overwhelmingly, the cancer maps used to communicate to members of the public and other non-expert audiences are choropleth maps.
+| Fig. | Atlas | Statistic | Data source | 
+|:----:|:-------------|:----------------------|:----------------|  
+|1a | The Environment and Health Atlas of England and Wales | relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2]| [ Office for National Statistics (ONS) (England) and from the Welsh Cancer Intelligence and Surveillance Unit (WCISU) ](http://www.envhealthatlas.co.uk/homepage/)|
+|1b | Globocan 2018: Estimated Cancer Incidence, Mortality and Prevalence Worldwide | age standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018 [@Globocan]| [World Health Organization's International Agency for Research on Cancer](https://www.iarc.fr/).
+|
+|1c | Atlas of Cancer in Queensland | the relative incidence ratio of lung cancer in males in the state of QLD within Australia based on data from 1998 to 2007, Queensland Cancer Council[@QLDcancerAtlas] | [Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/) |
+|1d | Bowel Cancer Australia Atlas | the percentage of Australian males between 50 - 54 years of age diagnosed with bowel cancer in 2016 in Australia [@Bowel].
+|  [Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)|
+|1e | United States Cancer Statistics: An Interactive Cancer Statistics Website | the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level [@USInteractive].|  *[Centers for Disease Control and Prevention](https://www.cdc.gov/)*, with data from state cancer registries.
+|
+|1f | Map of Cancer Mortality Rates in Spain | side by side maps of relative risk of lung cancer for men vs women for 2004 to 2008 [@cancerSpain].| [Map of cancer mortality rates in Spain](https://elpais.com/elpais/2014/10/06/media/1412612722_141933.html)|
+|1g | Atlas of Childhood Cancer in Ontario | the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
+| [The Pediatric Oncology Group of Ontario Networked Information System](https://www.pogo.ca/research-data/pogonis-childhood-cancer-database/data-requests/)|
 
 Epidemiologists and statisticians have developed the statistics used to communicate the burden of cancer over several decades.
-Table 1<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases.
+Table 2<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases.
 Mortality rates are commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations.
@@
 
-Table: 1<!--\label{tab:measures}--> Common measures for reporting cancer information.
+Table: 2<!--\label{tab:measures}--> Common measures for reporting cancer information.
 
@@
 Figure 1<!--\ref{fig:choropleth-grid-create}--> displays a subset of maps from these atlases, the selection varies in the geographies explored.
-Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows Globocan 2012 [@Globocan] which explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide using data sourced from cancer registries of each country.
+Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows Globocan 2018 [@Globocan] which explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide using data sourced from cancer registries of each country.
 The Bowel Cancer Australia Atlas in Figure 1<!--\ref{fig:choropleth-grid-create}-->d presents an example of a cancer specific atlas, it shows the average Standardized Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia [@Bowel].
@@
 The Environment and Health Atlas of England and Wales, Figure 1<!--\ref{fig:choropleth-grid-create}-->a shows the relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2] at a neighborhood (small-area) scale.
-The Atlas of Cancer in Queensland, Figure 1<!--\ref{fig:choropleth-grid-create}-->c shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 in the state of Queensland within Australia [@QLDcancerAtlas].
+The Atlas of Cancer in Queensland, Figure 1<!--\ref{fig:choropleth-grid-create}-->c shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2016] in the state of Queensland within Australia [@QLDcancerAtlas].
 
@@
 
-Australia presents an extreme case of an urban rural divide.
-The land mass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas [@ACTUC].
-This impedes reading information about disease incidence from choropleth maps.
+Australia presents an extreme case of an urban rural divide. The land mass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas [@ACTUC].
 
@@
 This is especially true for Australia.
+
+Choropleth maps colour each geographic unit to allow map users to measure the value of the statistic [@EI].
+Map users contrast the colours in neighbouring areas to understand the spatial distribution.
+Map creators must choose an attractive and effective colour scheme, but is a difficult task [@CB].
+The ColorBrewer system supports map creators in the decision making process, but this does not consider perceptual colour
+specifications.
+
+Brewer [@CB] also suggests the use and colors of borders and backgrounds can change the appearance of the colors representing the value of the statistics.
+Incorporating borders may allow a reference point to orient users as they compare colours of neighbouring areas.
 
@@
 
-Table: 2<!--\label{tab:usa}--> Maps used to present statistics for the United States of America.
+Table: 3<!--\label{tab:usa}--> Maps used to present statistics for the United States of America.
 The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
@@
 
-Figure 2<!--\ref{fig:ggcartograms}--> shows four different cartograms for the same data. The information in Table 2<!--\ref{tab:usa}--> summarizes what can be observed in the four types of cartograms.
+Figure 2<!--\ref{fig:ggcartograms}--> shows four different cartograms for the same data. The information in Table 3<!--\ref{tab:usa}--> summarizes what can be observed in the four types of cartograms.
 
@@
 
-Figure 3<!--\ref{fig:auscartograms}--> shows four main types of cartograms displaying melanoma incidence on Statistical Areas at Level 3 [@ABSmaps] areas in Australia.
+Figure 3<!--\ref{fig:auscartograms}--> shows four main types of cartograms displaying melanoma incidence on Statistical Areas at Level 3 [@abs2016] in Australia.
 The version of a contiguous cartogram (a) has expanded the highly populated areas while preserving the full shapes of rural areas.
@@
 Some mix of tiling, faceting or even micro maps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies.
-Table \ref{tab:methods} summarizes the key criteria for testing maps and alternative displays.
+Table 3 <!--\ref{tab:methods}--> summarizes the key criteria for testing maps and alternative displays.
 Moore and Carpenter [@SAMGIS] and Bell et al.
@@
 
-Table: \label{tab:methods} Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
+Table: 3<!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
 

### 2019-12-17 sa3_plots.R

@@ -58,17 +58,13 @@ cimar_sa3 <- cimar_sa3 %>%
 sa3lung <- sa3lung %>% 
   left_join(cimar_sa3, by = c("sa3_name_2016" = "Name")) %>% 
   st_as_sf() %>%
-  #filter(!is.na(Population)) %>%
-  mutate(Population = as.numeric(ifelse(is.na(Population), 1, Population)),
-         `Age-standardised rate (per 100,000)` = as.numeric(ifelse(is.na(`Age-standardised rate (per 100,000)`),
-                                                        1, `Age-standardised rate (per 100,000)`))) %>% 
+  filter(!is.na(Population)) %>%
+  mutate(`Age-standardised rate (per 100,000)` = as.numeric(ifelse(is.na(`Age-standardised rate (per 100,000)`), 1, `Age-standardised rate (per 100,000)`))) %>% 
   filter(!st_is_empty(geometry))
 
 sa3lung <- st_transform(sa3lung, 3112)
 
 b <- st_bbox(sa3lung)
-b["xmin"]<- -2181807
-b["xmax"]<- 1933081
 
 aus <- absmapsdata::state2016 %>% 
   filter(!(state_name_2016 == "Other Territories"))
@@ -111,6 +107,7 @@ cont <- sa3lung %>%
 save(cont, "data/auscont.rda")
 load("data/auscont.rda")
 aus_ggcont <- ggplot(cont) + 
+  geom_sf(data=sa3lung, fill = NA, colour = "grey", size = 0.5) +
   geom_sf(aes(fill = `Age-standardised rate (per 100,000)`)) + 
   scale_fill_distiller(type = "seq", palette = "Purples",  direction = 1) + 
   coord_sf(crs = CRS("+init=epsg:3112"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
@@ -129,31 +126,34 @@ sa3lung %>%
   ggplot(.) +
   geom_density(aes(x = sva)) + geom_vline(aes(xintercept = 7))
 
+
 ncont <- cartogram_ncont(sa3lung, k = 1/5,
   weight = "Population") %>% st_as_sf() %>% 
   rename(`Age-standardised rate (per 100,000)` = `Age.standardised.rate..per.100.000.`)
-aus_ggncont <- ggplot(ncont) + 
-  geom_sf(data=aus, fill = NA, colour = "black", size = 0.1) +
[…]

### 2019-12-17 review_article.Rmd
Inset map completed

@@
 
-| Fig. | Atlas | Statistic | Data source | 
-|:----:|:-------------|:----------------------|:----------------|  
+|Fig.| Atlas | Statistic | Data source | 
+|:---:|:------------|:----------------------|:----------------|  
 |1a | The Environment and Health Atlas of England and Wales | relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2]| [ Office for National Statistics (ONS) (England) and from the Welsh Cancer Intelligence and Surveillance Unit (WCISU) ](http://www.envhealthatlas.co.uk/homepage/)|

### 2019-12-17 sa3_plots.R

@@ -111,7 +111,7 @@ aus_ggcont <- ggplot(cont) +
   geom_sf(aes(fill = `Age-standardised rate (per 100,000)`)) + 
   scale_fill_distiller(type = "seq", palette = "Purples",  direction = 1) + 
   coord_sf(crs = CRS("+init=epsg:3112"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
-  invthm +  guides(fill=FALSE)
+  invthm + guides(fill=FALSE)
 aus_ggcont
 ggsave(filename = "figures/aus_ggcont.png", device = "png",  bg = "transparent", dpi = 300,  width = 7, height = 6)
 
@@ -127,14 +127,16 @@ sa3lung %>%
   geom_density(aes(x = sva)) + geom_vline(aes(xintercept = 7))
 
 
-ncont <- cartogram_ncont(sa3lung, k = 1/5,
+ncont <- cartogram_ncont(sa3lung, k = 1/3,
   weight = "Population") %>% st_as_sf() %>% 
   rename(`Age-standardised rate (per 100,000)` = `Age.standardised.rate..per.100.000.`)
 
 goldfields_box <- st_bbox(st_buffer(ncont %>% filter(sa3_name_2016 == "Goldfields"), dist=500))
 
 aus_ggncont <- ggplot(ncont) +  
-  geom_sf(aes(fill = `Age-standardised rate (per 100,000)`), colour = NA) + geom_rect(aes(xmin =goldfields_box[1] , xmax = goldfields_box[3] , ymin=goldfields_box [2], ymax=goldfields_box [4]), fill = NA, colour = "black") + 
+  geom_sf(aes(fill = `Age-standardised rate (per 100,000)`), colour = NA) + 
+  geom_rect(aes(xmin =goldfields_box[1], xmax = goldfields_box[3], 
+                ymin=goldfields_box[2], ymax=goldfields_box[4]), fill = NA, colour = "black") + 
   geom_sf(data=sa3lung, fill = NA, colour = "grey", size = 0.5) +
   scale_fill_distiller(type = "seq", palette = "Purples",  direction = 1) + 
   coord_sf(crs = CRS("+init=epsg:3112 +units=m"), xlim =
@@ -151,9 +153,9 @@ goldfields_ggncont
 library(cowplot)
[…]

### 2019-12-17 Response/Response.Rmd
colour scales

@@
 **Traditional approaches for cancer map displays:** This section discusses the choropleth map as the common map display used in cancer maps and atlases.
-*Limitations of current approaches:* This section presents the possible issues with displaying cancer statistics on a choropleth map.
-*Alternative visualisation methods:* We have shown that there are many alternative displays that highlight population dense regions.
+*Limitations of choropleth displays:* This section presents the possible issues with displaying cancer statistics on a choropleth map.
+*Contemporary alternatives to choropleth maps:* We have shown that there are many alternative displays that highlight population dense regions.
 *Challenges of alternative approaches:* Alternative displays also have limitations when displaying spatial distributions of cancer statistics.
@@
 *Changes in the text:*
+
 Cancer atlases communicate cancer statistics over geographic domains.
-These domains are subdivided by administrative areas such as countries, states or suburbs.
-When communicating human-related statistics in Australia, the geographic map base highlights the sparsely populated rural areas.
-The smaller geographic areas may not draw the attention of readers but they are important to consider if they are more densely populated.
-Alternative map displays can reveal patterns that are not obvious when the same data is shown on a choropleth map.
-Alternative displays have proven to be successful on online news sites, especially for the communication of election data, where votes are aggregated at the state level.
+They subdivide these domains into administrative regions such as countries, states, or suburbs.
+When communicating human-related statistics in Australia, the geographic map base draws attention to sparsely populated rural areas.
+The smaller geographic areas may not draw the attention of readers, but they are important to consider if they are more densely populated.
+Alternative map displays can be more effective than a choropleth map at revealing spatial patterns.
+Alternative displays have rarely been used in the context of online cancer atlases, but the Globocan atlas, [@Globocan] provides an example of how alternative statistical displays can be implemented.
 
@@
 *Reply 2:* 
-See *Reply 1* for changes to the abstract. Additionally, the conclusion has also been edited to be more direct in concluding the alternatives presented.
+Changes to the abstract have been made. Additionally, the conclusion has also been edited to be more direct in concluding the alternatives presented.
 
 *Changes in the text:*
+
+Cancer atlases communicate cancer statistics over geographic domains.
+They subdivide these domains into administrative regions such as countries, states, or suburbs.
+When communicating human-related statistics in Australia, the geographic map base draws attention to sparsely populated rural areas.
+The smaller geographic areas may not draw the attention of readers, but they are important to consider if they are more densely populated.
+Alternative map displays can be more effective than a choropleth map at revealing spatial patterns.
+Alternative displays have rarely been used in the context of online cancer atlases, but the Globocan atlas, [@Globocan] provides an example of how alternative statistical displays can be implemented.
 
@@
 
-*Reply :*
+*Reply 3:* 
+Thank you for this point. Subtantial time has been taken to improve the language used.
 

### 2019-12-17 review_article.Rmd
colour scales

@@
 Alternative map displays can be more effective than a choropleth map at revealing spatial patterns.
-Alternative displays are effective for the communication of election data aggregated at the state or electorate level on online news sites.
+Alternative displays have rarely been used in the context of online cancer atlases, but the Globocan atlas, [@Globocan] provides an example of how alternative statistical displays can be implemented.
 
@@
 Mortality rates are commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations.
-Howe [@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s.
+Stock developed the standardized mortality ratios through the 1930s, this is decribed by Howe [@HEDP].
 
@@
 Similarly, socioeconomic indicators, such as unemployment rates, poverty rates, remoteness, and education levels, can explain how the experience of cancer prevalence varies for various members of society.
-Few atlases provide this
-level of detail.
+Few atlases provide this level of detail.
+
 Introducing population and demographic information helps to interpret the rates in areas effectively, but there will still be uncertainty around the rates.
@@
 Map creators must choose an attractive and effective colour scheme, but is a difficult task [@CB].
-The ColorBrewer system supports map creators in the decision making process, but this does not consider perceptual colour
-specifications.
+The ColorBrewer system supports map creators in the decision making process, but this does not consider perceptual colour specifications.
+When communicating information using colour, a map creator should use a scheme that has a linear color gradient, with perceptually uniform color spaces to convey equal steps in data align with equal steps in the colour scheme [@PUCS].
 

### 2019-12-17 review_article.Rmd
colour section

@@
 Map users contrast the colours in neighbouring areas to understand the spatial distribution.
-Map creators must choose an attractive and effective colour scheme, but is a difficult task [@CB].
-The ColorBrewer system supports map creators in the decision making process, but this does not consider perceptual colour specifications.
-When communicating information using colour, a map creator should use a scheme that has a linear color gradient, with perceptually uniform color spaces to convey equal steps in data align with equal steps in the colour scheme [@PUCS].
+Map creators must choose an effective colour scheme, but it is a difficult task [@CB].
+The ColorBrewer system supports map creators in the decision making process, but this does not full provide perceptual colour specifications.
+When communicating information using colour, a map creator should use a scheme that has a linear color gradient, with perceptually uniform color spaces to convey that equal steps in data align with equal steps in the colour scheme [@PUCS].
 
-Brewer [@CB] also suggests the use and colors of borders and backgrounds can change the appearance of the colors representing the value of the statistics.
-Incorporating borders may allow a reference point to orient users as they compare colours of neighbouring areas.
+The use and colors of borders and backgrounds can also change the appearance of the colors representing the value of the statistics [@CB].
+These supports are important, though, to allow a reference point to orient users to the geographic regions.
 

### 2019-12-17 review_article.Rmd
conclusions

@@
 <!-- summary of mapping -->
-This paper provides an overview of mapping practices as used for cancer atlases and new approaches that could be adopted.
+This paper provides an overview of mapping practices as commonly used for cancer atlases and recommends new approaches, such as cartograms and hexagon tile maps that should be adopted going forward.
 The conventional approach is the choropleth map, and it is widely used.
@@
 <!-- interaction -->
-Interaction with maps is an important component of public atlases.
+Interaction with maps is an important component of public atlases, and is easy to add with today's technology.
 A purpose is to provide access to more information than is possible to display in a single map, without overwhelming the viewer.

### 2019-12-18 review_article.Rmd
abstract

@@
 
-Cancer atlases communicate cancer statistics over geographic domains.
+Cancer atlases communicate cancer statistics over geographic domains, typically with a choropleth map.
 They subdivide these domains into administrative regions such as countries, states, or suburbs.
-When communicating human-related statistics in Australia, the geographic map base draws attention to sparsely populated rural areas.
-The smaller geographic areas may not draw the attention of readers, but they are important to consider if they are more densely populated.
-Alternative map displays can be more effective than a choropleth map at revealing spatial patterns.
-Alternative displays have rarely been used in the context of online cancer atlases, but the Globocan atlas, [@Globocan] provides an example of how alternative statistical displays can be implemented.
+When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner city areas.
+The smaller geographic areas are important to consider if they are densely populated.
+Alternative map displays, such as a cartogram or a hexagon tile map, can be more effective at accurately communicating spatial patterns.
+It is recommended that more use of these alternative displays is made.
+In addition, with the ease of today's technology, user interaction with the displays is encouraged.
+Users should also be able to interactively display different statistics, such as incidence rate or relative incidence.
 

### 2019-12-18 Response/Response.Rmd
changes with Di

@@
 
-6. Please clearly state the primary aim of this paper and structure it accordingly.
-
+6. Please clearly state the primary aim of this paper and structure it accordingl=['
+]
 *Reply :*
@@
 
-*Reply :*
+*Reply 7:* 
+Small changes have been made to the conclusion that make the focus of the paper and recommendations clearer.The conclusion has three parts (1) mapping practices, and recommendations, (2) staistics applied and the pros and cons, and (3) user interaction with online maps and the benefits of incorporating this.
 
-*Changes in the text:*
+*Changes in the text 7:*
 

### 2019-12-18 Response/Response.Rmd
Review article and responses finalised

--- 2019-12-17 Response/Response.Rmd
@@
 
+The manuscript title has been changed:
+"Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"
+
 The manuscript now contains the following sections:
@@
 
+The following section outlines comments made by the reviewer and contains our responses after the suggested changes were implemented.
+
 *Reviewer A: *
@@
 *Reply 1:* 
-We agree with the reviewer. Alternative mapping displays have been implemented in political contexts. Choropleth maps have often been used.
+The advantages of alternative mapping displays, including cartograms, has been included in the abstract.
+It clearly includes the changes to the map display made by alternative visualisation methods, and how they benefit the map user as they build an understanding of the spatial distribution.
 
@@
 
-Cancer atlases communicate cancer statistics over geographic domains.
+Cancer atlases communicate cancer statistics over geographic domains, typically with a choropleth map.
 They subdivide these domains into administrative regions such as countries, states, or suburbs.
-When communicating human-related statistics in Australia, the geographic map base draws attention to sparsely populated rural areas.
-The smaller geographic areas may not draw the attention of readers, but they are important to consider if they are more densely populated.
-Alternative map displays can be more effective than a choropleth map at revealing spatial patterns.
-Alternative displays have rarely been used in the context of online cancer atlases, but the Globocan atlas, [@Globocan] provides an example of how alternative statistical displays can be implemented.
+When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner city areas.
+The smaller geographic areas are important to consider if they are densely populated.
+Alternative map displays, such as a cartogram or a hexagon tile map, can shift the attention of map users from the large rural areas by decreasing their size on the map display.
+This means alternative displays can be more effective at accurately communicating spatial patterns across spatial areas, by highlighting areas of interest.
+It is recommended that alternative displays are included in cancer atlases.
+In addition, with the ease of today's technology, user interaction with the displays is encouraged.
+Users should also be able to interactively display different statistics, such as incidence rate or relative incidence.
 
@@
 *Reply 2:* 
-Changes to the abstract have been made. Additionally, the conclusion has also been edited to be more direct in concluding the alternatives presented.
+Changes to the abstract have been made. Additionally, the conclusion has also been edited to be more direct in recommending the use or consideration of the alternatives presented.
 
 *Changes in the text:* 
-
-Cancer atlases communicate cancer statistics over geographic domains.
-They subdivide these domains into administrative regions such as countries, states, or suburbs.
-When communicating human-related statistics in Australia, the geographic map base draws attention to sparsely populated rural areas.
-The smaller geographic areas may not draw the attention of readers, but they are important to consider if they are more densely populated.
-Alternative map displays can be more effective than a choropleth map at revealing spatial patterns.
-Alternative displays have rarely been used in the context of online cancer atlases, but the Globocan atlas, [@Globocan] provides an example of how alternative statistical displays can be implemented.
+Please see the response to *Comment 1*.
 
@@
 *Reply 3:* 
-Thank you for this point. Subtantial time has been taken to improve the language used.
-
-*Changes in the text:*
+Thank you for this point. Substantial time has been taken to improve the language used in this paper.
 
@@
 
-*Reply :*
-
-*Changes in the text:*
+*Reply:* 
+The referencing system was changed to implement the springer-vancouver style. We apologise for the issues this caused when exporting to a Word Document, the text has been modified appropriately.
 
@@
 
-*Reply :*
-
-*Changes in the text:*
+*Reply 5:* 
+this begins the discussion, andNumbers for figures and tables have been corrected.
+This paper was produced using Rmarkdown, this introduced errors with numbering figures, sections and tables.
+All numbering has now been manually defined.
 
@@
 
-6. Please clearly state the primary aim of this paper and structure it accordingly.
+6. Please clearly state the primary aim of this paper and structure it accordingly
 
-*Reply :*
-
-*Changes in the text:*
+*Reply 6:* 
+The primary aim of this paper is to present alternative map displays as viable visualisation methods for presenting spatial distributions in cancer atlases.
+A choropleth map is commonly used in practice, it is described and it's limitations are highlighted.
+The alternative map displays address the limitations of the choropleth map.
+The methods and the limitations of these alternative displays are discussed.
 
@@
 
-*Reply :*
-
-*Changes in the text:*
+*Reply 7:* 
+Small changes have been made to the conclusion that make the focus of the paper and recommendations clearer.
+The conclusion has three parts (1) mapping practices, and recommendations, (2) statistics applied and the pros and cons, and (3) user interaction with online maps and the benefits of incorporating this.
 
@@
 
-We thank you for your consideration.
+All minor comments were addressed in the process of substantial editing.
+
+We thank you for your consideration of our changes.

### 2019-12-18 review_article.Rmd
Review article and responses finalised

@@
 The smaller geographic areas are important to consider if they are densely populated.
-Alternative map displays, such as a cartogram or a hexagon tile map, can be more effective at accurately communicating spatial patterns.
-It is recommended that more use of these alternative displays is made.
+Alternative map displays, such as a cartogram or a hexagon tile map, can shift the attention of map users from the large rural areas by decreasing their size on the map display.
+This means alternative displays can be more effective at accurately communicating spatial patterns across spatial areas, by highlighting areas of interest.
+It is recommended that alternative displays are included in cancer atlases.
 In addition, with the ease of today's technology, user interaction with the displays is encouraged.
@@
 The Dorling cartogram presents all areas but many are difficult to compare.
-The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%', cache=FALSE}
+The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%', cache=FALSE, fig.width=8, fig.height=6, dpi=300}
 
@@
 library(cowplot)
-ggdraw() +
-  draw_plot(rasterGrob(aus1), 0, .55, 0.5, 0.50) +
-  draw_plot(rasterGrob(aus2), 0.5, 0.55, 0.5, 0.50) +
+ggdraw(xlim = c(0,1), ylim = c(0,1.2)) +
+  draw_plot(rasterGrob(aus1), 0, .65, 0.5, 0.50) +
+  draw_plot(rasterGrob(aus2), 0.5, 0.65, 0.5, 0.50) +
   draw_plot(rasterGrob(aus3), 0.0, 0.1, 0.5, 0.55) +
@@
     c(0, .5, 0, 0.5), 
-    c(1, 1, .5, .5), size = 20)
+    c(1.15, 1.15, .6, .6), size = 20)
 ```
@@
 Zooming in on a high-resolution output shows it does preserve the areas.
-The Dorling cartogram has lost all geographic context.
-The hexagon tile map provides a reasonable spatial distribution despite having too much white-space in the outback areas.
+The Dorling cartogram and the hexagon tile map provide reasonable displays of the spatial distribution, despite having too much white-space in the outback areas.
 
@@
 The conventional approach is the choropleth map, and it is widely used.
-When there are small areas, as occurs in Australia where the population is concentrated on the coast, the information about cancer can be lost, and alternatives are needed.
-Making an inset can clarify congested regions but this breaks the viewers' attention because they need to shift focus from the map to the inset, and if there are many congested areas, many insets would be needed.
-The map alternatives, like cartograms and their variations, can be useful to allow the spatial distribution of cancer data to be digested.
+The choropleth map suffers when there are small geographic units, as occurs in Australia where the population is concentrated on the coast, the information about the burden of cancer on those communities can be hidden.
+Making an inset can clarify congested regions but this breaks the viewers' attention as they shift focus from the map to the inset, and if there are many congested areas, many insets would be needed.
+The map alternatives implement tradeoffs between the familiar shapes, and the importance of the geographic areas in the context of the areas.
+Given the population or a cancer statistic for each area, the geographic size or shape will change.
+Alternative displays allow the spatial distribution of cancer data to be digested by map users.
 
 <!-- statistics -->
-Many different statistics are commonly used for cancer displays.
-The most basic is the incidence rate, which is easy to understand.
-It is common to see relative rates in many maps which measure how far a region is above or below the average.
+Many statistics are commonly used in cancer displays.
+The most basic is the incidence rate.
+It is common to see relative rates which measure how far a region is above or below the average.
 The purpose of using a relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates.
-However, we lose the incidence rate information and thus interpretability.
-A region might be much higher than average, but it may not be close to a health concern, because all regions have a low incidence.
+<!--However, we lose the incidence rate information and thus interpretability.
+--> A region might be much higher than average, but it may not be close to a health concern, because all regions have a low incidence.
+Supplementary materials can allow map users to recognise when this occurs.
 
@@
 Interaction with maps is an important component of public atlases, and is easy to add with today's technology.
-A purpose is to provide access to more information than is possible to display in a single map, without overwhelming the viewer.
-Too many choices though can similarly overwhelm a viewer, and thus decisions do need to be made about content to provide, or hide, for accurate and comprehensive communication of information.
+The purpose is to provide access to more information than is possible to display in a single map, without overwhelming the viewer.
+Too many choices can similarly overwhelm a viewer, and thus decisions do need to be made about content to provide for accurate and comprehensive communication of information.
 Similarly, providing ways for users to interact with the display encourages engagement, and creative, efficient, elegant, interactive tools elicit curiosity about the data.

### 2019-12-18 sa3_plots.R

@@ -153,9 +153,9 @@ goldfields_ggncont
 library(cowplot)
 full_ggncont <- ggdraw(aus_ggncont, xlim = c(0,1), ylim = c(0,1)) + 
   draw_plot(goldfields_ggncont, 0.38, 0.06, 0.2, 0.2) + 
-  draw_line(x = c(0.425, 0.385, 0.535), 
+  draw_line(x = c(0.405, 0.345, 0.555), 
             y = c(0.25, 0.44, 0.25)) + 
-  draw_line(x = c(0.425, 0.385), y = c(0.07, 0.44))
+  draw_line(x = c(0.405, 0.345), y = c(0.07, 0.44))
 full_ggncont
 ggsave(filename = "figures/aus_ggncont.png", plot = full_ggncont,
        device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

### 2019-12-18 review_article.Rmd
viridis reference

@@
 Map users contrast the colours in neighbouring areas to understand the spatial distribution.
-Map creators must choose an effective colour scheme, but it is a difficult task [@CB].
-The ColorBrewer system supports map creators in the decision making process, but this does not full provide perceptual colour specifications.
-When communicating information using colour, a map creator should use a scheme that has a linear color gradient, with perceptually uniform color spaces to convey that equal steps in data align with equal steps in the colour scheme [@PUCS].
-
-The use and colors of borders and backgrounds can also change the appearance of the colors representing the value of the statistics [@CB].
-These supports are important, though, to allow a reference point to orient users to the geographic regions.
+The ColorBrewer system [@CB] and viridis [@viridis] palettes provide effective colour schemes for qualitative, sequential and diverging data.
+When communicating information using colour, a map creator should use a scheme that has a linear color gradient, with perceptually uniform color spaces that match equal steps in data space with equal steps in the colour space [@PUCS].
+The use of borders and backgrounds, and their colours, can also change the appearance of the colors representing the value of the statistics [@CB].
+These supports can be used to implement a reference point in the colour scheme as well as orient users to the geographic regions.
 

### 2019-12-18 review_article.Rmd
formatting and image output

@@
     toc_depth: 3
-  word_document:
-    toc: no
-    toc_depth: '3'
+    dev: jpeg
+header-includes:
+    - \usepackage{setspace}\doublespacing
+    - \usepackage{fancyhdr}\pagestyle{fancy}\fancyhead[R]{\thepage}
 link-citations: yes
 fontsize: 11pt
+geometry: margin = 1.2in
 csl: springer-vancouver.csl
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+fig_path <- glue::glue("figs/")
+knitr::opts_chunk$set(echo = FALSE, 
+                      warning = FALSE,
+                      message = FALSE,
+                      fig.path = fig_path)
 library(ggplot2)
@@
 Atlases are key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work-related exposures, and high-risk diets [@MACM].
-The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization.
-Cancer maps and atlases dating back to Haviland's maps in 1875, and early work in US cancer atlases appearing in 1971 [@burbank].
+The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization.Cancer maps and atlases dating back to Haviland's maps in 1875, and early work in US cancer atlases appearing in 1971 [@burbank].
 The presentation of cancer statistics has increased with greater access to computational power and the availability of geographic information systems software [@SE].
@@
 Choropleth maps can inhibit visual inference when presenting human-related statistics as the display may draw attention away from the ‘potentially more important results in the more populous communities’ that are geographically smaller [@SE].
-Inset maps like in Brisbane city in Figure Figure 1<!--\ref{fig:choropleth-grid-create}-->c of the state of Queensland are commonly used to reduce distorted interpretations, but it is not always an appropriate solution.
+Inset maps like in Brisbane city in Figure Figure 1<!--\ref{fig:choropleth-grid-create}c of the state of Queensland are commonly used to reduce distorted interpretations, but it is not always an appropriate solution.
 As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
@@
 A contiguous cartogram alters the choropleth according to a statistic and maintains connectivity of the map regions.
-Min Ouyang and Revesz [@ACA] present three algorithms for creating value-by-area cartograms. They implement 'map deformation' to account for the value assigned to each area.
+Min Ouyang and Revesz [@ACA] present three algorithms for creating value-by-area cartograms.
+They implement 'map deformation' to account for the value assigned to each area.
 Other methods include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [@ACTUC], Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and Tikunov’s Line Integral Method, Constraint-Based Method (Kocmoud and House) [@CBATCC].
 
-Figure 2<!--\ref{fig:ggcartograms}--> a shows a population contiguous cartogram of the United States. All states are visible and the shape of the United States overall is still recognizable.
+Figure 2<!--\ref{fig:ggcartograms}--> a shows a population contiguous cartogram of the United States.
+All states are visible and the shape of the United States overall is still recognizable.
 In contrast, Figure 3<!--\ref{fig:auscartograms}--> a shows an Australian contiguous cartogram also based on population.
@@
 The squares are tessellated to create a grid.
-This method has been used broadly in the media by FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
 Each area is represented by a square of the same dimensions, each tile is usually one unit of measurement, this could be geographic regions such as states or population-based that use a consistent measure of population for each tile.
@@
 A similar method to a rectangular cartogram represents each geographic area using a square of the same dimensions.
-The squares tessellate to create a grid.
 There are online media sources using this method, these include [NPR](https://www.npr.org/2016/10/18/498406765/npr-battleground-map-hillary-clinton-is-winning-and-its-not-close), [FiveThirtyEight](https://projects.fivethirtyeight.com/2016-election-forecast/), [The Wall Street Journal](http://graphics.wsj.com/elections/2016/2016-electoral-college-map-predictions/), [The Washington Post](https://www.washingtonpost.com/graphics/politics/2016-election/50-state-poll/).
@@
 These maps group areas based on their value for one variable, and additional columns provide displays that contrast the areas in each group by other variables.
-The display juxtaposes choropleth maps and statistical plots;
-it shows one map per group of the key separating variable, in a row with each additional statistical plot.
-Linked micro maps predominantly use the choropleth map for displays of spatial relationships. These maps show spatial relationships by allotting spatial neighbors to the same group.
+The display juxtaposes choropleth maps and statistical plots; it shows one map per group of the key separating variable, in a row with each additional statistical plot.
+Linked micro maps predominantly use the choropleth map for displays of spatial relationships.
+These maps show spatial relationships by allotting spatial neighbors to the same group.
 It is one of several alternative displays that allow maps to become bivariate displays, commonly used to present both an estimate and the associated uncertainty.
@@
 Interactive design features, found in online cancer maps, include tool tips, drop-down menus, data selection, zooming, and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
-The use of these supports can be found in various online cancer maps [@roberts2019communication].
+The use of these supports can be found in various online cancer maps and are shown in Figure 5 [@roberts2019communication].
 
@@
 
-Figure 7 <!--\ref{fig:animating}--> shows two examples of more sophisticated interactive maps.
+Figure 6 <!--\ref{fig:animating}--> shows two examples of more sophisticated interactive maps.
 The Spanish Cancer map (left) contains a linked display between a choropleth map and time series plots of cancer change.

### 2019-12-18 review_article.Rmd
final proofing

@@
 ---
-title: "Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"
-author: "Stephanie Kobakian, Jessie Roberts and Dianne Cook"
-bibliography: reviewbib.bib
+title: Cancer Applications of Choropleth Maps, and the Potential of Cartograms and
+  Alternative Map Displays
+author: Stephanie Kobakian, Jessie Roberts and Dianne Cook
+csl: springer-vancouver.csl
 output:
+  word_document:
+    toc: no
+    toc_depth: '3'
   pdf_document:
+    dev: jpeg
     fig_caption: yes
@@
     toc_depth: 3
-    dev: jpeg
+geometry: margin = 1.2in
 header-includes:
@@
 fontsize: 11pt
-geometry: margin = 1.2in
-csl: springer-vancouver.csl
+bibliography: reviewbib.bib
 yolo: yes
@@
 Alternative map displays, such as a cartogram or a hexagon tile map, can shift the attention of map users from the large rural areas by decreasing their size on the map display.
-This means alternative displays can be more effective at accurately communicating spatial patterns across spatial areas, by highlighting areas of interest.
+This means alternative displays can be more effective at accurately communicating spatial patterns across spatial areas.
 It is recommended that alternative displays are included in cancer atlases.
 In addition, with the ease of today's technology, user interaction with the displays is encouraged.
-Users should also be able to interactively display different statistics, such as incidence rate or relative incidence.
+Users should also be able to interactively display different statistics, such as incidence rate or relative incidence, or filtered by demographic variables.
 
-# Introduction
+# 1. Introduction
 <!-- 436 words -->
@@
 Examples of typical geographical units include states, provinces, local government areas, and post/zip codes.
-It is easy to provide counts of the diagnoses or excess deaths within these areas.
-However, the amount of people living in each of the areas in a set of areas will be different.
-Calculating incidence rates involves dividing the count by the population for each area, for example, the rate per 100000 people.
-Counts and population datasets are collected routinely for public health reasons and may be made available to the general public as a service to the community.
+It is easy to provide counts or incidence rates of the diagnoses of these areas.
+This type of data is routinely collected for public health reasons and may be made available to the general public as a service to the community.
 
@@
 This paper considers the current visualization techniques to communicate statistics to the public and their applications to cancer statistics.
-Alternative approaches are also considered, and the limitations of the visualization methods, highlighting the differences and historic use of these displays.
+Alternative approaches are posed because they may be more effective than contemporary techniques.
+The limitations of the visualization methods, highlighting the differences and historic use of these displays is discussed.
 
@@
 Section 5 presents the limitations of the production and use of alternative displays.
-Disease maps are more useful when made interactive, Section 6 describes common options, along with a discussion of benefits and disadvantages.
+Disease maps are more useful when made interactive, and common options are described in Section 6, along with a discussion of benefits and disadvantages.
 
-# Traditional approaches for cancer map displays {#ch:choropleth_maps}
+# 2. Traditional approaches for cancer map displays {#ch:choropleth_maps}
 <!-- 381 words -->
@@
 
-## Cancer atlases {#ch:public}
+## 2.1 Cancer atlases {#ch:public}
 <!-- 360 words -->
@@
 A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
-Atlases are key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work-related exposures, and high-risk diets [@MACM].
-The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization.Cancer maps and atlases dating back to Haviland's maps in 1875, and early work in US cancer atlases appearing in 1971 [@burbank].
+Atlases are key to developing hypotheses regarding areas with unusually high rates, and geographic correlations [@MACM].
+The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization.
+Cancer maps and atlases date back to Haviland's maps in 1875, and early work in US cancer atlases appearing in 1971 [@burbank].
 The presentation of cancer statistics has increased with greater access to computational power and the availability of geographic information systems software [@SE].
@@
 Walter [@DMAHP], discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
-Overwhelmingly, choropleth maps are visulisation chosen to communicate cancer statistics to members of the public and other non-expert audiences.
+Overwhelmingly, choropleth maps are visualisations chosen to communicate cancer statistics to members of the public and other non-expert audiences.
 
-```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "A selection of choropleth cancer maps from online atlases that are publicly available.
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Figure 1: A selection of choropleth cancer maps from online atlases that are publicly available.
 Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends.
@@
 
-Table: 1<!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases.
+Table: Table 1: <!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases.
 
@@
 Mortality rates are commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations.
-Stock developed the standardized mortality ratios through the 1930s, this is decribed by Howe [@HEDP].
+As described in Howe [@HEDP], Englishman P.
+Stocks advanced the field of mortality statistics by introducing the standardized mortality ratios in the 1930s, which is an improvement on crude death rates.
 
-Table: 2<!--\label{tab:measures}--> Common measures for reporting cancer information.
+Table: Table 2: <!--\label{tab:measures}--> Common measures for reporting cancer information.
 
@@
 Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows Globocan 2018 [@Globocan] which explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide using data sourced from cancer registries of each country.
-The Bowel Cancer Australia Atlas in Figure 1<!--\ref{fig:choropleth-grid-create}-->d presents an example of a cancer specific atlas, it shows the average Standardized Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia [@Bowel].
+The Bowel Cancer Australia Atlas in Figure 1<!--\ref{fig:choropleth-grid-create}-->d presents an example of a cancer specific atlas -- it shows the average Standardized Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 [@Bowel].
 Like many of the atlases examined, there is a choice of gender displayed in the Bowel Cancer Atlas.
-An alternative approach used in the Map of Cancer Mortality Rates in Spain, Figure 1<!--\ref{fig:choropleth-grid-create}-->f shows side-by-side maps of the relative risk of lung cancer separately for men and women [@cancerSpain].
+Gender is displayed in side-by-side maps in the Map of Cancer Mortality Rates in Spain (Figure 1f)<!--\ref{fig:choropleth-grid-create}--> [@cancerSpain].
 
-There are vast differences in the geographies considered by online atlases. 
-Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows global age-standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018.
-The United States Cancer Statistics [@USInteractive] presented in Figure 1<!--\ref{fig:choropleth-grid-create}-->e contains the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level.
-The Environment and Health Atlas of England and Wales, Figure 1<!--\ref{fig:choropleth-grid-create}-->a shows the relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2] at a neighborhood (small-area) scale.
-The Atlas of Cancer in Queensland, Figure 1<!--\ref{fig:choropleth-grid-create}-->c shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2016] in the state of Queensland within Australia [@QLDcancerAtlas].
+Resolution of the maps varies greatly.
+Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows global information at a national level.
+The United States Cancer Statistics [@USInteractive] shows data aggregated at the state level.
+The Environment and Health Atlas of England and Wales [@EnvEnglandWales2] (Figure 1<!--\ref{fig:choropleth-grid-create}-->a) shows the relative risk for women developing lung cancer  at a neighborhood (small-area) scale.
+The Atlas of Cancer in Queensland (Figure 1<!--\ref{fig:choropleth-grid-create}-->c) shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2016] in the state of Queensland within Australia [@QLDcancerAtlas].
 
-Age-specific atlases are less common, Figure 1<!--\ref{fig:choropleth-grid-create}-->g displays Atlas of Childhood Cancer in Ontario, this communicates the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
-This atlas communicates information regarding a small geographic region.
+Age-specific atlases are less common.
+Figure 1<!--\ref{fig:choropleth-grid-create}-->g displays Atlas of Childhood Cancer in Ontario, this communicates the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
 
-## Additional considerations
+## 2.2 Additional considerations
 <!-- 479 words -->
@@
 World atlases can allow for displays of data aggregated into continents, countries, states, provinces and congressional districts [@USInteractive].
-Each population area will probably have a different number of people.
+Each population area will probably have a different number of people, which is typically used to calibrate the statistic.
 Cancer atlases may also communicate the distribution of the population living in all areas in a table or histogram display [@NICR_1].
 Atlases can connect the population to the land available to them by communicating population density.
-Section  4<!--\ref{ch:alternatives}--> considers alternatives to choropleth that introduce population into the visualizations.
 
-Map users can define the population further using the demographics of the individuals within each area.
-Demographics that can influence cancer rates include information regarding the age and sex distribution of the areas presented.
-Some digital atlases surveyed allow users to control the display to select subsets such as males, females or those aged over 65.
-Similarly, socioeconomic indicators, such as unemployment rates, poverty rates, remoteness, and education levels, can explain how the experience of cancer prevalence varies for various members of society.
+Maps can also be used to focus on demographic strata, such as age and sex.
+Some of the digital atlases surveyed allow subsets such as males, females, or those aged over 65, to be selected for display.
+Similarly, socioeconomic indicators, such as unemployment rates, poverty rates, remoteness, and education levels, can be used to filter data, in order to communicate how cancer prevalence varies for different members of society.
 Few atlases provide this level of detail.
@@
 Introducing population and demographic information helps to interpret the rates in areas effectively, but there will still be uncertainty around the rates.
-To address this, a cancer atlases often communicates uncertainty about the value of a statistic.
+To address this, a cancer atlas often communicates uncertainty about the value of a statistic.
 There are several potential sources of uncertainty: sampling error, errors arising from the disease reporting process (or data collection), and errors arising from the statistical modeling or simulation process.
-The most common measures used to represent uncertainty were credible or confidence intervals (CIs).
+The most common measures used to present uncertainty are credible or confidence intervals (CIs).
 Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it is difficult to display effectively.
 The map focuses on displaying the statistic and lacks additional space to represent the uncertainty.
-Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two approaches.
+Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two common solutions.
 
-# Limitations of choropleth displays {#ch:chorolimit}
+## 2.3 Limitations of choropleth displays {#ch:chorolimit}
 
@@
 
-# Contemporary alternatives to choropleth maps {#ch:alternatives}
+# 3 Contemporary alternatives to choropleth maps {#ch:alternatives}
 
-## Cartograms
+## 3.1 Cartograms
 <!-- 1496 words inc table-->
@@
 When a map creator desires a uniform population density of the map base, the purposeful distortion of the map space is beneficial.
-The "population distribution is often extremely uneven in former British colonies" [@ACTUC], and this makes the distortion necessary, [@CTTMB] population will then become a uniformly distributed background for the statistic presented [@GOINO].
-An area cartogram [@NAC], or population-by-area cartogram [@TAAM] occurs after the distortion of the geographical shape according to population.
-Event cartograms change the area of regions on a map depending on the amount of disease-related events, but this does not consider the effects of land area and population [@VSSDCUC].
+The "population distribution is often extremely uneven", making a distortion necessary so that population is more faithfully represented as a uniformly distributed background for the statistic to be presented [@ACTUC] [@CTTMB] [@GOINO].
+An area cartogram [@NAC], or population-by-area cartogram [@TAAM] is produced from the distortion of the geographical shape according to population.
+Event cartograms [@VSSDCUC] change the area of regions on a map depending on the amount of disease-related events, rather than population.
 
@@
 Cartograms provide an alternative visualization method for statistical and geographical information.
-The key difference between a choropleth and a cartogram is the desirable augmentation of the size, shape, or distance of geographical areas [@ACTUC].
-Monmonier [@HTLWM] suggests that map creators can use white lies to create useful displays by distorting the geometry and suppressing features.
-It is easy for the average person to disregard the impact of transformations used to create cartograms.
-One disadvantage of the conventional map is interpreting spatial patterns is difficult because of the sparsely populated rural areas being emphasized in a choropleth map, and the inner city areas are geographically small.
-The distortion of each area when making a cartogram is proportional to a value other than the actual earth size area [@NAC].
-The distortion in an area cartogram accounts for the population density, preventing it from obscuring the spatial patterns [@TAAM].
+Monmonier [@HTLWM] suggests that map creators can use white lies to create useful spatial displays.
+It is easy for the reader to disregard the impact of transformations used to create cartograms, for the benefit of reading the statistical distribution more accurately with approximate geographic information.
 The spatial transformation of map regions relative to the data emphasizes the data distribution instead of land size [@CBATCC].
-When visualizing population statistics, Dorling [@ACTUC] considers this design 'more socially just', or honest [@NISCC], giving equitable representation and attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
+When visualizing population statistics, Dorling  considers this design 'more socially just' [@ACTUC], or honest [@NISCC], giving equitable representation and attention to all members of the population and reducing the visual impact of large areas with small populations [@DMAHP].
 Howe [@HEDP] suggests that 'cancer occurs in people, not in geographical areas' and that spatial socio-economic data, like cancer rates, are best presented on a cartogram for urban areas as the population map base avoids allocating 'undue prominence' to rural areas [@CTTMB].
@@
 
-```{r ggcartograms, fig.cap = "Common alternatives to maps, showing the same information for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous).
+```{r ggcartograms, fig.cap = "Figure 2: Common alternatives to maps, showing the same information for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous).
 Maps (a) - (c) are created by resizing and reshaping the states of the USA to match the 2015 population of the state.
@@
 
-Table: 3<!--\label{tab:usa}--> Maps used to present statistics for the United States of America.
+Table: Table 3: <!--\label{tab:usa}--> Maps used to present statistics for the United States of America.
 The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
@@
 
-### Contiguous
+### 3.1.1 Contiguous
 
@@
 They implement 'map deformation' to account for the value assigned to each area.
-Other methods include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [@ACTUC], Radial Expansion Method of Selvin et al., Rubber Sheet Method of Dougenik et al., Gusein-Zade and Tikunov’s Line Integral Method, Constraint-Based Method (Kocmoud and House) [@CBATCC].
+Other methods include Tobler’s Pseudo-Cartogram Method, Dorling’s Cellular Automaton Method [@ACTUC], Radial Expansion Method, Rubber Sheet Method, Line Integral Method, Constraint-Based Method [@CBATCC].
 
-Figure 2<!--\ref{fig:ggcartograms}--> a shows a population contiguous cartogram of the United States.
+Figure 2a<!--\ref{fig:ggcartograms}--> shows a population contiguous cartogram of the United States.
 All states are visible and the shape of the United States overall is still recognizable.
 In contrast, Figure 3<!--\ref{fig:auscartograms}--> a shows an Australian contiguous cartogram also based on population.
-The south east is completely distorted, areas with low population are still large very the map, as their initial area was used to determine the area allocated in the contiguous cartogram.
+The south east is enlarged, but high population areas are still small, and low population areas are still large on the map.
+The algorithm doesn't fully reach an optimal configuration where area matches population -- Australia is too heterogeneous for the algorithm to handle.
 
@@
 
-### Non-contiguous
+### 3.1.2 Non-contiguous
 
@@
 Olson [@NAC] discusses the creation of these maps and the significance of the empty areas left between the geographic boundaries and the new shape.
-It is the 'difference from the original map that is the real message' of these displays.
 
-The distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition, and the non-contiguous cartogram eliminates this hurdle, and presents the meaningful empty-space property [@ECGC, @NAC].
+The white space presents the meaningful empty-space property [@ECGC] [@NAC] but it also distracts the reader from the data, with a low data density [@tufte].
 
-### Dorling
+### 3.1.3 Dorling
 
-Daniel Dorling presents an alternative display engineered to highlight the spatial distribution and neighborhood relationships without complex distortions of borders and boundaries.
-Daniel Dorling puts forward a simple question and his approach opposes preserving the intricate shape details [@ACTUC]:
+Daniel Dorling presents an alternative display engineered to highlight the spatial distribution and neighborhood relationships without complex distortions of borders and boundaries [@ACTUC]:
 
@@
 He acknowledged the sophistication of contiguous cartograms but critiqued their 'very complex shapes,' he answers this with his implementation of maps created using 'the simplest of all shapes'.
-Circular cartograms use the same simple shapes for every region represented and resize the shapes according to the statistic represented or the population for a base map.
-This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms.
-Contiguous cartograms use a gravity model to apply distortions to create the 'nonsense' shapes that have 'no meaning' [@NISCC].
-This method applies a gravity model to produce a compelling map, that avoids overlaps and keep spatial relationships with neighboring areas over many iterations.
-This implementation can work for up to ‘one hundred thousand’ areas.
+Circular cartograms use the same circle shape for every region represented, resized  according to the statistic represented or the population.
+This simple shape may be more effective for understanding the spatial distribution than contiguous cartograms.
+Contiguous cartograms create 'nonsense' shapes that have 'no meaning' [@NISCC].
+Both methods applies a gravity model to produce a layout, that avoids overlaps and keep spatial relationships with neighboring areas over many iterations.
+The circular cartogram is relatively fast to compute.
 
-Raisz laid the groundwork for this approach in the mid-1930s, and his rectangular cartograms provide dramatic comparisons and are especially useful for correcting misconceptions communicated by geographic maps.
-@TFYCC quotes the official definition of Value-Area Cartograms, the simplistic displays which represent each area as a single rectangle, sized according to the value of the statistic.
+Raisz [@RSCW] laid the groundwork for this approach in the mid-1930s, drawing rectangular cartograms that provide simple comparisons, effective for correcting misconceptions communicated by geographic maps.
+Tobler @TFYCC names and defines these as Value-Area Cartograms.
 This rectangular display may sacrifice contiguity but allows for tiling where geographic neighbors placed in suitable relative positions also share borders [@CDWCS].
-Rectangular cartograms communicate bi-variate displays of the population by the size of each rectangular, and they use color to communicate a second variable [@ORC].
+Rectangular cartograms communicate bivariate displays of the population by the size of each rectangular, and they use color to communicate a second variable [@ORC].
 
-## Tile Map
+## 3.2 Tile Map
 <!-- 173 words -->
@@
 The squares are tessellated to create a grid.
-Each area is represented by a square of the same dimensions, each tile is usually one unit of measurement, this could be geographic regions such as states or population-based that use a consistent measure of population for each tile.
-Regions with over four neighbors require some necessary displacement.
-The tile map uses color to represent a value of a statistic for each area.
-A similar method to a rectangular cartogram represents each geographic area using a square of the same dimensions.
-There are online media sources using this method, these include [NPR](https://www.npr.org/2016/10/18/498406765/npr-battleground-map-hillary-clinton-is-winning-and-its-not-close), [FiveThirtyEight](https://projects.fivethirtyeight.com/2016-election-forecast/), [The Wall Street Journal](http://graphics.wsj.com/elections/2016/2016-electoral-college-map-predictions/), [The Washington Post](https://www.washingtonpost.com/graphics/politics/2016-election/50-state-poll/).
-Tile maps may be difficult to create as they are best created manually, they require additional time and care as the number of geographic areas to include increases.
+Each area is represented by a square of the same dimensions, each tile is usually one unit of measurement.
+This could be geographic regions such as states or population-based that use a consistent measure of population for each tile.
+Regions with more than four neighbors introduce some displacement, because only four squares can be adjoining.
+Like the chorpleth and cartograms, the tile map uses color to represent a value of a statistic for each area.
+There are online media sources using this method, which include [NPR](https://www.npr.org/2016/10/18/498406765/npr-battleground-map-hillary-clinton-is-winning-and-its-not-close), [FiveThirtyEight](https://projects.fivethirtyeight.com/2016-election-forecast/), [The Wall Street Journal](http://graphics.wsj.com/elections/2016/2016-electoral-college-map-predictions/), [The Washington Post](https://www.washingtonpost.com/graphics/politics/2016-election/50-state-poll/).
+Tile maps often need to be created manually, requiring more time and care as the number of geographic areas increases.
 <!-- cite mosaic cartograms -->
@@
 Cano and others [@MDAC] define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area or the color of them can communicate the statistic of regions.
-When using several tile per regions, map makers can adjust the complexity of the boundaries in the resulting display.
+When using several tiles per region, map makers can adjust the complexity of the boundaries in the resulting display.
 They can also make a trade-off between boundary complexity and simplicity by the size of the tiles used.
 
-## Geofacet
+## 3.3 Geofacet
 <!-- 166 words -->
-Hafen [@IGF] formalizes the term geofaceting to describe a grid display, the arrangement of tiles to create a grid that mimics the geographic topology of the set of areas.
-Like tile maps, they can reuse the arrangement of the areas for other spatial distribution visualizations.
-Geofaceting has the functionality of facets, often used to replicate visualizations for each subset of the data.
-Here, the data subsets are geographic.
-A tile map communicates one value per region in a visualization, geofaceting is a more flexible visualization for communication as it increases the amount of information displayed.
-Virtually any plot display can be shown in the tile representing the areas, allowing displays of multiple variables or values per geographic entity.
-Creating a tile map or a geofacet both require manual creation of the grid, except for the grids that other have already contributed.
+Hafen [@IGF] introduces the term geofacet to describe a grid display of small plots.
+The arrangement of tiles mimics the geographic topology.
+Geofaceting has the functionality that a statistical plot can be constructed in each facet for each geographic area.
+A tile map can communicate only one value per region in a visualization, while geofaceting is a more flexible visualization for communication as it increases the amount of information displayed.
+Virtually any type of plot can be shown in the tile, allowing displays of multiple variables or values per geographic entity.
+Creating the layout of a geofacet is manual, but once created can be used for any data on that geographic base.
 
-```{r ggtilemap, fig.cap = "Two alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
+```{r ggtilemap, fig.cap = "Figure 3: Two alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
 In the tile map, the layout approximates spatial location, with each state being an equal box filled with color representing cancer incidence.
@@
 
-## Multivariate displays
+## 3.4 Multivariate displays
 <!-- 222 words -->
 
-Pickle and others [@MMST] present linked micro map plots to match geographic and statistical data visually,  this serves as a solution to  multi-dimensionality issues.
-An online atlas of US states uses this display.
+Pickle and others [@MMST] present linked micromap plots to match geographic and statistical data visually,  this serves as a solution to  multi-dimensionality issues.
 These maps group areas based on their value for one variable, and additional columns provide displays that contrast the areas in each group by other variables.
 The display juxtaposes choropleth maps and statistical plots; it shows one map per group of the key separating variable, in a row with each additional statistical plot.
-Linked micro maps predominantly use the choropleth map for displays of spatial relationships.
+Linked micromaps predominantly use the choropleth map for displays of spatial relationships.
 These maps show spatial relationships by allotting spatial neighbors to the same group.
@@
 
-Bivariate choropleth maps [@VUADBC] blend color schemes to convey the intersection of categorized levels of an estimate and the associated uncertainty for each spatial area.
-Lucchesi and Wikle [@VUADBC] also suggest map pixilation, it breaks each region into small pixels, and  allocates values to the individual pixels to create texture.
+Lucchesi and Wikle [@VUADBC] present bivariate choropleth maps blend color schemes to convey the intersection of categorized levels of an estimate and the associated uncertainty for each spatial area.
+They also suggest map pixilation, which breaks each region into small pixels, and allocates values to the individual pixels to create texture.
 This reflects the uncertainty around the area's estimate by randomly sampling from the confidence interval of the estimate of the area.
@@
 
-# Comparison and critique of alternative displays {#sec:compare}
+# 4 Comparison and critique of alternative displays {#sec:compare}
 <!-- 408 words -->
 
-```{r auscartograms, fig.cap = "Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map.
+```{r auscartograms, fig.cap = "Figure 4: Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map.
 The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas.
@@
 
-## Neither choropleth maps or cartograms perform well for Australia
+## 4.1 Neither choropleth maps or cartograms perform well for Australia
 
-Figure 3<!--\ref{fig:auscartograms}--> shows four main types of cartograms displaying melanoma incidence on Statistical Areas at Level 3 [@abs2016] in Australia.
+Figure 4<!--\ref{fig:auscartograms}--> shows four main types of cartograms using melanoma incidence on Australian Statistical Areas at Level 3 [@abs2016].
 The version of a contiguous cartogram (a) has expanded the highly populated areas while preserving the full shapes of rural areas.
@@
 The shape-preserved cartogram is unreadable, and it has reduced all areas to tiny spots on the map.
-Zooming in on a high-resolution output shows it does preserve the areas.
+Zooming in on a high-resolution output shows it does preserve the shapes.
 The Dorling cartogram and the hexagon tile map provide reasonable displays of the spatial distribution, despite having too much white-space in the outback areas.
 
-## Limitations of alternative displays
+## 4.2 Limitations of alternative displays
 
-Cartograms provide the spatial distortion to more accurately convey the statistical distribution, focuses on the human impact of the disease.
+Cartograms provide the spatial distortion to more accurately convey the statistical distribution, focusing on the human impact of the disease.
 However, the transformation of contiguous cartograms often occurs at the expense of the shape of areas [@CBATCC, @NAC, @TAAM].
@@
 
-Some mix of tiling, faceting or even micro maps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies.
+Some mix of tiling, faceting or even micromaps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies.
 Table 3 <!--\ref{tab:methods}--> summarizes the key criteria for testing maps and alternative displays.
@@
 
-Table: 3<!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
+Table: Table 3: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
 
-|                     | Choropleth | Contiguous | Non-contig | Dorling | tile maps | Geofacets |
+|                     | Choropleth | Contiguous | Non-contig | Dorling | Tile maps | Geofacets |
 |:-----------|:------------:|:-----------:|:---------------:|:---------:|:-----------:|:----------:|
@@
 
-# User interaction {#ch:interacting}
+# 5 User interaction {#ch:interacting}
 <!-- 634 words -->
@@
 
-One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
-It can be a juggling act for a diverse audience, with experts probably preferring more detail [@cliburn2002design].
+One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which they become confused.
+It can be a juggling act for a diverse audience, with experts probably preferring more detail [@cliburn2002design] while a simpler display is more broadly readable.
 Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
@@
 Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display.
-Interactive design features, found in online cancer maps, include tool tips, drop-down menus, data selection, zooming, and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
+Interactive design features, found in online cancer maps, include tool tips, drop-down menus, data selection, zooming, and panning to allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
 The use of these supports can be found in various online cancer maps and are shown in Figure 5 [@roberts2019communication].
@@
 Animation, in contrast to interactivity, usually involves pre-computing views and showing these in a sequence.
-An overview of animation in association with the R package `gganimate` [@gganimate], with some focus on maps, explaining transitions between frames in a linear, to provide a fixed narrative to passive users [@TGA].
+Lin Pedersen [@TGA] provides an overview of animation for maps using the R package `gganimate` [@gganimate].
 Animations are used to communicate a message by capturing and directing users' attention.
 It is most often employed to show changes over time.
-The controls for basic animation are usually placed outside of the plot space [@TGA], the map image is updated/replaced as the animation progresses.
+The controls for basic animation are usually placed outside of the plot space [@TGA], and the map image is updated/replaced as the animation progresses.
 
-Weather maps are a thoroughly developed example of animation of spatial displays to communicate information to the general public [@CPISACA].
-The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period.
+Weather maps are a thoroughly developed examples of animation of spatial displays to communicate information to the general public [@CPISACA].
+The movement of a weather system will follow a forecasted path.
+All map users can follow the animated path of the weather system across the geography over a specified period.
 
@@
 
-## Under-utilized techniques
+<!-- ## 5.1 Under-utilized techniques-->
 
@@
 
-```{r interacting, fig.cap = "Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
+```{r interacting, fig.cap = "Figure 5: Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
 i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png') # b in choro grid
@@
 
-# Conclusions
+# 6. Conclusions
 <!-- 307 words -->

### 2019-12-18 review_article.Rmd
election mapping references

--- 2019-12-18 review_article.Rmd
@@
 title: "Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"
-author: "Stephanie Kobakian, Jessie Roberts and Dianne Cook"
-bibliography: reviewbib.bib
+author: "Stephanie Kobakian^[Science and Engineering Faculty, Queensland University of Technology.], Dianne Cook^[Faculty of Econometrics and Business Statistics, Monash University.] and Jessie Roberts^[Queensland University of Technology.]"
 output:
@@
     - \usepackage{setspace}\doublespacing
-    - \usepackage{fancyhdr}\pagestyle{fancy}\fancyhead[R]{\thepage}
 link-citations: yes
+bibliography: reviewbib.bib
 fontsize: 11pt
@@
 A similar method to a rectangular cartogram represents each geographic area using a square of the same dimensions.
-There are online media sources using this method, these include [NPR](https://www.npr.org/2016/10/18/498406765/npr-battleground-map-hillary-clinton-is-winning-and-its-not-close), [FiveThirtyEight](https://projects.fivethirtyeight.com/2016-election-forecast/), [The Wall Street Journal](http://graphics.wsj.com/elections/2016/2016-electoral-college-map-predictions/), [The Washington Post](https://www.washingtonpost.com/graphics/politics/2016-election/50-state-poll/).
+There are online media sources using this method, these include [@NPR], [@FiveThirtyEight], [@WSJ], [@WP].
 Tile maps may be difficult to create as they are best created manually, they require additional time and care as the number of geographic areas to include increases.

### 2019-12-18 review_article.Rmd

--- 2019-12-18 review_article.Rmd
@@
 ---
-title: Cancer Applications of Choropleth Maps, and the Potential of Cartograms and
-  Alternative Map Displays
-author: Stephanie Kobakian, Jessie Roberts and Dianne Cook
-csl: springer-vancouver.csl
+title: "Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"
+author: "Stephanie Kobakian^[Science and Engineering Faculty, Queensland University of Technology.], Dianne Cook^[Faculty of Econometrics and Business Statistics, Monash University.] and Jessie Roberts^[Queensland University of Technology.]"
 output:
@@
     - \usepackage{setspace}\doublespacing
-- \usepackage{fancyhdr}\pagestyle{fancy}\fancyhead[R]{\thepage}
 link-citations: yes
+csl: springer-vancouver.csl
 fontsize: 11pt
@@
 The paper is structured as follows.
-The next section describes the choropleth map, which is the common approach to disease maps and presents examples of atlases in use today.
-Section 3 discusses the limitations of the choropleth map.
-Section 4 describes alternative displays, including the cartogram, which is useful when the map has heterogeneously sized geographic units.
-Section 5 presents the limitations of the production and use of alternative displays.
-Disease maps are more useful when made interactive, and common options are described in Section 6, along with a discussion of benefits and disadvantages.
+The next section describes the choropleth map, which is the common approach to disease maps and presents examples of atlases in use today and discusses the limitations of the choropleth map.
+Section 3 describes alternative displays, including the cartogram, which is useful when the map has heterogeneously sized geographic units.
+Section 4 presents the limitations of the production and use of alternative displays.
+Disease maps are more useful when made interactive, and common options are described in Section 5, along with a discussion of benefits and disadvantages.
 
@@
 
-# 3 Contemporary alternatives to choropleth maps {#ch:alternatives}
+# 3. Contemporary alternatives to choropleth maps {#ch:alternatives}
 
@@
 
-There are many issues to consider when creating alternative map displays: the intended audience of the map, and its purpose are key points in cartogram use and creation as 'there is no "best" cartogram or method of creating cartograms just as there is no "best" map' (Monmonier and Schnell, 1988) [@ACTUC].
+There are two key issues to consider when creating alternative map displays, (1) the intended audience of the map, and (2) its purpose.
 Nusrat and Kobourov [@SAIC] provided a framework to investigate implementations of the many algorithms presented, and the "statistical accuracy, geographical accuracy, and topological accuracy".
@@
 
-The white space presents the meaningful empty-space property [@ECGC] [@NAC] but it also distracts the reader from the data, with a low data density [@tufte].
+The white space presents the meaningful empty-space property [@ECGC] [@NAC] but it also distracts the reader from the data, with a low data density [@TVDQI].
 
@@
 The squares are tessellated to create a grid.
-Each area is represented by a square of the same dimensions, each tile is usually one unit of measurement.
-This could be geographic regions such as states or population-based that use a consistent measure of population for each tile.
-Regions with more than four neighbors introduce some displacement, because only four squares can be adjoining.
-Like the chorpleth and cartograms, the tile map uses color to represent a value of a statistic for each area.
-There are online media sources using this method, which include [NPR](https://www.npr.org/2016/10/18/498406765/npr-battleground-map-hillary-clinton-is-winning-and-its-not-close), [FiveThirtyEight](https://projects.fivethirtyeight.com/2016-election-forecast/), [The Wall Street Journal](http://graphics.wsj.com/elections/2016/2016-electoral-college-map-predictions/), [The Washington Post](https://www.washingtonpost.com/graphics/politics/2016-election/50-state-poll/).
-Tile maps often need to be created manually, requiring more time and care as the number of geographic areas increases.
-<!-- cite mosaic cartograms -->
+Each area is represented by a square of the same dimensions, each tile is usually one unit of measurement, this could be geographic regions such as states or population-based that use a consistent measure of population for each tile.
+Regions with over four neighbors require some necessary displacement.
+The tile map uses color to represent a value of a statistic for each area.
+A similar method to a rectangular cartogram represents each geographic area using a square of the same dimensions.
+There are online media sources using this method, these include [@NPR], [@FiveThirtyEight], [@WSJ], [@WP].
+Tile maps may be difficult to create as they are best created manually, they require additional time and care as the number of geographic areas to include increases.
 
@@
 
-# 4 Comparison and critique of alternative displays {#sec:compare}
+# 4. Comparison and critique of alternative displays {#sec:compare}
 <!-- 408 words -->
@@
 
-# 5 User interaction {#ch:interacting}
+# 5. User interaction {#ch:interacting}
 <!-- 634 words -->

### 2019-12-19 Response/Response.Rmd
figure creation script

@@
 fromqualifications: BA
-fromtitle: Research Masters Student
 fromemail: stephanie.kobakian@monash.edu

### 2019-12-19 review_article.Rmd
figure creation script

@@
                       warning = FALSE,
-                      message = FALSE,
-                      fig.path = fig_path)
+                      message = FALSE)
 library(ggplot2)
@@
 These atllases are described in Table 1.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
-library(cowplot)
-
-# Import the images
-plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
-plot2 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
-plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
-plot4 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
-plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
-plot6 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
-plot7 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
-
-ggdraw() +
-  draw_plot(rasterGrob(plot1), 0, .75, 0.33, .25) +
-  draw_plot(rasterGrob(plot2), 0.33, 0.75, 0.33, .25) +
-  draw_plot(rasterGrob(plot3), 0.66, 0.5, .33, .5) +
-  draw_plot(rasterGrob(plot4), 0, 0.5, .33, .25) +
-  draw_plot(rasterGrob(plot5), 0.33, 0.5, .33, .25) +
-  draw_plot(rasterGrob(plot6), 0, 0, .66, .5) +
-  draw_plot(rasterGrob(plot7), .66, 0, .33, .5) +
-  draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
-    c(0.015, 0.31, 0.66, 0.015, 0.31, 0.014, 0.66), 
-    c(0.99, 0.99, 0.99, 0.74, 0.74, 0.46, 0.46), size = 20)
-
-ggsave(filename = "figures/choropleth_grid.png", 
-  device = "png", width = 6, height = 5, dpi = 300)
+knitr::include_graphics("figures/choropleth_grid.png")
 ```
@@
 Map creators give each state equal size and thus equal emphasis in (d) the hexagon tile map.", message=FALSE, warning=FALSE, out.width='100%'}
-g1 <- png::readPNG("figures/ggcont.png")
-g2 <- png::readPNG("figures/ggncont.png")
-g3 <- png::readPNG("figures/ggdorl.png")
-g4 <- png::readPNG("figures/gghexmap.png")
-#knitr::include_graphics("figures/usa_grid.png")
-
-load("figures/usa_legend.rda")
-
-ggdraw() +
-  draw_plot(rasterGrob(g1), 0, .55, 0.5,  0.45) +
-  draw_plot(rasterGrob(g2), 0.5, 0.55, 0.5, 0.45) +
-  draw_plot(rasterGrob(g3), 0.0, 0.1, 0.5, 0.45) +
-  draw_plot(rasterGrob(g4), 0.5, 0.1, 0.5, 0.45) +
-  draw_plot(usa_legend, 0, 0, 1, 0.1, scale = 2) +
-  draw_plot_label(c("a", "b", "c", "d"), 
-    c(0, .5, 0, 0.5), 
-    c(1, 1, .5, .5), size = 20)
+knitr::include_graphics("figures/usa_grid.png")
 ```
@@
 The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%', cache=FALSE, fig.width=8, fig.height=6, dpi=300}
-
-aus1 <- png::readPNG("figures/aus_ggcont.png")
-aus2 <- png::readPNG("figures/aus_ggncont.png")
-aus3 <- png::readPNG("figures/aus_ggdorl.png")
-aus4 <- png::readPNG("figures/aus_gghexmap.png")
-load("aus_legend.rda")
-load("figures/aus_legend.rda")
-
-library(cowplot)
-ggdraw(xlim = c(0,1), ylim = c(0,1.2)) +
-  draw_plot(rasterGrob(aus1), 0, .65, 0.5, 0.50) +
-  draw_plot(rasterGrob(aus2), 0.5, 0.65, 0.5, 0.50) +
-  draw_plot(rasterGrob(aus3), 0.0, 0.1, 0.5, 0.55) +
-  draw_plot(rasterGrob(aus4), 0.5, 0.1, 0.5, 0.55) +
-  draw_plot(aus_legend, 0, 0, 1, .1) +
-  draw_plot_label(c("a", "b", "c", "d"), 
-    c(0, .5, 0, 0.5), 
-    c(1.15, 1.15, .6, .6), size = 20)
+knitr::include_graphics("figures/auscartograms.png")
 ```
@@
 ```{r interacting, fig.cap = "Figure 5: Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
-i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png') # b in choro grid
-i2 <- png::readPNG('cancer_map_images/interactivity/15.interactivity.png') # d in choro grid
-i3 <- png::readPNG('cancer_map_images/interactivity/4.interactivity.png')
-i4 <- png::readPNG('cancer_map_images/interactivity/5.interactivity.png')
-i5 <- png::readPNG('cancer_map_images/interactivity/7.interactivity.png') # e in choro grid
-
-ggdraw() +
-  draw_plot(rasterGrob(i1), 0,  0.13,   .28,  .83) + #a
-  draw_plot(rasterGrob(i2), 0.32, 0.56, .68, .4) + #b
-  draw_plot(rasterGrob(i3), 0.32, 0.41, .68, .13) + #c
-  draw_plot(rasterGrob(i4), 0.32, 0.14, .68, .25) + #d
-  draw_plot(rasterGrob(i5), 0, 0, 1, .13) + #e
-  draw_plot_label(c("i", "ii", "iii", "iv", "v"), 
-    c(0.01, 0.3, 0.3, 0.3, 0.01), 
-    c(0.99, 0.99, 0.57, 0.40, 0.15), size = 20)
-
-ggsave(filename = "figures/supp_grid.png", device = "png", 
-  width = 12, height = 9, dpi = 300)
+knitr::include_graphics("figures/interacting.png")
 ```
@@
 Linked maps and time-series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b.
-A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 9}
-a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
-a2 <- png::readPNG("cancer_map_images/animation/2.animation.png")
-
-ggdraw() +
-  draw_plot(rasterGrob(a1), 0, 0, 0.3, 1) +
-  draw_plot(rasterGrob(a2), 0.3, 0, 0.7, 1) +
-  draw_plot_label(c("I", "II"), 
-    c(0, .3), c(1, 1), size = 20)
+A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 7}
+knitr::include_graphics("figures/animating.png")
 ```

### 2019-12-19 figure_patchwork.R

@@ -1,0 +1,2 @@
+# Figure patchworking
+

### 2019-12-19 review_figures.R

@@ -1,0 +1,107 @@
+# Figures
+
+library(cowplot)
+
+# Import the images
+plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
+plot2 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
+plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
+plot4 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
+plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
+plot6 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
+plot7 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
+
+ggdraw(xlim = c(0,1), ylim = c(0,1)) +
+  draw_plot(rasterGrob(plot1), 0, .75, 0.33, .26) +
+  draw_plot(rasterGrob(plot2), 0.33, 0.75, 0.33, .25) +
+  draw_plot(rasterGrob(plot3), 0.66, 0.5, .33, .5) +
+  draw_plot(rasterGrob(plot4), 0, 0.5, .33, .25) +
+  draw_plot(rasterGrob(plot5), 0.33, 0.5, .33, .25) +
+  draw_plot(rasterGrob(plot6), 0, 0, .66, .5) +
+  draw_plot(rasterGrob(plot7), .66, 0, .33, .5) +
+  draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
+    c(0.025, 0.33, 0.66, 0.025, 0.33, 0.025, 0.66), 
+    c(0.99, 0.99, 0.99, 0.74, 0.74, 0.46, 0.46), size = 20)
+
+ggsave(filename = "figures/choropleth_grid.png", 
+  device = "png", width = 12, height = 10, dpi = 300)
+
+
+
+
+g1 <- png::readPNG("figures/ggcont.png")
+g2 <- png::readPNG("figures/ggncont.png")
+g3 <- png::readPNG("figures/ggdorl.png")
+g4 <- png::readPNG("figures/gghexmap.png")
+
+load("figures/usa_legend.rda")
+
+ggdraw() +
+  draw_plot(rasterGrob(g1), 0, .55, 0.5,  0.45) +
[…]

### 2019-12-30 lit_review.Rmd
pdf for thesis

--- 2019-12-19 review_article.Rmd
@@
 output:
-  word_document:
-    toc: no
-    toc_depth: '3'
   pdf_document:
@@
     toc_depth: 3
+  word_document:
+    toc: no
+    toc_depth: '3'
 geometry: margin = 1.2in
@@
     - \usepackage{setspace}\doublespacing
-link-citations: yes
+link-citations: false
 csl: springer-vancouver.csl
@@
 
-# 1. Introduction
+#  Introduction
 <!-- 436 words -->
@@
 
-# 2. Traditional approaches for cancer map displays {#ch:choropleth_maps}
+#  Traditional approaches for cancer map displays {#ch:choropleth_maps}
 <!-- 381 words -->
@@
 
-## 2.1 Cancer atlases {#ch:public}
+##  Cancer atlases {#ch:public}
 <!-- 360 words -->
@@
 
-```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Figure 1: A selection of choropleth cancer maps from online atlases that are publicly available.
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "  A selection of choropleth cancer maps from online atlases that are publicly available.
 Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends.
-These atllases are described in Table 1.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
+These atlases are described in Table 1.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
 knitr::include_graphics("figures/choropleth_grid.png")
@@
 
-Table: Table 1: <!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases.
+Table:  <!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases.
 
@@
 
-Table: Table 2: <!--\label{tab:measures}--> Common measures for reporting cancer information.
+Table:  <!--\label{tab:measures}--> Common measures for reporting cancer information.
 
@@
 
-## 2.2 Additional considerations
+##  Additional considerations
 <!-- 479 words -->
@@
 
-## 2.3 Limitations of choropleth displays {#ch:chorolimit}
+##  Limitations of choropleth displays {#ch:chorolimit}
 
@@
 
-# 3. Contemporary alternatives to choropleth maps {#ch:alternatives}
+#  Contemporary alternatives to choropleth maps {#ch:alternatives}
 
-## 3.1 Cartograms
+##  Cartograms
 <!-- 1496 words inc table-->
@@
 
-```{r ggcartograms, fig.cap = "Figure 2: Common alternatives to maps, showing the same information for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous).
+```{r ggcartograms, fig.cap = "  Common alternatives to maps, showing the same information for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous).
 Maps (a) - (c) are created by resizing and reshaping the states of the USA to match the 2015 population of the state.
@@
 
-Table: Table 3: <!--\label{tab:usa}--> Maps used to present statistics for the United States of America.
+Table:  <!--\label{tab:usa}--> Maps used to present statistics for the United States of America.
 The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
@@
 
-### 3.1.1 Contiguous
+###  Contiguous
 
@@
 
-### 3.1.2 Non-contiguous
+###  Non-contiguous
 
@@
 
-### 3.1.3 Dorling
+###  Dorling
 
@@
 
-## 3.2 Tile Map
+##  Tile Map
 <!-- 173 words -->
@@
 
-## 3.3 Geofacet
+##  Geofacet
 <!-- 166 words -->
@@
 
-```{r ggtilemap, fig.cap = "Figure 3: Two alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
+```{r ggtilemap, fig.cap = "  Two alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
 In the tile map, the layout approximates spatial location, with each state being an equal box filled with color representing cancer incidence.
@@
 
-## 3.4 Multivariate displays
+##  Multivariate displays
 <!-- 222 words -->
@@
 
-# 4. Comparison and critique of alternative displays {#sec:compare}
+#  Comparison and critique of alternative displays {#sec:compare}
 <!-- 408 words -->
 
-```{r auscartograms, fig.cap = "Figure 4: Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map.
+```{r auscartograms, fig.cap = "  Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map.
 The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas.
@@
 
-## 4.1 Neither choropleth maps or cartograms perform well for Australia
+##  Neither choropleth maps or cartograms perform well for Australia
 
@@
 
-## 4.2 Limitations of alternative displays
+##  Limitations of alternative displays
 
@@
 
-Table: Table 3: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
+Table:  <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
 
@@
 
-# 5. User interaction {#ch:interacting}
+#  User interaction {#ch:interacting}
 <!-- 634 words -->
@@
 
-<!-- ## 5.1 Under-utilized techniques-->
+<!-- ##  Under-utilized techniques-->
 
@@
 
-```{r interacting, fig.cap = "Figure 5: Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
+```{r interacting, fig.cap = "  Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
 knitr::include_graphics("figures/interacting.png")
@@
 
-# 6. Conclusions
+#  Conclusions
 <!-- 307 words -->
@@
 # References
-
-<!--
-## counts versus predictions
-
-@SEMA
-'Disease mapping is carried out to summarise spatial and spatio-temporal variation in risk.'
-Descriptive statistic of the relationship between geography and a disease.
-population, exposure, and health data may have point data or aggregated summaries (count data)
-potentially be subject to a number of inaccuracies
-diagnostic error or misclassification
-
-Not only the counts, but population, migration, deaths and births are modelled across census years.
-
-rapid visual summary of complex geographical information, descriptive purposes, to generate hypotheses, for surveillance, and to aid policy formation and resource allocation.
-
-When using counts:
- small populations imply large variability in estimated rates
-
-Decisions taken at this stage may be critical as apparent ‘clusters’ may depend crucially on the boundaries chosen in time or space: ‘The more narrowly the underlying population is defined, the less will be the number of expected cases, the greater will be the estimate of the excess rate, and often the more pronounced will be the statistical significance’
-
-Using prediction:
-assuming that neighboring μi are correlated leads to a ‘clustering’ prior
-
--->
-
-<!-- presents forms of low level interactions with a user interface, these are the control menus for several online cancer atlases.
-The user interface for these online atlases allow combinations of cancer type, gender, and other factors to be shown on the map display and the user to move between and/or compare different views of these variables.
-These controls may also change the geographies shown, such as continents or countries, states or counties.
-These displays can show familiar shapes for users to orient themselves on the map, then zoom in to explore densely populated communities.
-Interactions with online atlases can also change the amount of detail shown.-->
-
-<!-- Interactivity can be implemented within modern mapping methods to incorporate additional information and complexity without overloading the user.
-Adding too much information to a map may create cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
-Complexity and density of visualization methods can overwhelm novice decision makers however, experts are able to use the detail more readily when making-decisions [@cliburn2002design]. 
-Supplementary information can be incorporated into online atlases without cluttering the display.
-Interactive design features found in online cancer maps include tool tip features, drop-down menus, data selection, zooming and panning.
-Allowing users to explore the map for more information and providing flexibility in the display [@HTLWM].
-The use of these supports were found in various online cancer maps identified by @roberts2019communication.
-Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
-Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information [@maceachren1992visualizing, @van1994visualization].
-For example, changing the population, age or other demographic variables.
-The controls for basic interactive features are often placed outside of the plot space [@TGA], the map image is updated/replaced as the user interacts with the controls. 
-Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactive capabilities.
-
-Relationships between spatial areas and diseases can be explored with sophistication in nontraditional but still 'cognitively accessible' ways [@TNTEA].
-This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-
-The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
-
-Interactivity and animation provide an alternative to the traditional approach of a single map for a single use, and enable the map to become more exploratory.
-Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tooltips, and data selection tools [@TGA].
-tooltips allow users to obtain additional information regarding specific geographic areas.
-
-Development and adoption of this technology has allowed map users more control of the displayed message. 
--->
-
-<!--to expand the communicative ability of a disease map.
-It is important to identify the primary questions that might be asked by the users of a map, and allow the display to change to best answer these questions.
-@CPISACA provide three questions that are generally asked when using a disease map:
-
-- What is the mortality rate in a certain area?
-- Are there geographic trends in the data, or regions of unusually high or low rates?
-- Is the lung cancer mortality pattern similar to the pattern of smoking prevalence shown in a companion map?
-
-A single map display may not effectively answer all three at the same time.
-This is especially true when some areas are small and difficult to identify clearly in a single display or resolution.
-However, with the ability to interact with a map, or animate the display, map creators can specify and answer these questions by directing the attention of a map user to the areas of interest.
--->

### 2020-01-14 01-intro.Rmd
statements

---
title: "Chapter 1: Introduction"
author: "Stephanie"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, cache=TRUE)
# Load any R packages you need here
library(forecast)
library(ggplot2)
```

# Introduction {#ch:intro}

There are many visualisation methods used to present geospatial data. The design of the visualisation chosen can hinder or improve the communication of the spatial distribution. A choropleth map is the most common display used to present geographical data.Maps contribute to understanding spatial distributions of disease occurrence, and locating disease clusters. Disease data is often aggregated by political areas. One reason for this is privacy, another the responsibility on the political entity to respond. The typical visualisation for aggregated spatial data is a choropleth map, where areas are coloured by the numerical value.

Choropleth maps do a disservice to the map reader, as the attention of the map user is distributed according to the size of the area. Using a choropleth map to get a broad perspective of Australia can be misleading, when the use of geographical areas misrepresents the spatial distribution of a dataset. This is not practical if each area is considered equally important. In Australia, the population is not equally dispersed across the geographic map base. 
Instead, the communities are densely populated in the inner city areas, especially around the capital cities. There are several visualisation methods that have been developed to emphasise the population dense areas. These alternatives should be considered when planning the communication of geospatial statistics. Visualisations should be chosen to best represent the spatial distribution, The work is motivated by the Cancer Atlas of Australia,
which presents the spatial patterns of many cancers in Australia. The aim of this thesis is to contribute an algorithm that creates effective visualisations for the communication of geospatial population statistics.

## The Australian Cancer Atlas

This work was motivated by the Australian Cancer Atlas. An online, interactive web tool for exploring the impact of cancer on Australian communities. The prominent display used by the Australian Cancer Atlas is a visualisation of Incidence Rates or Excess Death Rates. The set of geographic units used is Australian Statisical Areas, at Level 2. There are almost 2,200 individual areas, ranging in size from

The choropleth map display used in the Australian Cancer Atlas is familiar to the general public of Australia. It is appropriate to use this display as users can orient themselves on the map base and find geographic areas relevant to them.
However, when the intention of the map user is to convey the whole spatial distribution the information derived visually from the colours can be misleading.
The rural areas are over emphasised, and the densely populated inner city areas are not given enough attention.

## Visual Inference

Visual inference will be used to determine if the communication of population geospatial statistics is more effective when using an alternative display.
Buja et al [@GIIV] provide the 'lineup' protocol as a formal framework for testing visual statistical methods. Implementing this framework will allow new alternative visualisation method to be tested.

The lineup protocol will be used to test if a visualisation is effective, a visualisation displaying a real population based distribution can be hidden in a collection of visualisations that display null distributions [@chowd]. 
It takes inspiration from a police lineup.
The witness in this regard is a participant recruited from a crowdsource platform, such as Figure-Eight.
The visualisation containing a real distribution is considered the 'accused'.
It is put in a lineup of innocent displays that do not show a real population based distribution.
If the 'witness' chooses the 'accussed' as different from the innocent plots, it can be considered that there is a specific pattern displayed that is not present in the others. 
In this protocol, the null hypothesis can be rejected in favour of the alternative when it is chosen in the lineup. The null hypothesis fails to be rejected when it is not selected in the linup.

## Aims and Objectives

This work aims to provide a solution to presenting geospatial data regarding populations. 
It considers the visualisation methods developed over the past two centuries that shift the focus from the geographic map base.

1. *Algorithm for creating hexagon tile maps of Australia:* The algorithm will take geospatial areas and create an alternative visualisation of the spatially distribution.

2. *Test the effectiveness of the hexagon tile map relative to the choropleth map:* The hexagon tile map produced by the algorithm will be contrasted with the traditional choropleth map, applying the same colour methods to represent the data. The maps will be used in an experiment to test the effectiveness by asking for users to spot spatial distributions.

3. *Communicating the relationship between the hexmap and choropleth map through animation:* Maximise the benefits of both displays when communicating to the public. The use of animations may control how people follow a recognisable map of Australia into an alternative visualisation for inference.

## Research Contributions

This research contributes a new algorithm for creating hexagon tile map displays. It contributes an R [@R] package which implements the algorithm and allows R users to create their own visualisations.
It presents a case study that contributes to a growing field of visual inference studies, applied to spatial data by comparing a choropleth map to a hexagon tile map display.
It also shows how it can be used in practice to effectively communicate cancer distributions.

## Thesis Structure

The thesis is structured as follows: Chapter two contains a literature review. 
The literature reviews considers the current peer reviewed literature and published books that explore spatial distributions of cancer across the globe.
It also considers how to evaluate the visualisation methods used for spatial data.

Chapter three explores the algorithm to create hexagon tile maps and the code used to create a small example of Tasmania in Australia.
Chapter four is a visual inference study that contains the methods and results that compare the use of a choropleth map and a hexagon tile map on the same data sets.
Chapter five provides a conclusion of the results of the visual inference study and how the hexagon tile map may be used in practice.

### 2020-01-16 review_article.Rmd
Remove numbering

@@
                       warning = FALSE,
-                      message = FALSE)
+                      message = FALSE,
+  out.width = "\\textwidth")
 library(ggplot2)
@@
 
-# 1. Introduction
+# Introduction
 <!-- 436 words -->
@@
 
-# 2. Traditional approaches for cancer map displays {#ch:choropleth_maps}
+# Traditional approaches for cancer map displays {#ch:choropleth-maps}
 <!-- 381 words -->
 <!-- what is a choropleth map -->
-A choropleth map displays the geographical distribution of data over a set of spatial units by shading areas of a map [@EI, @BCM].
+A choropleth map displays the geographical distribution of data over a set of spatial units by shading areas of a map [@EI], [@BCM].
 Faithful rendering of the geography, when combined with an appropriate color scheme, can reveal spatial patterns among data values.
@@
 
-## 2.1 Cancer atlases {#ch:public}
+## Cancer atlases {#ch:public}
 <!-- 360 words -->
@@
 Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends.
-These atllases are described in Table 1.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
+These atllases are described in Table 1.", fig.align='center', fig.width=12, fig.height=10}
 knitr::include_graphics("figures/choropleth_grid.png")
@@
 
-Table: Table 1: <!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases.
+Table 1: <!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases.
 
@@
 
-Table: Table 2: <!--\label{tab:measures}--> Common measures for reporting cancer information.
+Table 2: <!--\label{tab:measures}--> Common measures for reporting cancer information.
 
@@
 
-## 2.2 Additional considerations
+## Additional considerations
 <!-- 479 words -->
@@
 To address this, a cancer atlas often communicates uncertainty about the value of a statistic.
-There are several potential sources of uncertainty: sampling error, errors arising from the disease reporting process (or data collection), and errors arising from the statistical modeling or simulation process.
+There are several potential sources of uncertainty: sampling error, errors arising from the disease reporting process (or data collection), and errors arising from the statistical modelling or simulation process.
 The most common measures used to present uncertainty are credible or confidence intervals (CIs).
@@
 
-## 2.3 Limitations of choropleth displays {#ch:chorolimit}
+## Limitations of choropleth displays {#ch:chorolimit}
 
@@
 
-# 3. Contemporary alternatives to choropleth maps {#ch:alternatives}
+# Contemporary alternatives to choropleth maps {#ch:alternatives}
 
-## 3.1 Cartograms
+## Cartograms
 <!-- 1496 words inc table-->
@@
 ```{r choroCRS, fig.cap = "Four choropleth maps average age-adjusted rate of incidence for lung and bronchus in the United States averaged from sex and years 2012-2016, using four different map projections.
-The map projections alter the shapes and angles of the boundaries of each state.", message=FALSE, warning=FALSE, out.width='100%', fig.align='center'}
+The map projections alter the shapes and angles of the boundaries of each state.", message=FALSE, warning=FALSE, fig.align='center'}
 #knitr::include_graphics("figures/ggchoroCRS.png")
@@
 This provides a better sense of the extent of disease relative to the population in the country and can help ease losing information about physically small but population-dense states.
-Map creators give each state equal size and thus equal emphasis in (d) the hexagon tile map.", message=FALSE, warning=FALSE, out.width='100%'}
+Map creators give each state equal size and thus equal emphasis in (d) the hexagon tile map.", message=FALSE, warning=FALSE}
 knitr::include_graphics("figures/usa_grid.png")
@@
 
-Table: Table 3: <!--\label{tab:usa}--> Maps used to present statistics for the United States of America.
+Table 3: <!--\label{tab:usa}--> Maps used to present statistics for the United States of America.
 The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
@@
 Dorling| Circles are used to represent each state, but the population of the state determines the size in 2015.
-The North-East states remain closer to their neighbors and are slghtly displaced from their geographic location.
+The North-East states remain closer to their neighbors and are slightly displaced from their geographic location.
 It highlights the sparsity of the population in the Mid West by the distance between the circles at the geographic centroids.|
@@
 
-### 3.1.1 Contiguous
+### Contiguous
 
@@
 To be able to recognize the significant changes, a reader will usually have to know the initial geography to find the differences in the new cartogram layout [@NAC].
-Tobler's Conformal mapping means to preserve angles locally so that the shapes of small areas on a traditional map and a cartogram would be similar.
-@CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes.
+The shapes of small areas on a choropleth map and a cartogram are preserved using Tobler's Conformal mapping method.
+Koccmoud and House [@CBATCC] present this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes.
 
-### 3.1.2 Non-contiguous
+### Non-contiguous
 
@@
 
-### 3.1.3 Dorling
+### Dorling
 
@@
 
-## 3.2 Tile Map
+## Tile Map
 <!-- 173 words -->
@@
 
-## 3.3 Geofacet
+## Geofacet
 <!-- 166 words -->
@@
 
-## 3.4 Multivariate displays
+## Multivariate displays
 <!-- 222 words -->
@@
 
-# 4. Comparison and critique of alternative displays {#sec:compare}
+# Comparison and critique of alternative displays {#sec:compare}
 <!-- 408 words -->
@@
 The Dorling cartogram presents all areas but many are difficult to compare.
-The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%', cache=FALSE, fig.width=8, fig.height=6, dpi=300}
+The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
 knitr::include_graphics("figures/auscartograms.png")
@@
 
-## 4.1 Neither choropleth maps or cartograms perform well for Australia
+## Neither choropleth maps or cartograms perform well for Australia
 
@@
 
-## 4.2 Limitations of alternative displays
+## Limitations of alternative displays
 
@@
 
-<!-- Cartograms, sacrifice geography, focus on the outliers/ relationships in data space 
-Cartographers were historically the creators of geographic displays, and epidemiologists utilized the maps that had been prepared in advance.
-Cartograms with the population variable used to distortion the geographic map regions and boundaries are density equalizing maps, that reduce the visual prominence of large, but low-population areas [@VSSDCUC].
-Cartogram displays incorporate the statistics and population into the design of the display,
-shifting and sometimes sacrificing familiar boundaries to draw attention to large outliers in the data space or population density.
-The unusual shapes are not longer familiar and may be difficult to compare.
-This can prevent effective comparison of the statistic.
-Cartograms also favor extreme statistics in the positive direction.
-Small extreme values are often lost due to the resulting small sized areas the do not draw user's attention.
-The change in the map based on one variable provides the opportunity for bi-variate displays, using the change in area for one variable, and color for another.
-The difference between the familiar map display helps highlight the impact of the disease on communities.
--->
-
-<!-- Alternatives, remove geographic shapes, keep spatial relationship, focus on the outliers/ relationships in population space 
-Alternative maps may replace the unexpected boundaries and shapes with familiar, simple shapes. This makes the spatial distribution the primary concept communicated by the display.
-Areas that are geographically close will maintain connectedness in some way, but the population and the statistic will dominate.
-tilemap displays may only allow one variable to be visualized if each region is given consistent map space, this occurs in geofaceting. 
-Dorlings provide bi-variate displays as they use the size of the shape for one variable, and color for another.
-These alternatives are especially helpful for data aggregations where administrative boundaries break populations into groups.
-The experience of each community may be worth considering, as the experiences shared by the population within them may be similar due to the services and facilities they share.
-This display allows a more equitable view of each community, and does not minimise those that operate on a smaller geographic scale.
-
-Faceting provides an opportunity to contrast the overall distribution of variables across a geographic space or collection of geographic areas.
-This display makes is difficult to compare specific regions.-->
-
-Table: Table 3: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
+Table 3: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
 
@@
 
-# 5. User interaction {#ch:interacting}
+# User interaction {#ch:interacting}
 <!-- 634 words -->
-
-<!-- Interactivity and animation can add value to online atlases, by enabling additional information to be presented on-the-fly with one display.
-It also invites the viewer to engage with the display, and the information.
-
-The word "interactive" is used, and perhaps, confused for many different graphics actions.
-At one extreme, it is synonymous with direct manipulation of the display, such as selecting an area, mousing over, or panning and zooming, using mouse action on the plot.
-At the other extreme, it refers to graphical user interfaces (GUI) such as selecting an item on a menu to change the variable being displayed, or checking a toggle box, or dragging a scrollbar.
-@RPIG provides a good overview of types of interactivity with plots.
-a key difference from an animation is that interactivity typically involves on-the-fly computations, and changes to the plot.
-
-The primary interactive tools found in online cancer maps include
-
-- drop-down menus to select features,
-- variable selection, 
-- tool tips, 
-- zooming and panning.
-
-The latter two would be considered to be direct manipulation on the map display, and the former two are GUI interfaces. The effects of these actions are rapid, incremental, and reversible [@DMIV]. 
--->
 
@@
 
-<!-- 
-XXX This part needs to go to interactivity!
-One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information.
-These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design].
-Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.
+<!-- ## Under-utilized techniques-->
 
-Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display.
-Interactive design features, found in online cancer maps, include tooltips, drop-down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM].
-The use of these supports were found in various online cancer maps identified by @roberts2019communication.
-The controls for basic interactive features are often placed outside of the plot space [@TGA], thus the map image is updated/replaced as the user interacts with the controls.
-For example, changing the population age or other demographic variables.
-Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection tools.
-
-Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction].
-Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).
-
-These interactive features provide an opportunity for users to explore the additional information available.
-This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
-This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accessible' ways [@TNTEA].
-The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
-@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
--->
-
-<!-- ## 5.1 Under-utilized techniques-->
+```{r interacting, fig.cap = "Figure 5: Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9, results = "asis"}
+knitr::include_graphics("figures/interacting.png")
+```
 
@@
 
-<!-- It is uncommon to see linked displays utilized in online atlases, but there are several frameworks that support interactive displays of multivariate data.
-@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
-Linked displays allow for direct manipulation, @RPIG uses Swayne and Klinke's definition of high level interaction that occurs via data selection or brushing tools that highlight subsets of graphical elements such as polygons or points.
-Selections can be made statistically through interactions which allow subsets to be selected from statistical displays, such as tables of data.
-Monmonier implemented linking choropleth maps with statistical plots, adding a geographic component to scatter plots [@MCRG], the use of link allows spatial dimensions to be reserved for geographic locations of areas.
-The statistical variation of additional variables can be presented in other appropriate displays.
-Low level interaction is more common, this occurs via the user interface or the map.
-Interacting with the map returns values or information used in the graphical display without referring to, of changing, the original data source of the map.
--->
-
-<!-- Animations -->
-<!--Animations communicate a story, showcasing a critical region by defining and progressing through a small set of keyframes, or waypoints [@ESRI].
-@TGA suggests the use of moving objects demands attention by taking advantage of the behaviour of the visual cognition system.
-They can be produced to not tolerate or require user input, creating useful, directed descriptions of processes, such as highlighting a particular region or time period [@HTLWM].
-Animations ensure all users follow the same interpretive path but creators cannot control the level of engagement.
--->
-
-<!-- Animations can be implemented to allow users varying degrees of control.
-Figure \ref{fig:animating} a gives an example of how time can be included through animating displays.
-This is a basic form of interactivity with animation allowing users to play, pause and rewind films, where the film is intended to be watched in one direction but the position in the animation can be controlled.
-@CPISACA provide weather maps as a thoroughly developed example of animation of spatial displays to communicate information to the general public.
-The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period of time.
-Where the message is most important, static or animated graphics allow control over the display and interpretation.
-User interaction more allow clicks to progress between stages of more complex animations.
-The Australian Cancer Atlas [@TACA] provides [tours](
-https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
-This implementation of animation gives users tools to plan their own exploration.
--->
-
-```{r interacting, fig.cap = "Figure 5: Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
-knitr::include_graphics("figures/interacting.png")
-```
-
 ```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a.
 Linked maps and time-series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b.
-A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 7}
+A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 7, results = "asis"}
 knitr::include_graphics("figures/animating.png")
@@
 
-# 6. Conclusions
+# Conclusions
 <!-- 307 words -->
@@
 Making an inset can clarify congested regions but this breaks the viewers' attention as they shift focus from the map to the inset, and if there are many congested areas, many insets would be needed.
-The map alternatives implement tradeoffs between the familiar shapes, and the importance of the geographic areas in the context of the areas.
+The map alternatives implement trade-offs between the familiar shapes, and the importance of the geographic areas in the context of the areas.
 Given the population or a cancer statistic for each area, the geographic size or shape will change.

### 2020-01-16 review_article.Rmd
figure images

@@
 It is recommended that alternative displays are included in cancer atlases.
-In addition, with the ease of today's technology, user interaction with the displays is encouraged.
+In addition, with the ease of today’s technology, user interaction with the displays is encouraged.
 Users should also be able to interactively display different statistics, such as incidence rate or relative incidence, or filtered by demographic variables.
@@
 
-```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Figure 1: A selection of choropleth cancer maps from online atlases that are publicly available.
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = " A selection of choropleth cancer maps from online atlases that are publicly available.
 Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends.
@@
 
-```{r ggcartograms, fig.cap = "Figure 2: Common alternatives to maps, showing the same information for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous).
+```{r ggcartograms, fig.cap = " Common alternatives to maps, showing the same information for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous).
 Maps (a) - (c) are created by resizing and reshaping the states of the USA to match the 2015 population of the state.
@@
 
-```{r ggtilemap, fig.cap = "Figure 3: Two alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
+```{r ggtilemap, fig.cap = " Two alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
 In the tile map, the layout approximates spatial location, with each state being an equal box filled with color representing cancer incidence.
@@
 
-```{r auscartograms, fig.cap = "Figure 4: Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map.
+```{r auscartograms, fig.cap = " Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map.
 The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas.
@@
 
-<!-- ## Under-utilized techniques-->
-
-```{r interacting, fig.cap = "Figure 5: Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9, results = "asis"}
+```{r interacting, fig.cap = " Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for Globocan 2018 [@Globocan], (ii) Menus for variable selection and zooming on Bowel Cancer Australia Atlas, (iii) Menus for choosing variables and countries in The Cancer Atlas, (iv) Tabs for different indicators and cancer types in Global Cancer Map, (v) Menus and toggles for variable and subset selection in United States Cancer Statistics: Data Visualizations.", results = "asis", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
 knitr::include_graphics("figures/interacting.png")
@@
 ```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a.
-Linked maps and time-series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b.
-A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 7, results = "asis"}
+Linked maps and time-series line plots, with temporal animation in Map of Cancer Mortality Rates in Spain, b.
+A highly responsive magnifying glass on a map of Breast Cancer Mortality in Canada.", results = "asis", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
 knitr::include_graphics("figures/animating.png")

### 2020-01-16 literature.Rmd
Table number corrected

--- 2020-01-16 review_article.Rmd
@@
 
-Table 3: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
+Table 4: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
 

### 2020-04-03 literature.Rmd
page numbering
begin letter to editors using reviewer feedback

@@
 ---
-title: "Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"
+title: "Mapping Cancer: the Potential of Cartograms and Alternative Map Displays"
 author: "Stephanie Kobakian^[Science and Engineering Faculty, Queensland University of Technology.], Dianne Cook^[Faculty of Econometrics and Business Statistics, Monash University.] and Jessie Roberts^[Queensland University of Technology.]"
@@
 
+\setcounter{page}{13}
+
 \section*{Abstract}
@@
 Cancer atlases communicate cancer statistics over geographic domains, typically with a choropleth map.
-They subdivide these domains into administrative regions such as countries, states, or suburbs.
-When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner city areas.
+A choropleth map subdivides these domains into administrative regions such as countries, states, or suburbs.
+When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner- city areas.
 The smaller geographic areas are important to consider if they are densely populated.
@@
 This means alternative displays can be more effective at accurately communicating spatial patterns across spatial areas.
+This study summarises current practices for cancer atlases and investigates the alternative map displays that could be used to accurately represent the distribution of cancer statistics across a population, as many cancer atlases lack appropriate displays for population statistics.
 It is recommended that alternative displays are included in cancer atlases.
 In addition, with the ease of today’s technology, user interaction with the displays is encouraged.
-Users should also be able to interactively display different statistics, such as incidence rate or relative incidence, or filtered by demographic variables.
+Users should also be able to interactively display different statistics, such as incidence rate or relative incidence, or filter by demographic variables.
 
@@
 Examples of typical geographical units include states, provinces, local government areas, and post/zip codes.
-It is easy to provide counts or incidence rates of the diagnoses of these areas.
+Counts and incidence rates are often obtained by state health registries.
+For this data privacy and ethics are key concerns, even when combined to small area statistics.
 This type of data is routinely collected for public health reasons and may be made available to the general public as a service to the community.
@@
 <!-- Ways to visualize, whats our hypothesis -->
-To visualize and communicate geospatial cancer statistics over geographic domain, a choropleth map is the common display.
-Choropleth maps show polygons representing the geographic units, where each polygon is shaded with a color according to the area-specific values of the statistic being conveyed.
+A choropleth map is the common display to visualize and communicate geospatial cancer statistics over geographic domain.
+Choropleth maps are a type of thematic map that show polygons for each of the groups of data points, representing the geographic units, where each polygon is shaded with a color according to the area-specific values of the statistic being conveyed.
 Visualizing this data is helpful as geographic patterns of disease may be obscured when reported in a table [@SAMGIS].
@@
 
-# Traditional approaches for cancer map displays {#ch:choropleth-maps}
-<!-- 381 words -->
-<!-- what is a choropleth map -->
-A choropleth map displays the geographical distribution of data over a set of spatial units by shading areas of a map [@EI], [@BCM].
-Faithful rendering of the geography, when combined with an appropriate color scheme, can reveal spatial patterns among data values.
-Identifying and explaining spatial structures, patterns, and processes involve considering the individuals and organizing them into representable units of communities [@SAMGIS].
-Early versions of choropleth maps used symbols or patterns instead of color.
-Choropleth maps can be used for displaying disease data [@DMAHP], including cancer data [@CPISACA].
-In epidemiology, choropleth maps are often used as a tool to study the spatial distribution of cancer incidence and mortality.
-
-<!-- psychology -->
-Displaying familiar state boundaries can make a map easier to read [@CIBMUK] and allow viewers to infer the spatial relationships visually in the data using their mental model of the geography.
-The map users of disease displays may include researchers, the public, policymakers, and the media [@CPISACA].
-For these users, the familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
-
-## Cancer atlases {#ch:public}
+# Cancer atlases {#ch:public}
 <!-- 360 words -->
@@
 | [The Pediatric Oncology Group of Ontario Networked Information System](https://www.pogo.ca/research-data/pogonis-childhood-cancer-database/data-requests/)|
+
+# Measures mapped
 
@@
 
-## Additional considerations
-<!-- 479 words -->
-Cancer atlases often display supplementary graphs and plots to add more information.
-Additional materials such as tables, graphs, and text explanations support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers [@CPISACA].
-The many displays of statistical summaries, including dot plots, bar plots, box plots, cumulative distribution plots, scatter plots, and normal probability plots, can provide alternative views of the cancer statistics.
-These can also display supporting statistics such as error, confidence intervals, distributions, sample or population sizes, and standard deviation.
+# Traditional approaches for cancer map displays {#ch:choropleth-maps}
+<!-- 381 words -->
+<!-- what is a choropleth map -->
+A choropleth map displays the geographical distribution of data over a set of spatial units by shading areas of a map [@EI], [@BCM].
+Faithful rendering of the geography, when combined with an appropriate color scheme, can reveal spatial patterns among data values.
+Identifying and explaining spatial structures, patterns, and processes involve considering the individuals and organizing them into representable units of communities [@SAMGIS].
+Early versions of choropleth maps used symbols or patterns instead of color.
+Choropleth maps can be used for displaying disease data [@DMAHP], including cancer data [@CPISACA].
+In epidemiology, choropleth maps are often used as a tool to study the spatial distribution of cancer incidence and mortality.
 
-The statistics communicated in atlases are often used to describe differences between areas.
-This can occur at different levels of aggregation.
-Aggregation of global health statistics occurs within administrative and arbitrarily defined regions, such as those used by the World Health Organization and the United Nations [@IARC_3].
-World atlases can allow for displays of data aggregated into continents, countries, states, provinces and congressional districts [@USInteractive].
-Each population area will probably have a different number of people, which is typically used to calibrate the statistic.
-Cancer atlases may also communicate the distribution of the population living in all areas in a table or histogram display [@NICR_1].
-Atlases can connect the population to the land available to them by communicating population density.
-
-Maps can also be used to focus on demographic strata, such as age and sex.
-Some of the digital atlases surveyed allow subsets such as males, females, or those aged over 65, to be selected for display.
-Similarly, socioeconomic indicators, such as unemployment rates, poverty rates, remoteness, and education levels, can be used to filter data, in order to communicate how cancer prevalence varies for different members of society.
-Few atlases provide this level of detail.
-
-Introducing population and demographic information helps to interpret the rates in areas effectively, but there will still be uncertainty around the rates.
-To address this, a cancer atlas often communicates uncertainty about the value of a statistic.
-There are several potential sources of uncertainty: sampling error, errors arising from the disease reporting process (or data collection), and errors arising from the statistical modelling or simulation process.
-The most common measures used to present uncertainty are credible or confidence intervals (CIs).
-Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it is difficult to display effectively.
-The map focuses on displaying the statistic and lacks additional space to represent the uncertainty.
-Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two common solutions.
+<!-- psychology -->
+Displaying familiar state boundaries can make a map easier to read [@CIBMUK] and allow viewers to infer the spatial relationships visually in the data using their mental model of the geography.
+The map users of disease displays may include researchers, the public, policymakers, and the media [@CPISACA].
+For these users, the familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
 
@@
 
+## Additional considerations
+<!-- 479 words -->
+Cancer atlases often display supplementary graphs and plots to add more information.
+Additional materials such as tables, graphs, and text explanations support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers [@CPISACA].
+The many displays of statistical summaries, including dot plots, bar plots, box plots, cumulative distribution plots, scatter plots, and normal probability plots, can provide alternative views of the cancer statistics.
+These can also display supporting statistics such as error, confidence intervals, distributions, sample or population sizes, and standard deviation.
+
+The statistics communicated in atlases are often used to describe differences between areas.
+This can occur at different levels of aggregation.
+Aggregation of global health statistics occurs within administrative and arbitrarily defined regions, such as those used by the World Health Organization and the United Nations [@IARC_3].
+World atlases can allow for displays of data aggregated into continents, countries, states, provinces and congressional districts [@USInteractive].
+Each population area will probably have a different number of people, which is typically used to calibrate the statistic.
+Cancer atlases may also communicate the distribution of the population living in all areas in a table or histogram display [@NICR_1].
+Atlases can connect the population to the land available to them by communicating population density.
+
+Maps can also be used to focus on demographic strata, such as age and sex.
+Some of the digital atlases surveyed allow subsets such as males, females, or those aged over 65, to be selected for display.
+Similarly, socioeconomic indicators, such as unemployment rates, poverty rates, remoteness, and education levels, can be used to filter data, in order to communicate how cancer prevalence varies for different members of society.
+Few atlases provide this level of detail.
+
+Introducing population and demographic information helps to interpret the rates in areas effectively, but there will still be uncertainty around the rates.
+To address this, a cancer atlas often communicates uncertainty about the value of a statistic.
+There are several potential sources of uncertainty: sampling error, errors arising from the disease reporting process (or data collection), and errors arising from the statistical modelling or simulation process.
+The most common measures used to present uncertainty are credible or confidence intervals (CIs).
+Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it is difficult to display effectively.
+The map focuses on displaying the statistic and lacks additional space to represent the uncertainty.
+Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two common solutions.
+
 # User interaction {#ch:interacting}

### 2020-04-03 submission2/submission2.Rmd
begin letter to editors using reviewer feedback

---
from: Stephanie Kobakian
fromqualifications: B.Com and B.Eco
fromemail: stephanie.kobakian@monash.edu
fromphone: +61 433 699 797
fromsignature: sigfile.png
toname: Professor Peter Baade and Dr Susanna Cramb
opening: Dear Professor Peter Baade and Dr Susanna Cramb,
closing: Warm regards
spacing: 1.4
output: MonashEBSTemplates::letter
---

We recently received the document titled *ACE-19-31-R1-from-the-reviewers*, we thank you for the recommendations and changes made within this document.

Your feedback regarding our submission included a title change, please accept this resubmission with the new title "Mapping Cancer: the Potential of Cartograms and Alternative Map Displays", for inclusion in the Focused Issue on Spatial Patterns in Cancer Epidemiology in the Annals of Cancer Epidemiology.

This review paper presents an exploration of cancer mapping methods, focusing on variations on choropleths for online cancer atlases.

**Reviewer suggestion:**
**Add a sentence about the study aims (in the abstract).**

We have included the following sentence in the abstract:

This study summarises current practices for cancer atlases and investigates the alternative map displays that could be used to accurately represent the distribution of cancer statistics across a population, as many cancer atlases lack appropriate displays for population statistics.

**Reviewer suggestion:**
**Given the ethical and privacy considerations for small area data, it is not a straightforward process to obtain and report area-specific counts. I would suggest deleting this sentence and the previous one. **

We acknowledge that counts and incidence rates for small area data are priveledged information and have removed the sentences:

It is easy to provide counts or incidence rates of the diagnoses of these areas. This type of data is routinely collected for public health reasons and may be made available to the general public as a service to the community.

**Reviewer suggestion:**

**We would suggest that the aim of the manuscript could be phrased to highlight the current approaches and limitations of typical visualisation methods for cancer maps, along with outlining possible alternative methods and challenges in implementing these?**

We have restructed our original submission to provide a clear path from current approaches used, to the alternative approaches that should be explored. We highlight the limitations of using current display methods, to encourage the use of alternative display methods. We also present the challenges that may be faced when implementing the alternative designs.

The manuscript title has been changed:
"Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"

The manuscript now contains the following sections:

**Traditional approaches for cancer map displays:** This section discusses the choropleth map as the common map display used in cancer maps and atlases.
*Limitations of choropleth displays:* This section presents the possible issues with displaying cancer statistics on a choropleth map.
*Contemporary alternatives to choropleth maps:* We have shown that there are many alternative displays that highlight population dense regions.
*Challenges of alternative approaches:* Alternative displays also have limitations when displaying spatial distributions of cancer statistics.

We have inclued changes to reflect your minor notes:

*Editor suggestion:*
*The authors mention a focus on Atlases published between 2010 and 2015, but then include a bowel cancer atlas from 2016. *

*Also, within the bowel cancer atlas data, the reporting of percent of males aged 50-54 with bowel cancer in 2016 is incorrect; it should be the % of the ERP in each area that are males aged 50-54 years. *

We have changed the image used from the Bowel Cancer Atlas. To reflect this change the description of the statistic used in the bowel cancer atlas of Australia ahs been corrected:

Bowel Cancer Australia Atlas:
...shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia. It is published by Bowel Cancer Australia https://www.bowelcanceraustralia.org/.

The following section outlines comments made by the reviewer and contains our responses after the suggested changes were implemented.

*Reviewer A: *
*Major*
*Comment 1: *
The Abstract should be more focused- perhaps give an example of how cartogram mapping has been applied to data visualisation. What are the benefits of this approach and suggestions for the future.

*Reply 1:* 
The advantages of alternative mapping displays, including cartograms, has been included in the abstract. It clearly includes the changes to the map display made by alternative visualisation methods, and how they benefit the map user as they build an understanding of the spatial distribution.

*Changes in the text:*

Cancer atlases communicate cancer statistics over geographic domains, typically with a choropleth map. They subdivide these domains into administrative regions such as countries, states, or suburbs. When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner city areas. The smaller geographic areas are important to consider if they are densely populated. Alternative map displays, such as a cartogram or a hexagon tile map, can shift the attention of map users from the large rural areas by decreasing their size on the map display. This means alternative displays can be more effective at accurately communicating spatial patterns across spatial areas, by highlighting areas of interest. It is recommended that alternative displays are included in cancer atlases. In addition, with the ease of today's technology, user interaction with the displays is encouraged. Users should also be able to interactively display different statistics, such as incidence rate or relative incidence.

*Comment 2: *

2. The key message seems to get lost in the Abstract. The Conclusion is vague.

*Reply 2:* 
Changes to the abstract have been made. Additionally, the conclusion has also been edited to be more direct in recommending the use or consideration of the alternatives presented.

*Changes in the text:* 
Please see the response to *Comment 1*.

*Comment 3: *

3. Entire paper needs substantial English language editing

*Reply 3:* 
Thank you for this point. Substantial time has been taken to improve the language used in this paper.

*Comment 4: *

4. Please avoid the use of sentences which have been used throughout the paper where you start with a reference number: such as “[4] discusses the use of (Map displays for disease data, Line 15-16) They do not make any sense and must be modified.

*Reply:* 
The referencing system was changed to implement the springer-vancouver style. We apologise for the issues this caused when exporting to a Word Document, the text has been modified appropriately.

*Comment 5: *

5. Please provide numbers for Figures and Tables

*Reply 5:* 
this begins the discussion, andNumbers for figures and tables have been corrected. This paper was produced using Rmarkdown, this introduced errors with numbering figures, sections and tables. All numbering has now been manually defined.

*Comment 6: *

6. Please clearly state the primary aim of this paper and structure it accordingly

*Reply 6:* 
The primary aim of this paper is to present alternative map displays as viable visualisation methods for presenting spatial distributions in cancer atlases. A choropleth map is commonly used in practice, it is described and it's limitations are highlighted. The alternative map displays address the limitations of the choropleth map. The methods and the limitations of these alternative displays are discussed.

*Comment 7: *

7. Conclusions must discuss the key focus and conclusions of this paper.

*Reply 7:* 
Small changes have been made to the conclusion that make the focus of the paper and recommendations clearer. The conclusion has three parts (1) mapping practices, and recommendations, (2) statistics applied and the pros and cons, and (3) user interaction with online maps and the benefits of incorporating this.

*Minor Comments*

- Please check the use of past and present tense
- There are many minor grammatical errors that need to checked.
- Please edit the last section of the Introduction. Either number each section and then refer to them or find another way to appropriately refer to each section.

All minor comments were addressed in the process of substantial editing.

We thank you for your consideration of our changes.

### 2020-05-14 literature.Rmd
begin letter to editors using reviewer feedback
New figure 1 (include ACA)

hopeful

Table 1 and 2 switched, stats updated

@@
 ---
-title: "Mapping Cancer: the Potential of Cartograms and Alternative Map Displays"
-author: "Stephanie Kobakian^[Science and Engineering Faculty, Queensland University of Technology.], Dianne Cook^[Faculty of Econometrics and Business Statistics, Monash University.] and Jessie Roberts^[Queensland University of Technology.]"
+title: 'Mapping Cancer: the Potential of Cartograms and Alternative Map Displays'
+author: Stephanie Kobakian^[Science and Engineering Faculty, Queensland University
+  of Technology.], Dianne Cook^[Faculty of Econometrics and Business Statistics, Monash
+  University.] and Jessie Roberts^[Queensland University of Technology.]
+csl: springer-vancouver.csl
 output:
+  word_document:
+    toc: no
+    toc_depth: '3'
   pdf_document:
@@
 geometry: margin = 1.2in
-header-includes:
-    - \usepackage{setspace}\doublespacing
+header-includes: \usepackage{setspace}\doublespacing
 link-citations: yes
-csl: springer-vancouver.csl
 fontsize: 11pt
@@
 
-\setcounter{page}{13}
-
 \section*{Abstract}
@@
 A choropleth map subdivides these domains into administrative regions such as countries, states, or suburbs.
-When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner- city areas.
+When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner - city areas.
 The smaller geographic areas are important to consider if they are densely populated.
@@
 A choropleth map is the common display to visualize and communicate geospatial cancer statistics over geographic domain.
-Choropleth maps are a type of thematic map that show polygons for each of the groups of data points, representing the geographic units, where each polygon is shaded with a color according to the area-specific values of the statistic being conveyed.
+Choropleth maps are a type of thematic map that show polygons for each of the groups of data points representing the geographic units, where each polygon is shaded with a color according to the area-specific values of the statistic being conveyed.
 Visualizing this data is helpful as geographic patterns of disease may be obscured when reported in a table [@SAMGIS].
 Providing a visual representation of cancer outcomes allows identification of geographic patterns of the disease that can then be addressed with public health policy and actions.
-The spatial distribution of the disease incidence can be examined using a choropleth and may reveal a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease.
+The spatial distribution of the disease incidence can be examined using a choropleth map display and may reveal a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease.
 One of the key challenges with mapping spatial patterns of disease is the design of visualizations [@SE].
 It is important to consider the strengths and weaknesses of designs, as visualizing diseases on maps is often the first step in exploratory spatial data analysis and helps in the formulation of hypotheses.
+
+<!-- # Cancer atlases -->
+<!-- Displaying of cancer data on choropleth maps -->
+A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
+Atlases are key to developing hypotheses regarding areas with unusually high rates, and geographic correlations [@MACM].
+The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization.
+Cancer maps and atlases date back to Haviland's maps in 1875, with more modern atlases directly evolving from early work in US cancer atlases, appearing in 1971 [@burbank].
+The presentation of cancer statistics has increased with greater access to computational power and the availability of geographic information systems software [@SE].
+
+<!-- Measures reported in cancer atlases -->
+<!-- Structure of the paper -->
+
 This paper considers the current visualization techniques to communicate statistics to the public and their applications to cancer statistics.
@@
 
-<!-- Structure of the paper -->
 The paper is structured as follows.
-The next section describes the choropleth map, which is the common approach to disease maps and presents examples of atlases in use today and discusses the limitations of the choropleth map.
+The next section introduces cancer atlases and the worldwide implementations.
+Choropleths are the common approach to disease maps and presents examples of atlases in use today and discusses the limitations of the choropleth map.
 Section 3 describes alternative displays, including the cartogram, which is useful when the map has heterogeneously sized geographic units.
@@
 
-# Cancer atlases {#ch:public}
-<!-- 360 words -->
-<!-- Displaying of cancer data on choropleth maps -->
-A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
-Atlases are key to developing hypotheses regarding areas with unusually high rates, and geographic correlations [@MACM].
-The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization.
-Cancer maps and atlases date back to Haviland's maps in 1875, and early work in US cancer atlases appearing in 1971 [@burbank].
-The presentation of cancer statistics has increased with greater access to computational power and the availability of geographic information systems software [@SE].
-
-<!-- Measures reported in cancer atlases -->
-Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the public and others not trained in statistical analysis.
-These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analysis.
-Cruickshank (1947) as cited by S.
-D.
-Walter [@DMAHP], discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
-Overwhelmingly, choropleth maps are visualisations chosen to communicate cancer statistics to members of the public and other non-expert audiences.
-
-```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = " A selection of choropleth cancer maps from online atlases that are publicly available.
-Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends.
-These atllases are described in Table 1.", fig.align='center', fig.width=12, fig.height=10}
-knitr::include_graphics("figures/choropleth_grid.png")
-```
-
-Table 1: <!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases.
-
-|Fig.| Atlas | Statistic | Data source | 
-|:---:|:------------|:----------------------|:----------------|  
-|1a | The Environment and Health Atlas of England and Wales | relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2]| [ Office for National Statistics (ONS) (England) and from the Welsh Cancer Intelligence and Surveillance Unit (WCISU) ](http://www.envhealthatlas.co.uk/homepage/)|
-|1b | Globocan 2018: Estimated Cancer Incidence, Mortality and Prevalence Worldwide | age standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018 [@Globocan]| [World Health Organization's International Agency for Research on Cancer](https://www.iarc.fr/).
-|
-|1c | Atlas of Cancer in Queensland | the relative incidence ratio of lung cancer in males in the state of QLD within Australia based on data from 1998 to 2007, Queensland Cancer Council[@QLDcancerAtlas] | [Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/) |
-|1d | Bowel Cancer Australia Atlas | the percentage of Australian males between 50 - 54 years of age diagnosed with bowel cancer in 2016 in Australia [@Bowel].
-|  [Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)|
-|1e | United States Cancer Statistics: An Interactive Cancer Statistics Website | the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level [@USInteractive].|  *[Centers for Disease Control and Prevention](https://www.cdc.gov/)*, with data from state cancer registries.
-|
-|1f | Map of Cancer Mortality Rates in Spain | side by side maps of relative risk of lung cancer for men vs women for 2004 to 2008 [@cancerSpain].| [Map of cancer mortality rates in Spain](https://elpais.com/elpais/2014/10/06/media/1412612722_141933.html)|
-|1g | Atlas of Childhood Cancer in Ontario | the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
-| [The Pediatric Oncology Group of Ontario Networked Information System](https://www.pogo.ca/research-data/pogonis-childhood-cancer-database/data-requests/)|
-
 # Measures mapped
@@
 Epidemiologists and statisticians have developed the statistics used to communicate the burden of cancer over several decades.
-Table 2<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases.
+Table 1<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases.
 Mortality rates are commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations.
@@
 
-Table 2: <!--\label{tab:measures}--> Common measures for reporting cancer information.
+XXX
+Some of the measures used to present cancer statistics are aggregations.
+The counts of cases in small areas can be difficult to obtain as they are often protected for privacy reasons.
+The information released and the statistics presented in worldwide cancer atlases are often modelled estimates.
+The measures described in table use Incidence as the statistic of interest, this is the amount of new cases and the same statistics can be used to describe the mortality, in this case Standardised Incidence Rate is analogous to the Standarised Mortality Rate.
 
-|Measure |Details| 
-|:--------------------|:------------------------------------------------|  
-|1. Count| Crude cancer counts|
-|2. Rate per 100,000| Cancer incidence per 100,000 population|  
-|3. IR (Incidence Ratio)| $(IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}$, |
-|| The cancer incidence rate in region $i$ over the average cancer incidence rate for all of the regions|
-|4. Age-Adjusted Rate per 100,000| Standardized by age structure or region|   
-|5. Age-Adjusted Relative Risk| Standardized by age structure in each region $i$|  
-|6. SIR (Standardized Incidence Ratio)| Incidence standardized by population at risk in each region $i$|
-|7. Below or above Expected| An alternative expression of the SIR|
-|8. RER | $RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}$|
-|(Relative Excess Risk)| Represents the estimate of cancer-related mortality within five years of diagnosis. Also referred to as 'excess hazard ratio'|
+<!--
+Cancer mortality is the number of deaths from cancer in a specific population within a specific period of time, usually a year.
+It usually only includes deaths where cancer is mentioned as an underlying cause of death on death certificates.
 
-<!-- 416 words -->
-Roberts [@roberts2019communication] identified 33 cancer atlases published between 2010 and 2018.
-Each of these online atlases uses choropleth maps.
-All except one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships.
-Figure 1<!--\ref{fig:choropleth-grid-create}--> displays a subset of maps from these atlases, the selection varies in the geographies explored.
-Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows Globocan 2018 [@Globocan] which explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide using data sourced from cancer registries of each country.
-The Bowel Cancer Australia Atlas in Figure 1<!--\ref{fig:choropleth-grid-create}-->d presents an example of a cancer specific atlas -- it shows the average Standardized Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 [@Bowel].
-Like many of the atlases examined, there is a choice of gender displayed in the Bowel Cancer Atlas.
-Gender is displayed in side-by-side maps in the Map of Cancer Mortality Rates in Spain (Figure 1f)<!--\ref{fig:choropleth-grid-create}--> [@cancerSpain].
+Cancer mortality rates are a standard measure of the frequency of deaths within a specific period of time relative to a fixed population size, usually 100,000 people or expressed per million if rare.
+-->
 
-Resolution of the maps varies greatly.
-Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows global information at a national level.
-The United States Cancer Statistics [@USInteractive] shows data aggregated at the state level.
-The Environment and Health Atlas of England and Wales [@EnvEnglandWales2] (Figure 1<!--\ref{fig:choropleth-grid-create}-->a) shows the relative risk for women developing lung cancer  at a neighborhood (small-area) scale.
-The Atlas of Cancer in Queensland (Figure 1<!--\ref{fig:choropleth-grid-create}-->c) shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2016] in the state of Queensland within Australia [@QLDcancerAtlas].
+Table 1: <!--\label{tab:measures}--> Common measures for reporting cancer information.
 
-Age-specific atlases are less common.
-Figure 1<!--\ref{fig:choropleth-grid-create}-->g displays Atlas of Childhood Cancer in Ontario, this communicates the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
+|Measure |Details| Fig. |
+|:--------------------|:------------------------------------------------|:--------------------|
+|1. Count| Cancer incidence counts| |
+|2.
+Incidence Rate (per 100,000) | The frequency of cases relative to a fixed population size, usually 100,000 people, within a specific period of time XXX [cite it](https://www.cancerresearchuk.org/health-professional/cancer-statistics/cancer-stats-explained/statistics-terminology-explained#heading-Eleven)|  |
+|3. IR (Incidence Ratio)| The incidence rate in a region divided by the average  incidence rate for all of the regions| |
+|4. Age-Standardized Rate per 100,000| Standardized by age structure or region|   |
+|5. Age-Standardized Relative Risk| Standardized by age structure|  |
+|6. SIR (Standardized Incidence Ratio)| The observed number of cases in a region divided by the expected number of cases, if that region had the same age-specific rates as a reference population.  |
+|7. Excess Hazard Ratio | The estimate of cancer-related mortality within five years of diagnosis. |  |
 
@@
 Displaying familiar state boundaries can make a map easier to read [@CIBMUK] and allow viewers to infer the spatial relationships visually in the data using their mental model of the geography.
-The map users of disease displays may include researchers, the public, policymakers, and the media [@CPISACA].
+The users of disease maps may include researchers, the public, policy makers, and the media [@CPISACA].
 For these users, the familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
 
-## Limitations of choropleth displays {#ch:chorolimit}
+Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the public and others not trained in statistical analysis.
+These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analysis.
+Cruickshank (1947) as cited by S.
+D.
+Walter [@DMAHP], discusses using visuals as a 'formal statistical assessment of the spatial pattern'.
+Overwhelmingly, choropleth maps are visualisations chosen to communicate cancer statistics to members of the public and other non-expert audiences.
 
-Australia presents an extreme case of an urban rural divide. The land mass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas [@ACTUC].
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = " A selection of choropleth cancer maps from online atlases that are publicly available.
+Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends.
+These atlases are described in Table 2.", fig.align='center', fig.width=12, fig.height=10}
+knitr::include_graphics("figures/choropleth_grid.png")
+```
 
+Table 2: <!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases displayed in Figure 1.
+
+| Fig. | Atlas | Data | Statistic | Source |
+|------|---------------------------------------------------------------------------------------------------------|-----------------------------------------------------|------------------------------------------------------|---------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
+| a | The Environment and Health<br>Atlas of England and Wales<br>[@EnvEnglandWales2] | Sex: Women<br>Year: 2010<br>Cancer: Lung | Relative risk | [Office for National Statistics<br>(ONS) (England) and from the<br>Welsh Cancer Intelligence and<br>Surveillance Unit (WCISU)](http://www.envhealthatlas.co.uk/homepage/) |
+| b | Globocan 2018: Estimated Cancer<br>Incidence, Mortality and <br><br>prevalence Worldwide<br>[@Globocan] | Sex: Both<br>Year: 2018<br>Cancer: All invasive | Age standardized<br>incidence rates<br>(per 100,000) | [World Health Organization's<br>International Agency for<br>Research on Cancer](https://www.iarc.fr/<br> |
+| c | Atlas of Cancer in Queensland<br> [@QLDcancerAtlas] | Sex: Males<br>Year: 1998 - 2007<br>Cancer: Lung | Relative incidence<br>ratio | [Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/) |
+| d | Australian Cancer Atlas<br>[@ACA] | Sex: Males<br>Year: 2010 - 2014<br>Cancer: Lung | Average standardised<br>incidence rate | [Australian Cancer Atlas](https://atlas.cancer.org.au/) |
+| e | United States Cancer<br>Statistics: An Interactive<br>Cancer Statistics Website<br> | Sex: Both<br>Year: 2016<br>Cancer: All types | Incidence rate per<br>100,000 | With data from state cancer<br>registries: <br>[Centers for Disease Control and Prevention](https://www.cdc.gov/) |
+| f | Map of Cancer Mortality<br>Rates in Spain<br>[@cancerSpain] | Sex: Both<br>Year: 2004 - 2008<br>Cancer: All types | Relative risk | <br>[Map of cancer mortality rates<br>in Spain](https://elpais.com/elpais/2014/10/06/media/1412612722_141933.html |
+| g | Atlas of Childhood<br>Cancer in Ontario<br>[@OntarioPediatric] | Sex: Both<br>Year: 1995 - 2004<br>Cancer: All types | Incidence rate per<br>100,00 | <br>[The Pediatric Oncology Group<br>of Ontario Networked Information<br>System](https://www.pogo.ca/research-data/pogonis-childhood-cancer-database/data-requests/) |
+
+<!-- 416 words -->
+A review of modern cancer atlases [@roberts2019communication] identified 33 cancer atlases published between 2010 and 2016.
+Atlases published between 2016 and 2018 have also been considered.
+Each of these online atlases uses choropleth maps.
+All except one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships.
+Figure 1<!--\ref{fig:choropleth-grid-create}--> displays a subset of maps from these atlases, the selection varies in the geographies explored.
+Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows Globocan 2018 [@Globocan] which explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide using estimates based on available country cancer registries.
+Many atlases allow users to view sex-specific distributions, however sex is displayed in side-by-side maps in the Map of Cancer Mortality Rates in Spain (Figure 1f)<!--\ref{fig:choropleth-grid-create}--> [@cancerSpain].
+
+There is large variation in the resolution of the maps.
+Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows global information at a national level.
+The United States Cancer Statistics [@USInteractive] shows data aggregated at the state level (n = 51).
+The Environment and Health Atlas of England and Wales [@EnvEnglandWales2] (Figure 1<!--\ref{fig:choropleth-grid-create}-->a) shows the relative risk for women developing lung cancer  at a neighborhood (small-area) scale (n = 8850).
+The Asutralian Cancer Atlas (Figure 1<!--\ref{fig:choropleth-grid-create}-->d) shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2016] (n = 2292).
+The Atlas of Cancer in Queensland (Figure 1<!--\ref{fig:choropleth-grid-create}-->c) shows a subset of the Australian Statistical Areas at Level 2 [@abs2016] located in the state of Queensland within Australia [@QLDcancerAtlas] (n = 2292).
+
+Age-specific atlases are less common.
+Figure 1<!--\ref{fig:choropleth-grid-create}-->g displays Atlas of Childhood Cancer in Ontario, this communicates the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
+
+Australia presents an extreme case of an urban rural divide.
+The land mass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas [@ACTUC].
 Choropleth maps provide a familiar display, which shows data in a geographically recognisable way.

### 2020-05-14 submission2/submission2.Rmd
begin letter to editors using reviewer feedback
New figure 1 (include ACA)

hopeful

Table 1 and 2 switched, stats updated

@@
 **Reviewer suggestion:**
-**Given the ethical and privacy considerations for small area data, it is not a straightforward process to obtain and report area-specific counts.
-I would suggest deleting this sentence and the previous one.
-**
+**Could we structure paper as:**
+**Introduction (also introduces cancer atlases)**
+**Measures mapped (info around Table 2)**
+**Visualisation approaches**
+**- Choropleth (mention advant and disadvant, along with figures from cancer atlases)**
+**- Cartograms**
+**- Tile maps**
+**- Geofacets**
+**- Multivariate maps**
+**Comparison & critique of alternative methods**
+**Additional considerations**
+**- Associated graphs**
+**- Displaying uncertainty measures**
+**- Interactivity**
+**Recommendations for mapping cancer**
+**Conclusion**
 
-We acknowledge that counts and incidence rates for small area data are priveledged information and have removed the sentences:
-
-It is easy to provide counts or incidence rates of the diagnoses of these areas.
-This type of data is routinely collected for public health reasons and may be made available to the general public as a service to the community.
+We have rearranged the sections from the "Introduction" to "Traditional approaches for cancer map displays" in this submission at the request of the reviewer.
+The discussion of choropleth maps considers the advantages and disadvantages now occurs in the same section.
 
 **Reviewer suggestion:**
+**Suggest also noting development of statistical methods and software**
 
-**We would suggest that the aim of the manuscript could be phrased to highlight the current approaches and limitations of typical visualisation methods for cancer maps, along with outlining possible alternative methods and challenges in implementing these?**
+**Reviewer suggestion:**
+**Give the actual reference for Cruikshank 1947**
 
-We have restructed our original submission to provide a clear path from current approaches used, to the alternative approaches that should be explored.
-We highlight the limitations of using current display methods, to encourage the use of alternative display methods.
-We also present the challenges that may be faced when implementing the alternative designs.
+This quote has been removed.
 
-The manuscript title has been changed:
-"Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"
+**Reviewer suggestion:**
+**Suggest including Australian Cancer Atlas**
+We have replaced the Bowel Cancer Australia Atlas with the Australian Cancer Atlas.
 
-The manuscript now contains the following sections:
+**Reviewer suggestion:**
+**This table might benefit from an additional column, which gives an example of a cancer map using this measure. Also, some of these measures can be calculated.**
 
-**Traditional approaches for cancer map displays:** This section discusses the choropleth map as the common map display used in cancer maps and atlases.
-*Limitations of choropleth displays:* This section presents the possible issues with displaying cancer statistics on a choropleth map.
-*Contemporary alternatives to choropleth maps:* We have shown that there are many alternative displays that highlight population dense regions.
-*Challenges of alternative approaches:* Alternative displays also have limitations when displaying spatial distributions of cancer statistics.
-
-We have inclued changes to reflect your minor notes:
-
-*Editor suggestion:*
-*The authors mention a focus on Atlases published between 2010 and 2015, but then include a bowel cancer atlas from 2016. *
-
-*Also, within the bowel cancer atlas data, the reporting of percent of males aged 50-54 with bowel cancer in 2016 is incorrect; it should be the % of the ERP in each area that are males aged 50-54 years.
-*
-
-We have changed the image used from the Bowel Cancer Atlas. To reflect this change the description of the statistic used in the bowel cancer atlas of Australia ahs been corrected:
-
-Bowel Cancer Australia Atlas:
-...shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia.
-It is published by Bowel Cancer Australia https://www.bowelcanceraustralia.org/.
-
-The following section outlines comments made by the reviewer and contains our responses after the suggested changes were implemented.
-
-*Reviewer A: *
-*Major*
-*Comment 1: *
-The Abstract should be more focused- perhaps give an example of how cartogram mapping has been applied to data visualisation. What are the benefits of this approach and suggestions for the future.
-
-*Reply 1:* 
-The advantages of alternative mapping displays, including cartograms, has been included in the abstract.
-It clearly includes the changes to the map display made by alternative visualisation methods, and how they benefit the map user as they build an understanding of the spatial distribution.
-
-*Changes in the text:*
-
-Cancer atlases communicate cancer statistics over geographic domains, typically with a choropleth map.
-They subdivide these domains into administrative regions such as countries, states, or suburbs.
-When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner city areas.
-The smaller geographic areas are important to consider if they are densely populated.
-Alternative map displays, such as a cartogram or a hexagon tile map, can shift the attention of map users from the large rural areas by decreasing their size on the map display.
-This means alternative displays can be more effective at accurately communicating spatial patterns across spatial areas, by highlighting areas of interest.
-It is recommended that alternative displays are included in cancer atlases.
-In addition, with the ease of today's technology, user interaction with the displays is encouraged.
-Users should also be able to interactively display different statistics, such as incidence rate or relative incidence.
-
-*Comment 2: *
-
-2. The key message seems to get lost in the Abstract. The Conclusion is vague.
-
-*Reply 2:* 
-Changes to the abstract have been made. Additionally, the conclusion has also been edited to be more direct in recommending the use or consideration of the alternatives presented.
-
-*Changes in the text:* 
-Please see the response to *Comment 1*.
-
-*Comment 3: *
-
-3. Entire paper needs substantial English language editing
-
-*Reply 3:* 
-Thank you for this point. Substantial time has been taken to improve the language used in this paper.
-
-*Comment 4: *
-
-4.
-Please avoid the use of sentences which have been used throughout the paper where you start with a reference number: such as “[4] discusses the use of (Map displays for disease data, Line 15-16) They do not make any sense and must be modified.
-
-*Reply:* 
-The referencing system was changed to implement the springer-vancouver style. We apologise for the issues this caused when exporting to a Word Document, the text has been modified appropriately.
-
-*Comment 5: *
-
-5. Please provide numbers for Figures and Tables
-
-*Reply 5:* 
-this begins the discussion, andNumbers for figures and tables have been corrected.
-This paper was produced using Rmarkdown, this introduced errors with numbering figures, sections and tables.
-All numbering has now been manually defined.
-
-*Comment 6: *
-
-6. Please clearly state the primary aim of this paper and structure it accordingly
-
-*Reply 6:* 
-The primary aim of this paper is to present alternative map displays as viable visualisation methods for presenting spatial distributions in cancer atlases.
-A choropleth map is commonly used in practice, it is described and it's limitations are highlighted.
-The alternative map displays address the limitations of the choropleth map.
-The methods and the limitations of these alternative displays are discussed.
-
-*Comment 7: *
-
-7. Conclusions must discuss the key focus and conclusions of this paper.
-
-*Reply 7:* 
-Small changes have been made to the conclusion that make the focus of the paper and recommendations clearer.
-The conclusion has three parts (1) mapping practices, and recommendations, (2) statistics applied and the pros and cons, and (3) user interaction with online maps and the benefits of incorporating this.
-
-*Minor Comments*
-
-- Please check the use of past and present tense
-- There are many minor grammatical errors that need to checked.
-- Please edit the last section of the Introduction. Either number each section and then refer to them or find another way to appropriately refer to each section.
+Several suggestions were made regarding Table 2: Common measures for reporting cancer information.
+The additional information requested is communicated more effectively using a paragraph style format.
 

### 2020-05-14 review_figures.R

@@ -1,12 +1,13 @@
 # Figures
-
+library(ggplot2)
+library(grid)
 library(cowplot)
 
 # Import the images
 plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
 plot2 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
 plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
-plot4 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
+plot4 <- png::readPNG('cancer_map_images/choropleths/aca.choropleth.png')
 plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
 plot6 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
 plot7 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
@@ -19,6 +20,7 @@ ggdraw(xlim = c(0,1), ylim = c(0,1)) +
   draw_plot(rasterGrob(plot5), 0.33, 0.5, .33, .25) +
   draw_plot(rasterGrob(plot6), 0, 0, .66, .5) +
   draw_plot(rasterGrob(plot7), .66, 0, .33, .5) +
+  draw_plot(rasterGrob(plot8), .66, 0, .33, .5) +
   draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
     c(0.025, 0.33, 0.66, 0.025, 0.33, 0.025, 0.66), 
     c(0.99, 0.99, 0.99, 0.74, 0.74, 0.46, 0.46), size = 20)

### 2020-05-15 literature.Rmd
Edited according to reviewers suggestions up to Limitations of choropleth displays

@@
 
-<!-- Ways to visualize, whats our hypothesis -->
-A choropleth map is the common display to visualize and communicate geospatial cancer statistics over geographic domain.
-Choropleth maps are a type of thematic map that show polygons for each of the groups of data points representing the geographic units, where each polygon is shaded with a color according to the area-specific values of the statistic being conveyed.
-Visualizing this data is helpful as geographic patterns of disease may be obscured when reported in a table [@SAMGIS].
-Providing a visual representation of cancer outcomes allows identification of geographic patterns of the disease that can then be addressed with public health policy and actions.
-The spatial distribution of the disease incidence can be examined using a choropleth map display and may reveal a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease.
-One of the key challenges with mapping spatial patterns of disease is the design of visualizations [@SE].
-It is important to consider the strengths and weaknesses of designs, as visualizing diseases on maps is often the first step in exploratory spatial data analysis and helps in the formulation of hypotheses.
-
 <!-- # Cancer atlases -->
 <!-- Displaying of cancer data on choropleth maps -->
-A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries.
+A cancer atlas is a map, or collection of maps, commoly representing cancer incidence or mortality patterns across a country, or group of countries.
 Atlases are key to developing hypotheses regarding areas with unusually high rates, and geographic correlations [@MACM].
@@
 
+# Visualisation approaches
+
 # Traditional approaches for cancer map displays {#ch:choropleth-maps}
+
+<!-- Ways to visualize, whats our hypothesis -->
+A choropleth map is the common display to visualize and communicate geospatial cancer statistics over geographic domain.
+Choropleth maps are a type of thematic map that show polygons for each of the groups of data points representing the geographic units, where each polygon is shaded with a color according to the area-specific values of the statistic being conveyed.
+Visualizing this data is helpful as geographic patterns of disease may be obscured when reported in a table [@SAMGIS].
+Providing a visual representation of cancer outcomes allows identification of geographic patterns of the disease that can then be addressed with public health policy and actions.
+The spatial distribution of the disease incidence can be examined using a choropleth map display and may reveal a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease.
+One of the key challenges with mapping spatial patterns of disease is the design of visualizations [@SE].
+It is important to consider the strengths and weaknesses of designs, as visualizing diseases on maps is often the first step in exploratory spatial data analysis and helps in the formulation of hypotheses.
+
 <!-- 381 words -->
@@
 The Environment and Health Atlas of England and Wales [@EnvEnglandWales2] (Figure 1<!--\ref{fig:choropleth-grid-create}-->a) shows the relative risk for women developing lung cancer  at a neighborhood (small-area) scale (n = 8850).
-The Asutralian Cancer Atlas (Figure 1<!--\ref{fig:choropleth-grid-create}-->d) shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2016] (n = 2292).
-The Atlas of Cancer in Queensland (Figure 1<!--\ref{fig:choropleth-grid-create}-->c) shows a subset of the Australian Statistical Areas at Level 2 [@abs2016] located in the state of Queensland within Australia [@QLDcancerAtlas] (n = 2292).
+The Australian Cancer Atlas (Figure 1<!--\ref{fig:choropleth-grid-create}-->d) shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2016] (n = 2292).
+The Atlas of Cancer in Queensland (Figure 1<!--\ref{fig:choropleth-grid-create}-->c) shows a subset of the Australian Statistical Local Areas (SLAs) located in the state of Queensland within Australia [@QLDcancerAtlas] (n = 478).
 
@@
 Maps can also be used to focus on demographic strata, such as age and sex.
-Some of the digital atlases surveyed allow subsets such as males, females, or those aged over 65, to be selected for display.
-Similarly, socioeconomic indicators, such as unemployment rates, poverty rates, remoteness, and education levels, can be used to filter data, in order to communicate how cancer prevalence varies for different members of society.
+Some of the digital atlases surveyed allow subsets to be selected for display, The Australian Cancer Atlas [@ACA] allows filtering for the selection of statistics regarding males and females specifically.
+Similarly, remoteness [@ACA], socioeconomic indicators, such as unemployment rates, poverty rates, and education levels, can be used to filter data, in order to communicate how cancer prevalence varies for different population subgroups.
 Few atlases provide this level of detail.
+XXX
 
@@
 To address this, a cancer atlas often communicates uncertainty about the value of a statistic.
-There are several potential sources of uncertainty: sampling error, errors arising from the disease reporting process (or data collection), and errors arising from the statistical modelling or simulation process.
-The most common measures used to present uncertainty are credible or confidence intervals (CIs).
+There are several potential sources of uncertainty: sampling error, errors arising from the disease reporting or data collection processes, and uncertainty arising from the statistical modelling or simulation process.
+The most common measures used to present uncertainty are credible intervals or confidence intervals.
 Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it is difficult to display effectively.
-The map focuses on displaying the statistic and lacks additional space to represent the uncertainty.
 Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two common solutions.

### 2020-05-15 submission2/submission2.Rmd
Edited according to reviewers suggestions up to Limitations of choropleth displays

@@
 
+The following reviewer suggestions relate to the introduction:
+
+**Reviewer suggestion:**
+**Given the ethical and privacy considerations for small area data, it is not a straightforward process to obtain and report area-specific counts.**
+
+We agree that it is not easy to obtain and report area-specific counts.
+The statement has been removed.
+
+**Reviewer suggestion:**
+**Suggest moving paragraph to proposed choropleth maps section**
+
+This paragraph has been moved the choropleth section under visualisation approches.
+
+**Reviewer suggestion:**
+**Can you include mention of how this is different to a thematic map?**
+
+We have included in the choropleth section:
+Choropleth maps are a type of thematic map that show polygons for each of the groups of data points representing the geographic units, where each polygon is shaded with a color according to the area-specific values of the statistic being conveyed.
+
 **Reviewer suggestion:**
 **Suggest also noting development of statistical methods and software**
+The paragraph that began the *Cancer Atlases* section has been moved to the introduction. This paragraph describes cancer atlases, and the developments oof statistical methods and software.
 
@@
 Several suggestions were made regarding Table 2: Common measures for reporting cancer information.
-The additional information requested is communicated more effectively using a paragraph style format.
+Additional columns have been included.
+
+XXX
+
+**Reviewer suggestion:**
+**This is the wrong reference.**
+The incorrect reference to the use of Statistical Areas at Level 2 for the Atlas of Cancer in Queensland has been corrected to reflect that the atlas used Statistical Local Areas.
+
+**Reviewer suggestion:**
+The following paragraph was included at the suggestion of the reviewer:
+
+As data becomes more sparse, either by increasing geographic resolution or stratifying by age/sex, obtaining reliable estimates becomes more difficult.
+Many atlases analyzing smaller areas use statistical modelling to produce estimates, although some continue to use simple calculations and suppress regions where estimates are unstable (e.g.
+(mention some)).
+XXX
+
+**Reviewer suggestion:**
+**These are very different terms, so ideally should not have the same abbreviation. **
+The terms credible intervals and confidence intervals are no longer abbreviated.
+
+**Reviewer suggestion:**
+**Which map?**
+The following sentence has been removed: The map focuses on displaying the statistic and lacks additional space to represent the uncertainty. 
+As the Australian Cancer Atlas example presented in Figure 1 does include uncertainty using transparency.
 

### 2020-05-17 literature.Rmd
Adjusted heading levels

@@
 
-# Traditional approaches for cancer map displays {#ch:choropleth-maps}
+## Traditional approaches for cancer map displays {#ch:choropleth-maps}
 
@@
 
-# Contemporary alternatives to choropleth maps {#ch:alternatives}
+## Contemporary alternatives to choropleth maps {#ch:alternatives}
 
-## Cartograms
+### Cartograms
 <!-- 1496 words inc table-->
@@
 
-## Tile Map
+### Tile Map
 <!-- 173 words -->
@@
 
-## Geofacet
+### Geofacet
 <!-- 166 words -->
@@
 
-## Multivariate displays
+### Multivariate displays
 <!-- 222 words -->
@@
 
-## Neither choropleth maps or cartograms perform well for Australia
+## Performance of choropleth maps or cartograms for Australia
 

### 2020-05-17 literature.Rmd
confirmed use and description of statistics in atlas examples

@@
 
-Table 1: <!--\label{tab:measures}--> Common measures for reporting cancer information.
+Table 1: <!--\label{tab:measures}--> Common measures for reporting cancer information. Each measure is given along with a description and the Atlases in Figure 1 that display each measure.
 
-|Measure |Details| Fig. |
-|:--------------------|:------------------------------------------------|:--------------------|
-|1. Count| Cancer incidence counts| |
-|2.
-Incidence Rate (per 100,000) | The frequency of cases relative to a fixed population size, usually 100,000 people, within a specific period of time XXX [cite it](https://www.cancerresearchuk.org/health-professional/cancer-statistics/cancer-stats-explained/statistics-terminology-explained#heading-Eleven)|  |
-|3. IR (Incidence Ratio)| The incidence rate in a region divided by the average  incidence rate for all of the regions| |
-|4. Age-Standardized Rate per 100,000| Standardized by age structure or region|   |
-|5. Age-Standardized Relative Risk| Standardized by age structure|  |
-|6. SIR (Standardized Incidence Ratio)| The observed number of cases in a region divided by the expected number of cases, if that region had the same age-specific rates as a reference population.  |
-|7. Excess Hazard Ratio | The estimate of cancer-related mortality within five years of diagnosis. |  |
+|Measure | Details | Fig. |
+|:--------------------|:---------------------------------|:----|
+|1. Cancer incidence counts| the number of new cases occurring in a specified period and geographic area [@Globocan]| |
+|2. Incidence Rate (per 100,000) | The frequency of cases relative to a fixed population size, usually 100,000 people, within a specific period of time [@Globocan] | e, g |
+|3. Incidence Ratio | The incidence rate in a region divided by the average incidence rate for all of the regions| c |
+|4.
+Age-Standardized Incidence Rate (per 100,000) | The weighted mean of the age-specific incidence rates; the weights are taken from population distribution of the standard population.
+The most frequently used standard population is the World Standard Population.
+[@Globocan] | b, d |
+|5.
+Age-Standardized Relative Risk | The relative risks represent the risk of an area (ward) relative to average risk, adjusted for age, deprivation and chance fluctuations due to small numbers.
+[@EnvEnglandWales2] | a, f |
+
+Table describes the common measures used to present cancer incidence in atlases.
+The calculations used to produce these measures often are also applied to mortality statistics.The Excess Hazard Ratio and Standardized Incidence Ratio are additional measures that can be found in atlases, but are not used in the images displayed in Figure 1.
 
@@
 | Fig. | Atlas | Data | Statistic | Source |
-|------|---------------------------------------------------------------------------------------------------------|-----------------------------------------------------|------------------------------------------------------|---------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
-| a | The Environment and Health<br>Atlas of England and Wales<br>[@EnvEnglandWales2] | Sex: Women<br>Year: 2010<br>Cancer: Lung | Relative risk | [Office for National Statistics<br>(ONS) (England) and from the<br>Welsh Cancer Intelligence and<br>Surveillance Unit (WCISU)](http://www.envhealthatlas.co.uk/homepage/) |
-| b | Globocan 2018: Estimated Cancer<br>Incidence, Mortality and <br><br>prevalence Worldwide<br>[@Globocan] | Sex: Both<br>Year: 2018<br>Cancer: All invasive | Age standardized<br>incidence rates<br>(per 100,000) | [World Health Organization's<br>International Agency for<br>Research on Cancer](https://www.iarc.fr/<br> |
-| c | Atlas of Cancer in Queensland<br> [@QLDcancerAtlas] | Sex: Males<br>Year: 1998 - 2007<br>Cancer: Lung | Relative incidence<br>ratio | [Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/) |
-| d | Australian Cancer Atlas<br>[@ACA] | Sex: Males<br>Year: 2010 - 2014<br>Cancer: Lung | Average standardised<br>incidence rate | [Australian Cancer Atlas](https://atlas.cancer.org.au/) |
-| e | United States Cancer<br>Statistics: An Interactive<br>Cancer Statistics Website<br> | Sex: Both<br>Year: 2016<br>Cancer: All types | Incidence rate per<br>100,000 | With data from state cancer<br>registries: <br>[Centers for Disease Control and Prevention](https://www.cdc.gov/) |
-| f | Map of Cancer Mortality<br>Rates in Spain<br>[@cancerSpain] | Sex: Both<br>Year: 2004 - 2008<br>Cancer: All types | Relative risk | <br>[Map of cancer mortality rates<br>in Spain](https://elpais.com/elpais/2014/10/06/media/1412612722_141933.html |
-| g | Atlas of Childhood<br>Cancer in Ontario<br>[@OntarioPediatric] | Sex: Both<br>Year: 1995 - 2004<br>Cancer: All types | Incidence rate per<br>100,00 | <br>[The Pediatric Oncology Group<br>of Ontario Networked Information<br>System](https://www.pogo.ca/research-data/pogonis-childhood-cancer-database/data-requests/) |
+|------|------------|-----------|----------------------|-----------------------------|
+| a | The Environment and Health<br> Atlas of England and Wales<br> [@EnvEnglandWales2] | Sex: Women<br> Year: 2010<br> Cancer: Lung | Relative risk | [Office for National Statistics<br> (ONS) (England) and from the<br> Welsh Cancer Intelligence and<br> Surveillance Unit (WCISU)](http://www.envhealthatlas.co.uk/homepage/) |
+| b | Globocan 2018: Estimated Cancer<br> Incidence, Mortality and <br> <br> prevalence Worldwide<br> [@Globocan] | Sex: Both<br> Year: 2018<br> Cancer: All invasive | Age standardized<br> incidence rates<br> (per 100,000) | [World Health Organization's<br> International Agency for<br> Research on Cancer](https://www.iarc.fr/<br>  |
+| c | Atlas of Cancer in Queensland<br>  [@QLDcancerAtlas] | Sex: Males<br> Year: 1998 - 2007<br> Cancer: Lung | Relative incidence<br> ratio | [Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/) |
+| d | Australian Cancer Atlas<br> [@ACA] | Sex: Males<br> Year: 2010 - 2014<br> Cancer: Lung | Average standardised<br> incidence rate | [Australian Cancer Atlas](https://atlas.cancer.org.au/) |
+| e | United States Cancer<br> Statistics: An Interactive<br> Cancer Statistics Website<br>  | Sex: Both<br> Year: 2016<br> Cancer: All types | Incidence rate per<br> 100,000 | With data from state cancer<br> registries: <br> [Centers for Disease Control and Prevention](https://www.cdc.gov/) |
+| f | Map of Cancer Mortality<br> Rates in Spain<br> [@cancerSpain] | Sex: Both<br> Year: 2004 - 2008<br> Cancer: All types | Relative risk | <br> [Map of cancer mortality rates<br> in Spain](https://elpais.com/elpais/2014/10/06/media/1412612722_141933.html |
+| g | Atlas of Childhood<br> Cancer in Ontario<br> [@OntarioPediatric] | Sex: Both<br> Year: 1995 - 2004<br> Cancer: All types | Incidence rate per<br> 100,00 | <br> [The Pediatric Oncology Group<br> of Ontario Networked Information<br> System](https://www.pogo.ca/research-data/pogonis-childhood-cancer-database/data-requests/) |
 
@@
 When communicating information using colour, a map creator should use a scheme that has a linear color gradient, with perceptually uniform color spaces that match equal steps in data space with equal steps in the colour space [@PUCS].
+It is possible to allow for data to progress uniformly in both positive and negative directions from a mid point, such as the mean of the data.
+These diverging colour schemes pair two sequential schemes that use a common light colour at the mid point, each sequential scheme progresses to a dark hue at the extreme value in each direction [@DCS].
 The use of borders and backgrounds, and their colours, can also change the appearance of the colors representing the value of the statistics [@CB].
@@
 Maps can also be used to focus on demographic strata, such as age and sex.
-Some of the digital atlases surveyed allow subsets to be selected for display, The Australian Cancer Atlas [@ACA] allows filtering for the selection of statistics regarding males and females specifically.
-Similarly, remoteness [@ACA], socioeconomic indicators, such as unemployment rates, poverty rates, and education levels, can be used to filter data, in order to communicate how cancer prevalence varies for different population subgroups.
-Few atlases provide this level of detail.
-XXX
+Some of the digital atlases surveyed allow subsets to be selected for display, for example the Australian Cancer Atlas [@ACA] allows filtering for the selection of statistics regarding males and females specifically.
 
@@
 The most common measures used to present uncertainty are credible intervals or confidence intervals.
-Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it is difficult to display effectively.
+Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it is difficult to display effectively, the Australian Cancer Atlas [@ACA] uses transparency to communicate uncertainty.
 Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two common solutions.

### 2020-05-19 letter2/submission2.Rmd
reorganised repo

--- 2020-05-15 submission2/submission2.Rmd
@@
 **Reviewer suggestion:**
+**Are you referring to area-based characteristics?**
+This socioeconomic indicators mentioned refer to the characteristics of small areas presented in cancer atlases.
+The Australian cancer atlas allows filtering using socioeconomic indicators, but this statement has been removed as it is not common in other atlases.
+
+**Reviewer suggestion:**
 **These are very different terms, so ideally should not have the same abbreviation. **
-The terms credible intervals and confidence intervals are no longer abbreviated.
+The terms credible intervals and confidence intervals are no longer abbreviated to avoid dual use of the abbreviation.
 
@@
 
+**Reviewer suggestion:**
+**Can you clarify whether this recommendation holds when using a ratio measure – which has 1 as the average, and then diverging values higher and lower than the average (eg an SIR).**
+Diverging colour schemes have been suggested for this use by Cynthia Brewer, a reference has been included:
+
+It is possible to allow for data to progress uniformly in both positive and negative directions from a mid point, such as the mean of the data.
+These diverging colour schemes pair two sequential schemes that use a common light colour at the mid point, each sequential scheme progresses to a dark hue at the extreme value in each direction XXX.
+
 All minor comments were addressed in the process of substantial editing.

### 2020-05-19 literature.Rmd
reorganised repo

@@
 |            | Choropleth | Contiguous | Non-contig | Dorling | Tile maps | Geofacets |
-|:-----------|:------------:|:-----------:|:---------------:|:---------:|:-----------:|:----------:|
-|Spatial distortion   |  N  |  Y  |  Y  |  Y  |  Y  |  Y  |
-|Preserves neighbors  |  Y  |  Y  |  Y  |  S  |  S  |  S  |
-|Conceals small areas |  Y  |  S  |  N  |  N  |  N  |  N  |
-|Uniform shape        |  N  |  N  |  N  |  Y  |  Y  |  Y  |
-|Univariate only      |  Y  |  Y  |  Y  |  S  |  S  |  N  |
-|Manual construction  |  N  |  N  |  N  |  N  |  Y  |  Y  |
+|:------------|:----------:|:------------:|:---------:|:-----------:|:----------:|:---------:|
+|Spatial<br> distortion   |  N  |  Y  |  Y  |  Y  |  Y  |  Y  |
+|Preserves<br> neighbors  |  Y  |  Y  |  Y  |  S  |  S  |  S  |
+|Conceals<br> small areas |  Y  |  S  |  N  |  N  |  N  |  N  |
+|Uniform<br> shapes        |  N  |  N  |  N  |  Y  |  Y  |  Y  |
+|Univariate      |  Y  |  Y  |  Y  |  S  |  S  |  N  |
+|Manual<br> construction  |  N  |  N  |  N  |  N  |  Y  |  Y  |
 

### 2020-05-21 letter2/submission2.Rmd
Changes according to reviewers comments, up to table of characteristics

--- 2020-05-15 submission2/submission2.Rmd
@@
 
-All minor comments were addressed in the process of substantial editing.
+**Reviewer suggestion:**
+**Can you clarify whether this recommendation holds when using a ratio measure – which has 1 as the average, and then diverging values higher and lower than the average (eg an SIR).**
+Diverging colour schemes have been suggested for this use by Cynthia Brewer, however a ratio measure requires rescaling of the colour space:
+
+When communicating information using colour, a map creator should use a scheme with perceptually uniform color spaces that match equal steps in data space with equal steps in the colour space [@PUCS], a linear color gradient is appropriate for incidence counts and rates, but may not be appropriate for ratio measures.
+
+**Reviewer suggestion:**
+**Suggest deleting, or rewording to reflect scientific language. **
+The following section was removed:
+*Inset maps like in Brisbane city in Figure 1c of the state of Queensland are commonly used to reduce distorted interpretations, but it is a bandaid remedy.
+For Australia, many, many inset maps would be needed.*
+
+**Reviewer suggestion:**
+**Can this sentence be reworded for clarity?**
+The following section was removed and the concept of cartograms has been explained in the paragraph that introduced cartograms:
+
+*The “population distribution is often extremely uneven”, making a distortion necessary so that population is more faithfully represented as a uniformly distributed background for the statistic to be presented *
+
+**Reviewer suggestion:**
+**Ensure all these refs are included in the list.**
+References have been added to credit the original publications of the alternative displays dicussed by Tobler.
+
+**Reviewer suggestion:**
+**The information in Figure 2 and Table 3 are complementary, and would be better to be combined.**
+The details of the variables and their use in the creation of the cartograms has now been included in the caption of figure 2.
+
+**Reviewer suggestion:**
+**This text could be moved to the main text / methods.**
+The impact of the changes to the maps has now been included in a paragraph structure.
+
+**Reviewer suggestion:**
+**This is a subjective comment – maybe “a very large amount of white space”?**
+Thank you for the suggestion, the previous text has been replaced with your suggestion.
+
+**Reviewer suggestion:**
+
+All minor comments including spell checking, grammar  were addressed in the process of substantial editing.
 

### 2020-05-21 literature.Rmd
Changes according to reviewers comments, up to table of characteristics

--- 2020-05-17 literature.Rmd
@@
 The ColorBrewer system [@CB] and viridis [@viridis] palettes provide effective colour schemes for qualitative, sequential and diverging data.
-When communicating information using colour, a map creator should use a scheme that has a linear color gradient, with perceptually uniform color spaces that match equal steps in data space with equal steps in the colour space [@PUCS].
+When communicating information using colour, a map creator should use a scheme with perceptually uniform color spaces that match equal steps in data space with equal steps in the colour space [@PUCS], a linear color gradient is appropriate for incidence counts and rates, but may not be appropriate for ratio measures.
 It is possible to allow for data to progress uniformly in both positive and negative directions from a mid point, such as the mean of the data.
@@
 -->
-
-Inset maps like in Brisbane city in Figure 1<!--\ref{fig:choropleth-grid-create}-->c of the state of Queensland are commonly used to reduce distorted interpretations, but it is a bandaid remedy.
-For Australia, many, many inset maps would be needed.
 
@@
 
-<!-- Denominators -->
-When a map creator desires a uniform population density of the map base, the purposeful distortion of the map space is beneficial.
-The "population distribution is often extremely uneven", making a distortion necessary so that population is more faithfully represented as a uniformly distributed background for the statistic to be presented [@ACTUC] [@CTTMB] [@GOINO].
-An area cartogram [@NAC], or population-by-area cartogram [@TAAM] is produced from the distortion of the geographical shape according to population.
-Event cartograms [@VSSDCUC] change the area of regions on a map depending on the amount of disease-related events, rather than population.
-
+<!-- Denominators 
+Through the purposeful distortion of the map space, a map can present each region using a size that reflects the amount of population, rather than the amount of land the region encompasses.
+The borders or boundaries can be adjusted and the distortion of map areas can allow for equal representation of the members of population, and provide a uniform population display for the statistic to be presented [@ACTUC] [@CTTMB] [@GOINO].
+-->
 <!-- Why transform?-->
@@
 Monmonier [@HTLWM] suggests that map creators can use white lies to create useful spatial displays.
+An area cartogram [@NAC], or population-by-area cartogram [@TAAM] is produced from the distortion of the geographical shape according to population.
+Event cartograms [@VSSDCUC] change the area of regions on a map depending on the amount of disease-related events, rather than population.
 It is easy for the reader to disregard the impact of transformations used to create cartograms, for the benefit of reading the statistical distribution more accurately with approximate geographic information.
@@
 The creation of cartograms was historically in the hands of professional cartographers [@CD].
-Early approaches by John Hunter and Jonathan Young (1968) and Durham's wooden tile method, Skoda and Robertson's (1972) steel ball-bearing approach and Tobler's (1973) computer programs [@ACTUC].
+Early approaches by John Hunter and Jonathan Young's [@pam] wooden tile methods, Skoda and Robertson's [@imoc] steel ball-bearing approach and Tobler's [@ctufd] computer programs.
 Howe [@HEDP] discusses the impact of electronic computer-assisted techniques.
@@
 Nusrat and Kobourov [@SAIC] provided a framework to investigate implementations of the many algorithms presented, and the "statistical accuracy, geographical accuracy, and topological accuracy".
+The alternative map displays in Figure 2 a - c are created by resizing and reshaping the states of the USA to match the 2015 population of the state.
+This provides a better sense of the extent of disease relative to the population in the country and can help prevent overlooking the impact of the disease within physically small but population-dense states.
+Map creators give each state equal size and thus equal emphasis in Figure 2d the hexagon tile map.
 
-```{r ggcartograms, fig.cap = " Common alternatives to maps, showing the same information for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous).
-Maps (a) - (c) are created by resizing and reshaping the states of the USA to match the 2015 population of the state.
-This provides a better sense of the extent of disease relative to the population in the country and can help ease losing information about physically small but population-dense states.
-Map creators give each state equal size and thus equal emphasis in (d) the hexagon tile map.", message=FALSE, warning=FALSE}
+```{r ggcartograms, fig.cap = "Common alternatives to maps, showing the same information for the United States of America.
+The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016: (a) contiguous cartogram distorted each state's shape by the population of the state in 2015, (b) non-contiguous cartogram preserves the shape of the cartogram, but the size now reflects the population of the state in 2015 (c) Dorling cartogram are non-contiguous, circles are used to represent each state and the population of the state determines the size in 2015 (d) hexagon tilemap (non-contiguous) uses a hexagon of equal size for each state, and colours the state by the average age-adjusted rate of incidence.
+", message=FALSE, warning=FALSE}
 knitr::include_graphics("figures/usa_grid.png")
@@
 
-Table 3: <!--\label{tab:usa}--> Maps used to present statistics for the United States of America.
-The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Figure 2<!--\ref{fig:ggcartograms}--> shows four different cartograms for the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.
+Each cartogram highlights different aspects of the population and relation to the average age-adjusted rate of incidence.
 
-|Map display |Details| 
-|:--------------------|:------------------------------------------------|  
-|a.
-Contiguous| It has distorted each state's shape according to the population of the state in 2015.
-The state of California has become much larger because of the large population density.
-This draws attention to the densely populated North-East region and detracts from the less populated Mid West.|
-|b.
-Non-contiguous| It maintains the geographic shape of the states, but the size has altered according to the population of the state in 2015.
-The state of California has remained closer to its original size than its surrounding states.
+In the contiguous cartogram map the state of California has become much larger because of the large population density.
+This draws attention to the densely populated North-East region and detracts from the less populated Mid West.
+
+In the non-contiguous the state of California has remained closer to its original size than its surrounding states.
 The North-East states have remained closer to their geographical size, for Massachusetts and Connecticut.
-This draws attention to the densely populated North-East region and the sparse Mid West.|
-|c.
-Dorling| Circles are used to represent each state, but the population of the state determines the size in 2015.
-The North-East states remain closer to their neighbors and are slightly displaced from their geographic location.
-It highlights the sparsity of the population in the Mid West by the distance between the circles at the geographic centroids.|
-|d.
-Hexagon Tessellation| A hexagon of equal size represents each state.
-It is easy to contrast the neighboring states however the North-East regions have been displaced from their geographic location.
-It highlights the sparsity of the population in the Mid West by the light yellow color, the Age-Adjusted rate in Kentucky is the darkest and its neighbors are similar.|
+This draws attention to the densely populated North-East region and the sparse Mid West.
 
-Figure 2<!--\ref{fig:ggcartograms}--> shows four different cartograms for the same data. The information in Table 3<!--\ref{tab:usa}--> summarizes what can be observed in the four types of cartograms.
+In Dorling cartograms the North-East states remain closer to their neighbors and are slightly displaced from their geographic location.
+It highlights the sparsity of the population in the Mid West by the distance between the circles at the geographic centroids.
+
+In the Hexagon Tessellation it is simple to contrast the neighboring states however the North-East regions have been displaced from their geographic location.
+It highlights the sparsity of the population in the Mid West by the light yellow color, the Age-Adjusted rate in Kentucky is the darkest and its neighbors are similar.
 
@@
 
-Figure 2a<!--\ref{fig:ggcartograms}--> shows a population contiguous cartogram of the United States.
-All states are visible and the shape of the United States overall is still recognizable.
-In contrast, Figure 3<!--\ref{fig:auscartograms}--> a shows an Australian contiguous cartogram also based on population.
-The south east is enlarged, but high population areas are still small, and low population areas are still large on the map.
-The algorithm doesn't fully reach an optimal configuration where area matches population -- Australia is too heterogeneous for the algorithm to handle.
+Figure 2a<!--\ref{fig:ggcartograms}--> shows a population contiguous cartogram of the United States. All states are visible and the shape of the United States overall is still recognizable.
 
@@
 Olson [@NAC] discusses the creation of these maps and the significance of the empty areas left between the geographic boundaries and the new shape.
-
-The white space presents the meaningful empty-space property [@ECGC] [@NAC] but it also distracts the reader from the data, with a low data density [@TVDQI].
+The white space presents the meaningful empty-space property [@ECGC] [@NAC].
 
@@
 
-He acknowledged the sophistication of contiguous cartograms but critiqued their 'very complex shapes,' he answers this with his implementation of maps created using 'the simplest of all shapes'.
+He acknowledged the sophistication of contiguous cartograms but critiqued their 'very complex shapes,' he answered this with his implementation of maps created using 'the simplest of all shapes'.
 Circular cartograms use the same circle shape for every region represented, resized  according to the statistic represented or the population.
@@
 Contiguous cartograms create 'nonsense' shapes that have 'no meaning' [@NISCC].
-Both methods applies a gravity model to produce a layout, that avoids overlaps and keep spatial relationships with neighboring areas over many iterations.
+Both methods apply a gravity model to produce a layout, that avoids overlaps and keep spatial relationships with neighboring areas over many iterations.
 The circular cartogram is relatively fast to compute.
@@
 The version of a contiguous cartogram (a) has expanded the highly populated areas while preserving the full shapes of rural areas.
-It has not fully resolved the population transformation of areas, and if it had accurately sized areas by population, the country would be unrecognizable.
+The south east is enlarged, but high population areas are still small, and low population areas are still large on the map.
+It has not fully resolved the population transformation of areas, the algorithm doesn't reach an optimal configuration where area matches population -- Australia is too heterogeneous for the algorithm to handle.
 The shape-preserved cartogram is unreadable, and it has reduced all areas to tiny spots on the map.
 Zooming in on a high-resolution output shows it does preserve the shapes.
-The Dorling cartogram and the hexagon tile map provide reasonable displays of the spatial distribution, despite having too much white-space in the outback areas.
+The Dorling cartogram and the hexagon tile map provide reasonable displays of the spatial distribution, despite having a very large amount of white space in the outback areas.
 
@@
 Dorling [@ACTUC] has a cartogram showing the 1966 general election results, which looked very little like the geographical shape of Australia.
+XXX
+
+The most common aesthetics employed in alternative map displays are shape, colour and size.
+Each alternative alternative display allows for some combination of these.
+Colour is used most often to represent the variable of interest.
+The size and shape are often used to scale the areas to draw attention to regions of interest.
 
 Some mix of tiling, faceting or even micromaps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies.
-Table 3 <!--\ref{tab:methods}--> summarizes the key criteria for testing maps and alternative displays.
 Moore and Carpenter [@SAMGIS] and Bell et al.
@@
 
-Table 4: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).
+Table 4: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics.
+Yes means that the map dispay always employs the characteristic.
+No means that the display does not employ the characteristic.
+Possible means it is possible to employ the characteristic when using the display, but it is not necessary or required.
 
 |           | Choropleth | Contiguous | Non-contig | Dorling | Tile maps | Geofacets |
-|:-----------|:------------:|:-----------:|:---------------:|:---------:|:-----------:|:----------:|
-|Spatial distortion   |  N  |  Y  |  Y  |  Y  |  Y  |  Y  |
-|Preserves neighbors  |  Y  |  Y  |  Y  |  S  |  S  |  S  |
-|Conceals small areas |  Y  |  S  |  N  |  N  |  N  |  N  |
-|Uniform shape        |  N  |  N  |  N  |  Y  |  Y  |  Y  |
-|Univariate only      |  Y  |  Y  |  Y  |  S  |  S  |  N  |
-|Manual construction  |  N  |  N  |  N  |  N  |  Y  |  Y  |
-
-<!-- Creating maps of diseases now involves more decisions to be made by map makers, rather than cartographers. 
-Technology has played an enormous role in increasing the opportunities for map makers.
-The computation and graphics power have made creation, alteration and interactivity possible.
-as these options have expanded and it is the objectives of the investigator that will drive the choices.
-@CPISACA and @SAMGIS have provided suggestions and comments to help map creators best communicate their health data and spatial analyses.-->
+|:----------|:---------:|:---------:|:-----------:|:---------:|:---------:|:---------:|
+|Spatial<br> distortion   | No  | Yes      | Yes | Yes      | Yes      | Yes      |
+|Preserves<br> neighbors  | Yes | Yes      | Yes | Possible | Possible | Possible |
+|Conceals<br> small areas | Yes | Possible | No  | No       | No       | No       |
+|Uniform<br> shapes       | No  | No       | No  | Yes      | Yes      | Yes      |
+|Univariate<br> data      | Yes | Yes      | Yes | Possible | Possible | No       |
+|Manual<br> construction  | No  | No       | No  | No       | Yes      | Yes      |
 

### 2020-05-22 letter2/submission2.Rmd
Conclusions and recommendations

@@
 **Reviewer suggestion:**
+**Could you show an example? (of a completed contiguous cartogram that fully resolved the population transformation of areas)**
+Sadly I could not find any packages in R that created a complete contiguous cartogram.
+Allowing the algorithm to run for further iterations saw the error rise, we chose the amount of iterations that led to the minimum error.
+
+**Reviewer suggestion:**
 **This is a subjective comment – maybe “a very large amount of white space”?**
@@
 **Reviewer suggestion:**
+**Can you summarise these here? (suggestions and comments to help map creators best communicate their health data and spatial analysis)**
 
-All minor comments including spell checking, grammar  were addressed in the process of substantial editing.
+An additional paragraph has been included that dicusses the recommendations made by Bell et al. for choosing desirable alternative map diplays.
+
+*Some mix of tiling, faceting or even micromaps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies.
+Bell et al.
+[@CPISACA] provide suggestions and comments to help map creators best communicate their health data and spatial analysis.
+The authors highlighted that the map design chosen should be tested on a representative sample of potential consumers, to ensure that the target audience is not misinformed by the display.
+The authors encourage the consideration of map types beyond the traditional classed choropleth map, but warn that sound cartographic principles must be employed to ensure effective communication to the public.
+A clear definition of the purpose of the display can help map creators to select the design that best communicates the statistic of interest [@CPISACA].
+Table 4 lists several features, or limitations, of each alternative display in contrast to the commonly used choropleth map.
+The desirable features of each display can be contrasted within the table, this can be used to help inform the choice of map creators as they consider each alternative display.
+Map creators should choose a display that best communicates the statistic according to the purpose of their display.*
+
+**Reviewer suggestion:**
+**Not clear what “sometimes” means (in table 3).**
+The use of Yes, No and Possible in the calls of Table 3 has been explained in the new caption:
+
+*Summary of features and constraints of common mapping methods used to display cancer statistics.
+Yes means that the map dispay always employs the characteristic.
+No means that the display does not employ the characteristic.
+Possible means it is possible to employ the characteristic when using the display, but it is not necessary to produce the display.*
+
+**Reviewer suggestion:**
+**Suggest either a separate recommendations section or very clearly laying out the recommendations here.
+Do you recommend always (and only) using alternative visualisations, regardless of the audience, or are you suggesting supplying multiple forms of visualisations?
+Does your recommendation differ depending on the country/data?
+Ensure your final recommendations are clearly reflected in your abstract wording also.
+**
+
+The following paragraph has been included to make clear the recommendations of the authors to the readers.
+
+*Choropleth map displays are familiar and useful for map users to orient themselves to the statistic and additional features of the display.
+A choropleth map display may be the most effective display for the data considering budget, time, and maintenance contraints.
+Map creators should consider the shapes and discrepancies in size within the set of geographic units used for analysis, and whether the features of the dataset warrant transformation of the geographic units.
+Additional mapping methods should be considered by map creators during the development of a cancer atlas, as alternative displays may align better with the purpose of the display.
+Incorporating the population of the geographic units may better inform the map readers of the distribution of the statistic.
+The importance and potential impact of utilising a display that emphasises the underlying population should be considered during the design phase.
+Cartograms and tile maps should be adopted as primary displays going forward, or secondary displays if the familiarity of the choropleth map will benefit the users.
+Clear explanations of the interpretation of the geogephic units in alternative displays should also be included, along with the features of the distribution of the statistic.
+*
+
+**Reviewer suggestion:**
+**Isn’t the most ‘basic’ the number of cancers?**
+We have removed the statement: The most basic is the incidence rate.
+
+**Reviewer suggestion:**
+**Crude or age standardized?**
+We have clarified that the age standardized rates are of interest here.
+
+**Reviewer suggestion:** XXX
+I would suggest replacing this section with a discussion of relative versus absolute measures.
+
+All minor comments including spell checking and grammar fixes were addressed in the process of editing.
 

### 2020-05-22 literature.Rmd
Conclusions and recommendations

@@
 Some mix of tiling, faceting or even micromaps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies.
-Moore and Carpenter [@SAMGIS] and Bell et al.
+Bell et al.
 [@CPISACA] provide suggestions and comments to help map creators best communicate their health data and spatial analysis.
+The authors highlighted that the map design chosen should be tested on a representative sample of potential consumers, to ensure that the target audience is not misinformed by the display.
+The authors encourage the consideration of map types beyond the traditional classed choropleth map, but warn that sound cartographic principles must be employed to ensure effective communication to the public.
+A clear definition of the purpose of the display can help map creators to select the design that best communicates the statistic of interest [@CPISACA].
+Table 3 lists several features, or limitations, of each alternative display in contrast to the commonly used choropleth map.
+The desirable features of each display can be contrasted within the table, this can be used to help inform the choice of map creators as they consider each alternative display.
+Map creators should choose a display that best communicates the statistic according to the purpose of their display.
 
-Table 4: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics.
+Table 3: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics.
 Yes means that the map dispay always employs the characteristic.
 No means that the display does not employ the characteristic.
-Possible means it is possible to employ the characteristic when using the display, but it is not necessary or required.
+Possible means it is possible to employ the characteristic when using the display, but it is not necessary to produce the display.
 
@@
 
-Monmonier [@HTLWM] says that interactivity can be used to allow users to explore the map for more information and provides flexibility for the display.
+Monmonier [@HTLWM] recommends using interactivityto allow users to explore the map for more information and provide flexibility for the display.
 The user can toggle between different variables, map views or even multiple realizations of future scenarios [@goodchild1994introduction].
@@
 The Australian Cancer Atlas [@TACA] provides [tours](
-https://atlas.cancer.org.au/app/tour/lungcancer) that change the display to draw users' attention to areas on the map that are relevant to the story.
+https://atlas.cancer.org.au/app/tour/lungcancer) that change the display to draw users' attention to areas on the map that are relevant to the interpretation of the statistic displayed.
 This implementation of animation gives users tools to plan their exploration.
@@
 <!-- summary of mapping -->
-This paper provides an overview of mapping practices as commonly used for cancer atlases and recommends new approaches, such as cartograms and hexagon tile maps that should be adopted going forward.
+This paper provides an overview of mapping practices as commonly used for cancer atlases.
 The conventional approach is the choropleth map, and it is widely used.
@@
 
+Choropleth map displays are familiar and useful for map users to orient themselves to the statistic and additional features of the display.
+A choropleth map display may be the most effective display for the data considering budget, time, and maintenance contraints.
+Map creators should consider the shapes and discrepancies in size within the set of geographic units used for analysis, and whether the features of the dataset warrant transformation of the geographic units.
+Additional mapping methods should be considered by map creators during the development of a cancer atlas, as alternative displays may align better with the purpose of the display.
+Incorporating the population of the geographic units may better inform the map readers of the distribution of the statistic.
+The importance and potential impact of utilising a display that emphasises the underlying population should be considered during the design phase.
+Cartograms and tile maps should be adopted as primary displays going forward, or secondary displays if the familiarity of the choropleth map will benefit the users.
+Clear explanations of the interpretation of the geogephic units in alternative displays should also be included, along with the features of the distribution of the statistic.
+
 <!-- statistics -->
 Many statistics are commonly used in cancer displays.
-The most basic is the incidence rate.
-It is common to see relative rates which measure how far a region is above or below the average.
-The purpose of using a relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates.
+It is common to see incidence rates, or relative rates which measure how far a region is above or below the average.
+The purpose of using an age standardised relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates.
 <!--However, we lose the incidence rate information and thus interpretability.
@@
 <!-- interaction -->
-Interaction with maps is an important component of public atlases, and is easy to add with today's technology.
-The purpose is to provide access to more information than is possible to display in a single map, without overwhelming the viewer.
+Interaction with maps is an important component of public atlases, and is becoming increasingly straightforward to add with today's technology.
+The purpose of interaction in public atlasses is to provide access to more information than is possible to display in a single map, without overwhelming the viewer.
 Too many choices can similarly overwhelm a viewer, and thus decisions do need to be made about content to provide for accurate and comprehensive communication of information.

### 2020-05-22 literature.Rmd
revisions by di

@@
 
-Epidemiologists and statisticians have developed the statistics used to communicate the burden of cancer over several decades.
+Epidemiologists and statisticians have developed a range of statistics to communicate the burden of cancer, and the choice of statistics used in maps has changed in recent decades.
 Table 1<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases.
-Mortality rates are commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations.
-As described in Howe [@HEDP], Englishman P.
-Stocks advanced the field of mortality statistics by introducing the standardized mortality ratios in the 1930s, which is an improvement on crude death rates.
+Mortality rates are commonly presented as relative rates of risk across the population, and age-adjusted to correct for the higher prevalence of cancers in older populations.
+As described by Howe [@HEDP], Englishman P.
+Stocks advanced the field of mortality statistics by introducing the standardized mortality ratios in the 1930s, which was an improvement on crude death rates.
 
-XXX
-Some of the measures used to present cancer statistics are aggregations.
+The measures displayed are typically aggregations of data values over small areas or model estimates.
+This to protect patient privacy, and for numerical stability.
 The counts of cases in small areas can be difficult to obtain as they are often protected for privacy reasons.
-The information released and the statistics presented in worldwide cancer atlases are often modelled estimates.
-The measures described in table use Incidence as the statistic of interest, this is the amount of new cases and the same statistics can be used to describe the mortality, in this case Standardised Incidence Rate is analogous to the Standarised Mortality Rate.
+The information released and the statistics presented in worldwide cancer atlases are often model estimates.
+The measures described in Table 1 use incidence as the statistic of interest, which reports the number of new cases.
+These same statistics can be used to describe the mortality, that is, Standardised Incidence Rate is analogous to the Standarised Mortality Rate.
 
@@
 The south east is enlarged, but high population areas are still small, and low population areas are still large on the map.
-It has not fully resolved the population transformation of areas, the algorithm doesn't reach an optimal configuration where area matches population -- Australia is too heterogeneous for the algorithm to handle.
+It has not fully resolved the population transformation of areas, because the algorithm can't reach an optimal configuration where area matches population -- Australia is too heterogeneous.
 The shape-preserved cartogram is unreadable, and it has reduced all areas to tiny spots on the map.
@@
 When the population density of the geographic units is highly dissonant with geographic density, the cartogram will lose all spatial context.
-Dorling [@ACTUC] has a cartogram showing the 1966 general election results, which looked very little like the geographical shape of Australia.
-XXX
+Dorling [@ACTUC] contains a cartogram showing the 1966 general election results for Australia, produced Hughes and Savage [@HughesSavage], which looked very little like the geographical shape of Australia.
+The reader is encouraged to access the freely available pdf of Dorling's book, and this image can be found on page 41.
 
@@
 Table 3: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics.
-Yes means that the map dispay always employs the characteristic.
-No means that the display does not employ the characteristic.
-Possible means it is possible to employ the characteristic when using the display, but it is not necessary to produce the display.
+("Yes" indicates the characteristic is always used, "No" indicates never used, and "Possible" it may be employed, but it is not necessary.)
 
@@
 
-Choropleth map displays are familiar and useful for map users to orient themselves to the statistic and additional features of the display.
-A choropleth map display may be the most effective display for the data considering budget, time, and maintenance contraints.
-Map creators should consider the shapes and discrepancies in size within the set of geographic units used for analysis, and whether the features of the dataset warrant transformation of the geographic units.
-Additional mapping methods should be considered by map creators during the development of a cancer atlas, as alternative displays may align better with the purpose of the display.
-Incorporating the population of the geographic units may better inform the map readers of the distribution of the statistic.
-The importance and potential impact of utilising a display that emphasises the underlying population should be considered during the design phase.
-Cartograms and tile maps should be adopted as primary displays going forward, or secondary displays if the familiarity of the choropleth map will benefit the users.
-Clear explanations of the interpretation of the geogephic units in alternative displays should also be included, along with the features of the distribution of the statistic.
+<!-- recommendation -->
+Additional mapping methods should be considered by map creators during the development of a cancer atlas, as alternative displays may align better with the communication purpose.
+Other considerations need to be taken into account, including audience, budget, time, maintenance.
+Of primary importance is that information about cancer statistics be effectively communicated to the public.
+The choropleth has an advantage in that it is more familiar in form for more people, but we have seen that it can give an incorrect perception of the information.
+Public atlases can be useful educational tools.
+A combination of the choropleth alongside an alternative display, could provide a balance of the familiar along with a perceptually accurate display, and provide an opportunity to educate the reader.
+This is especially recommended for Australia, because of the vast difference between spatial area and population density.
 

### 2020-05-22 letter2/submission2.Rmd
letter to reviewers

@@
 toname: Professor Peter Baade and Dr Susanna Cramb
-opening: Dear Professor Peter Baade and Dr Susanna Cramb,
+opening: Dear Professor Peter and Dr Susanna,
 closing: Warm regards
@@
 
-This study summarises current practices for cancer atlases and investigates the alternative map displays that could be used to accurately represent the distribution of cancer statistics across a population, as many cancer atlases lack appropriate displays for population statistics.
+*This study summarises current practices for cancer atlases and investigates the alternative map displays that could be used to accurately represent the distribution of cancer statistics across a population, as many cancer atlases lack appropriate displays for population statistics.*
 
@@
 Additional columns have been included.
-
-XXX
 

### 2020-05-22 literature.Rmd
letter to reviewers

--- 2020-05-22 literature.Rmd
@@
 | Fig. | Atlas | Data | Statistic | Source |
-|------|------------|-----------|----------------------|-----------------------------|
+|------|------------|-----------|----------------------|------------------|
 | a | The Environment and Health<br> Atlas of England and Wales<br> [@EnvEnglandWales2] | Sex: Women<br> Year: 2010<br> Cancer: Lung | Relative risk | [Office for National Statistics<br> (ONS) (England) and from the<br> Welsh Cancer Intelligence and<br> Surveillance Unit (WCISU)](http://www.envhealthatlas.co.uk/homepage/) |

### 2020-05-22 Response/Response.Rmd

@@
 
-We have restructed our original submission to provide a clear path from current approaches used, to the alternative approaches that should be explored.
+We have restructured our original submission to provide a clear path from current approaches used, to the alternative approaches that should be explored.
 We highlight the limitations of using current display methods, to encourage the use of alternative display methods.

### 2020-05-22 letter2/submission2.Rmd
aligned table 2, changed links to be 1a-1g

@@
 
-Several suggestions were made regarding Table 2: Common measures for reporting cancer information.
-Additional columns have been included.
+Several suggestions were made regarding Table 1: Common measures for reporting cancer information.
+this table was previously Table 2 and an additional column has been included to indicate the cancer atlas examples in Figure 1 that display each statistic.
 
@@
 As data becomes more sparse, either by increasing geographic resolution or stratifying by age/sex, obtaining reliable estimates becomes more difficult.
-Many atlases analyzing smaller areas use statistical modelling to produce estimates, although some continue to use simple calculations and suppress regions where estimates are unstable (e.g.
-(mention some)).
-XXX
+Many atlases analyzing smaller areas use statistical modelling to produce estimates, although some continue to use simple calculations and suppress regions where estimates are unstable.
 

### 2020-05-22 literature.Rmd
aligned table 2, changed links to be 1a-1g

@@
 This study summarises current practices for cancer atlases and investigates the alternative map displays that could be used to accurately represent the distribution of cancer statistics across a population, as many cancer atlases lack appropriate displays for population statistics.
-It is recommended that alternative displays are included in cancer atlases.
-In addition, with the ease of today’s technology, user interaction with the displays is encouraged.
+It is recommended that alternative displays are included in cancer atlases for a perceptually accurate display of the distribution of the burden of cancer over the population.
+If it possible given the time and budget constraints, atlases should also present countries or regions containing hetergeneous spatial areas in an alternative display, in addition to the familiar choropleth map.
+With the ease of today’s technology, user interaction with the displays is also encouraged.
 Users should also be able to interactively display different statistics, such as incidence rate or relative incidence, or filter by demographic variables.
@@
 Mortality rates are commonly presented as relative rates of risk across the population, and age-adjusted to correct for the higher prevalence of cancers in older populations.
-As described by Howe [@HEDP], Englishman P.
+As described by Howe [@HEDP], the Englishman P.
 Stocks advanced the field of mortality statistics by introducing the standardized mortality ratios in the 1930s, which was an improvement on crude death rates.
@@
 These same statistics can be used to describe the mortality, that is, Standardised Incidence Rate is analogous to the Standarised Mortality Rate.
+
+Obtaining reliable estimates becomes more difficult as data becomes more sparse, either by increasing geographic resolution or stratifying by age/sex.
+Many atlases analyzing smaller areas such as the Australian Cancer Atlas [@ACA], use statistical modelling to produce estimates, although some continue to use simple calculations and suppress regions where estimates are unstable.
 
@@
 [@EnvEnglandWales2] | a, f |
-
-Table describes the common measures used to present cancer incidence in atlases.
-The calculations used to produce these measures often are also applied to mortality statistics.The Excess Hazard Ratio and Standardized Incidence Ratio are additional measures that can be found in atlases, but are not used in the images displayed in Figure 1.
 
@@
 | Fig. | Atlas | Data | Statistic | Source |
-|------|------------|-----------|----------------------|------------------|
-| a | The Environment and Health<br> Atlas of England and Wales<br> [@EnvEnglandWales2] | Sex: Women<br> Year: 2010<br> Cancer: Lung | Relative risk | [Office for National Statistics<br> (ONS) (England) and from the<br> Welsh Cancer Intelligence and<br> Surveillance Unit (WCISU)](http://www.envhealthatlas.co.uk/homepage/) |
-| b | Globocan 2018: Estimated Cancer<br> Incidence, Mortality and <br> <br> prevalence Worldwide<br> [@Globocan] | Sex: Both<br> Year: 2018<br> Cancer: All invasive | Age standardized<br> incidence rates<br> (per 100,000) | [World Health Organization's<br> International Agency for<br> Research on Cancer](https://www.iarc.fr/<br>  |
-| c | Atlas of Cancer in Queensland<br>  [@QLDcancerAtlas] | Sex: Males<br> Year: 1998 - 2007<br> Cancer: Lung | Relative incidence<br> ratio | [Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/) |
-| d | Australian Cancer Atlas<br> [@ACA] | Sex: Males<br> Year: 2010 - 2014<br> Cancer: Lung | Average standardised<br> incidence rate | [Australian Cancer Atlas](https://atlas.cancer.org.au/) |
-| e | United States Cancer<br> Statistics: An Interactive<br> Cancer Statistics Website<br>  | Sex: Both<br> Year: 2016<br> Cancer: All types | Incidence rate per<br> 100,000 | With data from state cancer<br> registries: <br> [Centers for Disease Control and Prevention](https://www.cdc.gov/) |
-| f | Map of Cancer Mortality<br> Rates in Spain<br> [@cancerSpain] | Sex: Both<br> Year: 2004 - 2008<br> Cancer: All types | Relative risk | <br> [Map of cancer mortality rates<br> in Spain](https://elpais.com/elpais/2014/10/06/media/1412612722_141933.html |
-| g | Atlas of Childhood<br> Cancer in Ontario<br> [@OntarioPediatric] | Sex: Both<br> Year: 1995 - 2004<br> Cancer: All types | Incidence rate per<br> 100,00 | <br> [The Pediatric Oncology Group<br> of Ontario Networked Information<br> System](https://www.pogo.ca/research-data/pogonis-childhood-cancer-database/data-requests/) |
+|----|-----------|-------------|-----------|--------------|
+| a | The Environment and Health<br> Atlas of England and Wales<br> [@EnvEnglandWales2] | Sex: Women<br> Year: 2010<br> Cancer: Lung | Relative risk | Office for National Statistics<br> (ONS) (England) and from the<br> Welsh Cancer Intelligence and<br> Surveillance Unit (WCISU) [1a](http://www.envhealthatlas.co.uk/homepage/) |
+| b | Globocan 2018: Estimated Cancer<br> Incidence, Mortality and <br> <br> prevalence Worldwide<br> [@Globocan] | Sex: Both<br> Year: 2018<br> Cancer: All invasive | Age standardized<br> incidence rates<br> (per 100,000) | World Health Organization's<br> International Agency for<br> Research on Cancer [1b](https://www.iarc.fr/<br>)  |
+| c | Atlas of Cancer in Queensland<br>  [@QLDcancerAtlas] | Sex: Males<br> Year: 1998 - 2007<br> Cancer: Lung | Relative incidence<br> ratio | Queensland Cancer Registry [1c](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/) |
+| d | Australian Cancer Atlas<br> [@ACA] | Sex: Males<br> Year: 2010 - 2014<br> Cancer: Lung | Average standardised<br> incidence rate | Australian Cancer Atlas [1d](https://atlas.cancer.org.au/) |
+| e | United States Cancer<br> Statistics: An Interactive<br> Cancer Statistics Website<br>  | Sex: Both<br> Year: 2016<br> Cancer: All types | Incidence rate per<br> 100,000 | With data from state cancer<br> registries: <br> Centers for Disease Control and Prevention [1e](https://www.cdc.gov/) |
+| f | Map of Cancer Mortality<br> Rates in Spain<br> [@cancerSpain] | Sex: Both<br> Year: 2004 - 2008<br> Cancer: All types | Relative risk | <br> Map of cancer mortality rates<br> in Spain [1f](https://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) |
+| g | Atlas of Childhood<br> Cancer in Ontario<br> [@OntarioPediatric] | Sex: Both<br> Year: 1995 - 2004<br> Cancer: All types | Incidence rate per<br> 100,00 | <br> The Pediatric Oncology Group<br> of Ontario Networked Information<br> System [1g](https://www.pogo.ca/research-data/pogonis-childhood-cancer-database/data-requests/) |
 
@@
 Figure 1<!--\ref{fig:choropleth-grid-create}--> displays a subset of maps from these atlases, the selection varies in the geographies explored.
-Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows Globocan 2018 [@Globocan] which explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide using estimates based on available country cancer registries.
+Figure 1b<!--\ref{fig:choropleth-grid-create}--> shows Globocan 2018 [@Globocan] which explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide using estimates based on available country cancer registries.
 Many atlases allow users to view sex-specific distributions, however sex is displayed in side-by-side maps in the Map of Cancer Mortality Rates in Spain (Figure 1f)<!--\ref{fig:choropleth-grid-create}--> [@cancerSpain].
@@
 The United States Cancer Statistics [@USInteractive] shows data aggregated at the state level (n = 51).
-The Environment and Health Atlas of England and Wales [@EnvEnglandWales2] (Figure 1<!--\ref{fig:choropleth-grid-create}-->a) shows the relative risk for women developing lung cancer  at a neighborhood (small-area) scale (n = 8850).
-The Australian Cancer Atlas (Figure 1<!--\ref{fig:choropleth-grid-create}-->d) shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2016] (n = 2292).
-The Atlas of Cancer in Queensland (Figure 1<!--\ref{fig:choropleth-grid-create}-->c) shows a subset of the Australian Statistical Local Areas (SLAs) located in the state of Queensland within Australia [@QLDcancerAtlas] (n = 478).
+The Environment and Health Atlas of England and Wales [@EnvEnglandWales2] (Figure 1a<!--\ref{fig:choropleth-grid-create}-->) shows the relative risk for women developing lung cancer  at a neighborhood (small-area) scale (n = 8850).
+The Australian Cancer Atlas (Figure 1d<!--\ref{fig:choropleth-grid-create}-->) shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2016] (n = 2292).
+The Atlas of Cancer in Queensland (Figure 1c<!--\ref{fig:choropleth-grid-create}-->) shows a subset of the Australian Statistical Local Areas (SLAs) located in the state of Queensland within Australia [@QLDcancerAtlas] (n = 478).
 
 Age-specific atlases are less common.
-Figure 1<!--\ref{fig:choropleth-grid-create}-->g displays Atlas of Childhood Cancer in Ontario, this communicates the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
+Figure 1g<!--\ref{fig:choropleth-grid-create}--> displays Atlas of Childhood Cancer in Ontario, this communicates the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric].
 
@@
 Choropleth maps can inhibit visual inference when presenting human-related statistics as the display may draw attention away from the ‘potentially more important results in the more populous communities’ that are geographically smaller [@SE].
-Inset maps like in Brisbane city in Figure Figure 1<!--\ref{fig:choropleth-grid-create}c of the state of Queensland are commonly used to reduce distorted interpretations, but it is not always an appropriate solution.
+Inset maps like in Brisbane city in Figure 1c<!--\ref{fig:choropleth-grid-create} of the state of Queensland are commonly used to reduce distorted interpretations, but it is not always an appropriate solution.
 As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
@@
 
-In the contiguous cartogram map the state of California has become much larger because of the large population density.
+In the contiguous cartogram map (2a) the state of California has become much larger because of the large population density.
 This draws attention to the densely populated North-East region and detracts from the less populated Mid West.
 
-In the non-contiguous the state of California has remained closer to its original size than its surrounding states.
+In the non-contiguous (2b) the state of California has remained closer to its original size than its surrounding states.
 The North-East states have remained closer to their geographical size, for Massachusetts and Connecticut.
@@
 
-In Dorling cartograms the North-East states remain closer to their neighbors and are slightly displaced from their geographic location.
+In the Dorling cartogram (2c) the North-East states remain closer to their neighbors and are slightly displaced from their geographic location.
 It highlights the sparsity of the population in the Mid West by the distance between the circles at the geographic centroids.
 
-In the Hexagon Tessellation it is simple to contrast the neighboring states however the North-East regions have been displaced from their geographic location.
+In the Hexagon Tessellation (2d) it is simple to contrast the neighboring states however the North-East regions have been displaced from their geographic location.
 It highlights the sparsity of the population in the Mid West by the light yellow color, the Age-Adjusted rate in Kentucky is the darkest and its neighbors are similar.

### 2020-05-22 letter2/submission2.Rmd
Figure numbering and section numbering, and responses to comments on doc

@@
 
-**Reviewer suggestion:**
-**Add a sentence about the study aims (in the abstract).**
-
-We have included the following sentence in the abstract:
-
-*This study summarises current practices for cancer atlases and investigates the alternative map displays that could be used to accurately represent the distribution of cancer statistics across a population, as many cancer atlases lack appropriate displays for population statistics.*
+The first suggestion pertains to the structure of the article. The rest of the suggestions are listed in the order of the comments made in the document  *ACE-19-31-R1-from-the-reviewers*.
 
@@
 
-The following reviewer suggestions relate to the introduction:
+**Reviewer suggestion:**
+**Add a sentence about the study aims (in the abstract).**
+
+We have included the following sentence in the abstract:
+
+*This study summarises current practices for cancer atlases and investigates the alternative map displays that could be used to accurately represent the distribution of cancer statistics across a population, as many cancer atlases lack appropriate displays for population statistics.*
+
+**Reviewer suggestion:**
+**I would suggest deleting this sentence and the previous one. **
+These sentences have been removed
 
@@
 **Reviewer suggestion:**
+**Could move to the introduction** and
 **Suggest also noting development of statistical methods and software**
-The paragraph that began the *Cancer Atlases* section has been moved to the introduction. This paragraph describes cancer atlases, and the developments oof statistical methods and software.
+The paragraph that began the *Cancer Atlases* section has been moved to the introduction. This paragraph describes cancer atlases, and the developments of statistical methods and software.
 
@@
 **Give the actual reference for Cruikshank 1947**
-
 This quote has been removed.
@@
 **Are you referring to area-based characteristics?**
-This socioeconomic indicators mentioned refer to the characteristics of small areas presented in cancer atlases.
+The socioeconomic indicators mentioned refer to the characteristics of small areas presented in cancer atlases.
 The Australian cancer atlas allows filtering using socioeconomic indicators, but this statement has been removed as it is not common in other atlases.
@@
 **Ensure all these refs are included in the list.**
-References have been added to credit the original publications of the alternative displays dicussed by Tobler.
+References have been added to credit the original publications of the alternative displays discussed by Tobler.
 

### 2020-05-22 literature.Rmd
Figure numbering and section numbering, and responses to comments on doc

@@
 
-# Introduction
+# 1. Introduction
 <!-- 436 words -->
@@
 
-# Measures mapped
+# 2. Measures mapped
 
@@
 
-# Visualisation approaches
+# 3. Visualisation approaches
 
-## Traditional approaches for cancer map displays {#ch:choropleth-maps}
+## 3.1 Traditional approaches for cancer map displays {#ch:choropleth-maps}
 
@@
 
-```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = " A selection of choropleth cancer maps from online atlases that are publicly available.
+```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Figure 1.
+A selection of choropleth cancer maps from online atlases that are publicly available.
 Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends.
@@
 
-## Contemporary alternatives to choropleth maps {#ch:alternatives}
+## 3.2 Contemporary alternatives to choropleth maps {#ch:alternatives}
 
-### Cartograms
+### 3.2.1 Cartograms
 <!-- 1496 words inc table-->
@@
 
-```{r ggcartograms, fig.cap = "Common alternatives to maps, showing the same information for the United States of America.
+```{r ggcartograms, fig.cap = "Figure 2.
+Common alternatives to maps, showing the same information for the United States of America.
 The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016: (a) contiguous cartogram distorted each state's shape by the population of the state in 2015, (b) non-contiguous cartogram preserves the shape of the cartogram, but the size now reflects the population of the state in 2015 (c) Dorling cartogram are non-contiguous, circles are used to represent each state and the population of the state determines the size in 2015 (d) hexagon tilemap (non-contiguous) uses a hexagon of equal size for each state, and colours the state by the average age-adjusted rate of incidence.
@@
 
-### Contiguous
+### 3.2.2 Contiguous
 
@@
 
-### Non-contiguous
+### 3.2.3 Non-contiguous
 
@@
 
-### Dorling
+### 3.2.4 Dorling
 
@@
 
-### Tile Map
+### 3.2.5 Tile Map
 <!-- 173 words -->
@@
 
-### Geofacet
+### 3.2.6 Geofacet
 <!-- 166 words -->
@@
 
-```{r ggtilemap, fig.cap = " Two alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
+```{r ggtilemap, fig.cap = "Figure 3.
+Two alternative displays, tile map (left) and geofaceted map (right), showing state age-adjusted rate of incidence for lung and bronchus in the USA.
 In the tile map, the layout approximates spatial location, with each state being an equal box filled with color representing cancer incidence.
@@
 
-### Multivariate displays
+### 3.2.7 Multivariate displays
 <!-- 222 words -->
@@
 
-# Comparison and critique of alternative displays {#sec:compare}
+# 4. Comparison and critique of alternative displays {#sec:compare}
 <!-- 408 words -->
 
-```{r auscartograms, fig.cap = " Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map.
+```{r auscartograms, fig.cap = "Figure 4.
+Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map.
 The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas.
@@
 
-## Performance of choropleth maps or cartograms for Australia
+## 4.1 Performance of choropleth maps or cartograms for Australia
 
@@
 
-## Limitations of alternative displays
+## 4.2 Limitations of alternative displays
 
@@
 
-## Additional considerations
+## 4.3 Additional considerations
 <!-- 479 words -->
@@
 
-# User interaction {#ch:interacting}
+# 5. User interaction {#ch:interacting}
 <!-- 634 words -->
@@
 
-```{r interacting, fig.cap = " Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for Globocan 2018 [@Globocan], (ii) Menus for variable selection and zooming on Bowel Cancer Australia Atlas, (iii) Menus for choosing variables and countries in The Cancer Atlas, (iv) Tabs for different indicators and cancer types in Global Cancer Map, (v) Menus and toggles for variable and subset selection in United States Cancer Statistics: Data Visualizations.", results = "asis", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
+```{r interacting, fig.cap = "Figure 5.
+Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for Globocan 2018 [@Globocan], (ii) Menus for variable selection and zooming on Bowel Cancer Australia Atlas, (iii) Menus for choosing variables and countries in The Cancer Atlas, (iv) Tabs for different indicators and cancer types in Global Cancer Map, (v) Menus and toggles for variable and subset selection in United States Cancer Statistics: Data Visualizations.", results = "asis", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
 knitr::include_graphics("figures/interacting.png")
@@
 
-```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a.
+```{r animating, fig.cap = "Figure 6.
+Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a.
 Linked maps and time-series line plots, with temporal animation in Map of Cancer Mortality Rates in Spain, b.

### 2020-05-25 literature.Rmd
Final changes for resubmission

@@
 They often present statistics to the public as aggregated values for geopolitical areas.
+Counts and incidence rates are often obtained by state health registries, as data privacy and ethics are key concerns, even when combined to small area statistics.
 Presenting these statistics requires aggregating individual observations for the geographical units, especially for privacy protection, but also for political and policy purposes.
 Examples of typical geographical units include states, provinces, local government areas, and post/zip codes.
-Counts and incidence rates are often obtained by state health registries.
-For this data privacy and ethics are key concerns, even when combined to small area statistics.
 This type of data is routinely collected for public health reasons and may be made available to the general public as a service to the community.
@@
 Alternative approaches are posed because they may be more effective than contemporary techniques.
-The limitations of the visualization methods, highlighting the differences and historic use of these displays is discussed.
+The differences and historic use of these displays is discussed, highlighting the potential and the limitations of the visualization methods.
 
@@
 The most frequently used standard population is the World Standard Population.
-[@Globocan] | b, d |
+[@Globocan] | b |
+|4.
+Age-Standardized Incidence Ratio | The ratio of the observed Age-Standardized Incidence Rate, to the expected number of cases.
+Having adjusted for differences in population and differences in age structure of the population within each area [@DACA] | d |
 |5.
@@
 | c | Atlas of Cancer in Queensland<br>  [@QLDcancerAtlas] | Sex: Males<br> Year: 1998 - 2007<br> Cancer: Lung | Relative incidence<br> ratio | Queensland Cancer Registry [1c](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/) |
-| d | Australian Cancer Atlas<br> [@ACA] | Sex: Males<br> Year: 2010 - 2014<br> Cancer: Lung | Average standardised<br> incidence rate | Australian Cancer Atlas [1d](https://atlas.cancer.org.au/) |
+| d | Australian Cancer Atlas<br> [@ACA] | Sex: Males<br> Year: 2010 - 2014<br> Cancer: Lung | Age-standardised<br> incidence ratio | Australian Cancer Atlas [1d](https://atlas.cancer.org.au/) |
 | e | United States Cancer<br> Statistics: An Interactive<br> Cancer Statistics Website<br>  | Sex: Both<br> Year: 2016<br> Cancer: All types | Incidence rate per<br> 100,000 | With data from state cancer<br> registries: <br> Centers for Disease Control and Prevention [1e](https://www.cdc.gov/) |
@@
 Map users contrast the colours in neighbouring areas to understand the spatial distribution.
-The ColorBrewer system [@CB] and viridis [@viridis] palettes provide effective colour schemes for qualitative, sequential and diverging data.
-When communicating information using colour, a map creator should use a scheme with perceptually uniform color spaces that match equal steps in data space with equal steps in the colour space [@PUCS], a linear color gradient is appropriate for incidence counts and rates, but may not be appropriate for ratio measures.
+Pickle's [@USMA] suggestions for choropleth map displays include directions to categorise the statistic presented on the map according to percentiles.
+A colour scheme that communicates high and low rates can be useful in displays of cancer statistics, using a double ended colour scheme such as those provided by the ColorBrewer system [@CB] and viridis [@viridis] palettes provide effective colour schemes for qualitative, sequential and diverging data.
+When communicating information using colour, a map creator should use a scheme with perceptually uniform color spaces that match equal steps in data space with equal steps in the colour space [@PUCS].
 It is possible to allow for data to progress uniformly in both positive and negative directions from a mid point, such as the mean of the data.
-These diverging colour schemes pair two sequential schemes that use a common light colour at the mid point, each sequential scheme progresses to a dark hue at the extreme value in each direction [@DCS].
+These diverging colour schemes pair two sequential schemes that use a common light colour at the mid point, each sequential scheme progresses to a dark hue at the extreme value in each direction [@DCS], it prefereable that the scheme use darker or warmer colors for higher cancer rates [@USMA].
+A linear color gradient is appropriate for incidence counts and rates.
+The linear gradient can also be transformed using the log scale, to show appropriate colours for ratio measures.
+The Australian Cancer Atlas example in Figure 1d implements dark red for areas with an SIR value with a risk level more than 50% (eg.
+SIR = 1.5) above the Australian average.
+Areas were coloured dark blue if they had an SIR value below the inverse of the risk value specified (eg.
+≈ 0.67).
 The use of borders and backgrounds, and their colours, can also change the appearance of the colors representing the value of the statistics [@CB].
@@
 Many statistics are commonly used in cancer displays.
-It is common to see incidence rates, or relative rates which measure how far a region is above or below the average.
-The purpose of using an age standardised relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates.
+It is common to see incidence rates, or ratios which displays how far a region is above or below the average.
+The purpose of using an age standardised ratio is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates.
+Ratios may negatively impact the interpretability of the actual rates of incidence or mortality.
+However this impact is offset by the value of seeing the relativity of the values in reference to the mean, especially when the mean value is also given on the display.
+This helps to put individual areas in perspective,
 <!--However, we lose the incidence rate information and thus interpretability.
---> A region might be much higher than average, but it may not be close to a health concern, because all regions have a low incidence.
-Supplementary materials can allow map users to recognise when this occurs.
+--> as a region might have a value higher than the average, but it may not be a health concern if all regions have a low incidence rate.
+Supplementary materials such as displaying the mean, or a distribution plot, can allow map users to recognise when this occurs.
 

### 2020-05-26 literature.Rmd
lots of fixes in both word and rmd file and changes to code for fig 2

--- 2020-05-22 literature.Rmd
@@
 This study summarises current practices for cancer atlases and investigates the alternative map displays that could be used to accurately represent the distribution of cancer statistics across a population, as many cancer atlases lack appropriate displays for population statistics.
-It is recommended that alternative displays are included in cancer atlases for a perceptually accurate display of the distribution of the burden of cancer over the population.
-If it possible given the time and budget constraints, atlases should also present countries or regions containing hetergeneous spatial areas in an alternative display, in addition to the familiar choropleth map.
+It is recommended that alternative displays are included in cancer atlases for a perceptually accurate display of the distribution of the burden of cancer over the population, in addition to the familiar choropleth map, if possible given time and budget constraints.
 With the ease of today’s technology, user interaction with the displays is also encouraged.
@@
 Counts and incidence rates are often obtained by state health registries.
-For this data privacy and ethics are key concerns, even when combined to small area statistics.
+For this data, privacy and ethics are key concerns, even when combined to small area statistics.
 This type of data is routinely collected for public health reasons and may be made available to the general public as a service to the community.
@@
 <!-- Displaying of cancer data on choropleth maps -->
-A cancer atlas is a map, or collection of maps, commoly representing cancer incidence or mortality patterns across a country, or group of countries.
+A cancer atlas is a map, or collection of maps, commonly representing cancer incidence or mortality patterns across a country, or group of countries.
 Atlases are key to developing hypotheses regarding areas with unusually high rates, and geographic correlations [@MACM].
@@
 The paper is structured as follows.
-The next section introduces cancer atlases and the worldwide implementations.
-Choropleths are the common approach to disease maps and presents examples of atlases in use today and discusses the limitations of the choropleth map.
-Section 3 describes alternative displays, including the cartogram, which is useful when the map has heterogeneously sized geographic units.
-Section 4 presents the limitations of the production and use of alternative displays.
+The next section describes the common statistics displayed in disease mapping.
+Section 3 focuses on disease map visualizations.
+It describes cancer atlases, presents examples of atlases in use today.
+It discusses the limitations of the most commonly used technique for disease mapping, the choropleth map.
+This section also describes alternative displays, including the cartogram, which are useful when the map has heterogeneously sized geographic units.
+Section 4 presents the limitations in the production and use of alternative displays.
 Disease maps are more useful when made interactive, and common options are described in Section 5, along with a discussion of benefits and disadvantages.
+Section 6 summarizes this survey of the literature, and provides recommendations.
 
@@
 The measures described in Table 1 use incidence as the statistic of interest, which reports the number of new cases.
-These same statistics can be used to describe the mortality, that is, Standardised Incidence Rate is analogous to the Standarised Mortality Rate.
+These same statistics can be used to describe the mortality, that is, standardized Incidence Rate is analogous to the Standardized Mortality Rate.
 
@@
 | c | Atlas of Cancer in Queensland<br>  [@QLDcancerAtlas] | Sex: Males<br> Year: 1998 - 2007<br> Cancer: Lung | Relative incidence<br> ratio | Queensland Cancer Registry [1c](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/) |
-| d | Australian Cancer Atlas<br> [@ACA] | Sex: Males<br> Year: 2010 - 2014<br> Cancer: Lung | Average standardised<br> incidence rate | Australian Cancer Atlas [1d](https://atlas.cancer.org.au/) |
+| d | Australian Cancer Atlas<br> [@ACA] | Sex: Males<br> Year: 2010 - 2014<br> Cancer: Lung | Average standardized<br> incidence rate | Australian Cancer Atlas [1d](https://atlas.cancer.org.au/) |
 | e | United States Cancer<br> Statistics: An Interactive<br> Cancer Statistics Website<br>  | Sex: Both<br> Year: 2016<br> Cancer: All types | Incidence rate per<br> 100,000 | With data from state cancer<br> registries: <br> Centers for Disease Control and Prevention [1e](https://www.cdc.gov/) |
@@
 
-There are two key issues to consider when creating alternative map displays, (1) the intended audience of the map, and (2) its purpose.
+The intended audience and the communication purpose are important to consider when creating alternative map displays.
 Nusrat and Kobourov [@SAIC] provided a framework to investigate implementations of the many algorithms presented, and the "statistical accuracy, geographical accuracy, and topological accuracy".
@@
 
-## 4.1 Performance of choropleth maps or cartograms for Australia
+## 4.1 Performance of cartograms for Australia
 
@@
 
-Monmonier [@HTLWM] recommends using interactivityto allow users to explore the map for more information and provide flexibility for the display.
+Monmonier [@HTLWM] recommends using interactivity to allow users to explore the map for more information and provide flexibility for the display.
 The user can toggle between different variables, map views or even multiple realizations of future scenarios [@goodchild1994introduction].
@@
 The conventional approach is the choropleth map, and it is widely used.
-The choropleth map suffers when there are small geographic units, as occurs in Australia where the population is concentrated on the coast, the information about the burden of cancer on those communities can be hidden.
-Making an inset can clarify congested regions but this breaks the viewers' attention as they shift focus from the map to the inset, and if there are many congested areas, many insets would be needed.
-The map alternatives implement trade-offs between the familiar shapes, and the importance of the geographic areas in the context of the areas.
+The choropleth map suffers when there are small geographic units, as occurs in Australia.
+The population of Australia is concentrated in small areas on the coast, and a choropleth map can hide information about the burden of cancer on those communities.
+Making an inset can clarify congested regions but this breaks the viewers’ attention as they shift focus from the map to the inset, and if there are many congested areas, many insets would be needed.
+The map alternatives implement trade-offs between the familiar shapes, and the importance of the geographic areas.
 Given the population or a cancer statistic for each area, the geographic size or shape will change.
@@
 It is common to see incidence rates, or relative rates which measure how far a region is above or below the average.
-The purpose of using an age standardised relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates.
+The purpose of using an age standardized relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates.
 <!--However, we lose the incidence rate information and thus interpretability.
@@
 Interaction with maps is an important component of public atlases, and is becoming increasingly straightforward to add with today's technology.
-The purpose of interaction in public atlasses is to provide access to more information than is possible to display in a single map, without overwhelming the viewer.
+The purpose of interaction in public atlases is to provide access to more information than is possible to display in a single map, without overwhelming the viewer.
 Too many choices can similarly overwhelm a viewer, and thus decisions do need to be made about content to provide for accurate and comprehensive communication of information.

### 2020-05-26 cancer_map_images/usa_plots.R

@@ -69,7 +69,24 @@ ggchoro <- cancer %>%
 ggchoro
 ggsave(filename = "figures/ggchoro.png", device = "png", dpi = 300, width = 7, height = 6, bg = "transparent")
 
-usa_legend <- get_legend(ggchoro)
+invthm_black <- theme_void() + theme(
+  panel.background = element_rect(fill = "transparent", colour = NA), 
+  plot.background = element_rect(fill = "transparent", colour = NA),
+  legend.background = element_rect(fill = "transparent", colour = NA),
+  legend.key = element_rect(fill = "transparent", colour = NA),
+  text = element_text(colour = "black", size = 20)
+)
+ggchoro2 <- cancer %>% 
+  mutate(`Age adjusted rate\n(per 100,000)` = AgeAdjustedRate) %>% 
+  ggplot() + 
+  geom_sf(aes(fill = `Age adjusted rate\n(per 100,000)`), colour = NA) +
+  scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
+  ggtitle(title) +
+  #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
+  invthm +
+  theme(legend.position ="bottom",
+        legend.text = element_text(size = 8))
+usa_legend <- get_legend(ggchoro2)
 save(usa_legend, file = "figures/usa_legend.rda")
 
 ###############################################################################
@@ -160,7 +177,7 @@ ggsave(filename = "figures/ggdorl.png", device = "png", dpi = 300, width = 7, he
 ggchoro1 <- ggplot(st_transform(cancer, 3857)) + 
   geom_sf(aes(fill = AgeAdjustedRate), colour = NA) +
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
-  theme_void()+ 
+  theme_void()
 ggsave(filename = "figures/gg3857.png", plot = ggchoro1,
        device = "png", dpi = 300, width = 7, height = 6, bg = "transparent")
 
@@ -168,7 +185,7 @@ ggsave(filename = "figures/gg3857.png", plot = ggchoro1,
 ggchoro2 <- ggplot(st_transform(cancer, 2163)) + 
[…]

### 2020-05-26 letter2/submission2.Rmd

@@
 **Can you clarify whether this recommendation holds when using a ratio measure – which has 1 as the average, and then diverging values higher and lower than the average (eg an SIR).**
-Diverging colour schemes have been suggested for this use by Cynthia Brewer, however a ratio measure requires rescaling of the colour space:
-
-When communicating information using colour, a map creator should use a scheme with perceptually uniform color spaces that match equal steps in data space with equal steps in the colour space [@PUCS], a linear color gradient is appropriate for incidence counts and rates, but may not be appropriate for ratio measures.
+A linear colour gradient can be transformed to a log scale for ratio measures.
+This has been used in the Australian Cancer Atlas, a small example has been included in the paragraph to explain the transformation.
 
@@
 
-**Reviewer suggestion:** XXX
-I would suggest replacing this section with a discussion of relative versus absolute measures.
+**Reviewer suggestion:**
+**I would suggest replacing this section with a discussion of relative versus absolute measures.**
+We have included additional comments discussing of relative measures.
+
+Please pass our warm thanks to the reviewers for their thorough and thoughtful recommendations. Their input has greatly improved this piece of writing.
 
@@
 
-We thank you for your consideration of our changes.
+We thank you for your patience and your consideration of our changes.

### 2020-05-26 literature.Rmd

@@
 <!-- What data are we working with, why, what are the purposes -->
+
 Researchers, health authorities, governments, not-for-profits and the media are common communicators of cancer statistics.
 They often present statistics to the public as aggregated values for geopolitical areas.
-Presenting these statistics requires aggregating individual observations for the geographical units, especially for privacy protection, but also for political and policy purposes.
+Counts and incidence rates are often obtained by state health registries.
+Data privacy and ethics are key concerns even when combined to small area statistics.
+Presenting these statistics requires aggregating individual observations for the geographical units, also for political and policy purposes.
 Examples of typical geographical units include states, provinces, local government areas, and post/zip codes.
-Counts and incidence rates are often obtained by state health registries.
-For this data, privacy and ethics are key concerns, even when combined to small area statistics.
 This type of data is routinely collected for public health reasons and may be made available to the general public as a service to the community.
@@
 
-<!-- statistics -->
 Many statistics are commonly used in cancer displays.
-It is common to see incidence rates, or relative rates which measure how far a region is above or below the average.
-The purpose of using an age standardized relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates.
-<!--However, we lose the incidence rate information and thus interpretability.
---> A region might be much higher than average, but it may not be close to a health concern, because all regions have a low incidence.
-Supplementary materials can allow map users to recognise when this occurs.
+It is common to see incidence rates, or ratios which displays how far a region is above or below the average.
+The purpose of using an age standardized ratio is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates.
+Ratios may negatively impact the interpretability of the actual rates of incidence or mortality.
+However, this impact is offset by the value of seeing the relativity of the values in reference to the mean, especially when the mean value is also given on the display.
+This helps to put individual areas in perspective,
+as a region might have a value higher than the average, but it may not be a health concern if all regions have a low incidence rate.
+Supplementary materials such as displaying the mean, or a distribution plot, can allow map users to recognise when this occurs.
 

### 2020-05-26 cancer_map_images/usa_plots.R
fig 2 revised

@@ -83,7 +83,7 @@ ggchoro2 <- cancer %>%
   scale_fill_distiller(type = "seq", palette = "RdPu",  direction = 1) + 
   ggtitle(title) +
   #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
-  invthm +
+  invthm_black +
   theme(legend.position ="bottom",
         legend.text = element_text(size = 8))
 usa_legend <- get_legend(ggchoro2)
@@ -277,8 +277,6 @@ usa_gridl <- ggdraw() +
 usa_gridl
 
 ggsave(filename = "figures/usa_grid.png", plot = usa_gridl,
-  device = "png", dpi = 300, width = 7, height = 5, bg = "transparent")
-ggsave(filename = "figures/usa_grid2.png", plot = usa_gridl,
        device = "png", dpi = 300, width = 11, height = 6, bg = "transparent")
 
 ###############################################################################

### 2020-05-26 Response_to_Review2/author_response_to_review.Rmd
response directory name change

---
from: Stephanie Kobakian
fromqualifications: B.Com and B.Eco
fromemail: stephanie.kobakian@monash.edu
fromphone: +61 433 699 797
fromsignature: sigfile.png
toname: "Editors, Focused Issue on Spatial Patterns in Cancer Epidemiology"
opening: Dear Professor Baade and Dr Cramb,
closing: Warm regards
spacing: 1.4
output: MonashEBSTemplates::letter
---

Thank you for the reviews of our revised paper. We have completed our revision, and a point-by-point summary of the changes made are is provided here. We hope that the revision can be included in the Focused Issue on Spatial Patterns in Cancer Epidemiology in the Annals of Cancer Epidemiology.

The **title change**, has been changed, as requested, to "Mapping Cancer: the Potential of Cartograms and Alternative Map Displays".

The **structure of the article has been changed** as requested, and the remaining changes are in order of the comments made in the review  (provided in the document *ACE-19-31-R1-from-the-reviewers*).

**Reviewer suggestion:**
**Could we structure paper as:**

- **Introduction (also introduces cancer atlases)**
- **Measures mapped (info around Table 2)**
- **Visualisation approaches**
    - **Choropleth (mention advant and disadvant, along with figures from cancer atlases)**
    - **Cartograms**
    - **Tile maps**
    - **Geofacets**
    - **Multivariate maps**
- **Comparison & critique of alternative methods**
- **Additional considerations**
    - **Associated graphs**
    - **Displaying uncertainty measures**
    - **Interactivity**
- **Recommendations for mapping cancer**
- **Conclusion**

We have rearranged the sections from the "Introduction" to "Traditional approaches for cancer map displays" in this submission at the request of the reviewer. The discussion of choropleth maps considers the advantages and disadvantages now occurs in the same section.

**Reviewer suggestion:**
**Add a sentence about the study aims (in the abstract).**

We have included the following sentence in the abstract:

*This study summarises current practices for cancer atlases and investigates the alternative map displays that could be used to accurately represent the distribution of cancer statistics across a population, as many cancer atlases lack appropriate displays for population statistics.*

**Reviewer suggestion:**
**I would suggest deleting this sentence and the previous one. **
These sentences have been removed

**Reviewer suggestion:**
**Given the ethical and privacy considerations for small area data, it is not a straightforward process to obtain and report area-specific counts.**

We agree that it is not easy to obtain and report area-specific counts.
The statement has been removed.

**Reviewer suggestion:**
**Suggest moving paragraph to proposed choropleth maps section**

This paragraph has been moved the choropleth section under visualisation approaches.

**Reviewer suggestion:**
**Can you include mention of how this is different to a thematic map?**

We have included in the choropleth section:
Choropleth maps are a type of thematic map that show polygons for each of the groups of data points representing the geographic units, where each polygon is shaded with a color according to the area-specific values of the statistic being conveyed.

**Reviewer suggestion:**
**Could move to the introduction** and
**Suggest also noting development of statistical methods and software**

The paragraph that began the *Cancer Atlases* section has been moved to the introduction. This paragraph describes cancer atlases, and the developments of statistical methods and software.

**Reviewer suggestion:**
**Give the actual reference for Cruikshank 1947**

This quote has been removed.

**Reviewer suggestion:**
**Suggest including Australian Cancer Atlas**

We have replaced the Bowel Cancer Australia Atlas with the Australian Cancer Atlas.

**Reviewer suggestion:**
**This table might benefit from an additional column, which gives an example of a cancer map using this measure. Also, some of these measures can be calculated.**

Several suggestions were made regarding Table 1: Common measures for reporting cancer information. this table was previously Table 2 and an additional column has been included to indicate the cancer atlas examples in Figure 1 that display each statistic.

**Reviewer suggestion:**
**This is the wrong reference.**

The incorrect reference to the use of Statistical Areas at Level 2 for the Atlas of Cancer in Queensland has been corrected to reflect that the atlas used Statistical Local Areas.

**Reviewer suggestion:**

The following paragraph was included at the suggestion of the reviewer:

As data becomes more sparse, either by increasing geographic resolution or stratifying by age/sex, obtaining reliable estimates becomes more difficult. Many atlases analyzing smaller areas use statistical modelling to produce estimates, although some continue to use simple calculations and suppress regions where estimates are unstable.

**Reviewer suggestion:**
**Are you referring to area-based characteristics?**

The socioeconomic indicators mentioned refer to the characteristics of small areas presented in cancer atlases.
The Australian cancer atlas allows filtering using socio-economic indicators, but this statement has been removed as it is not common in other atlases.

**Reviewer suggestion:**
**These are very different terms, so ideally should not have the same abbreviation. **

The terms credible intervals and confidence intervals are no longer abbreviated to avoid dual use of the abbreviation.

**Reviewer suggestion:**
**Which map?**

The following sentence has been removed: The map focuses on displaying the statistic and lacks additional space to represent the uncertainty. 
As the Australian Cancer Atlas example presented in Figure 1 does include uncertainty using transparency.

**Reviewer suggestion:**
**Can you clarify whether this recommendation holds when using a ratio measure – which has 1 as the average, and then diverging values higher and lower than the average (eg an SIR).**

A linear colour gradient can be transformed to a log scale for ratio measures. This has been used in the Australian Cancer Atlas, a small example has been included in the paragraph to explain the transformation.

**Reviewer suggestion:**
**Suggest deleting, or rewording to reflect scientific language. **

The following section was removed:

*Inset maps like in Brisbane city in Figure 1c of the state of Queensland are commonly used to reduce distorted interpretations, but it is a bandaid remedy. For Australia, many, many inset maps would be needed.*

**Reviewer suggestion:**
**Can this sentence be reworded for clarity?**

The following section was removed and the concept of cartograms has been explained in the paragraph that introduced cartograms:

*The “population distribution is often extremely uneven”, making a distortion necessary so that population is more faithfully represented as a uniformly distributed background for the statistic to be presented *

**Reviewer suggestion:**
**Ensure all these refs are included in the list.**

References have been added to credit the original publications of the alternative displays discussed by Tobler.

**Reviewer suggestion:**
**The information in Figure 2 and Table 3 are complementary, and would be better to be combined.**

The details of the variables and their use in the creation of the cartograms has now been included in the caption of figure 2.

**Reviewer suggestion:**
**This text could be moved to the main text / methods.**

The impact of the changes to the maps has now been included in a paragraph structure.

**Reviewer suggestion:**
**Could you show an example? (of a completed contiguous cartogram that fully resolved the population transformation of areas)**

Sadly I could not find any packages in R that created a complete contiguous cartogram. Allowing the algorithm to run for further iterations saw the error rise, we chose the amount of iterations that led to the minimum error.

**Reviewer suggestion:**
**This is a subjective comment – maybe “a very large amount of white space”?**

Thank you for the suggestion, the previous text has been replaced with your suggestion.

**Reviewer suggestion:**
**Can you summarise these here? (suggestions and comments to help map creators best communicate their health data and spatial analysis)**

An additional paragraph has been included that discusses the recommendations made by Bell et al. for choosing desirable alternative map displays.

*Some mix of tiling, faceting or even micromaps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies. Bell et al. [@CPISACA] provide suggestions and comments to help map creators best communicate their health data and spatial analysis. The authors highlighted that the map design chosen should be tested on a representative sample of potential consumers, to ensure that the target audience is not misinformed by the display. The authors encourage the consideration of map types beyond the traditional classed choropleth map, but warn that sound cartographic principles must be employed to ensure effective communication to the public. A clear definition of the purpose of the display can help map creators to select the design that best communicates the statistic of interest [@CPISACA]. Table 4 lists several features, or limitations, of each alternative display in contrast to the commonly used choropleth map. The desirable features of each display can be contrasted within the table, this can be used to help inform the choice of map creators as they consider each alternative display. Map creators should choose a display that best communicates the statistic according to the purpose of their display.*

**Reviewer suggestion:**
**Not clear what “sometimes” means (in table 3).**

The use of Yes, No and Possible in the calls of Table 3 has been explained in the new caption:

*Summary of features and constraints of common mapping methods used to display cancer statistics. Yes means that the map display always employs the characteristic. No means that the display does not employ the characteristic. Possible means it is possible to employ the characteristic when using the display, but it is not necessary to produce the display.*

**Reviewer suggestion:**
**Suggest either a separate recommendations section or very clearly laying out the recommendations here. Do you recommend always (and only) using alternative visualisations, regardless of the audience, or are you suggesting supplying multiple forms of visualisations? Does your recommendation differ depending on the country/data? Ensure your final recommendations are clearly reflected in your abstract wording also. **

The following paragraph has been included to make clear the recommendations of the authors to the readers.

*Choropleth map displays are familiar and useful for map users to orient themselves to the statistic and additional features of the display. A choropleth map display may be the most effective display for the data considering budget, time, and maintenance constraints. Map creators should consider the shapes and discrepancies in size within the set of geographic units used for analysis, and whether the features of the dataset warrant transformation of the geographic units. Additional mapping methods should be considered by map creators during the development of a cancer atlas, as alternative displays may align better with the purpose of the display. Incorporating the population of the geographic units may better inform the map readers of the distribution of the statistic. The importance and potential impact of utilising a display that emphasises the underlying population should be considered during the design phase. Cartograms and tile maps should be adopted as primary displays going forward, or secondary displays if the familiarity of the choropleth map will benefit the users. Clear explanations of the interpretation of the geographic units in alternative displays should also be included, along with the features of the distribution of the statistic. *

**Reviewer suggestion:**
**Isn’t the most ‘basic’ the number of cancers?**

We have removed the statement: The most basic is the incidence rate.

**Reviewer suggestion:**
**Crude or age standardized?**

We have clarified that the age standardized rates are of interest here.

**Reviewer suggestion:**
**I would suggest replacing this section with a discussion of relative versus absolute measures.**

We have included additional comments discussing relative measures. All minor comments including spell checking and grammar fixes were addressed in the process of editing.

Please pass our warmest thanks to the reviewers for their thorough and thoughtful recommendations. Their input has greatly improved this paper.
We thank you for your patience and your consideration of our changes.

### 2020-05-26 literature.Rmd
renamed fig files for submisson, and added zip file

@@
 ```{r setup, include=FALSE}
-fig_path <- glue::glue("figs/")
 knitr::opts_chunk$set(echo = FALSE, 
@@
                       message = FALSE,
+  fig_path = "figs/",
   out.width = "\\textwidth")
@@
 When the population density of the geographic units is highly dissonant with geographic density, the cartogram will lose all spatial context.
-Dorling [@ACTUC] contains a cartogram showing the 1966 general election results for Australia, produced Hughes and Savage [@HughesSavage], which looked very little like the geographical shape of Australia.
+Dorling [@ACTUC] contains a cartogram showing the 1966 general election results for Australia, produced Hughes and Savage [@HS], which looked very little like the geographical shape of Australia.
 The reader is encouraged to access the freely available pdf of Dorling's book, and this image can be found on page 41.

### 2020-08-26 literature.Rmd
Final changes requested

@@
 author: Stephanie Kobakian^[Science and Engineering Faculty, Queensland University
-  of Technology.], Dianne Cook^[Faculty of Econometrics and Business Statistics, Monash
-  University.] and Jessie Roberts^[Queensland University of Technology.]
+  of Technology, Queensland, Australia.], Dianne Cook^[Faculty of Econometrics and Business Statistics, Monash University, Victoria, Australia.] and Jessie Roberts^[Science and Engineering Faculty, Queensland University of Technology, Queensland, Australia.]
 csl: springer-vancouver.csl
@@
 ```
+
+(I) Conception and design: All authors; 
+(II) Administrative support: None; 
+(III) Provision of study materials or patients: J Roberts, S Kobakian; 
+(IV) Collection and assembly of data: None; 
+(V) Data analysis and interpretation: All authors; 
+(VI) Manuscript writing: All authors; 
+(VII) Final approval of manuscript: All authors.
+
+Correspondence to: Stephanie Kobakian, MPhil. Monash University, Melbourne VIC 3800, AUS. Email: stephanie.kobakian@monash.edu.
+
+\section*{Keywords}
+Cancer; cancer atlas; choropleth; cartogram; spatial;
 
@@
 | a | The Environment and Health<br> Atlas of England and Wales<br> [@EnvEnglandWales2] | Sex: Women<br> Year: 2010<br> Cancer: Lung | Relative risk | Office for National Statistics<br> (ONS) (England) and from the<br> Welsh Cancer Intelligence and<br> Surveillance Unit (WCISU) [1a](http://www.envhealthatlas.co.uk/homepage/) |
-| b | Globocan 2018: Estimated Cancer<br> Incidence, Mortality and <br> <br> prevalence Worldwide<br> [@Globocan] | Sex: Both<br> Year: 2018<br> Cancer: All invasive | Age standardized<br> incidence rates<br> (per 100,000) | World Health Organization's<br> International Agency for<br> Research on Cancer [1b](https://www.iarc.fr/<br>)  |
+| b | Globocan 2018: Estimated Cancer<br> Incidence, Mortality and <br> <br> Prevalence Worldwide<br> [@Globocan] | Sex: Both<br> Year: 2018<br> Cancer: All invasive | Age standardized<br> incidence rates<br> (per 100,000) | World Health Organization's<br> International Agency for<br> Research on Cancer [1b](https://gco.iarc.fr/today<br>)  |
 | c | Atlas of Cancer in Queensland<br>  [@QLDcancerAtlas] | Sex: Males<br> Year: 1998 - 2007<br> Cancer: Lung | Relative incidence<br> ratio | Queensland Cancer Registry [1c](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/) |
-| d | Australian Cancer Atlas<br> [@ACA] | Sex: Males<br> Year: 2010 - 2014<br> Cancer: Lung | Average standardized<br> incidence rate | Australian Cancer Atlas [1d](https://atlas.cancer.org.au/) |
-| e | United States Cancer<br> Statistics: An Interactive<br> Cancer Statistics Website<br>  | Sex: Both<br> Year: 2016<br> Cancer: All types | Incidence rate per<br> 100,000 | With data from state cancer<br> registries: <br> Centers for Disease Control and Prevention [1e](https://www.cdc.gov/) |
-| f | Map of Cancer Mortality<br> Rates in Spain<br> [@cancerSpain] | Sex: Both<br> Year: 2004 - 2008<br> Cancer: All types | Relative risk | <br> Map of cancer mortality rates<br> in Spain [1f](https://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) |
+| d | Map of Cancer Mortality<br> Rates in Spain<br> [@cancerSpain] | Sex: Males<br> Year: 1975 - 2017<br> Cancer: Lung | Age Standardized Rate | <br> Map of cancer mortality rates<br> in Spain [1f](http://ariadna.cne.isciii.es/MapP/) |
+| e | United States Cancer<br> Statistics: An Interactive<br> Cancer Statistics Website<br> [@USInteractive] | Sex: Both<br> Year: 2016<br> Cancer: All types | Incidence rate per<br> 100,000 | With data from state cancer<br> registries: <br> Centers for Disease Control and Prevention [1e](https://gis.cdc.gov/Cancer/USCS/DataViz.html) |
+| f | Australian Cancer Atlas<br> [@ACA] | Sex: Males<br> Year: 2010 - 2014<br> Cancer: Lung | Average standardized<br> incidence rate | Australian Cancer Atlas [1d](https://atlas.cancer.org.au/) |
 | g | Atlas of Childhood<br> Cancer in Ontario<br> [@OntarioPediatric] | Sex: Both<br> Year: 1995 - 2004<br> Cancer: All types | Incidence rate per<br> 100,00 | <br> The Pediatric Oncology Group<br> of Ontario Networked Information<br> System [1g](https://www.pogo.ca/research-data/pogonis-childhood-cancer-database/data-requests/) |
@@
 Figure 1b<!--\ref{fig:choropleth-grid-create}--> shows Globocan 2018 [@Globocan] which explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide using estimates based on available country cancer registries.
-Many atlases allow users to view sex-specific distributions, however sex is displayed in side-by-side maps in the Map of Cancer Mortality Rates in Spain (Figure 1f)<!--\ref{fig:choropleth-grid-create}--> [@cancerSpain].
+Most atlases allow users to view sex-specific distributions.
 
@@
 The Environment and Health Atlas of England and Wales [@EnvEnglandWales2] (Figure 1a<!--\ref{fig:choropleth-grid-create}-->) shows the relative risk for women developing lung cancer  at a neighborhood (small-area) scale (n = 8850).
-The Australian Cancer Atlas (Figure 1d<!--\ref{fig:choropleth-grid-create}-->) shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2016] (n = 2292).
+The Australian Cancer Atlas (Figure 1f<!--\ref{fig:choropleth-grid-create}-->) shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2011] (n = 2292).
 The Atlas of Cancer in Queensland (Figure 1c<!--\ref{fig:choropleth-grid-create}-->) shows a subset of the Australian Statistical Local Areas (SLAs) located in the state of Queensland within Australia [@QLDcancerAtlas] (n = 478).
@@
 The resulting display can communicate the impact of the disease more accurately across the population, as recorded by the statistic, at the sacrifice of geographic accuracy.
-
-<!--
-```{r choroCRS, fig.cap = "Four choropleth maps average age-adjusted rate of incidence for lung and bronchus in the United States averaged from sex and years 2012-2016, using four different map projections.
-The map projections alter the shapes and angles of the boundaries of each state.", message=FALSE, warning=FALSE, fig.align='center'}
-#knitr::include_graphics("figures/ggchoroCRS.png")
-c1 <- png::readPNG("figures/gg3857.png")
-c2 <- png::readPNG("figures/gg2163.png")
-c3 <- png::readPNG("figures/gg4326.png")
-c4 <- png::readPNG("figures/gg2955.png")
-#knitr::include_graphics("figures/usa_grid.png")
-load("figures/usa_legend.rda")
-
-ggdraw() +
-  draw_plot(rasterGrob(c1), 0, .55, 0.5,    0.45) +
-  draw_plot(rasterGrob(c2), 0.5, 0.55, 0.5, 0.45) +
-  draw_plot(rasterGrob(c3), 0.0, 0.1, 0.5, 0.45) +
-  draw_plot(rasterGrob(c4), 0.5, 0.1, 0.5,   0.45) +
-  draw_plot(usa_legend, 0, 0, 1, 0.1, scale = 2) +
-  draw_plot_label(c("a", "b", "c", "d"), 
-    c(0, .5, 0, 0.5), 
-    c(1, 1, .5, .5), size = 20)
-```
--->
 
@@
 The user can toggle between different variables, map views or even multiple realizations of future scenarios [@goodchild1994introduction].
-This provides additional mechanisms for the users to digest the uncertainty of the available information [@maceachren1992visualizing, @van1994visualization].
+This provides additional mechanisms for the users to digest the uncertainty of the available information [@maceachren1992visualizing], [@van1994visualization].
 When the needs of the audience are changeable and are also the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions.
@@
 
+# Footnote
+
+*Conflicts of Interest:* All authors have completed the ICMJE uniform disclosure form.
+The series “Spatial Patterns in Cancer Epidemiology” was commissioned by the editorial office without any funding or sponsorship.
+The authors have no other conflicts of interest to declare.
+
+*Ethical Statement:* The authors are accountable for all aspects of the work in ensuring that questions related to the accuracy or integrity of any part of the work are appropriately investigated and resolved.
+
 # References

### 2020-08-26 cancer_map_images/review_figures.R

@@ -7,23 +7,22 @@ library(cowplot)
 plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
 plot2 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
 plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
-plot4 <- png::readPNG('cancer_map_images/choropleths/aca.choropleth.png')
+plot4 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
 plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
-plot6 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
+plot6 <- png::readPNG('cancer_map_images/choropleths/aca.choropleth.png')
 plot7 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
 
 ggdraw(xlim = c(0,1), ylim = c(0,1)) +
-  draw_plot(rasterGrob(plot1), 0, .75, 0.33, .26) +
-  draw_plot(rasterGrob(plot2), 0.33, 0.75, 0.33, .25) +
-  draw_plot(rasterGrob(plot3), 0.66, 0.5, .33, .5) +
-  draw_plot(rasterGrob(plot4), 0, 0.5, .33, .25) +
-  draw_plot(rasterGrob(plot5), 0.33, 0.5, .33, .25) +
-  draw_plot(rasterGrob(plot6), 0, 0, .66, .5) +
-  draw_plot(rasterGrob(plot7), .66, 0, .33, .5) +
-  draw_plot(rasterGrob(plot8), .66, 0, .33, .5) +
+  draw_plot(rasterGrob(plot1), 0.02, .73, 0.31, .25) +
+  draw_plot(rasterGrob(plot2), 0.33, 0.73, 0.33, .25) +
+  draw_plot(rasterGrob(plot3), 0.66, 0.47, .33, .5) +
+  draw_plot(rasterGrob(plot4), 0, 0.465, .35, .26) +
+  draw_plot(rasterGrob(plot5), 0.33, 0.465, .33, .25) +
+  draw_plot(rasterGrob(plot6), 0.005, 0, .66, .5) +
+  draw_plot(rasterGrob(plot7), 0.68, 0, 0.26, .48) +
[…]

### 2020-08-26 qldsa2.R

@@ -76,7 +76,8 @@ nac_colours <- c("A" = "#33809d",
                  "B" = "#aec6c7",
                  "C" = "#fff4bc",
                  "D" = "#ff9a64",
-                 "E" = "#ff3500")
+                 "E" = "#ff3500",
+                 "NA" = "#D3D3D3")
 
 colours <- c(   "Much lower\nthan average" = "#33809d",
                 "Lower than\naverage" = "#aec6c7",
@@ -143,13 +144,26 @@ qld_ggchoro <- ggplot(sa2_mel_fort) +
   geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon),
                    fill = p50_col), colour = "lightgrey") +
   scale_fill_manual(values = nac_colours) + 
-  coord_sf(crs = CRS("+init=epsg:3112"), 
+  coord_sf(crs = 3112, 
            xlim = c(b["xmin"], b["xmax"]), 
            ylim = c(b["ymin"], b["ymax"])) +
-  theme(legend.position ="bottom") +
+  guides(fill = FALSE) +
   labs(fill = "Melanoma Cancer (Males):\nAge-standardised rate\n(per 100,000)") +
   invthm
 
+bris_box <- c(xmin = 1808445, ymin = -3297649,        
+              xmax = 1877776, ymax = -3190617)
+
+bris_ggchoro <- ggplot(sa2_mel_fort) + 
+  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon),
+                   fill = p50_col), colour = "lightgrey") +
+  scale_fill_manual(values = nac_colours) + 
+  coord_sf(crs = 3112, 
+           xlim = c(bris_box["xmin"], bris_box["xmax"]), 
+           ylim = c(bris_box["ymin"], bris_box["ymax"])) +
+  guides(fill = FALSE) +
+  labs(fill = "Melanoma Cancer (Males):\nAge-standardised rate\n(per 100,000)") + invthm 
+  
 qld_ggchoro
 ggsave(filename = "figures/qld_ggchoro.png", device = "png",
        bg = "transparent",  dpi = 300,  width = 7, height = 6)
@@ -161,10 +175,13 @@ save(qld_legend, file = "figures/qld_legend.rda")
 # Cartograms
 
 # Contiguous Cartograms
-#qldcont <- sa2_mel %>% 
[…]

### 2020-12-08 ACE-19-31.Rmd
Clean repo for publication

--- 2020-08-26 literature.Rmd
@@
                       message = FALSE,
-  fig_path = "figs/",
   out.width = "\\textwidth")
@@
 These atlases are described in Table 2.", fig.align='center', fig.width=12, fig.height=10}
-knitr::include_graphics("figures/choropleth_grid.png")
+knitr::include_graphics("figs/F1 choropleth_grid.png")
 ```
@@
 ", message=FALSE, warning=FALSE}
-knitr::include_graphics("figures/usa_grid.png")
+knitr::include_graphics("figs/F2 usa_grid.png")
 ```
@@
 This display allows the presentation of multiple variables for each geographic area.", out.width="100%"}
-knitr::include_graphics("figures/gggrids.png")
+knitr::include_graphics("figs/F3 grids.png")
 ```
@@
 The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
-knitr::include_graphics("figures/auscartograms.png")
+knitr::include_graphics("figs/F4 auscartograms.png")
 ```
@@
 Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for Globocan 2018 [@Globocan], (ii) Menus for variable selection and zooming on Bowel Cancer Australia Atlas, (iii) Menus for choosing variables and countries in The Cancer Atlas, (iv) Tabs for different indicators and cancer types in Global Cancer Map, (v) Menus and toggles for variable and subset selection in United States Cancer Statistics: Data Visualizations.", results = "asis", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
-knitr::include_graphics("figures/interacting.png")
+knitr::include_graphics("figs/F5 interacting.png")
 ```
@@
 A highly responsive magnifying glass on a map of Breast Cancer Mortality in Canada.", results = "asis", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
-knitr::include_graphics("figures/animating.png")
+knitr::include_graphics("figs/F6 animating.png")
 ```

### 2020-12-08 R/sa3_plots.R

@@ -1,0 +1,246 @@
+#sa3_plots
+
+# Create a choropleth of Australia
+library(sugarbag)
+library(tidyverse)
+library(readxl)
+library(spData)
+library(ggthemes)
+library(maptools)
+library(sf)
+library(cartogram)
+library(cowplot)
+
+
+# Filter for centroids only within longitudes
+checkgeom <- function(x){
+  xval <- st_coordinates(x)[[2]]
+  
+  if (xval < 109){return(FALSE)}
+  if (xval > 155){return(FALSE)}
+  return(TRUE)
+}
+
+sa3lung <- absmapsdata::sa32016 %>% 
+  filter(!st_is_empty(geometry)) %>%
+  filter(checkgeom(x = geometry))
+  
+sa3lung <- sa3lung %>% 
+  rmapshaper::ms_simplify(keep = 0.05, keep_shapes = TRUE)
+
+sa3lung_fort <- fortify_sfc(sa3lung)
+
+# Join with cancer data from AIHW
+#https://www.aihw.gov.au/getmedia/df6732ee-5246-4c9f-a458-df85c88fc512/aihw-can-108-cancer-SA3-incidence-mortality.xls.aspx
+
+cimar <- read_excel("~/review/data/cimar-sa3.xlsx", 
+  sheet = "Incidence Persons", range = "A3:Z337")
+
+cimar_sa3 <- cimar %>%  
+  # select lung cancer variables
+  rename(Code = `...1`,
+    Name = `...2`,
+    `Total incidence` = `...22`,
+    Population = `...23`,
+    `Crude rate (per 100,000)` = `...24`,
+    `Age-standardised rate (per 100,000)` = `...25`,
+    `Rate ratio (relative to all of Australia)` = `...26`
+  ) %>% 
+  filter(!is.na(Name)) %>% 
+  filter(!(Name %in% c("Name", "Australia"))) %>% 
+  mutate(Population = as.numeric(Population))
+
+# Filter for desired areas
+cimar_sa3 <- cimar_sa3 %>% 
+  filter((Name %in% sa3lung$sa3_name_2016))
+
+# Join with sa3 sf object
+sa3lung <- sa3lung %>% 
+  left_join(cimar_sa3, by = c("sa3_name_2016" = "Name")) %>% 
+  st_as_sf() %>%
+  #filter(!is.na(Population)) %>%
+  mutate(Population = as.numeric(ifelse(is.na(Population), 1, Population)),
[…]

### 2020-12-08 R/usa_plots.R

@@ -1,0 +1,325 @@
+#USA Cancer Data
+
+# United States Cancer Statistics: Data Visualizations
+# The official federal statistics on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
+
+# https://gis.cdc.gov/cancer/USCS/DataViz.html
+
+
+# United States
+# Rate of New Cancers
+# Female and Male
+# Lung and Bronchus
+# 2012-2016
+
+# Export
+library(sugarbag)
+library(tidyverse)
+library(spData)
+library(ggthemes)
+library(maptools)
+library(sf)
+library(cartogram)
+library(cowplot)
+library(grid)
+library(png)
+library(gridExtra)
+
+# read in data
+cancer <- read_csv("~/review/data/USCSlung.csv") 
+
+cancer <-  cancer %>% 
+  rename(NAME = Area)
+
+
+# Join polygons to data
+cancer <- left_join(us_states, cancer, by = c("NAME"))
+cancer <- st_transform(cancer, 3857)
+
+b <- st_bbox(cancer)
+
+# Histogram
+ggplot(cancer, aes(x = total_pop_15)) + geom_histogram() 
+ggplot(cancer, aes(x = total_pop_15 )) + geom_density()
+
+title <- ""
+  #"Average rate of incidence for lung and bronchus for females and males in the United States 2012-2016, \n where the size of each state has been adjusted for population"
+
+###############################################################################
+
+invthm <- theme_void() + theme(
+  panel.background = element_rect(fill = "transparent", colour = NA), 
+  plot.background = element_rect(fill = "transparent", colour = NA),
+  legend.background = element_rect(fill = "transparent", colour = NA),
+  legend.key = element_rect(fill = "transparent", colour = NA),
+  text = element_text(colour = "white", size = 20)
+)
+
+# Create a choropleth
+ggchoro <- cancer %>% 
+  mutate(`Age adjusted rate\n(per 100,000)` = AgeAdjustedRate) %>% 
+  ggplot() + 
+  geom_sf(aes(fill = `Age adjusted rate\n(per 100,000)`), colour = NA) +
[…]

### 2020-12-08 cancer_map_images/review_figures.R

@@ -1,108 +1,0 @@
-# Figures
-library(ggplot2)
-library(grid)
-library(cowplot)
-
-# Import the images
-plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
-plot2 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
-plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
-plot4 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
-plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
-plot6 <- png::readPNG('cancer_map_images/choropleths/aca.choropleth.png')
-plot7 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')
-
-ggdraw(xlim = c(0,1), ylim = c(0,1)) +
-  draw_plot(rasterGrob(plot1), 0.02, .73, 0.31, .25) +
-  draw_plot(rasterGrob(plot2), 0.33, 0.73, 0.33, .25) +
-  draw_plot(rasterGrob(plot3), 0.66, 0.47, .33, .5) +
-  draw_plot(rasterGrob(plot4), 0, 0.465, .35, .26) +
-  draw_plot(rasterGrob(plot5), 0.33, 0.465, .33, .25) +
-  draw_plot(rasterGrob(plot6), 0.005, 0, .66, .5) +
-  draw_plot(rasterGrob(plot7), 0.68, 0, 0.26, .48) +
-  draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
-    c(0, 0.32, 0.66, 0, 0.32, 0, 0.66), 
-    c(0.99, 0.99, 0.99, 0.74, 0.74, 0.5, 0.5), size = 20)
-
-ggsave(filename = "figures/choropleth_grid.png", 
-  device = "png", width = 12, height = 10, dpi = 300)
-
-
-
-
-g1 <- png::readPNG("figures/ggcont.png")
-g2 <- png::readPNG("figures/ggncont.png")
-g3 <- png::readPNG("figures/ggdorl.png")
-g4 <- png::readPNG("figures/gghexmap.png")
-
-load("figures/usa_legend.rda")
-
-ggdraw() +
-  draw_plot(rasterGrob(g1), 0, .55, 0.5,  0.45) +
[…]

### 2020-12-08 cancer_map_images/sa3_plots.R

@@ -1,252 +1,0 @@
-#sa3_plots
-
-# Create a choropleth of Australia
-library(sugarbag)
-library(tidyverse)
-library(readxl)
-library(spData)
-library(ggthemes)
-library(maptools)
-library(sf)
-library(cartogram)
-library(cowplot)
-
-
-# Filter for centroids only within longitudes
-checkgeom <- function(x){
-  xval <- st_coordinates(x)[[2]]
-  
-  if (xval < 109){return(FALSE)}
-  if (xval > 155){return(FALSE)}
-  return(TRUE)
-}
-
-sa3lung <- absmapsdata::sa32016 %>% 
-  filter(!st_is_empty(geometry)) %>%
-  filter(checkgeom(x = geometry))
-  
-sa3lung <- sa3lung %>% 
-  rmapshaper::ms_simplify(keep = 0.05, keep_shapes = TRUE)
-
-sa3lung_fort <- fortify_sfc(sa3lung)
-
-# Join with cancer data from AIHW
-#https://www.aihw.gov.au/getmedia/df6732ee-5246-4c9f-a458-df85c88fc512/aihw-can-108-cancer-SA3-incidence-mortality.xls.aspx
-
-cimar <- read_excel("~/review/data/cimar-sa3.xlsx", 
-  sheet = "Incidence Persons", range = "A3:Z337")
-
-cimar_sa3 <- cimar %>%  
-  # select lung cancer variables
-  rename(Code = `...1`,
-    Name = `...2`,
-    `Total incidence` = `...22`,
-    Population = `...23`,
-    `Crude rate (per 100,000)` = `...24`,
-    `Age-standardised rate (per 100,000)` = `...25`,
-    `Rate ratio (relative to all of Australia)` = `...26`
-  ) %>% 
-  filter(!is.na(Name)) %>% 
-  filter(!(Name %in% c("Name", "Australia"))) %>% 
-  mutate(Population = as.numeric(Population))
-
-# Filter for desired areas
-cimar_sa3 <- cimar_sa3 %>% 
-  filter((Name %in% sa3lung$sa3_name_2016))
-
-# Join with sa3 sf object
-sa3lung <- sa3lung %>% 
-  left_join(cimar_sa3, by = c("sa3_name_2016" = "Name")) %>% 
-  st_as_sf() %>%
-  filter(!is.na(Population)) %>%
[…]

### 2020-12-08 cancer_map_images/usa_plots.R

@@ -1,345 +1,0 @@
-#USA Cancer Data
-
-# United States Cancer Statistics: Data Visualizations
-# The official federal statistics on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
-
-# https://gis.cdc.gov/cancer/USCS/DataViz.html
-
-
-# United States
-# Rate of New Cancers
-# Female and Male
-# Lung and Bronchus
-# 2012-2016
-
-# Export
-library(sugarbag)
-library(tidyverse)
-library(spData)
-library(ggthemes)
-library(maptools)
-library(sf)
-library(cartogram)
-library(cowplot)
-library(grid)
-library(png)
-library(gridExtra)
-
-# read in data
-cancer <- read_csv("~/review/data/USCSlung.csv") 
-
-cancer <-  cancer %>% 
-  rename(NAME = Area)
-
-
-# Join polygons to data
-cancer <- left_join(us_states, cancer, by = c("NAME"))
-cancer <- st_transform(cancer, 3857)
-
-b <- st_bbox(cancer)
-
-# Histogram
-ggplot(cancer, aes(x = total_pop_15)) + geom_histogram() 
-ggplot(cancer, aes(x = total_pop_15 )) + geom_density()
-
-title <- ""
-  #"Average rate of incidence for lung and bronchus for females and males in the United States 2012-2016, \n where the size of each state has been adjusted for population"
-
-###############################################################################
-
-invthm <- theme_void() + theme(
-  panel.background = element_rect(fill = "transparent", colour = NA), 
-  plot.background = element_rect(fill = "transparent", colour = NA),
-  legend.background = element_rect(fill = "transparent", colour = NA),
-  legend.key = element_rect(fill = "transparent", colour = NA),
-  text = element_text(colour = "white", size = 20)
-)
-
-# Create a choropleth
-ggchoro <- cancer %>% 
-  mutate(`Age adjusted rate\n(per 100,000)` = AgeAdjustedRate) %>% 
-  ggplot() + 
-  geom_sf(aes(fill = `Age adjusted rate\n(per 100,000)`), colour = NA) +
[…]

### 2020-12-08 qldsa2.R

@@ -1,526 +1,0 @@
-# Queensland maps
-
-# Create a choropleth of Australia
-library(sugarbag)
-library(tidyverse)
-library(readxl)
-library(spData)
-library(ggthemes)
-library(maptools)
-library(sf)
-library(cartogram)
-library(cowplot)
-
-# SA2 populations
-#https://www.qgso.qld.gov.au/issues/5496/estimated-resident-population-sa2-qld-2006-2019p.csv
-pop <- read_csv("data/estimated-resident-population-sa2-qld-2006-2019p.csv",
-                skip = 3) %>% select(SA2_name = "X2", pop = `2011`) %>% 
-                drop_na()
-
-# SIR data from Australian Cancer Atlases
-SIR <- read_csv("data/SIR Downloadable Data.csv")
-
-SIR <- SIR %>%
-  filter(SA2_code >30000, SA2_code <40000) %>% 
-  filter(Year == "2010-2014") 
-
-SIR_fem <- SIR %>% 
-  filter(Sex_name == "Females") %>%  
-  select(Year, SA2_NAME11 = SA2_name,
-         p50, Cancer_name) %>% 
-  spread(Cancer_name, p50)
-
-SIR_male <- SIR %>%
-  filter(Year == "2010-2014") %>% 
-  filter(Sex_name == "Males") %>%  
-  select(Year, SA2_NAME11 = SA2_name,
-         p50, Cancer_name) %>% 
-  spread(Cancer_name, p50)
-
-SIR_persons <- SIR %>%
-  filter(Year == "2010-2014") %>% 
-  filter(Sex_name == "Persons") %>%  
-  select(Year, SA2_NAME11 = SA2_name,
-         p50, Cancer_name) %>% 
-  spread(Cancer_name, p50)
- 
-
-# Create a colour scheme
-nac_colours <- c("A" = "#33809d",
-                 "B" = "#aec6c7",
-                 "C" = "#fff4bc",
-                 "D" = "#ff9a64",
-                 "E" = "#ff3500")
-
-colour_cat <- function(num){
-  case_when(
-    num < 0.75  ~ "A",
-    num >= 0.75 & num < 1  ~ "B",
-    num >= 1 & num < 1.25  ~ "C",
-    num >= 1.25 & num < 1.5  ~ "D",
-    num > 1.5  ~ "E")
-} 
-
-
-# Filter for centroids only within longitudes
-checkgeom <- function(x){
[…]

### 2020-12-08 version1/aus_plots.R

@@ -1,69 +1,0 @@
-# Create a choropleth of Australia
-library(sugarbag)
-library(tidyverse)
-
-# 
-# # Filter islands
-# aus_sf <- aus_sf %>% 
-#   filter(!(SA2_NAME11 %in% 
-#       c("Christmas Island", "Cocos (Keeling) Islands", "Lord Howe Island")))
-
-# Join with cancer data from AIHW
-#https://www.aihw.gov.au/getmedia/df6732ee-5246-4c9f-a458-df85c88fc512/aihw-can-108-cancer-SA3-incidence-mortality.xls.aspx
-
-library(readxl)
-cimar_sa3 <- read_excel("data/cimar-sa3.xls", 
-  sheet = "Incidence Persons", range = "A3:Z337")
-
-
-SIR <- SIR %>%
-  filter(Year == "2010-2014") 
-
-SIR_fem <- SIR %>% 
-  filter(Sex_name == "Females") %>%  
-  select(Year, SA2_NAME11 = SA2_name,
-    p50, Cancer_name) %>% 
-  spread(Cancer_name, p50)
-
-SIR_male <- SIR %>%
-  filter(Year == "2010-2014") %>% 
-  filter(Sex_name == "Males") %>%  
-  select(Year, SA2_NAME11 = SA2_name,
-    p50, Cancer_name) %>% 
-  spread(Cancer_name, p50)
-
-SIR_persons <- SIR %>%
-  filter(Year == "2010-2014") %>% 
-  filter(Sex_name == "Persons") %>%  
-  select(Year, SA2_NAME11 = SA2_name,
-    p50, Cancer_name) %>% 
-  spread(Cancer_name, p50)
-
-aus_sf <- aus_sf %>% 
-  left_join(SIR_persons) 
-
-# Create a colour scheme
-nac_colours <- c("A" = "#33809d",
-  "B" = "#aec6c7",
-  "C" = "#fff4bc",
-  "D" = "#ff9a64",
-  "E" = "#ff3500")
-
-colour_cat <- function(num){
-  case_when(
-    num < 0.75  ~ "A",
-    num >= 0.75 & num < 1  ~ "B",
-    num >= 1 & num < 1.25  ~ "C",
-    num >= 1.25 & num < 1.5  ~ "D",
-    num > 1.5  ~ "E")
-} 
-
-
-mel_data <- aus_sf %>% 
-  mutate(p50_col = factor(colour_cat(Melanoma), 
-    levels = c("A", "B", "C", "D", "E")))
-
-
[…]

### 2020-12-08 version1/figure_patchwork.R

@@ -1,2 +1,0 @@
-# Figure patchworking
-

### 2020-12-08 version1/sa3_plots.R

@@ -1,246 +1,0 @@
-#sa3_plots
-
-# Create a choropleth of Australia
-library(sugarbag)
-library(tidyverse)
-library(readxl)
-library(spData)
-library(ggthemes)
-library(maptools)
-library(sf)
-library(cartogram)
-library(cowplot)
-
-
-# Filter for centroids only within longitudes
-checkgeom <- function(x){
-  xval <- st_coordinates(x)[[2]]
-  
-  if (xval < 109){return(FALSE)}
-  if (xval > 155){return(FALSE)}
-  return(TRUE)
-}
-
-sa3lung <- absmapsdata::sa32016 %>% 
-  filter(!st_is_empty(geometry)) %>%
-  filter(checkgeom(x = geometry))
-  
-sa3lung <- sa3lung %>% 
-  rmapshaper::ms_simplify(keep = 0.05, keep_shapes = TRUE)
-
-sa3lung_fort <- fortify_sfc(sa3lung)
-
-# Join with cancer data from AIHW
-#https://www.aihw.gov.au/getmedia/df6732ee-5246-4c9f-a458-df85c88fc512/aihw-can-108-cancer-SA3-incidence-mortality.xls.aspx
-
-cimar <- read_excel("~/review/data/cimar-sa3.xlsx", 
-  sheet = "Incidence Persons", range = "A3:Z337")
-
-cimar_sa3 <- cimar %>%  
-  # select lung cancer variables
-  rename(Code = `...1`,
-    Name = `...2`,
-    `Total incidence` = `...22`,
-    Population = `...23`,
-    `Crude rate (per 100,000)` = `...24`,
-    `Age-standardised rate (per 100,000)` = `...25`,
-    `Rate ratio (relative to all of Australia)` = `...26`
-  ) %>% 
-  filter(!is.na(Name)) %>% 
-  filter(!(Name %in% c("Name", "Australia"))) %>% 
-  mutate(Population = as.numeric(Population))
-
-# Filter for desired areas
-cimar_sa3 <- cimar_sa3 %>% 
-  filter((Name %in% sa3lung$sa3_name_2016))
-
-# Join with sa3 sf object
-sa3lung <- sa3lung %>% 
-  left_join(cimar_sa3, by = c("sa3_name_2016" = "Name")) %>% 
-  st_as_sf() %>%
-  #filter(!is.na(Population)) %>%
-  mutate(Population = as.numeric(ifelse(is.na(Population), 1, Population)),
[…]

### 2020-12-08 version1/usa_plots.R

@@ -1,325 +1,0 @@
-#USA Cancer Data
-
-# United States Cancer Statistics: Data Visualizations
-# The official federal statistics on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).
-
-# https://gis.cdc.gov/cancer/USCS/DataViz.html
-
-
-# United States
-# Rate of New Cancers
-# Female and Male
-# Lung and Bronchus
-# 2012-2016
-
-# Export
-library(sugarbag)
-library(tidyverse)
-library(spData)
-library(ggthemes)
-library(maptools)
-library(sf)
-library(cartogram)
-library(cowplot)
-library(grid)
-library(png)
-library(gridExtra)
-
-# read in data
-cancer <- read_csv("~/review/data/USCSlung.csv") 
-
-cancer <-  cancer %>% 
-  rename(NAME = Area)
-
-
-# Join polygons to data
-cancer <- left_join(us_states, cancer, by = c("NAME"))
-cancer <- st_transform(cancer, 3857)
-
-b <- st_bbox(cancer)
-
-# Histogram
-ggplot(cancer, aes(x = total_pop_15)) + geom_histogram() 
-ggplot(cancer, aes(x = total_pop_15 )) + geom_density()
-
-title <- ""
-  #"Average rate of incidence for lung and bronchus for females and males in the United States 2012-2016, \n where the size of each state has been adjusted for population"
-
-###############################################################################
-
-invthm <- theme_void() + theme(
-  panel.background = element_rect(fill = "transparent", colour = NA), 
-  plot.background = element_rect(fill = "transparent", colour = NA),
-  legend.background = element_rect(fill = "transparent", colour = NA),
-  legend.key = element_rect(fill = "transparent", colour = NA),
-  text = element_text(colour = "white", size = 20)
-)
-
-# Create a choropleth
-ggchoro <- cancer %>% 
-  mutate(`Age adjusted rate\n(per 100,000)` = AgeAdjustedRate) %>% 
-  ggplot() + 
-  geom_sf(aes(fill = `Age adjusted rate\n(per 100,000)`), colour = NA) +
[…]
